### 2020-05-31 problem-19.17/Text.tex
first commit

\documentclass[a4paper]{article}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{bussproofs}
\usepackage{mathtools}
\usepackage{verbatim}
\usepackage{dsfont}
\usepackage{mathabx}
\usepackage[all, 2cell]{xy}
\usepackage[all]{xy}
\usepackage{wasysym}
\usepackage{rotating}
\usepackage{geometry}
\usepackage{trfsigns}
\usepackage{cmll}
\usepackage{authblk}
\usepackage{hyperref}
\usepackage{cleveref}
\usepackage{lipsum}
\usepackage{extpfeil}
\usepackage{soul}
\usepackage{graphicx}

\newcommand\mapsfrom{\mathrel{\reflectbox{\ensuremath{\mapsto}}}}

\theoremstyle{defin}
\newtheorem{defin}{Definition}

\theoremstyle{theorem}
\newtheorem{theorem}{Theorem}

\theoremstyle{prop}
\newtheorem{prop}{Proposition}

\theoremstyle{lemma}
\newtheorem{lemma}{Lemma}

\theoremstyle{ex}
\newtheorem{ex}{Example}

\theoremstyle{col}
\newtheorem{col}{Corollary}

\author{Daniel Rogozin}
\date{}
\title{The finite base property for some relation algebras subreducts}

\begin{document}

\maketitle

\nocite{*}

\section{The Relation Algebras Background}

We describe the basic definitions and results about relation algebras \cite{hirsch2002relation} \cite{maddux2006relation}.

\begin{defin}
  $ $

  \begin{enumerate}
    \item A relation algebra is an algebra $\mathcal{R} = \langle R, 0, 1, \wedge, \vee, \neg, ;, {}^{\smile}, {\bf 1 }\rangle$ such that $\langle R, 0, 1, \wedge, \vee, \neg \rangle$ is a Boolean algebra and the following
    equations hold, for each $a, b, c \in R$:
    \begin{enumerate}
      \item $a ; (b ; c) = (a ; b) ; c$
      \item $(a \vee b) ; c = (a ; c) \vee (b ; c)$
      \item $a ; {\bf 1} = a$
      \item $a^{\smile \smile} = a$
      \item $(a \vee b)^{\smile} = a^{\smile} \vee b^{\smile}$
      \item $(a ; b)^{\smile} = b^{\smile} ; a^{\smile}$
      \item $a^{\smile} ; (\neg (a ; b)) \leq \neg b$
    \end{enumerate}
    where $a \leq b$ iff $a \wedge b = a$ iff $a \vee b = b$. ${\bf RA}$ denotes the class of all relation
    algebras.
    \item A proper relation algebra is an algebra $\mathcal{R} = \langle R, 0, 1, \wedge, \vee, \neg, ;, {}^{\smile}, {\bf 1 }\rangle$ such that $R \subseteq \mathcal{P}(W)$, where $W$ is an equivalence relation; $0 = \emptyset$; $1 = W$; $\wedge$, $\vee$, $\neg$ are set-theoretic intersection, union, and
    complement respectively; $;$ is relation composition, ${}^{\smile}$ is relation converse, ${\bf 1}$ is a
    diagonal relation restricted to $W$, that is:
    \begin{enumerate}
    \item $a ; b = \{ \langle x, z \rangle \: | \: \exists y \: \langle x, y \rangle \in a \: \& \: \langle y, z \rangle \in b \}$
    \item $a^{\smile} = \{ \langle x, y \rangle \: | \: \langle y, x \rangle \in a \}$
    \item ${\bf 1} = \{ \langle x, y \rangle \: | \: x = y \}$
    \end{enumerate}
      The class of all proper relation algebras is denoted as ${\bf PRA}$. ${\bf Rs}$ is the class of all relation set algebras, proper relation algebra with a diagonal subrelation as an identity. ${\bf RRA}$ is the class of all representable relation algebras, that is, the closure of ${\bf PRA}$ under isomorphic copies. That is,
      ${\bf RRA} = {\bf IPRA}$.
  \end{enumerate}
\end{defin}
Note that the (quasi)equational theories of those classes coincide, that is
\begin{center}
  ${\bf IPRA} = {\bf RRA} = {\bf SP}{\bf Rs}$
\end{center}

One may express residuals in every $\mathcal{R} \in {\bf RA}$ as follows, for every $a, b \in \mathcal{R}$:

\begin{enumerate}
  \item $a \setminus b = \neg (a^{\smile} ; \neg b)$
  \item $a / b = \neg (\neg a ; b^{\smile})$
\end{enumerate}

Those residuals have the following interpretation in $\mathcal{R} \in {\bf PRA}$ (as well as in ${\bf RRA}$), for every $a, b \in \mathcal{R}$:
\begin{enumerate}
  \item $a \setminus b = \{ \langle x, y \rangle \: | \: \forall z \: (z, x) \in a \Rightarrow (z, y) \in b \}$
  \item $a / b = \{ \langle x, y \rangle \: | \: \forall z \: (y, z) \in b \Rightarrow (x, z) \in a \}$
\end{enumerate}
One may illustrate composition and residuals in ${\bf PRA}$ and ${\bf RRA}$ via the following triangles:

\xymatrix{
&&& \exists y \ar@{-->}[ddr]^{b} &&& \forall z \ar@{-->}[ddl]_{a} \ar@{-->}[ddr]^{b} &&& \forall z \\
&&&&&& \Rightarrow &&& \Leftarrow \\
&& x \ar@{-->}[uur]^{a} \ar[rr]_{a;c} && z & x \ar[rr]_{a \setminus b} && y & x \ar@{-->}[uur]^{a} \ar[rr]_{a / b} && y \ar@{-->}[uul]_{b}
}

Given a subset of definable operations in ${\bf RA}$ $\tau$, we denote the class of subalgebras of the
$\tau$-reducts by ${\bf R}(\tau)$. The algebras containing to this class are defined as restrictions of
elements belonging to ${\bf Rs}$ to operations of $\tau$. By ${\bf Q}(\tau)$ we mean a quasivariety generated by $R(\tau)$. As in \cite{hirsch2011positive}, we put ${\bf Q}(\tau)$ as the closure of ${\bf R}(\tau)$ under subalgebras and products assuming that ${\bf R}(\tau)$ is already closed under ultraproducts.

\section{The Finite Base Property}

We recall the underlying definitions according to \cite[Section 19]{hirsch2002relation}

\begin{defin}
  Let ${\bf K}$ be a class of algebras of a signature $\Omega$, ${\bf K}$ has the finite algebra property, if
  if any first-order $\Omega$-sentence that is true in all finite algebras in ${\bf K}$ is true in every algebra in ${\bf K}$.
\end{defin}

The finite base property is a version of the finite algebra property if ${\bf K}$ is a class of representable algebras:

\begin{defin}
Let ${\bf K}$ be a class of representable algebras of a signature $\Omega$

  \begin{enumerate}
  \item ${\bf K}$ has the finite base property if any first-order $\Omega$-sentence that is true in every algebra in ${\bf K}$ having a representation over a finite base set is valid in ${\bf K}$.
  \item ${\bf K}$ has the finite algebra on finite base property if any finite algebra in ${\bf K}$ has a representation with finite base.
  \item ${\bf K}$ has the finite algebra property for equations/quasi-identites if any equation/quasi-identity that is true in all finite algebras is true in every algebra in ${\bf K}$. The finite base property for equations/quasi-identites is defined similarly.
\end{enumerate}
\end{defin}

The following statements were shown in \cite{andreka1999finite}. This lemma connects finite base property with finite algebra on finite base and finite algebra properties as follows:
\begin{lemma}
  Let ${\bf K}$ be a class of representable $\Omega$-algebras:
  \begin{enumerate}
    \item If ${\bf K}$ has the finite algebra property, then it has the finite algebra and the finite base properties for equations/quasi-identites.
    \item The finite algebra on finite base and the finite algebra properties implies the finite base property for ${\bf K}$. The same holds for equations/quasi-identites.
    \item If any representation of an infinite algebra has an infinite base, then the finite base property implies the finite algebra one for ${\bf K}$.
    \item Suppose $\Omega$ is finite and any subalgebra of a representable algebra is representable on the same base. Then the finite base property implies the finite algebra on finite base property.
  \end{enumerate}
\end{lemma}

\section{The Relation Residuated Semigroups Background}

\subsection{The underlying definitions and results}

 A \emph{relation structure} (${\bf RS}$) is an arbitrary algebra of the signature $\Omega = \langle \cdot, \setminus, /, \leq \rangle$, where $\cdot, \setminus, /$ are binary function symbols and $\leq$ is a binary relation symbol.

\begin{defin}
  A residuated semigroup is an algebra $\mathcal{S} = \langle S, \cdot, \leq, \setminus, / \rangle$ such that $\langle S, \cdot, \leq, \rangle$ is an ordered residuated semigroup ${\bf ORS}$ an the following equivalences hold for each $a, b, c \in S$:

  \begin{center}
    $b \leq a \setminus c \Leftrightarrow a \cdot b \leq c \Leftrightarrow a \leq c / b$
  \end{center}
\end{defin}

It is clear that ${\bf ORS}$ is a quasivariety and, thus \cite{mal2012algebraic}, is closed under subalgebras, products, and ultraproducts, i. e.,
${\bf ORS} = {\bf S}{\bf P}{\bf Up} \: {\bf ORS}$.

\begin{defin} \label{rrs}
  Let $A$ be a set of binary relations on some base set $W$ such that $R = \cup A$ is transitive and $\{ x, y \: | \: x R y \} = W$. A relation residuated semigroup is an algebra $\mathcal{A} = \langle A, ;, \setminus, /, \subseteq \rangle$ where for each $r, s \in A$
  \begin{enumerate}
    \item $r ; s = \{ \langle a, c \rangle \: | \: \exists b \in W \: (\langle a, b \rangle \in r \: \& \: \langle b, c \rangle \in s) \}$
    \item $r \setminus s = \{ \langle a, c \rangle \: | \: \forall b \in W \: (\langle b, a \rangle \in r \Rightarrow \langle b, c \rangle \in s)\}$
    \item $r / s = \{ \langle a, c \rangle \: | \: \forall b \in W \: (\langle c, b \rangle \in s \Rightarrow \langle a, b \rangle \in r)\}$
  \end{enumerate}
\end{defin}
Relation residuated semigroup are also called representable relativised relational structure (${\bf RRS}$).

Andr\'{e}ka and Mikul\'{a}s proved the following representation theorem for ${\bf ORS}$ in \cite{andreka1994lambek} that implies relational completeness of the Lambek calculus, the logic of ${\bf ORS}$:

\begin{theorem}\label{ors=irrs}
  ${\bf ORS} = {\bf IRRS}$, where ${\bf IRRS}$ is a closure of ${\bf RRS}$ under isomorphic copies.
\end{theorem}

\subsection{The finite base property for ${\bf RRS}$}

\begin{defin}
  A relativised representation
\end{defin}

\begin{defin}
  The standard translation
\end{defin}

TODO: take a look at relativised representations and loosely guarded fragments in general
TODO: realise whether it makes sense to use the technique similar to \cite[Theorem 19.13]{hirsch2002relation}
used for weakly associative algebras.

\begin{theorem}
  Let $\mathcal{A}$ be a finite residuated semigroup and
  $|\mathcal{A}| < \omega$, then $\mathcal{A}$ has a finite relativised representation.
\end{theorem}

\begin{theorem}
  Let $\mathcal{A}$ be a finite representable residuated semigroup, then $\mathcal{A}$ is isomorphic to
  representable residuated semigroup a domain of which is finite.
\end{theorem}

\begin{theorem}
  That might follow from the previous theorem, Theorem~\label{ors=irrs}, and something else.
\end{theorem}

\begin{col}
  The Lambek calculus has the fmp and the universal theory of ${\bf IRRS}$ is NP-complete.
\end{col}

The hypothetical plan is the following one:
\begin{enumerate}
\item Define properly relativised representation for residuated semigroups, that should look like ternary Kripke frames for the basic Lambek calculus or arrow logic.
\item Define the standard translation to such first-order relation structures. TODO: take a look at the loosely guarded fragment stuff.
\item Every finite residuated semigroup has a finite relativised representation.
\item If every $\Pi_1$-statement $\varphi$ of the language of residuated semigroups that is valid in every
residuated semingroup is valid in algebra having a finite relativised representation (one may use here Theorem~\ref{ors=irrs} somehow), then $\varphi$ is valid in ${\bf ORS}$ as well as in ${\bf IRRS}$.
\item Every finite residuated semigroup should have a finite relativised representation.
\item Construct a finitely based relation residuated semigroup from that (an analogue of complex algebra or smth like that). This item is the most non-trivial one.
\item As a corollary, the first-order universal first-order theory of ${\bf IRRS}$ should be decidable and
(it seems so) $\operatorname{NP}$-complete (that should follow from the results in \cite{pentus2006lambek}).
\end{enumerate}

\section{Join-semilattice ordered semigroups}

\begin{defin} A join-semilattice ordered semigroup (${\bf ASOS}$) is an algebra $\mathcal{S} = \langle S, \cdot, \vee \rangle$ such that $\langle S, \cdot \rangle$ is a semigroup, $\langle S, \vee \rangle$ is a join-semilattice and the following equations hold for each $a, b. c \in S$:

  \begin{enumerate}
    \item $a \cdot (b \vee c) = (a \cdot b) \vee (a \cdot c)$
    \item $(a \vee b) \cdot c = (a \cdot c) \vee (b \cdot c)$
  \end{enumerate}
\end{defin}
This class is clearly a variety since ${\bf ASOS}$ has the equational definition so far as $\vee$ is defined as an associative, idempotent, and commutative operation.

Let $A$ be a set of binary relations on some base set $W$ such that $R = \cup A$ is transitive and $\{ x, y \: | \: x R y \} = W$ as in Definition~\ref{rrs}. A relation join-semilattice ordered semigroup (${\bf RJSOS}$) is an algebra of binary relations $\mathcal{A} = \langle A, ;, \cup\rangle$ such that $;$ is a relation composition as above and $\cup$ is the set-theoretic union. It is known that the class of all representable join-semilattice ordered semigroups has no finite axiomatisation \cite{andreka1989union}.

\begin{theorem}
  The equational and quasiequational theories of $R(;,\vee)$ is not finitely based.
\end{theorem}

\begin{proof}
TODO: use games and rainbow stuff
\end{proof}

\subsection{The finite base property for ${\bf RJSOS}$}

\subsection{The finite algebra property for ${\bf RJSOS}$}

\subsection{The finite algebra on finite base for ${\bf RJSOS}$}

\bibliographystyle{plain}
\bibliography{Text}

\end{document}

### 2020-05-31 problem-9.4/Text.tex
first commit

\documentclass[a4paper]{article}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{bussproofs}
\usepackage{mathtools}
\usepackage{verbatim}
\usepackage{dsfont}
\usepackage{mathabx}
\usepackage[all, 2cell]{xy}
\usepackage[all]{xy}
\usepackage{wasysym}
\usepackage{rotating}
\usepackage{geometry}
\usepackage{trfsigns}
\usepackage{cmll}
\usepackage{authblk}
\usepackage{hyperref}
\usepackage{cleveref}
\usepackage{lipsum}
\usepackage{extpfeil}
\usepackage{soul}
\usepackage{graphicx}

\newcommand\mapsfrom{\mathrel{\reflectbox{\ensuremath{\mapsto}}}}

\theoremstyle{defin}
\newtheorem{defin}{Definition}

\theoremstyle{theorem}
\newtheorem{theorem}{Theorem}

\theoremstyle{prop}
\newtheorem{prop}{Proposition}

\theoremstyle{lemma}
\newtheorem{lemma}{Lemma}

\theoremstyle{ex}
\newtheorem{ex}{Example}

\theoremstyle{col}
\newtheorem{col}{Corollary}

\author{Daniel Rogozin}
\date{}
\title{Model-theoretic aspects of relativised cylindric set algebras}

\begin{document}

\maketitle

\nocite{*}

\section{Model-theoretic and universal algebraic preliminaries}

\subsection{Ultraproducts}

Here are the required notions and facts from model theory and universal algebra \cite{hirsch2002relation} \cite{hodges1997shorter} \cite{sankappanavar1981course}.

Let $A$ be a non-empty set, an \emph{ultrafilter} on $A$ is a set of subsets $U \subseteq \mathcal{P}(\mathcal{P})(A)$ such that $A$ is closed under intersections, $\subseteq$-upwardly closed, and either $X \in U$ or $- X \in U$, where $X \subseteq A$. An ultrafilter is called principal if it has the form $\uparrow X = \{ Y \in \mathcal{P}(A) \: | \: X \subseteq Y \}$.

Let $\Omega = \langle \operatorname{Cnst}, \operatorname{Fn}, \operatorname{Pred} \rangle$ be a signature and $\Lambda$ an index set, and let $\{ M_{\lambda}\}_{\lambda \in \Lambda}$ be an indexed set of $\Omega$ structures. The $\Omega$-structure
\begin{center}
  $M = \prod \limits_{\lambda \in \Lambda} M_{\lambda}$
\end{center}
is called a \emph{product} that defined as follows. Its domain is the Cartesian product of the domains of $M_{\lambda}$. $a \in M$ is a function $\Lambda \to \bigcup \limits_{\lambda \in \Lambda} \operatorname{dom}(M_{\lambda})$ such that $a(\lambda) \in M_{\lambda}$ for each $\lambda \in \Lambda$.
Given $\lambda \in \Lambda$ and $a_{\lambda} \in M_{\lambda}$, we denote the function mapping $\lambda$ to $a_{\lambda}$ as
$\langle a_{\lambda} \: | \: \lambda \in \Lambda \rangle$. We define the interpretation of $\Omega$-symbols as

\begin{enumerate}
  \item If $c \in \operatorname{Cnst}$, then $c^{M} = \langle c^{M_{\lambda}} \: | \: \lambda \in \Lambda \rangle$
  \item If $f \in \operatorname{Fn}$ is an $n$-ary function symbol, then $f^{M}(\overline{a}) = \langle f^{M_{\lambda}}(\overline{a}) \: | \: \lambda \in \Lambda \rangle$, where $\overline{n} \in M^{n}$
  \item If $R \in \operatorname{Pred}$ is an $n$-ary relation symbol and $\overline{n} \in M^{a}$, then $R^{M}(\overline{a}) = \langle R^{M_{\lambda}}(\overline{a}) \: | \: \lambda \in \Lambda \rangle$
\end{enumerate}

Given $\lambda \in \Lambda$, we define the $\lambda$th projection as $\pi_{\lambda} : M \to M_{\lambda}$ such that $\pi_{\lambda}(a) = a(\lambda)$.

Let $\Lambda$ be an index set and $D$ an ultrafilter on the Boolean algebra $\langle \mathcal{P}(\Lambda), \cup, -, \Lambda, \emptyset \rangle$. Consider the product $M = \prod \limits_{\lambda \in \Lambda} M_{\lambda}$ of the $\Omega$-structures $\{ M_{\lambda}\}_{\lambda \in \Lambda}$ and the equivalence relation on $\operatorname{dom}(M)$ defined as
\begin{center}
  $a_1 \sim a_2 \Leftrightarrow \{ \lambda \in \Lambda \: | \: a_1(\lambda) = a_2(\lambda)\} \in D$
\end{center}
Let us denote $\operatorname{dom}(M) / \sim$ as $U$ and $[a]_{\sim}$ as $a / D$, where $a \in \operatorname{dom}(M)$. We also denote the \emph{ultraproduct} of $\{ M_{\lambda} \}_{\lambda}$ as $\prod \limits_{\lambda \in \Lambda} M_{\lambda} / D$, or,
for brevity, as $\prod_{D} M_{\lambda}$. The $\Omega$-symbols have the following interpretation
\begin{enumerate}
  \item If $c \in \operatorname{Cnst}$, then $c^{U} = c^{M} / D$
  \item If $f \in \operatorname{Fn}$ is an $n$-ary function symbol and $\overline{a} \in M^{n}$, then $f^{U}(\overline{a}) = f^{M}(x) = f^{M}(\overline{a}) / D$
  \item If $R \in \operatorname{Fn}$ is an $n$-ary relation symbol and $\overline{a} \in M^{n}$, then $U \models R(\overline{a} / D)$ iff
  $\{ \lambda \in \Lambda \: | \: M_{\lambda} \models R(\overline{a} (\lambda))\} \in D$
\end{enumerate}
The ultraproduct is principal if $D$ is a principal filter.

\begin{defin}
$ $

  \begin{enumerate}
    \item Let $\{ M_{\lambda}\}_{\lambda \in \Lambda}$ be a set of $\Omega$-structures such that every
    $M_{\lambda}$ is isomorphic to the single structure $M$, then their ultraproduct over $D$ is called the ultrapower over $D$.
    The denotation is $\prod \limits_{D} M$ or $M^{\Lambda} / D$.
    \item If $\prod \limits_{D} M \cong N$ for some structure $N$, then $M$ is an ultraroot of $N$.
  \end{enumerate}
\end{defin}

\begin{theorem}[\L{}os]

  Let $\{ M_{\lambda}\}_{\lambda \in \Lambda}$ be $\Omega$-structures and $D$ an ultrafilter on $\Lambda$, and let
  $U = \prod \limits_{D} M_{\lambda}$ be an ultraproduct of $\{ M_{\lambda}\}_{\lambda \in \Lambda}$ over $D$. For each first-order formula
  $\varphi(x_1, \dots, x_n)$ and for each $a_1 / D, \dots, a_n / D \in U$:

\begin{center}
  $U \models \varphi(a_1 / D, \dots, a_n / D)$ iff $\{ \lambda \in \Lambda \: | \: \varphi(a_1(\lambda), \dots, a_n(\lambda))\} \in D$
\end{center}
\end{theorem}

The \L{}os has the following helpful corollary:

\begin{col}
  Let $\prod \limits_{D} M$ be an ultrapower of $M$. For $a \in M$, let us define a function $\overline{a} : a \mapsto a / D$.
  Then such a map is an elementary embedding of $M$ into $\prod \limits_{D} M$.
\end{col}

Moreover, any elementary equivalent structures have isomorphic ultrapowers.

Recall that a class of $\Omega$-structures ${\bf K}$ is called \emph{elementary}, if ${\bf K} = \operatorname{Mod}(\operatorname{T})$ for some first-order theory ${\bf T}$. In that case, $\operatorname{T}$ is an axiomatisation of ${\bf K}$.

\begin{theorem}
  Let ${\bf K}$ be a class $\Omega$-structures,
  ${\bf K}$ is elementary iff ${\bf K}$ is closed under isomorphic copies, ultraroots, and ultrapowers.
\end{theorem}

\subsection{Preliminaries from universal algebra}

\begin{defin}
  Let ${\bf K}$ be a class of $\Omega$-structures, then ${\bf K}$ is a variety, if it is defined by some set of equations.
  The variety generated by ${\bf K}$ is the smallest variety contatining ${\bf K}$. ${\bf K}$ is a quasi-variety, if it is defined by some
  set of quasi-identities.
\end{defin}

Given a class ${\bf K}$ of $\Omega$-structures, then ${\bf I}({\bf K})$, ${\bf S}({\bf K})$, ${\bf H}({\bf K})$, and ${\bf P}({\bf K})$ are
the classes of isomorphic copies, algebras isomorphic to subalgebras belonging to ${\bf K}$,
algebras isomorphic to homomorphic images belonging to ${\bf K}$, and algebras isomorphic to direct products belonging to ${\bf K}$.
We claim that ${\bf I}({\bf K}) \subseteq {\bf S}({\bf K})$. ${\bf Up}({\bf K})$ is the class of algebras isomorphic to ultraproducts
belonging to ${\bf K}$.

\begin{theorem} Let ${\bf K}$ be a class of $\Omega$-structures

  \begin{enumerate}
    \item ${\bf K}$ is a variety iff ${\bf H}({\bf K}), {\bf S}({\bf K}), {\bf P}({\bf K}) \subseteq {\bf K}$
    \item ${\bf HSP}({\bf K}) = {\bf H}({\bf S}({\bf P}({\bf K})))$ is the smallest variety containg ${\bf K}$
    \item ${\bf K}$ is a quasi-variety iff it is closed under subalgebras, products, and ultraproducts, iff ${\bf SPUp}({\bf K}) = {\bf K}$.
  \end{enumerate}
\end{theorem}

\subsection{Subdirect products}

\begin{defin}
  $ $

  \begin{enumerate}
    \item Let $\{ A \}_{i \in I}$ be $\Omega$-structures, a subdirect product of $\langle A_i \: | \: i \in I \rangle$ is a
    subalgebra $B$ of $\prod \limits_{i \in I} A_i$ such that for each $i \in I$, a projection map $\pi_i : B \to A_i$ is a surjection.
    \item A subdirect representation of an $\Omega$-structure is an embedding $f : A \to \prod \limits_{i \in I} A_i$ for some $I$ and
    $\{ A_i \}_{i \in I}$ such that $f \circ \pi_i : A \to A_i$ is a surjection.
    \item An $\Omega$-structure $A$ is subdirectly irreducible if for every subdirect representation $f : A \to \prod \limits_{i \in I} A_i$
    there exists a projection $\pi_i$ such that $f \circ \pi_i$ is an isomorphism.
    \item ${\bf Sir}({\bf K})$ is the class of subdirectly irreducible structures belonging to ${\bf K}$.
    \item A subdirect decomposition of $A$ if there exists a subdirect representation $f : A \to \prod \limits_{i \in I} A_i$ such that
    every $A_i$ is subdirectly irreducible.
  \end{enumerate}
\end{defin}

It is known that every Boolean algebra with operators has a subdirect demcomposition. Moreover, that implies:
\begin{theorem}
$ $

  \begin{enumerate}
  \item If ${\bf K}$ is a variety, then every element of ${\bf K}$ has a subdirect decomposition with some subdirect irreducible elements of
  ${\bf K}$.
  \item If ${\bf K}$ is a variety and $\varepsilon$ is an equation,
  ${\bf Sir}({\bf K}) \models \varepsilon \Leftrightarrow {\bf K} \models \varepsilon$.
\end{enumerate}
\end{theorem}

\subsection{Pseudo-elementary classes}

\section{Cylindric algebras}

\subsection{(Representable) cylindric algebras and cylindric set algebras}

Let $\alpha$ be an ordinal. Let $U^{\alpha}$ be the set of all functions mapping $\alpha$ to a non-empty set $U$. We denote $x(i) = x_i$ for
$x \in U^{\alpha}$ and $i < \alpha$.

\begin{defin}
$ $

  \begin{enumerate}
    \item A subset of $U^{\alpha}$ is an $\alpha$-ry relation on $U$. For $i, j < \alpha$, the $i,j$-diagonal $D_{ij}$ is the set of all
    elements of $U$ such that $y_i = y_j$. If $i < \alpha$ and $X$ is an $\alpha$-ry relation on $U$, then
    the $i$-th cylindrification $C_i X$ is the set of all elements of $U$ that agree with some element of $X$ on each coordinate except the
    $i$-th one. To be more precise,
    $C_i X = \{ y \in U^{\alpha} \: | \: \exists x \in X \forall i < \alpha \: (i \neq j \Rightarrow y_j = x_j)\}$.
    \item A cylindic set algebra of dimension $\alpha$ is an algebra consisting of a set $S$ of $\alpha$-ry relation on some base set $U$
    with the constants and operations $0 = \emptyset$, $1 = U^{\alpha}$, $\cap$, $-$, the diagonal elements $\{ D_{ij} \}_{i, j < \alpha}$,
    the cylindrifications $\{ C \}_{i < \alpha}$. A generalised cylindric set algebra of dimension $\alpha$ is a subdirect of
    cylindric algebras that have dimension $\alpha$
    \item A cylindric algebra of dimension $\alpha$ is an algebra $\mathcal{C} = \langle \mathcal{B}, \{ c_i \}_{i < \alpha}, \{ d_{ij} \}_{i, j < \alpha} \rangle$ such that
    \begin{itemize}
      \item $\mathcal{B}$ is a Boolean algebra, for each $i, j < \alpha$ $c_i$ is an operator and $d_{ij} \in \mathcal{B}$
      \item For each $i < \alpha$, $a \leq c_i a$, $c_i (a \land c_i b) = c_i a \land c_i b$ and $d_{ii} = 1$
      \item For every $i, j < \alpha$, $c_i c_j a = c_j c_i a$
      \item If $k \neq i, j < \alpha$, then $d_{ij} = c_k (d_{ij} \land d_{jk})$
      \item If $i \neq j$, then $c_i (d_{ij} \land a) \land c_i (d_{ij} \land - a) = 0$
    \end{itemize}
    ${\bf CA}_{\alpha}$ is the class of all cylindric algebras of dimension $\alpha$
    \item An $\alpha$-dimensional cylindric algebra $C$ is representable, if it is isomorphic to a generalised cylindric set algebra
    of dimension $\alpha$. Such is isomorphism is a representation of $C$.
    ${\bf RCA}_{\alpha}$ is the class of all representable cylindric algebras that have dimension $\alpha$.
  \end{enumerate}
\end{defin}

\subsection{Substitution in cylindric algebras}

\begin{defin}
  Given a cylindric algebra of dimension $\alpha$ $C$, let $x$ be a term of its signature, the substitution operator $s^{i}_{j}$ have the following definition:
  \begin{center}
  $s^{i}_{j} x = \begin{cases} x, \text{if } i = j \\ c_i (d_{ij} \land x), \text{otherwise} \end{cases}$
  \end{center}
\end{defin}

\begin{prop} Let $\alpha$ be an ordinal and let $i, j, k, l < \alpha$. The following facts hold in ${\bf CA}_{\alpha}$

  \begin{enumerate}
    \item $s^{i}_j x \leq c_i x$.
    \item $s^{i}_j (x \land y) = s^{i}_j x \land s^{i}_j y$, $s^{i}_j (x \lor y) = s^{i}_j x \lor s^{i}_j y$, $- s^{i}_j x = s^{i}_j (- x)$.
    Moreover, $s^{i}_j$ is completely additive.
    \item $i \neq k, l$ implies $s^i_j d_{ik} = d_{jk}$ and $s^i_j d_{kl} = d_{kl}$.
    \item $d_{jk} \land s^{i}_j = d_{jk} \land s^{i}_k$.
    \item $s^i_j c_i x = c_i x$.
    \item $k \neq i,j$ implies $s^i_j c_i x = c_i s^i_j x$.
    \item $c_j s^i_j x = c_i s^j_i x$.
    \item $i \neq j$ implies $c_i s^i_j x = s^i_j x$.
    \item $i \neq k$ implies $s^i_j s^i_k = s^i_k x$.
    \item If either $i \notin \{ k, l\}$ and $k \notin \{ i, j\}$, or $j = l$, then $s^i_j s^k_l x = s^k_l s^i_j x$.
    \item $s^i_j s^j_i x = s^i_j x$.
    \item $s^i_k s^j_i x = s^i_k s^j_k x = s^j_k s^i_j x$
  \end{enumerate}
\end{prop}

\section{${\bf IG}_{\omega}$ and ultraproducts}

\section{${\bf IG}_{\omega}$ is (not) a variety; is (not) (pseudo-)elementary}

\bibliographystyle{plain}
\bibliography{Text}

\end{document}

### 2020-06-01 problem-19.17/Text.tex
update pdf
add bib files

fixes

update

update

fixes

@@
 \begin{defin}
-  A residuated semigroup is an algebra $\mathcal{S} = \langle S, \cdot, \leq, \setminus, / \rangle$ such that $\langle S, \cdot, \leq, \rangle$ is an ordered residuated semigroup ${\bf ORS}$ an the following equivalences hold for each $a, b, c \in S$:
+  A residuated semigroup is an algebra $\mathcal{S} = \langle S, \cdot, \leq, \setminus, / \rangle$ such that $\langle S, \cdot, \leq, \rangle$ is an ordered residuated semigroup and the following equivalences hold for each $a, b, c \in S$:
 
@@
   \end{center}
+  ${\bf ORS}$ is the class of all residuated semigroups.
 \end{defin}
-
-It is clear that ${\bf ORS}$ is a quasivariety and, thus \cite{mal2012algebraic}, is closed under subalgebras, products, and ultraproducts, i. e.,
-${\bf ORS} = {\bf S}{\bf P}{\bf Up} \: {\bf ORS}$.
 
@@
 
-\begin{theorem}
-  That might follow from the previous theorem, Theorem~\label{ors=irrs}, and something else.
-\end{theorem}
+\begin{proof}
+  That might follow from the previous theorem, Theorem~\ref{ors=irrs}, and something else.
+\end{proof}
 
@@
 \item Define properly relativised representation for residuated semigroups, that should look like ternary Kripke frames for the basic Lambek calculus or arrow logic.
-\item Define the standard translation to such first-order relation structures. TODO: take a look at the loosely guarded fragment stuff.
+\item Define the standard translation to such first-order relation structures. TODO: take a look at loosely guarded fragment stuff.
 \item Every finite residuated semigroup has a finite relativised representation.
@@
 (it seems so) $\operatorname{NP}$-complete (that should follow from the results in \cite{pentus2006lambek}).
+The Lambek calculus is decidable that was shown syntactically via cut elimination and subformula property.
+Here we would have an alternative way of showing decidability for some substructural logics.
 \end{enumerate}
@@
 
-\subsection{The finite base property for ${\bf RJSOS}$}
-
-\subsection{The finite algebra property for ${\bf RJSOS}$}
-
-\subsection{The finite algebra on finite base for ${\bf RJSOS}$}
+\subsection{The finite algebra on finite base for ${\bf RJSOS}$ (or its failure)}
 

### 2020-06-06 tensor-product-stuff/Text.tex
initial commit for tensor product stuff

--- 2020-05-31 problem-9.4/Text.tex
@@
 \date{}
-\title{Model-theoretic aspects of relativised cylindric set algebras}
+\title{Stuff related to tensor products of modal algebras}
 
@@
 
-\section{Model-theoretic and universal algebraic preliminaries}
+\section{Preliminaries}
 
-\subsection{Ultraproducts}
-
-Here are the required notions and facts from model theory and universal algebra \cite{hirsch2002relation} \cite{hodges1997shorter} \cite{sankappanavar1981course}.
-
-Let $A$ be a non-empty set, an \emph{ultrafilter} on $A$ is a set of subsets $U \subseteq \mathcal{P}(\mathcal{P})(A)$ such that $A$ is closed under intersections, $\subseteq$-upwardly closed, and either $X \in U$ or $- X \in U$, where $X \subseteq A$.
-An ultrafilter is called principal if it has the form $\uparrow X = \{ Y \in \mathcal{P}(A) \: | \: X \subseteq Y \}$.
-
-Let $\Omega = \langle \operatorname{Cnst}, \operatorname{Fn}, \operatorname{Pred} \rangle$ be a signature and $\Lambda$ an index set, and let $\{ M_{\lambda}\}_{\lambda \in \Lambda}$ be an indexed set of $\Omega$ structures.
-The $\Omega$-structure
-\begin{center}
-  $M = \prod \limits_{\lambda \in \Lambda} M_{\lambda}$
-\end{center}
-is called a \emph{product} that defined as follows.
-Its domain is the Cartesian product of the domains of $M_{\lambda}$.
-$a \in M$ is a function $\Lambda \to \bigcup \limits_{\lambda \in \Lambda} \operatorname{dom}(M_{\lambda})$ such that $a(\lambda) \in M_{\lambda}$ for each $\lambda \in \Lambda$.
-Given $\lambda \in \Lambda$ and $a_{\lambda} \in M_{\lambda}$, we denote the function mapping $\lambda$ to $a_{\lambda}$ as
-$\langle a_{\lambda} \: | \: \lambda \in \Lambda \rangle$. We define the interpretation of $\Omega$-symbols as
-
-\begin{enumerate}
-  \item If $c \in \operatorname{Cnst}$, then $c^{M} = \langle c^{M_{\lambda}} \: | \: \lambda \in \Lambda \rangle$
-  \item If $f \in \operatorname{Fn}$ is an $n$-ary function symbol, then $f^{M}(\overline{a}) = \langle f^{M_{\lambda}}(\overline{a}) \: | \: \lambda \in \Lambda \rangle$, where $\overline{n} \in M^{n}$
-  \item If $R \in \operatorname{Pred}$ is an $n$-ary relation symbol and $\overline{n} \in M^{a}$, then $R^{M}(\overline{a}) = \langle R^{M_{\lambda}}(\overline{a}) \: | \: \lambda \in \Lambda \rangle$
-\end{enumerate}
-
-Given $\lambda \in \Lambda$, we define the $\lambda$th projection as $\pi_{\lambda} : M \to M_{\lambda}$ such that $\pi_{\lambda}(a) = a(\lambda)$.
-
-Let $\Lambda$ be an index set and $D$ an ultrafilter on the Boolean algebra $\langle \mathcal{P}(\Lambda), \cup, -, \Lambda, \emptyset \rangle$.
-Consider the product $M = \prod \limits_{\lambda \in \Lambda} M_{\lambda}$ of the $\Omega$-structures $\{ M_{\lambda}\}_{\lambda \in \Lambda}$ and the equivalence relation on $\operatorname{dom}(M)$ defined as
-\begin{center}
-  $a_1 \sim a_2 \Leftrightarrow \{ \lambda \in \Lambda \: | \: a_1(\lambda) = a_2(\lambda)\} \in D$
-\end{center}
-Let us denote $\operatorname{dom}(M) / \sim$ as $U$ and $[a]_{\sim}$ as $a / D$, where $a \in \operatorname{dom}(M)$.
-We also denote the \emph{ultraproduct} of $\{ M_{\lambda} \}_{\lambda}$ as $\prod \limits_{\lambda \in \Lambda} M_{\lambda} / D$, or,
-for brevity, as $\prod_{D} M_{\lambda}$. The $\Omega$-symbols have the following interpretation
-\begin{enumerate}
-  \item If $c \in \operatorname{Cnst}$, then $c^{U} = c^{M} / D$
-  \item If $f \in \operatorname{Fn}$ is an $n$-ary function symbol and $\overline{a} \in M^{n}$, then $f^{U}(\overline{a}) = f^{M}(x) = f^{M}(\overline{a}) / D$
-  \item If $R \in \operatorname{Fn}$ is an $n$-ary relation symbol and $\overline{a} \in M^{n}$, then $U \models R(\overline{a} / D)$ iff
-  $\{ \lambda \in \Lambda \: | \: M_{\lambda} \models R(\overline{a} (\lambda))\} \in D$
-\end{enumerate}
-The ultraproduct is principal if $D$ is a principal filter.
-
-\begin{defin}
-$ $
-
-  \begin{enumerate}
-    \item Let $\{ M_{\lambda}\}_{\lambda \in \Lambda}$ be a set of $\Omega$-structures such that every
-    $M_{\lambda}$ is isomorphic to the single structure $M$, then their ultraproduct over $D$ is called the ultrapower over $D$.
-    The denotation is $\prod \limits_{D} M$ or $M^{\Lambda} / D$.
-    \item If $\prod \limits_{D} M \cong N$ for some structure $N$, then $M$ is an ultraroot of $N$.
-  \end{enumerate}
+\begin{defin} A normal modal logic
 \end{defin}
 
-\begin{theorem}[\L{}os]
-
-  Let $\{ M_{\lambda}\}_{\lambda \in \Lambda}$ be $\Omega$-structures and $D$ an ultrafilter on $\Lambda$, and let
-  $U = \prod \limits_{D} M_{\lambda}$ be an ultraproduct of $\{ M_{\lambda}\}_{\lambda \in \Lambda}$ over $D$. For each first-order formula
-  $\varphi(x_1, \dots, x_n)$ and for each $a_1 / D, \dots, a_n / D \in U$:
-
-\begin{center}
-  $U \models \varphi(a_1 / D, \dots, a_n / D)$ iff $\{ \lambda \in \Lambda \: | \: \varphi(a_1(\lambda), \dots, a_n(\lambda))\} \in D$
-\end{center}
-\end{theorem}
-
-The \L{}os has the following helpful corollary:
-
-\begin{col}
-  Let $\prod \limits_{D} M$ be an ultrapower of $M$. For $a \in M$, let us define a function $\overline{a} : a \mapsto a / D$.
-  Then such a map is an elementary embedding of $M$ into $\prod \limits_{D} M$.
-\end{col}
-
-Moreover, any elementary equivalent structures have isomorphic ultrapowers.
-
-Recall that a class of $\Omega$-structures ${\bf K}$ is called \emph{elementary}, if ${\bf K} = \operatorname{Mod}(\operatorname{T})$ for some first-order theory ${\bf T}$.
-In that case, $\operatorname{T}$ is an axiomatisation of ${\bf K}$.
-
-\begin{theorem}
-  Let ${\bf K}$ be a class $\Omega$-structures,
-  ${\bf K}$ is elementary iff ${\bf K}$ is closed under isomorphic copies, ultraroots, and ultrapowers.
-\end{theorem}
-
-\subsection{Preliminaries from universal algebra}
-
-\begin{defin}
-  Let ${\bf K}$ be a class of $\Omega$-structures, then ${\bf K}$ is a variety, if it is defined by some set of equations.
-  The variety generated by ${\bf K}$ is the smallest variety contatining ${\bf K}$. ${\bf K}$ is a quasi-variety, if it is defined by some
-  set of quasi-identities.
+\begin{defin} A Kripke frame, a Kripke model, a general frame
 \end{defin}
 
-Given a class ${\bf K}$ of $\Omega$-structures, then ${\bf I}({\bf K})$, ${\bf S}({\bf K})$, ${\bf H}({\bf K})$, and ${\bf P}({\bf K})$ are
-the classes of isomorphic copies, algebras isomorphic to subalgebras belonging to ${\bf K}$,
-algebras isomorphic to homomorphic images belonging to ${\bf K}$, and algebras isomorphic to direct products belonging to ${\bf K}$.
-We claim that ${\bf I}({\bf K}) \subseteq {\bf S}({\bf K})$. ${\bf Up}({\bf K})$ is the class of algebras isomorphic to ultraproducts
-belonging to ${\bf K}$.
-
-\begin{theorem} Let ${\bf K}$ be a class of $\Omega$-structures
-
-  \begin{enumerate}
-    \item ${\bf K}$ is a variety iff ${\bf H}({\bf K}), {\bf S}({\bf K}), {\bf P}({\bf K}) \subseteq {\bf K}$
-    \item ${\bf HSP}({\bf K}) = {\bf H}({\bf S}({\bf P}({\bf K})))$ is the smallest variety containg ${\bf K}$
-    \item ${\bf K}$ is a quasi-variety iff it is closed under subalgebras, products, and ultraproducts, iff ${\bf SPUp}({\bf K}) = {\bf K}$.
-  \end{enumerate}
-\end{theorem}
-
-\subsection{Subdirect products}
-
-\begin{defin}
-  $ $
-
-  \begin{enumerate}
-    \item Let $\{ A \}_{i \in I}$ be $\Omega$-structures, a subdirect product of $\langle A_i \: | \: i \in I \rangle$ is a
-    subalgebra $B$ of $\prod \limits_{i \in I} A_i$ such that for each $i \in I$, a projection map $\pi_i : B \to A_i$ is a surjection.
-    \item A subdirect representation of an $\Omega$-structure is an embedding $f : A \to \prod \limits_{i \in I} A_i$ for some $I$ and
-    $\{ A_i \}_{i \in I}$ such that $f \circ \pi_i : A \to A_i$ is a surjection.
-    \item An $\Omega$-structure $A$ is subdirectly irreducible if for every subdirect representation $f : A \to \prod \limits_{i \in I} A_i$
-    there exists a projection $\pi_i$ such that $f \circ \pi_i$ is an isomorphism.
-    \item ${\bf Sir}({\bf K})$ is the class of subdirectly irreducible structures belonging to ${\bf K}$.
-    \item A subdirect decomposition of $A$ if there exists a subdirect representation $f : A \to \prod \limits_{i \in I} A_i$ such that
-    every $A_i$ is subdirectly irreducible.
-  \end{enumerate}
+\begin{defin} A modal algebra/a complex algebra
 \end{defin}
 
-It is known that every Boolean algebra with operators has a subdirect demcomposition. Moreover, that implies:
-\begin{theorem}
-$ $
-
-  \begin{enumerate}
-  \item If ${\bf K}$ is a variety, then every element of ${\bf K}$ has a subdirect decomposition with some subdirect irreducible elements of
-  ${\bf K}$.
-  \item If ${\bf K}$ is a variety and $\varepsilon$ is an equation,
-  ${\bf Sir}({\bf K}) \models \varepsilon \Leftrightarrow {\bf K} \models \varepsilon$.
-\end{enumerate}
-\end{theorem}
-
-\subsection{Pseudo-elementary classes}
-
-\section{Cylindric algebras}
-
-\subsection{(Representable) cylindric algebras and cylindric set algebras}
-
-Let $\alpha$ be an ordinal. Let $U^{\alpha}$ be the set of all functions mapping $\alpha$ to a non-empty set $U$. We denote $x(i) = x_i$ for
-$x \in U^{\alpha}$ and $i < \alpha$.
-
-\begin{defin}
-$ $
-
-  \begin{enumerate}
-    \item A subset of $U^{\alpha}$ is an $\alpha$-ry relation on $U$. For $i, j < \alpha$, the $i,j$-diagonal $D_{ij}$ is the set of all
-    elements of $U$ such that $y_i = y_j$. If $i < \alpha$ and $X$ is an $\alpha$-ry relation on $U$, then
-    the $i$-th cylindrification $C_i X$ is the set of all elements of $U$ that agree with some element of $X$ on each coordinate except the
-    $i$-th one. To be more precise,
-    $C_i X = \{ y \in U^{\alpha} \: | \: \exists x \in X \forall i < \alpha \: (i \neq j \Rightarrow y_j = x_j)\}$.
-    \item A cylindic set algebra of dimension $\alpha$ is an algebra consisting of a set $S$ of $\alpha$-ry relation on some base set $U$
-    with the constants and operations $0 = \emptyset$, $1 = U^{\alpha}$, $\cap$, $-$, the diagonal elements $\{ D_{ij} \}_{i, j < \alpha}$,
-    the cylindrifications $\{ C \}_{i < \alpha}$. A generalised cylindric set algebra of dimension $\alpha$ is a subdirect of
-    cylindric algebras that have dimension $\alpha$
-    \item A cylindric algebra of dimension $\alpha$ is an algebra $\mathcal{C} = \langle \mathcal{B}, \{ c_i \}_{i < \alpha}, \{ d_{ij} \}_{i, j < \alpha} \rangle$ such that
-    \begin{itemize}
-      \item $\mathcal{B}$ is a Boolean algebra, for each $i, j < \alpha$ $c_i$ is an operator and $d_{ij} \in \mathcal{B}$
-      \item For each $i < \alpha$, $a \leq c_i a$, $c_i (a \land c_i b) = c_i a \land c_i b$ and $d_{ii} = 1$
-      \item For every $i, j < \alpha$, $c_i c_j a = c_j c_i a$
-      \item If $k \neq i, j < \alpha$, then $d_{ij} = c_k (d_{ij} \land d_{jk})$
-      \item If $i \neq j$, then $c_i (d_{ij} \land a) \land c_i (d_{ij} \land - a) = 0$
-    \end{itemize}
-    ${\bf CA}_{\alpha}$ is the class of all cylindric algebras of dimension $\alpha$
-    \item An $\alpha$-dimensional cylindric algebra $C$ is representable, if it is isomorphic to a generalised cylindric set algebra
-    of dimension $\alpha$. Such is isomorphism is a representation of $C$.
-    ${\bf RCA}_{\alpha}$ is the class of all representable cylindric algebras that have dimension $\alpha$.
-  \end{enumerate}
+\begin{defin} the logic of a frame/class of frames.
 \end{defin}
 
-\subsection{Substitution in cylindric algebras}
-
-\begin{defin}
-  Given a cylindric algebra of dimension $\alpha$ $C$, let $x$ be a term of its signature, the substitution operator $s^{i}_{j}$ have the following definition:
-  \begin{center}
-  $s^{i}_{j} x = \begin{cases} x, \text{if } i = j \\ c_i (d_{ij} \land x), \text{otherwise} \end{cases}$
-  \end{center}
+\begin{defin} product of frames/product of logics
 \end{defin}
 
-\begin{prop} Let $\alpha$ be an ordinal and let $i, j, k, l < \alpha$. The following facts hold in ${\bf CA}_{\alpha}$
+\begin{defin} discrete duality between perfect modal algebras and Kripke frames
+\end{defin}
 
-  \begin{enumerate}
-    \item $s^{i}_j x \leq c_i x$.
-    \item $s^{i}_j (x \land y) = s^{i}_j x \land s^{i}_j y$, $s^{i}_j (x \lor y) = s^{i}_j x \lor s^{i}_j y$, $- s^{i}_j x = s^{i}_j (- x)$.
-    Moreover, $s^{i}_j$ is completely additive.
-    \item $i \neq k, l$ implies $s^i_j d_{ik} = d_{jk}$ and $s^i_j d_{kl} = d_{kl}$.
-    \item $d_{jk} \land s^{i}_j = d_{jk} \land s^{i}_k$.
-    \item $s^i_j c_i x = c_i x$.
-    \item $k \neq i,j$ implies $s^i_j c_i x = c_i s^i_j x$.
-    \item $c_j s^i_j x = c_i s^j_i x$.
-    \item $i \neq j$ implies $c_i s^i_j x = s^i_j x$.
-    \item $i \neq k$ implies $s^i_j s^i_k = s^i_k x$.
-    \item If either $i \notin \{ k, l\}$ and $k \notin \{ i, j\}$, or $j = l$, then $s^i_j s^k_l x = s^k_l s^i_j x$.
-    \item $s^i_j s^j_i x = s^i_j x$.
-    \item $s^i_k s^j_i x = s^i_k s^j_k x = s^j_k s^i_j x$
-  \end{enumerate}
-\end{prop}
+\begin{defin} A finitely axiomatisable normal modal logic
+\end{defin}
 
-\section{${\bf IG}_{\omega}$ and ultraproducts}
+TODO: descibe related underlying results
 
-\section{${\bf IG}_{\omega}$ is (not) a variety; is (not) (pseudo-)elementary}
+\section{Tensor products of modal algebras. The basic definitions and results}
+
+\begin{defin} A commutative associative algebra
+\end{defin}
+
+\begin{defin} Tensor product of them
+\end{defin}
+
+\begin{defin} Tensor product of modal algebras
+\end{defin}
+
+\begin{defin} Tensor product of general frames
+\end{defin}
+
+TODO: descibe related underlying results
+
+\section{Note on incomplete modal logics}
+
+TODO: consider the system containing ${\bf GL}$, McKinsey, seriality, and linearity.
+
+\section{Solution of Problem 1}
+
+\section{Solution of Problem 4}
 

### 2020-06-07 tensor-product-stuff/Text.tex
fix
update tensor stuff

@@
 
-\begin{defin} A normal modal logic
+The basic definitions according to \cite{chagrov1997modal}.
+
+\begin{defin} An $n$-modal algebra is an algebra $\mathcal{B} = \langle B, \vee, \neg, \bot, \Diamond_1, \dots, \Diamond_n \rangle$, where $\mathcal{B} = \langle B, \vee, \neg, \bot \rangle$ is a Boolean algebra and for each $i = 1,\dots,n$ $\Diamond_i$ is an order-preserving unary function obeying:
+  \begin{enumerate}
+    \item $\Diamond_i p \vee \Diamond_i q = \Diamond_i (p \vee q)$
+    \item $\Diamond_i \bot = \bot$
+  \end{enumerate}
 \end{defin}
 
-\begin{defin} A Kripke frame, a Kripke model, a general frame
+\begin{defin} An $n$-normal modal logic is a set of formulas containing Boolean tautologies, formulas $\Diamond_i p \vee \Diamond_i q \leftrightarrow \Diamond_i (p \vee q)$ and $\Diamond_i \bot \leftrightarrow \bot$ for $i = 1,\dots,n$, and is closed under ${\bf MP}$, ${\bf Sub}$, and the monotonicity rule: from $\varphi \to \psi$ infer $\Diamond_i \varphi \to \Diamond_i \psi$.
 \end{defin}
 
-\begin{defin} A modal algebra/a complex algebra
+\begin{defin}
+$ $
+\begin{enumerate}
+  \item A Kripke $n$-frame is a structure $\mathcal{F} = \langle W, R_1, \dots, R_n \rangle$, where $R_i \subseteq W \times W$ for each $i = 1,\dots, n$.
+  \item Let $\mathcal{F}$ be a Kripke frame.
+A Kripke model over $\mathcal{F}$ is a tuple $\mathcal{M} = \langle \mathcal{F}, \vartheta \rangle$, where $\vartheta : \operatorname{PV} \to 2^W$ is a valuation.
+Modalised formulas have the following semantics:
+  \begin{center}
+    $\mathcal{M}, x \models \Diamond_i \varphi \Leftrightarrow \exists y \in R_i(x) \: \mathcal{M}, y \models \varphi$.
+  \end{center}
+  A formula $\varphi$ is true in $\mathcal{M}$ iff $||\varphi||_{\mathcal{M}} = W$.
+$\varphi$ is valid in $\mathcal{F} = \langle W, R_1, \dots, R_n \rangle$ iff $||\varphi||_{\mathcal{M}} = W$ for every model $\mathcal{M}$ over $\mathcal{F}$.
+  \item The complex algebra of a Kripke frame $\mathcal{F} = \langle W, R_1, \dots, R_n \rangle$ is a complete modal algebra defined as $\mathcal{F}^{+} = \langle \mathcal{P}(W), \cup, \neg, \emptyset, R_1^{-1}, \dots, R_n^{-1} \rangle$.
+  \item A general $n$-frame is a structure $\mathcal{F} = \langle W, R_1, \dots, R_n, \mathcal{A} \rangle$, where $\mathcal{A}$ is a subalgebra of $\langle W, R_1, \dots, R_n \rangle^{+}$.
+\end{enumerate}
 \end{defin}
 
-\begin{defin} the logic of a frame/class of frames.
+\begin{defin}
+$ $
+  Let $\mathcal{F}$ be an $n$-frame, then $\operatorname{Log}(\mathcal{F}) = \{ \varphi \: | \: \mathcal{F} \models \varphi \}$. If $\mathbb{F}$ is a class of $n$-frames,
+  then $\operatorname{Log}(\mathbb{F}) = \bigcap \limits_{\mathcal{F} \in \mathbb{F}} \operatorname{Log}(\mathcal{F})$. We use the same notation for logics of general frames and modal algebras.
 \end{defin}
 
-\begin{defin} product of frames/product of logics
+We discuss the background on products of modal logics and products of Kripke frames \cite{kurucz2007combining} \cite{kurucz2003many}.
+
+\begin{defin} Let $\mathcal{F}_1 = \langle W_1, R_1^1, \dots, R_1^n\rangle$ be an $n$-frame and $\mathcal{F}_2 = \langle W_2, R_2^1, \dots, R_2^m\rangle$.
+The product frame of $\mathcal{F}_1$ and $\mathcal{F}_2$ is an $n+m$-frame of the form
+\begin{center}
+$\mathcal{F}_1 \times \mathcal{F}_2 = \langle W_1 \times W_2, R_h^1, \dots, R_h^n, R_v^1, \dots, R_v^m \rangle$
+\end{center}
+such that for all $u_1, u_2 \in W_1$ and for all $v_1, v_2 \in W_2$,
+\begin{center}
+  $\langle u_1, v_1 \rangle R^i_h \langle u_2, v_2 \rangle$ iff $u_1 R_1^i u_2$ and $v_1 = v_2$ for
+  $1 \leq i \leq n$.
+
+  $\langle u_1, v_1 \rangle R^j_v \langle u_2, v_2 \rangle$ iff $u_1 = u_2$ and $u_1 R_2^j u_2$ for $1 \leq j \leq m$.
+\end{center}
 \end{defin}
 
-\begin{defin} discrete duality between perfect modal algebras and Kripke frames
-\end{defin}
+Let $\mathcal{L}_1$ be a normal $n$-modal logic and $\mathcal{L}_2$ a normal $m$-modal logic, the product of $\mathcal{L}_1$ and $\mathcal{L}_2$ is defined as
 
-\begin{defin} A finitely axiomatisable normal modal logic
-\end{defin}
+\begin{center}
+  $\mathcal{L}_1 \times \mathcal{L}_2 = \operatorname{Log}(\operatorname{Frames}(\mathcal{L}_1 ) \times \operatorname{Frames}(\mathcal{L}_2))$
+\end{center}
 
-TODO: descibe related underlying results
+\begin{prop}
+  $ $
+
+  \begin{enumerate}
+    \item Let $\mathcal{L}_1$, $\mathcal{L}_2$ be modal logics, then $\mathcal{L}_1 * \mathcal{L}_2 \subseteq \mathcal{L}_1 \times \mathcal{L}_2$.
+    \item Let $\mathcal{L}_1$, $\mathcal{L}_2$ be Kripke complete modal logics, then $\mathcal{L}_1 \times \mathcal{L}_2 = \operatorname{Log}(\operatorname{Frames}_r(\mathcal{L}_1) \times \operatorname{Frames}_r(\mathcal{L}_2))$, where $\operatorname{Frames}_r(\mathcal{L}_i) = \{\mathcal{F} \in \operatorname{Frames}(\mathcal{L}_i) \: | \: \mathcal{F} \text{ is rooted }\}$ for $i = 1,2$.
+  \end{enumerate}
+\end{prop}
+
+\subsection{Axiomatising products}
+
+The following properties hold for a product frame having the form $\mathcal{F} = \langle W, R_1, \dots, R_n \rangle$:
+
+\begin{enumerate}
+  \item commutativity 1
+  \item commutativity 2
+  \item confluence
+\end{enumerate}
 
@@
 
-TODO: descibe related underlying results
+TODO: describe related underlying results
 
@@
 
+\begin{defin} A finitely axiomatisable normal modal logic
+\end{defin}
+
 \bibliographystyle{plain}

### 2020-06-08 tensor-product-stuff/Text.tex
update tensor stuff

@@
 \end{defin}
+This operation on Kripke frames commutes with disjoint unions, $p$-morphic images, and generated subframes.
 
@@
 \begin{enumerate}
-  \item commutativity 1
-  \item commutativity 2
-  \item confluence
+  \item (left commutativity) $\forall x, y, z \in W \: (x R_j y \: \& \: y R_i z \Rightarrow \exists u \in W \: (x R_i u \: \& \: u R_j z))$
+  \item (right commutativity) $\forall x, y, z \in W \: (x R_i y \: \& \: y R_j z \Rightarrow \exists u \in W \: (x R_j u \: \& \: u R_i z))$
+  \item (Confluence) $\forall x, y, z \in W (x R_j y \: \& \: x R_i z \Rightarrow \exists u \in W (y R_i u \: \& \: z R_j u))$
 \end{enumerate}
+
+The properties are expressed as modal formulas as well:
+\begin{enumerate}
+  \item ${\bf comm}^{\bf l}_{ij} = \Diamond_j \Diamond_i p \to \Diamond_i \Diamond_j p$
+  \item ${\bf comm}^{\bf r}_{ij} = \Diamond_i \Diamond_j p \to \Diamond_j \Diamond_i p$
+  \item ${\bf cr}_{ij} = \Diamond_i \Box_j p \to \Box_i \Diamond_j p$
+\end{enumerate}
+
+\begin{defin} Given unimodal modals logics $\mathcal{L}_1, \dots, \mathcal{L}_n$, the commutator
+
+  \begin{center}
+    $[\mathcal{L}_1, \dots, \mathcal{L}_n]$
+  \end{center}
+  is the smallest $n$-modal logic containing $\mathcal{L}_i$ and axioms ${\bf comm}^{\bf l}_{ij}$, ${\bf comm}^{\bf r}_{ij}$, and ${\bf cr}_{ij}$ for $i, j \in \{ 1, \dots, n\}$ with $i \neq j$.
+\end{defin}
+
+Since ${\bf comm}^{\bf l}_{ij}$, ${\bf comm}^{\bf r}_{ij}$, and ${\bf comm}^{\bf r}_{ij}$ are Salqvist formulas, one has
+
+\begin{prop}
+  If $\mathcal{L}_1, \dots, \mathcal{L}_n$ are canonical, then $[\mathcal{L}_1, \dots, \mathcal{L}_n]$ is canonical, and thus, elementary and Kripke complete.
+\end{prop}
+
+Moreover, one has
+
+\begin{prop}
+  $[\mathcal{L}_1, \dots, \mathcal{L}_n] \subseteq \mathcal{L}_1 \times \dots \times \mathcal{L}_n$.
+\end{prop}
+
+If the converse inclusion holds, then $\mathcal{L}_1, \dots, \mathcal{L}_n$ are \emph{product-mathcing}, see \cite{gabbay1998products}.
+The examples of product-matching logics are Horn axiomatisable ones \cite{kurucz2003many}.
+In particular, the following equality holds for Kripke complete and Horn axiomatisable logics $\mathcal{L}_1, \mathcal{L}_2, \mathcal{L}_3$:
+
+\begin{center}
+  $\mathcal{L}_1 \times \mathcal{L}_2 \times \mathcal{L}_3 = (\mathcal{L}_1 \times \mathcal{L}_2) \times \mathcal{L}_3 = \mathcal{L}_1 \times (\mathcal{L}_2 \times \mathcal{L}_3)$
+\end{center}
+
+Generally, this is an open question whether the product of modal logics is associative.
 

### 2020-06-23 filtration-stuff/Text.tex
initial commit for filtration notes

\documentclass[a4paper]{article}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{bussproofs}
\usepackage{mathtools}
\usepackage{verbatim}
\usepackage{dsfont}
\usepackage{mathabx}
\usepackage[all, 2cell]{xy}
\usepackage[all]{xy}
\usepackage{wasysym}
\usepackage{rotating}
\usepackage{geometry}
\usepackage{trfsigns}
\usepackage{cmll}
\usepackage{authblk}
\usepackage{hyperref}
\usepackage{cleveref}
\usepackage{lipsum}
\usepackage{extpfeil}
\usepackage{soul}
\usepackage{graphicx}

\newcommand\mapsfrom{\mathrel{\reflectbox{\ensuremath{\mapsto}}}}

\theoremstyle{defin}
\newtheorem{defin}{Definition}

\theoremstyle{theorem}
\newtheorem{theorem}{Theorem}

\theoremstyle{prop}
\newtheorem{prop}{Proposition}

\theoremstyle{lemma}
\newtheorem{lemma}{Lemma}

\theoremstyle{ex}
\newtheorem{ex}{Example}

\theoremstyle{col}
\newtheorem{col}{Corollary}

\author{Daniel Rogozin}
\date{}
\title{Note on filtration of logics containing {\bf K5}}

\begin{document}

\maketitle

\nocite{*}

\section{Preliminaries}

\begin{defin}
  An $n$-normal modal logic is a set of formulas that contains all Boolean tautologies, formulas $\Diamond_i p \lor \Diamond_i q \leftrightarrow \Diamond_i (p \lor q)$ and $\Diamond_i \bot \leftrightarrow \bot$ for $i \leq n$, and is closed under modus ponens, substitution, and monotonicity: from $\varphi \rightarrow \psi$ infer $\Diamond_i \varphi \rightarrow \Diamond_i \psi$ for $i \leq n$.
\end{defin}

\begin{defin} An $n$-Kripke model is a triple $\mathcal{M} = \langle W, R_1, \dots, R_n, \vartheta \rangle$, where $R_i \subseteq W \times W$, $\vartheta : \operatorname{PV} \to 2^W$, and the connectives have the following semantics:

  \begin{enumerate}
    \item $\mathcal{M}, w \models p \Leftrightarrow w \in \vartheta(p)$
    \item $\mathcal{M}, w \models \not \varphi \Leftrightarrow \mathcal{M}, w \nvDash \varphi$
    \item $\mathcal{M}, w \models \varphi \lor \psi \Leftrightarrow \mathcal{M}, w \models \varphi \text{ or } \mathcal{M}, w \models \psi$
    \item $\mathcal{M}, w \models \Diamond_i \varphi \Leftrightarrow \exists v \in R_i(w) \: \mathcal{M}, v \models \varphi$
  \end{enumerate}
\end{defin}

By ${\bf K5}$ we mean the logic ${\bf K} \oplus {A5}$, where ${A5} = \Diamond p \to \Box \Diamond p$. It is known that ${\bf K5}$ is the modal logic of all Euqlidean frames. A frame is called Euqlidean if for each $x, y, z$ one has $x R y$ and $x R z$

Let $\mathcal{M} = \langle W, R_1, \dots, R_n, \vartheta \rangle$ be a Kripke model and $\Gamma$ a set of formulas closed under subformulas. An equivalence relation $\sim$ is set to have a finite index if the quotient set $W / \sim$ is finite. The equivalence relation $\sim_{\Gamma}$ induced by $\Gamma$ is defined as

\begin{center}
  $w \sim_{\Gamma} v \Leftrightarrow \forall \varphi \in \Gamma \: (\mathcal{M}, w \models \varphi \Leftrightarrow \mathcal{M}, v \models \varphi)$.
\end{center}

If $\Gamma$ is finite, then $\sim_{\Gamma}$ has a finite index. An equivalence relation $\sim$ respects $\sim_{\Gamma}$, if $w \sim v$ implies $w \sim_{\Gamma} v$.

\begin{defin}
  Let $\mathcal{M} = \langle W, R_1, \dots, R_n, \vartheta \rangle$ be a Kripke model and $\Gamma$ be a Sub-closed set formulas. A $\Gamma$-filtration of $\mathcal{M}$ is a model
  $\widehat{\mathcal{M}} = \langle \widehat{W}, \widehat{R_1}, \dots, \widehat{R_n}, \widehat{\vartheta} \rangle$ such that:
  \begin{enumerate}
    \item $\widehat{W} = W / \sim$, where $\sim$ is an equivalence relation having a finite index that respects $\Gamma$
    \item $\widehat{\vartheta}(p) = \{ [x]_{\sim} \: | \: x \in W \: \& \: x \in \vartheta(p)\}$
    \item For each $i \in I$ one has $\widehat{R}_i^{\text{min}} \subseteq \widehat{R}_i \subseteq \widehat{R}_i^{\text{max}}$. $\widehat{R}_{i, \sim}^{\text{min}}$ is the $i$-th minimal filtered relation on $\widehat{W}$ defined as
    \begin{center}
      $\hat{x} \widehat{R}_{i, \sim}^{\text{min}} \hat{y} \Leftrightarrow \exists x' \sim x \: \exists y' \sim y \: x R_i y$
    \end{center}
    $\widehat{R}_{\Gamma,i}^{\text{max}}$ is the $i$-th maximal filtered relation on $\widehat{W}$ induced by $\Gamma$ defined as
    \begin{center}
      $\hat{x} \widehat{R}_{\Gamma,i}^{\text{max}} \hat{y} \Leftrightarrow \forall \Box_i \varphi \in \Gamma (\mathcal{M}, x \models \Box_i \varphi \Rightarrow \mathcal{M}, y \models \varphi)$
    \end{center}
  \end{enumerate}
\end{defin}

If $\Phi$ is finite subset of $\Gamma$ and $\sim = \sim_{\Gamma}$, then $\widehat{M}$ is a definable $\Gamma$-filtration of $\mathcal{M}$ through $\Phi$.

\begin{lemma}
  Let $\Gamma$ be a finite set of formulas closed under subformulas and $\widehat{M}$ a filtration of $\mathcal{M}$ through $\Gamma$, then for each $x \in W$ and for each $\varphi \in \Gamma$ one has
  \begin{center}
    $\mathcal{M}, x \models \varphi \Leftrightarrow \widehat{M}, \hat{x} \models \varphi$
  \end{center}
\end{lemma}

\section{Filtration of Euclidian logics}

\bibliographystyle{plain}
\bibliography{Text}

\end{document}

### 2020-06-23 filtration-stuff/Text.tex
update

@@
 By ${\bf K5}$ we mean the logic ${\bf K} \oplus {A5}$, where ${A5} = \Diamond p \to \Box \Diamond p$.
-It is known that ${\bf K5}$ is the modal logic of all Euqlidean frames.
-A frame is called Euqlidean if for each $x, y, z$ one has $x R y$ and $x R z$
+It is known that ${\bf K5}$ is the modal logic of all Euclidean frames.
+A frame is called Euclidean if for each $x, y, z$, $x R y$ and $x R z$ implies $y R z$.
+
+\begin{prop}
+  Let $\mathcal{F} = \langle W, R \rangle$ be an Euclidean frame.
+
+  \begin{enumerate}
+    \item For each $x, y, z \in W$, $x R y$ and $x R z$ implies either $y R z$ or $z R y$.
+    \item $R \subseteq R ; R$, that is, $R$ is dense.
+    \item For each $x \in W$, $R^{*}(x) = \{ x\} \cup R(R(x))$.
+  \end{enumerate}
+\end{prop}
 
@@
     \begin{center}
-      $\hat{x} \widehat{R}_{\Gamma,i}^{\text{max}} \hat{y} \Leftrightarrow \forall \Box_i \varphi \in \Gamma (\mathcal{M}, x \models \Box_i \varphi \Rightarrow \mathcal{M}, y \models \varphi)$
+      $\hat{x} \widehat{R}_{\Gamma,i}^{\text{max}} \hat{y} \Leftrightarrow \forall \Box_i \varphi \in \Gamma \: (\mathcal{M}, x \models \Box_i \varphi \Rightarrow \mathcal{M}, y \models \varphi)$
     \end{center}
@@
 
-If $\Phi$ is finite subset of $\Gamma$ and $\sim = \sim_{\Gamma}$, then $\widehat{M}$ is a definable $\Gamma$-filtration of $\mathcal{M}$ through $\Phi$.
+If $\Phi$ is finite subset of $\Gamma$ and $\sim = \sim_{\Phi}$, then $\widehat{M}$ is a definable $\Gamma$-filtration of $\mathcal{M}$ through $\Phi$.
+If $\sim = \sim_{\Gamma}$, then such a filtration by means of the definiton above is called \emph{strict}.
 
@@
 
-\section{Filtration of Euclidian logics}
+\begin{defin} Let $\mathbb{F}$ be a class of Kripke frames and $\Gamma$ a finite set of formulas closed under subformulas.
+If for every model $\mathcal{M}$ over $\mathcal{F} \in \mathbb{F}$ there exists a model that is a $\Gamma$-definable filtration of $\mathcal{M}$, then $\mathbb{F}$ admits definable filtration.
+A class of models $\mathbb{M}$ admits definable filtration if for every $\mathcal{M} \in \mathbb{M}$ there exists a model belonging to the same class that is a definable $\Gamma$-filtration of $\mathcal{M}$.
+\end{defin}
+
+\begin{lemma}
+  $ $
+
+\begin{enumerate}
+  \item Let $\mathcal{L}$ be a complete normal modal logic. If $\operatorname{Frames}(\mathcal{L})$ admits filtration, then $\mathcal{L}$ has the finite model property.
+  \item If the class of models $\operatorname{Mod}(\mathcal{L})$ admits filtration, then $\mathcal{L}$ has the finite model property and Kripke complete as well.
+\end{enumerate}
+\end{lemma}
+
+\section{Filtration of Euclidean logics}
+
+First of all, let us ensure that a filtration of an Euclidean frame is not necessary finite.
+Let $[x] \sim_{\Gamma} [y]$ and $[x] \sim_{\Gamma} [z]$. Then for some $x' \in [x]$ $y' \in [y]$, one has
+$x' R y'$ and $x'' R z'$ for some $x'' \in [x]$ and $z' \in [z]$. Clearly, we cannot claim that $x' = x''$ in general. Thus, minimal filtration does not preserve the required property.
 

### 2020-06-23 tensor-product-stuff/Text.tex
update

--- 2020-06-07 tensor-product-stuff/Text.tex
@@
 
-\begin{defin} Let $\mathcal{F}_1 = \langle W_1, R_1^1, \dots, R_1^n\rangle$ be an $n$-frame and $\mathcal{F}_2 = \langle W_2, R_2^1, \dots, R_2^m\rangle$.
-The product frame of $\mathcal{F}_1$ and $\mathcal{F}_2$ is an $n+m$-frame of the form
-\begin{center}
-$\mathcal{F}_1 \times \mathcal{F}_2 = \langle W_1 \times W_2, R_h^1, \dots, R_h^n, R_v^1, \dots, R_v^m \rangle$
-\end{center}
-such that for all $u_1, u_2 \in W_1$ and for all $v_1, v_2 \in W_2$,
-\begin{center}
-  $\langle u_1, v_1 \rangle R^i_h \langle u_2, v_2 \rangle$ iff $u_1 R_1^i u_2$ and $v_1 = v_2$ for
-  $1 \leq i \leq n$.
-
-  $\langle u_1, v_1 \rangle R^j_v \langle u_2, v_2 \rangle$ iff $u_1 = u_2$ and $u_1 R_2^j u_2$ for $1 \leq j \leq m$.
-\end{center}
+\begin{defin} Let $\mathcal{F}_1 = \langle W_1, R^1_1, \dots, R^1_n\rangle$ be an $n$-frame and $\mathcal{F}_2 = \langle W_2, R^2_1, \dots, R^2_m\rangle$ an $m$-frame
 \end{defin}
 
-Let $\mathcal{L}_1$ be a normal $n$-modal logic and $\mathcal{L}_2$ a normal $m$-modal logic, the product of $\mathcal{L}_1$ and $\mathcal{L}_2$ is defined as
+\begin{defin} discrete duality between perfect modal algebras and Kripke frames
+\end{defin}
 
-\begin{center}
-  $\mathcal{L}_1 \times \mathcal{L}_2 = \operatorname{Log}(\operatorname{Frames}(\mathcal{L}_1 ) \times \operatorname{Frames}(\mathcal{L}_2))$
-\end{center}
+\begin{defin} A finitely axiomatisable normal modal logic
+\end{defin}
 
-\begin{prop}
-  $ $
-
-  \begin{enumerate}
-    \item Let $\mathcal{L}_1$, $\mathcal{L}_2$ be modal logics, then $\mathcal{L}_1 * \mathcal{L}_2 \subseteq \mathcal{L}_1 \times \mathcal{L}_2$.
-    \item Let $\mathcal{L}_1$, $\mathcal{L}_2$ be Kripke complete modal logics, then $\mathcal{L}_1 \times \mathcal{L}_2 = \operatorname{Log}(\operatorname{Frames}_r(\mathcal{L}_1) \times \operatorname{Frames}_r(\mathcal{L}_2))$, where $\operatorname{Frames}_r(\mathcal{L}_i) = \{\mathcal{F} \in \operatorname{Frames}(\mathcal{L}_i) \: | \: \mathcal{F} \text{ is rooted }\}$ for $i = 1,2$.
-  \end{enumerate}
-\end{prop}
-
-\subsection{Axiomatising products}
-
-The following properties hold for a product frame having the form $\mathcal{F} = \langle W, R_1, \dots, R_n \rangle$:
-
-\begin{enumerate}
-  \item commutativity 1
-  \item commutativity 2
-  \item confluence
-\end{enumerate}
+TODO: describe related underlying results
 
@@
 
-\begin{defin} A finitely axiomatisable normal modal logic
-\end{defin}
+\newpage
 

### 2020-06-24 filtration-stuff/Text.tex
started sections with clusters
update clusters stuff

@@
 
+\subsection{Clusters}
+
+Let $\mathcal{F} = \langle W, R \rangle$ be a transitive frame.
+Let us put $x R^{\bullet} y \Leftrightarrow x R y \: \& \: \neg (x R y)$. A point $x$ is proper if $x R x$. Let us define the following
+equivalence relation:
+
+\begin{center}
+  $x \equiv y \Leftrightarrow x R y \: \& \: y R x \lor x = y$.
+\end{center}
+A cluster is an element of the quotient set $W / \equiv$. Given $x \in W$, $C_x$ is a cluster containing $x$. Thus $C_x =
+\{ x\} \cup \{ y \: | x R y x \}$. The original relation lifts to the antisymmetric transitive relation on $W / \equiv$ defined as
+$C_x R C_y$ iff $x R y$. A cluster $C$ is called maximal if $C R C'$ implies $C = C'$. A point is $R$-maximal if $C_x$ is a maximal cluster,
+that is, $R^{\bullet}(x) = \emptyset$.
+A degenerated cluster is a singleton $\{ x \}$ with $\neg (x R x)$.
+A cluster is called simple if it has the form $\{ x \}$ with $x R x$.
+If $\langle W', R' \rangle$ is an inner substructure of $\langle W, R \rangle$, then every $R'$-cluster is an $R$-cluster and every $R$-cluster that intersects $W'$ is a subset of $W'$ and is an $R'$-cluster itself.
+Given a Kripke model $\mathcal{M}$, a set of formulas $\Gamma$ is satisfied by a cluster $C$ if every member of $\Gamma$ is true at some point of $C$.
+
+\section{Transitive closure stuff}
+
+\section{PDLisation of Euclidean logics}
+
+\section{Transitive closure and fusion}
+
 \bibliographystyle{plain}

### 2020-07-02 problem-19.17/Text.tex
problem-19.17

@@
 \end{center}
+Moreover, ${\bf RRA}$ is a variety, but it cannot be defined by any set of first-order formulas \cite{monk1964representable} \cite{}.
 
@@
 
-\begin{defin} A join-semilattice ordered semigroup (${\bf ASOS}$) is an algebra $\mathcal{S} = \langle S, \cdot, \vee \rangle$ such that $\langle S, \cdot \rangle$ is a semigroup, $\langle S, \vee \rangle$ is a join-semilattice and the following equations hold for each $a, b.
+\begin{defin} A join-semilattice ordered semigroup (${\bf OS^{\vee}}$) is an algebra $\mathcal{S} = \langle S, \cdot, \vee \rangle$ such that $\langle S, \cdot \rangle$ is a semigroup, $\langle S, \vee \rangle$ is a join-semilattice and the following equations hold for each $a, b.
 c \in S$:
@@
 \end{defin}
-This class is clearly a variety since ${\bf ASOS}$ has the equational definition so far as $\vee$ is defined as an associative, idempotent, and commutative operation.
+This class is clearly a variety since ${\bf OS^{\vee}}$ has the equational definition so far as $\vee$ is defined as an associative, idempotent, and commutative operation.
 
 Let $A$ be a set of binary relations on some base set $W$ such that $R = \cup A$ is transitive and $\{ x, y \: | \: x R y \} = W$ as in Definition~\ref{rrs}.
-A relation join-semilattice ordered semigroup (${\bf RJSOS}$) is an algebra of binary relations $\mathcal{A} = \langle A, ;, \cup\rangle$ such that $;$ is a relation composition as above and $\cup$ is the set-theoretic union.
-It is known that the class of all representable join-semilattice ordered semigroups has no finite axiomatisation \cite{andreka1989union}.
+A relation join-semilattice ordered semigroup (${\bf ROS^{\vee}}$) is an algebra of binary relations $\mathcal{A} = \langle A, |, \cup \rangle$ such that $;$ is a relation composition as above and $\cup$ is the set-theoretic union.
+Let us define a representation of
+\begin{defin}
+  Let $\mathcal{S} = \langle S, \cdot, \vee \rangle$ a join-semilattice ordered semigroup. A representation of $\mathcal{S}$ is a map
+  $h : S \to 2^{D \times D}$ for some base set $D$ satisfying the following conditions
+  \begin{itemize}
+    \item $h(a) = h(b) \Rightarrow a = b$
+    \item $h(a \vee b) = h(a) \vee h(b)$
+    \item $h(a ; b) = h(a) | h(b) = \{ (e,d) \in D \times D \: | \: \exists f \in D \: (e,f) \in h(a) \: \& \: (f,d) \in h(b) \}$
+  \end{itemize}
+\end{defin}
+It is known that the class of all representable join-semilattice ordered semigroups has no finite axiomatisation \cite{andreka1989union}. In other words,
 
@@
 
+Let us provide a proof of this fact using the rainbow technique \cite{hirsch2002relation} to show that the complement of ${\bf ROS^{\vee}}$ is not closed ultraproducts.
+This is (more or less) a standard way, see \cite{hodkinson2000axiomatizability}.
+We note that representability is not decidable for finite relation algebras \cite{hirsch2001representability}.
+Moreover, representability is undecidable for lattice-ordered semigroups and ordered complemented semigroups \cite{neuzerling2016undecidability}.
+
+First of all, we recall several definitions such as colourings.
+We define a sequence of relation algebras $\{ \mathfrak{A}_n \}_{n < \omega}$ each of which belongs to ${\bf RA}$.
+We need these algebras to show that their $\{;, \vee\}$-reducts are not representable.
+That is, we are seeking to show that
+
+Given $n < \omega$, the set of atoms $\operatorname{At}(\mathfrak{A}_n)$ consists of the following elements:
+\begin{itemize}
+\item identity: ${\bf 1}$, an atom with no colour
+\item greens: ${\bf g}_i$ for $0 \leq i \leq 2^n$
+\item yellow: ${\bf y}$
+\item black: ${\bf b}$
+\item whites: ${\bf b}_{ij}$ for $0 \leq i \leq j \leq 2^n$
+\item reds: ${\bf r}_i$ for $0 <i \leq 2^n$
+\end{itemize}
+We claim that every atom is self-converse ($a^{\smile} = a$). Given $x, y, z \in \mathfrak{A}_n$, a triple $(x,y,z)$ is an  inconsistent triangle if
+\begin{center}
+$x \land (y ; z) = y \land (z ; x) = z \land (x ; y) = 0$
+\end{center}
+We define the set of inconsistent triangles explicitly as follows.
+\begin{itemize}
+  \item $({\bf g}_i, {\bf g}_i, {\bf g}_i)$ for $0 \leq i \leq 2^n$
+  \item $({\bf y}, {\bf y}, {\bf y})$ for $0 \leq i \leq 2^n$
+  \item $({\bf g}_i, {\bf g}_i, {\bf w}_{kj})$ for $0 \leq i \leq 2^n$ and $0 \leq k \leq j \leq 2^n$
+  \item $({\bf r}_i, {\bf r}_j, {\bf r}_k)$ unless $i + k = j$ or $i + k = j$ or $j + k = i$
+  \item $({\bf g}_i, {\bf g}_{i + 1}, {\bf r})$ unless $j = 1$
+  \item $({\bf g}_i, {\bf y}, {\bf w}_{jk})$ unless $i \in \{ j, k \}$
+\end{itemize}
+$({\bf g}_i, {\bf g}_i, {\bf w}_{kj})$ stands for ${\bf g}_i \land ({\bf g}_i ; {\bf w}_{kj}) = {\bf g}_i \land (z ; {\bf g}_i) = {\bf w}_{kj} \land ({\bf g}_i ; {\bf g}_i) = 0$, and so on.
+
+\begin{lemma}
+  For each $n < \omega$, $\mathfrak{A}_n$ does not belong ${\bf RRA}$. The $(\vee, ;)$-reduct $\mathfrak{S}_n$ of $\mathfrak{A}_n$ is not representable as well.
+\end{lemma}
+
 \begin{proof}
-TODO: use games and rainbow stuff
+  See \cite{hodkinson2000axiomatizability} to have a proof that $\mathfrak{A}_n \notin {\bf RRA}$.
+
+  We prove that  
 \end{proof}

### 2020-07-03 problem-19.17/Text.tex
R(;, \vee)-update

@@
 A relation join-semilattice ordered semigroup (${\bf ROS^{\vee}}$) is an algebra of binary relations $\mathcal{A} = \langle A, |, \cup \rangle$ such that $;$ is a relation composition as above and $\cup$ is the set-theoretic union.
-Let us define a representation of
-\begin{defin}
-  Let $\mathcal{S} = \langle S, \cdot, \vee \rangle$ a join-semilattice ordered semigroup. A representation of $\mathcal{S}$ is a map
-  $h : S \to 2^{D \times D}$ for some base set $D$ satisfying the following conditions
-  \begin{itemize}
-    \item $h(a) = h(b) \Rightarrow a = b$
-    \item $h(a \vee b) = h(a) \vee h(b)$
-    \item $h(a ; b) = h(a) | h(b) = \{ (e,d) \in D \times D \: | \: \exists f \in D \: (e,f) \in h(a) \: \& \: (f,d) \in h(b) \}$
-  \end{itemize}
-\end{defin}
+
+Recall that a class of structures ${\bf K}$ is called finitely axiomatisable iff both ${\bf K}$ and its complement are closed
+ultraproducts and isomorphic copies.
+
 It is known that the class of all representable join-semilattice ordered semigroups has no finite axiomatisation \cite{andreka1989union}. In other words,
@@
 \begin{lemma}
-  For each $n < \omega$, $\mathfrak{A}_n$ does not belong ${\bf RRA}$. The $(\vee, ;)$-reduct $\mathfrak{S}_n$ of $\mathfrak{A}_n$ is not representable as well.
+  For each $n < \omega$, $\mathfrak{A}_n$ does not belong ${\bf RRA}$.
+The $(\vee, ;)$-reduct $\mathfrak{S}_n$ of $\mathfrak{A}_n$ is not representable as well.
+For each $n < \omega$, there is an equation valid in set algebras failing in $\mathfrak{S}_n$.
 \end{lemma}
@@
 
-  We prove that  
+  We prove that $\mathfrak{S}_n$ is not representable by contradiction.
+  Suppose $h$ is an isomorphism of $\mathfrak{S}_n$ to a set relation of relations having similarity type $\{ ;, \vee \}$. Let $0$ be a zero element of $\mathfrak{A}_n$.
 \end{proof}
+
+TODO: define games and networks. Take a look at \cite{hirsch1997step}.
+
+\begin{lemma}
+  Any non-trivial ultraproduct of $\{ \mathfrak{A}_n \}_{n < \omega}$ is representable.
+\end{lemma}
+
+\begin{lemma}
+  TODO: one needs to realise when $\exists$ has a winning strategy
+\end{lemma}
 

### 2020-07-03 filtration-stuff/Text.tex
the definition of a game and the definition of a network

@@
 \date{}
-\title{Note on filtration of logics containing {\bf K5}}
+\title{Notes on filtration of logics containing {\bf K5}}
 
@@
 
+If clusters coincide then their poitns have the same theory in the original model:
+
+\begin{lemma}
+  $C_x = C_y$ implies $\mathcal{M}, x \models \Box \varphi \Leftrightarrow \mathcal{M}, y \models \Box \varphi$
+\end{lemma}
+
+Let us describe the bulldozing technique allowing one to eliminate nondegenerated clusters \cite{bull1984basic}.
+Let $\mathcal{L}$ be a transitive logic and $\mathcal{F}$ its frame.
+We construct first a frame $\mathcal{F}^0 = \langle W^{0}, R^{0} \rangle$ replacing every nondegenerated frame $C$ of $W$ by $C^0 = \{ \langle x, i \rangle \: | \: x \in C, i < \omega \}$.
+We also replace each degenerated cluster $C$ by $\{ \langle x,0 \rangle\}$.
+Elements of these subsets form $W^0$. The relation $R^{0}$ is defined as
+\begin{center}
+  $\langle x, i \rangle R^{0} \langle y, j \rangle \Leftrightarrow x R^{\bullet} y \text{ or } (x \equiv y \: \& \: i < j) \text{ or } i = j \: \& \: x <_C y$
+\end{center}
+where $<_C$ is an arbitrary strict ordering on the proper cluster $C$ containing $x$ and $y$.
+
+Each nondegenrated cluster $C$ is replaced by an infinite set $C_0$ such that $\langle C_0, R_0 \rangle$ is a strict linear order.
+Moreover, $\langle y, j \rangle$, a copy of $y$, occurs after $\langle y, j \rangle$, a copy of $x$.
+
+Bulldozing might be extended to models $\mathcal{M} = \langle W, R, \vartheta \rangle$ defining $\vartheta^0$ as follows
+\begin{center}
+  $\vartheta^{0}(p_i) = \{ \langle x, i \rangle \: | \: x \in \vartheta(p_i), i < \omega \}$.
+\end{center}
+
+One may show inductively the following fact.
+\begin{lemma}
+  $\mathcal{M}, x \models \varphi \Leftrightarrow \mathcal{M}^{0}, \langle x, i \rangle \models \varphi$
+\end{lemma}
+
+Let us concretise the case of transitive Euclidean frames. First of all, we consider clusters in ${\bf K}45$ frames.
+
 \section{Transitive closure stuff}

### 2020-07-03 problem-19.17/Text.tex
update
the definition of a game and the definition of a network

@@
 
+Let $\mathcal{A}$ be a relation algebra
+
+\begin{defin}
+  A network is a complete directed finite graph with edges labelled by elements of $\mathcal{A}$.
+Such a graph have the following form.
+$N = \langle E_N, l_N \rangle$, where $E_N = U_N \times U_N$ for some finite base set and $l_n : E_N \to \operatorname{At}(\mathcal{A})$ is function mapping each edge to some atom of $A$.
+This function obey the following requirements:
+  \begin{enumerate}
+    \item $l_N(x,y) \leq {\bf 1}$ iff $x = y$
+    \item $l_N(x, y) ; l_N(y,z) \geq l_N(x,z)$
+  \end{enumerate}
+\end{defin}
+Given two networks $N = \langle E_N, l_N \rangle$ and $N^{'} = \langle E_{N^{'}}, l_{N^{'}} \rangle$, $N$ is a subnetwork of $N^{'}$ ($N \subseteq N^{'}$) if $E_N \subseteq E_{N^{'}}$ and for each $x,y \in U_N$, $l_{N^{'}}(x,y) = l_{N}(x,y)$.
+
+\begin{defin}
+  Let $n < \omega$.
+We define a game $\mathcal{G}_n(\mathcal{A})$ for two players $\forall$ (Abelard) and $\exists$ (H\'{e}lo\"{i}se).
+Abelard and H\'{e}lo\"{i}se build a finite chain of networks $N_0 \subseteq N_1 \dots \subseteq N_n$ as follows.
+In the first round $\forall$ picks an atom $\alpha$ and $\exists$ plays a network $N_0$ containing an edge $(m_0, n_0)$ such that $l_n(m_0, n_0) = \alpha$.
+If $\alpha \leq {\bf 1}$, then $m_0 = n_0$, otherwise $m_0 \neq n_0$.
+If $m_0 \neq n_0$, the edges $(m_0, n_0)$ and $(n_0, m_0)$ belong to Abelard.
+Suppose $N_{i-1}$ for $i < n$ has been played, then
+  \begin{itemize}
+    \item $\forall$ chooses an edge $(m, n) \in E_{N_{i - 1}}$ and atoms $x, y \in \operatorname{At}(\mathcal{A})$ such that $l_{N_{i - 1}}(m,n) \leq x ; y$.
+    \item $\exists$ provides a network $N_i = \langle E_{N_i}, l_{N_i} \rangle \supseteq N_{i - 1}$ such that there exists
+    $l \in U_{N_i}$ such that $l_{N_i}(m, l) = x$ and $l_{N_i}(l, n) = y$.
+  \end{itemize}
+  If $(m,n) \in E_i$ such that $m \neq n$ and $m, n \in U_{N_{i - 1}}$, then the owner of this edge is the same as in the previous round.
+  The edges $(m, l)$ and $(l, n)$ and their converses belong to Abelard.
+The rest of the irreflexive edges belongs to H\'{e}lo\"{i}se.
+$\exists$ wins a match of the game $\mathcal{G}_n(\mathcal{A})$ if she can provide a network $N_i$ for each move of $\forall$ for each $i \leq n$.
+  $\exists$ has a winning strategy if she can win all matches.
+\end{defin}
+
+This lemma has been proved by Hirsch and Hodkinson here \cite{hirsch1997step}:
+
 \begin{lemma}
-  Any non-trivial ultraproduct of $\{ \mathfrak{A}_n \}_{n < \omega}$ is representable.
+  Let $\mathcal{A}$ be an atomic relation algebra.
+Then $\exists$ has a winning strategy in $\mathcal{G}_n({\mathcal{A}})$ for each $n < \omega$ iff $\mathcal{A}$ is elementary equivalent to some completely representable relation algebra.
+If $\exists$ has a winning strategy, then $\mathcal{A}$ is representable since ${\bf RRA}$ is elementary.
+\end{lemma}
+
+\begin{lemma}
+  Any non-trivial ultraproduct of $\{ \mathfrak{A}_n \}_{n < \omega}$ is representable. The same statement for non-trivial ultraproduct of reducts $\{ \mathfrak{S}_n \}_{n < \omega}$.
 \end{lemma}

### 2020-07-04 problem-19.17/Text.tex
fix several mistakes in the definition of a rainbow algebra

@@
 \date{}
-\title{The finite base property for some relation algebras subreducts}
+\title{The finite base property and lacking of finite axiomatisation for some subreducts of representable relation algebras}
 
@@
 \item black: ${\bf b}$
-\item whites: ${\bf b}_{ij}$ for $0 \leq i \leq j \leq 2^n$
+\item whites: ${\bf w}$, ${\bf w}_{ij}$ for $0 \leq i \leq j \leq 2^n$
 \item reds: ${\bf r}_i$ for $0 <i \leq 2^n$
@@
 \begin{itemize}
-  \item $({\bf g}_i, {\bf g}_i, {\bf g}_i)$ for $0 \leq i \leq 2^n$
-  \item $({\bf y}, {\bf y}, {\bf y})$ for $0 \leq i \leq 2^n$
-  \item $({\bf g}_i, {\bf g}_i, {\bf w}_{kj})$ for $0 \leq i \leq 2^n$ and $0 \leq k \leq j \leq 2^n$
+  \item $({\bf 1}, x, y)$ unless $x = y$
+  \item $({\bf g}_i, {\bf g}_k, {\bf g}_l)$ for $i, k, l \leq 2^n$
+  \item $({\bf y}, {\bf y}, {\bf y})$
+  \item $({\bf y}, {\bf y}, {\bf b})$
+  \item $({\bf g}_i, {\bf g}_j, {\bf w}_{kl})$ for $i, j \leq 2^n$ and $k, j \leq 2^n$
   \item $({\bf r}_i, {\bf r}_j, {\bf r}_k)$ unless $i + k = j$ or $i + k = j$ or $j + k = i$
@@
 \end{itemize}
-$({\bf g}_i, {\bf g}_i, {\bf w}_{kj})$ stands for ${\bf g}_i \land ({\bf g}_i ; {\bf w}_{kj}) = {\bf g}_i \land (z ; {\bf g}_i) = {\bf w}_{kj} \land ({\bf g}_i ; {\bf g}_i) = 0$, and so on.
+$({\bf g}_i, {\bf g}_i, {\bf w}_{kj})$ stands for ${\bf g}_i \land ({\bf g}_i ; {\bf w}_{kj}) = {\bf g}_i \land ({\bf w}_{kj} ; {\bf g}_i) = {\bf w}_{kj} \land ({\bf g}_i ; {\bf g}_i) = 0$, and so on.
+
+$C$ is the set of consistent triangles defined as
+
+\begin{center}
+  $C = \{ (a, b, c ) \: | \: a,b,c \in \operatorname{At}(\mathfrak{A}_n), (a,b,c) \text{ is consistent}\}$
+\end{center}
+
+Let us ensure briefly that $\mathfrak{A}_n$ is indeed a relation algebra defining the multiplication table for composition as in \cite{lyndon1950representation}.
+For that, it is enough to show associativity for atoms $a, x, y, z$ such that $a \leq (x ; y) ; z$.
+For that, one needs to find an atom $b$ such that $a \leq x ; b$ and $b \leq y ; z$.
+If $x = y$, then we put $b = {\bf 1}$.
+Otherwise, we try to use a white atom or a black atom for $b$ unless $x, a$ are green and $y, z$ are yellow.
+Suppose $x = {\bf g}_i$, $a = {\bf g}_j$, $y = {\bf y}_k$, and $z = {\bf y}_l$, then we put
 
@@
 
-TODO: define games and networks. Take a look at \cite{hirsch1997step}.
-
-Let $\mathcal{A}$ be a relation algebra
+Let us define networks and games according to \cite{hirsch1997step}.
 
 \begin{defin}
+  Let $\mathcal{A}$ be a relation algebra.
 A network is a complete directed finite graph with edges labelled by elements of $\mathcal{A}$.
@@
 \end{defin}
-Given two networks $N = \langle E_N, l_N \rangle$ and $N^{'} = \langle E_{N^{'}}, l_{N^{'}} \rangle$, $N$ is a subnetwork of $N^{'}$ ($N \subseteq N^{'}$) if $E_N \subseteq E_{N^{'}}$ and for each $x,y \in U_N$, $l_{N^{'}}(x,y) = l_{N}(x,y)$.
+Given two networks $N = \langle E_N, l_N \rangle$ and $N^{'} = \langle E_{N^{'}}, l_{N^{'}} \rangle$, $N$ is a subnetwork of $N^{'}$ ($N \subseteq N^{'}$, or $N^{'}$ refines $N$) if $E_N \subseteq E_{N^{'}}$ and for each $x,y \in U_N$, $l_{N^{'}}(x,y) = l_{N}(x,y)$.
 
@@
 
-This lemma has been proved by Hirsch and Hodkinson here \cite{hirsch1997step}:
+This lemma has been proved by Hirsch and Hodkinson here \cite{hirsch1997step}. This lemma provide a criterion of representability for relation algebras.
 
@@
 \begin{lemma}
-  Any non-trivial ultraproduct of $\{ \mathfrak{A}_n \}_{n < \omega}$ is representable. The same statement for non-trivial ultraproduct of reducts $\{ \mathfrak{S}_n \}_{n < \omega}$.
+  Any non-trivial ultraproduct of $\{ \mathfrak{A}_n \}_{n < \omega}$ is representable, that is, belongs to ${\bf RRA}$.
+The same statement for non-trivial ultraproduct of reducts $\{ \mathfrak{S}_n \}_{n < \omega}$ that belongs to $R(;, \vee)$.
 \end{lemma}
@@
 
-\subsection{The finite algebra on finite base for ${\bf RJSOS}$ (or its failure)}
+\subsection{The finite algebra on finite base for $R(;, \vee)$ (or its failure)}
 

### 2020-07-06 filtration-stuff/Text.tex
strict filtrations for K5

@@
     \item For each $x \in W$, $R^{*}(x) = \{ x\} \cup R(R(x))$.
+    \item $R^{-1} ; R \subseteq R$.
   \end{enumerate}
@@
 
-If $\Phi$ is finite subset of $\Gamma$ and $\sim = \sim_{\Phi}$, then $\widehat{M}$ is a definable $\Gamma$-filtration of $\mathcal{M}$ through $\Phi$.
+If $\Phi$ is finite subset of $\Gamma$ and $\sim = \sim_{\Phi}$, then $\widehat{\mathcal{M}}$ is a definable $\Gamma$-filtration of $\mathcal{M}$ through $\Phi$.
 If $\sim = \sim_{\Gamma}$, then such a filtration by means of the definiton above is called \emph{strict}.
@@
 \begin{lemma}
-  Let $\Gamma$ be a finite set of formulas closed under subformulas and $\widehat{M}$ a filtration of $\mathcal{M}$ through $\Gamma$, then for each $x \in W$ and for each $\varphi \in \Gamma$ one has
+  Let $\Gamma$ be a finite set of formulas closed under subformulas and $\widehat{\mathcal{M}}$ a filtration of $\mathcal{M}$ through $\Gamma$, then for each $x \in W$ and for each $\varphi \in \Gamma$ one has
   \begin{center}
-    $\mathcal{M}, x \models \varphi \Leftrightarrow \widehat{M}, \hat{x} \models \varphi$
+    $\mathcal{M}, x \models \varphi \Leftrightarrow \widehat{\mathcal{M}}, \hat{x} \models \varphi$
   \end{center}
@@
 $x' R y'$ and $x'' R z'$ for some $x'' \in [x]$ and $z' \in [z]$. Clearly, we cannot claim that $x' = x''$ in general. Thus, minimal filtration does not preserve the required property.
+
+\subsection{Strict filtration for logics containing ${\bf K5}$}
+
+\begin{defin}
+  Let $\mathcal{F} = \langle W, R, \vartheta \rangle$ be a Kripke model. Then its Euclidean Horn closure ($R^E$, $E$-closure) is defined as
+  \begin{center}
+    $R^E = (\bigcup \limits_{n < \omega} (R^{-1})^n); R$
+  \end{center}
+  or, equivalently, $R^{E} = \bigcup \limits_{n < \omega} R_n$, where $R^{0} = R$ and $R_{n + 1} = R_n \cup (R^{-1} ; R_n)$.
+\end{defin}
+
+\begin{lemma}
+  Let $\mathcal{M} = \langle W, R, \vartheta \rangle$ be an Euclidean model, $\Gamma$ a set of $\operatorname{Sub}$-closed formulas, and $\sim$ an equivalence relation having a finite index that respects $\Gamma$.
+Then its $E$-closure of the minimal filtration of $R$ is a filtration itself.
+\end{lemma}
+
+\begin{proof}
+  Let us show that $\widehat{R^{E}} = \bigcup \limits_{n < \omega} (R_n)^{\text{min}}_{\sim}$ is a filtration and $\langle W / \sim, \widehat{R}^{E}, \widehat{\vartheta} \rangle$ is an Euclidean model.
+
+  Let $\widehat{\mathcal{M}} = \langle \widehat{W}, \widehat{R}, \widehat{\vartheta} \rangle$ be a minimal filtration of an Euclidean model through $\sim$.
+
+  \begin{enumerate}
+    \item Suppose $x R y$, then $[x] \widehat{R} [y]$ by the definition of a minimal filtration. $\widehat{R}$ is clearly a subrelation of $\widehat{R}^{E}$, thus $[x] \widehat{R}^{E} [y]$.
+    \item Suppose $[x] \widehat{R^{E}} [y]$ and $\mathcal{M}, x \models \Box \varphi$ for $\Box \varphi \in \Gamma$.
+then $([x], [y]) \in \bigcup \limits_{n < \omega} R_n$.
+Then $([x], [y]) \in R_n$ for some $n < \omega$.
+    There are two cases:
+    \begin{enumerate}
+      \item $n = 0$, then $[x] \widehat{R} [y]$, so, obviously, one has $\mathcal{M}, y \models \varphi$
+      \item $n = m + 1$. Suppose $[x] {\widehat{R}}_{m + 1} [y]$. Thus, $[x] \widehat{R}_n \cup (\widehat{R}^{-1} ; \widehat{R}_n) [y]$.
+      There are the following two cases:
+      \begin{enumerate}
+        \item $([x], [y]) \in \widehat{R}_n$, then the statement holds by IH.
+        \item $([x], [y]) \in \widehat{R}^{-1} ; \widehat{R}_n$.
+        Then there exists $[z]$ such that $[x] \widehat{R}^{-1} [z]$ and $[z] \widehat{R}_m [y]$, that is,
+        $[z] \widehat{R} [x]$ and $[z] \widehat{R}_m [y]$. One has $\mathcal{M}, x \models \Box \varphi$, then
+        $\widehat{\mathcal{M}}, [x] \models \Box \varphi$.
+So $\widehat{\mathcal{M}}, [z] \models \Diamond \Box \varphi$.
+On the other hand, $\mathcal{M}, z \models \Diamond \Box \varphi \to \Box \varphi$, so $\mathcal{M}, z \models \Box \varphi$ and,
+        from that, $\widehat{\mathcal{M}}, [z] \models \Box \varphi$, and, thus, $\widehat{\mathcal{M}}, [y] \models \varphi$ and $\mathcal{M}, y \models \varphi$ by IH and the definition of a minimal filtration.
+      \end{enumerate}
+    \end{enumerate}
+  \end{enumerate}
+\end{proof}
+
+\begin{col}
+  ${\bf K}5$ admit strict filtrations.
+\end{col}
 
@@
 
+\subsection{}
+
 \section{Transitive closure stuff}

### 2020-07-08 problem-19.17/Text.tex
strict filtrations for K5
wip: rainbow construction for R(;, \vee)

@@
 \end{center}
-Moreover, ${\bf RRA}$ is a variety, but it cannot be defined by any set of first-order formulas \cite{monk1964representable} \cite{}.
+Moreover, ${\bf RRA}$ is a variety, but it cannot be defined by any set of first-order formulas \cite{monk1964representable}.
 
@@
 Let $A$ be a set of binary relations on some base set $W$ such that $R = \cup A$ is transitive and $\{ x, y \: | \: x R y \} = W$ as in Definition~\ref{rrs}.
-A relation join-semilattice ordered semigroup (${\bf ROS^{\vee}}$) is an algebra of binary relations $\mathcal{A} = \langle A, |, \cup \rangle$ such that $;$ is a relation composition as above and $\cup$ is the set-theoretic union.
+A representable join semilattice-ordered semigroup is an algebra isomorphic to some join semilattice-ordered semigroup having the form $\mathcal{A} = \langle A, |, \cup \rangle$ such that $;$ is a relation composition as above and $\cup$ is the set-theoretic union.
+If $\mathcal{A}$ is representable, then $R(;, \vee)$ Let us recall some of underlying facts about representable join semilattice-ordered semigroups \cite{andreka1991representations}:
+
+\begin{prop}
+  $ $
+
+  \begin{enumerate}
+    \item Let $\mathcal{A} = \langle A, \vee, \cdot \rangle$ be a join semilattice-ordered semigroup such that, for all $a, b \in A$:
+    \begin{enumerate}
+      \item If $a \nleq b$, then there exists an atom $c \leq a$ and $c \nleq b$.
+      \item If $c \leq a \cdot b$ and $c$ is an atom, then there exists an atom $a' \leq a$ such that $c \leq a' \cdot b$.
+    \end{enumerate}
+    then $\mathcal{A}$ is representable.
+    \item Let $\mathcal{A} = \langle A, \cdot \rangle$ be a posemigroup, then $\mathcal{A}$ is representable and such a representation preserve any existing finite suprema and infima, if
+    \begin{enumerate}
+      \item The set of atoms is closed under $\cdot$.
+      \item $\mathcal{A}$ has enough atoms, that is, if $x \in \operatorname{At}(A)$ and $z, w \in A$, then $x \leq z \cdot w$ implies there exist atoms $z_1 \leq z$ and $w_1 \leq w$ such that $x \leq z_1 \cdot w_1$.
+If $z \nleq w$, then there exists an atom $x$ such that $x \leq z$ and $x \nleq w$.
+    \end{enumerate}
+  \end{enumerate}
+\end{prop}
 
@@
 Let us provide a proof of this fact using the rainbow technique \cite{hirsch2002relation} to show that the complement of ${\bf ROS^{\vee}}$ is not closed ultraproducts.
-This is (more or less) a standard way, see \cite{hodkinson2000axiomatizability}.
-We note that representability is not decidable for finite relation algebras \cite{hirsch2001representability}.
+This construction sometimes exploits the similar construction used by Andr{\'e}ka \cite{andreka1991representations} and by Maddux \cite{maddux_1989}.
+We note that representability is not decidable for finite relation algebras \cite{hirsch2001representability} and this result has several generalisations \cite{hirsch2012undecidability}.
 Moreover, representability is undecidable for lattice-ordered semigroups and ordered complemented semigroups \cite{neuzerling2016undecidability}.
+We use(more or less) a standard way of showing that the class of certain reducts of representable relation algebras has no finite axiomatisation, see \cite{hodkinson2000axiomatizability} \cite{hirsch1997step}.
 
@@
 \item whites: ${\bf w}$, ${\bf w}_{ij}$ for $0 \leq i \leq j \leq 2^n$
-\item reds: ${\bf r}_i$ for $0 <i \leq 2^n$
+\item reds: ${\bf r}_i$ for $0 < i \leq 2^n$
 \end{itemize}
@@
 \begin{itemize}
-  \item $({\bf 1}, x, y)$ unless $x = y$
+  \item $(x, y, {\bf 1})$ unless $x = y$
   \item $({\bf g}_i, {\bf g}_k, {\bf g}_l)$ for $i, k, l \leq 2^n$
@@
 \begin{proof}
-  See \cite{hodkinson2000axiomatizability} to have a proof that $\mathfrak{A}_n \notin {\bf RRA}$.
+  See \cite{hodkinson2000axiomatizability} to have a proof that $\mathfrak{A}_n \notin {\bf RRA}$. However, we leave a proof sketch.
 

### 2020-07-11 problem-19.17/Text.tex
problem 19-17 update

@@
 \date{}
-\title{The finite base property and lacking of finite axiomatisation for some subreducts of representable relation algebras}
+\title{The finite base property for some subreducts of representable relation algebras}
 
@@
 
+See this paper to have a proof of the following theorem \cite{goldblatt2006kripke}:
+
+\begin{theorem}
+  Every finite residuated semigroup is isomorphic to some residuated subsemigroup of some finite residuated lattice.
+\end{theorem}
+
 \begin{defin} \label{rrs}
@@
     \item $r / s = \{ \langle a, c \rangle \: | \: \forall b \in W \: (\langle c, b \rangle \in s \Rightarrow \langle a, b \rangle \in r)\}$
+    \item $r \leq s$ iff $r \subseteq s$.
   \end{enumerate}
@@
 
+See \cite{brown1993representation}
+\begin{theorem}
+  Every complete residuated semigroup $\mathcal{A}$ (quantale) is isomorphic to relational complete residuated semigroup on the underlying set of $\mathcal{A}$.
+\end{theorem}
+
+\begin{defin}
+  Let $\mathcal{A} = \langle A, \leq, \cdot, \setminus, / \rangle$ be a residuated semigroup. A representation of $\mathcal{A}$ is an inclusion map $h : \mathcal{A} \to \mathcal{A}'$,
+  where $\mathcal{A}' \in {\bf RRS}$ such that:
+  \begin{enumerate}
+    \item $dom(\mathcal{A}') = \{ \hat{a} \: | \: a \in A\}$, where $\hat{a} = \{ (b, c) \: | \: b \leq a \cdot c \}$.
+  \end{enumerate}
+\end{defin}
+
+Such a map preserves order, residuals, and composition.
 Andr\'{e}ka and Mikul\'{a}s proved the following representation theorem for ${\bf ORS}$ in \cite{andreka1994lambek} that implies relational completeness of the Lambek calculus, the logic of ${\bf ORS}$:
@@
 
-\subsection{The finite base property for ${\bf RRS}$}
-
-\begin{defin}
-  A relativised representation
-\end{defin}
-
-\begin{defin}
-  The standard translation
-\end{defin}
-
-TODO: take a look at relativised representations and loosely guarded fragments in general
-TODO: realise whether it makes sense to use the technique similar to \cite[Theorem 19.13]{hirsch2002relation}
-used for weakly associative algebras.
+\begin{col}
+  Every finite representable residuated semigroup is isomorphic to representable residuated subsemigroup of some finite residuated lattice.
+\end{col}
 
 \begin{theorem}
-  Let $\mathcal{A}$ be a finite residuated semigroup and
-  $|\mathcal{A}| < \omega$, then $\mathcal{A}$ has a finite relativised representation.
+  Let $\mathcal{A} \in {\bf RRS}$ and $|\mathcal{A}| < \omega$, then there exists a set $W$, a set $A$ of binary relations on $W$, $R = \cup A$ with $dom(R) = A$ such that
+  $\mathcal{A} \cong \langle A, |, \setminus, /, \subseteq \rangle$.
 \end{theorem}
@@
 \begin{theorem}
-  Let $\mathcal{A}$ be a finite representable residuated semigroup, then $\mathcal{A}$ is isomorphic to
-  representable residuated semigroup a domain of which is finite.
+  The Lambek calculus is complete w.r.t finite relational models (has the fmp).
 \end{theorem}
-
-\begin{proof}
-  That might follow from the previous theorem, Theorem~\ref{ors=irrs}, and something else.
-\end{proof}
-
-\begin{col}
-  The Lambek calculus has the fmp and the universal theory of ${\bf IRRS}$ is NP-complete.
-\end{col}
-
-The hypothetical plan is the following one:
-\begin{enumerate}
-\item Define properly relativised representation for residuated semigroups, that should look like ternary Kripke frames for the basic Lambek calculus or arrow logic.
-\item Define the standard translation to such first-order relation structures. TODO: take a look at loosely guarded fragment stuff.
-\item Every finite residuated semigroup has a finite relativised representation.
-\item If every $\Pi_1$-statement $\varphi$ of the language of residuated semigroups that is valid in every
-residuated semingroup is valid in algebra having a finite relativised representation (one may use here Theorem~\ref{ors=irrs} somehow), then $\varphi$ is valid in ${\bf ORS}$ as well as in ${\bf IRRS}$.
-\item Every finite residuated semigroup should have a finite relativised representation.
-\item Construct a finitely based relation residuated semigroup from that (an analogue of complex algebra or smth like that). This item is the most non-trivial one.
-\item As a corollary, the first-order universal first-order theory of ${\bf IRRS}$ should be decidable and
-(it seems so) $\operatorname{NP}$-complete (that should follow from the results in \cite{pentus2006lambek}).
-The Lambek calculus is decidable that was shown syntactically via cut elimination and subformula property.
-Here we would have an alternative way of showing decidability for some substructural logics.
-\end{enumerate}
 

### 2020-07-11 filtration-stuff/Text.tex
wip: K5 admit strict filtration

@@
 
+\begin{prop} \label{K5proves}
+  ${\bf K5}$ proves
+  \begin{enumerate}
+    \item $\Box^3 p \leftrightarrow \Box^2 p$
+    \item $\Box^2 \Diamond p \leftrightarrow \Box \Diamond p$
+    \item $\Box \Diamond \Box p \leftrightarrow \Box \Box p$
+    \item $\Box \Diamond^2 p \leftrightarrow \Box \Diamond p$
+  \end{enumerate}
+\end{prop}
+
+\begin{prop}
+  Let $\mathcal{M}$ be a ${\bf K}5$ model, $x R y$ for some $x, y \in W$ then one has
+  \begin{center}
+    $\mathcal{M}, x \models \Diamond \Box \varphi$ iff $\mathcal{M}, y \models \Diamond \Box \varphi$.
+  \end{center}
+\end{prop}
+
+\subsection{Filtrations: general definitions}
+
 Let $\mathcal{M} = \langle W, R_1, \dots, R_n, \vartheta \rangle$ be a Kripke model and $\Gamma$ a set of formulas closed under subformulas.
@@
 
-\subsection{Strict filtration for logics containing ${\bf K5}$}
+\begin{lemma}
+  ${\bf K5}$ admit filtration.
+\end{lemma}
+
+\begin{proof}
+  Let $\mathcal{M}$ be a ${\bf K5}$-model and $\Gamma_0$ a finite set of formulas closed under subformulas.
+  Let us put $\Gamma = \Gamma_0 \cup \operatorname{Sub}(\{ \Diamond \Box \psi \: | \: \Box \psi \in \Gamma_0 \}) \cup \Psi$, where $\Psi = \nabla_1 \nabla_2 \dots \nabla_n \Box \psi$ for $\Box \psi \in \Gamma_0$ and
+  $\nabla_i \in \{ \Diamond, \Box \}$. By Proposition~\ref{K5proves}, any element of $\Phi$ has one of the four forms. Thus, $W \sim_{\equiv_{\Gamma}}$ has a finite index.
+  We put $\widehat{R} = R^{\operatorname{max}}_{\Gamma}$.
+\end{proof}
+
+\begin{defin} A relation $R$ is called Horn, if it is defined with some first-order formula having the form
+
+  \begin{center}
+    $\forall x_1, \dots, x_n (x_{i_1} R x_{j_1} \land \dots \land x_{i_s} R x_{j_s} \rightarrow x_k R x_l)$
+  \end{center}
+\end{defin}
 
 \begin{defin}
-  Let $\mathcal{F} = \langle W, R, \vartheta \rangle$ be a Kripke model. Then its Euclidean Horn closure ($R^E$, $E$-closure) is defined as
-  \begin{center}
-    $R^E = (\bigcup \limits_{n < \omega} (R^{-1})^n); R$
-  \end{center}
-  or, equivalently, $R^{E} = \bigcup \limits_{n < \omega} R_n$, where $R^{0} = R$ and $R_{n + 1} = R_n \cup (R^{-1} ; R_n)$.
+  Let $H$ be a Horn property and $\langle W, R \rangle$ a Kripke frame. A Horn closure of a binary relation $R$ is the minimal relation $R^{H}$ containing $R$ and satisfying $H$.
 \end{defin}
+
+\begin{lemma}
+  $R^{H} = \bigcup \limits_{n < \omega} R_n$ where
+
+  \begin{enumerate}
+    \item $R_0 = R$.
+    \item $R_{n + 1} = R_n \cup \{ (a, b) \in W \: | \: \exists \vec{c} \in W \: P(a, b, \vec{c})\}$, where $P$ is a premise of $H$.
+  \end{enumerate}
+\end{lemma}
+
+$E$-closure (an Euclidean Horn closure of a binary relation) has the following equivalent definitions:
+\begin{lemma}
+  Let $\mathcal{F} = \langle W, R \rangle$ be a Kripke frame.
+  The following conditions are equivalent:
+
+  \begin{enumerate}
+    \item $R^{E}$ is the smallest Euclidean relation containing $R$.
+    \item $R^{E} = \bigcup \limits_{i < \omega} R_i$, where
+    \begin{itemize}
+      \item $R_0 = R$
+      \item $R_{n + 1} = R_n \cup (R^{-1}_n ; R_n)$
+    \end{itemize}
+    \item $x R^E y$ iff there exists $n < \omega$ such that
+    $\exists x_1, \dots, x_n$ with $z_1 R x$ and $z_{n - 1} R y$ and for each $1 < i \leq n$ one has
+    either $z_{i - 1} R z_i$ or $z_i R z_{i - 1}$.
+    \item $R^{E} = R \cup \bigcup \limits_{i < \omega} (R^{-1} ; (R ; R^{-1})^n ; R)$.
+  \end{enumerate}
+\end{lemma}
+
+\begin{proof}
+$ $
+
+\begin{enumerate}
+  \item $(1) \Rightarrow (2)$
+  Let us show that if $R^{E}$ is the smallest Euclidean relation containing $R$, then $R^{E} = \bigcup \limits_{i < \omega} R_i$. There are two inclusions:
+  \begin{itemize}
+    \item $R^{E} \subseteq \bigcup \limits_{i < \omega} R_i$. It is enough to show that
+    $x R^{E} y$ implies $x R^n y$ for some $n < \omega$.
+    \item $\bigcup \limits_{i < \omega} R_i \subseteq R^{E}$.
+    Let us show that $x R_n y$ for each $n < \omega$ implies $x R^E y$ by induction on $n$.
+    If $n = 0$, then $x R y$, thus, $x R^{E} y$, since $R$ is a subrelation of $R^{E}$.
+    Suppose $n = m + 1$ and $x R_{m + 1} y$. Let us show that $x R^{E} y$.
+    From $x R_{m + 1} y$, one has $(x, y) \in R^n \cup (R^{-1}_n ; R_n)$. There are two cases:
+    \begin{itemize}
+      \item $x R^n y$, one needs to merely apply the IH.
+      \item $x R^{-1}_n ; R_n y$. Then $\exists z \in W \: x R^{-1}_n z \: \& \: z R_n$. That is,
+      $z R_n x$ and $z R_n y$ for some $z$. $R_n$ is already a subrelation of $R^E$. Thus,
+      $z R^{E} x$ and $z R^{E} y$. That implies $x R^{E} y$.
+    \end{itemize}
+  \end{itemize}
+  \item $(2) \Rightarrow (3)$
+  Let $x R^E y$, then $(x, y) \in \bigcup \limits_{i < n} R_n$. That is, $(x, y) \in R_n \cup (R_n^{-1} ; R_n)$ for some $n < \omega$. There are two cases.
+  \begin{itemize}
+    \item $(x, y) \in R_n$ and that's it.
+    \item $(x, y) \in R_n^{-1} ; R_n$, then $(z, x), (z, y) \in R_n$.
+  \end{itemize}
+  \item $(3) \Rightarrow (4)$
+  \item $(4) \Rightarrow (1)$
+\end{enumerate}
+\end{proof}
 

### 2020-07-12 problem-19.17/Text.tex
R(;, |): update the definition of a rainbow algebra according to Andreka/Maddux

@@
 
-\nocite{*}
-
 \section{The Relation Algebras Background}
@@
 
-  \begin{enumerate}
-    \item A relation algebra is an algebra $\mathcal{R} = \langle R, 0, 1, \wedge, \vee, \neg, ;, {}^{\smile}, {\bf 1 }\rangle$ such that $\langle R, 0, 1, \wedge, \vee, \neg \rangle$ is a Boolean algebra and the following
+    A relation algebra is an algebra $\mathcal{R} = \langle R, 0, 1, +, -, ;, {}^{\smile}, {\bf 1 }\rangle$ such that $\langle R, 0, 1, +, - \rangle$ is a Boolean algebra and the following
     equations hold, for each $a, b, c \in R$:
@@
       \item $a ; (b ; c) = (a ; b) ; c$
-      \item $(a \vee b) ; c = (a ; c) \vee (b ; c)$
+      \item $(a + b) ; c = (a ; c) + (b ; c)$
       \item $a ; {\bf 1} = a$
       \item $a^{\smile \smile} = a$
-      \item $(a \vee b)^{\smile} = a^{\smile} \vee b^{\smile}$
+      \item $(a + b)^{\smile} = a^{\smile} + b^{\smile}$
       \item $(a ; b)^{\smile} = b^{\smile} ; a^{\smile}$
-      \item $a^{\smile} ; (\neg (a ; b)) \leq \neg b$
+      \item $a^{\smile} ; (- (a ; b)) \leq - b$
     \end{enumerate}
-    where $a \leq b$ iff $a \wedge b = a$ iff $a \vee b = b$. ${\bf RA}$ denotes the class of all relation
+    where $a \leq b$ iff $a + b = b$. ${\bf RA}$ denotes the class of all relation
     algebras.
-    \item A proper relation algebra is an algebra $\mathcal{R} = \langle R, 0, 1, \wedge, \vee, \neg, ;, {}^{\smile}, {\bf 1 }\rangle$ such that $R \subseteq \mathcal{P}(W)$, where $W$ is an equivalence relation; $0 = \emptyset$; $1 = W$; $\wedge$, $\vee$, $\neg$ are set-theoretic intersection, union, and
-    complement respectively; $;$ is relation composition, ${}^{\smile}$ is relation converse, ${\bf 1}$ is a
+\end{defin}
+A relation algebra is called symmetric, if every element is self-converse. A relation algebra is called integral, if
+\begin{center}
+  $a ; b = 0 \Rightarrow a = 0 \text{ or } b = 0$.
+\end{center}
+
+\begin{defin}
+    A proper relation algebra is an algebra $\mathcal{R} = \langle R, 0, 1, \cup, -, |, {}^{\smile}, {\bf 1 }\rangle$ such that $R \subseteq \mathcal{P}(W)$, where $W$ is an equivalence relation; $0 = \emptyset$; $1 = W$; $\cap$, $\cup$, $-$ are set-theoretic intersection, union, and
+    complement respectively; $|$ is relation composition, ${}^{\smile}$ is relation converse, ${\bf 1}$ is a
     diagonal relation restricted to $W$, that is:
     \begin{enumerate}
-    \item $a ; b = \{ \langle x, z \rangle \: | \: \exists y \: \langle x, y \rangle \in a \: \& \: \langle y, z \rangle \in b \}$
+    \item $a | b = \{ \langle x, z \rangle \: | \: \exists y \: \langle x, y \rangle \in a \: \& \: \langle y, z \rangle \in b \}$
     \item $a^{\smile} = \{ \langle x, y \rangle \: | \: \langle y, x \rangle \in a \}$
@@
       ${\bf RRA} = {\bf IPRA}$.
-  \end{enumerate}
 \end{defin}
@@
 \begin{enumerate}
-  \item $a \setminus b = \neg (a^{\smile} ; \neg b)$
-  \item $a / b = \neg (\neg a ; b^{\smile})$
+  \item $a \setminus b = -(a^{\smile} ; -b)$
+  \item $a / b = - (- a ; b^{\smile})$
 \end{enumerate}
@@
 elements belonging to ${\bf Rs}$ to operations of $\tau$.
-By ${\bf Q}(\tau)$ we mean a quasivariety generated by $R(\tau)$.
+By ${\bf Q}(\tau)$ we mean a quasivariety generated by ${\bf R}(\tau)$.
 As in \cite{hirsch2011positive}, we put ${\bf Q}(\tau)$ as the closure of ${\bf R}(\tau)$ under subalgebras and products assuming that ${\bf R}(\tau)$ is already closed under ultraproducts.
@@
 
- A \emph{relation structure} (${\bf RS}$) is an arbitrary algebra of the signature $\Omega = \langle \cdot, \setminus, /, \leq \rangle$, where $\cdot, \setminus, /$ are binary function symbols and $\leq$ is a binary relation symbol.
+ A \emph{relation structure} (${\bf RS}$) is an arbitrary algebra of the signature $\Omega = \langle \circ, \setminus, /, \leq \rangle$, where $\circ, \setminus, /$ are binary function symbols and $\leq$ is a binary relation symbol.
 
 \begin{defin}
-  A residuated semigroup is an algebra $\mathcal{S} = \langle S, \cdot, \leq, \setminus, / \rangle$ such that $\langle S, \cdot, \leq, \rangle$ is an ordered residuated semigroup and the following equivalences hold for each $a, b, c \in S$:
+  A residuated semigroup is an algebra $\mathcal{S} = \langle S, \circ, \leq, \setminus, / \rangle$ such that $\langle S, \circ, \leq, \rangle$ is an ordered residuated semigroup and the following equivalences hold for each $a, b, c \in S$:
 
   \begin{center}
-    $b \leq a \setminus c \Leftrightarrow a \cdot b \leq c \Leftrightarrow a \leq c / b$
+    $b \leq a \setminus c \Leftrightarrow a \circ b \leq c \Leftrightarrow a \leq c / b$
   \end{center}
@@
 \begin{defin}
-  Let $\mathcal{A} = \langle A, \leq, \cdot, \setminus, / \rangle$ be a residuated semigroup. A representation of $\mathcal{A}$ is an inclusion map $h : \mathcal{A} \to \mathcal{A}'$,
+  Let $\mathcal{A} = \langle A, \leq, ;, \setminus, / \rangle$ be a residuated semigroup. A representation of $\mathcal{A}$ is an inclusion map $h : \mathcal{A} \to \mathcal{A}'$,
   where $\mathcal{A}' \in {\bf RRS}$ such that:
   \begin{enumerate}
-    \item $dom(\mathcal{A}') = \{ \hat{a} \: | \: a \in A\}$, where $\hat{a} = \{ (b, c) \: | \: b \leq a \cdot c \}$.
+    \item $dom(\mathcal{A}') = \{ \hat{a} \: | \: a \in A\}$, where $\hat{a} = \{ (b, c) \: | \: b \leq a ; c \}$.
   \end{enumerate}
@@
 
-\begin{defin} A join-semilattice ordered semigroup (${\bf OS^{\vee}}$) is an algebra $\mathcal{S} = \langle S, \cdot, \vee \rangle$ such that $\langle S, \cdot \rangle$ is a semigroup, $\langle S, \vee \rangle$ is a join-semilattice and the following equations hold for each $a, b.
+\begin{defin} A join-semilattice ordered semigroup (${\bf OS^{+}}$) is an algebra $\mathcal{S} = \langle S, ;, + \rangle$ such that $\langle S, ; \rangle$ is a semigroup, $\langle S, + \rangle$ is a join-semilattice and the following equations hold for each $a, b.
 c \in S$:
@@
   \begin{enumerate}
-    \item $a \cdot (b \vee c) = (a \cdot b) \vee (a \cdot c)$
-    \item $(a \vee b) \cdot c = (a \cdot c) \vee (b \cdot c)$
+    \item $a ; (b + c) = (a ; b) + (a ; c)$
+    \item $(a + b) ; c = (a ; c) + (b ; c)$
   \end{enumerate}
 \end{defin}
-This class is clearly a variety since ${\bf OS^{\vee}}$ has the equational definition so far as $\vee$ is defined as an associative, idempotent, and commutative operation.
+This class is clearly a variety since ${\bf OS^{+}}$ has the equational definition so far as $+$ is defined as an associative, idempotent, and commutative operation.
 
@@
 A representable join semilattice-ordered semigroup is an algebra isomorphic to some join semilattice-ordered semigroup having the form $\mathcal{A} = \langle A, |, \cup \rangle$ such that $;$ is a relation composition as above and $\cup$ is the set-theoretic union.
-If $\mathcal{A}$ is representable, then $R(;, \vee)$ Let us recall some of underlying facts about representable join semilattice-ordered semigroups \cite{andreka1991representations}:
+If $\mathcal{A}$ is representable, then $\mathcal{A} \in {\bf I}({\bf R}(\cup, |))$.
+Let us recall some of underlying facts about representable join semilattice-ordered semigroups \cite{andreka1991representations}:
 
@@
   \begin{enumerate}
-    \item Let $\mathcal{A} = \langle A, \vee, \cdot \rangle$ be a join semilattice-ordered semigroup such that, for all $a, b \in A$:
+    \item Let $\mathcal{A} = \langle A, +, ; \rangle$ be a join semilattice-ordered semigroup such that, for all $a, b \in A$:
     \begin{enumerate}
       \item If $a \nleq b$, then there exists an atom $c \leq a$ and $c \nleq b$.
-      \item If $c \leq a \cdot b$ and $c$ is an atom, then there exists an atom $a' \leq a$ such that $c \leq a' \cdot b$.
+      \item If $c \leq a ; b$ and $c$ is an atom, then there exists an atom $a' \leq a$ such that $c \leq a' \cdot b$.
     \end{enumerate}
     then $\mathcal{A}$ is representable.
-    \item Let $\mathcal{A} = \langle A, \cdot \rangle$ be a posemigroup, then $\mathcal{A}$ is representable and such a representation preserve any existing finite suprema and infima, if
+    \item Let $\mathcal{A} = \langle A, ; \rangle$ be a posemigroup, then $\mathcal{A}$ is representable and such a representation preserve any existing finite suprema and infima, if
     \begin{enumerate}
-      \item The set of atoms is closed under $\cdot$.
-      \item $\mathcal{A}$ has enough atoms, that is, if $x \in \operatorname{At}(A)$ and $z, w \in A$, then $x \leq z \cdot w$ implies there exist atoms $z_1 \leq z$ and $w_1 \leq w$ such that $x \leq z_1 \cdot w_1$.
+      \item The set of atoms is closed under $;$.
+      \item $\mathcal{A}$ has enough atoms, that is, if $x \in \operatorname{At}(A)$ and $z, w \in A$, then $x \leq z ; w$ implies there exist atoms $z_1 \leq z$ and $w_1 \leq w$ such that $x \leq z_1 ; w_1$.
 If $z \nleq w$, then there exists an atom $x$ such that $x \leq z$ and $x \nleq w$.
@@
 \begin{theorem}
-  The equational and quasiequational theories of $R(;,\vee)$ is not finitely based.
+  ${\bf R}(\cup, |)$ is not finitely axiomatisable.
 \end{theorem}
 
-Let us provide a proof of this fact using the rainbow technique \cite{hirsch2002relation} to show that the complement of ${\bf ROS^{\vee}}$ is not closed ultraproducts.
+Let us provide a proof of this fact using the rainbow technique \cite{hirsch2002relation} to show that the complement of ${\bf ROS^{+}}$ is not closed ultraproducts.
 This construction sometimes exploits the similar construction used by Andr{\'e}ka \cite{andreka1991representations} and by Maddux \cite{maddux_1989}.
@@
 Moreover, representability is undecidable for lattice-ordered semigroups and ordered complemented semigroups \cite{neuzerling2016undecidability}.
-We use(more or less) a standard way of showing that the class of certain reducts of representable relation algebras has no finite axiomatisation, see \cite{hodkinson2000axiomatizability} \cite{hirsch1997step}.
+We use (more or less) a standard way of showing that the class of certain reducts of representable relation algebras has no finite axiomatisation, see \cite{hodkinson2000axiomatizability} \cite{hirsch1997step}.
 
 First of all, we recall several definitions such as colourings.
-We define a sequence of relation algebras $\{ \mathfrak{A}_n \}_{n < \omega}$ each of which belongs to ${\bf RA}$.
-We need these algebras to show that their $\{;, \vee\}$-reducts are not representable.
-That is, we are seeking to show that
+We provide a sequence of symmetric, integral, finite relation algebras $\{ \mathfrak{A}_n \}_{n < \omega}$ such that $\mathfrak{A}_n \notin {\bf RRA}$.
+The statement has been proved by Andreka \cite{andreka1991representations} and reproduced here \cite{andreka2011axiomatizability}.
 
@@
 \begin{itemize}
-\item identity: ${\bf 1}$, an atom with no colour
-\item greens: ${\bf g}_i$ for $0 \leq i \leq 2^n$
-\item yellow: ${\bf y}$
-\item black: ${\bf b}$
-\item whites: ${\bf w}$, ${\bf w}_{ij}$ for $0 \leq i \leq j \leq 2^n$
-\item reds: ${\bf r}_i$ for $0 < i \leq 2^n$
+\item identity: ${\bf 1}$, an atom with no colour,
+\item white: ${\bf w}$,
+\item greens: ${\bf g}_i$ for $1 \leq i \leq n$,
+\item yellows: ${\bf y}_i$ for $1 \leq i \leq n$,
+\item ivory: ${\bf i}$,
+\item reds: ${\bf r}_i$ for $1 \leq i \leq n$,
+\item blacks: ${\bf b}_i$ for $1 \leq i \leq n$.
 \end{itemize}
-We claim that every atom is self-converse ($a^{\smile} = a$). Given $x, y, z \in \mathfrak{A}_n$, a triple $(x,y,z)$ is an  inconsistent triangle if
+
+We have the following steps:
+
+\vspace{\baselineskip}
+
+{\bf Step 1}.
+Let $\mathcal{A}_n$ be the upper semilattice presented with the set $\operatorname{At}(\mathfrak{A}_n) \cup \{ 0 \}$ and the following relations for each $x \in \operatorname{At}(\mathfrak{A}_n)$ and for each $1 \leq i \leq n$.
+\begin{enumerate}
+  \item ${\bf w} \leq {\bf g}_i + {\bf y}_i$,
+  \item ${\bf i} \leq {\bf y}_i + {\bf r}_i$,
+  \item $x + 0 = x$
+\end{enumerate}
+
+{\bf Step 2}. We define $S$, the set of two element subsets of $\mathcal{A}_n$:
 \begin{center}
-$x \land (y ; z) = y \land (z ; x) = z \land (x ; y) = 0$
-\end{center}
-We define the set of inconsistent triangles explicitly as follows.
-\begin{itemize}
-  \item $(x, y, {\bf 1})$ unless $x = y$
-  \item $({\bf g}_i, {\bf g}_k, {\bf g}_l)$ for $i, k, l \leq 2^n$
-  \item $({\bf y}, {\bf y}, {\bf y})$
-  \item $({\bf y}, {\bf y}, {\bf b})$
-  \item $({\bf g}_i, {\bf g}_j, {\bf w}_{kl})$ for $i, j \leq 2^n$ and $k, j \leq 2^n$
-  \item $({\bf r}_i, {\bf r}_j, {\bf r}_k)$ unless $i + k = j$ or $i + k = j$ or $j + k = i$
-  \item $({\bf g}_i, {\bf g}_{i + 1}, {\bf r})$ unless $j = 1$
-  \item $({\bf g}_i, {\bf y}, {\bf w}_{jk})$ unless $i \in \{ j, k \}$
-\end{itemize}
-$({\bf g}_i, {\bf g}_i, {\bf w}_{kj})$ stands for ${\bf g}_i \land ({\bf g}_i ; {\bf w}_{kj}) = {\bf g}_i \land ({\bf w}_{kj} ; {\bf g}_i) = {\bf w}_{kj} \land ({\bf g}_i ; {\bf g}_i) = 0$, and so on.
-
-$C$ is the set of consistent triangles defined as
-
-\begin{center}
-  $C = \{ (a, b, c ) \: | \: a,b,c \in \operatorname{At}(\mathfrak{A}_n), (a,b,c) \text{ is consistent}\}$
+  $S = \{ \{ {\bf w}, {\bf r}_1 \} \} \cup \{ \{ {\bf g}_i, {\bf b}_i \} \: | \: 1 \leq i \leq n \} \cup
+  \{ \{ {\bf y}_i, {\bf r}_i \} \: | \: 1 \leq i < n \} \cup \{ \{ {\bf y}_n, {\bf i}\} \}$.
 \end{center}
 
-Let us ensure briefly that $\mathfrak{A}_n$ is indeed a relation algebra defining the multiplication table for composition as in \cite{lyndon1950representation}.
-For that, it is enough to show associativity for atoms $a, x, y, z$ such that $a \leq (x ; y) ; z$.
-For that, one needs to find an atom $b$ such that $a \leq x ; b$ and $b \leq y ; z$.
-If $x = y$, then we put $b = {\bf 1}$.
-Otherwise, we try to use a white atom or a black atom for $b$ unless $x, a$ are green and $y, z$ are yellow.
-Suppose $x = {\bf g}_i$, $a = {\bf g}_j$, $y = {\bf y}_k$, and $z = {\bf y}_l$, then we put
+{\bf Step 3}. The operations on $\mathfrak{A}_n$:
+\begin{enumerate}
+  \item $1 = \sum \operatorname{At}(\mathcal{A}_n) \cup \{ 0 \}$,
+  \item $0 = \O$,
+  \item $x = x^{\smile}$,
+  \item $0 ; x = 0 ; x = 0$,
+  \item ${\bf 1} ; x = {\bf 1} ; x = x$,
+  \item $x ; y = \begin{cases} {\bf i}, \text{ if $\{x, y\} \in S$} \\ 1, \text{ otherwise } \end{cases}$ \\ unless $x, y \in \{ 0, {\bf 1}\}$.
+\end{enumerate}
+
+{\bf Step 4}. Define the following quasi-idenity:
+\begin{center}
+  $\begin{array}{lll}
+  & q_n = \bigwedge \limits_{1 \leq i \leq n} ( (x \leq x^{'}_i + x^{''}_i) \land (y \leq y^{'}_i + y^{''}_i)) \to & \\
+  & x ; y \leq x ; y^{'}_1 + \sum \limits_{1 \leq i < n} (x^{'}_i ; y^{''}_i + x^{''}_i ; y^{'}_{i + 1}) + x^{'}_n ; y^{''}_n + x^{''}_n ; y&
+  \end{array}$
+\end{center}
 
 \begin{lemma}
-  For each $n < \omega$, $\mathfrak{A}_n$ does not belong ${\bf RRA}$.
-The $(\vee, ;)$-reduct $\mathfrak{S}_n$ of $\mathfrak{A}_n$ is not representable as well.
-For each $n < \omega$, there is an equation valid in set algebras failing in $\mathfrak{S}_n$.
+
+  \begin{enumerate}
+    \item $q_n$ is valid in ${\bf RRA}$ for each $n < \omega$.
+    \item $q_n$ fails in $\mathfrak{A}_n$.
+  \end{enumerate}
 \end{lemma}
@@
 \begin{proof}
-  See \cite{hodkinson2000axiomatizability} to have a proof that $\mathfrak{A}_n \notin {\bf RRA}$. However, we leave a proof sketch.
-
-  We prove that $\mathfrak{S}_n$ is not representable by contradiction.
-  Suppose $h$ is an isomorphism of $\mathfrak{S}_n$ to a set relation of relations having similarity type $\{ ;, \vee \}$. Let $0$ be a zero element of $\mathfrak{A}_n$.
+  The valuation $\vartheta$ defined as:
+  \begin{enumerate}
+    \item $\vartheta(x) = {\bf w}$
+    \item $\vartheta(x_i^{'}) = {\bf g}_i$
+    \item $\vartheta(x_i^{''}) = {\bf y}_i$
+    \item $\vartheta(y) = {\bf i}$
+    \item $\vartheta(y_i^{'}) = {\bf r}_i$
+    \item $\vartheta(y_i^{''}) = {\bf b}_i$
+  \end{enumerate}
+  falsifies $q_n$ in $\mathfrak{A}_n$
 \end{proof}
 
+The second is to show that any non-trivial ultraproduct $\prod \limits_{D} \mathfrak{A}_n \in {\bf RRA}$, where $D$ is an ultrafilter over $\mathcal{P}(\omega)$.
+We show that via the rainbow technique.
 Let us define networks and games according to \cite{hirsch1997step}.
@@
   Any non-trivial ultraproduct of $\{ \mathfrak{A}_n \}_{n < \omega}$ is representable, that is, belongs to ${\bf RRA}$.
-The same statement for non-trivial ultraproduct of reducts $\{ \mathfrak{S}_n \}_{n < \omega}$ that belongs to $R(;, \vee)$.
+The same statement for non-trivial ultraproduct of reducts $\{ \mathfrak{S}_n \}_{n < \omega}$ that belongs to ${\bf R}(\cup, |)$.
 \end{lemma}
@@
 
-\subsection{The finite algebra on finite base for $R(;, \vee)$ (or its failure)}
+\subsection{The finite algebra on finite base for ${\bf R}(\cup, |)$ (or its failure)}
 

### 2020-07-12 problem-19.17/Text.tex
formulate the claim for representability of the ultraproduct

@@
 \newtheorem{col}{Corollary}
+
+\theoremstyle{claim}
+\newtheorem{claim}{Claim}
 
@@
 
+\subsection{The rainbow construction}
+
 Let us provide a proof of this fact using the rainbow technique \cite{hirsch2002relation} to show that the complement of ${\bf ROS^{+}}$ is not closed ultraproducts.
@@
   \end{enumerate}
-  falsifies $q_n$ in $\mathfrak{A}_n$
+  falsifies $q_n$ in $\mathfrak{A}_n$.
+
+  TODO: visualise the reason for non-representability.
 \end{proof}
 
-The second is to show that any non-trivial ultraproduct $\prod \limits_{D} \mathfrak{A}_n \in {\bf RRA}$, where $D$ is an ultrafilter over $\mathcal{P}(\omega)$.
-We show that via the rainbow technique.
-Let us define networks and games according to \cite{hirsch1997step}.
+\subsection{Networks and games}
 
@@
 
-\begin{lemma}
+\begin{lemma} \label{win}
   Let $\mathcal{A}$ be an atomic relation algebra.
@@
 \end{lemma}
+
+\subsection{The ultraproduct}
+
+The second is to show that any non-trivial ultraproduct $\prod \limits_{D} \mathfrak{A}_n \in {\bf RRA}$, where $D$ is an ultrafilter over $\mathcal{P}(\omega)$.
+We show that via the rainbow technique.
+Let us define networks and games according to \cite{hirsch1997step}.
 
@@
 
-\begin{lemma}
-  TODO: one needs to realise when $\exists$ has a winning strategy
-\end{lemma}
+According to the following claim, $\exists$ has a winning strategy on cofinitely many algebras that allows her to win a game on the ultraproduct.
+Thus, according to Lemma~\ref{win}, the ultraproduct belongs to ${\bf RRA}$.
+
+\begin{claim}
+  Let $l < \omega$. $\exists$ has a winning strategy for $G_l(\mathfrak{A}_n)$ for cofinitely many algebras beloging to the sequence $\{ \mathfrak{A}_n \}_{n < \omega}$.
+\end{claim}
+
+\begin{proof}
+  Let $n \geq l$. We show that $\exists$ has a winning strategy for $G_l(\mathfrak{A}_n)$.
+\end{proof}
 

### 2020-07-13 filtration-stuff/Text.tex
the proper proof that K5 admits strict filtration

@@
     \item For each $x, y, z \in W$, $x R y$ and $x R z$ implies either $y R z$ or $z R y$.
-    \item $R \subseteq R ; R$, that is, $R$ is dense.
     \item For each $x \in W$, $R^{*}(x) = \{ x\} \cup R(R(x))$.
-    \item $R^{-1} ; R \subseteq R$.
+    \item $R^{-1} \circ R \subseteq R$.
   \end{enumerate}
@@
 
-\begin{prop}
-  Let $\mathcal{M}$ be a ${\bf K}5$ model, $x R y$ for some $x, y \in W$ then one has
+\begin{prop} \label{k5useful}
+  Let $\mathcal{M}$ be a ${\bf K}5$ model, $x R y$ for $x, y \in W$ then one has
   \begin{center}
@@
 \end{prop}
+
+\begin{proof}
+$ $
+
+  \begin{enumerate}
+    \item Suppose $\mathcal{M}, x \models \Diamond \Box \varphi$.
+One also has $\mathcal{M}, x \models \Diamond \Box \varphi \to \Box \Diamond \Box \varphi$, so $\mathcal{M}, x \models \Box \Diamond \Box \varphi$.
+Thus, $\mathcal{M}, y \models \Diamond \Box \varphi$ since $y \in R(x)$.
+    \item Suppose $\mathcal{M}, y \models \Diamond \Box \varphi$, then $\mathcal{M}, y \models \Box \varphi$, so $\mathcal{M}, x \models \Diamond \Box \varphi$.
+  \end{enumerate}
+\end{proof}
 
@@
 
-\begin{defin} A relation $R$ is called Horn, if it is defined with some first-order formula having the form
+\begin{defin} A first-order formula is called Horn if it has the following form:
 
@@
 $E$-closure (an Euclidean Horn closure of a binary relation) has the following equivalent definitions:
-\begin{lemma}
+\begin{lemma} \label{equivHorn}
   Let $\mathcal{F} = \langle W, R \rangle$ be a Kripke frame.
@@
       \item $R_0 = R$
-      \item $R_{n + 1} = R_n \cup (R^{-1}_n ; R_n)$
+      \item $R_{n + 1} = R_n \cup (R^{-1}_n \circ R_n)$
     \end{itemize}
     \item $x R^E y$ iff there exists $n < \omega$ such that
-    $\exists x_1, \dots, x_n$ with $z_1 R x$ and $z_{n - 1} R y$ and for each $1 < i \leq n$ one has
+    either $x R y$ or $\exists z_1, \dots, z_n$ with $z_1 R x$ and $z_{n - 1} R y$ and for each $1 < i \leq n$ one has
     either $z_{i - 1} R z_i$ or $z_i R z_{i - 1}$.
-    \item $R^{E} = R \cup \bigcup \limits_{i < \omega} (R^{-1} ; (R ; R^{-1})^n ; R)$.
+    \item $R^{E} = R \cup \bigcup \limits_{i < \omega} (R^{-1} \circ (R \circ R^{-1})^n \circ R)$.
   \end{enumerate}
@@
   \begin{itemize}
-    \item $R^{E} \subseteq \bigcup \limits_{i < \omega} R_i$. It is enough to show that
-    $x R^{E} y$ implies $x R^n y$ for some $n < \omega$.
+    \item $R^{E} \subseteq \bigcup \limits_{i < \omega} R_i$. Recall that $R^{E}$ has the form (?):
+    \begin{center}
+      $R^{E} = \bigcap \{ R' \: | \: R \subseteq R', \forall a, b \in W \: R'(a,b) \Rightarrow \exists x \in W \: R'(x, a) \: \& \: R'(x, b)\}$
+    \end{center}
+
     \item $\bigcup \limits_{i < \omega} R_i \subseteq R^{E}$.
@@
     Suppose $n = m + 1$ and $x R_{m + 1} y$. Let us show that $x R^{E} y$.
-    From $x R_{m + 1} y$, one has $(x, y) \in R^n \cup (R^{-1}_n ; R_n)$. There are two cases:
+    From $x R_{m + 1} y$, one has $(x, y) \in R^n \cup (R^{-1}_n \circ R_n)$. There are two cases:
     \begin{itemize}
       \item $x R^n y$, one needs to merely apply the IH.
-      \item $x R^{-1}_n ; R_n y$. Then $\exists z \in W \: x R^{-1}_n z \: \& \: z R_n$. That is,
+      \item $x R^{-1}_n \circ R_n y$. Then $\exists z \in W \: x R^{-1}_n z \: \& \: z R_n$. That is,
       $z R_n x$ and $z R_n y$ for some $z$. $R_n$ is already a subrelation of $R^E$. Thus,
@@
   \item $(2) \Rightarrow (3)$
-  Let $x R^E y$, then $(x, y) \in \bigcup \limits_{i < n} R_n$. That is, $(x, y) \in R_n \cup (R_n^{-1} ; R_n)$ for some $n < \omega$. There are two cases.
-  \begin{itemize}
-    \item $(x, y) \in R_n$ and that's it.
-    \item $(x, y) \in R_n^{-1} ; R_n$, then $(z, x), (z, y) \in R_n$.
-  \end{itemize}
+  Let $(x, y) \in R_m$, let us the statement by induction on $m$.
+  \begin{enumerate}
+    \item Suppose $m = 0$, then $x R y$, and the statement is shown putting $n = 0$.
+    \item Suppose $m = p + 1$ and $x R_{p + 1} y$.
+    Assume that either $x R y$ or $\exists z_1, \dots, z_p$ with $z_1 R x$ and $z_{p - 1} R y$ and for each $1 < i \leq p$ one has
+    either $z_{i - 1} R z_i$ or $z_i R z_{i - 1}$.
+
+    $x R_{p + 1} y$ implies $(x, y) \in R_{p} \cup (R^{-1}_p \circ R_p)$. If $(x, y) \in R_{p}$, then we merely apply the IH.
+    Suppose $(x, y) \in R^{-1}_p \circ R_p$, then $(z, x) \in R_p$ and $(z, y) \in R_p$
+  \end{enumerate}
   \item $(3) \Rightarrow (4)$
+  Suppose either $x R y$ or there exist $n \geq 1$ and $z_1, \dots, z_n$ with $z_1 R x$ and $z_{n - 1} R y$ and for each $1 < i \leq n$ one has either $z_{i - 1} R z_i$ or $z_i R z_{i - 1}$.
+  If $x R y$, then we are done. Otherwise there exists $n \geq 1$ with the condition above. Then $(x, y) \in R_{n + 1}$ that follows from the condition.
   \item $(4) \Rightarrow (1)$
@@
 \begin{lemma}
-  Let $\mathcal{M} = \langle W, R, \vartheta \rangle$ be an Euclidean model, $\Gamma$ a set of $\operatorname{Sub}$-closed formulas, and $\sim$ an equivalence relation having a finite index that respects $\Gamma$.
-Then its $E$-closure of the minimal filtration of $R$ is a filtration itself.
+  Let $\mathcal{M} = \langle W, R, \vartheta \rangle$ be an Euclidean model, $\Gamma$ a set of $\operatorname{Sub}$-closed formulas, and $\sim$ an equivalence relation having a finite index that respects $\Gamma$, then $\widehat{R} = (R^{min}_{\Phi})^{E} \subseteq R^{max}_{\Gamma}$, where $\Phi = \Gamma \cup \{ \Diamond \Box \varphi \: | \: \Box \varphi \in \Gamma \}$.
+
+  Thus, ${\bf K}5$ admits strict filtrations.
 \end{lemma}
@@
 \begin{proof}
-  Let us show that $\widehat{R^{E}} = \bigcup \limits_{n < \omega} (R_n)^{\text{min}}_{\sim}$ is a filtration and $\langle W / \sim, \widehat{R}^{E}, \widehat{\vartheta} \rangle$ is an Euclidean model.
+  Recall that $(R^{min}_{\Phi})^{E}$ has the form $(R^{min}_{\Phi})^{E} = \bigcup \limits_{n < \omega} (R^{min}_{\Phi})_n$, where
+  \begin{enumerate}
+    \item $(R^{min}_{\Phi})_0 = R^{min}_{\Phi}$
+    \item $(R^{min}_{\Phi})_{m + 1} = (R^{min}_{\Phi})_n \cup (((R^{min}_{\Phi})_n)^{-1} \circ (R^{min}_{\Phi})_n)$
+  \end{enumerate}
+  One needs to show that for each $n < \omega$ $(R^{min}_{\Phi})_n \subseteq R^{max}_{\Gamma}$. We prove this by induction.
+  Suppose $\mathcal{M}, x \models \Box \varphi$ for $\Box \varphi \in \Phi$ and $[x] (R^{min}_{\Phi})^{E} [y]$. We need $\mathcal{M}, y \models \varphi$.
+  \begin{enumerate}
+    \item $([x], [y]) \in (R^{min}_{\Phi})_0$, then $([x], [y]) \in R^{min}_{\Phi}$.
+Then there exist $x' \in [x]$ and $y' \in [y]$ such that $x' R y'$.
+So $\mathcal{M}, x' \models \Box \varphi$ and, thus,
+    $\mathcal{M}, y' \models \varphi$. Then $\mathcal{M}, y' \models \varphi$ as well since $y' \in [y]$.
+    \item $([x], [y]) \in (R^{min}_{\Phi})_{m + 1}$, then
+    $([x], [y]) \in (R^{min}_{\Phi})_{m} \cup (((R^{min}_{\Phi})_{m})^{-1} \circ R^{min}_{\Phi})_{m})$.
 
-  Let $\widehat{\mathcal{M}} = \langle \widehat{W}, \widehat{R}, \widehat{\vartheta} \rangle$ be a minimal filtration of an Euclidean model through $\sim$.
+    If $([x], [y]) \in (R^{min}_{\Phi})_{m}$, then we apply the IH.
 
-  \begin{enumerate}
-    \item Suppose $x R y$, then $[x] \widehat{R} [y]$ by the definition of a minimal filtration. $\widehat{R}$ is clearly a subrelation of $\widehat{R}^{E}$, thus $[x] \widehat{R}^{E} [y]$.
-    \item Suppose $[x] \widehat{R^{E}} [y]$ and $\mathcal{M}, x \models \Box \varphi$ for $\Box \varphi \in \Gamma$.
-then $([x], [y]) \in \bigcup \limits_{n < \omega} R_n$.
-Then $([x], [y]) \in R_n$ for some $n < \omega$.
-    There are two cases:
-    \begin{enumerate}
-      \item $n = 0$, then $[x] \widehat{R} [y]$, so, obviously, one has $\mathcal{M}, y \models \varphi$
-      \item $n = m + 1$. Suppose $[x] {\widehat{R}}_{m + 1} [y]$. Thus, $[x] \widehat{R}_n \cup (\widehat{R}^{-1} ; \widehat{R}_n) [y]$.
-      There are the following two cases:
-      \begin{enumerate}
-        \item $([x], [y]) \in \widehat{R}_n$, then the statement holds by IH.
-        \item $([x], [y]) \in \widehat{R}^{-1} ; \widehat{R}_n$.
-        Then there exists $[z]$ such that $[x] \widehat{R}^{-1} [z]$ and $[z] \widehat{R}_m [y]$, that is,
-        $[z] \widehat{R} [x]$ and $[z] \widehat{R}_m [y]$. One has $\mathcal{M}, x \models \Box \varphi$, then
-        $\widehat{\mathcal{M}}, [x] \models \Box \varphi$.
-So $\widehat{\mathcal{M}}, [z] \models \Diamond \Box \varphi$.
-On the other hand, $\mathcal{M}, z \models \Diamond \Box \varphi \to \Box \varphi$, so $\mathcal{M}, z \models \Box \varphi$ and,
-        from that, $\widehat{\mathcal{M}}, [z] \models \Box \varphi$, and, thus, $\widehat{\mathcal{M}}, [y] \models \varphi$ and $\mathcal{M}, y \models \varphi$ by IH and the definition of a minimal filtration.
-      \end{enumerate}
-    \end{enumerate}
+    Suppose $([x], [y]) \in (R^{min}_{\Phi})_{m})^{-1} \circ (R^{min}_{\Phi})_{m}$,
+    then there exists $[z] \in W / \sim_{\Phi}$ such that $([z], [x]) \in (R^{min}_{\Phi})_{m})$ and $([z], [y]) \in (R^{min}_{\Phi})_{m}$.
+
+    Then one has the following picture (using Lemma~\ref{equivHorn}):
+
+    \xymatrix{
+&&    [z] & \ar[l]_{R^{min}_{\Phi}} [z_1] \ar@{-}[r]^{R'} & [z_2] \ar@{-}[r]^{R'} & \ar@{-}[r]^{R'} \dots & [z_{m - 1}] \ar@{-}[r]^{R'} & [z_m] \ar[r]^{R^{min}_{\Phi}} & [x] \\
+&&    [z] & \ar[l]^{R^{min}_{\Phi}} [z_1^{'}] \ar@{-}[r]_{R'} & [z_2^{'}] \ar@{-}[r]_{R'} & \ar@{-}[r]_{R'} \dots & [z_{m - 1}] \ar@{-}[r]_{R'} & [z_m^{'}] \ar[r]_{R^{min}_{\Phi}} & [y]
+    }
+  Where $R'$ is either $R^{min}_{\Phi}$ or its converse.
+  One has $\mathcal{M}, x \models \Box \varphi$ for $\Box \varphi \in \Phi$, where $\widehat{M}$ is the minimal filtration of $\mathcal{M}$ through $\Phi$.
+One has $[z_m] R^{min}_{\Phi} [x]$, then $a_m R a$ for some $a_m \in [z_n]$ and $a \in [x]$.
+Thus, $\mathcal{M}, a_m \models \Diamond \Box \varphi$ and, thus, $\widehat{\mathcal{M}}, [z_m] \models \Diamond \Box \varphi$.
+
+  Applying Proposition~\ref{k5useful} several times, one may show that $\widehat{\mathcal{M}}, [z_1] \models \Diamond \Box \varphi$.
+  One has $[z_1] R^{min}_{\Phi} [z]$, then for some $a \in [z]$ and $a_1 \in [z_1]$ we have $a_1 R a$.
+
+  Then $\mathcal{M}, a \models \Box \varphi$ and $\widehat{M}, [z] \models \Box \varphi$.
+
+  We have $[z_1^{'}] R^{min}_{\Phi} [z]$, thus, $a_1^{'} R a^{'}$ for some $a_1^{'} \in [z_1^{'}]$ and $a^{'} \in [z]$. Then
+  $\mathcal{M}, a_1^{'} \models \Diamond \Box \varphi$, and, thus, $\widehat{M}, [z_1^{'}] \models \Diamond \Box \varphi$.
+
+  One may show that $\widehat{M}, [z_{m}^{'}] \models \Diamond \Box \varphi$ in the same way via Lemma~\ref{k5useful}.
+Thus, $\mathcal{M}, z_{m}^{'} \models \Diamond \Box \varphi$.
+We also have $\mathcal{M}, z_{m}^{'} \models \Diamond \Box \varphi \to \Box \varphi$, and, thus,
+  $\mathcal{M}, z_{m}^{'} \models \Box \varphi$. Then $\widehat{\mathcal{M}}, [z_m^{'}] \models \Box \varphi$.
+
+  One has $[z_m^{'}] R^{min}_{\Phi} [y]$, then $a_m^{'} R y'$ for some $a_m^{'} \in [z_m^{'}]$ and $y' \in [y]$.
+Then $\mathcal{M}, y' \models \varphi$.
+But $y' \sim_{\Phi} y$, so $\mathcal{M}, y \models \varphi$.
   \end{enumerate}
 \end{proof}
-
-\begin{col}
-  ${\bf K}5$ admit strict filtrations.
-\end{col}
 

### 2020-07-16 problem-19.17/FRP.tex
solution for the problem 19.17 of Hirsch-Hodkinson-2002
started the section with the signatures containing Boolean meet

the FRP: the proof sketch for the fact that R(;, ., ^{\smile}) and R(;, ., 1') have no the frp

update and definitions

\documentclass[a4paper]{article}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{bussproofs}
\usepackage{mathtools}
\usepackage{verbatim}
\usepackage{dsfont}
\usepackage{mathabx}
\usepackage[all, 2cell]{xy}
\usepackage[all]{xy}
\usepackage{wasysym}
\usepackage{rotating}
\usepackage{geometry}
\usepackage{trfsigns}
\usepackage{cmll}
\usepackage{authblk}
\usepackage{hyperref}
\usepackage{cleveref}
\usepackage{lipsum}
\usepackage{extpfeil}
\usepackage{soul}
\usepackage{graphicx}

\newcommand\mapsfrom{\mathrel{\reflectbox{\ensuremath{\mapsto}}}}

\theoremstyle{defin}
\newtheorem{defin}{Definition}

\theoremstyle{theorem}
\newtheorem{theorem}{Theorem}

\theoremstyle{prop}
\newtheorem{prop}{Proposition}

\theoremstyle{lemma}
\newtheorem{lemma}{Lemma}

\theoremstyle{ex}
\newtheorem{ex}{Example}

\theoremstyle{col}
\newtheorem{col}{Corollary}

\theoremstyle{claim}
\newtheorem{claim}{Claim}

\author{Daniel Rogozin}
\date{}
\title{The finite base property for subreducts of representable relation algebras}

\begin{document}

\maketitle

\section{Relation algebras and their reducts}

We recall the basic definitions and results about relation algebras \cite{hirsch2002relation} \cite{maddux2006relation}.

\begin{defin}
  $ $

    A relation algebra is an algebra $\mathcal{R} = \langle R, 0, 1, +, -, ;, {}^{\smile}, {\bf 1 }\rangle$ such that $\langle R, 0, 1, +, - \rangle$ is a Boolean algebra and the following
    equations hold, for each $a, b, c \in R$:
    \begin{enumerate}
      \item $a ; (b ; c) = (a ; b) ; c$
      \item $(a + b) ; c = (a ; c) + (b ; c)$
      \item $a ; {\bf 1} = a$
      \item $a^{\smile \smile} = a$
      \item $(a + b)^{\smile} = a^{\smile} + b^{\smile}$
      \item $(a ; b)^{\smile} = b^{\smile} ; a^{\smile}$
      \item $a^{\smile} ; (- (a ; b)) \leq - b$
    \end{enumerate}
    where $a \leq b$ iff $a + b = b$. ${\bf RA}$ denotes the class of all relation
    algebras.
\end{defin}
A relation algebra is called symmetric, if every element is self-converse. A relation algebra is called integral, if
\begin{center}
  $a ; b = 0 \Rightarrow a = 0 \text{ or } b = 0$.
\end{center}

\begin{defin}
    A proper relation algebra is an algebra $\mathcal{R} = \langle R, 0, 1, \cup, -, |, {}^{\smile}, {\bf 1 }\rangle$ such that $R \subseteq \mathcal{P}(W)$, where $W$ is an equivalence relation; $0 = \emptyset$; $1 = W$; $\cap$, $\cup$, $-$ are set-theoretic intersection, union, and
    complement respectively; $|$ is relation composition, ${}^{\smile}$ is relation converse, ${\bf 1}$ is a
    diagonal relation restricted to $W$, that is:
    \begin{enumerate}
    \item $a | b = \{ \langle x, z \rangle \: | \: \exists y \: \langle x, y \rangle \in a \: \& \: \langle y, z \rangle \in b \}$
    \item $a^{\smile} = \{ \langle x, y \rangle \: | \: \langle y, x \rangle \in a \}$
    \item ${\bf 1} = \{ \langle x, y \rangle \: | \: x = y \}$
    \end{enumerate}
      The class of all proper relation algebras is denoted as ${\bf PRA}$. ${\bf Rs}$ is the class of all relation set algebras, proper relation algebra with a diagonal subrelation as an identity. ${\bf RRA}$ is the class of all representable relation algebras, that is, the closure of ${\bf PRA}$ under isomorphic copies. That is,
      ${\bf RRA} = {\bf IPRA}$.
\end{defin}
Note that the (quasi)equational theories of these classes coincide, that is
\begin{center}
  ${\bf IPRA} = {\bf RRA} = {\bf SP}{\bf Rs}$
\end{center}
It is also known that ${\bf RA}$ and ${\bf RRA}$ are varieties, but ${\bf RRA}$ is not finitely axiomatisable.

We consider signatures include the equality predicate and are subsets of
$\Omega = \{ =, \leq, 0, 1, {\bf 1}, +, \cdot, -, \smile, ;, \setminus, / \}$. A subset $\tau \subseteq \Omega$ is called a $\tau$-signature.

\begin{defin}
  An interpretation $R$ of $\tau$-algebra $\mathcal{A}$ over a base $X$ is a map that assigns a binary relation $a^R \subseteq X \times X$
  for each $a \in X$. An interpretation $R$ is a $\tau$-interpretation if the following hold
  (for the relevant predicate, function or constant belonging to $\tau$):
  \begin{enumerate}
    \item $a^R = b^R$ iff $a = b$,
    \item $a^R \subseteq b^R$ iff $a \leq b$
    \item $0^{R} = \emptyset$,
    \item $(a + b)^{R} = a^R \cup b^R$,
    \item $(a \cdot b)^{R} = a^R \cap b^R$,
    \item $(- a)^R = \{ (x, y) \: | \: \exists b \in \mathcal{A} \: (x, y) \in b^R \setminus a^R \}$,
    \item $({\bf 1}) = \{ (a, a) \: | \: a \in \mathcal{A} \}$,
    \item $(a;b)^R = \{ (x, y) \: | \: \exists z \in X \: (x, z) \in a^R \: \& \: (z, x) \in b^R \}$,
    \item $(a \setminus b)^{R} = \{ (x, y) \: | \: \forall z \in X \: ((z, x) \in a^R \Rightarrow (z, y) \in b^R)\}$,
    \item $(a / b)^{R} = \{ (x, y) \: | \: \forall z \in X \: ((y, z) \in a^R \Rightarrow (x, z) \in b^R)\}$.
  \end{enumerate}
\end{defin}

We will the following notations according to, for example, \cite{hirsch2011positive}. Let $\tau$ be a subsets of operations definable in ${\bf RRA}$. ${\bf R}(\tau)$ is the class of subalgebras of $\tau$-reducts of algebras belonging to ${\bf Rs}$.

A $\tau$-structure is \emph{representable} if its isomorphic to some algebra of relations of $\tau$-signature. A representable
finite $\tau$-structure has a \emph{finite representation over a finite base} if it is isomoprhic to some finite representable over a
finite base.
${\bf R}(\tau)$ has the finite representation property if every $\mathcal{A} \in {\bf R}(\tau)$ has a finite representation over a
finite base.

\subsection{Residuals}

One may express residuals in every $\mathcal{R} \in {\bf RA}$ as follows using Boolean negation, inversion, and composition as follows:

\begin{enumerate}
  \item $a \setminus b = -(a^{\smile} ; -b)$
  \item $a / b = - (- a ; b^{\smile})$
\end{enumerate}

Those residuals have the following interpretation in $\mathcal{R} \in {\bf PRA}$ (as well as in ${\bf RRA}$), for every $a, b \in \mathcal{R}$:
\begin{enumerate}
  \item $a \setminus b = \{ \langle x, y \rangle \: | \: \forall z \: (z, x) \in a \Rightarrow (z, y) \in b \}$
  \item $a / b = \{ \langle x, y \rangle \: | \: \forall z \: (y, z) \in b \Rightarrow (x, z) \in a \}$
\end{enumerate}
One may illustrate composition and residuals in ${\bf PRA}$ and ${\bf RRA}$ via the following triangles:

\xymatrix{
&&& \exists y \ar@{-->}[ddr]^{b} &&& \forall z \ar@{-->}[ddl]_{a} \ar@{-->}[ddr]^{b} &&& \forall z \\
&&&&&& \Rightarrow &&& \Leftarrow \\
&& x \ar@{-->}[uur]^{a} \ar[rr]_{a;c} && z & x \ar[rr]_{a \setminus b} && y & x \ar@{-->}[uur]^{a} \ar[rr]_{a / b} && y \ar@{-->}[uul]_{b}
}

\section{The ${\bf R}(;, \setminus, /, \leq)$ case}

 A \emph{relation semigroup} (${\bf RS}$) is a algebra of the signature $\tau = \langle ;, \setminus, /, \leq \rangle$, where $;, \setminus, /$ are binary function symbols and $\leq$ is a binary relation symbol.

\begin{defin}
  A residuated semigroup is an algebra $\mathcal{A} = \langle A, ;, \leq, \setminus, / \rangle$ such that $\langle A, ;, \leq, \rangle$ is
  an ordered residuated semigroup and the following equivalences hold for each $a, b, c \in A$:

  \begin{center}
    $b \leq a \setminus c \Leftrightarrow a ; b \leq c \Leftrightarrow a \leq c / b$
  \end{center}
  ${\bf ORS}$ is the class of all residuated semigroups.
\end{defin}

Historically, residuated structures were introduced by Krull to study ideals of rings \cite{krull1968idealtheorie}. Such algebras further has been considered within problems of semantics of substructural logics, see \cite{jipsen2002survey}.

Let us define the class ${\bf R}(;, \setminus, /, \leq)$ explicitly:

\begin{defin} \label{rrs}
  Let $A$ be a set of binary relations on some base set $W$ such that $R = \cup A$ is transitive and $\{ x, y \: | \: x R y \} = W$. A relation residuated semigroup is an algebra $\mathcal{A} = \langle A, ;, \setminus, /, \subseteq \rangle$ where for each $r, s \in A$
  \begin{enumerate}
    \item $r ; s = \{ \langle a, c \rangle \: | \: \exists b \in W \: (\langle a, b \rangle \in r \: \& \: \langle b, c \rangle \in s) \}$
    \item $r \setminus s = \{ \langle a, c \rangle \: | \: \forall b \in W \: (\langle b, a \rangle \in r \Rightarrow \langle b, c \rangle \in s)\}$
    \item $r / s = \{ \langle a, c \rangle \: | \: \forall b \in W \: (\langle c, b \rangle \in s \Rightarrow \langle a, b \rangle \in r)\}$
    \item $r \leq s$ iff $r \subseteq s$.
  \end{enumerate}
\end{defin}
Relation residuated semigroup are also called representable relativised relational structure (${\bf RRS}$).
Andr\'{e}ka and Mikul\'{a}s proved the following representation theorem for ${\bf ORS}$ in \cite{andreka1994lambek} that implies relational completeness of the Lambek calculus, the logic of ${\bf ORS}$:

\begin{theorem}\label{ors=irrs}
  ${\bf ORS} = {\bf IRRS}$, where ${\bf IRRS}$ is a closure of ${\bf RRS}$ under isomorphic copies.
\end{theorem}
One may reformulate this result as $\mathcal{A} \in {\bf R}(;, \setminus, /, \leq)$ iff $\mathcal{A}$ is a residuated semigroup.

\subsection{Quantales and quantic nuclei}

A quantic nucleus is a closure operator on an ordered semigroup allowing one to define subalgebras. Such an operator is a generalisation of a well-known nucleus on a Heyting algebra, see \cite{bezhanishvili2016locales}. The following definition and the proposition below are given according to \cite{galatos2007residuated} and \cite{rosenthal1990quantales}.
\begin{defin}
  A quantic nucleus on a partially ordered semigroup $\langle A, ;, \leq \rangle$ is a map $j : A \to A$ such that $j$ a closure operator such that $j a ; j b \leq j (a ; b)$.
\end{defin}

\begin{prop} \label{subsemi}
  Let $\mathcal{A} = \langle A, ;, \leq \rangle$ be a partially ordered semigroup and $j$ a quantic conucleus, then
  $\mathcal{A}_j = \{ a \in A \: | \: j a \leq a \}$ is a partially ordered subsemigroup, where $a ;_j b = j(a ; b)$.
\end{prop}

\begin{defin}
  A quantale is an algebra $\mathcal{Q} = \langle Q, \cdot, \bigvee \rangle$ such that $\mathcal{Q} = \langle Q, \bigvee \rangle$ is a complete lattice, $\langle Q, \cdot \rangle$ is a semigroup, and the following conditions hold for each $a \in Q$ and $A \subseteq Q$:
  \begin{enumerate}
    \item $a \cdot \bigvee A = \bigvee \{ a \cdot q \: | \: q \in A \}$,
    \item $\bigvee A \cdot a = \bigvee \{ q \cdot a \: | \: q \in A \}$.
  \end{enumerate}
\end{defin}
Note that any quantale is a residuated semigroup as well. One may express residuals uniquely with suprema and products as follows:
\begin{enumerate}
  \item $a \setminus b = \bigvee \{ c \: | \: a \cdot c \leq b \}$,
  \item $a / b = \bigvee \{ c \: | \: b \cdot c \leq a \}$.
\end{enumerate}

One may embed any residuated semigroup into a quantale via the Dedekind-MacNeille completion (see, for example, \cite{theunissen2007macneille}) as follows. According to Goldblatt \cite{goldblatt2006kripke}, residuated semigroups have the following representation that exploits quantic conuclei (a closure operator over ordered semigroup) and Galois connection:
\begin{theorem} \label{orsRep}
  Every residuated semigroup has an isomorpic embedding to the subalgebra of some quantale.
\end{theorem}
\begin{proof}
  We provide a proof sketch.

  Let $\mathcal{A} = \langle A, \leq, \cdot, \setminus, / \rangle$ be a residuated semigroup.

  Let $X \subseteq A$. We put $lX$ and $uX$ as the sets of lower and upper bounds of $X$ in $A$. We also put $m X = lu X$.
  Note that the lower cone of an arbitrary $x$, $\downarrow x = \{ y \: | \: S \ni y \leq x\}$, is $m$-closed, that is,
  $m (\downarrow x) = \downarrow x$. Moreover, $m : \mathcal{P}(A) \to \mathcal{P}(A)$ is a closure operator and
  $\langle (\mathcal{P}(A))_m, \subseteq \rangle$ (where $(\mathcal{P}(A))_m = \{ X \in \mathcal{P}(S) \: | \: m X = X\}$ ) is a complete
  lattice with $\bigvee_{m} \mathcal{X} = m ( \bigcup \mathcal{X})$ and $\bigwedge_{m} = \bigcap \mathcal{X}$ \cite{davey2002introduction}.
  Moreover, according to Proposition~\ref{subsemi}, $\langle (\mathcal{P}(A))_m, \subseteq, ;_m \rangle$ is a subquantale of
  $\langle (\mathcal{P}(S))_m, \subseteq, ; \rangle$, since $m$ is a quantic nucleus. Here $X ; Y = \{ x ; y \: | \: x \in X, y \in Y \}$ and $X ;_m Y = m (X ; Y)$.

  We define a map $f_m : \mathcal{A} \to (\mathcal{P}(A))_m$ such that $f_m : a \mapsto \downarrow a$. Note that $f_m$ preserves residuals.
\end{proof}

Note that this (more or less) construction holds for residuated monoids as well, see \cite{goldblatt2011grishin}.

In their turn, quantales have a relational representation. First of all, let us define a relational quantale.
The notion of relational quantale was introduced by Brown and Gurr to represent quantales as algebras of relations and study relational semantics of the full Lambek calculus, see \cite{brown1993representation} and \cite{brown1995relations}.
\begin{defin}
  Let $A$ be a set. A relational quantale on $A$ is an algebra $\langle R, \subseteq, ; \rangle$, where
  \begin{enumerate}
    \item $R \subseteq \mathcal{P}(A \times A)$,
    \item $\langle R, \subseteq \rangle$ is a complete join-semilattice,
    \item $;$ is a relational composition that respects suprema in each coordinate.
  \end{enumerate}
\end{defin}

The following representation theorem for quantale has been proved by Brown and Gurr \cite{brown1993representation}.
\begin{theorem} \label{quantaleRep}
  Every complete residuated semigroup $\mathcal{A}$ (quantale) is isomorphic to relational quantale on the underlying set of $\mathcal{A}$.
\end{theorem}

\begin{proof}
  Let us describe a proof sketch. Let $\mathcal{Q}$ be a quantale and $\mathcal{G}(\mathcal{Q})$ a set of its generators. We define:

  \begin{center}
    $\hat{a} = \{ \langle g, q \rangle \: | g \in \mathcal{G}(\mathcal{Q}), q \in \mathcal{Q}, g \leq a \cdot q \}$

    $\widehat{\mathcal{Q}} = \{ \hat{a} \: | \: a \in \mathcal{Q} \}$
  \end{center}

  The rest of the sketch consists of the following claims.

  {\bf Claim 1} $a \leq b$ iff $\hat{a} \subseteq \hat{b}$.

  {\bf Claim 2} $\widehat{\bigvee A} = \bigvee \widehat{A}$, $\hat{a} ; \hat{b} = \widehat{a \cdot b}$, and $\langle \widehat{\mathcal{Q}}, \subseteq \rangle$ is a complete semilattice.

  {\bf Claim 3}. $\langle \widehat{\mathcal{Q}}, \subseteq, ; \rangle$ is a relational quantale.

  {\bf Claim 4}. $\mathcal{Q}$ is isomorphic to $\langle \widehat{\mathcal{Q}}, \subseteq, ; \rangle$.
\end{proof}

Theorems \ref{orsRep} and \ref{quantaleRep} imply the following statement.
\begin{col} \label{orsRep2}
  Every residuated semigroup is isomorphic to some subalgebra of some relational quantale.
\end{col}

\subsection{Solution for Problem 19.17 \cite{hirsch2002relation}}

In \cite{hirsch2002relation}, Hirsch and Hoskidson left the problem about having the finite representation property. Given a finite algebra $\mathcal{A} \in {\bf RRS}$, we show that $\mathcal{A}$ is isomorphic to some algebra $\mathcal{B}$ that has the finite base.

First of all, let us show that
\begin{lemma} \label{rrsLemma}
  Let $\mathcal{A} = \langle A, \cdot, \setminus, /, \leq \rangle$ be a residuated semigroup, then its representation with relational quantales belongs to ${\bf RRS}$.
\end{lemma}

\begin{proof}
  By Corollary~\ref{orsRep2}, $\mathcal{A}$ is isomorphic subalgebra of $\widehat{\mathcal{Q}_{\mathcal{A}}}$, where
  $\widehat{\mathcal{Q}_{\mathcal{A}}} = \langle R, ;, \bigvee \rangle$ is a quantale of Theorem~\ref{orsRep}.
  $\mathcal{Q}_{\mathcal{A}}$ is the quantale of Galois stable subsets of $\mathcal{A}$.

  $\widehat{\mathcal{Q}_{\mathcal{A}}}$ is the relational quantale of binary relations on $\mathcal{Q}_{\mathcal{A}}$ as a base set and
  $R = \{ \hat{A} \: | \: A \in \mathcal{Q}_{\mathcal{A}}\}$.

  Let us show that $W = \bigcup R$ is a transitive relation.
  Suppose $\widehat{X}, \widehat{Y}, \widehat{Z} \in R$ for $X, Y, Z \in \mathcal{Q}_{\mathcal{A}}$ with
  $\widehat{X} W \widehat{Y} W \widehat{Z}$.

  $\widehat{X} W \widehat{Y}$ means that $(\widehat{X}, \widehat{Y}) \in W$, i.e., $(\widehat{X}, \widehat{Y}) \in R_1$ for some
  $R_1 \in R$.

  $\widehat{Y} W \widehat{Z}$ means that $(\widehat{Y}, \widehat{Z}) \in W$, i.e., $(\widehat{Y}, \widehat{Z}) \in R_2$ for some
  $R_2 \in R$.

  We need $(\widehat{X}, \widehat{Z}) \in W$, that is, there exists $R_3 \in W$ such that $(\widehat{X}, \widehat{Z}) \in R_3$.
  But $(\widehat{X}, \widehat{Z}) \in R_1 ; R_2$. But $m(R_1 ; R_2) \in R$ as well, thus, $(\widehat{X}, \widehat{Z}) \in W$.
  We put $R_3 = m(R_1 ; R_2)$, where $m$ is the associated quantic nucleus on $\mathcal{P}(A)$.

  Thus, a representation of $\mathcal{A}$ by means of Theorem~\ref{orsRep} is representable in terms of ``usual'' representations.
  Thus, ${\bf RRS}$ is closed under such representations.
\end{proof}

\begin{col}\label{finite}
  Let $\mathcal{A}$ be a finite residuated semigroup, then its representation with the corresponding relational quantale has the finite base.
\end{col}

\begin{proof}
  The base of $\widehat{\mathcal{Q}_{\mathcal{A}}}$ is the set of Galois stable subsets of $\mathcal{A}$, the cardinality of which is clearly finite.
\end{proof}

Theorem \ref{orsRep}, Theorem \ref{quantaleRep}, Lemma \ref{rrsLemma}, and Corollary \ref{finite} provide the solution for Problem 19.17 of \cite{hirsch2002relation}.
\begin{theorem} \label{solution}
  Let $\mathcal{A} \in {\bf RRS}$ and $|\mathcal{A}| < \omega$, then there exists a set $W$ such that $|W| < \omega$, a set $A$ of binary relations on $W$, $R = \cup A$ with $dom(R) = W$ such that
  $\mathcal{A} \cong \langle A, ;, \setminus, /, \subseteq \rangle$.
\end{theorem}
Clearly that a desired $W$ is the domain of $\mathcal{A}$.

The main corollary of Theorem~\ref{solution} is that the Lambek calculus has the finite model property. Thus, we have a semantical proof of decidability of the Lambek calculus. Before that, there were several algebraic proofs that the Lambek caclulus has the fmp \cite{buszkowski2008infinitary}, but the authors considered arbitrary algebras, not representable ones.
Alternatively, one may show that the Lambek calculus is decidable syntactically, that is, via cut elimination and the subformula property \cite{lambek1958mathematics}.
\begin{col}
  The Lambek calculus is complete w.r.t finite relational models (has the fmp).
\end{col}

Modulo the famous result by Pentus \cite{pentus2006lambek}, one may also conclude that

\begin{col}
  The theory of ${\bf IRRS}$ is $\operatorname{NP}$-complete.
\end{col}

\subsection{The second proof for Problem 19.17 (the case of residuated monoids)}

The construction above does not work properly for residuated monoids since the identity has the representation $\hat{{\bf 1}} = \{ (a,b) \: | \: a \leq {\bf 1} ; b, a \in \mathcal{G}(\mathcal{Q}), b \in \mathcal{Q}\} = \{ (a,b) \: | \: a \leq b, a \in \mathcal{G}(\mathcal{Q}), b \in \mathcal{Q}\}$, which is not the identity relation at all. Thus, the finite representation property for residuated monoids has to be proved differently.

First of all, let us recall we can prove the finite representation property might be shown for (ordered) monoids.

Let us put $\tau = \{ ;, {\bf 1} \}$. Recall that $\mathcal{A} \in {\bf R}(\tau)$ iff $\mathcal{A} = \langle A, ;, {\bf 1} \rangle$ is a monoid. That is, every monoid is representable. The finite representation property for finite monoids might be shown with a Cayley representation putting $A$ as a base set and defining:
\begin{center}
  $a^R = \{ (b, b ; a) \: | \: b \in A \}$
\end{center}
Thus, we have
\begin{lemma}
  Let $\tau \subseteq \{ ;,  \}$ ${\bf R}(\tau)$ has the finite representation property.
\end{lemma}

We also note that $\mathcal{A} \in {\bf R}(\leq)$ iff $\mathcal{A}$ is a poset defining a representation over the set of upper cones
$\operatorname{Up}(\mathcal{A})$ as
\begin{center}
  $a^R = \{ (X,X) \: | \: a \in X \in \operatorname{Up}(\mathcal{A}) \}$
\end{center}

In contrast to usual monoids, the universal theory of representable partially ordered monoids is not finitely axiomatisable \cite{hirsch2005class}.

\begin{lemma}
  ${\bf R}(;, {\bf 1}, \leq)$ has (?) the finite representation property.
\end{lemma}

\begin{proof}

\end{proof}

\begin{theorem}
  ${\bf R}(;, \setminus, /, {\bf 1}, \leq)$ has (?) the finite representation property.
\end{theorem}

\begin{proof}
\end{proof}

\section{Signatures containing either Boolean meet or negation}

It is known that the class ${\bf R}(\tau)$ has no the finite representation property where $\tau$ is either $\{ ;, \cdot, \smile\}$ or $\{ ;, \cdot, {\bf 1} \}$ \cite{hirsch2004finite}. That is, they have representations only over infinite bases. This fact has the following proof (a proof sketch, to be more precise).
\begin{proof}
Consider the set of atoms $\operatorname{A}(\mathcal{Q}) = \{ {\bf 1}, <, > \}$, where ${\bf 1}$ is the identity, $<^{\smile} = >$ and vice versa. The composition is defined with the following equations:
\begin{enumerate}
  \item $< ; < = <$,
  \item $< ; > = {\bf 1} + < + >$
\end{enumerate}

Let $R$ be a $\tau$-representation of $\mathcal{Q} \upharpoonright_{\tau}$ over the base $X$. $\leq$ is definable and $0$ is below $<$. Note that $0^{R} \leq <^R$, but it is not necessary that $0^R = \emptyset$. $< \neq 0$, thus, $(x,y) \in <^R \setminus 0^R$. If $(u, v) \in <^R \setminus 0^R$, then $u \neq v$.
The reason for $u \neq v$ is the following one. Suppose ${\bf 1} \in \tau$, then $(u,u) \in {\bf 1}^R$, so $u = v$ implies $(u, v) \in <^R \cdot {\bf 1} = 0^R$. Suppose $\smile \in \tau$,
then $u = v$ implies $(u,v) \in (< \cdot >)^R = 0$. Since $< \: \leq \: < ; <$ and $(u, v) \in <^R \setminus 0^R$, then there exists $w \in X$ such that $(u, w) \in <^R$ and $(u, w) \in <^R$. Hence, $(u, v), (v, W) \in <^R \setminus 0^R$. Thus, $<^R \setminus 0^R$ is reflexive and dense, but it is not transitive.

For each $n < \omega$ there exist $x_0, \dots, x_{n - 1} \in X$ such that for each $i < j < n$ one has $(x_i, x_j) \in <^R \setminus 0^R$. Thus, $n \leq |X|$ for each $n < \omega$.
\end{proof}

Moreover, the stronger theorem holds that has been proved by Maddux \cite{maddux2016finite}:
\begin{theorem}
  There exists a finite algebra $\mathcal{A} \in {\bf R}(\tau)$, where $\tau \subseteq \{ ;, \cdot \}$, such that $\mathcal{A}$ has no finite representation.
\end{theorem}

\begin{theorem}
  There exists a finite algebra $\mathcal{A} \in {\bf R}(\tau)$, where $\{ ;, \neg \} \subseteq$, such that $\mathcal{A}$ has no finite representation.
\end{theorem}

\begin{proof}
  Let $\operatorname{At}(\mathcal{A}) = \{ {\bf 1}, p, q \}$ be the set of atoms of a finite relation algebra $\mathcal{A}$ having the cardinality $8$, where ${\bf 1}$ is the identity.

  Let $\mathcal{A} \upharpoonright_{\tau}$ be the $\tau$-subreduct of $\mathcal{A}$ and $R$ its representation over the base $U$.

  {\bf Claim} For each $n < \omega$ there exists $x_0, \dots, x_{n - 1}$ such that for each $i < j < n$, $(x_i, x_j) \in X$, thus $|X| \geq \omega$.
\end{proof}

\begin{col}
  If $\tau$ contains composition and either Boolean meet or complement, then ${\bf R}(\tau)$ fails to have the finite representation property.
\end{col}

\begin{theorem}
  Suppose ${\bf R}(\tau)$ fails to have the finite representation property, then $\tau$ contains composition either meet or negation.
\end{theorem}

\bibliographystyle{plain}
\bibliography{Text}

\end{document}

### 2020-07-16 problem-19.17/Text.tex
solution for the problem 19.17 of Hirsch-Hodkinson-2002
started the section with the signatures containing Boolean meet

the FRP: the proof sketch for the fact that R(;, ., ^{\smile}) and R(;, ., 1') have no the frp

update and definitions

@@
 \date{}
-\title{The finite base property for some subreducts of representable relation algebras}
+\title{The finite base property for subreducts of representable relation algebras}
 
@@
 
-We describe the basic definitions and results about relation algebras \cite{hirsch2002relation} \cite{maddux2006relation}.
+We recall the basic definitions and results about relation algebras \cite{hirsch2002relation} \cite{maddux2006relation}.
 
@@
 \end{center}
-Moreover, ${\bf RRA}$ is a variety, but it cannot be defined by any set of first-order formulas \cite{monk1964representable}.
-
-One may express residuals in every $\mathcal{R} \in {\bf RA}$ as follows, for every $a, b \in \mathcal{R}$:
-
-\begin{enumerate}
-  \item $a \setminus b = -(a^{\smile} ; -b)$
-  \item $a / b = - (- a ; b^{\smile})$
-\end{enumerate}
-
-Those residuals have the following interpretation in $\mathcal{R} \in {\bf PRA}$ (as well as in ${\bf RRA}$), for every $a, b \in \mathcal{R}$:
-\begin{enumerate}
-  \item $a \setminus b = \{ \langle x, y \rangle \: | \: \forall z \: (z, x) \in a \Rightarrow (z, y) \in b \}$
-  \item $a / b = \{ \langle x, y \rangle \: | \: \forall z \: (y, z) \in b \Rightarrow (x, z) \in a \}$
-\end{enumerate}
-One may illustrate composition and residuals in ${\bf PRA}$ and ${\bf RRA}$ via the following triangles:
-
-\xymatrix{
-&&& \exists y \ar@{-->}[ddr]^{b} &&& \forall z \ar@{-->}[ddl]_{a} \ar@{-->}[ddr]^{b} &&& \forall z \\
-&&&&&& \Rightarrow &&& \Leftarrow \\
-&& x \ar@{-->}[uur]^{a} \ar[rr]_{a;c} && z & x \ar[rr]_{a \setminus b} && y & x \ar@{-->}[uur]^{a} \ar[rr]_{a / b} && y \ar@{-->}[uul]_{b}
-}
-
-Given a subset of definable operations in ${\bf RA}$ $\tau$, we denote the class of subalgebras of the
-$\tau$-reducts by ${\bf R}(\tau)$. The algebras containing to this class are defined as restrictions of
-elements belonging to ${\bf Rs}$ to operations of $\tau$.
-By ${\bf Q}(\tau)$ we mean a quasivariety generated by ${\bf R}(\tau)$.
-As in \cite{hirsch2011positive}, we put ${\bf Q}(\tau)$ as the closure of ${\bf R}(\tau)$ under subalgebras and products assuming that ${\bf R}(\tau)$ is already closed under ultraproducts.
-
-\section{The Finite Base Property}
-
-We recall the underlying definitions according to \cite[Section 19]{hirsch2002relation}
-
-\begin{defin}
-  Let ${\bf K}$ be a class of algebras of a signature $\Omega$, ${\bf K}$ has the finite algebra property, if
-  if any first-order $\Omega$-sentence that is true in all finite algebras in ${\bf K}$ is true in every algebra in ${\bf K}$.
-\end{defin}
-
-The finite base property is a version of the finite algebra property if ${\bf K}$ is a class of representable algebras:
-
-\begin{defin}
-Let ${\bf K}$ be a class of representable algebras of a signature $\Omega$
-
-  \begin{enumerate}
-  \item ${\bf K}$ has the finite base property if any first-order $\Omega$-sentence that is true in every algebra in ${\bf K}$ having a representation over a finite base set is valid in ${\bf K}$.
-  \item ${\bf K}$ has the finite algebra on finite base property if any finite algebra in ${\bf K}$ has a representation with finite base.
-  \item ${\bf K}$ has the finite algebra property for equations/quasi-identites if any equation/quasi-identity that is true in all finite algebras is true in every algebra in ${\bf K}$.
-The finite base property for equations/quasi-identites is defined similarly.
-\end{enumerate}
-\end{defin}
-
-The following statements were shown in \cite{andreka1999finite}. This lemma connects finite base property with finite algebra on finite base and finite algebra properties as follows:
-\begin{lemma}
-  Let ${\bf K}$ be a class of representable $\Omega$-algebras:
-  \begin{enumerate}
-    \item If ${\bf K}$ has the finite algebra property, then it has the finite algebra and the finite base properties for equations/quasi-identites.
-    \item The finite algebra on finite base and the finite algebra properties implies the finite base property for ${\bf K}$. The same holds for equations/quasi-identites.
-    \item If any representation of an infinite algebra has an infinite base, then the finite base property implies the finite algebra one for ${\bf K}$.
-    \item Suppose $\Omega$ is finite and any subalgebra of a representable algebra is representable on the same base. Then the finite base property implies the finite algebra on finite base property.
-  \end{enumerate}
-\end{lemma}
-
-\section{The Relation Residuated Semigroups Background}
-
-\subsection{The underlying definitions and results}
-
- A \emph{relation structure} (${\bf RS}$) is an arbitrary algebra of the signature $\Omega = \langle \circ, \setminus, /, \leq \rangle$, where $\circ, \setminus, /$ are binary function symbols and $\leq$ is a binary relation symbol.
-
-\begin{defin}
-  A residuated semigroup is an algebra $\mathcal{S} = \langle S, \circ, \leq, \setminus, / \rangle$ such that $\langle S, \circ, \leq, \rangle$ is an ordered residuated semigroup and the following equivalences hold for each $a, b, c \in S$:
-
-  \begin{center}
-    $b \leq a \setminus c \Leftrightarrow a \circ b \leq c \Leftrightarrow a \leq c / b$
-  \end{center}
-  ${\bf ORS}$ is the class of all residuated semigroups.
-\end{defin}
-
-See this paper to have a proof of the following theorem \cite{goldblatt2006kripke}:
-
-\begin{theorem}
-  Every finite residuated semigroup is isomorphic to some residuated subsemigroup of some finite residuated lattice.
-\end{theorem}
-
-\begin{defin} \label{rrs}
-  Let $A$ be a set of binary relations on some base set $W$ such that $R = \cup A$ is transitive and $\{ x, y \: | \: x R y \} = W$.
-A relation residuated semigroup is an algebra $\mathcal{A} = \langle A, ;, \setminus, /, \subseteq \rangle$ where for each $r, s \in A$
-  \begin{enumerate}
-    \item $r ; s = \{ \langle a, c \rangle \: | \: \exists b \in W \: (\langle a, b \rangle \in r \: \& \: \langle b, c \rangle \in s) \}$
-    \item $r \setminus s = \{ \langle a, c \rangle \: | \: \forall b \in W \: (\langle b, a \rangle \in r \Rightarrow \langle b, c \rangle \in s)\}$
-    \item $r / s = \{ \langle a, c \rangle \: | \: \forall b \in W \: (\langle c, b \rangle \in s \Rightarrow \langle a, b \rangle \in r)\}$
-    \item $r \leq s$ iff $r \subseteq s$.
-  \end{enumerate}
-\end{defin}
-Relation residuated semigroup are also called representable relativised relational structure (${\bf RRS}$).
-
-See \cite{brown1993representation}
-\begin{theorem}
-  Every complete residuated semigroup $\mathcal{A}$ (quantale) is isomorphic to relational complete residuated semigroup on the underlying set of $\mathcal{A}$.
-\end{theorem}
-
-\begin{defin}
-  Let $\mathcal{A} = \langle A, \leq, ;, \setminus, / \rangle$ be a residuated semigroup. A representation of $\mathcal{A}$ is an inclusion map $h : \mathcal{A} \to \mathcal{A}'$,
-  where $\mathcal{A}' \in {\bf RRS}$ such that:
-  \begin{enumerate}
-    \item $dom(\mathcal{A}') = \{ \hat{a} \: | \: a \in A\}$, where $\hat{a} = \{ (b, c) \: | \: b \leq a ; c \}$.
-  \end{enumerate}
-\end{defin}
-
-Such a map preserves order, residuals, and composition.
-Andr\'{e}ka and Mikul\'{a}s proved the following representation theorem for ${\bf ORS}$ in \cite{andreka1994lambek} that implies relational completeness of the Lambek calculus, the logic of ${\bf ORS}$:
-
-\begin{theorem}\label{ors=irrs}
-  ${\bf ORS} = {\bf IRRS}$, where ${\bf IRRS}$ is a closure of ${\bf RRS}$ under isomorphic copies.
-\end{theorem}
-
-\begin{col}
-  Every finite representable residuated semigroup is isomorphic to representable residuated subsemigroup of some finite residuated lattice.
-\end{col}
-
-\begin{theorem}
-  Let $\mathcal{A} \in {\bf RRS}$ and $|\mathcal{A}| < \omega$, then there exists a set $W$, a set $A$ of binary relations on $W$, $R = \cup A$ with $dom(R) = A$ such that
-  $\mathcal{A} \cong \langle A, |, \setminus, /, \subseteq \rangle$.
-\end{theorem}
-
-\begin{theorem}
-  The Lambek calculus is complete w.r.t finite relational models (has the fmp).
-\end{theorem}
+${\bf RRA}$ is a variety, but it cannot be defined by any set of first-order formulas.
+It is also known that ${\bf RA}$ and ${\bf RRA}$ are equational classes, but ${\bf RRA}$ is not finitely axiomatisable.
 
@@
 \item yellows: ${\bf y}_i$ for $1 \leq i \leq n$,
-\item ivory: ${\bf i}$,
 \item reds: ${\bf r}_i$ for $1 \leq i \leq n$,
-\item blacks: ${\bf b}_i$ for $1 \leq i \leq n$.
+\item blacks: ${\bf b}_i$ for $1 \leq i \leq n$,
+\item ivory: ${\bf i}$.
+\item orange: ${\bf o}$
 \end{itemize}
@@
 {\bf Step 1}.
-Let $\mathcal{A}_n$ be the upper semilattice presented with the set $\operatorname{At}(\mathfrak{A}_n) \cup \{ 0 \}$ and the following relations for each $x \in \operatorname{At}(\mathfrak{A}_n)$ and for each $1 \leq i \leq n$.
+Let $\mathcal{A}_n$ be the Boolean algebra presented with the set of generators $\operatorname{At}(\mathfrak{A}_n)$ and the following relations for each $x \in \operatorname{At}(\mathfrak{A}_n)$ and for each $1 \leq i \leq n$.
 \begin{enumerate}
-  \item ${\bf w} \leq {\bf g}_i + {\bf y}_i$,
-  \item ${\bf i} \leq {\bf y}_i + {\bf r}_i$,
-  \item $x + 0 = x$
+  \item ${\bf w} + {\bf g}_i + {\bf y}_i = {\bf g}_i + {\bf y}_i$,
+  \item ${\bf i} + {\bf y}_i + {\bf r}_i = {\bf y}_i + {\bf r}_i$.
 \end{enumerate}
@@
 \begin{enumerate}
-  \item $1 = \sum \operatorname{At}(\mathcal{A}_n) \cup \{ 0 \}$,
-  \item $0 = \O$,
+  \item $1 = \sum \operatorname{At}(\mathcal{A}_n)$,
   \item $x = x^{\smile}$,
@@
   \item ${\bf 1} ; x = {\bf 1} ; x = x$,
-  \item $x ; y = \begin{cases} {\bf i}, \text{ if $\{x, y\} \in S$} \\ 1, \text{ otherwise } \end{cases}$ \\ unless $x, y \in \{ 0, {\bf 1}\}$.
+  \item $x ; y = \begin{cases} {\bf o}, \text{ if $\{x, y\} \in S$} \\ 1, \text{ otherwise } \end{cases}$ \\ unless $x, y \in \{ 0, {\bf 1}\}$.
 \end{enumerate}
@@
 Such a graph have the following form.
-$N = \langle E_N, l_N \rangle$, where $E_N = U_N \times U_N$ for some finite base set and $l_n : E_N \to \operatorname{At}(\mathcal{A})$ is function mapping each edge to some atom of $A$.
+$N = \langle U_N, E_N, l_N \rangle$, where $E_N = U_N \times U_N$ for some finite base set and $l_n : E_N \to \mathcal{A}$ is function mapping each edge to some atom of $A$.
 This function obey the following requirements:
@@
   \end{enumerate}
+  A network is called atomic if all the edges are labelled by atoms.
 \end{defin}
@@
 
-\begin{proof}
-  Let $n \geq l$. We show that $\exists$ has a winning strategy for $G_l(\mathfrak{A}_n)$.
-\end{proof}
+Let us consider the following sequence of networks $\{ \mathcal{N}_n \}_{n < \omega}$ labelled by atoms of $\mathfrak{A}_n$.
+
+\begin{enumerate}
+  \item $n = 0$, then the $N_0 = \langle U_0, E_0, l_0 \rangle$ has the form:
+\end{enumerate}
+
+\xymatrix{
+\\
+&&&&& u \ar@(ul,ur) \ar@{-}[rr] && v \ar@(ul,ur)
+}
+
+\begin{claim}
+\end{claim}
 

### 2020-10-15 problem-19.17/FRP.tex
cylindric stuff: atom structures and canonical extensions

@@
 
-A $\tau$-structure is \emph{representable} if its isomorphic to some algebra of relations of $\tau$-signature. A representable
-finite $\tau$-structure has a \emph{finite representation over a finite base} if it is isomoprhic to some finite representable over a
-finite base.
+A $\tau$-structure is \emph{representable} if its isomorphic to some algebra of relations of $\tau$-signature.
+A representable finite $\tau$-structure has a \emph{finite representation over a finite base} if it is isomoprhic to some finite representable over a finite base.
 ${\bf R}(\tau)$ has the finite representation property if every $\mathcal{A} \in {\bf R}(\tau)$ has a finite representation over a
@@
 One may reformulate this result as $\mathcal{A} \in {\bf R}(;, \setminus, /, \leq)$ iff $\mathcal{A}$ is a residuated semigroup.
+Thus, it is sufficient to show that any finite residuated semigroup has a representation over a finite base.
 
@@
 
-In \cite{hirsch2002relation}, Hirsch and Hoskidson left the problem about having the finite representation property.
-Given a finite algebra $\mathcal{A} \in {\bf RRS}$, we show that $\mathcal{A}$ is isomorphic to some algebra $\mathcal{B}$ that has the finite base.
+Hirsch and Hoskidson left the problem about having the finite representation property.
+Given a finite algebra $\mathcal{A} \in {\bf RRS}$, we show that $\mathcal{A}$ is isomorphic to some algebra over a finite base.
 
@@
 \end{lemma}
-
 \begin{proof}
@@
 \begin{lemma}
-  Let $\tau \subseteq \{ ;,  \}$ ${\bf R}(\tau)$ has the finite representation property.
+  Let $\tau \subseteq \{ ;, \leq \}$, then ${\bf R}(\tau)$ has the finite representation property.
 \end{lemma}
@@
 
+It is known that $\mathcal{A} \in {\bf R}(;, \leq)$ iff $\mathcal{A}$ is an ordered semigroup \cite{zaretskii1959representation} and the class ${\bf R}(;, \leq)$ is finitely axiomatisable.
+
 In contrast to usual monoids, the universal theory of representable partially ordered monoids is not finitely axiomatisable \cite{hirsch2005class}.
+However, the universal theory of representable residuated monoids coincide with the theory of the whole class of residuated monoids and, thus, is finitely axiomatisable \cite{andreka1994lambek}.
+Let us recall how the uniform representation for residuated monoids looks like.
+In fact, we reproduce the proof sketch of Theorem~\ref{ors=irrs} assuming that a given algebra is finite.
 
-\begin{lemma}
-  ${\bf R}(;, {\bf 1}, \leq)$ has (?) the finite representation property.
-\end{lemma}
+\begin{defin}
+  A network
+\end{defin}
 
-\begin{proof}
+TODO: a desired network
 
-\end{proof}
+TODO: inductive steps in a nutshell
+
+TODO: an interpretation
+
+TODO: extract the finite base from a representation
 
 \begin{theorem}
-  ${\bf R}(;, \setminus, /, {\bf 1}, \leq)$ has (?) the finite representation property.
+  ${\bf R}(;, \setminus, /, {\bf 1}, \leq)$ has the finite representation property.
 \end{theorem}

### 2020-10-15 problem-9.4/Text.tex
a small cylindric update
update
wip: rainbow construction for R(;, \vee)
cylindric stuff: atom structures and canonical extensions

@@
 
+\theoremstyle{fact}
+\newtheorem{fact}{Fact}
+
 \theoremstyle{ex}
@@
 
+\section{The problems themselves}
+
+\begin{enumerate}
+\item Suppose $\mathcal{C} \in {\bf RCA}_{\omega}$, whether $\mathcal{C}$ has a complete, $\omega$-dimensional representation?
+\item Is the class ${\bf IG}_{\omega}$ (the isomorphism-closure of the $\omega$-dimensional cylindric relativised set algebras in which the unit is closed under substitutions and permutations) a variety, or even
+a pseudo-elementary class? Is it closed under ultraproducts?
+\end{enumerate}
+
+\section{Boolean algebras with operators and cylindric algebras}
+
+\begin{defin}
+
+$ $
+
+  \begin{enumerate}
+    \item Let $\mathcal{B} = \langle B, +, -, 0, 1 \rangle$ be a Boolean algebra. An operator is an $n$-ary function $\Omega : B^n \to B$ satisfying the following conditions:
+    \begin{itemize}
+      \item Normality: for all $b_0, \dots, b_{n - 1} \in B$, if $b_1 = 0$ for some $i < n$, then
+
+\begin{center}
+      $\Omega(b_0, \dots, b_{i - 1}, 0, b_{i+1}, \dots, b_{n - 1}) = 0$
+\end{center}
+      \item Additivity: for all $b_0, \dots, b_{n - 1}, b, b' \in B$ we have
+      \begin{center}
+        $\Omega(b_0, \dots, b_{i - 1}, (b + b'), b_{i+1}, \dots, b_{n - 1}) = \Omega(b_0, \dots, b_{i - 1}, b, b_{i+1}, \dots, b_{n - 1}) + \Omega(b_0, \dots, b_{i - 1}, b', b_{i+1}, \dots, b_{n - 1})$
+      \end{center}
+    \end{itemize}
+    \item Let $I$ be an index set, a Boolean algebra with operators (BAO) is an algebra $\langle B, +, -, 0, 1, \{ \Omega_{i} \}_{i \in I} \rangle$ such that $\langle B, +, -, 0, 1 \rangle$ is a Boolean algebra and for each $i \in I$ $\Omega_{i}$ is an operator.
+  \end{enumerate}
+\end{defin}
+
+\begin{defin} Let $\mathcal{B} = \langle B, +, -, 0, 1, \{ \Omega_{i} \}_{i \in I} \rangle$ be a BAO, then
+
+  \begin{enumerate}
+    \item An opetator $\Omega$ is completely additive, if for each $b_0, \dots, b_{n - 1} \in B$ and $X \subseteq B$, one has
+
+    \begin{center}
+      $\Omega(b_0, \dots, b_{i-1}, \sum X, b_{i+1}, \dots, b_{n - 1}) = \sum \limits_{x \in X} \Omega(b_0, \dots, b_{i-1}, x, b_{i+1}, \dots, b_{n - 1})$
+    \end{center}
+    \item $\mathcal{B}$ is completely additive, if for each $i \in I$ $\Omega_{i}$ is additive,
+    \item A class $\mathcal{K}$ of BAOs is completely additive, if every $\mathcal{B} \in \mathcal{K}$ is completely additive.
+  \end{enumerate}
+\end{defin}
+
+\subsection{Atom structures and canonical extensions}
+
+\begin{defin}  Let $I$ be an index set and $\{ \Omega_i\}_{i \in I}$ a set of function symbols
+\begin{enumerate}
+  \item An atom structure is a relational structrure
+  $\mathcal{F} = \langle W, \{ R_{i} \}_{i \in I} \rangle$
+  such that $R_{i}$ is a $n+1$-ary relation symbol, if $\Omega_{i \in I}$ is an $n$-ary function symbol,
+  \item Let $\mathcal{B}$ be an atomic BAO of the signature $I$,
+the atom structure of $\mathcal{B}$, written as ${\bf At} \mathcal{B}$, is an atom structure $\langle \operatorname{At}(\mathcal{B}), \{ R_{i}\}_{i \in I} \rangle$ such that for each
+$a, b_0, \dots, b_{n + 1} \in \operatorname{At}(\mathcal{B})$ and for each $i \in I$
+\begin{center}
+  ${\bf At} \mathcal{B} \models R_{i}(a,  b_0, \dots, b_{n + 1})$ iff $\mathcal{B} \models a \leq \Omega_{i}(b_0, \dots, b_{n + 1})$
+\end{center}
+\item Let $\mathcal{F} = \langle W, \{ R_{i} \}_{i \in I} \rangle$ be an atom structure, the complex algebra of $\mathcal{F}$, written as $\bf{Cm} \mathcal{F}$, is a BAO
+$\langle \mathcal{P}(W), \cup, -, \emptyset, W, \{\Omega_{R_{i}}\}_{i \in I} \rangle$ such that
+for all $X_0, \dots, X_{n - 1} \subseteq W$ and for each $i \in I$
+\begin{center}
+  $\Omega_{R_{i}}(X_0, \dots, X_{n - 1}) = \{ a \in W \: | \: \exists b_0 \in X_0 \dots \exists b_{n -1} \in X_{n - 1} \: \mathcal{F} \models R_{i}(a, b_0, \dots, b_{n - 1})\}$
+\end{center}
+\end{enumerate}
+\end{defin}
+
+The following duality is due to Thomason \cite{thomason1975categories}.
+
+\begin{fact}
+  $ $
+
+  \begin{enumerate}
+    \item Let $\mathcal{B}$ be a complete atomic BAO, then $\mathcal{B} \cong \bf{Cm} (\bf{At}(\mathcal{B}))$,
+    \item Let $\mathcal{F}$ be an atom structure, then $\mathcal{F} \cong {\bf At}({\bf Cm}(\mathcal{B}))$.
+  \end{enumerate}
+\end{fact}
+
+Let $A$ be a non-empty subset of a Boolean algebra $\mathcal{B}$, $A$ is a \emph{filter}, if $A$ is closed under finite infima and upwardly closed.
+$A$ is an ultrafilter, if it has no non-trivial extensions.
+That is, if $A \subseteq A'$, then $A' = \mathcal{B}$.
+
+\begin{defin}
+  Let $\mathcal{B} = \langle B, +, -, 0, 1, \{ \Omega_i \}_{i \in I} \rangle$ be a BAO and ${\bf Uf}(\mathcal{B})$ the set of its ultrafilters.
+The ultrafilter frame of $\mathcal{B}$ (or canonical frame) is a relational structure $\mathcal{F}_{\mathcal{B}} = \langle {\bf Uf}(\mathcal{B}), R_{\Omega_i} \rangle$ such that for each ultrafilters $\beta_0, \dots, \beta_{n - 1}, \gamma$ one has
+  \begin{center}
+    ${\bf Uf}(\mathcal{B}) \models R_{\Omega_i}(\beta_0, \dots, \beta_{n - 1}, \gamma)$ iff $\{ \Omega(b_0, \dots, b_{n - 1}) \: | \: b_0 \in \beta_0, \dots, b_{n - 1} \in \beta_{n - 1}\} \subseteq \gamma$.
+  \end{center}
+\end{defin}
+
+\begin{defin} Let $\mathcal{B}$ be a BAO, then
+
+  \begin{enumerate}
+    \item The canonical extension of $\mathcal{B}$ is a complex algebra of the canonical frame ${\bf Cm}(\mathcal{F}_{\mathcal{B}})$ denoted as $\mathcal{B}^{+}$,
+    \item The class of BAOs is canonical, if it is closed under canonical extensions.
+  \end{enumerate}
+\end{defin}
+
+\begin{theorem} Let $\mathcal{A}$, $\mathcal{B}$ be BAOs,
+
+\begin{enumerate}
+  \item There exists $\iota : \mathcal{A} \hookrightarrow \mathcal{A}^{+}$ such that
+  $\iota : a \mapsto \{ \gamma \in {\bf Uf}(\mathcal{A}) \: | \: a \in \gamma \}$.
+  \item If $i : \mathcal{A} \hookrightarrow \mathcal{B}$, then this embedding might be extented to the embedding
+  $i^{+} : \mathcal{A}^{+} \hookrightarrow \mathcal{B}^{+}$
+\end{enumerate}
+\end{theorem}
+
+\begin{fact}
+\end{fact}
+
+\subsection{(Representable) cylindric algebras and cylindric set algebras}
+
+Let $\alpha$ be an ordinal. Let $U^{\alpha}$ be the set of all functions mapping $\alpha$ to a non-empty set $U$. We denote $x(i) = x_i$ for
+$x \in U^{\alpha}$ and $i < \alpha$.
+
+\begin{defin}
+$ $
+
+  \begin{enumerate}
+    \item A subset of $U^{\alpha}$ is an $\alpha$-ry relation on $U$. For $i, j < \alpha$, the $i,j$-diagonal $D_{ij}$ is the set of all
+    elements of $U$ such that $y_i = y_j$. If $i < \alpha$ and $X$ is an $\alpha$-ry relation on $U$, then
+    the $i$-th cylindrification $C_i X$ is the set of all elements of $U$ that agree with some element of $X$ on each coordinate except the
+    $i$-th one. To be more precise,
+    $C_i X = \{ y \in U^{\alpha} \: | \: \exists x \in X \forall i < \alpha \: (i \neq j \Rightarrow y_j = x_j)\}$.
+    \item A cylindic set algebra of dimension $\alpha$ is an algebra consisting of a set $S$ of $\alpha$-ry relation on some base set $U$
+    with the constants and operations $0 = \emptyset$, $1 = U^{\alpha}$, $\cap$, $-$, the diagonal elements $\{ D_{ij} \}_{i, j < \alpha}$,
+    the cylindrifications $\{ C \}_{i < \alpha}$. A generalised cylindric set algebra of dimension $\alpha$ is a subdirect of
+    cylindric algebras that have dimension $\alpha$
+    \item A cylindric algebra of dimension $\alpha$ is an algebra $\mathcal{C} = \langle \mathcal{B}, \{ c_i \}_{i < \alpha}, \{ d_{ij} \}_{i, j < \alpha} \rangle$ such that
+    \begin{itemize}
+      \item $\mathcal{B}$ is a Boolean algebra, for each $i, j < \alpha$ $c_i$ is an operator and $d_{ij} \in \mathcal{B}$
+      \item For each $i < \alpha$, $a \leq c_i a$, $c_i (a \land c_i b) = c_i a \land c_i b$ and $d_{ii} = 1$
+      \item For every $i, j < \alpha$, $c_i c_j a = c_j c_i a$
+      \item If $k \neq i, j < \alpha$, then $d_{ij} = c_k (d_{ij} \land d_{jk})$
+      \item If $i \neq j$, then $c_i (d_{ij} \land a) \land c_i (d_{ij} \land - a) = 0$
+    \end{itemize}
+    ${\bf CA}_{\alpha}$ is the class of all cylindric algebras of dimension $\alpha$
+    \item An $\alpha$-dimensional cylindric algebra $C$ is representable, if it is isomorphic to a generalised cylindric set algebra
+    of dimension $\alpha$. Such is isomorphism is a representation of $C$.
+    ${\bf RCA}_{\alpha}$ is the class of all representable cylindric algebras that have dimension $\alpha$. In particular, we are interested in the case when $\alpha = \omega$.
+  \end{enumerate}
+\end{defin}
+
+\subsection{Substitution in cylindric algebras}
+
+\begin{defin}
+  Given a cylindric algebra of dimension $\alpha$ $C$, let $x$ be a term of its signature, the substitution operator $s^{i}_{j}$ have the following definition:
+  \begin{center}
+  $s^{i}_{j} x = \begin{cases} x, \text{if } i = j \\ c_i (d_{ij} \land x), \text{otherwise} \end{cases}$
+  \end{center}
+\end{defin}
+
+\begin{prop} Let $\alpha$ be an ordinal and let $i, j, k, l < \alpha$. The following facts hold in ${\bf CA}_{\alpha}$
+
+  \begin{enumerate}
+    \item $s^{i}_j x \leq c_i x$.
+    \item $s^{i}_j (x \land y) = s^{i}_j x \land s^{i}_j y$, $s^{i}_j (x \lor y) = s^{i}_j x \lor s^{i}_j y$, $- s^{i}_j x = s^{i}_j (- x)$.
+    Moreover, $s^{i}_j$ is completely additive.
+    \item $i \neq k, l$ implies $s^i_j d_{ik} = d_{jk}$ and $s^i_j d_{kl} = d_{kl}$.
+    \item $d_{jk} \land s^{i}_j = d_{jk} \land s^{i}_k$.
+    \item $s^i_j c_i x = c_i x$.
+    \item $k \neq i,j$ implies $s^i_j c_i x = c_i s^i_j x$.
+    \item $c_j s^i_j x = c_i s^j_i x$.
+    \item $i \neq j$ implies $c_i s^i_j x = s^i_j x$.
+    \item $i \neq k$ implies $s^i_j s^i_k = s^i_k x$.
+    \item If either $i \notin \{ k, l\}$ and $k \notin \{ i, j\}$, or $j = l$, then $s^i_j s^k_l x = s^k_l s^i_j x$.
+    \item $s^i_j s^j_i x = s^i_j x$.
+    \item $s^i_k s^j_i x = s^i_k s^j_k x = s^j_k s^i_j x$
+  \end{enumerate}
+\end{prop}
+
 \section{Model-theoretic and universal algebraic preliminaries}
@@
 Here are the required notions and facts from model theory and universal algebra \cite{hirsch2002relation} \cite{hodges1997shorter} \cite{sankappanavar1981course}.
-
-Let $A$ be a non-empty set, an \emph{ultrafilter} on $A$ is a set of subsets $U \subseteq \mathcal{P}(\mathcal{P})(A)$ such that $A$ is closed under intersections, $\subseteq$-upwardly closed, and either $X \in U$ or $- X \in U$, where $X \subseteq A$.
-An ultrafilter is called principal if it has the form $\uparrow X = \{ Y \in \mathcal{P}(A) \: | \: X \subseteq Y \}$.
-
-Let $\Omega = \langle \operatorname{Cnst}, \operatorname{Fn}, \operatorname{Pred} \rangle$ be a signature and $\Lambda$ an index set, and let $\{ M_{\lambda}\}_{\lambda \in \Lambda}$ be an indexed set of $\Omega$ structures.
-The $\Omega$-structure
-\begin{center}
-  $M = \prod \limits_{\lambda \in \Lambda} M_{\lambda}$
-\end{center}
-is called a \emph{product} that defined as follows.
-Its domain is the Cartesian product of the domains of $M_{\lambda}$.
-$a \in M$ is a function $\Lambda \to \bigcup \limits_{\lambda \in \Lambda} \operatorname{dom}(M_{\lambda})$ such that $a(\lambda) \in M_{\lambda}$ for each $\lambda \in \Lambda$.
-Given $\lambda \in \Lambda$ and $a_{\lambda} \in M_{\lambda}$, we denote the function mapping $\lambda$ to $a_{\lambda}$ as
-$\langle a_{\lambda} \: | \: \lambda \in \Lambda \rangle$. We define the interpretation of $\Omega$-symbols as
-
-\begin{enumerate}
-  \item If $c \in \operatorname{Cnst}$, then $c^{M} = \langle c^{M_{\lambda}} \: | \: \lambda \in \Lambda \rangle$
-  \item If $f \in \operatorname{Fn}$ is an $n$-ary function symbol, then $f^{M}(\overline{a}) = \langle f^{M_{\lambda}}(\overline{a}) \: | \: \lambda \in \Lambda \rangle$, where $\overline{n} \in M^{n}$
-  \item If $R \in \operatorname{Pred}$ is an $n$-ary relation symbol and $\overline{n} \in M^{a}$, then $R^{M}(\overline{a}) = \langle R^{M_{\lambda}}(\overline{a}) \: | \: \lambda \in \Lambda \rangle$
-\end{enumerate}
-
-Given $\lambda \in \Lambda$, we define the $\lambda$th projection as $\pi_{\lambda} : M \to M_{\lambda}$ such that $\pi_{\lambda}(a) = a(\lambda)$.
 
@@
 
-\subsection{Preliminaries from universal algebra}
-
-\begin{defin}
-  Let ${\bf K}$ be a class of $\Omega$-structures, then ${\bf K}$ is a variety, if it is defined by some set of equations.
-  The variety generated by ${\bf K}$ is the smallest variety contatining ${\bf K}$. ${\bf K}$ is a quasi-variety, if it is defined by some
-  set of quasi-identities.
-\end{defin}
-
-Given a class ${\bf K}$ of $\Omega$-structures, then ${\bf I}({\bf K})$, ${\bf S}({\bf K})$, ${\bf H}({\bf K})$, and ${\bf P}({\bf K})$ are
-the classes of isomorphic copies, algebras isomorphic to subalgebras belonging to ${\bf K}$,
-algebras isomorphic to homomorphic images belonging to ${\bf K}$, and algebras isomorphic to direct products belonging to ${\bf K}$.
-We claim that ${\bf I}({\bf K}) \subseteq {\bf S}({\bf K})$. ${\bf Up}({\bf K})$ is the class of algebras isomorphic to ultraproducts
-belonging to ${\bf K}$.
-
-\begin{theorem} Let ${\bf K}$ be a class of $\Omega$-structures
-
-  \begin{enumerate}
-    \item ${\bf K}$ is a variety iff ${\bf H}({\bf K}), {\bf S}({\bf K}), {\bf P}({\bf K}) \subseteq {\bf K}$
-    \item ${\bf HSP}({\bf K}) = {\bf H}({\bf S}({\bf P}({\bf K})))$ is the smallest variety containg ${\bf K}$
-    \item ${\bf K}$ is a quasi-variety iff it is closed under subalgebras, products, and ultraproducts, iff ${\bf SPUp}({\bf K}) = {\bf K}$.
-  \end{enumerate}
-\end{theorem}
-
-\subsection{Subdirect products}
-
-\begin{defin}
-  $ $
-
-  \begin{enumerate}
-    \item Let $\{ A \}_{i \in I}$ be $\Omega$-structures, a subdirect product of $\langle A_i \: | \: i \in I \rangle$ is a
-    subalgebra $B$ of $\prod \limits_{i \in I} A_i$ such that for each $i \in I$, a projection map $\pi_i : B \to A_i$ is a surjection.
-    \item A subdirect representation of an $\Omega$-structure is an embedding $f : A \to \prod \limits_{i \in I} A_i$ for some $I$ and
-    $\{ A_i \}_{i \in I}$ such that $f \circ \pi_i : A \to A_i$ is a surjection.
-    \item An $\Omega$-structure $A$ is subdirectly irreducible if for every subdirect representation $f : A \to \prod \limits_{i \in I} A_i$
-    there exists a projection $\pi_i$ such that $f \circ \pi_i$ is an isomorphism.
-    \item ${\bf Sir}({\bf K})$ is the class of subdirectly irreducible structures belonging to ${\bf K}$.
-    \item A subdirect decomposition of $A$ if there exists a subdirect representation $f : A \to \prod \limits_{i \in I} A_i$ such that
-    every $A_i$ is subdirectly irreducible.
-  \end{enumerate}
-\end{defin}
-
-It is known that every Boolean algebra with operators has a subdirect demcomposition. Moreover, that implies:
-\begin{theorem}
-$ $
-
-  \begin{enumerate}
-  \item If ${\bf K}$ is a variety, then every element of ${\bf K}$ has a subdirect decomposition with some subdirect irreducible elements of
-  ${\bf K}$.
-  \item If ${\bf K}$ is a variety and $\varepsilon$ is an equation,
-  ${\bf Sir}({\bf K}) \models \varepsilon \Leftrightarrow {\bf K} \models \varepsilon$.
-\end{enumerate}
-\end{theorem}
-
-\subsection{Pseudo-elementary classes}
-
-\section{Cylindric algebras}
-
-\subsection{(Representable) cylindric algebras and cylindric set algebras}
-
-Let $\alpha$ be an ordinal. Let $U^{\alpha}$ be the set of all functions mapping $\alpha$ to a non-empty set $U$. We denote $x(i) = x_i$ for
-$x \in U^{\alpha}$ and $i < \alpha$.
-
-\begin{defin}
-$ $
-
-  \begin{enumerate}
-    \item A subset of $U^{\alpha}$ is an $\alpha$-ry relation on $U$. For $i, j < \alpha$, the $i,j$-diagonal $D_{ij}$ is the set of all
-    elements of $U$ such that $y_i = y_j$. If $i < \alpha$ and $X$ is an $\alpha$-ry relation on $U$, then
-    the $i$-th cylindrification $C_i X$ is the set of all elements of $U$ that agree with some element of $X$ on each coordinate except the
-    $i$-th one. To be more precise,
-    $C_i X = \{ y \in U^{\alpha} \: | \: \exists x \in X \forall i < \alpha \: (i \neq j \Rightarrow y_j = x_j)\}$.
-    \item A cylindic set algebra of dimension $\alpha$ is an algebra consisting of a set $S$ of $\alpha$-ry relation on some base set $U$
-    with the constants and operations $0 = \emptyset$, $1 = U^{\alpha}$, $\cap$, $-$, the diagonal elements $\{ D_{ij} \}_{i, j < \alpha}$,
-    the cylindrifications $\{ C \}_{i < \alpha}$. A generalised cylindric set algebra of dimension $\alpha$ is a subdirect of
-    cylindric algebras that have dimension $\alpha$
-    \item A cylindric algebra of dimension $\alpha$ is an algebra $\mathcal{C} = \langle \mathcal{B}, \{ c_i \}_{i < \alpha}, \{ d_{ij} \}_{i, j < \alpha} \rangle$ such that
-    \begin{itemize}
-      \item $\mathcal{B}$ is a Boolean algebra, for each $i, j < \alpha$ $c_i$ is an operator and $d_{ij} \in \mathcal{B}$
-      \item For each $i < \alpha$, $a \leq c_i a$, $c_i (a \land c_i b) = c_i a \land c_i b$ and $d_{ii} = 1$
-      \item For every $i, j < \alpha$, $c_i c_j a = c_j c_i a$
-      \item If $k \neq i, j < \alpha$, then $d_{ij} = c_k (d_{ij} \land d_{jk})$
-      \item If $i \neq j$, then $c_i (d_{ij} \land a) \land c_i (d_{ij} \land - a) = 0$
-    \end{itemize}
-    ${\bf CA}_{\alpha}$ is the class of all cylindric algebras of dimension $\alpha$
-    \item An $\alpha$-dimensional cylindric algebra $C$ is representable, if it is isomorphic to a generalised cylindric set algebra
-    of dimension $\alpha$. Such is isomorphism is a representation of $C$.
-    ${\bf RCA}_{\alpha}$ is the class of all representable cylindric algebras that have dimension $\alpha$.
-  \end{enumerate}
-\end{defin}
-
-\subsection{Substitution in cylindric algebras}
-
-\begin{defin}
-  Given a cylindric algebra of dimension $\alpha$ $C$, let $x$ be a term of its signature, the substitution operator $s^{i}_{j}$ have the following definition:
-  \begin{center}
-  $s^{i}_{j} x = \begin{cases} x, \text{if } i = j \\ c_i (d_{ij} \land x), \text{otherwise} \end{cases}$
-  \end{center}
-\end{defin}
-
-\begin{prop} Let $\alpha$ be an ordinal and let $i, j, k, l < \alpha$. The following facts hold in ${\bf CA}_{\alpha}$
-
-  \begin{enumerate}
-    \item $s^{i}_j x \leq c_i x$.
-    \item $s^{i}_j (x \land y) = s^{i}_j x \land s^{i}_j y$, $s^{i}_j (x \lor y) = s^{i}_j x \lor s^{i}_j y$, $- s^{i}_j x = s^{i}_j (- x)$.
-    Moreover, $s^{i}_j$ is completely additive.
-    \item $i \neq k, l$ implies $s^i_j d_{ik} = d_{jk}$ and $s^i_j d_{kl} = d_{kl}$.
-    \item $d_{jk} \land s^{i}_j = d_{jk} \land s^{i}_k$.
-    \item $s^i_j c_i x = c_i x$.
-    \item $k \neq i,j$ implies $s^i_j c_i x = c_i s^i_j x$.
-    \item $c_j s^i_j x = c_i s^j_i x$.
-    \item $i \neq j$ implies $c_i s^i_j x = s^i_j x$.
-    \item $i \neq k$ implies $s^i_j s^i_k = s^i_k x$.
-    \item If either $i \notin \{ k, l\}$ and $k \notin \{ i, j\}$, or $j = l$, then $s^i_j s^k_l x = s^k_l s^i_j x$.
-    \item $s^i_j s^j_i x = s^i_j x$.
-    \item $s^i_k s^j_i x = s^i_k s^j_k x = s^j_k s^i_j x$
-  \end{enumerate}
-\end{prop}
-
 \section{${\bf IG}_{\omega}$ and ultraproducts}
 
-\section{${\bf IG}_{\omega}$ is (not) a variety; is (not) (pseudo-)elementary}
+\section{${\bf IG}_{\omega}$ is (not) (pseudo-)elementary; is (not) a variety}
 

### 2020-10-15 problem-9.4/Text.tex
cylindric background

@@
 \nocite{*}
+
+\section{Intro}
+...
+It is known that the equational theory of ${\bf RCA}_{\omega}$ for $\alpha \leq \omega$ is decidable \cite{nemeti1996fine}.
+...
 
@@
 
-Let $\alpha$ be an ordinal. Let $U^{\alpha}$ be the set of all functions mapping $\alpha$ to a non-empty set $U$. We denote $x(i) = x_i$ for
-$x \in U^{\alpha}$ and $i < \alpha$.
+Let $\alpha$ be an ordinal. Let $\prescript{\alpha}{}U$ be the set of all functions mapping $\alpha$ to a non-empty set $U$. We denote $x(i) = x_i$ for
+$x \in \prescript{\alpha}{}U$ and $i < \alpha$.
 
@@
   \begin{enumerate}
-    \item A subset of $U^{\alpha}$ is an $\alpha$-ry relation on $U$. For $i, j < \alpha$, the $i,j$-diagonal $D_{ij}$ is the set of all
-    elements of $U$ such that $y_i = y_j$. If $i < \alpha$ and $X$ is an $\alpha$-ry relation on $U$, then
+    \item A subset of $\prescript{\alpha}{}U$ is an $\alpha$-ry relation on $U$. For $i, j < \alpha$, the $i,j$-diagonal $D_{ij}$ is the set of all
+    elements of $U$ such that $y_i = y_j$.
+
+    If $i < \alpha$ and $X$ is an $\alpha$-ry relation on $U$, then
     the $i$-th cylindrification $C_i X$ is the set of all elements of $U$ that agree with some element of $X$ on each coordinate except the
     $i$-th one. To be more precise,
-    $C_i X = \{ y \in U^{\alpha} \: | \: \exists x \in X \forall i < \alpha \: (i \neq j \Rightarrow y_j = x_j)\}$.
+    $C_i X = \{ y \in \prescript{\alpha}{}U \: | \: \exists x \in X \forall i < \alpha \: (i \neq j \Rightarrow y_j = x_j)\}$.
     \item A cylindic set algebra of dimension $\alpha$ is an algebra consisting of a set $S$ of $\alpha$-ry relation on some base set $U$
-    with the constants and operations $0 = \emptyset$, $1 = U^{\alpha}$, $\cap$, $-$, the diagonal elements $\{ D_{ij} \}_{i, j < \alpha}$,
-    the cylindrifications $\{ C \}_{i < \alpha}$. A generalised cylindric set algebra of dimension $\alpha$ is a subdirect of
+    with the constants and operations $0 = \emptyset$, $1 = \prescript{\alpha}{}U$, $\cap$, $-$, the diagonal elements $\{ D_{ij} \}_{i, j < \alpha}$,
+    the cylindrifications $\{ C \}_{i < \alpha}$.
+
+    A generalised cylindric set algebra of dimension $\alpha$ is a subdirect of
     cylindric algebras that have dimension $\alpha$
@@
     of dimension $\alpha$. Such is isomorphism is a representation of $C$.
+
     ${\bf RCA}_{\alpha}$ is the class of all representable cylindric algebras that have dimension $\alpha$. In particular, we are interested in the case when $\alpha = \omega$.
   \end{enumerate}
+\end{defin}
+
+It is well-known that ${\bf RCA}_{\alpha}$ is a variety, ${\bf RCA}_{\alpha}$ ($\alpha \leq 2$) is finitely axiomatisable and ${\bf RCA}_{\alpha}$ ($2 < \alpha < \omega$) has no finite axiomatisation, see \cite{Henkin1988-HENCAP-4}.
+
+Let $\mathcal{A} \in {\bf C}_{\omega}$, then $\mathcal{A}$ has a \emph{complete representation}, if this representation preserves all existing suprema.
+
+Let us discuss the connection between representability and canonical extensions.
+
+\begin{defin}
+  neat embeddable
 \end{defin}

### 2020-11-13 problem-9.4/Text.tex
neat embedding stuff (wip)

@@
 \begin{enumerate}
-\item Suppose $\mathcal{C} \in {\bf RCA}_{\omega}$, whether $\mathcal{C}$ has a complete, $\omega$-dimensional representation?
+\item Suppose $\mathcal{C} \in {\bf RCA}_{\omega}$, whether $\mathcal{C}^{+}$ has a complete, $\omega$-dimensional representation?
 \item Is the class ${\bf IG}_{\omega}$ (the isomorphism-closure of the $\omega$-dimensional cylindric relativised set algebras in which the unit is closed under substitutions and permutations) a variety, or even
@@
 
-It is well-known that ${\bf RCA}_{\alpha}$ is a variety, ${\bf RCA}_{\alpha}$ ($\alpha \leq 2$) is finitely axiomatisable and ${\bf RCA}_{\alpha}$ ($2 < \alpha < \omega$) has no finite axiomatisation, see \cite{Henkin1988-HENCAP-4}.
+It is well known that ${\bf RCA}_{\alpha}$ is a variety, ${\bf RCA}_{\alpha}$ ($\alpha \leq 2$) is finitely axiomatisable and ${\bf RCA}_{\alpha}$ ($2 < \alpha < \omega$) has no finite axiomatisation, see \cite{Henkin1988-HENCAP-4}.
 
@@
 
+Let us concretise the definition of a canonical extension for $\alpha$-dimensional cylindric algebras:
+
+\begin{defin}
+  Let $\mathcal{C} = \langle C, +, -, 0, 1, \{ d_{ij} \}_{i, j < \omega}, \{c_i\}_{i < \omega} \rangle$ be an $\omega$-dimensional cylindric algebra.
+Let ${\bf Uf}(\mathcal{C})$ be the set of all ultrafilters of $\mathfrak{B}\mathcal{C}$, a Boolean reduct of $\mathcal{C}$.
+
+  Let us define ${\bf C}_{i} : {\bf Uf}(\mathcal{C}) \to {\bf Uf}(\mathcal{C})$ for each $i, j < \omega$ as
+  \begin{enumerate}
+\item ${\bf C}_{i} \mathcal{X} = \{ \mathcal{F} \in {\bf Uf}(\mathcal{C}) \: | \: \exists \mathcal{F}' \in {\bf Uf}(\mathcal{C}) \: (a \in \mathcal{F} \Rightarrow c_i a \in \mathcal{F}' R)\}$,
+\item $D_{ij} = \{ \mathcal{F} \in {\bf Uf}(\mathcal{C}) \: | \: d_{ij} \in \mathcal{F} \}$.
+  \end{enumerate}
+  The structure $\mathcal{C}^{+} = \langle  {\bf Uf}(\mathcal{C}), \cup, -, \emptyset, C, {\bf C}_{i < \omega}, \{D_{ij}\}_{i, j < \omega} \rangle$ is called the canonical extension of $\mathcal{C}$.
+\end{defin}
+
 Let us discuss the connection between representability and canonical extensions.
 
+The following definitions and facts are due to Henkin, Monk, and Tarski \cite{henkin1971cylindric}.
+
+Let $\mathcal{C} = \langle C, +, -, 0, 1, \{d_{ij}\}_{i, j < \beta}, \{ c \}_{c < \beta} \rangle$ be a $\beta$-dimensional cylindic algebra and $\alpha \leq \beta$ an ordinal.
+The \emph{$\alpha$-th reduct} of $\mathcal{C}$, denoted as $\mathfrak{R}\mathfrak{d}_{\alpha}\mathcal{C}$, is an algebra having the form
+\begin{center}
+$\mathfrak{R}\mathfrak{d}_{\alpha}\mathcal{C} = \langle C, +, -, 0, 1, \{d_{ij}\}_{i, j < \alpha}, \{ c \}_{c < \alpha} \rangle$
+\end{center}
+
+$\mathcal{B}$ is a subreduct of $\mathcal{C}$, denoted as $\mathcal{B} \subseteq^{r} \mathcal{C}$, if $\mathcal{B} \subseteq \mathfrak{R}\mathfrak{d}_{\gamma}\mathcal{C}$ for some $\gamma \leq \beta$.
+
 \begin{defin}
-  neat embeddable
+  Let $\mathcal{C}$ be a $\beta$-dimensional cylindic algebra and $\alpha$ an ordinal such that $\alpha \leq \beta$.
+  The neat $\alpha$-reduct of $\mathcal{C}$, denoted as $\mathfrak{N}\mathfrak{r}_\alpha \mathcal{C}$,
+  is the subalgbera $\mathcal{A}$ of $\mathfrak{R}\mathfrak{d}_{\alpha}\mathcal{C}$ with $\mathcal{A} = \operatorname{Cl}_{\kappa}$ where $\alpha + \kappa = \beta$.
+
+  Let $\mathbb{K}$ be a class of $\beta$-dimensional cylindic algebras, then we put
+  \begin{center}
+    ${\bf Nr}_{\alpha}\mathbb{K} = \{ \mathfrak{N}\mathfrak{r}_\alpha \mathcal{C} \: | \: \mathcal{C} \in \mathbb{K} \}$
+  \end{center}
+
+  An algebra $\mathcal{B}$ is a neat subreduct of $\mathcal{C}$, or $\mathcal{B}$ is neatly embeddable to $\mathcal{C}$ if there exists an ordinal $\gamma \leq \alpha$ such that
+  $\mathcal{C} \subseteq \mathfrak{R}\mathfrak{d}_{\gamma}\mathcal{B}$.
 \end{defin}

### 2020-11-16 problem-9.4/Text.tex
update cylindric stuff

@@
 
-\nocite{*}
-
 \section{Intro}
@@
 \begin{enumerate}
-\item Suppose $\mathcal{C} \in {\bf RCA}_{\omega}$, whether $\mathcal{C}^{+}$ has a complete, $\omega$-dimensional representation?
+\item Suppose $\mathcal{C} \in {\bf RCA}_{\omega}$, whether $\mathcal{C}^{+}$ has a complete, $\omega$-dimensional representation? \cite{hirsch2002relation}
 \item Is the class ${\bf IG}_{\omega}$ (the isomorphism-closure of the $\omega$-dimensional cylindric relativised set algebras in which the unit is closed under substitutions and permutations) a variety, or even
-a pseudo-elementary class? Is it closed under ultraproducts?
+a pseudo-elementary class? Is it closed under ultraproducts? \cite{hirsch2002relation}
 \end{enumerate}
@@
 \begin{defin}
-  Let $\mathcal{C} = \langle C, +, -, 0, 1, \{ d_{ij} \}_{i, j < \omega}, \{c_i\}_{i < \omega} \rangle$ be an $\omega$-dimensional cylindric algebra.
-Let ${\bf Uf}(\mathcal{C})$ be the set of all ultrafilters of $\mathfrak{B}\mathcal{C}$, a Boolean reduct of $\mathcal{C}$.
+  Let $\mathcal{C} = \langle C, +, -, 0, 1, \{ d_{ij} \}_{i, j < \alpha}, \{c_i\}_{i < \omega} \rangle$ $\mathcal{A}$ be a BAO of type ${\bf CA}_{\alpha}$ Let ${\bf Uf}(\mathcal{C})$ be the set of all ultrafilters of $\mathfrak{B}\mathcal{C}$, the Boolean part of $\mathcal{C}$.
 
-  Let us define ${\bf C}_{i} : {\bf Uf}(\mathcal{C}) \to {\bf Uf}(\mathcal{C})$ for each $i, j < \omega$ as
+  Let us define ${\bf C}_{i} : {\bf Uf}(\mathcal{C}) \to {\bf Uf}(\mathcal{C})$ for each $i, j < \alpha$ as
   \begin{enumerate}
@@
   \end{enumerate}
-  The structure $\mathcal{C}^{+} = \langle  {\bf Uf}(\mathcal{C}), \cup, -, \emptyset, C, {\bf C}_{i < \omega}, \{D_{ij}\}_{i, j < \omega} \rangle$ is called the canonical extension of $\mathcal{C}$.
+  The structure $\mathcal{C}^{+} = \langle  {\bf Uf}(\mathcal{C}), \cup, -, \emptyset, C, {\bf C}_{i < \alpha}, \{D_{ij}\}_{i, j < \alpha} \rangle$ is called the canonical extension of $\mathcal{C}$.
 \end{defin}
@@
 The following definitions and facts are due to Henkin, Monk, and Tarski \cite{henkin1971cylindric}.
+
+Let $\mathcal{A} \in {\bf CA}_{\alpha}$ and $x \in \mathcal{A}$. Recall that \emph{the dimension of $x$}
+is the set of all ordinals $\gamma < \alpha$ such that $c_\gamma x \neq x$. More formally,
+\begin{center}
+  $\Delta x = \{ \gamma \: | \: \gamma < \alpha \: \& \: c_\gamma x \neq x\}$
+\end{center}
+
+Let us discuss some metamathematical intuitions standing behind the notion of a dimension.
+Let $\Theta$ be a first-order theory and $\mathcal{C}/\equiv_{\Theta}$ its Lindenbaum-Tarski algebra.
+Let $\varphi$ be a formula in the signature of $\Theta$.
+Then $\Delta(\varphi/\Theta)$ consists of all $\kappa < \alpha$ such that $\exists x_{\kappa} \varphi \leftrightarrow \varphi$ is not valid in $\Theta$.
+That is, $\Delta(\varphi/\Theta)$ contains ordinals $\kappa$ for which $x_{\kappa}$ is free in $\varphi$.
+Moreover, $\Delta(\varphi/\Theta)$ consists only of those ordinals for which $x_{\kappa}$ is free in every $\psi \in \varphi/\Theta$.
+
+In particular, an element $x$ is called \emph{zero-dimensional} if $\Delta x = 0$.
+Zero-dimensional elements reflect equivalence classes of sentences in the Lindenbaum-Tarski algebra of a given first-order theory.
+Thus, the set of zero-dimensional elements form a Boolean algebras of sentences associated with $\Theta$.
+
+\begin{defin}
+  Let $\mathcal{A}$ be an $\alpha$-dimensional cylindric algebra.
+Let $\alpha$ be an ordinal and $\Gamma$ a subset $\alpha$, then an element $x \in \mathcal{A}$ is $\Gamma$-closed if $\Delta x \cdot \Gamma = \emptyset$.
+Alternatively, $x$ is a $\Gamma$-cylinder.
+
+ $\operatorname{Cl}_{\Gamma}\mathcal{A}$ is the set of all $\Gamma$-closed elements.
+\end{defin}
+
+Metamathematically, $\Gamma$-closed elements reflect universal closures (is it correct?).
 
@@
   The neat $\alpha$-reduct of $\mathcal{C}$, denoted as $\mathfrak{N}\mathfrak{r}_\alpha \mathcal{C}$,
-  is the subalgbera $\mathcal{A}$ of $\mathfrak{R}\mathfrak{d}_{\alpha}\mathcal{C}$ with $\mathcal{A} = \operatorname{Cl}_{\kappa}$ where $\alpha + \kappa = \beta$.
+  is the subalgbera $\mathcal{A}$ of $\mathfrak{R}\mathfrak{d}_{\alpha}\mathcal{C}$ with $\mathcal{A} = \operatorname{Cl}_{\kappa} \mathcal{C}$ where $\alpha + \kappa = \beta$.
 
@@
 
+\begin{lemma}[Henkin, Monk, Tarski]\label{Neat}
+  $ $
+
+  Let $\mathcal{A}$ be a BAO of type ${\bf CA}_{\alpha}$ and $\mathcal{B}$ be a $\beta$-dimensional cylindric algebra such that $\beta \leq \alpha$ and $\mathcal{A}$ neatly embeds to $\mathcal{B}$ by a complete embedding.
+Then $\mathcal{A}^{+}$ neatly embeds to $\mathcal{B}^{+}$ by a complete embedding.
+\end{lemma}
+
+\begin{proof}
+\end{proof}
+
+\begin{theorem}[This assumption is by Ian Hodkinson]\label{Neat2}
+  $ $
+
+  Let $\mathcal{A}$ be a BAO of type ${\bf CA}_{\omega}$ such that $\mathcal{A}$ neatly embeds into ${\bf CA}_{\omega + \omega}$ by a complete embedding.
+Then $\mathcal{A}$ is completely representable as ${\bf CA}_{\omega}$.
+\end{theorem}
+
+\begin{proof}
+  Hmmmmm, I believe so.
+\end{proof}
+
+Lemma~\ref{Neat} and Theorem~\ref{Neat2} imply the following theorem.
+
+\begin{theorem}
+  Let $\mathcal{C} \in {\bf RCA}_{\omega}$, then $\mathcal{C}^{+} \in {\bf RCA}_{\omega}$. That is, ${\bf RCA}_{\omega}$ is closed under canonical extensions.
+\end{theorem}
+
+\begin{proof}
+  :monkahmm:
+\end{proof}
+
+\begin{comment}
 \subsection{Substitution in cylindric algebras}
@@
 \section{${\bf IG}_{\omega}$ is (not) (pseudo-)elementary; is (not) a variety}
+\end{comment}
 

### 2020-11-20 filtration-stuff/Text.tex
R^E is well defined

@@
 
-First of all, let us ensure that a filtration of an Euclidean frame is not necessary finite.
+First of all, let us ensure that a minimal filtration of an Euclidean frame is not necessary Euclidean.
 Let $[x] \sim_{\Gamma} [y]$ and $[x] \sim_{\Gamma} [z]$. Then for some $x' \in [x]$ $y' \in [y]$, one has
@@
   \item $(4) \Rightarrow (1)$
+\end{enumerate}
+\end{proof}
+
+\begin{lemma}
+ Let $\mathcal{F} = \langle W, R \rangle$ be a Kripke frame. Let us define $R^{E} = \bigcup \limits_{i < \omega} R_i$ where:
+
+ \begin{enumerate}
+   \item $R_0 = R$
+   \item $R_{n + 1} = R_n \cup ({R_n}^{-1} \circ R_n)$
+ \end{enumerate}
+ Then $R^{E}$ is Euclidean.
+\end{lemma}
+
+\begin{proof}
+  Let $(x, y), (x, z) \in R^{E}$, one needs to show that $(y, z) \in R^{E}$.
+  Clearly that $(x, y) \in R_i$ and $(x, y) \in R_j$ for some $i, j < \omega$. Thus, we need $(y, z) \in R_m$ for some $m$ depending on $i$ and $j$.
+
+  Let us consider the following cases:
+  \begin{enumerate}
+    \item $i = 0$ and $j = 0$
+
+    Suppose $(x, y), (x, z) \in R_0 = R$, then $(y, z) \in R^{-1} \circ R$. Thus, $(y, z) \in R_1$
+    \item $i = 0$ and $j = k + 1$
+
+    Suppose $(x, y) \in R$ and $(x, z) \in R_{k + 1} = R_{k} \cup ({R_k}^{-1} \circ R_k)$.
+    Clearly that $(x, y) \in R_{k + 1}$ as well. It is obviously that $(y, z) \in R_{k + 2}$ since
+    $(y, x) \in R_{k + 1}^{-1}$ and $(x, z) \in R_{k + 1}$.
+
+    \item The case with $i = k + 1$ and $j = 0$ is similar to the previous one.
+
+    \item Suppose $i = m + 1$ and $i = k + 1$. That is, $(x, y) \in R_{m + 1} = R_m \cup ({R_m}^{-1} \circ R_m)$ and $(x, z) \in R_{k + 1} = R_k \cup ({R_k}^{-1} \circ R_k)$.
+    Consider the following four subcases:
+    \begin{enumerate}
+      \item Suppose $(x, y) \in R_m$ and $(x, z) \in R_k$ and $m \leq k$ without loss of generality.
+      $m \leq k$ implies $R_m \subseteq R_k$ and $(x, y) \in R_k$ in particular. Thus, $(y, z) \in {R_k}^{-1} \circ R_k$, so $(y, z) \in R_{k + 1}$.
+      \item The rest of the cases are similar to the first one.
+    \end{enumerate}
     \end{enumerate}

### 2020-11-20 problem-9.4/Text.tex
R^E is well defined

@@
 
+One may define neat reducts alternatively as follows.
+Let $\mathcal{C}$ be a $\beta$-dimensional cylindic algebra and $\alpha$ an ordinal such that $\alpha \leq \beta$.
+The neat $\alpha$-reduct of $\mathcal{C}$ is the $\alpha$-dimensional cylindric algebra having the form
+
+\begin{center}
+  $\mathfrak{N}\mathfrak{r}_\alpha \mathcal{C} = \langle \{ a \in \mathcal{C} \: | \: \forall j (\alpha \leq j \: \& \: j < \beta \Rightarrow c_j a = a) \}, +, -, 0, 1, \{d_{ij}\}_{i,j < \alpha}, \{c_{\gamma}\}_{\gamma} \rangle$
+\end{center}
+
+We will adapting the following proof of the fact that ${\bf RRA}$ is canonical \footnote{This idea is by Ian Hodkinson} to our case.
+This proof is due to Monk, but that was describe in McKenzie's thesis \cite{mckenzie1968representation}.
+
+\begin{enumerate}
+  \item A relation algebra $\mathcal{A}$ is representable iff $\mathcal{A}$ neatly embeds to some omege-dimensional cylinric algebra,
+  \item If $\mathcal{A}$ neatly embeds in $\mathcal{A}$ then $\mathcal{A}^{+}$ neatly embeds in $\mathcal{B}^{+}$,
+  \item ${\bf CA}_{\alpha}$ is closed under canonical extensions,
+  \item Voil\'{a}.
+\end{enumerate}
+
+We descibe the following sketch of that proof published by Maddux in \cite{maddux2006relation}.
+
+\begin{theorem}
+  ${\bf RRA}$ is closed under canonical extensions.
+\end{theorem}
+
+\begin{proof}
+\end{proof}
+
 \begin{lemma}[Henkin, Monk, Tarski]\label{Neat}

### 2020-11-20 filtration-stuff/Text.tex
filtration for K4

@@
 
-\nocite{*}
-
 \section{Preliminaries}
@@
 A frame is called Euclidean if for each $x, y, z$, $x R y$ and $x R z$ implies $y R z$.
-
-\begin{prop}
-  Let $\mathcal{F} = \langle W, R \rangle$ be an Euclidean frame.
-
-  \begin{enumerate}
-    \item For each $x, y, z \in W$, $x R y$ and $x R z$ implies either $y R z$ or $z R y$.
-    \item For each $x \in W$, $R^{*}(x) = \{ x\} \cup R(R(x))$.
-    \item $R^{-1} \circ R \subseteq R$.
-  \end{enumerate}
-\end{prop}
 
@@
 
-\subsection{Clusters}
+\section{Filtration for {\bf K4}}
 
-Let $\mathcal{F} = \langle W, R \rangle$ be a transitive frame.
-Let us put $x R^{\bullet} y \Leftrightarrow x R y \: \& \: \neg (x R y)$. A point $x$ is proper if $x R x$. Let us define the following
-equivalence relation:
-
-\begin{center}
-  $x \equiv y \Leftrightarrow x R y \: \& \: y R x \lor x = y$.
-\end{center}
-A cluster is an element of the quotient set $W / \equiv$. Given $x \in W$, $C_x$ is a cluster containing $x$. Thus $C_x =
-\{ x\} \cup \{ y \: | x R y x \}$. The original relation lifts to the antisymmetric transitive relation on $W / \equiv$ defined as
-$C_x R C_y$ iff $x R y$. A cluster $C$ is called maximal if $C R C'$ implies $C = C'$. A point is $R$-maximal if $C_x$ is a maximal cluster,
-that is, $R^{\bullet}(x) = \emptyset$.
-A degenerated cluster is a singleton $\{ x \}$ with $\neg (x R x)$.
-A cluster is called simple if it has the form $\{ x \}$ with $x R x$.
-If $\langle W', R' \rangle$ is an inner substructure of $\langle W, R \rangle$, then every $R'$-cluster is an $R$-cluster and every $R$-cluster that intersects $W'$ is a subset of $W'$ and is an $R'$-cluster itself.
-Given a Kripke model $\mathcal{M}$, a set of formulas $\Gamma$ is satisfied by a cluster $C$ if every member of $\Gamma$ is true at some point of $C$.
-
-If clusters coincide then their poitns have the same theory in the original model:
+\begin{prop}
+  Let $R$ be a binary relation on $W \neq \emptyset$. Define $R^{+} = \bigcup \limits_{i < \omega} R_i$
+  \begin{enumerate}
+    \item $R_0 = R$
+    \item $R_{n + 1} = R_n \circ R$
+  \end{enumerate}
+  Then $R^{+}$ is transitive
+\end{prop}
 
 \begin{lemma}
-  $C_x = C_y$ implies $\mathcal{M}, x \models \Box \varphi \Leftrightarrow \mathcal{M}, y \models \Box \varphi$
+  Let $\mathcal{M} = \langle W, R, \vartheta \rangle$ be a transitive model and $\overline{\mathcal{M}} = \langle \overline{W}, \overline{R}, \overline{\vartheta} \rangle$ its minimal filtration through a finite $\operatorname{Sub}$-closed set of formulas $\Theta$.
+
+  Then $\overline{\mathcal{M}}^{+} = \langle \overline{W}, (\overline{R})^{+}, \overline{\vartheta} \rangle$ is a $\Theta$-filtration of $\mathcal{M}$.
 \end{lemma}
 
-Let us describe the bulldozing technique allowing one to eliminate nondegenerated clusters \cite{bull1984basic}.
-Let $\mathcal{L}$ be a transitive logic and $\mathcal{F}$ its frame.
-We construct first a frame $\mathcal{F}^0 = \langle W^{0}, R^{0} \rangle$ replacing every nondegenerated frame $C$ of $W$ by $C^0 = \{ \langle x, i \rangle \: | \: x \in C, i < \omega \}$.
-We also replace each degenerated cluster $C$ by $\{ \langle x,0 \rangle\}$.
-Elements of these subsets form $W^0$. The relation $R^{0}$ is defined as
+\begin{proof}
+  $(\overline{R})^{+}$ obviously contains $R$. By the previous proposition, $(\overline{R})^{+}$ is transitive.
+  Let us show that $(\overline{R})^{+} \subseteq R^{max}_{\Theta}$.
+
+  Let $\hat{x}, \hat{y} \in \overline{W}$ with $\hat{x} (\overline{R})^{+} \hat{y}$ and
+  $\Box \varphi \in \Theta$ with $\mathcal{M}, x \models \Box \varphi$. Let us show that $\mathcal{M}, y \models \varphi$.
+
+  If $\hat{x} (\overline{R})^{+} \hat{y}$, then there exist equivalence classes $\hat{x}_1, \dots, \hat{x}_n$ such that
   \begin{center}
-  $\langle x, i \rangle R^{0} \langle y, j \rangle \Leftrightarrow x R^{\bullet} y \text{ or } (x \equiv y \: \& \: i < j) \text{ or } i = j \: \& \: x <_C y$
-\end{center}
-where $<_C$ is an arbitrary strict ordering on the proper cluster $C$ containing $x$ and $y$.
-
-Each nondegenrated cluster $C$ is replaced by an infinite set $C_0$ such that $\langle C_0, R_0 \rangle$ is a strict linear order.
-Moreover, $\langle y, j \rangle$, a copy of $y$, occurs after $\langle y, j \rangle$, a copy of $x$.
-
-Bulldozing might be extended to models $\mathcal{M} = \langle W, R, \vartheta \rangle$ defining $\vartheta^0$ as follows
-\begin{center}
-  $\vartheta^{0}(p_i) = \{ \langle x, i \rangle \: | \: x \in \vartheta(p_i), i < \omega \}$.
+    $\hat{x} \overline{R} \hat{x}_1 \overline{R} \dots \overline{R} \hat{x}_n \overline{R} \hat{y}$
   \end{center}
 
-One may show inductively the following fact.
-\begin{lemma}
-  $\mathcal{M}, x \models \varphi \Leftrightarrow \mathcal{M}^{0}, \langle x, i \rangle \models \varphi$
-\end{lemma}
+  $\mathcal{M}, x \models \Box \varphi$ implies $\mathcal{M}, x \models \Box \varphi \to \Box \Box \varphi$. Thus, $\overline{M}, \hat{x} \models \Box \Box \varphi$.
 
-Let us concretise the case of transitive Euclidean frames. First of all, we consider clusters in ${\bf K}45$ frames.
+  $\hat{x} \overline{R} \hat{x}_1$, so there are $x_1 \in \hat{x}$ and $x_2 \in \hat{x}_1$ with $x_1 R x_2$. In particular, $\mathcal{M}, x_2 \models \Box \varphi$,
+  so $\overline{\mathcal{M}}, \hat{x}_2 \models \Box \varphi$, and et cetera.
 
-\subsection{}
+  For each $i \in \{1, \dots, n\}$ we have $\mathcal{M}, x_i \models \Box \varphi$ which is shown inductively:
 
-\section{Transitive closure stuff}
+  If $\mathcal{M}, x_i \models \Box \varphi$ for $x_i \in \hat{x}_i$, so $\mathcal{M}, x_i \models \Box \Box \varphi$, but there exist ${x_i}' \in \hat{x}_i$ and $x_{i + 1} \in \hat{x}_{i + 1}$, so
+  $\mathcal{M}, x_{i + 1} \models \Box \varphi$.
 
-\section{PDLisation of Euclidean logics}
-
-\section{Transitive closure and fusion}
+  Finally, we have $\mathcal{M}, x_n \models \Box \varphi$ for $x_n \in \hat{x}_n$, but $\hat{x}_n \overline{R} \hat{y}$, so $\mathcal{M}, y' \models \varphi$ for each $y' \in \hat{y}$.
+Thus, $\varphi$ is true at $y$ as well.
+\end{proof}
 

### 2020-11-21 filtration-stuff/Text.tex
a different proof that the transitive closure of a K4-model is a filtration

@@
 
-  $\mathcal{M}, x \models \Box \varphi$ implies $\mathcal{M}, x \models \Box \varphi \to \Box \Box \varphi$. Thus, $\overline{M}, \hat{x} \models \Box \Box \varphi$.
+  $\mathcal{M}, x \models \Box \varphi$ implies $\mathcal{M}, x \models \Box \Box \varphi$. Thus, $\overline{M}, \hat{x} \models \Box \Box \varphi$.
 
@@
 
+\begin{proof}
+  Let $\hat{x}, \hat{y} \in \overline{W}$ with $\hat{x} (\overline{R})^{+} \hat{y}$ and
+  $\Box \varphi \in \Theta$ with $\mathcal{M}, x \models \Box \varphi$. Let us show that $\mathcal{M}, y \models \varphi$.
+
+  If $\hat{x} (\overline{R})^{+} \hat{y}$, then there exist equivalence classes $\hat{x}_1, \dots, \hat{x}_n$ such that
+  \begin{center}
+    $\hat{x} \overline{R} \hat{x}_1 \overline{R} \dots \overline{R} \hat{x}_n \overline{R} \hat{y}$
+  \end{center}
+
+  Let us show that $\mathcal{M}, \hat{x}_i \models \Box \varphi$ inductively:
+  \begin{enumerate}
+    \item $n = 1$
+    We have the following sequence:
+    \begin{center}
+      $\hat{x} \overline{R} \hat{x}_1 \overline{R} \hat{y}$
+    \end{center}
+    $\hat{x} \overline{R} \hat{x}_1$, so there are $x' \in \hat{x}$ and $x_1' \in \hat{x}_1$ such that $x' R x_1'$. $\Box \varphi$ is true at $x'$, so is $\Box \Box \varphi$. Then
+    $\mathcal{M}, x_1' \models \Box \varphi$ since $x_1' \in R(x')$.
+    So $\overline{\mathcal{M}}, \hat{x}_1 \models \Box \varphi$.
+
+    \item $n = i + 1$
+    The case is the following:
+    \begin{center}
+      $\hat{x} \overline{R} \hat{x}_1 \overline{R} \dots \overline{R} \hat{x}_i \overline{R} \hat{x}_{i + 1} \overline{R} \hat{y}$
+    \end{center}
+
+    By IH, $\Box \varphi$ is true at $\hat{x}_i$, so is $\Box \Box \varphi$. Hence, we have $\overline{\mathcal{M}}, \hat{x}_{i + 1} \models \Box \varphi$ since
+    $\hat{x}_i \overline{R} \hat{x}_{i + 1}$.
+  \end{enumerate}
+
+  That is, for each $0 < n < \omega$, if we have a sequence of equivalence classes with $\hat{x} \overline{R} \hat{x}_1 \overline{R} \dots \overline{R} \hat{x}_n \overline{R} \hat{y}$ where $\overline{\mathcal{M}}, \hat{x} \models \Box \varphi$, then
+  $\overline{\mathcal{M}}, \hat{x}_n \models \Box \varphi$.
+
+  If $\hat{x}_n \overline{R} \hat{y}$,
+  then there are $x_n' \in \hat{x}_n$ and  $y' \in \hat{y}$ with $x_n' R y'$.
+$\mathcal{M}, x_n' \models \Box \varphi$ implies $\mathcal{M}, y' \models \varphi$, but $y'$ and $y$ are $\Gamma$-equivalent and $\varphi \in \Gamma$, so $\mathcal{M}, y \models \varphi$.
+\end{proof}
+
 \bibliographystyle{plain}

### 2020-11-29 problem-9.4/Text.tex
the proof that RRA is canonical using neat reducts of omega-cylindric algebras and fixes in filtration-stuff/Text.tex

@@
 \date{}
-\title{Model-theoretic aspects of relativised cylindric set algebras}
+\title{Representable cylindric algebras of dimension $\omega$: the aspects of canonicity}
 
@@
 \section{Intro}
-...
-It is known that the equational theory of ${\bf RCA}_{\omega}$ for $\alpha \leq \omega$ is decidable \cite{nemeti1996fine}.
-...
 
-\section{The problems themselves}
+\section{The problem itself}
 
-\begin{enumerate}
-\item Suppose $\mathcal{C} \in {\bf RCA}_{\omega}$, whether $\mathcal{C}^{+}$ has a complete, $\omega$-dimensional representation? \cite{hirsch2002relation}
-\item Is the class ${\bf IG}_{\omega}$ (the isomorphism-closure of the $\omega$-dimensional cylindric relativised set algebras in which the unit is closed under substitutions and permutations) a variety, or even
-a pseudo-elementary class? Is it closed under ultraproducts? \cite{hirsch2002relation}
-\end{enumerate}
+Suppose $\mathcal{C} \in {\bf RCA}_{\omega}$, whether $\mathcal{C}^{+}$ has a complete, $\omega$-dimensional representation? \cite{hirsch2002relation}
 
@@
 
-Let $\alpha$ be an ordinal. Let $\prescript{\alpha}{}U$ be the set of all functions mapping $\alpha$ to a non-empty set $U$. We denote $x(i) = x_i$ for
+\begin{defin}
+  $ $
+
+    A relation algebra is an algebra $\mathcal{R} = \langle R, 0, 1, +, -, ;, {}^{\smile}, {\bf 1 }' \rangle$ such that $\langle R, 0, 1, +, - \rangle$ is a Boolean algebra and the following
+    equations hold, for each $a, b, c \in R$:
+    \begin{enumerate}
+      \item $a ; (b ; c) = (a ; b) ; c$
+      \item $(a + b) ; c = (a ; c) + (b ; c)$
+      \item $a ; {\bf 1}' = a$
+      \item $a^{\smile \smile} = a$
+      \item $(a + b)^{\smile} = a^{\smile} + b^{\smile}$
+      \item $(a ; b)^{\smile} = b^{\smile} ; a^{\smile}$
+      \item $a^{\smile} ; (- (a ; b)) \leq - b$
+    \end{enumerate}
+    where $a \leq b$ iff $a + b = b$. ${\bf RA}$ denotes the class of all relation
+    algebras.
+\end{defin}
+
+Cylindric algebras provide a generalisation of relation algebras for relations of an arbitrary arity.
+Let $\alpha$ be an ordinal.
+Let $\prescript{\alpha}{}U$ be the set of all functions mapping $\alpha$ to a non-empty set $U$.
+We denote $x(i) = x_i$ for
 $x \in \prescript{\alpha}{}U$ and $i < \alpha$.
@@
 
+\begin{defin}
+  Given a cylindric algebra of dimension $\alpha$ $C$, let $x$ be a term of its signature, the substitution operator $s^{i}_{j}$ have the following definition:
+  \begin{center}
+  $s^{i}_{j} x = \begin{cases} x, \text{if } i = j \\ c_i (d_{ij} \land x), \text{otherwise} \end{cases}$
+  \end{center}
+\end{defin}
+
 It is well known that ${\bf RCA}_{\alpha}$ is a variety, ${\bf RCA}_{\alpha}$ ($\alpha \leq 2$) is finitely axiomatisable and ${\bf RCA}_{\alpha}$ ($2 < \alpha < \omega$) has no finite axiomatisation, see \cite{Henkin1988-HENCAP-4}.
@@
 \begin{defin}
-  Let $\mathcal{C} = \langle C, +, -, 0, 1, \{ d_{ij} \}_{i, j < \alpha}, \{c_i\}_{i < \omega} \rangle$ $\mathcal{A}$ be a BAO of type ${\bf CA}_{\alpha}$ Let ${\bf Uf}(\mathcal{C})$ be the set of all ultrafilters of $\mathfrak{B}\mathcal{C}$, the Boolean part of $\mathcal{C}$.
+  Let $\mathcal{C} = \langle C, +, -, 0, 1, \{ d_{ij} \}_{i, j < \alpha}, \{c_i\}_{i < \alpha} \rangle$ $\mathcal{A}$ be a BAO of type ${\bf CA}_{\alpha}$ Let ${\bf Uf}(\mathcal{C})$ be the set of all ultrafilters of $\mathfrak{B}\mathcal{C}$, the Boolean part of $\mathcal{C}$.
 
@@
 
-We descibe the following sketch of that proof published by Maddux in \cite{maddux2006relation}.
+\begin{defin}
+  Let $\mathcal{C} \in {\bf CA}_{\alpha}$, where $\alpha \geq 3$. The relation algebra reduct of $\mathcal{C}$, written as $\mathfrak{Ra}(\mathcal{C})$, is the algebra having the form
+
+  \begin{center}
+    $\langle \operatorname{dom}(\mathfrak{Nr}_2(\mathcal{C})), 0, 1, +, -, {\bf 1}', \smile, ; \rangle$
+  \end{center}
+  where:
+
+  \begin{enumerate}
+    \item $+$, $-$, $0$, and $1$ are defined as usual in $\mathcal{C}$,
+    \item ${\bf 1}' = d_{01} \in \mathfrak{Nr}_2(\mathcal{C})$,
+    \item $r^{\smile} = s^{2}_{0} s^{0}_{1} s^{1}_{2} r$ for $r \in \mathfrak{Nr}_2(\mathcal{C})$,
+    \item Let $r, s \in \mathfrak{Nr}_2(\mathcal{C})$, then $r ; s = c_2 (s^{1}_{2}r \cdot s^{0}_{2}s)$
+  \end{enumerate}
+\end{defin}
+
+Moreover, $\mathfrak{Nr}_{\beta}(\mathcal{C})$ and $\mathfrak{Ra}(\mathcal{C})$ are closed under these operations.
+There is also the following fact by due to Henkin, Monk, and Tarski \cite{Henkin1988-HENCAP-4}:
+\begin{theorem}
+  Let $\mathcal{C} \in {\bf CA}_{\alpha}$ for $\alpha \geq 4$, then $\mathfrak{Ra}(\mathcal{C})$ is a relation algebra.
+\end{theorem}
+
+The following characterisation results are by Henkin, Monk, and Tarski \cite{Henkin1988-HENCAP-4} as well:
+\begin{theorem} \label{char}
+  $ $
+
+  \begin{enumerate}
+    \item ${\bf RA} = {\bf S} \mathfrak{Ra} {\bf CA}_4$,
+    \item ${\bf RRA} = \bigcap \limits_{3 \leq n < \omega} {\bf S} \mathfrak{Ra} {\bf CA}_n =  {\bf S} \mathfrak{Ra} {\bf CA}_{\alpha}$, where $\alpha$ is an infinite ordinal.
+  \end{enumerate}
+\end{theorem}
+
+Let $\mathcal{C} \in \mathcal{CA}_{\alpha}$, then $\mathcal{R} \in {\bf RA}$ neatly embeds to $\mathcal{C}$, if $\mathcal{R}$ is isomorphic to some subalgebra of $\mathfrak{Ra}(\mathcal{C})$.
 
@@
 
-\begin{proof}
+\begin{proof} Let ${\bf R} \in {\bf RRA}$.
+  By the second item of~\ref{char}, every representable relation algebra is isomorphic to some subalgbera of the relation algebra reduct $\mathfrak{Ra}\mathcal{C}$ for some $\mathcal{C} \in {\bf CA}_{\omega}$.
+But neat embeddings repsect canonical extensions, so if ${\bf R} \hookrightarrow_{n} \mathcal{C}$, so is ${\bf R}^{+} \hookrightarrow_{n} \mathcal{C}^{+}$.
+  ${\bf CA}_{\alpha}$ is closed under canonical extensions, so is ${\bf RRA}$.
 \end{proof}
@@
 \subsection{Substitution in cylindric algebras}
-
-\begin{defin}
-  Given a cylindric algebra of dimension $\alpha$ $C$, let $x$ be a term of its signature, the substitution operator $s^{i}_{j}$ have the following definition:
-  \begin{center}
-  $s^{i}_{j} x = \begin{cases} x, \text{if } i = j \\ c_i (d_{ij} \land x), \text{otherwise} \end{cases}$
-  \end{center}
-\end{defin}
 

### 2020-12-02 filtration-stuff/Text.tex
the proof that RRA is canonical using neat reducts of omega-cylindric algebras and fixes in filtration-stuff/Text.tex
strict filtration for k5 (wip)

@@
   \begin{center}
-    $\mathcal{M}, x \models \Diamond \Box \varphi$ iff $\mathcal{M}, y \models \Diamond \Box \varphi$.
+    $\mathcal{M}, x \models \Box \Diamond \varphi$ iff $\mathcal{M}, y \models \Box \Diamond \varphi$.
   \end{center}
@@
   \begin{enumerate}
-    \item Suppose $\mathcal{M}, x \models \Diamond \Box \varphi$.
-One also has $\mathcal{M}, x \models \Diamond \Box \varphi \to \Box \Diamond \Box \varphi$, so $\mathcal{M}, x \models \Box \Diamond \Box \varphi$.
-Thus, $\mathcal{M}, y \models \Diamond \Box \varphi$ since $y \in R(x)$.
-    \item Suppose $\mathcal{M}, y \models \Diamond \Box \varphi$, then $\mathcal{M}, y \models \Box \varphi$, so $\mathcal{M}, x \models \Diamond \Box \varphi$.
+    \item Suppose $\mathcal{M}, x \models \Box \Diamond \varphi$. Then $\mathcal{M}, y \models \Diamond \varphi$ and $\mathcal{M}, y \models \Box \Diamond \varphi$
+    \item Suppose $\mathcal{M}, y \models \Box \Diamond \varphi$, then $\mathcal{M}, x \models \Diamond \Box \Diamond \varphi$, so $\mathcal{M}, x \models \Box \Diamond \varphi$.
   \end{enumerate}
@@
   Let $(x, y), (x, z) \in R^{E}$, one needs to show that $(y, z) \in R^{E}$.
-  Clearly that $(x, y) \in R_i$ and $(x, y) \in R_j$ for some $i, j < \omega$. Thus, we need $(y, z) \in R_m$ for some $m$ depending on $i$ and $j$.
+  Clearly that $(x, y) \in R_i$ and $(x, z) \in R_j$ for some $i, j < \omega$. Thus, we need $(y, z) \in R_m$ for some $m$ depending on $i$ and $j$.
 
@@
 
-    \item Suppose $i = m + 1$ and $i = k + 1$. That is, $(x, y) \in R_{m + 1} = R_m \cup ({R_m}^{-1} \circ R_m)$ and $(x, z) \in R_{k + 1} = R_k \cup ({R_k}^{-1} \circ R_k)$.
+    \item Suppose $i = m + 1$ and $j = k + 1$. That is, $(x, y) \in R_{m + 1} = R_m \cup ({R_m}^{-1} \circ R_m)$ and $(x, z) \in R_{k + 1} = R_k \cup ({R_k}^{-1} \circ R_k)$.
     Consider the following four subcases:
@@
 
-\begin{lemma}
-  Let $\mathcal{M} = \langle W, R, \vartheta \rangle$ be an Euclidean model, $\Gamma$ a set of $\operatorname{Sub}$-closed formulas, and $\sim$ an equivalence relation having a finite index that respects $\Gamma$, then $\widehat{R} = (R^{min}_{\Phi})^{E} \subseteq R^{max}_{\Gamma}$, where $\Phi = \Gamma \cup \{ \Diamond \Box \varphi \: | \: \Box \varphi \in \Gamma \}$.
+\begin{theorem}
+  Let $\mathcal{M} = \langle W, R, \vartheta \rangle$ be an Euclidean model, $\Gamma$ a set of $\operatorname{Sub}$-closed formulas, and $\sim$ an equivalence relation having a finite index that respects $\Gamma$, then $\widehat{R} = (R^{min}_{\Gamma})^{E} \subseteq R^{max}_{\Gamma}$.
 
   Thus, ${\bf K}5$ admits strict filtrations.
-\end{lemma}
+\end{theorem}
 
 \begin{proof}
-  Recall that $(R^{min}_{\Phi})^{E}$ has the form $(R^{min}_{\Phi})^{E} = \bigcup \limits_{n < \omega} (R^{min}_{\Phi})_n$, where
+  We put $R^{min}_{\Gamma} := \overline{R}$ for brevity.
+
+  Recall that $\overline{R}^{E}$ has the form $\bigcup \limits_{n < \omega} (\overline{R})_n$, where
   \begin{enumerate}
-    \item $(R^{min}_{\Phi})_0 = R^{min}_{\Phi}$
-    \item $(R^{min}_{\Phi})_{m + 1} = (R^{min}_{\Phi})_n \cup (((R^{min}_{\Phi})_n)^{-1} \circ (R^{min}_{\Phi})_n)$
+    \item $\overline{R}_0 = \widehat{R}$
+    \item $\overline{R}_{m + 1} = \overline{R}_n \cup ((\overline{R})^{-1} \circ \overline{R}_n)$
   \end{enumerate}
-  One needs to show that for each $n < \omega$ $(R^{min}_{\Phi})_n \subseteq R^{max}_{\Gamma}$. We prove this by induction.
-  Suppose $\mathcal{M}, x \models \Box \varphi$ for $\Box \varphi \in \Phi$ and $[x] (R^{min}_{\Phi})^{E} [y]$. We need $\mathcal{M}, y \models \varphi$.
+  One needs to show that for each $n < \omega$ $(\overline{R})_n \subseteq R^{max}_{\Gamma}$. We prove this by induction.
+  Suppose $\mathcal{M}, y \models \varphi$ for $\Diamond \varphi \in \Phi$ and $\hat{x} \overline{R}^{E} \hat{x}$. We need $\mathcal{M}, x \models \Diamond \varphi$.
   \begin{enumerate}
-    \item $([x], [y]) \in (R^{min}_{\Phi})_0$, then $([x], [y]) \in R^{min}_{\Phi}$.
-Then there exist $x' \in [x]$ and $y' \in [y]$ such that $x' R y'$.
-So $\mathcal{M}, x' \models \Box \varphi$ and, thus,
-    $\mathcal{M}, y' \models \varphi$. Then $\mathcal{M}, y' \models \varphi$ as well since $y' \in [y]$.
-    \item $([x], [y]) \in (R^{min}_{\Phi})_{m + 1}$, then
-    $([x], [y]) \in (R^{min}_{\Phi})_{m} \cup (((R^{min}_{\Phi})_{m})^{-1} \circ R^{min}_{\Phi})_{m})$.
+    \item The case of $n = 0$ is trivial and it follows directly from the definition of a minimal filtration.
+    \item Suppose $n = 1$. $\hat{x} \overline{R}_1 \hat{y}$ means that $(\hat{x}, \hat{y}) \in \overline{R} \cup (\overline{R}^{-1} \circ \overline{R})$.
+    Let us assume that $(\hat{x}, \hat{y}) \in \overline{R}^{-1} \circ \overline{R}$.
+That is, there exists $\hat{z} \in W / \sim_{\Gamma}$ such that $(\hat{z}, \hat{x}), (\hat{z}, \hat{y}) \in \overline{R}$.
 
-    If $([x], [y]) \in (R^{min}_{\Phi})_{m}$, then we apply the IH.
+    $\hat{z} \overline{R} \hat{y}$ means that there are $z' \in \hat{z}$ and $y' \in \hat{y}$ such that $z' R y'$. $\mathcal{M}, y \models \varphi$ implies $\mathcal{M}, y' \models \varphi$.
+    Then $\mathcal{M}, z' \models \Diamond \varphi$. $\hat{z} \overline{R} \hat{x}$ implies that there are $z'' \in \hat{z}$ and $x' \in \hat{x}$ with $z'' R x'$.
 
-    Suppose $([x], [y]) \in (R^{min}_{\Phi})_{m})^{-1} \circ (R^{min}_{\Phi})_{m}$,
-    then there exists $[z] \in W / \sim_{\Phi}$ such that $([z], [x]) \in (R^{min}_{\Phi})_{m})$ and $([z], [y]) \in (R^{min}_{\Phi})_{m}$.
+    $z'' \sim_{\Gamma} z'$ implies $\mathcal{M}, z'' \models \Diamond \varphi$.
+$\mathcal{M} \models {\bf K}5$, so $\mathcal{M}, z'' \models \Box \Diamond \varphi$.
+Thus, $\mathcal{M}, x' \models \Diamond \varphi$ since $x' \in R(z'')$.
+But $x' \in \hat{x}$, then $\mathcal{M}, x \models \Diamond \varphi$.
 
-    Then one has the following picture (using Lemma~\ref{equivHorn}):
+    \item Suppose $(\overline{R})_n \subseteq R^{max}_{\Gamma}$.
+Let us show that for $(\overline{R})_{n + 1}$.
+Then either $(\hat{x}, \hat{y}) \in (\overline{R})_n$ or $(\hat{z}, \hat{y}), (\hat{z}, \hat{x}) \in (\overline{R})_n$.
+The first alternative holds due to the IH.
+We consider only the second one.
+
+    By Lemma..., we may visualise $(\hat{z}, \hat{y}), (\hat{z}, \hat{x}) \in (\overline{R})_n$ with the following sequences for some $\widehat{u_1}, \dots, \widehat{u_n}, \widehat{t_1}, \dots, \widehat{t_n}$:
 
     \xymatrix{
-&&    [z] & \ar[l]_{R^{min}_{\Phi}} [z_1] \ar@{-}[r]^{R'} & [z_2] \ar@{-}[r]^{R'} & \ar@{-}[r]^{R'} \dots & [z_{m - 1}] \ar@{-}[r]^{R'} & [z_m] \ar[r]^{R^{min}_{\Phi}} & [x] \\
-&&    [z] & \ar[l]^{R^{min}_{\Phi}} [z_1^{'}] \ar@{-}[r]_{R'} & [z_2^{'}] \ar@{-}[r]_{R'} & \ar@{-}[r]_{R'} \dots & [z_{m - 1}] \ar@{-}[r]_{R'} & [z_m^{'}] \ar[r]_{R^{min}_{\Phi}} & [y]
+    &&& \hat{z} & \ar[l]_{\overline{R}} \widehat{u_1} \ar@{-}[r]^{R'} & \widehat{u_2} \ar@{-}[r]^{R'} & \dots \ar@{-}[r]^{R'} & \widehat{u_{n - 1}} \ar@{-}[r]^{R'} & \widehat{u_n} \ar[r]^{\overline{R}} & \hat{y} \\
+    &&& \hat{z} & \ar[l]^{\overline{R}} \widehat{t_1} \ar@{-}[r]_{R'} & \widehat{t_2} \ar@{-}[r]_{R'} & \dots \ar@{-}[r]_{R'} & \widehat{t_{n - 1}} \ar@{-}[r]_{R'} & \widehat{t_n} \ar[r]_{\overline{R}} & \hat{x}
     }
-  Where $R'$ is either $R^{min}_{\Phi}$ or its converse.
-  One has $\mathcal{M}, x \models \Box \varphi$ for $\Box \varphi \in \Phi$, where $\widehat{M}$ is the minimal filtration of $\mathcal{M}$ through $\Phi$.
-One has $[z_m] R^{min}_{\Phi} [x]$, then $a_m R a$ for some $a_m \in [z_n]$ and $a \in [x]$.
-Thus, $\mathcal{M}, a_m \models \Diamond \Box \varphi$ and, thus, $\widehat{\mathcal{M}}, [z_m] \models \Diamond \Box \varphi$.
+    where $R'$ is either $\overline{R}$ or $\overline{R}^{-1}$
 
-  Applying Proposition~\ref{k5useful} several times, one may show that $\widehat{\mathcal{M}}, [z_1] \models \Diamond \Box \varphi$.
-  One has $[z_1] R^{min}_{\Phi} [z]$, then for some $a \in [z]$ and $a_1 \in [z_1]$ we have $a_1 R a$.
+    $\mathcal{M}, y \models \varphi$ and $\widehat{u_n} \overline{R} \hat{y}$, so for some $u_n^{'} \in \widehat{u_n}$ and $y' \in \hat{y}$ with $u_n^{'} R y'$, so $\mathcal{M}, u_n^{'} \models \Diamond \varphi$ and $\widehat{\mathcal{M}}, \widehat{u_n} \models \Diamond \varphi$.
 
-  Then $\mathcal{M}, a \models \Box \varphi$ and $\widehat{M}, [z] \models \Box \varphi$.
+    We have either $\widehat{u_{n - 1}} \overline{R} \widehat{u_n}$ or $\widehat{u_n} \overline{R} \widehat{u_{n - 1}}$.
 
-  We have $[z_1^{'}] R^{min}_{\Phi} [z]$, thus, $a_1^{'} R a^{'}$ for some $a_1^{'} \in [z_1^{'}]$ and $a^{'} \in [z]$. Then
-  $\mathcal{M}, a_1^{'} \models \Diamond \Box \varphi$, and, thus, $\widehat{M}, [z_1^{'}] \models \Diamond \Box \varphi$.
+    If the first holds, then there are $u_{n - 1}^{'} \in \widehat{u_{n - 1}}$ and $u_{n}^{''} \in \widehat{u_{n}}$ with $u_{n - 1}^{'} R u_{n}^{''}$.
+    $\mathcal{M}, u_{n}^{''} \models \Diamond \varphi$. wtf?
 
-  One may show that $\widehat{M}, [z_{m}^{'}] \models \Diamond \Box \varphi$ in the same way via Lemma~\ref{k5useful}.
-Thus, $\mathcal{M}, z_{m}^{'} \models \Diamond \Box \varphi$.
-We also have $\mathcal{M}, z_{m}^{'} \models \Diamond \Box \varphi \to \Box \varphi$, and, thus,
-  $\mathcal{M}, z_{m}^{'} \models \Box \varphi$. Then $\widehat{\mathcal{M}}, [z_m^{'}] \models \Box \varphi$.
+    If the second holds, then there $u_{n - 1}^{'} \in \widehat{u_{n - 1}}$ and $u_{n}^{''} \in \widehat{u_{n}}$ with $u_{n}^{''} R u_{n - 1}^{'}$.
+$\mathcal{M}, u_{n}^{''} \models \Diamond \varphi$ implies $\mathcal{M}, u_{n}^{''} \models \Box \Diamond \varphi$, so $\mathcal{M}, u_{n - 1}^{'} \models \Diamond \varphi$.
 
-  One has $[z_m^{'}] R^{min}_{\Phi} [y]$, then $a_m^{'} R y'$ for some $a_m^{'} \in [z_m^{'}]$ and $y' \in [y]$.
-Then $\mathcal{M}, y' \models \varphi$.
-But $y' \sim_{\Phi} y$, so $\mathcal{M}, y \models \varphi$.
   \end{enumerate}

### 2020-12-10 frp-paper/Paper.tex
new proof for R(;,+) (wip)

\documentclass[a4paper]{article}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{bussproofs}
\usepackage{mathtools}
\usepackage{verbatim}
\usepackage{dsfont}
\usepackage{mathabx}
\usepackage[all, 2cell]{xy}
\usepackage[all]{xy}
\usepackage{wasysym}
\usepackage{rotating}
\usepackage{geometry}
\usepackage{trfsigns}
\usepackage{cmll}
\usepackage{authblk}
\usepackage{hyperref}
\usepackage{cleveref}
\usepackage{lipsum}
\usepackage{extpfeil}
\usepackage{soul}
\usepackage{graphicx}

\newcommand\mapsfrom{\mathrel{\reflectbox{\ensuremath{\mapsto}}}}

\theoremstyle{definition}
\newtheorem{definition}{Definition}

\theoremstyle{theorem}
\newtheorem{theorem}{Theorem}

\theoremstyle{proposition}
\newtheorem{proposition}{Proposition}

\theoremstyle{lemma}
\newtheorem{lemma}{Lemma}

\theoremstyle{ex}
\newtheorem{ex}{Example}

\theoremstyle{corollary}
\newtheorem{corollary}{Corollary}

\theoremstyle{claim}
\newtheorem{claim}{Claim}

\author{Daniel Rogozin \\ Institute for Information Transmission Problems, Russian Academy of Sciences \\ daniel.rogozin@serokell.io}
\date{}
\title{The finite representation property for some reducts of relation algebras}

\begin{document}

\maketitle

\begin{abstract}
In this paper, we show that the class of representable residuated semigroups has the finite representation property. That is, every finite representable residuated semigroup is isomorphic to some algebra over a finite base. This result gives a positive solution to Problem 19.17 from the monograph by Hirsch and Hodkinson \cite{hirsch2002relation}.
\end{abstract}

\section{Introduction}

Relation algebras are the kind of Boolean algebras with operators representing algebras of binary relations \cite{jonsson1951boolean}. One often emphasise the following two classes of relation algebras.
The first class called ${\bf RA}$ consists of algebras the signature of which is $\{ 0, 1, +, -, ;, {}^{\smile}, {\bf 1}' \}$ obeying the certain axioms that we define precisely below.
The second class called ${\bf RRA}$, the class of representable relation algebras, consists of algebras isomorphic to set relation algebras. ${\bf RRA}$ is a subclass of ${\bf RA}$, but the converse inclusion does not hold. That is, there exist relation algebras having no representation as set relation algebras \cite{lyndon1950representation}. Moreover, the class ${\bf RRA}$ is not finitely axiomatisable in contrast to ${\bf RA}$ \cite{monk1964representable}. The problem of determining whether a given finite relation algebra is representable is undecidable, see \cite{hirsch2001representability}.

Under these circumstances, one may conclude that relation algebras are quite badly behaved. The study of such reducts is mostly motivated by such ``bad behaviour'' in order to avoid these restrictions and determine the possible reasons for them.

There are several results on reducts of relation algebras having no finite axiomatisation such as ordered monoids \cite{hirsch2005class}, distributive residuated lattices \cite{andreka1994lambek}, join semilattice-ordered semigroups \cite{andreka2011axiomatizability}, algebras whose signature contains composition, meet, and converse \cite{hodkinson2000axiomatizability}, etc.

On the other hand, such classes as representable residuated semigroups and monoids \cite{andreka1994lambek}, and ordered domain algebras \cite{hirsch2013ordered} are finitely axiomatisable. There are also plenty of subsignatures for which the question of finite axiomatisability remains open, see, e. g., \cite{andreka2011axiomatizability}.

The other direction we emphasise is related to finite representability. A finite algebra of relations has the finite representation property if it is isomorphic to some algebra of relations on a finite base. The investigation is of interest to study such aspects as the decidability of membership of ${\bf R}(\tau)$. The finite representation property also implies recursivity of the class of all finite representable $\tau$-structures \cite{hirsch2004finite}, if the whole class is finitely axiomatisable.  Here, $\tau$ is a subsignature of operations definable in $\{ 0, 1, +, -, ;, {}^{\smile}, {\bf 1 } \}$. The examples of the class having the finite representation property are some classes of algebras \cite{hirsch2004finite} \cite{hirsch2013ordered} \cite{mclean2016finite}, the subsignature of which contains the domain and range operators. The other kind of algebras of binary relations having the finite representation property is semigroups with so-called demonic refinement has been recently studied by Hirsch and \v{S}emrl \cite{hirsch2020finite}.

Other algebras of relations such as weakly associative or relativised cylindric set algebras of a finite dimension have a similar property called the finite base property \cite{andreka1999finite}.

There are subsignatures of relation algebras $\tau$ such that
the class ${\bf R}(\tau)$ of representable reducts fails to have the finite representation property. For instance, $\{;, \cdot\}$, see \cite{maddux2016finite}. In general, (un)decidability of determining whether a finite relation algebra has a finite representation is an open question \cite[Problem 18.18]{hirsch2002relation}.

In this paper, we consider reducts of relation algebras the signature of which consists of composition, residuals, and the binary relation symbol denoting partial ordering, that is, the class of representable residuated semigroups. We show that ${\bf R}(;, \setminus, /, \leq)$ has the finite representation property. As result, Problem 19.17 of \cite{hirsch2002relation} has a positive solution. We also note that this result implies of membership decidability of ${\bf R}(;, \setminus, /, \leq)$ for finite structures and the class of finite representable residuated semigroups is recursive. The solution is surprisingly simple and based on the Dedekind-MacNeille completion and the relation representation of quantales. We consider a finite residuated semigroup and embed into a finite quantale mapping every element to its lower cone. After that, we apply the relational representation for quantales. As a result, the original finite residuated semigroup has a Zaretski-style representation \cite{zaretskii1959representation} and this satisfies the finite base requirement.

\section{Preliminaries}

Let us the basic definitions related to relation algebras \cite{hirsch2002relation}.
\begin{definition} A relation algebra is an algebra $\mathcal{R} = \langle R, 0, 1, +, -, ;, {}^{\smile}, {\bf 1 }\rangle$ such that $\langle R, 0, 1, +, - \rangle$ is a Boolean algebra and the following
    equations hold, for each $a, b, c \in R$:
    \begin{enumerate}
      \item $a ; (b ; c) = (a ; b) ; c$,
      \item $(a + b) ; c = (a ; c) + (b ; c)$,
      \item $a ; {\bf 1}' = a$,
      \item $a^{\smile \smile} = a$,
      \item $(a + b)^{\smile} = a^{\smile} + b^{\smile}$,
      \item $(a ; b)^{\smile} = b^{\smile} ; a^{\smile}$,
      \item $a^{\smile} ; (- (a ; b)) \leq - b$.
    \end{enumerate}
    Note that $a \leq b$ iff $a + b = b$ iff $a \cdot b = a$, where $a \cdot b = - (- a + - b)$. ${\bf RA}$ is the class of all relation algebras.
\end{definition}

\begin{definition}
    A proper relation algebra (or, a set relation algebra) is an algebra $\mathcal{R} = \langle R, 0, 1, \cup, -, ;, {}^{\smile}, {\bf 1 }\rangle$ such that $R \subseteq \mathcal{P}(W)$, where $X$ is a base set, $W \subseteq X \times X$ is an equivalence relation, $0 = \emptyset$, $1 = W$, $\cup$ and $-$ are set-theoretic union and complement respectively, $;$ is relation composition, ${}^{\smile}$ is relation converse,
    ${\bf 1}'$ is the identity relation restricted to $W$, that is:
    \begin{enumerate}
    \item $a ; b = \{ \langle x, z \rangle \in W \: | \: \exists y \: \langle x, y \rangle \in a \: \& \: \langle y, z \rangle \in b \}$
    \item $a^{\smile} = \{ \langle x, y \rangle \in W \: | \: \langle y, x \rangle \in a \}$
    \item ${\bf 1}' = \{ \langle x, y \rangle \in W \: | \: x = y \}$
    \end{enumerate}
       ${\bf PRA}$ is the class of all proper relation algebras. ${\bf RRA}$ is the class of all representable relation algebras, that is, the closure of ${\bf PRA}$ under isomorphic copies.
\end{definition}

We will use the following notation due to, for example, \cite{hirsch2011positive}. Let $\tau$ be a subset of operations and predicates definable in ${\bf RA}$. ${\bf R}(\tau)$ is the class of subalgebras of $\tau$-subreducts of algebras belonging to ${\bf RRA}$. We assume that ${\bf R}(\tau)$ is closed under isomorphic copies.

A $\tau$-structure is \emph{representable} if it is isomorphic to some algebra of relations of $\tau$-signature. A representable finite $\tau$-structure has a \emph{finite representation over a finite base} if it is isomoprhic to some finite representable over a finite base. ${\bf R}(\tau)$ has the finite representation property if every $\mathcal{A} \in {\bf R}(\tau)$ has a finite representation over a finite base.

One may express residuals in every $\mathcal{R} \in {\bf RA}$ as follows using Boolean negation, inversion, and composition as follows:

\begin{enumerate}
  \item $a \setminus b = -(a^{\smile} ; -b)$
  \item $a \: / \: b = - (- a ; b^{\smile})$
\end{enumerate}

These residuals have the following explicit definition in $\mathcal{R} \in {\bf PRA}$:
\begin{enumerate}
  \item $a \setminus b = \{ \langle x, y \rangle \: | \: \forall z \: (z, x) \in a \Rightarrow (z, y) \in b \}$
  \item $a \: / \: b = \{ \langle x, y \rangle \: | \: \forall z \: (y, z) \in b \Rightarrow (x, z) \in a \}$
\end{enumerate}

One may visualise residuals in ${\bf RRA}$ with the following triangles:

\xymatrix{
&& \exists y \ar@{-->}[ddr]^{b} &&& \forall z \ar@{-->}[ddl]_{a} \ar@{-->}[ddr]^{b} &&& \forall z \\
&&&&& \Rightarrow &&& \Leftarrow \\
& x \ar@{-->}[uur]^{a} \ar[rr]_{a;c} && z & x \ar[rr]_{a \setminus b} && y & x \ar@{-->}[uur]^{a} \ar[rr]_{a / b} && y \ar@{-->}[uul]_{b}
}

\section{The case of residuated semigroups}

The problem we are interested in is the following \cite[Problem 19.17]{hirsch2002relation}:

\begin{center}
  Does ${\bf R}(;, \setminus, /, \leq)$ have the finite representation property?
\end{center}

Let us introduce the notion of a residuated semigroup. Historically, residuated structures were introduced by Krull to study ideals of rings \cite{krull1968idealtheorie}. Structures of this kind has been considered further within semantic aspects of substructural logics, see, for example, \cite{jipsen2002survey}.

\begin{definition}
  A residuated semigroup is an algebra $\mathcal{A} = \langle A, ;, \leq, \setminus, / \rangle$ such that $\langle A, ;, \leq, \rangle$ is a partially ordered residuated semigroup and $\setminus, /$ are binary operations satisfying the residuation property:

  \begin{center}
    $b \leq a \setminus c \Leftrightarrow a ; b \leq c \Leftrightarrow a \leq c \: / \: b$
  \end{center}
  ${\bf RS}$ is the class of all residuated semigroups.
\end{definition}

The logic of such structures is the Lambek calculus \cite{lambek1958mathematics} allowing one to characterise inference in categorial grammars, the equivalent version of context-free grammars \cite{pentus1993lambek}. One may define the Lambek calculus as the following (cut-free) sequent calculus:

\begin{definition}
  $ $

  \begin{prooftree}
  \AxiomC{$ $}
  \RightLabel{\scriptsize{ax}}
  \UnaryInfC{$\varphi \vdash \varphi$}
  \end{prooftree}

\begin{minipage}{0.5\textwidth}
  \begin{flushleft}
        \begin{prooftree}
      \AxiomC{$\Gamma \vdash \varphi$}
      \AxiomC{$\Delta, \psi, \Theta \vdash \theta$}
      \RightLabel{$\backslash \vdash $}
      \BinaryInfC{$\Delta, \Gamma, \varphi \backslash \psi, \Theta \vdash \theta$}
    \end{prooftree}

    \begin{prooftree}
      \AxiomC{$\Gamma \vdash \varphi$}
      \AxiomC{$\Delta, \psi, \Theta \vdash \theta$}
      \RightLabel{$/ \vdash $}
      \BinaryInfC{$\Delta, \psi \: / \: \varphi, \Gamma, \Theta \vdash \theta$}
    \end{prooftree}

    \begin{prooftree}
      \AxiomC{$\Gamma, \varphi, \psi, \Delta \vdash \theta$}
      \RightLabel{$\bullet \vdash $}
      \UnaryInfC{$\Gamma, \varphi \bullet \psi, \Delta \vdash \theta$}
    \end{prooftree}
\end{flushleft}
\end{minipage}
\begin{minipage}{0.5\textwidth}
  \begin{flushright}
       \begin{prooftree}
      \AxiomC{$\varphi, \Pi \vdash \psi$}
      \RightLabel{$\vdash \backslash$}
      \UnaryInfC{$\Pi \vdash \varphi \backslash \psi$}
    \end{prooftree}

    \begin{prooftree}
      \AxiomC{$\Pi, \varphi \vdash \psi$}
      \RightLabel{$\vdash /$}
      \UnaryInfC{$\Pi \vdash \varphi \: / \: \psi$}
    \end{prooftree}

    \begin{prooftree}
      \AxiomC{$\Gamma \vdash \varphi$}
      \AxiomC{$\Delta \vdash \psi$}
      \RightLabel{$\vdash \bullet$}
      \BinaryInfC{$\Gamma, \Delta \vdash \varphi \bullet \psi$}
    \end{prooftree}
\end{flushright}
\end{minipage}
\end{definition}

The class ${\bf R}(;, \setminus, /, \leq)$ consists of the following structures:

\begin{definition} \label{rrs}
  Let $A$ be a set of binary relations on some base set $W$ such that $R = \cup A$ is transitive and $W$ is a domain of $R$. A relation residuated semigroup is an algebra $\mathcal{A} = \langle A, ;, \setminus, /, \subseteq \rangle$ where for each $a, b \in A$
  \begin{enumerate}
    \item $a ; b = \{ (x, z) \: | \: \exists y \in W \: ((x, y) \in a \: \& \: (y, z) \in b) \}$,
    \item $a \setminus b = \{ (x, y) \: | \: \forall z \in W \: ((z, x) \in a \Rightarrow (z, y) \in b)\}$,
    \item $a \: / \: b = \{ (x, y) \: | \: \forall z \in W \: ((y, z) \in b \Rightarrow (x, z) \in a)\}$,
    \item $a \leq b$ iff $a \subseteq b$.
  \end{enumerate}
\end{definition}
A residuated semigroup is called \emph{representable} if it is isomorphic to some algebra belonging to ${\bf R}(;, \setminus, /, \leq)$.

\begin{definition}
  Let $\tau = \{ ;, \setminus, /, \leq \}$, $\mathcal{A}$ a $\tau$-structure, and $X$ a base set. An \emph{interpretation} $R$ over a base $X$ maps every $a \in \mathcal{A}$ to a binary relation $a^R \subseteq X \times X$. A \emph{representation} of $\mathcal{A}$ is an interpretation $R$ satisfying the following conditions:

  \begin{enumerate}
    \item $a \leq b$ iff $a^R \subseteq b^R$,
    \item $(a;b)^R = \{ (x, y) \: | \: \exists z \in X \: (x, z) \in a^R \: \& \: (z, x) \in b^R \} = a^R ; b^R$,
    \item $(a \setminus b)^{R} = \{ (x, y) \: | \: \forall z \in X \: ((z, x) \in a^R \Rightarrow (z, y) \in b^R)\} = a^R \setminus b^R$,
    \item $(a \: / \: b)^{R} = \{ (x, y) \: | \: \forall z \in X \: ((y, z) \in a^R \Rightarrow (x, z) \in b^R)\} = a^R \: / \: b^R$.
  \end{enumerate}
\end{definition}

Andr\'{e}ka and Mikul\'{a}s proved the representation theorem for ${\bf RS}$ (\cite{andreka1994lambek}) in the fashion of step-by-step representation. See this paper to learn more about step-by-step representations in general \cite{hirsch1997step}. That also implies relational completeness of the Lambek calculus, the logic of ${\bf RS}$.

This fact also claims that the theory of ${\bf R}(;, \setminus, /, \leq)$ is finitely axiomatisable since their theories coincide, and the class of all residuated semigroup is finitely axiomatisable.

One may rephrase the result of the theorem by Andr\'{e}ka and Mikul\'{a}s as $\mathcal{A}$ is representable iff $\mathcal{A}$ is a residuated semigroup, where $\mathcal{A}$ is a structure of the signature $\{ ;, \setminus, /, \leq \}$. Thus, it is sufficient to show that any finite residuated semigroup has a representation over a finite base in order to show that ${\bf R}(;, \setminus, /, \leq)$ has the finite representation property.

Let us start with preliminary order-theoretic definitions. Recall that a \emph{closure operator} on a poset $\langle P, \leq \rangle$ is a monotone map $j : P \to P$ satisfying $a \leq j a = j j a$ for each $a \in P$. Given $a \in P$, a \emph{lower cone} generated by $a$ is a subset $\downarrow a = \{ x \in P \: | \: x \leq a \}$.

A quantale is a complete lattice-ordered semigroup. Quantales has been initially introduced by Mulvey to provide a noncommutative generalisation of locales, study the spectra of $C^{*}$-algebras, and classify Penrose tilings, see \cite{mulvey1986suppl} \cite{mulvey2005noncommutative}.

\begin{definition}
  A \emph{quantale} is a structure $\mathcal{Q} = \langle Q, ;, \Sigma \rangle$ such that $\mathcal{Q} = \langle Q, \Sigma \rangle$ is a complete lattice, $\langle Q, ; \rangle$ is a semigroup, and the following conditions hold for each $a \in Q$ and $A \subseteq Q$:
  \begin{enumerate}
    \item $a \: ; \: \Sigma A = \Sigma \{ a ; q \: | \: q \in A \}$,
    \item $\Sigma A \: ; \: a = \Sigma \{ q ; a \: | \: q \in A \}$.
  \end{enumerate}
\end{definition}

Note that any quantale is a residuated semigroup as well. Given a quantale $\mathcal{Q} = \langle Q, ;, \Sigma \rangle$, One may express residuals uniquely with supremum and product as follows for each $a, b \in Q$:
\begin{enumerate}
  \item $a \setminus b = \Sigma \{ c \in Q \: | \: a ; c \leq b \}$,
  \item $a \: / \: b = \Sigma \{ c \in Q\: | \: b ; c \leq a \}$.
\end{enumerate}

A quantic nucleus is a closure operator on a quantale allowing one to define subalgebras. Such an operator is a generalisation of a well-known nucleus operator in locale theory and pointfree topology, see, e. g., \cite{bezhanishvili2016locales}. The following definition and the proposition below are due to \cite[Theorem 3.1.1]{rosenthal1990quantales}.
\begin{definition}
  A \emph{quantic nucleus} on a quantale $\langle A, ;, \Sigma \rangle$ is a mapping $j : A \to A$ such that $j$ a closure operator satisfying $j a ; j b \leq j (a ; b)$.
\end{definition}

\begin{proposition} \label{subsemi}
  Let $\mathcal{A} = \langle A, ;, \Sigma \rangle$ be a quantale and $j$ a quantic nucleus, then
  $\mathcal{A}_j = \{ a \in A \: | \: j a = a \}$ forms a quantale, where $a ;_j b = j(a ; b)$ and $\Sigma_j A = j (\Sigma A)$ for each $a, b \in {A}_j$ and $A \subseteq \mathcal{A}_j$.
\end{proposition}

One may embed any residuated semigroup into some quantale with the Dedekind-MacNeille completion (see, for example, \cite{theunissen2007macneille}) as follows. According to Goldblatt \cite{goldblatt2006kripke}, residuated semigroups have the following representation based on quantic nuclei and the Galois connection.

We need this construction to solve the problem, let us discuss it briefly. See Goldblatt's paper to have a complete argument \cite{goldblatt2006kripke}.

Let $\mathcal{A} = \langle A, \leq, ;, \setminus, / \rangle$ be a residuated semigroup. Then $\langle \mathcal{P}(A), ;, \bigcup \rangle$ is a quantale, where the product operation on subsets is defined with the pairwise products of their elements.

Let $X \subseteq A$. We put $lX$ and $uX$ as the sets of lower and upper bounds of $X$ in $A$. We also put $m X = lu X$.
Note that the lower cone of an arbitrary $x$ is $m$-closed, that is,
$m (\downarrow x) = \downarrow x$.

$m : \mathcal{P}(A) \to \mathcal{P}(A)$ is a closure operator and the set

\begin{center}
$(\mathcal{P}(A))_m = \{ X \in \mathcal{P}(S) \: | \: m X = X\}$ )
\end{center}
forms a complete lattice with $\Sigma_{m} \mathcal{X} = m ( \bigcup \mathcal{X})$ and $\Pi_{m} = \bigcap \mathcal{X}$ \cite{davey2002introduction}.

The key observation is that $m$ is a quantic nucleus on $\mathcal{P}(A)$, that is, $m A ; m B \subseteq m (A ; B)$. We refer here to the Goldblatt's paper mentioned above.

Thus, according to Proposition~\ref{subsemi}, $\langle (\mathcal{P}(A))_m, \subseteq, ;_m \rangle$ is a quantale itself since $m$ is a quantic nucleus.

Let us define a mapping $f_m : \mathcal{A} \to (\mathcal{P}(A))_m$ such that $f_m : a \mapsto \downarrow a$. This map is well-defined since any lower cone generated by a point is $m$-closed. Moreover, $f_m$ preserves products, residuals, and existing suprema. In particular, $f_m$ is a residuated semigroup embedding.

As a result, we have the following representation theorem.

\begin{theorem} \label{orsRep}
  Every residuated semigroup has an isomorphic embedding to the subalgebra of some quantale.
\end{theorem}

In their turn, quantales have a relational representation. First of all, let us define a relational quantale.
The notion of a relational quantale was introduced by Brown and Gurr to represent quantales as algebras of relations and study relational semantics of the full Lambek calculus, the logic of bounded residuated lattices, see, e.g., \cite{brown1993representation}.
\begin{definition}
  Let $A$ be a non-empty set. A \emph{relational quantale} on $A$ is an algebra $\langle R, \subseteq, ; \rangle$, where
  \begin{enumerate}
    \item $R \subseteq \mathcal{P}(A \times A)$,
    \item $\langle R, \subseteq \rangle$ is a complete join-semilattice,
    \item $;$ is a relational composition that respects all suprema in both coordinates.
  \end{enumerate}
\end{definition}

The uniqueness of residuals in any quantale implies the following quite obvious fact.
\begin{proposition}\label{ok}
  Let $\mathcal{A}$ be a relational quantale over a base set $X$, then for each $a, b \in \mathcal{A}$
  \begin{enumerate}
    \item $a \setminus b = \{ (x, y) \in X^2 \: | \: \forall z \in X ( (z, x) \in a \Rightarrow (z, y) \in b) \}$,
    \item $a \: / \: b = \{ (x, y) \in X^2 \: | \: \forall z \in X ( (y, z) \in b \Rightarrow (x, z) \in b )\}$.
  \end{enumerate}
\end{proposition}

Let us discuss the relational representation of quantales. This construction is due to Brown and Gurr \cite{brown1993representation}.

Let $\mathcal{Q}$ be a quantale and $\mathcal{G}(\mathcal{Q})$ a set of its generators. We define:

\begin{center}
  $\hat{a} = \{ \langle g, q \rangle \: | \: g \in \mathcal{G}(\mathcal{Q}), q \in \mathcal{Q}, g \leq a ; q \} \:\:\:\: \widehat{\mathcal{Q}} = \{ \hat{a} \: | \: a \in \mathcal{Q} \}$
\end{center}

The mapping $a \mapsto \hat{a}$ satisfies the following conditions:

\begin{enumerate}
\item $a \leq b$ iff $\hat{a} \subseteq \hat{b}$,

\item $\widehat{\Sigma A} = \Sigma \widehat{A}$, $\hat{a} ; \hat{b} = \widehat{a ; b}$, and $\langle \widehat{\mathcal{Q}}, \subseteq, \Sigma \rangle$ is a complete lattice,

\item $\langle \widehat{\mathcal{Q}}, \subseteq, ; \rangle$ is a relational quantale,

\item $\mathcal{Q}$ is isomorphic to $\langle \widehat{\mathcal{Q}}, \subseteq, ; \rangle$,

\item $a \mapsto \hat{a}$ is a quantale isomorphism.
\end{enumerate}

\begin{theorem} \label{quantaleRep}
  Every quantale $\mathcal{Q} = \langle Q, ;, \Sigma \rangle$ is isomorphic to a relational quantale on $Q$ as a base set.
\end{theorem}

We describe how we use Theorem~\ref{orsRep}, Proposition~\ref{ok}, and Theorem~\ref{quantaleRep} and constructions from their proofs to obtain an interpretation of the signature of residuated semigroups on ${\bf R}(;, \setminus, /, \leq)$.

Let $\mathcal{A}$ be a residuated semigroup and $\mathcal{Q}_{\mathcal{A}}$ is a quantale of Galois closed subsets of $\mathcal{A}$. $\widehat{\mathcal{Q}_{\mathcal{A}}}$ is the corresponding relational quantale. Let us define an interpretation $R : \mathcal{A} \to \widehat{\mathcal{Q}_{\mathcal{A}}}$ such that

\begin{center}
  $R : a \mapsto a^{R} = \widehat{\downarrow a}$
\end{center}

According to the lemma below, such an interpretation is a representation. As we have already said above, the mapping $a \mapsto \downarrow a$ is order-preserving. Moreover, this mapping commutes with products and residuals.

\begin{lemma}
  Let $\tau$ be a signature of residuated semigroups. An interpretation $R : \mathcal{A} \to \widehat{\mathcal{Q}_{\mathcal{A}}}$ such that $R : a \mapsto a^{R} = \widehat{\downarrow a}$ is a $\tau$-representation.
\end{lemma}

\begin{proof}
  By Theorem~\ref{orsRep}, $\mathcal{Q}_{\mathcal{A}}$ is isomorphic to $\widehat{\mathcal{Q}_{\mathcal{A}}}$. The isomorphism is established with the mapping $\downarrow{a} \mapsto \widehat{\downarrow a}$ according to Theorem~\ref{quantaleRep}.
  Residuals in $\widehat{\mathcal{Q}_{\mathcal{A}}}$ are well-defined by Proposition~\ref{ok}.
\end{proof}

Theorem~\ref{orsRep}, Theorem~\ref{quantaleRep}, and the lemma above imply the following statement.
\begin{corollary} \label{orsRep2}
  Every residuated semigroup is isomorphic to the subalgebra of some relational quantale.
\end{corollary}

In particular, the representation we proposed implies the solution to \cite[Problem 19.17]{hirsch2002relation}.
\begin{theorem} \label{solution}
  ${\bf R}(;, \setminus, /, \leq)$ has the finite representation property.
\end{theorem}

\begin{proof}
  $ $

  Let $\mathcal{A}$ be a finite residuated semigroup.

  The representation of $\mathcal{A}$ as a subalgebra of a relational quantale clearly belongs to ${\bf R}(;, \setminus, /, \leq)$. This representation has the form

\begin{center}
  $\widehat{\mathcal{A}} = \langle \{ \widehat{\downarrow a} \}_{a \in \mathcal{A}}, ;, \setminus, /, \subseteq \rangle$.
\end{center}

  Moreover, such a representation with the corresponding relational quantale has the finite base, if the original algebra is finite. The base set of the quantale $\widehat{\mathcal{Q}_{\mathcal{A}}}$ is the set of Galois stable subsets of $\mathcal{A}$, the cardinality of which is finite.
\end{proof}

The main corollary of Theorem~\ref{solution} is that the Lambek calculus has the finite model property. Thus, we have a semantical proof of decidability of the Lambek calculus. Before that, there were several algebraic proofs that the Lambek caclulus has the FMP \cite{buszkowski2008infinitary}, but the authors used to consider arbitrary algebras, not representable ones.
Alternatively, one may show that the Lambek calculus is decidable syntactically, that is, via cut elimination and the subformula property \cite{lambek1958mathematics}.

\begin{corollary} \label{fmp}
  The Lambek calculus is complete w.r.t finite relational models (has the FMP).
\end{corollary}

Moreover, finite axiomatisability and having the finite representation property of ${\bf R}(;, \setminus, /, \leq)$ imply that the membership of ${\bf R}(\tau)_{fin}$ is decidable.

\section{The case of join semilattice-ordered semigroups}

A \emph{join semilattice-ordered semigroup} is an algebra $\mathcal{A} = \langle A, ;, + \rangle$ such that $\langle A, ; \rangle$ is a semigroup, $\langle A, + \rangle$ is a join semilattice, and the following identity holds for all $a, b, c \in A$ the identity $a ; (b + c) = a ; b + a ; c$ holds.

Let $\mathcal{A}$ be a join semilattice-ordered semigroup and $X, Y \subseteq \mathcal{A}$, then $X \: ; \: Y = \{ x \: ; \: y \: | \: x \in X, y \in Y \}$.

\begin{definition}
  A \emph{representation} $R$ of a join semilattice-ordered semigroup $\mathcal{A}$
  is an injection $R : \mathcal{A} \to 2^{D \times D}$ (where $D$ is a non-empty base set) such that
  \begin{enumerate}
    \item $(a + b)^R = a^R \cup b^R$
    \item $(a ; b)^R = a^R ; b^R$
  \end{enumerate}
\end{definition}

\section{Axiomatisating ${\bf R}(+, ;)$.}

It is known that the class of all representable join-semilattice ordered semigroups has no finite axiomatisation \cite{andreka1989union}. This proof has been reproduced in \cite{andreka2011axiomatizability} as well.

Let $\mathcal{A}$ is a join-semilattice ordered semigroup, a subset $F \subseteq \mathcal{A}$ is called \emph{prime filter}, if $F$ is upwardly closed and $x + y \in F$ implies $x \in F$ or $y \in F$. The spectrum of $\mathcal{A}$, denoted as $\operatorname{Spec}(\mathcal{A})$, is the set of all prime filters.

\begin{definition} Let $\mathcal{A}$ be a join-semilattice ordered semigroup. A prenetwork over $\mathcal{A}$ is a pair $(E, V, N)$, where $E$ is a set of nodes, $V$ is a set of nodes (an equivalence relation on $E$), and $N$ is a function $l : V \to \operatorname{Spec}(\mathcal{A})$.

  A prenetwork over $\mathcal{A}$ is a network if the following hold:
  \begin{enumerate}
    \item For every $u, v \in E$ and for every $x,y,z \in \mathcal{A}$ $z \in l(u, v)$ and $z \leq x \: ; \: y$ implies $x \in N(u, w)$ and $y \in l(w, v)$ for some $w \in E$,
    \item For every $u, v, w \in E$ one has $l(u, v) ; l(v, w) \subseteq l(u, w)$,
    \item For every $u, v \in E$ and for every $x,y,z \in \mathcal{A}$ $z \in l(u, v)$ and $z \leq x + y$ implies either $x \in l(u, v)$ or $y \in l(u, v)$.
  \end{enumerate}
\end{definition}

Let $\mathcal{N}_1 = (E_1, V_1, l_1)$ and $\mathcal{N}_2 = (E_2, V_2, l_2)$ be (pre)networks, then $\mathcal{N}_1$ is a \emph{subnetwork} of $\mathcal{N}_2$, written as $\mathcal{N}_1 \subseteq \mathcal{N}_2$,
if $E_1 \subseteq E_2$, $V_1 \subseteq V_2$, and for all $x, y \in E_1$ if $a \in l_1(x, y)$ then there is $b \in l_2(x, y)$ with $a \leq b$.

A $\mathcal{N}_1$ is an \emph{induced} sub(pre)network of $\mathcal{N}_2$, if $E_1 \subseteq E_2$ and for all $x, y \in E_1$ one has $l_1(x,y) = l_2(x,y)$.

Let $\{ \mathcal{N}_i \}_{i \in I}$ be an indexed set of prenetworks (where each $\mathcal{N}_i = (E_i, V_i, l_i)$), then $\mathcal{N} = \bigcup \limits_{i \in I} \mathcal{N}_i$ defined as $(E, V, l)$

\begin{enumerate}
  \item $E = \bigcup \limits_{i \in I} E_i$ and $V = \bigcup \limits_{i \in I} V_i$
  \item $l(x, y) = \bigcup \limits_{ i \in I } l_i(x, y)$ for each $x, y \in E$
\end{enumerate}

\begin{definition}
  Let $n \leq \omega$ and $\mathcal{A}$ a join semilattice-ordered semigroup. A game $\mathcal{G}_n(\mathcal{A})$ has $n$ rounds and consists of $n$ finitary networks.

  \begin{enumerate}
    \item Round $0$: $\forall$ picks a prime filter $F_0$ of $(\mathcal{A})$. $\exists$ responds with a prenetwork $\mathcal{N}_0 = (E_0 = \{ x_0, x_1 \}, V_0, l_0)$ such that the $l_0(x_0, x_1) = l_0(x_1, x_0) = F_0$ and $l_0(x_0, x_0) = l_0(x_1, x_1) = {\uparrow {\bf 1}}$.
    \item Round $n + 1$. Suppose, the prenetwork $\mathcal{N}_n = (E_n, V_n, l_n) \rangle$ has been played and $F_n \in \operatorname{Spec}(\mathcal{A})$.

    $\forall$ picks elements $a, b, c \in \mathcal{A}$ with $a \: ; \: b \geq c$

    $\exists$ responds with the network $\mathcal{N}_{n + 1} = (E_{n + 1}, V_{n + 1}, l_{n + 1})$
  \end{enumerate}

  $\exists$ wins for the match of the game $\mathcal{G}_n(\mathcal{A})$ if for each $0 \leq i \leq n$ $\mathcal{N}_i$ is a network. $\exists$ has a winning strategy, if she can win all the matches. Otherwise, $\forall$ wins the game.
\end{definition}

\begin{lemma} Let $\mathcal{A}$ be a join semilattice-ordered semigroup,

  \begin{enumerate}
    \item If $\mathcal{A}$ is representable then $\exists$ have a winning strategy in $\mathcal{G}_{\omega}(\mathcal{A})$.
    \item If $|\mathcal{A}| = \omega$ and $\exists$ have a winning strategy in $\mathcal{G}_{\omega}(\mathcal{A})$ then $\mathcal{A}$ is representable.
    \item For each $n < \omega$ there exists a first-order sentence $\sigma_n$ such that $\exists$ have a winning strategy in $\mathcal{G}_n(\mathcal{A})$ iff $\mathcal{A} \models \sigma_n$.
  \end{enumerate}
\end{lemma}

\begin{proof}
$ $

  \begin{enumerate}
    \item Let $h$ be a representation of $\mathcal{A}$
    \item
    \item
  \end{enumerate}
\end{proof}

\begin{theorem}
  ${\bf R}(+, ;)$ is axiomatised with the axioms of join semilattice-ordered semigroups and $\{ \sigma_n \}_{n < \omega}$.
\end{theorem}

\begin{theorem}
  ${\bf R}(+, ;)$ has the finite representation property.
\end{theorem}

\begin{proof}
  Let $\mathcal{A} \in {\bf R}(+, ;)$ such that $|\mathcal{A}| < \omega$ and $h$ a representation of $\mathcal{A}$
\end{proof}

\section{Acknowledgements}

The author is sincerely grateful to Robin Hirsch, Ian Hodkinson, Stepan Kuznetsov, Valentin Shehtman, Eugeny Zolin, and his supervisor Ilya Shapirovsky for valuable conversations.\\The previous version contained an error. Thanks to Jas Semrl for noticing that.

\bibliographystyle{plain}
\bibliography{Paper}
\end{document}

### 2020-12-21 filtration-stuff/Text.tex
K5 actually doesn't admit strict filtrations

@@
 \begin{theorem}
-  Let $\mathcal{M} = \langle W, R, \vartheta \rangle$ be an Euclidean model, $\Gamma$ a set of $\operatorname{Sub}$-closed formulas, and $\sim$ an equivalence relation having a finite index that respects $\Gamma$, then $\widehat{R} = (R^{min}_{\Gamma})^{E} \subseteq R^{max}_{\Gamma}$.
-
-  Thus, ${\bf K}5$ admits strict filtrations.
+  ${\bf K}45$ admits strict filtrations.
 \end{theorem}
@@
 \begin{proof}
-  We put $R^{min}_{\Gamma} := \overline{R}$ for brevity.
+  Let $\mathcal{M} = \langle W, R, \vartheta \rangle$ be a transitive Euclidean model and $\overline{\mathcal{M}} = \langle \overline{W}, \overline{R}, \overline{\vartheta} \rangle$ its minimal filtration through $\Gamma$, where $\Gamma$ is finite and Sub-closed.
+Let us put $\widehat{R} = \overline{R}^{+} \cup \overline{R}^{E}$.
+Let us show that $\overline{R}^{+} \cup \overline{R}^{E} \subseteq \overline{R}^{max}$.
 
-  Recall that $\overline{R}^{E}$ has the form $\bigcup \limits_{n < \omega} (\overline{R})_n$, where
+  That is, if $\mathcal{M}, y \models \varphi$ for $\Diamond \varphi \in \Gamma$ and $\hat{x} \widehat{R} \hat{y}$, then $\mathcal{M}, x \models \Diamond \varphi$.
+
+  Let $\hat{x} \widehat{R} \hat{y}$. Let us consider the case when $(\hat{x}, \hat{y}) \in \overline{R}^{E}$
+
   \begin{enumerate}
-    \item $\overline{R}_0 = \widehat{R}$
-    \item $\overline{R}_{m + 1} = \overline{R}_n \cup ((\overline{R})^{-1} \circ \overline{R}_n)$
-  \end{enumerate}
-  One needs to show that for each $n < \omega$ $(\overline{R})_n \subseteq R^{max}_{\Gamma}$. We prove this by induction.
-  Suppose $\mathcal{M}, y \models \varphi$ for $\Diamond \varphi \in \Phi$ and $\hat{x} \overline{R}^{E} \hat{x}$. We need $\mathcal{M}, x \models \Diamond \varphi$.
-  \begin{enumerate}
-    \item The case of $n = 0$ is trivial and it follows directly from the definition of a minimal filtration.
-    \item Suppose $n = 1$. $\hat{x} \overline{R}_1 \hat{y}$ means that $(\hat{x}, \hat{y}) \in \overline{R} \cup (\overline{R}^{-1} \circ \overline{R})$.
-    Let us assume that $(\hat{x}, \hat{y}) \in \overline{R}^{-1} \circ \overline{R}$.
-That is, there exists $\hat{z} \in W / \sim_{\Gamma}$ such that $(\hat{z}, \hat{x}), (\hat{z}, \hat{y}) \in \overline{R}$.
+    \item Suppose $(\hat{x}, \hat{y}) \in \overline{R}$, then $\mathcal{M}, x \models \Diamond \varphi$ holds trivially by the definition of the minimal filtration.
+    \item Suppose the statement holds $\overline{R}_n$ and $(\hat{x}, \hat{y}) \in \overline{R}_{n + 1} = \overline{R}_n \cup (\overline{R}_n^{-1} \circ \overline{R}_n)$.
+We consider the case of $(\hat{x}, \hat{y}) \in (\overline{R}_n^{-1} \circ \overline{R}_n)$.
 
-    $\hat{z} \overline{R} \hat{y}$ means that there are $z' \in \hat{z}$ and $y' \in \hat{y}$ such that $z' R y'$. $\mathcal{M}, y \models \varphi$ implies $\mathcal{M}, y' \models \varphi$.
-    Then $\mathcal{M}, z' \models \Diamond \varphi$. $\hat{z} \overline{R} \hat{x}$ implies that there are $z'' \in \hat{z}$ and $x' \in \hat{x}$ with $z'' R x'$.
+    Then there exists $\hat{z}$ such that $(\hat{z}, \hat{x}), (\hat{z}, \hat{y}) \in \overline{R}_n$.
 
-    $z'' \sim_{\Gamma} z'$ implies $\mathcal{M}, z'' \models \Diamond \varphi$.
-$\mathcal{M} \models {\bf K}5$, so $\mathcal{M}, z'' \models \Box \Diamond \varphi$.
-Thus, $\mathcal{M}, x' \models \Diamond \varphi$ since $x' \in R(z'')$.
-But $x' \in \hat{x}$, then $\mathcal{M}, x \models \Diamond \varphi$.
+    By IH, $\mathcal{M}, z \models \Diamond \varphi$.
 
-    \item Suppose $(\overline{R})_n \subseteq R^{max}_{\Gamma}$.
-Let us show that for $(\overline{R})_{n + 1}$.
-Then either $(\hat{x}, \hat{y}) \in (\overline{R})_n$ or $(\hat{z}, \hat{y}), (\hat{z}, \hat{x}) \in (\overline{R})_n$.
-The first alternative holds due to the IH.
-We consider only the second one.
+    $(\hat{z}, \hat{y}) \in \overline{R}_n$ iff there are $\hat{u}_1, \dots, \hat{u}_n$ such that
 
-    By Lemma..., we may visualise $(\hat{z}, \hat{y}), (\hat{z}, \hat{x}) \in (\overline{R})_n$ with the following sequences for some $\widehat{u_1}, \dots, \widehat{u_n}, \widehat{t_1}, \dots, \widehat{t_n}$:
+    \vspace{\baselineskip}
 
     \xymatrix{
-    &&& \hat{z} & \ar[l]_{\overline{R}} \widehat{u_1} \ar@{-}[r]^{R'} & \widehat{u_2} \ar@{-}[r]^{R'} & \dots \ar@{-}[r]^{R'} & \widehat{u_{n - 1}} \ar@{-}[r]^{R'} & \widehat{u_n} \ar[r]^{\overline{R}} & \hat{y} \\
-    &&& \hat{z} & \ar[l]^{\overline{R}} \widehat{t_1} \ar@{-}[r]_{R'} & \widehat{t_2} \ar@{-}[r]_{R'} & \dots \ar@{-}[r]_{R'} & \widehat{t_{n - 1}} \ar@{-}[r]_{R'} & \widehat{t_n} \ar[r]_{\overline{R}} & \hat{x}
+    &&& \hat{z} & \ar[l]^{\widehat{R}} \hat{u}_1 \ar[r]_{\widehat{R}'} & \hat{u}_2 \ar[r]_{\widehat{R}'} & \dots \ar[r]_{\widehat{R}'} & \hat{u}_{n - 1} \ar[r]_{\widehat{R}'} & \hat{u}_n \ar[r]_{\widehat{R}} & \hat{y}
     }
-    where $R'$ is either $\overline{R}$ or $\overline{R}^{-1}$
+    where $\widehat{R}'$ is either $\widehat{R}$ or $\widehat{R}^{-1}$.
+  \end{enumerate}
+  As it is known, $\Diamond \Diamond \varphi \to \Box \Diamond \varphi \in {\bf K}45$.
 
-    $\mathcal{M}, y \models \varphi$ and $\widehat{u_n} \overline{R} \hat{y}$, so for some $u_n^{'} \in \widehat{u_n}$ and $y' \in \hat{y}$ with $u_n^{'} R y'$, so $\mathcal{M}, u_n^{'} \models \Diamond \varphi$ and $\widehat{\mathcal{M}}, \widehat{u_n} \models \Diamond \varphi$.
+  \vspace{\baselineskip}
 
-    We have either $\widehat{u_{n - 1}} \overline{R} \widehat{u_n}$ or $\widehat{u_n} \overline{R} \widehat{u_{n - 1}}$.
+  $\hat{u}_1 \widehat \hat{z}$, that is, $u_1' R z'$ for some $u_1' \in \hat{u}_1$ and $z' \in \hat{z}$.
+  That is, $\mathcal{M}, u_1' \models \Diamond \Diamond \varphi$, so $\mathcal{M}, u_1' \models \Diamond \varphi$ and $\overline{M}, \hat{u}_1 \models \Diamond \varphi$.
 
-    If the first holds, then there are $u_{n - 1}^{'} \in \widehat{u_{n - 1}}$ and $u_{n}^{''} \in \widehat{u_{n}}$ with $u_{n - 1}^{'} R u_{n}^{''}$.
-    $\mathcal{M}, u_{n}^{''} \models \Diamond \varphi$. wtf?
+  We have $\hat{u}_1 \widehat{R}' \hat{u}_2$. Suppose $\mathcal{M}, u_1'' \models \Diamond \varphi$ and $u_1'' R u_2'$.
+  We also have $\mathcal{M}, u_1'' \models \Box \Diamond \varphi$, thus, $\mathcal{M}, u_2' \models \Diamond \varphi$.
 
-    If the second holds, then there $u_{n - 1}^{'} \in \widehat{u_{n - 1}}$ and $u_{n}^{''} \in \widehat{u_{n}}$ with $u_{n}^{''} R u_{n - 1}^{'}$.
-$\mathcal{M}, u_{n}^{''} \models \Diamond \varphi$ implies $\mathcal{M}, u_{n}^{''} \models \Box \Diamond \varphi$, so $\mathcal{M}, u_{n - 1}^{'} \models \Diamond \varphi$.
+  Suppose $\hat{u}_2 \widehat{R} \hat{u}_1$ and $u_2' R u_1''$, then $\mathcal{M}, u_2' \models \Diamond \varphi$.
 
-  \end{enumerate}
+  Similarly, we have $\mathcal{M}, u_i \models \Diamond \varphi$ iff $\mathcal{M}, u_{i + 1} \models \Diamond \varphi$, whenever $\hat{u}_i \widehat{R}' \hat{u}_{i + 1}$.
+
+  Finally, we have $\hat{u}_n \widehat{R} \hat{x}$.
+Thus, $u_n' R x'$ for some $u_n' \in \hat{u}_n$ and $x' \in \hat{x}$.
+$\mathcal{M}, u_n' \models \Diamond \varphi$, so $\mathcal{M}, u_n' \models \Box \Diamond \varphi$.
+Then $\mathcal{M}, x' \models \Diamond \varphi$.
 \end{proof}
 
+\begin{theorem}
+  ${\bf K}5$ does not admit strict filtrations.
+\end{theorem}
+
+\begin{proof}
+  Let us consider a ${\bf K}5$ model whose Euclidean closure of the minimal filtration does not give us a filtration.
+
+  Let us consider a frame called $\mathcal{F}_{bad}$. We define this frame with the following graph:
+
+\vspace{\baselineskip}
+
+  \xymatrix{
+  &&&&&&& w_5\ar@(ul,ur) \ar@<-.5ex>[d] \\
+  &&&&&& w_1\ar@(ul,ur)  & w_4 \ar@<.5ex>[u] \\
+  &&&&&& w_2 \ar[u] & w_3 \ar[u]
+  }
+\end{proof}
+
+Let us define a valuation $\vartheta$ such that $\vartheta(p) = \{ w_5\}$ and $\vartheta(q) = \{ w_1\}$.
+Let us consider a minimal filtration of $\mathcal{M}_{bad}$ through the Sub-closure of $\Gamma = \{ \neg p, \neg \Diamond p\}$.
+
+Clearly that $w_2 \sim_{\Gamma} w_3$, since $\neg p$ and $\neg \Diamond p$ are true both at $w_2$ and $w_3$.
+
+Moreover, $R_{min} \cup (R_{min}^{-1} \circ R_{min})$ is not a subset of $R_{max}$ since $(\hat{w_1}, \hat{w_5}) \in (R_{min}^{-1} \circ R_{min})$, but $\Diamond p$ is not true at $w_5$.
+
+Let us also note that strict filtrations of this model is not Euclidean.
+Suppose by contrary that $\widehat{R}^{\mathcal{E}}$ is a strict filtraction of that model.
+So $R_{min}^E \subseteq \widehat{R}^{\mathcal{E}}$, since $R_{min}^E$ is the minimal Euclidean relation containing $R_{min}$.
+On the other hand, $R_{min}^E \not\subseteq R_{max}$, so is not $\widehat{R}^{\mathcal{E}}$.
 \section{Filtration for {\bf K4}}

### 2020-12-21 frp-paper/Paper.tex
K5 actually doesn't admit strict filtrations

@@
 
+Let $\mathcal{A}$ is a join-semilattice ordered semigroup, a subset $F \subseteq \mathcal{A}$ is called \emph{prime filter}, if $F$ is upwardly closed, downwardly directed, and $x + y \in F$ implies $x \in F$ or $y \in F$.
+The spectrum of $\mathcal{A}$, denoted as $\operatorname{Spec}(\mathcal{A})$, is the set of all prime filters.
+Let $a \in \mathcal{A}$, then $\uparrow a$ is a prime filter generated by $a$.
+
+\begin{lemma}
+Let $\mathcal{A}$ be a join-semilattice ordered semigroup
+
+  \begin{enumerate}
+    \item Every element $a \in \mathcal{A}$ contains in some maximal filter,
+    \item Every maximal filter is prime,
+    \item If $a \not\leq b$ then there exist prime filter $F$ such that $a \in F$ and $b \not\in F$,
+    \item If $a \in F_1$ and $b \in F_2$, where $F_1$ and $F_2$ are prime filters, then there exists a prime filter $F$ such that $F_1 \: \* \: F_2 \subseteq F$.
+  \end{enumerate}
+\end{lemma}
+
+\begin{proof}
+$ $
+
+  \begin{enumerate}
+    \item
+    \item
+    \item
+    \item
+  \end{enumerate}
+\end{proof}
+
+\subsection{Axiomatisating ${\bf R}(+, ;)$ and the finite representation property}
+
 \begin{definition}
@@
 
-\section{Axiomatisating ${\bf R}(+, ;)$.}
-
 It is known that the class of all representable join-semilattice ordered semigroups has no finite axiomatisation \cite{andreka1989union}.
@@
 
-Let $\mathcal{A}$ is a join-semilattice ordered semigroup, a subset $F \subseteq \mathcal{A}$ is called \emph{prime filter}, if $F$ is upwardly closed and $x + y \in F$ implies $x \in F$ or $y \in F$.
-The spectrum of $\mathcal{A}$, denoted as $\operatorname{Spec}(\mathcal{A})$, is the set of all prime filters.
-
 \begin{definition} Let $\mathcal{A}$ be a join-semilattice ordered semigroup.
-A prenetwork over $\mathcal{A}$ is a pair $(E, V, N)$, where $E$ is a set of nodes, $V$ is a set of nodes (an equivalence relation on $E$), and $N$ is a function $l : V \to \operatorname{Spec}(\mathcal{A})$.
+A prenetwork over $\mathcal{A}$ is a pair $(E, V, N)$, where $E$ is a set of nodes, $V$ is a set of nodes such that $\langle E, V \rangle$ is a complete directed graph, and $N$ is a function $l : V \to \operatorname{Spec}(\mathcal{A})$.
 
@@
   \begin{enumerate}
-    \item Round $0$: $\forall$ picks a prime filter $F_0$ of $(\mathcal{A})$.
-$\exists$ responds with a prenetwork $\mathcal{N}_0 = (E_0 = \{ x_0, x_1 \}, V_0, l_0)$ such that the $l_0(x_0, x_1) = l_0(x_1, x_0) = F_0$ and $l_0(x_0, x_0) = l_0(x_1, x_1) = {\uparrow {\bf 1}}$.
-    \item Round $n + 1$. Suppose, the prenetwork $\mathcal{N}_n = (E_n, V_n, l_n) \rangle$ has been played and $F_n \in \operatorname{Spec}(\mathcal{A})$.
+    \item Round $0$: $\forall$ picks elements $a, b$ with $a \not\leq b$.
+By Lemma, there exists a prime filter such that $a \not\in F_0$ and $b \in F_0$.
+$\exists$ responds with a prenetwork $\mathcal{N}_0 = (E_0 = \{ x_0, x_1 \}, V_0 = \{ (x_0, x_1)\}, l_0)$ such that $l_0(x_0, x_1) = F_0$.
+    \item Round $n + 1$. Suppose, the prenetwork $\mathcal{N}_n = (E_n, V_n, l_n)$ has been played and $F_n \in \operatorname{Spec}(\mathcal{A})$.
 
-    $\forall$ picks elements $a, b, c \in \mathcal{A}$ with $a \: ; \: b \geq c$
+    $\forall$ has the following two options:
+    \begin{enumerate}
+      \item ({\bf Composition move}): $\forall$ can pick $x, y, z \in E_n$ with $a \in l_n(x, y)$ and $b \in l_n(y, z)$.
+Then $\exists$ responds with $\mathcal{N}_{n + 1} = (E_{n + 1}, V_{n + 1}, l_{n + 1})$ such that $\mathcal{N}_{n + 1}$
+      is the same as $\mathcal{N}_n$, but $l_{n + 1}(x, z) = F_{n+1}$, where $F_{n+1}$ is a prime filter that contains
+      $l_{n}(x, z) \cup \uparrow(a \: ; \: b)$
+      \item ({\bf Witness move}):
 
-    $\exists$ responds with the network $\mathcal{N}_{n + 1} = (E_{n + 1}, V_{n + 1}, l_{n + 1})$
+      $\forall$ picks nodes $x, y \in E_n$ and $a, b \in \mathcal{A}$ such that
+    \end{enumerate}
   \end{enumerate}
@@
     \item If $|\mathcal{A}| = \omega$ and $\exists$ have a winning strategy in $\mathcal{G}_{\omega}(\mathcal{A})$ then $\mathcal{A}$ is representable.
-    \item For each $n < \omega$ there exists a first-order sentence $\sigma_n$ such that $\exists$ have a winning strategy in $\mathcal{G}_n(\mathcal{A})$ iff $\mathcal{A} \models \sigma_n$.
   \end{enumerate}
@@
     \item
-    \item
   \end{enumerate}
+\end{proof}
+
+\begin{definition}
+  A term network
+\end{definition}
+
+\begin{lemma}
+  For each $n < \omega$ there exists a first-order sentence $\sigma_n$ such that $\exists$ have a winning strategy in $\mathcal{G}_n(\mathcal{A})$ iff $\mathcal{A} \models \sigma_n$.
+\end{lemma}
+
+\begin{proof}
 \end{proof}
@@
 
-The author is sincerely grateful to Robin Hirsch, Ian Hodkinson, Stepan Kuznetsov, Valentin Shehtman, Eugeny Zolin, and his supervisor Ilya Shapirovsky for valuable conversations.\\The previous version contained an error.
-Thanks to Jas Semrl for noticing that.
+The author is sincerely grateful to Robin Hirsch, Ian Hodkinson, Stepan Kuznetsov, Valentin Shehtman, Eugeny Zolin, and his supervisor Ilya Shapirovsky for valuable conversations.\\The previous version of this paper contained an error.
+Many thanks to Jas Semrl for noticing that.
 

### 2020-12-22 frp-paper/Paper.tex
the R(;,+) case (wip): fixes and thoughts

--- 2020-12-10 frp-paper/Paper.tex
@@
 
+Let $\langle P, \leq \rangle$ be a partial order.
+An upper cone generated by $x$ is the set $\uparrow x = \{ a \in P \: | \: x \leq a \}$.
+Let $A \subseteq P$, then $\uparrow A = \bigcup \limits_{x \in A} \uparrow x = \{ a \in P \: | \: \exists x \in P \: x \leq a \}$.
+The set of all upper cones of a poset $\langle P, \leq \rangle$ is denoted as $\operatorname{Up}(P)$.
+
+\subsection{Axiomatisating ${\bf R}(+, ;)$ and the finite representation property}
+
 \begin{definition}
@@
 
-Let $\mathcal{A}$ is a join-semilattice ordered semigroup, a subset $F \subseteq \mathcal{A}$ is called \emph{prime filter}, if $F$ is upwardly closed and $x + y \in F$ implies $x \in F$ or $y \in F$.
-The spectrum of $\mathcal{A}$, denoted as $\operatorname{Spec}(\mathcal{A})$, is the set of all prime filters.
-
 \begin{definition} Let $\mathcal{A}$ be a join-semilattice ordered semigroup.
-A prenetwork over $\mathcal{A}$ is a pair $(E, V, N)$, where $E$ is a set of nodes, $V$ is a set of nodes (an equivalence relation on $E$), and $N$ is a function $l : V \to \operatorname{Spec}(\mathcal{A})$.
+A prenetwork over $\mathcal{A}$ is a pair $(E, V, N)$, where $E$ is a set of nodes, $V$ is a set of nodes such that $\langle E, V \rangle$ is a complete directed graph, and $N$ is a function $l : V \to \operatorname{Up}(\mathcal{A})$.
 
@@
   \begin{enumerate}
-    \item For every $u, v \in E$ and for every $x,y,z \in \mathcal{A}$ $z \in l(u, v)$ and $z \leq x \: ; \: y$ implies $x \in N(u, w)$ and $y \in l(w, v)$ for some $w \in E$,
-    \item For every $u, v, w \in E$ one has $l(u, v) ; l(v, w) \subseteq l(u, w)$,
-    \item For every $u, v \in E$ and for every $x,y,z \in \mathcal{A}$ $z \in l(u, v)$ and $z \leq x + y$ implies either $x \in l(u, v)$ or $y \in l(u, v)$.
+    \item For every $u, v \in E$ and for every $x,y,z \in \mathcal{A}$ $z \in l(u, v)$ and $z \leq x \: ; \: y$ implies $x \in N(u, w)$ and $y \in l(w, v)$ for some $w \in E$.
+    \item For every $u, v, w \in E$ one has $l(u, v) ; l(v, w) \subseteq l(u, w)$.
+    \item For every $u, v \in E$ $l(u,v)$ is join-prime. That is, for each $a, b \in \mathcal{A}$ if $a + b \in l(u,v)$, then either $a \in l(u,v)$ or $b \in l(u,v)$.
   \end{enumerate}
@@
 Let $\mathcal{N}_1 = (E_1, V_1, l_1)$ and $\mathcal{N}_2 = (E_2, V_2, l_2)$ be (pre)networks, then $\mathcal{N}_1$ is a \emph{subnetwork} of $\mathcal{N}_2$, written as $\mathcal{N}_1 \subseteq \mathcal{N}_2$,
-if $E_1 \subseteq E_2$, $V_1 \subseteq V_2$, and for all $x, y \in E_1$ if $a \in l_1(x, y)$ then there is $b \in l_2(x, y)$ with $a \leq b$.
+if $E_1 \subseteq E_2$, $V_1 \subseteq V_2$, and for all $x, y \in E_1$ if $a \in l_1(x, y)$ then there is $b \in l_2(x, y)$ with $b \leq a$.
 
@@
   \begin{enumerate}
-    \item Round $0$: $\forall$ picks a prime filter $F_0$ of $(\mathcal{A})$.
-$\exists$ responds with a prenetwork $\mathcal{N}_0 = (E_0 = \{ x_0, x_1 \}, V_0, l_0)$ such that the $l_0(x_0, x_1) = l_0(x_1, x_0) = F_0$ and $l_0(x_0, x_0) = l_0(x_1, x_1) = {\uparrow {\bf 1}}$.
-    \item Round $n + 1$. Suppose, the prenetwork $\mathcal{N}_n = (E_n, V_n, l_n) \rangle$ has been played and $F_n \in \operatorname{Spec}(\mathcal{A})$.
+    \item Round $0$: $\forall$ picks elements $a, b \in \mathcal{A}$ with $a \not\leq b$.
+$\exists$ responds with a prenetwork $\mathcal{N}_0 = (E_0 = \{ x_0, x_1 \}, V_0 = \{ (x_0, x_1)\}, l_0)$ such that $l_0(x_0, x_1) = \uparrow b$.
+    \item Round $n + 1$. Suppose, the prenetwork $\mathcal{N}_n = (E_n, V_n, l_n)$ has been played.
 
-    $\forall$ picks elements $a, b, c \in \mathcal{A}$ with $a \: ; \: b \geq c$
+    $\forall$ has the following three options:
+    \begin{enumerate}
+      \item ({\bf Composition move}): $\forall$ can pick $x, y, z \in E_n$ with $a \in l_n(x, y)$ and $b \in l_n(y, z)$.
+We denote such a move as $N(x,y,z,a,b)$.
+Then $\exists$ responds with $\mathcal{N}_{n + 1} = (E_{n + 1}, V_{n + 1}, l_{n + 1})$ such that $\mathcal{N}_{n + 1}$
+      is the same as $\mathcal{N}_n$, but $l_{n + 1}(x, z) = \uparrow (l_{n}(x, z) \cup \{ a \: ; \: b\})$.
+      \item ({\bf Witness move}):
 
-    $\exists$ responds with the network $\mathcal{N}_{n + 1} = (E_{n + 1}, V_{n + 1}, l_{n + 1})$
+      $\forall$ picks nodes $x, y \in E_n$ and $a, b \in \mathcal{A}$ such that $c \leq a ; b$ for $c \in l_n(x, y)$.
+$\exists$ has to find a witness.
+She responds with a fresh node $z$.
+We define the prenetwork $T$ , the edges of which are $x, y, z$ with the following labelling:
+      \begin{enumerate}
+        \item $l_T(x, z) = \uparrow a$
+        \item $l_T(z, y) = \uparrow b$
+      \end{enumerate}
+      For her response, $\exists$ plays $N_{n + 1} = N_n \cup T$
+      \item ({\bf Join move}):
+
+      $\forall$ picks nodes $x, y \in E_n$ and $a + b$ for $a, b \in \mathcal{A}$. $\exists$ has the following two alternatives for her response:
+      \begin{enumerate}
+        \item $\exists$ chooses $a$ and responds with the prenetwork $\mathcal{N}_{n + 1} = \langle E_n, V_{n}, l_{n + 1} \rangle$, where $l_{n+1}(x, y) = \uparrow (l_n(x,y) \cup \{ a \})$.
+        \item $\exists$ chooses $b$. The response is similar but $l_{n + 1}(x, y) = \uparrow (l_n(x,y) \cup \{ b \}$.
+      \end{enumerate}
+    \end{enumerate}
   \end{enumerate}
@@
   \begin{enumerate}
-    \item If $\mathcal{A}$ is representable then $\exists$ have a winning strategy in $\mathcal{G}_{\omega}(\mathcal{A})$.
-    \item If $|\mathcal{A}| = \omega$ and $\exists$ have a winning strategy in $\mathcal{G}_{\omega}(\mathcal{A})$ then $\mathcal{A}$ is representable.
-    \item For each $n < \omega$ there exists a first-order sentence $\sigma_n$ such that $\exists$ have a winning strategy in $\mathcal{G}_n(\mathcal{A})$ iff $\mathcal{A} \models \sigma_n$.
+    \item If $\mathcal{A}$ is representable then $\exists$ has a winning strategy in $\mathcal{G}_{\omega}(\mathcal{A})$.
+    \item If $|\mathcal{A}| = \omega$ and $\exists$ has a winning strategy in $\mathcal{G}_{\omega}(\mathcal{A})$ then $\mathcal{A}$ is representable.
   \end{enumerate}
@@
   \begin{enumerate}
-    \item Let $h$ be a representation of $\mathcal{A}$
-    \item
-    \item
+    \item Let $h$ be a representation of $\mathcal{A}$ over some domain $D$.
+$\forall$ plays $(a, b)$ with $a \not\leq b$ for the initial round.
+Let $N = \langle E, V, l \rangle$ be a network being played.
+
+    TODO: define an embedding $\iota : E \to D$ and check the presence of a winning strategy explicitly
+
+    \item Let $a, b \in \mathcal{A}$ such that $a \not\leq b$.
+We consider a play of the game where $\forall$ pick the pair $(a, b)$ in the initial round and plays $(\mathcal{N},x, y, z, c, d)$ in the further rounds for each $x, y, z \in \operatorname{Nodes}(\mathcal{N})$ and $c,d \in \mathcal{A}$.
+Here, $c \in l_N(x, y)$ and $d \in l_N(y, z)$.
+
+    $\forall$ also plays all rounds $(\mathcal{N},x,y,c,d)$ for each $x, y \in \operatorname{Nodes}(\mathcal{N})$ and $c,d \in \mathcal{A}$ such that there is $e \in \mathcal{A}$ such that
+    $e \leq c ; d$ and $e \in l_N(x,y)$.
+
+    $\forall$ picks also $c + d$ and vertices $x,y \in \operatorname{Nodes}(\mathcal{N})$ for $c, d \in \mathcal{A}$.
+
+    We have the following play of a game that we define as a sequence of prenetworks where $\exists$ has a winning strategy for each $i < \omega$:
+
+    \begin{center}
+      $\mathcal{N}_0 \subseteq \mathcal{N}_1 \subseteq \mathcal{N}_2 \dots$
+    \end{center}
+
+    Let us put $\mathcal{N} = \bigcup \limits_{i < \omega} \mathcal{N}_i$.
+
+    Clearly that every $l_{\mathcal{N}}(x, y)$ for each $x, y \in \operatorname{Nodes}(\mathcal{N})$ is join-prime. Let $a + b \in l_{\mathcal{N}}(x, y)$, then there exists $i < \omega$ with
+    $a + b \in l_{\mathcal{N}_i}(x, y)$
+
+    $\forall$ makes all possible join moves, so either $a \in l_{\mathcal{N}_{i + 1}}(x, y)$ or $b \in l_{\mathcal{N}_{i + 1}}(x, y)$ depending on what $\exists$ chooses.
+
+    Let us ensure that the labelling function $l_{\mathcal{N}}$ respects $;$. Here, the argument is similar to the proof of \cite[Proposition 5]{hirsch2005class}.
+
+    The map $rep$ defined as follows is a representation:
+    \begin{center}
+      $rep(a) = \{ (x, y) \: | \: a \in l_{\mathcal{N}}(x, y)\}$
+    \end{center}
+    This map is well-defined.
+
+    $rep(a + b) = \{ (x, y) \: | \: a + b \in l_{\mathcal{N}}(x, y)\} \subseteq \{ (x, y) \: | \: a \in l_{\mathcal{N}}(x, y) \lor b \in l_{\mathcal{N}}(x, y) \} = rep(a) \cup rep(b)$.
+
+    The converse inclusion:
+
+    Suppose $(x,y) \in rep(a) \cup rep(b)$, so either $a \in l_{\mathcal{N}}(x,y)$ or $b \in l_{\mathcal{N}}(x,y)$.
+If $a \in l_{\mathcal{N}}(x,y)$, then $a + b \in l_{\mathcal{N}}(x,y)$ since $a \leq a + b$ and $l_{\mathcal{N}}(x,y)$ is an upset.
+The last holds because
+    $l_{\mathcal{N}}(x,y) = \bigcup \limits_{i < \omega} l_{\mathcal{N}_i}(x,y)$ and the union of upsets is an upset.
+
+    The second alternative ($b \in l_{\mathcal{N}}(x,y)$) is similar to the second one.
+
+    Thus, $rep(a + b) = rep(a) \cup rep(b)$.
+
+\medskip
+
+Let us check that $rep(a ; b) = rep(a) | rep(b)$.
+
+$rep(a;b) = \{ (x,y) \: | \: a;b \in l_{\mathcal{N}}(x,y) \}$
+
+Suppose $(x,y) \in rep(a) ; rep(b)$, we need $(x, y) \in rep(a ; b)$, i.e.
+$a;b \in l_{\mathcal{N}}(x,y)$.
+Thus, there exists $z \in \operatorname{Nodes}(\mathcal{N})$ such that $(x,z) \in rep(a)$ and $(z,y) \in rep(b)$, i.e.
+$a \in l_{\mathcal{N}}(x,z)$ and $b \in l_{\mathcal{N}}(z, y)$.
+$\mathcal{N}$ is a network, so $l_{\mathcal{N}}(x,z) ; l_{\mathcal{N}}(z,y) \subseteq l_{\mathcal{N}}(x,y)$.
+Thus, $a ; b \in l_{\mathcal{N}}(x,y)$.
+
+Suppose for the converse, $(x, y) \in rep(a ; b)$, i.e.
+$a;b \in l_{\mathcal{N}}(x,y)$.
+We need some $z$ such that $(x, z) \in rep(a)$ and $(z, y) \in rep(b)$, i.e.
+$a \in l_{\mathcal{N}}(x,z)$ and $b \in l_{\mathcal{N}}(z,y)$.
+(Ideally, we need some $c$ with $c \leq a ; b$, but $l_{\mathcal{N}}(x,y)$ is not downward directed)
   \end{enumerate}
+\end{proof}
+
+\begin{definition}
+  Let $V = \{ v_0, v_1, ... \}$ be a countable set of variables. The set of terms $T$ is generated by the following grammar:
+
+  \begin{center}
+    $T ::= V \: | \: (T + T) \: | \: (T ; T)$
+  \end{center}
+\end{definition}
+
+\begin{definition}
+  A term network is a graph $\langle E, V, l \rangle$, where $\langle E, V \rangle$ is a complete directed graph and $l : V \to 2^T$ is a labelling function.
+\end{definition}
+
+Let $\mathcal{A}$ be a join semilattice-ordered semigroup and $\vartheta : V \to \underline{{\mathcal{A}}}$ a valuation. The value of complex terms is defined inductively for $a, b \in T$:
+
+\begin{enumerate}
+  \item $(a ; b)^{\vartheta} = a^{\vartheta} ; b^{\vartheta}$
+  \item $(a + b)^{\vartheta} = a^{\vartheta} + b^{\vartheta}$
+\end{enumerate}
+
+Let $\mathcal{N} = \langle E, V, l \rangle$ be a term prenetwork, $\mathcal{A}$ be a join-semilattice ordered semigroup and $\vartheta : V \to \underline{{\mathcal{A}}}$ a valuation.
+Let us define the prenetwork $\mathcal{N}^{\vartheta}$ with the same edges and vertices with the labelling $l^{\vartheta}(x, y) = \{ t^{\vartheta} \: | \: t \in l(x, y)\}$.
+
+\begin{lemma}
+  For each $n < \omega$ there exists a first-order sentence $\sigma_n$ such that $\exists$ has a winning strategy in $\mathcal{G}_n(\mathcal{A})$ iff $\mathcal{A} \models \sigma_n$.
+\end{lemma}
+
+\begin{proof}
+
 \end{proof}
@@
 \begin{proof}
-  Let $\mathcal{A} \in {\bf R}(+, ;)$ such that $|\mathcal{A}| < \omega$ and $h$ a representation of $\mathcal{A}$
+  Let $\mathcal{A} \in {\bf R}(+, ;)$ such that $|\mathcal{A}| < \omega$ and $h$ a representation of $\mathcal{A}$. monkahmm
 \end{proof}
 
+\begin{comment}
 \section{Acknowledgements}
 
-The author is sincerely grateful to Robin Hirsch, Ian Hodkinson, Stepan Kuznetsov, Valentin Shehtman, Eugeny Zolin, and his supervisor Ilya Shapirovsky for valuable conversations.\\The previous version contained an error.
-Thanks to Jas Semrl for noticing that.
+The author is sincerely grateful to Robin Hirsch, Ian Hodkinson, Stepan Kuznetsov, Valentin Shehtman, Eugeny Zolin, and his supervisor Ilya Shapirovsky for valuable conversations.\\The previous version of this paper contained an error.
+Many thanks to Jas Semrl for noticing that.
+\end{comment}
 

### 2021-01-03 frp-paper/Paper.tex
R(;, +): monkahmm

@@
 
-    Let us put $\mathcal{N} = \bigcup \limits_{i < \omega} \mathcal{N}_i$.
+    Let us put $\mathcal{N}^{*}(a, b) = \bigcup \limits_{i < \omega} \mathcal{N}_i$. It is clear that $\mathcal{N}^{*}$ is a network.
 
-    Clearly that every $l_{\mathcal{N}}(x, y)$ for each $x, y \in \operatorname{Nodes}(\mathcal{N})$ is join-prime. Let $a + b \in l_{\mathcal{N}}(x, y)$, then there exists $i < \omega$ with
-    $a + b \in l_{\mathcal{N}_i}(x, y)$
+    Let us put the following network:
 
+    \begin{center}
+      $\mathcal{N} = \bigcup \limits_{\substack{a, b \in \mathcal{A} \\ a \not\leq b}} \mathcal{N}^{*}(a,b)$
+    \end{center}
+
+    $\mathcal{N}$ is really a network since it is the disjoint union of networks.
+
+    The labelling function $l_{\mathcal{N}}$ respects $;$. Here, the argument is similar to the proof of \cite[Proposition 5]{hirsch2005class}.
+
+    Let us define the $rep : \mathcal{A} \to \operatorname{Nodes}(\mathcal{N})$:
+    \begin{center}
+      $rep(a) = \{ (x, y) \: | \: \exists b \in \downarrow a \:\: b \in l_{\mathcal{N}}(x, y)\}$
+    \end{center}
+
+    Let us ensure that $rep$ is a representation.
+
+    \begin{itemize}
+  \item Let us show that $rep(a + b) = rep(a) \cup rep(b)$
+
+  Suppose $(x, y) \in rep(a + b)$. That is, there exists $c \leq a + b$ with $c \in l_{\mathcal{N}}(x, y)$, so does $a + b$ since $l_{\mathcal{N}}$ is an upper cone.
+
+  $a + b \in l_{\mathcal{N}}(x, y)$, that is,
+
+\begin{center}
+$a + b \in \bigcup \limits_{ \substack{a_1, a_0 \in \mathcal{A} \\ a_1 \not\leq a_0}} l_{\mathcal{N}^{*}(a_1, a_0)}(x,y)$
+\end{center}
+
+That is, there are $a_1, a_0 \in \mathcal{A}$ with $a_1 \not\leq a_0$ such that $a + b \in l_{\mathcal{N}^{*}(a_1, a_0)}(x,y)$.
+This means $a + b \in \bigcup \limits_{i < \omega} \mathcal{N}_i$.
 $\forall$ makes all possible join moves, so either $a \in l_{\mathcal{N}_{i + 1}}(x, y)$ or $b \in l_{\mathcal{N}_{i + 1}}(x, y)$ depending on what $\exists$ chooses.
+Thus, $(x, y) \in rep(a) \cup rep(b)$.
 
-    Let us ensure that the labelling function $l_{\mathcal{N}}$ respects $;$. Here, the argument is similar to the proof of \cite[Proposition 5]{hirsch2005class}.
+Suppose for the converse, $(x, y) \in rep(a)$. We need $(x, y) \in rep(a + b)$.
+In other words, we have some $c \in \mathcal{A}$ with $c \leq a$ and $c \in l_{\mathcal{N}}(x, y)$.
+We have $c \leq a \leq a + b$, so $(x, y) \in rep(a + b)$.
 
-    The map $rep$ defined as follows is a representation:
-    \begin{center}
-      $rep(a) = \{ (x, y) \: | \: a \in l_{\mathcal{N}}(x, y)\}$
-    \end{center}
-    This map is well-defined.
+\item Let us show that $rep(a ; b) = rep(a) ; rep(b)$.
 
-    $rep(a + b) = \{ (x, y) \: | \: a + b \in l_{\mathcal{N}}(x, y)\} \subseteq \{ (x, y) \: | \: a \in l_{\mathcal{N}}(x, y) \lor b \in l_{\mathcal{N}}(x, y) \} = rep(a) \cup rep(b)$.
+Suppose $(x, y) \in rep(a ; b)$. We need some $z$ with $(x, z) \in rep(a)$ and $(z, y) \in rep(b)$.
 
-    The converse inclusion:
+There is $c \leq a ; b$ with $c \in l_{\mathcal{N}}(x,y)$.
+That is, there are $a_1, a_0 \in \mathcal{A}$ and $\mathcal{N}_i$ such that $c \in l_{\mathcal{N}_i}(x, y)$ where $\forall$ plays $(a_1, a_0)$ for the initial round.
 
-    Suppose $(x,y) \in rep(a) \cup rep(b)$, so either $a \in l_{\mathcal{N}}(x,y)$ or $b \in l_{\mathcal{N}}(x,y)$.
-If $a \in l_{\mathcal{N}}(x,y)$, then $a + b \in l_{\mathcal{N}}(x,y)$ since $a \leq a + b$ and $l_{\mathcal{N}}(x,y)$ is an upset.
-The last holds because
-    $l_{\mathcal{N}}(x,y) = \bigcup \limits_{i < \omega} l_{\mathcal{N}_i}(x,y)$ and the union of upsets is an upset.
+By the condition, $\forall$ makes all witness moves and $\exists$ responds with a witness.
+Her response is a fresh node $z$ such that $l_{\mathcal{N}_{i + 1}}(x, z) =  \uparrow (l_{\mathcal{N}_i}(x, z) \cup \{ a \})$ and $l_{\mathcal{N}_{i + 1}}(z, y) = \uparrow (l_{\mathcal{N}_i}(z, y) \cup \{ b\})$.
 
-    The second alternative ($b \in l_{\mathcal{N}}(x,y)$) is similar to the second one.
+Let us show the following subitems:
+\begin{itemize}
+  \item $(x, z) \in rep(a)$ ??
+  \item $(z, y) \in rep(b)$ ??
+\end{itemize}
 
-    Thus, $rep(a + b) = rep(a) \cup rep(b)$.
-
-\medskip
-
-Let us check that $rep(a ; b) = rep(a) | rep(b)$.
-
-$rep(a;b) = \{ (x,y) \: | \: a;b \in l_{\mathcal{N}}(x,y) \}$
-
-Suppose $(x,y) \in rep(a) ; rep(b)$, we need $(x, y) \in rep(a ; b)$, i.e.
-$a;b \in l_{\mathcal{N}}(x,y)$.
-Thus, there exists $z \in \operatorname{Nodes}(\mathcal{N})$ such that $(x,z) \in rep(a)$ and $(z,y) \in rep(b)$, i.e.
-$a \in l_{\mathcal{N}}(x,z)$ and $b \in l_{\mathcal{N}}(z, y)$.
-$\mathcal{N}$ is a network, so $l_{\mathcal{N}}(x,z) ; l_{\mathcal{N}}(z,y) \subseteq l_{\mathcal{N}}(x,y)$.
-Thus, $a ; b \in l_{\mathcal{N}}(x,y)$.
-
-Suppose for the converse, $(x, y) \in rep(a ; b)$, i.e.
-$a;b \in l_{\mathcal{N}}(x,y)$.
-We need some $z$ such that $(x, z) \in rep(a)$ and $(z, y) \in rep(b)$, i.e.
-$a \in l_{\mathcal{N}}(x,z)$ and $b \in l_{\mathcal{N}}(z,y)$.
-(Ideally, we need some $c$ with $c \leq a ; b$, but $l_{\mathcal{N}}(x,y)$ is not downward directed)
+\end{itemize}
   \end{enumerate}

### 2021-01-03 frp-paper/Paper.tex
if Heloise has a winning strategy on a countable structure then it's representable

@@
 
-By the condition, $\forall$ makes all witness moves and $\exists$ responds with a witness.
+By the condition, $\forall$ makes the witness moves and $\exists$ responds with a witness.
 Her response is a fresh node $z$ such that $l_{\mathcal{N}_{i + 1}}(x, z) =  \uparrow (l_{\mathcal{N}_i}(x, z) \cup \{ a \})$ and $l_{\mathcal{N}_{i + 1}}(z, y) = \uparrow (l_{\mathcal{N}_i}(z, y) \cup \{ b\})$.
+The inclusion $rep(a ; b) \subseteq rep(a) ; rep(b)$ holds since all witness moves have been played and for every such a move there exists a witness.
 
-Let us show the following subitems:
-\begin{itemize}
-  \item $(x, z) \in rep(a)$ ??
-  \item $(z, y) \in rep(b)$ ??
-\end{itemize}
+\item Suppose $(x, y) \in rep(a) ; rep(b)$.
+We need $(x, y) \in rep(a;b)$.
+That is, there exists $z \in \mathcal{Nodes}(\mathcal{N})$ with $(x, z) \in rep(a)$ and $(z, y) \in rep(b)$.
+So, there are $c, d$ such that $c \leq a$ with $c \in l_{\mathcal{N}}(x, z)$ and $d \leq b$ with $d \in l_{\mathcal{N}}(z, y)$.
+We also know that $l_{\mathcal{N}}(x, z) ; l_{\mathcal{N}}(z, y) \subseteq l_{\mathcal{N}}(x, y)$ because all composition moves have been played.
+So $c ; d \in l_{\mathcal{N}}(x, y)$. That makes $(x, y) \in rep(a;b)$ since $c ; d \leq a ; b$.
 
+\item Injectivity: TODO.
 \end{itemize}

### 2021-01-23 frp-paper/Paper.tex
wip

@@
 The logic of such structures is the Lambek calculus \cite{lambek1958mathematics} allowing one to characterise inference in categorial grammars, the equivalent version of context-free grammars \cite{pentus1993lambek}.
-One may define the Lambek calculus as the following (cut-free) sequent calculus:
-
-\begin{definition}
-  $ $
-
-  \begin{prooftree}
-  \AxiomC{$ $}
-  \RightLabel{\scriptsize{ax}}
-  \UnaryInfC{$\varphi \vdash \varphi$}
-  \end{prooftree}
-
-\begin{minipage}{0.5\textwidth}
-  \begin{flushleft}
-        \begin{prooftree}
-      \AxiomC{$\Gamma \vdash \varphi$}
-      \AxiomC{$\Delta, \psi, \Theta \vdash \theta$}
-      \RightLabel{$\backslash \vdash $}
-      \BinaryInfC{$\Delta, \Gamma, \varphi \backslash \psi, \Theta \vdash \theta$}
-    \end{prooftree}
-
-    \begin{prooftree}
-      \AxiomC{$\Gamma \vdash \varphi$}
-      \AxiomC{$\Delta, \psi, \Theta \vdash \theta$}
-      \RightLabel{$/ \vdash $}
-      \BinaryInfC{$\Delta, \psi \: / \: \varphi, \Gamma, \Theta \vdash \theta$}
-    \end{prooftree}
-
-    \begin{prooftree}
-      \AxiomC{$\Gamma, \varphi, \psi, \Delta \vdash \theta$}
-      \RightLabel{$\bullet \vdash $}
-      \UnaryInfC{$\Gamma, \varphi \bullet \psi, \Delta \vdash \theta$}
-    \end{prooftree}
-\end{flushleft}
-\end{minipage}
-\begin{minipage}{0.5\textwidth}
-  \begin{flushright}
-       \begin{prooftree}
-      \AxiomC{$\varphi, \Pi \vdash \psi$}
-      \RightLabel{$\vdash \backslash$}
-      \UnaryInfC{$\Pi \vdash \varphi \backslash \psi$}
-    \end{prooftree}
-
-    \begin{prooftree}
-      \AxiomC{$\Pi, \varphi \vdash \psi$}
-      \RightLabel{$\vdash /$}
-      \UnaryInfC{$\Pi \vdash \varphi \: / \: \psi$}
-    \end{prooftree}
-
-    \begin{prooftree}
-      \AxiomC{$\Gamma \vdash \varphi$}
-      \AxiomC{$\Delta \vdash \psi$}
-      \RightLabel{$\vdash \bullet$}
-      \BinaryInfC{$\Gamma, \Delta \vdash \varphi \bullet \psi$}
-    \end{prooftree}
-\end{flushright}
-\end{minipage}
-\end{definition}
+One may define the Lambek calculus as the following (cut-free) sequent calculus.
 
@@
     \item If $\mathcal{A}$ is representable then $\exists$ has a winning strategy in $\mathcal{G}_{\omega}(\mathcal{A})$.
-    \item If $|\mathcal{A}| = \omega$ and $\exists$ has a winning strategy in $\mathcal{G}_{\omega}(\mathcal{A})$ then $\mathcal{A}$ is representable.
+    \item If $\mathcal{A}$ is infinite and $\exists$ has a winning strategy in $\mathcal{G}_{\omega}(\mathcal{A})$ then $\mathcal{A}$ is representable.
+    \item If $\mathcal{A}$ is infinite and $\forall$ has a winning strategy in $\mathcal{G}_{\omega}(\mathcal{A})$ then $\mathcal{A}$ is not representable.
   \end{enumerate}
@@
     \item Let $h$ be a representation of $\mathcal{A}$ over some domain $D$.
-$\forall$ plays $(a, b)$ with $a \not\leq b$ for the initial round.
-Let $N = \langle E, V, l \rangle$ be a network being played.
+$\forall$ picks $(a, b)$ with $a \not\leq b$ for the initial round.
+Let $\mathcal{N} = \langle E, V, l \rangle$ be a network being played.
+To win the game $\exists$ maintains a map $\iota : E \to D$ such that $(\iota(x_0), \iota(x_1)) \in h(b) \setminus h(a)$ and for each $x, y \in E$ and
+    $a \in l_{\mathcal{N}}(x, y)$ one has $(\iota(x), \iota(y)) \in h(a)$. The argument repeats the proof of the first item of \cite[Proposition 5]{hirsch2005class}.
 
-    TODO: define an embedding $\iota : E \to D$ and check the presence of a winning strategy explicitly
+    \item
+    Firstly, let us suppose that $|\mathcal{A}| = \omega$.
 
-    \item Let $a, b \in \mathcal{A}$ such that $a \not\leq b$.
-We consider a play of the game where $\forall$ pick the pair $(a, b)$ in the initial round and plays $(\mathcal{N},x, y, z, c, d)$ in the further rounds for each $x, y, z \in \operatorname{Nodes}(\mathcal{N})$ and $c,d \in \mathcal{A}$.
+    Let $a, b \in \mathcal{A}$ such that $a \not\leq b$.
+We consider a play of the game where $\forall$ pick the pair $(a, b)$ in the initial round and plays $(\mathcal{N}, x, y, z, c, d)$ in the further rounds for each $x, y, z \in \operatorname{Nodes}(\mathcal{N})$ and $c,d \in \mathcal{A}$.
 Here, $c \in l_N(x, y)$ and $d \in l_N(y, z)$.
@@
 We need $(x, y) \in rep(a;b)$.
-That is, there exists $z \in \mathcal{Nodes}(\mathcal{N})$ with $(x, z) \in rep(a)$ and $(z, y) \in rep(b)$.
+That is, there exists $z \in \operatorname{Nodes}(\mathcal{N})$ with $(x, z) \in rep(a)$ and $(z, y) \in rep(b)$.
 So, there are $c, d$ such that $c \leq a$ with $c \in l_{\mathcal{N}}(x, z)$ and $d \leq b$ with $d \in l_{\mathcal{N}}(z, y)$.
@@
 
-\item Injectivity: TODO.
+\item Injectivity: TODO
+
 \end{itemize}
+Now let us assume that $|\mathcal{A}| \geq \aleph_1$. By elementary chain theorem (see \cite[Theorem 2.5.2]{hodges1997shorter})
+
+\item
   \end{enumerate}
 \end{proof}
+
+\begin{proposition}
+  Let $\mathcal{A}$ be a join semilattice-ordered semigroup, then there exists a game such that $\mathcal{A}$ is representable iff $\exists$ has a winning strategy in the all finite lenght games.
+\end{proposition}
+
+\begin{proof}
+$ $
+
+  \begin{enumerate}
+    \item $(\Rightarrow)$
+    \item $(\Leftarrow)$
+  \end{enumerate}
+\end{proof}
+
+Now we are seeking to obtain the explicit axiomatisation of ${\bf R}(;, +)$.
+The idea is to axiomatise this class with the additional axioms claiming that $\exists$ has a winning strategy in certain games.
+See \cite[Chapter 9]{hirsch2002relation} for the discussion in detail to have a more general methodology.
+Our purpose is to provide the recursively enumerable axiomatisation of ${\bf R}(;, +)$.
+Thus, according to famous Craig's trick \cite{craig1953axiomatizability}, ${\bf R}(;, +)$ will have the decidable axiomatisation.
+
+\begin{definition}
+
+\end{definition}
 
@@
 
-\begin{lemma}
+\begin{lemma}\label{ax}
   For each $n < \omega$ there exists a first-order sentence $\sigma_n$ such that $\exists$ has a winning strategy in $\mathcal{G}_n(\mathcal{A})$ iff $\mathcal{A} \models \sigma_n$.
@@
 \begin{proof}
+  As usual, for each $n < \omega$ a formula $\sigma_n$ claims that $\exists$ has a winning strategy in the game of lenght $n$.
 
+  Let $\mathcal{N}$ be a term network.
+
+  ...
+
+  We define the sequence of formulas $\{ \sigma_n \}_{n < \omega}$ recursively:
+  \begin{enumerate}
+    \item $\sigma_0(\mathcal{N}, v) = \bigwedge \limits_{a \in l_{\mathcal{N}}(x, y)} \neg (a \leq v)$
+
+    \item $\sigma_{n + 1} = \sigma_{n} \land $
+  \end{enumerate}
 \end{proof}
@@
 \begin{theorem}
-  ${\bf R}(+, ;)$ is axiomatised with the axioms of join semilattice-ordered semigroups and $\{ \sigma_n \}_{n < \omega}$.
+  ${\bf R}(+, ;)$ is axiomatised with the axioms of join semilattice-ordered semigroups and $\{ \sigma_n \}_{n < \omega}$ defined in the proof of Lemma~\ref{ax}.
 \end{theorem}
+
+\subsection{The finite representation property for ${\bf R}(+, ;)$}
 
@@
 \end{proof}
+
+The following corollary follows from finite representability and finite axiomatisability.
+
+\begin{corollary}
+  The class ${\bf R}_{fin}(;, +) = \{ \mathcal{A} \: | \: |\mathcal{A}| < \omega \}$ is recursive. That is, the problem of membership for this class is decidable.
+\end{corollary}
 

### 2021-01-25 frp-paper/Paper.tex
wip

@@
 
-\subsection{Axiomatisating ${\bf R}(+, ;)$ and the finite representation property}
+\subsection{Axiomatisating ${\bf R}(+, ;)$}
 
@@
 \end{itemize}
-Now let us assume that $|\mathcal{A}| \geq \aleph_1$. By elementary chain theorem (see \cite[Theorem 2.5.2]{hodges1997shorter})
+Now let us assume that $|\mathcal{A}| \geq \aleph_1$. By Tarksi-Vaught theorem on elementary chain union (see \cite[Theorem 2.5.2]{hodges1997shorter})
 
@@
 
+The following proposition is a version of \cite[Proposition 7.24]{hirsch2002relation} and the right-to-left part is proved using K\"{o}nig's lemma \cite[Exercise 5.6.3]{hodges1997shorter}.
+
 \begin{proposition}
@@
 
-\begin{proof}
-$ $
-
-  \begin{enumerate}
-    \item $(\Rightarrow)$
-    \item $(\Leftarrow)$
-  \end{enumerate}
-\end{proof}
-
 Now we are seeking to obtain the explicit axiomatisation of ${\bf R}(;, +)$.
-The idea is to axiomatise this class with the additional axioms claiming that $\exists$ has a winning strategy in certain games.
 See \cite[Chapter 9]{hirsch2002relation} for the discussion in detail to have a more general methodology.
@@
 Thus, according to famous Craig's trick \cite{craig1953axiomatizability}, ${\bf R}(;, +)$ will have the decidable axiomatisation.
-
-\begin{definition}
-
-\end{definition}
+According to the proposition above, it is sufficient to extract the axioms claiming that $\exists$ has a winning strategy in all finitary games.
 
@@
 Let $\mathcal{N} = \langle E, V, l \rangle$ be a term prenetwork, $\mathcal{A}$ be a join-semilattice ordered semigroup and $\vartheta : V \to \underline{{\mathcal{A}}}$ a valuation.
-Let us define the prenetwork $\mathcal{N}^{\vartheta}$ with the same edges and vertices with the labelling $l^{\vartheta}(x, y) = \{ t^{\vartheta} \: | \: t \in l(x, y)\}$.
+Let us define the prenetwork $\mathcal{N}^{\vartheta}$ with the same edges and vertices with labelling $l^{\vartheta}(x, y) = \{ t^{\vartheta} \: | \: t \in l_{\mathcal{N}}(x, y)\}$.
 
@@
 
-  Let $\mathcal{N}$ be a term network.
-
+  Let $\mathcal{N}$ be a term network. We define the following three extensions of $\mathcal{N}$ that reflect the composition, witness, and join moves respectively:
+  \begin{enumerate}
+    \item $\mathcal{N}_c = $
+    \item $\mathcal{N}_w = $
+    \item $\mathcal{N}_j = $
+  \end{enumerate}
   ...
@@
 
-    \item $\sigma_{n + 1} = \sigma_{n} \land $
+This formula merely claims that $\exists$ has a winning strategy in the zero length game.
+
+    \item $\sigma_{n + 1}(\mathcal{N}, v) = \\ \sigma_{n}(\mathcal{N}_c, ..) \land \sigma_{n}(\mathcal{N}_w, ..) \land \sigma_{n}(\mathcal{N}_j, ..)$
   \end{enumerate}
@@
   ${\bf R}(+, ;)$ is axiomatised with the axioms of join semilattice-ordered semigroups and $\{ \sigma_n \}_{n < \omega}$ defined in the proof of Lemma~\ref{ax}.
+Moreover, ${\bf R}(+, ;)$ has the decidable axiomatisation.
 \end{theorem}
@@
 \begin{corollary}
-  The class ${\bf R}_{fin}(;, +) = \{ \mathcal{A} \: | \: |\mathcal{A}| < \omega \}$ is recursive. That is, the problem of membership for this class is decidable.
+  The class ${\bf R}_{fin}(;, +) = \{ \mathcal{A} \in {\bf R}(;, +) \: | \: |\mathcal{A}| < \omega \}$ is recursive. That is, the problem of membership for this class is decidable.
 \end{corollary}
+
+\begin{proof}
+  Let $\mathcal{A}$ be a finite semilattice-ordered semigroup. We check axioms with the algorithm that yields ``no'', if $\mathcal{A}$ is not representable. So,
+  In parallel, we check whether $\mathcal{A}$ has a finite representation.
+If there exists such $n < \omega$ such that $\mathcal{A}$ is representable over size $n$.
+Both sets are domains of corresponding computable functions, so they are enumerable.
+That makes ${\bf R}_{fin}(;, +)$ decidable by Post theorem.
+\end{proof}
 

### 2021-01-30 filtration-stuff/Text.tex
diagram sketch

@@
 
-\begin{theorem}
+\begin{theorem}\label{sad}
   ${\bf K}5$ does not admit strict filtrations.
@@
 
+\section{``Finite'' canonical models}
+
+Let $\mathcal{L}$ be a normal modal logic, $\mathcal{M}_{\mathcal{L}}$ its canonical model, and $\varphi$.
+Let us put $\Gamma = \operatorname{Sub}(\varphi) \cup \{ \neg \psi \: | \: \psi \in \operatorname{Sub}(\varphi) \}$.
+
+A subset $\Delta \subseteq \Gamma$ is a \emph{finite $\mathcal{L}$-theory} if $\bigwedge \Delta \notin \mathcal{L}$. A subset $\Delta$ is maximal, if (the following are obviouslyequivalent):
+\begin{enumerate}
+  \item $\Delta$ is maximal amongst finite $\mathcal{L}$-theories,
+  \item For each $\psi \in \operatorname{Sub}(\varphi)$ either $\psi \in \Delta$ or
+  $\neg \psi \in \Delta$.
+\end{enumerate}
+
+Every finite $\mathcal{L}$-theory is clearly can be extended to some maximal one. It is the finite version of Lindenbaum's lemma.
+
+\begin{defin}
+  Let $\mathcal{L}$ be a normal modal logic and $\varphi \not\in \mathcal{L}$.
+A ``finite'' canonical model is a triple $\mathcal{M}_{\mathcal{L}}^{\varphi} = \langle W_{\mathcal{L}}^{\varphi}, R_{\mathcal{L}}^{\varphi}, \vartheta_{\mathcal{L}}^{\varphi} \rangle$, where
+  \begin{enumerate}
+    \item $W_{\mathcal{L}}^{\varphi}$ is the set all maximal theories that extend finite $\mathcal{L}$-theories
+    \item $\Delta_1 R \Delta_2$ iff $\Diamond \psi \in \Delta_2 \Rightarrow \psi \in \Delta_1$
+    \item $\vartheta_{\mathcal{L}}^{\varphi}(p) = \{ \Delta \in W_{\mathcal{L}}^{\varphi} \: | \: p \in \Delta \}$.
+  \end{enumerate}
+\end{defin}
+
+\begin{defin} Let $\varphi$ be a modal formula, $\mathcal{L}$ a logic that admits strict filtration, and $\mathcal{M}$ an $\mathcal{L}$-model that refutes $\varphi$.
+A \emph{filtration diagram} of a model $\widehat{\mathcal{M}}$ (a filtration of $\mathcal{M}$ through $\varphi$) is a formula $\operatorname{FDiagram}(\mathcal{M}_{\mathcal{L}}^{\operatorname{Sub}(\varphi)}) = \bigwedge \Gamma \wedge \bigwedge \Delta$, where
+  \begin{itemize}
+    \item $\Gamma =
+    \{ \psi \in \operatorname{Sub}(\varphi) \: | \: \exists w \in \underline{\mathcal{M}} \: \widehat{M}, \hat{w} \models \psi \}$
+    \item $\Delta = \{ \neg \psi \in \operatorname{Sub}(\varphi) \: | \: \psi \in \operatorname{Sub}(\varphi) \: \& \: \mathcal{M} \not\models \psi \}$
+  \end{itemize}
+\end{defin}
+
+\begin{lemma}
+  Let $\varphi$ be a modal formula, $\mathcal{L}$ a logic that admits strict filtration, and $\mathcal{M}$ an $\mathcal{L}$-model that refutes $\varphi$.
+Then $\mathcal{M}_{\mathcal{L}}^{\operatorname{Sub}(\varphi)} \models \operatorname{FDiagram}(\mathcal{M}_{\mathcal{L}}^{\operatorname{Sub}(\varphi)})$
+\end{lemma}
+
+\begin{defin}
+  Let $\mathcal{M}$ be a model and $\Psi$ a set of formulas, then a $\Psi$-theory of a model, denoted as $\operatorname{Th}_{\Psi}$, is the set $\operatorname{Th}_{\Psi} = \{ \varphi \in \Psi \: | \: \mathcal{M} \models \varphi \}$.
+\end{defin}
+
+\begin{lemma}
+  Let $\mathcal{L}$ be a modal logic that admits strict filtrations and $\varphi \not\in \mathcal{L}$.
+  Let $\mathcal{M}$ be a model that refutes $\varphi$ and $\mathcal{N}$ be a different model such that $\mathcal{N} \models \mathcal{L}$.
+If $\mathcal{N} \models \operatorname{FDiagram}(\mathcal{M}_{\mathcal{L}}^{\operatorname{Sub}(\varphi)})$, then $\operatorname{Th}_{\operatorname{Sub}(\varphi)}(\mathcal{M}_{\mathcal{L}}^{\operatorname{Sub}(\varphi)}) = \operatorname{Th}_{\operatorname{Sub}(\varphi)}(\mathcal{N})$.
+In particular, $\mathcal{M}$ refutes $\varphi$.
+\end{lemma}
+
+\begin{theorem} Let $\mathcal{L}$ be a complete normal modal logic, then the following are equivalent:
+  \begin{enumerate}
+    \item $\mathcal{L}$ admits strict filtrations
+    \item If $\varphi \not\in \mathcal{L}$, then the ``finite'' canonical model $\mathcal{M}_{\mathcal{L}}^{\varphi}$ refutes $\varphi$
+  \end{enumerate}
+\end{theorem}
+
+\begin{proof}
+  $ $
+
+  \begin{enumerate}
+    \item $(\Rightarrow)$
+
+    Let $\varphi \not\in \mathcal{L}$, there exists a model $\mathcal{M}$ that refutes $\varphi$, so there exists $w \in \underline{\mathcal{M}}$ such that $\mathcal{M}, w \not\models \varphi$.
+But $\mathcal{L}$ admits strict filtration, so there exists a model $\widehat{\mathcal{M}} = \langle \widehat{W}, \widehat{R}, \widehat{\vartheta} \rangle$, where $\widehat{W} = W / \sim_{\varphi}$ and $\hat{R}^{min} \subseteq \widehat{R} \subseteq \hat{R}^{max}$.
+In particular, $\widehat{\mathcal{M}}, \widehat{w} \not\models \varphi$.
+
+    By lemma above, $\widehat{\mathcal{M}} \models \operatorname{FDiagram}(\widehat{\mathcal{M}})$
+
+    Consider a ``finite'' canonical model
+    $\mathcal{M}_{\mathcal{L}}^{\operatorname{Sub}(\varphi)} = \langle W_{\mathcal{L}}^{\operatorname{Sub}(\varphi)}, R_{\mathcal{L}}^{\operatorname{Sub}(\varphi)}, \vartheta_{\mathcal{L}}^{\operatorname{Sub}(\varphi)} \rangle$.
+    Let us show that $\mathcal{M}_{\mathcal{L}}^{\operatorname{Sub}}$ refutes $\varphi$.
+For that, we show that $\mathcal{M}_{\mathcal{L}}^{\operatorname{Sub}(\varphi)} \models \operatorname{FDiagram}(\widehat{\mathcal{M}})$.
+
+    \item $(\Leftarrow)$
+
+    Let $\mathcal{M}_{\mathcal{L}}^{\varphi} = $ be a ``finite'' canonical model that refutes $\varphi$.
+
+    Consider $W_{\mathcal{L}}^{\varphi} / \sim_{\operatorname{Sub}(\varphi)}$. Clearly $W_{\mathcal{L}}^{\varphi} = W_{\mathcal{L}}^{\varphi} / \sim_{\operatorname{Sub}(\varphi)}$ since
+
+    \begin{center}
+      $\Delta_1 \sim_{\operatorname{Sub}(\varphi)} \Delta_2$ iff $\forall \psi \in \operatorname{Sub}(\varphi) \:\: W_{\mathcal{L}}^{\varphi}, \Delta_1 \models \psi \Leftrightarrow W_{\mathcal{L}}^{\varphi}, \Delta_2 \models \psi$ iff $\Delta_1 = \Delta_2$.
+    \end{center}
+
+    $R_{\mathcal{L}}^{\varphi}$ also trivially satisfies the required defined.
+Therefore, $\mathcal{L}$ admits strict filtrations since $\mathcal{M}_{\mathcal{L}}^{\varphi}$ is already finite.
+In particular, ${\bf K}5$ does not have the ``finite'' canonical property that follows from the contraposition of this statement and Theorem~\ref{sad}.
+
+  \end{enumerate}
+\end{proof}
+
 \bibliographystyle{plain}

### 2021-02-10 filtration-stuff/Text.tex
fix finite 'canonical' stuff

@@
 
-\section{``Finite'' canonical models}
+\section{Finite ``canonical`` models}
 
@@
   Let $\mathcal{L}$ be a normal modal logic and $\varphi \not\in \mathcal{L}$.
-A ``finite'' canonical model is a triple $\mathcal{M}_{\mathcal{L}}^{\varphi} = \langle W_{\mathcal{L}}^{\varphi}, R_{\mathcal{L}}^{\varphi}, \vartheta_{\mathcal{L}}^{\varphi} \rangle$, where
+A finite ``canonical'' model is a triple $\mathcal{M}_{\mathcal{L}}^{\varphi} = \langle W_{\mathcal{L}}^{\varphi}, R_{\mathcal{L}}^{\varphi}, \vartheta_{\mathcal{L}}^{\varphi} \rangle$, where
   \begin{enumerate}
     \item $W_{\mathcal{L}}^{\varphi}$ is the set all maximal theories that extend finite $\mathcal{L}$-theories
-    \item $\Delta_1 R \Delta_2$ iff $\Diamond \psi \in \Delta_2 \Rightarrow \psi \in \Delta_1$
+    \item $R_{\mathcal{L}}^{\varphi}$ is a relation such that
+    $\langle W_{\mathcal{L}}^{\varphi}, R_{\mathcal{L}}^{\varphi} \rangle$ is an $\mathcal{L}$-frame
+    and
+    \begin{center}
+    $\forall \Box \psi \operatorname{Sub}(\varphi) \:\:
+    \forall \Delta_1 \in W_{\mathcal{L}}^{\varphi} \:\: (\Box \psi \in \Delta_1 \Leftrightarrow
+    \forall \Delta_2 \in R_{\mathcal{L}}^{\varphi}(\Delta_1) \:\: \psi \in \Delta_2)$
+    \end{center}
     \item $\vartheta_{\mathcal{L}}^{\varphi}(p) = \{ \Delta \in W_{\mathcal{L}}^{\varphi} \: | \: p \in \Delta \}$.
@@
     \item $\mathcal{L}$ admits strict filtrations
-    \item If $\varphi \not\in \mathcal{L}$, then the ``finite'' canonical model $\mathcal{M}_{\mathcal{L}}^{\varphi}$ refutes $\varphi$
+    \item If $\varphi \not\in \mathcal{L}$, then the finite ``canonical'' model $\mathcal{M}_{\mathcal{L}}^{\varphi}$ refutes $\varphi$
   \end{enumerate}
@@
 
-    Consider a ``finite'' canonical model
+    Consider a finite ``canonical'' model
     $\mathcal{M}_{\mathcal{L}}^{\operatorname{Sub}(\varphi)} = \langle W_{\mathcal{L}}^{\operatorname{Sub}(\varphi)}, R_{\mathcal{L}}^{\operatorname{Sub}(\varphi)}, \vartheta_{\mathcal{L}}^{\operatorname{Sub}(\varphi)} \rangle$.
@@
 
-    Let $\mathcal{M}_{\mathcal{L}}^{\varphi} = $ be a ``finite'' canonical model that refutes $\varphi$.
+    Let $\mathcal{M}_{\mathcal{L}}^{\varphi}$ be a finite ``canonical'' model that refutes $\varphi$.
 
@@
 
-    $R_{\mathcal{L}}^{\varphi}$ also trivially satisfies the required defined.
+    $R_{\mathcal{L}}^{\varphi}$ also satisfies the required definition.
+    Suppose $\Box \psi \in \varphi$ and
+    $\mathcal{M}_{\mathcal{L}}^{\varphi}, \Delta_1 \models \Box \psi$ and
+    $\Delta_1 R_{\mathcal{L}}^{\varphi} \Delta_2$. Clearly that $\psi \in \Delta_2$ by the definition
+    of a finite ``canonical'' model.
+
     Therefore, $\mathcal{L}$ admits strict filtrations since $\mathcal{M}_{\mathcal{L}}^{\varphi}$ is already finite.
-In particular, ${\bf K}5$ does not have the ``finite'' canonical property that follows from the contraposition of this statement and Theorem~\ref{sad}.
+In particular, ${\bf K}5$ fails to have the ``canonical'' model property that follows from the contraposition of this statement and Theorem~\ref{sad}.
 

### 2021-02-11 filtration-stuff/Text.tex
fix

@@
     \begin{center}
-    $\forall \Box \psi \operatorname{Sub}(\varphi) \:\:
+    $\forall \Box \psi \in \operatorname{Sub}(\varphi) \:\:
     \forall \Delta_1 \in W_{\mathcal{L}}^{\varphi} \:\: (\Box \psi \in \Delta_1 \Leftrightarrow
@@
     \item $\mathcal{L}$ admits strict filtrations
-    \item If $\varphi \not\in \mathcal{L}$, then the finite ``canonical'' model $\mathcal{M}_{\mathcal{L}}^{\varphi}$ refutes $\varphi$
+    \item If $\varphi \not\in \mathcal{L}$, there exists a finite ``canonical'' model $\mathcal{M}_{\mathcal{L}}^{\varphi}$ such that
+    $\mathcal{M}_{\mathcal{L}}^{\varphi} \models \mathcal{L}$ and
+    $\mathcal{M}_{\mathcal{L}}^{\varphi}$ refutes $\varphi$.
   \end{enumerate}

### 2021-02-11 filtration-stuff/Text.tex
another fix
fix

@@
 If $\mathcal{N} \models \operatorname{FDiagram}(\mathcal{M}_{\mathcal{L}}^{\operatorname{Sub}(\varphi)})$, then $\operatorname{Th}_{\operatorname{Sub}(\varphi)}(\mathcal{M}_{\mathcal{L}}^{\operatorname{Sub}(\varphi)}) = \operatorname{Th}_{\operatorname{Sub}(\varphi)}(\mathcal{N})$.
-In particular, $\mathcal{M}$ refutes $\varphi$.
+In particular, $\mathcal{N}$ refutes $\varphi$.
 \end{lemma}
@@
     Therefore, $\mathcal{L}$ admits strict filtrations since $\mathcal{M}_{\mathcal{L}}^{\varphi}$ is already finite.
-In particular, ${\bf K}5$ fails to have the ``canonical'' model property that follows from the contraposition of this statement and Theorem~\ref{sad}.
+In particular, ${\bf K}5$ fails to have the finite ``canonical'' model property that follows from the contraposition of this statement and Theorem~\ref{sad}.
 

### 2021-02-13 filtration-stuff/Text.tex
wip

@@
 \begin{defin}
-  An $n$-normal modal logic is a set of formulas that contains all Boolean tautologies, formulas $\Diamond_i p \lor \Diamond_i q \leftrightarrow \Diamond_i (p \lor q)$ and $\Diamond_i \bot \leftrightarrow \bot$ for $i \leq n$, and is closed under modus ponens, substitution, and monotonicity: from $\varphi \rightarrow \psi$ infer $\Diamond_i \varphi \rightarrow \Diamond_i \psi$ for $i \leq n$.
+  A normal modal logic is a set of formulas that contains all Boolean tautologies,
+  formulas $\Diamond p \lor \Diamond q \leftrightarrow \Diamond (p \lor q)$ and
+  $\Diamond \bot \leftrightarrow \bot$, and is closed under Modus Ponens, Substitution, and
+  Monotonicity:
+  from $\varphi \rightarrow \psi$ infer $\Diamond \varphi \rightarrow \Diamond \psi$.
 \end{defin}
 
-\begin{defin} An $n$-Kripke model is a triple $\mathcal{M} = \langle W, R_1, \dots, R_n, \vartheta \rangle$, where $R_i \subseteq W \times W$, $\vartheta : \operatorname{PV} \to 2^W$, and the connectives have the following semantics:
+\begin{defin} An Kripke model is a triple $\mathcal{M} = \langle W, R, \vartheta \rangle$,
+  where $R \subseteq W \times W$, $\vartheta : \operatorname{PV} \to 2^W$, and
+  the connectives have the following semantics:
 
@@
     \item $\mathcal{M}, w \models p \Leftrightarrow w \in \vartheta(p)$
-    \item $\mathcal{M}, w \models \not \varphi \Leftrightarrow \mathcal{M}, w \nvDash \varphi$
+    \item $\mathcal{M}, w \models \neg \varphi \Leftrightarrow \mathcal{M}, w \nvDash \varphi$
     \item $\mathcal{M}, w \models \varphi \lor \psi \Leftrightarrow \mathcal{M}, w \models \varphi \text{ or } \mathcal{M}, w \models \psi$
-    \item $\mathcal{M}, w \models \Diamond_i \varphi \Leftrightarrow \exists v \in R_i(w) \: \mathcal{M}, v \models \varphi$
+    \item $\mathcal{M}, w \models \Diamond \varphi \Leftrightarrow \exists v \in R(w) \: \mathcal{M}, v \models \varphi$
   \end{enumerate}
 \end{defin}
+
+\begin{defin}
+  Let $\mathcal{M} = \langle W, R, \vartheta \rangle$ be a Kripke model.
+  A Kripke model $\mathcal{M}' = \langle W', R', \vartheta' \rangle$ is a generated submodel of
+  $\mathcal{M}$, where:
+  \begin{enumerate}
+    \item $\emptyset \neq W' \subseteq W$ and $W'$ is $R$-closed, i.e., for each $u \in W'$ and $v \in W$,
+    $u R v$ implies $v \in W'$
+    \item $R' = R \cap W' \times W'$
+    \item for every propositional variable $p$, $\vartheta(p)' = \vartheta(p) \cap W'$
+  \end{enumerate}
+\end{defin}
+
+\begin{fact}
+  Let $\mathcal{M} = \langle W, R, \vartheta \rangle$ be a Kripke model and
+  $\mathcal{M}' = \langle W', R', \vartheta' \rangle$ its generated submodel,
+  then for each $w \in W'$
+  \begin{center}
+    $\mathcal{M}, w \models \varphi$ iff $\mathcal{M}', w \models \varphi$
+  \end{center}
+\end{fact}
 
@@
 
-\section{Finite ``canonical`` models}
+\section{Finite ``canonical'' models}
 
@@
     \{ \psi \in \operatorname{Sub}(\varphi) \: | \: \exists w \in \underline{\mathcal{M}} \: \widehat{M}, \hat{w} \models \psi \}$
-    \item $\Delta = \{ \neg \psi \in \operatorname{Sub}(\varphi) \: | \: \psi \in \operatorname{Sub}(\varphi) \: \& \: \mathcal{M} \not\models \psi \}$
+    \item $\Delta = \{ \neg \psi \in \operatorname{Sub}(\varphi) \: | \: \psi \in \operatorname{Sub}(\varphi) \: \& \: \mathcal{M} \nvDash \psi \}$
   \end{itemize}
@@
 
-\begin{theorem} Let $\mathcal{L}$ be a complete normal modal logic, then the following are equivalent:
+\begin{theorem} Let $\mathcal{L}$ be a canonical logic, then the following are equivalent:
   \begin{enumerate}
@@
 
-    Let $\varphi \not\in \mathcal{L}$, there exists a model $\mathcal{M}$ that refutes $\varphi$, so there exists $w \in \underline{\mathcal{M}}$ such that $\mathcal{M}, w \not\models \varphi$.
+    Let $\varphi \not\in \mathcal{L}$, there exists a model $\mathcal{M}$ that refutes $\varphi$, so there exists $w \in \underline{\mathcal{M}}$ such that $\mathcal{M}, w \nvDash \varphi$.
 But $\mathcal{L}$ admits strict filtration, so there exists a model $\widehat{\mathcal{M}} = \langle \widehat{W}, \widehat{R}, \widehat{\vartheta} \rangle$, where $\widehat{W} = W / \sim_{\varphi}$ and $\hat{R}^{min} \subseteq \widehat{R} \subseteq \hat{R}^{max}$.
-In particular, $\widehat{\mathcal{M}}, \widehat{w} \not\models \varphi$.
+In particular, $\widehat{\mathcal{M}}, \widehat{w} \nvDash \varphi$.
 
@@
 
-    Let $\mathcal{M}_{\mathcal{L}}^{\varphi}$ be a finite ``canonical'' model that refutes $\varphi$.
+    Let $\varphi \not\in \mathcal{L}$ and $\mathcal{M}_{\mathcal{L}}^{\varphi} = \langle W_{\mathcal{L}}^{\varphi}, R_{\mathcal{L}}^{\varphi}, \vartheta_{\mathcal{L}}^{\varphi} \rangle$
+    be a finite ``canonical'' model that refutes $\varphi$.
+    Let $\mathcal{M}$ be an $\mathcal{L}$-model that also refutes $\varphi$.
+    Let us show that there exists a model $\widehat{\mathcal{M}}$ such that
+    $\widehat{\mathcal{M}}$ is a
+    strict filtration of $\mathcal{M}$ through $\operatorname{Sub}(\varphi)$.
 
-    Consider $W_{\mathcal{L}}^{\varphi} / \sim_{\operatorname{Sub}(\varphi)}$. Clearly $W_{\mathcal{L}}^{\varphi} = W_{\mathcal{L}}^{\varphi} / \sim_{\operatorname{Sub}(\varphi)}$ since
+    Let $\sim_{\operatorname{Sub}(\varphi)}$ be an equivalence relation on $\underline{\mathcal{M}}$ defined as usual:
 
     \begin{center}
-      $\Delta_1 \sim_{\operatorname{Sub}(\varphi)} \Delta_2$ iff $\forall \psi \in \operatorname{Sub}(\varphi) \:\: W_{\mathcal{L}}^{\varphi}, \Delta_1 \models \psi \Leftrightarrow W_{\mathcal{L}}^{\varphi}, \Delta_2 \models \psi$ iff $\Delta_1 = \Delta_2$.
+      $u \sim_{\operatorname{Sub}(\varphi)} v$ iff $\forall \psi \in \operatorname{Sub}(\varphi)$
+      $\mathcal{M}, u \models \psi \Leftrightarrow \mathcal{M}, v \models \psi$
     \end{center}
 
-    $R_{\mathcal{L}}^{\varphi}$ also satisfies the required definition.
-    Suppose $\Box \psi \in \varphi$ and
-    $\mathcal{M}_{\mathcal{L}}^{\varphi}, \Delta_1 \models \Box \psi$ and
-    $\Delta_1 R_{\mathcal{L}}^{\varphi} \Delta_2$. Clearly that $\psi \in \Delta_2$ by the definition
-    of a finite ``canonical'' model.
+    Let $[u] \in W/\sim_{\operatorname{Sub}(\varphi)}$, where $W = \underline{\mathcal{M}}$,
+    consider its truth set $||[u]|| = \{ \psi \in \operatorname{Sub}(\varphi) \: | \: \mathcal{M}, u \models \psi \}$.
+    $||[u]||$ is clearly a finite consistent theory and it consists in some $\Delta \in W_{\mathcal{L}}^{\varphi}$ by the finite analogue of Lindenbaum's lemma.
+Such a $\Delta$ is clearly unique as well.
+    Consider a map
+    \begin{center}
+    $\iota : W/\sim_{\operatorname{Sub}(\varphi)} \to W_{\mathcal{L}}^{\varphi}$ with
+    $\iota : [u] \mapsto \Delta \supseteq ||[u]||$
+    \end{center}
 
-    Therefore, $\mathcal{L}$ admits strict filtrations since $\mathcal{M}_{\mathcal{L}}^{\varphi}$ is already finite.
-In particular, ${\bf K}5$ fails to have the finite ``canonical'' model property that follows from the contraposition of this statement and Theorem~\ref{sad}.
+    Consider the set $\iota(W/\sim_{\operatorname{Sub}(\varphi)})$ denoted as $\operatorname{Im}(\iota)$. Let us show that $\operatorname{Im}(\iota)$ is $R_{\mathcal{L}}^{\varphi}$-closed.
 
+    Let $\iota([u]) R_{\mathcal{L}}^{\varphi} \Delta$, let us show that $\Delta \in \iota(W/\sim_{\operatorname{Sub}(\varphi)})$, that is, $\Delta = \iota([v])$ for some $v \in W$.
+
+    Consider $\Delta$ as a finite ultrafilter.
+Thus, $\Delta$ is principal and it 's generated by $\psi_1 \land \dots \land \psi_n$, where each $\psi_i \in \operatorname{Sub}(\varphi)$.
+So $\Diamond (\psi_1 \land \dots \land \psi_n) \in \iota([u])$.
+    Then $\mathcal{M}, u \models \Diamond (\psi_1 \land \dots \land \psi_n)$. That implies $\mathcal{M}, v \models \psi_1 \land \dots \land \psi_n$ for some $v \in R(u)$.
+    This is clearly a desired $v$ since $||[v]|| \subseteq \Delta$ and such a $\Delta$ is unique.
+
+    Thus, $\langle \operatorname{Im}(\iota), R_{\mathcal{L}}^{\varphi} \cap \operatorname{Im}(\iota) \times \operatorname{Im}(\iota), {\vartheta'}_{\mathcal{L}}^{\varphi} \rangle$, where ${\vartheta'}_{\mathcal{L}}^{\varphi}(p) = \vartheta_{\mathcal{L}}^{\varphi}(p) \cap \operatorname{Im}(\iota)$ is a generated submodel of $\mathcal{M}_{\mathcal{L}}^{\varphi}$.
+
+    Consider a model $\widehat{\mathcal{M}} = \langle W/\sim_{\operatorname{Sub}(\varphi)}, \widehat{R}, \widehat{\vartheta} \rangle$, where $\widehat{R}$ is defined as
+    \begin{center}
+      $[u] \widehat{R} [v]$ iff $||[u]|| {R'}_{\mathcal{L}}^{\varphi} ||[v]||$, where ${R'}_{\mathcal{L}}^{\varphi} R_{\mathcal{L}}^{\varphi} \cap \operatorname{Im}(\iota) \times \operatorname{Im}(\iota)$.
+    \end{center}
+    Let us prove that $\widehat{\mathcal{M}}$ is a filtration of $\mathcal{M}$ through $\operatorname{Sub}(\varphi)$.
+
+    Let us show that $R_{min} \subseteq \widehat{R} \subseteq R_{max}$.
+
+    Suppose that there are $u' \in [u]$ and $v' \in [v]$ with $u' R v'$. Let us show that
+    $[u] \widehat{R} [v]$. That is, let us show that $\Diamond \iota(v) \subseteq \iota(v)$.
+    But if $\mathcal{M}, v' \models \psi$ for $\Diamond \psi \in \operatorname{Sub}(\varphi)$, then
+    $\mathcal{M}, u' \models \Diamond \psi$. Recall that $u' \sim_{\operatorname{Sub}(\varphi)} u$ and $v' \sim_{\operatorname{Sub}(\varphi)} v$, so we have $\Diamond ||[v]|| \subseteq ||[u]||$.
+    The same inclusion clearly holds for the maximal extensions of $\Diamond ||[v]||$ and $||[u]||$.
+
+    The second inclusion follows from the definition of a relation in a finite ``canonical'' model.
   \end{enumerate}
 \end{proof}
+
+In particular, ${\bf K}5$ fails to have the finite ``canonical'' model property that
+follows from the contraposition of the theorem above and Theorem~\ref{sad}.
 

### 2021-02-20 frp-paper/Paper.tex
the axiomatisation of R(;, +)

@@
 
-The logic of such structures is the Lambek calculus \cite{lambek1958mathematics} allowing one to characterise inference in categorial grammars, the equivalent version of context-free grammars \cite{pentus1993lambek}.
-One may define the Lambek calculus as the following (cut-free) sequent calculus.
-
 The class ${\bf R}(;, \setminus, /, \leq)$ consists of the following structures:
@@
 See this paper to learn more about step-by-step representations in general \cite{hirsch1997step}.
-That also implies relational completeness of the Lambek calculus, the logic of ${\bf RS}$.
+As it well known, the logic of such structures is the Lambek calculus \cite{lambek1958mathematics}, so we also have the relational completeness of the Lambek calculus.
 
@@
         \item $\exists$ chooses $a$ and responds with the prenetwork $\mathcal{N}_{n + 1} = \langle E_n, V_{n}, l_{n + 1} \rangle$, where $l_{n+1}(x, y) = \uparrow (l_n(x,y) \cup \{ a \})$.
-        \item $\exists$ chooses $b$. The response is similar but $l_{n + 1}(x, y) = \uparrow (l_n(x,y) \cup \{ b \}$.
+        \item $\exists$ chooses $b$. The response is similar but $l_{n + 1}(x, y) = \uparrow (l_n(x,y) \cup \{ b \})$.
       \end{enumerate}
@@
 
+Let us define the equivalence relation $\approx$ on $T$ as follows for $t_1$ and $t_2$ are terms:
+\begin{center}
+$t_1 \approx t_2$ iff $\theta(t_1) = \theta(t_2)$ for each interpretation of variables in $\theta : V \to F(\omega)$.
+\end{center}
+where $F(\omega)$ is the free countably generated algebra of the variety of all join semilattice-ordered group semigroup. We have the following partial order on the $T$ defined as
+\begin{center}
+  $t_1 \preceq t_2$ iff $t_1 + t_2 \approx t_2$, that is, $F(\omega) \models t_1 + t_2 = t_2$
+\end{center}
+A subset $A \subseteq T$ of terms is called upward closed if for every term $t \in A$ one has $\uparrow t \subseteq A$, where $\uparrow t = \{ t_1 \in T \: | \: t \preceq t_1 \}$.
+
 \begin{definition}
-  A term network is a graph $\langle E, V, l \rangle$, where $\langle E, V \rangle$ is a complete directed graph and $l : V \to 2^T$ is a labelling function.
+  A term network is a graph $\langle E, V, l \rangle$, where $\langle E, V \rangle$ is a complete directed graph and $l : V \to \operatorname{Up}(F(\omega), \preceq)$ is a labelling function.
 \end{definition}
@@
 
-Let $\mathcal{N} = \langle E, V, l \rangle$ be a term prenetwork, $\mathcal{A}$ be a join-semilattice ordered semigroup and $\vartheta : V \to \underline{{\mathcal{A}}}$ a valuation.
+Let $\mathcal{N} = \langle E, V, l \rangle$ a term network, $\mathcal{A}$ be a join-semilattice ordered semigroup and $\vartheta : V \to \underline{{\mathcal{A}}}$ a valuation.
 Let us define the prenetwork $\mathcal{N}^{\vartheta}$ with the same edges and vertices with labelling $l^{\vartheta}(x, y) = \{ t^{\vartheta} \: | \: t \in l_{\mathcal{N}}(x, y)\}$.
+
+Let us ensure that $\mathcal{N}^{\vartheta}$ is well-defined.
+Let us consider $\vartheta$ as a homomorphism $F(\omega) \to \mathcal{A}$.
+We would like to know whether ${\uparrow t^{\vartheta}} \subseteq l^{\vartheta}(x,y)$ for $t^{\vartheta} \in l^{\vartheta}(x,y)$.
+So $t \in l(x,y)$.
+Suppose $t \preceq t_1$, then $t_1$ clearly belongs to $l(x,y)$.
+So $l^{\vartheta}(x,y)$ is upward closed by the construction.
+
+Let $\mathcal{N}$ be a term network. We define the following three extensions of $\mathcal{N}$ reflecting the composition, witness, and join moves respectively:
+\begin{enumerate}
+  \item Let $x, y \in \operatorname{Nodes}(\mathcal{N})$ and $t$ a term. $\mathcal{N}_c$ is the extension of $\mathcal{N}$,
+  where $\operatorname{Nodes}(\mathcal{N}_c) = \operatorname{Nodes}(\mathcal{N})$ and $l_{\mathcal{N}}(x, y) \cup {\uparrow t}$ and $l_{\mathcal{N}_c}(u, v) = l_{\mathcal{N}}(u, v)$ for each $u \neq x$ and $v \neq y$.
+We denote this network as $\mathcal{N}_c(\mathcal{N}, x, y, t)$.
+  \item
+  Let $x, y \in \operatorname{Nodes}(\mathcal{N})$, $z$ a fresh node, and $t_1$, $t_2$ any terms.
+Let us define a network $T$ such that $\operatorname{Nodes}(T) = \{ x, y, z\}$.
+We define labelling as $l_{T}(x, y) = {\uparrow t_1}$ and $l_{T}(y, z) = {\uparrow t_2}$.
+So we put $\mathcal{N}_w = \mathcal{N} \cup T$.
+  We denote this network as $\mathcal{N}_w(\mathcal{N}, x,y,z, t_1, t_2)$.
+  \item Let $x, y \in \operatorname{Nodes}(\mathcal{N})$ and let $t_1, t_2$ be terms
+  Let us define $T_i = \langle \{ x, y\}, \{ (x, y) \}, l_{T_i} \rangle$, where $l_{T_i}(x, y) = \uparrow (l_{\mathcal{N}}(x, y) \cup t_i)$ for $i = 1,2$.
+So $\mathcal{N}_{j_1} = \mathcal{N} \cup T_1$ and $\mathcal{N}_{j_2} = \mathcal{N} \cup T_2$.
+  We denote this as $\mathcal{N}_j(\mathcal{N}, x, y, t_1, t_2)$.
+\end{enumerate}
+
+\newpage
 
@@
 \begin{proof}
-  As usual, for each $n < \omega$ a formula $\sigma_n$ claims that $\exists$ has a winning strategy in the game of lenght $n$.
-
-  Let $\mathcal{N}$ be a term network. We define the following three extensions of $\mathcal{N}$ that reflect the composition, witness, and join moves respectively:
-  \begin{enumerate}
-    \item $\mathcal{N}_c = $
-    \item $\mathcal{N}_w = $
-    \item $\mathcal{N}_j = $
-  \end{enumerate}
-  ...
+  As usual, for each $n < \omega$ we construct a formula $\sigma_n$ claiming that $\exists$ has a winning strategy in the game of lenght $n$. To be more precise, our purpose is to have
+  \begin{center}
+    $\exists$ has a winning strategy in $\mathcal{G}_n(\mathcal{N}^{\vartheta}, \mathcal{A}, \vartheta(v))$ if and only if $\mathcal{A} \models \sigma_{n}(\mathcal{N}, v)$
+  \end{center}
+  where $\mathcal{A}$ is a join semilattice-ordered semigroup and $\vartheta : V \to \mathcal{A}$ is a variable assignment.
 
@@
 
-    \item $\sigma_{n + 1}(\mathcal{N}, v) = \\ \sigma_{n}(\mathcal{N}_c, ..) \land \sigma_{n}(\mathcal{N}_w, ..) \land \sigma_{n}(\mathcal{N}_j, ..)$
+    \item Suppose $\sigma_{n}(\mathcal{N}, v)$ for $n < \omega$.
+Let us define a formula $\sigma_{n + 1}$ claiming that $\exists$ always has a proper response for a network $\mathcal{N}$ being played (which is defined by $\sigma_{n}(\mathcal{N}, v)$).
+
+    $\sigma_{n + 1}(\mathcal{N}, v)$ is defined as follows:
+    \begin{center}
+      $\sigma_{n + 1}(\mathcal{N}, v) = {\sigma_{n + 1}}_c(\mathcal{N}, v) \land {\sigma_{n + 1}}_w(\mathcal{N}, v) \land {\sigma_{n + 1}}_j(\mathcal{N}, v)$
+    \end{center}
+    where
+    \begin{itemize}
+      \item ${\sigma_{n + 1}}_c(\mathcal{N}, v) = \bigwedge \limits_{x, y, z \in \operatorname{Nodes}(\mathcal{N}), t_1 \in l_{\mathcal{N}}(x, y), t_2 \in l_{\mathcal{N}}(y, z)} \sigma_{n}(\mathcal{N}_c(x, z, t_1, t_2), v)$
+      \item ${\sigma_{n + 1}}_w(\mathcal{N}, v) = \bigwedge \limits_{x,y \in \operatorname{Nodes}(\mathcal{N}), t \in l_{\mathcal{N}}(x,y)} \forall u_1, u_2 (t \leq u_1 ; u_2 \rightarrow \bigvee \mathcal{N}_c(x,y,z, t_1, t_2))$, where $z$ is a fresh node.
+      \item ${\sigma_{n + 1}}_j(\mathcal{N}, v) = \bigwedge \limits_{x, y \in \operatorname{Nodes}(\mathcal{N})} \sigma_{n}(\mathcal{N}_{j_1}(\mathcal{N}, x, y, t_1, t_2), v) \lor \sigma_{n}(\mathcal{N}_{j_2}(\mathcal{N}, x, y, t_1, t_2), v)$
+    \end{itemize}
   \end{enumerate}
+
+  So, $\exists$ has a winning strategy iff these formulas are true under the valuation $\vartheta$ since the formulas $\{ \sigma_n \}_{n < \omega}$ encode the presence of a winning strategy for $\exists$ on every finite round.
 \end{proof}
 
+This inductive sequence of formulas provides us the explicit axiomatisation of the class of representable join semilattice-ordered semigroups.
+
 \begin{theorem}
-  ${\bf R}(+, ;)$ is axiomatised with the axioms of join semilattice-ordered semigroups and $\{ \sigma_n \}_{n < \omega}$ defined in the proof of Lemma~\ref{ax}.
-Moreover, ${\bf R}(+, ;)$ has the decidable axiomatisation.
+  ${\bf R}(+, ;)$ is axiomatised with the axioms of join semilattice-ordered semigroups and the coutable sequences of universal formulas $\{ \rho_n \}_{n < \omega}$.
+Moreover, ${\bf R}(+, ;)$ has a recursively enumerable axiomatisation.
 \end{theorem}
+
+\begin{proof}
+  Being the class of ${\bf RRA}$-reducts, ${\bf R}(+, ;)$ forms a pseudo-elementary class (see \cite{eklof1977ultraproducts} to have a rigorous definition).
+So we apply \cite[Theorem 9.28]{hirsch2002relation}.
+Therefore, a join semilattice-ordered semigroup $\mathcal{A}$ is representable iff $\mathcal{A} \models \{ \sigma_n \}_{n < \omega}$.
+Every $\sigma_n$ for $n < \omega$ defined in the proof of Lemma~\ref{ax} is equivalent to the universal formula $\rho_n$ by bringing all universal quantifiers into the prefix of $\sigma_n$.
+\end{proof}
 
@@
 Both sets are domains of corresponding computable functions, so they are enumerable.
-That makes ${\bf R}_{fin}(;, +)$ decidable by Post theorem.
+That makes ${\bf R}_{fin}(;, +)$ decidable by Post's theorem.
 \end{proof}

### 2021-02-24 filtration-stuff/Text.tex
some thoughts

@@
 
-\begin{defin} Let $\varphi$ be a modal formula, $\mathcal{L}$ a logic that admits strict filtration, and $\mathcal{M}$ an $\mathcal{L}$-model that refutes $\varphi$.
-A \emph{filtration diagram} of a model $\widehat{\mathcal{M}}$ (a filtration of $\mathcal{M}$ through $\varphi$) is a formula $\operatorname{FDiagram}(\mathcal{M}_{\mathcal{L}}^{\operatorname{Sub}(\varphi)}) = \bigwedge \Gamma \wedge \bigwedge \Delta$, where
-  \begin{itemize}
-    \item $\Gamma =
-    \{ \psi \in \operatorname{Sub}(\varphi) \: | \: \exists w \in \underline{\mathcal{M}} \: \widehat{M}, \hat{w} \models \psi \}$
-    \item $\Delta = \{ \neg \psi \in \operatorname{Sub}(\varphi) \: | \: \psi \in \operatorname{Sub}(\varphi) \: \& \: \mathcal{M} \nvDash \psi \}$
-  \end{itemize}
-\end{defin}
-
-\begin{lemma}
-  Let $\varphi$ be a modal formula, $\mathcal{L}$ a logic that admits strict filtration, and $\mathcal{M}$ an $\mathcal{L}$-model that refutes $\varphi$.
-Then $\mathcal{M}_{\mathcal{L}}^{\operatorname{Sub}(\varphi)} \models \operatorname{FDiagram}(\mathcal{M}_{\mathcal{L}}^{\operatorname{Sub}(\varphi)})$
-\end{lemma}
-
-\begin{defin}
-  Let $\mathcal{M}$ be a model and $\Psi$ a set of formulas, then a $\Psi$-theory of a model, denoted as $\operatorname{Th}_{\Psi}$, is the set $\operatorname{Th}_{\Psi} = \{ \varphi \in \Psi \: | \: \mathcal{M} \models \varphi \}$.
-\end{defin}
-
-\begin{lemma}
-  Let $\mathcal{L}$ be a modal logic that admits strict filtrations and $\varphi \not\in \mathcal{L}$.
-  Let $\mathcal{M}$ be a model that refutes $\varphi$ and $\mathcal{N}$ be a different model such that $\mathcal{N} \models \mathcal{L}$.
-If $\mathcal{N} \models \operatorname{FDiagram}(\mathcal{M}_{\mathcal{L}}^{\operatorname{Sub}(\varphi)})$, then $\operatorname{Th}_{\operatorname{Sub}(\varphi)}(\mathcal{M}_{\mathcal{L}}^{\operatorname{Sub}(\varphi)}) = \operatorname{Th}_{\operatorname{Sub}(\varphi)}(\mathcal{N})$.
-In particular, $\mathcal{N}$ refutes $\varphi$.
-\end{lemma}
-
 \begin{theorem} Let $\mathcal{L}$ be a canonical logic, then the following are equivalent:
@@
     Let $\varphi \not\in \mathcal{L}$, there exists a model $\mathcal{M}$ that refutes $\varphi$, so there exists $w \in \underline{\mathcal{M}}$ such that $\mathcal{M}, w \nvDash \varphi$.
-But $\mathcal{L}$ admits strict filtration, so there exists a model $\widehat{\mathcal{M}} = \langle \widehat{W}, \widehat{R}, \widehat{\vartheta} \rangle$, where $\widehat{W} = W / \sim_{\varphi}$ and $\hat{R}^{min} \subseteq \widehat{R} \subseteq \hat{R}^{max}$.
+But $\mathcal{L}$ admits strict filtration, so there exists a model
+    $\widehat{\mathcal{M}} = \langle \widehat{W}, \widehat{R}, \widehat{\vartheta} \rangle$, where $\widehat{W} = W / \sim_{\varphi}$ and $\hat{R}^{min} \subseteq \widehat{R} \subseteq \hat{R}^{max}$.
 In particular, $\widehat{\mathcal{M}}, \widehat{w} \nvDash \varphi$.
+    Let us put $T = \operatorname{Th}(\widehat{\mathcal{M}})$.
 
-    By lemma above, $\widehat{\mathcal{M}} \models \operatorname{FDiagram}(\widehat{\mathcal{M}})$
+    ...
 
-    Consider a finite ``canonical'' model
-    $\mathcal{M}_{\mathcal{L}}^{\operatorname{Sub}(\varphi)} = \langle W_{\mathcal{L}}^{\operatorname{Sub}(\varphi)}, R_{\mathcal{L}}^{\operatorname{Sub}(\varphi)}, \vartheta_{\mathcal{L}}^{\operatorname{Sub}(\varphi)} \rangle$.
-    Let us show that $\mathcal{M}_{\mathcal{L}}^{\operatorname{Sub}}$ refutes $\varphi$.
-For that, we show that $\mathcal{M}_{\mathcal{L}}^{\operatorname{Sub}(\varphi)} \models \operatorname{FDiagram}(\widehat{\mathcal{M}})$.
+    On the other hand, $\mathcal{L}$ is canonical. Therefore, $\mathcal{F}_{\mathcal{L}} \models \mathcal{L}$ as well as $\mathcal{F}_{T} \models \mathcal{L}$.
+
+    Here $\mathcal{F}_{T}$ is a frame $\langle W_T, R_T \rangle$, where $W_T$ is the set of all maximal consistent theories that contain $T$ and $R_T = R_{\mathcal{L}} \cap W_T \times W_T$.
+
+    $\mathcal{M}_{T} = \langle \mathcal{F}_{\widehat{\mathcal{M}}}, \vartheta_{\widehat{\mathcal{M}}} \rangle$,
+    being a generated submodel of $\mathcal{M}_{\mathcal{L}}$, does not have to be a ``finite canonical'' model according to our definition.
+
+    Let us put $W_{\mathcal{L}}^{\varphi} = \{ \Gamma \subseteq \operatorname{Sub}(\varphi) \cup \neg \operatorname{Sub}(\varphi) \: | \: \Gamma \text{ is maximal consistent} \}$.
+
+    ...
 

### 2021-02-27 filtration-stuff/Text.tex
refactor proof

@@
 If $\sim = \sim_{\Gamma}$, then such a filtration by means of the definiton above is called \emph{strict}.
+A class of models $\mathbb{M}$ admits strict filtrations for models (ASF), if for every $\operatorname{Sub}$-closed set $\Gamma$ and for every $\mathcal{M} \in \mathbb{M}$ there exists a $\Gamma$ filtration of $\mathcal{M}$.
+A class of frames $\mathbb{F}$ admits strict filtrations for frames, if for every $\operatorname{Sub}$-closed set $\Gamma$ and for every frame $\mathcal{F} \in \mathbb{F}$ and every model $\mathcal{M}$ over $\mathcal{F}$ there exists a $\Gamma$ filtration of $\mathcal{M}$.
+If $\mathcal{L}$ is canonical, then the ASF property for frames and ASF property for models are equivalent \cite[Theorem 2.10]{kikot2020completeness}.
 
@@
 
-A subset $\Delta \subseteq \Gamma$ is a \emph{finite $\mathcal{L}$-theory} if $\bigwedge \Delta \notin \mathcal{L}$. A subset $\Delta$ is maximal, if (the following are obviouslyequivalent):
+A subset $\Delta \subseteq \Gamma$ is a \emph{finite $\mathcal{L}$-consistent set} if $\neg \bigwedge \Delta \notin \mathcal{L}$.
+A subset $\Delta$ is maximal, if (the following are obviously equivalent):
 \begin{enumerate}
-  \item $\Delta$ is maximal amongst finite $\mathcal{L}$-theories,
+  \item $\Delta$ is maximal amongst finite $\mathcal{L}$-consistent sets,
   \item For each $\psi \in \operatorname{Sub}(\varphi)$ either $\psi \in \Delta$ or
@@
 
-\begin{defin}
+\begin{defin}\label{fin}
   Let $\mathcal{L}$ be a normal modal logic and $\varphi \not\in \mathcal{L}$.
@@
 
-\begin{theorem} Let $\mathcal{L}$ be a canonical logic, then the following are equivalent:
+\begin{theorem} $\mathcal{L}$ be a modal logic, then the following are equivalent:
   \begin{enumerate}
@@
     \item $(\Rightarrow)$
+    Let $\varphi \not\in \mathcal{L}$, let us put $\Gamma = \operatorname{Sub}(\varphi)$.
+    $\mathcal{L}$.
+    In particular, $\mathcal{M}_{\mathcal{L}} \nvDash \varphi$, where $\mathcal{M}_{\mathcal{L}}$ is the canonical model of $\mathcal{L}$.
+$\mathcal{L}$ admits strict filtrations, so the filtration of $\mathcal{M}_{\mathcal{L}}$ through $\Gamma$ is also an $\mathcal{L}$-model.
+    The underlying set of $\mathcal{M}_{\mathcal{L}} / \sim_{\Gamma}$ consists of maximal $\mathcal{L}$ theories up to $\Gamma$-equivalence and this quotient set is finite.
 
-    Let $\varphi \not\in \mathcal{L}$, there exists a model $\mathcal{M}$ that refutes $\varphi$, so there exists $w \in \underline{\mathcal{M}}$ such that $\mathcal{M}, w \nvDash \varphi$.
-But $\mathcal{L}$ admits strict filtration, so there exists a model
-    $\widehat{\mathcal{M}} = \langle \widehat{W}, \widehat{R}, \widehat{\vartheta} \rangle$, where $\widehat{W} = W / \sim_{\varphi}$ and $\hat{R}^{min} \subseteq \widehat{R} \subseteq \hat{R}^{max}$.
-In particular, $\widehat{\mathcal{M}}, \widehat{w} \nvDash \varphi$.
-    Let us put $T = \operatorname{Th}(\widehat{\mathcal{M}})$.
-
-    ...
-
-    On the other hand, $\mathcal{L}$ is canonical. Therefore, $\mathcal{F}_{\mathcal{L}} \models \mathcal{L}$ as well as $\mathcal{F}_{T} \models \mathcal{L}$.
-
-    Here $\mathcal{F}_{T}$ is a frame $\langle W_T, R_T \rangle$, where $W_T$ is the set of all maximal consistent theories that contain $T$ and $R_T = R_{\mathcal{L}} \cap W_T \times W_T$.
-
-    $\mathcal{M}_{T} = \langle \mathcal{F}_{\widehat{\mathcal{M}}}, \vartheta_{\widehat{\mathcal{M}}} \rangle$,
-    being a generated submodel of $\mathcal{M}_{\mathcal{L}}$, does not have to be a ``finite canonical'' model according to our definition.
-
-    Let us put $W_{\mathcal{L}}^{\varphi} = \{ \Gamma \subseteq \operatorname{Sub}(\varphi) \cup \neg \operatorname{Sub}(\varphi) \: | \: \Gamma \text{ is maximal consistent} \}$.
-
-    ...
+    It is readily checked that the quotient model $\mathcal{M}_{\mathcal{L}} / \sim_{\Gamma}$ satifsfies Definition~\ref{fin}.
 
@@
     Let us show that there exists a model $\widehat{\mathcal{M}}$ such that
-    $\widehat{\mathcal{M}}$ is a
-    strict filtration of $\mathcal{M}$ through $\operatorname{Sub}(\varphi)$.
+    $\widehat{\mathcal{M}}$ is a strict filtration of $\mathcal{M}$ through $\operatorname{Sub}(\varphi)$.
 
@@
     consider its truth set $||[u]|| = \{ \psi \in \operatorname{Sub}(\varphi) \: | \: \mathcal{M}, u \models \psi \}$.
-    $||[u]||$ is clearly a finite consistent theory and it consists in some $\Delta \in W_{\mathcal{L}}^{\varphi}$ by the finite analogue of Lindenbaum's lemma.
-Such a $\Delta$ is clearly unique as well.
+    $||[u]||$ is clearly a finite $\mathcal{L}$-consistent set and it is maximal.
     Consider a map
@@
     $\iota : W/\sim_{\operatorname{Sub}(\varphi)} \to W_{\mathcal{L}}^{\varphi}$ with
-    $\iota : [u] \mapsto \Delta \supseteq ||[u]||$
+    $\iota : [u] \mapsto ||[u]||$
     \end{center}
 
-    Consider the set $\iota(W/\sim_{\operatorname{Sub}(\varphi)})$ denoted as $\operatorname{Im}(\iota)$. Let us show that $\operatorname{Im}(\iota)$ is $R_{\mathcal{L}}^{\varphi}$-closed.
+    Consider the set $\iota(W/\sim_{\operatorname{Sub}(\varphi)})$ that we denote as $\operatorname{Im}(\iota)$. Let us show that $\operatorname{Im}(\iota)$ is $R_{\mathcal{L}}^{\varphi}$-closed.
 
@@
 
-    Consider $\Delta$ as a finite ultrafilter.
-Thus, $\Delta$ is principal and it 's generated by $\psi_1 \land \dots \land \psi_n$, where each $\psi_i \in \operatorname{Sub}(\varphi)$.
-So $\Diamond (\psi_1 \land \dots \land \psi_n) \in \iota([u])$.
+    Note that $\bigwedge \limits_{\psi \in \Delta} \psi \in \Delta$ since $\Delta$ is finite.
+    This conjuction has the form $\bigwedge \limits_{\psi \in \Delta} \psi \in \Delta = \psi_1 \land \dots \land \psi_n$ for some $n < \omega$ and for some $\psi_i \in \operatorname{Sub}(\varphi) \cup \neg \operatorname{Sub}(\varphi)$, where $0 < i < \omega$.
+
+    Therefore $\mathcal{M}^{\varphi}_{\mathcal{L}}, \iota([u]) \models \Diamond (\psi_1 \land \dots \land \psi_n)$.
     Then $\mathcal{M}, u \models \Diamond (\psi_1 \land \dots \land \psi_n)$. That implies $\mathcal{M}, v \models \psi_1 \land \dots \land \psi_n$ for some $v \in R(u)$.
@@
 
-    Thus, $\langle \operatorname{Im}(\iota), R_{\mathcal{L}}^{\varphi} \cap \operatorname{Im}(\iota) \times \operatorname{Im}(\iota), {\vartheta'}_{\mathcal{L}}^{\varphi} \rangle$, where ${\vartheta'}_{\mathcal{L}}^{\varphi}(p) = \vartheta_{\mathcal{L}}^{\varphi}(p) \cap \operatorname{Im}(\iota)$ is a generated submodel of $\mathcal{M}_{\mathcal{L}}^{\varphi}$.
+    Thus, $\langle \operatorname{Im}(\iota), R_{\mathcal{L}}^{\varphi}, {\vartheta'}_{\mathcal{L}}^{\varphi} \rangle$, where ${\vartheta'}_{\mathcal{L}}^{\varphi}(p) = \vartheta_{\mathcal{L}}^{\varphi}(p) \cap \operatorname{Im}(\iota)$
+    is a generated submodel of $\mathcal{M}_{\mathcal{L}}^{\varphi}$.
 
@@
     \begin{center}
-      $[u] \widehat{R} [v]$ iff $||[u]|| {R'}_{\mathcal{L}}^{\varphi} ||[v]||$, where ${R'}_{\mathcal{L}}^{\varphi} R_{\mathcal{L}}^{\varphi} \cap \operatorname{Im}(\iota) \times \operatorname{Im}(\iota)$.
+      $[u] \widehat{R} [v]$ iff $||[u]|| {R}_{\mathcal{L}}^{\varphi} ||[v]||$
     \end{center}
-    Let us prove that $\widehat{\mathcal{M}}$ is a filtration of $\mathcal{M}$ through $\operatorname{Sub}(\varphi)$.
 
@@
     $\mathcal{M}, u' \models \Diamond \psi$. Recall that $u' \sim_{\operatorname{Sub}(\varphi)} u$ and $v' \sim_{\operatorname{Sub}(\varphi)} v$, so we have $\Diamond ||[v]|| \subseteq ||[u]||$.
-    The same inclusion clearly holds for the maximal extensions of $\Diamond ||[v]||$ and $||[u]||$.
 

### 2021-03-04 filtration-stuff/Text.tex
fix the proof of Theorem 3

@@
 
-Let $\mathcal{L}$ be a normal modal logic, $\mathcal{M}_{\mathcal{L}}$ its canonical model, and $\varphi$.
-Let us put $\Gamma = \operatorname{Sub}(\varphi) \cup \{ \neg \psi \: | \: \psi \in \operatorname{Sub}(\varphi) \}$.
+Let $\mathcal{L}$ be a normal modal logic, $\mathcal{M}_{\mathcal{L}}$ its canonical model, and $\Gamma$ a finite $\operatorname{Sub}$-closed set of formulas.
+Let us put $\Gamma' = \operatorname{Sub}(\varphi) \cup \{ \neg \psi \: | \: \psi \in \operatorname{Sub}(\varphi) \}$.
 
-A subset $\Delta \subseteq \Gamma$ is a \emph{finite $\mathcal{L}$-consistent set} if $\neg \bigwedge \Delta \notin \mathcal{L}$.
+A subset $\Delta \subseteq '\Gamma$ is a \emph{finite $\mathcal{L}$-consistent set} if $\neg \bigwedge \Delta \notin \mathcal{L}$.
 A subset $\Delta$ is maximal, if (the following are obviously equivalent):
@@
 \begin{defin}\label{fin}
-  Let $\mathcal{L}$ be a normal modal logic and $\varphi \not\in \mathcal{L}$.
-A finite ``canonical'' model is a triple $\mathcal{M}_{\mathcal{L}}^{\varphi} = \langle W_{\mathcal{L}}^{\varphi}, R_{\mathcal{L}}^{\varphi}, \vartheta_{\mathcal{L}}^{\varphi} \rangle$, where
+  Let $\mathcal{L}$ be a modal logic and $\Gamma$ be a finite $\operatorname{Sub}$-closed set of formulas.
+A finite ``canonical'' model is a triple $\mathcal{M}_{\mathcal{L}}^{\Gamma} = \langle W_{\mathcal{L}}^{\Gamma}, R_{\mathcal{L}}^{\Gamma}, \vartheta_{\mathcal{L}}^{\Gamma} \rangle$, where
   \begin{enumerate}
-    \item $W_{\mathcal{L}}^{\varphi}$ is the set all maximal theories that extend finite $\mathcal{L}$-theories
-    \item $R_{\mathcal{L}}^{\varphi}$ is a relation such that
-    $\langle W_{\mathcal{L}}^{\varphi}, R_{\mathcal{L}}^{\varphi} \rangle$ is an $\mathcal{L}$-frame
+    \item $W_{\mathcal{L}}^{\Gamma}$ is the set all maximal theories that extend finite $\mathcal{L}$-theories
+    \item $R_{\mathcal{L}}^{\Gamma}$ is a relation such that
+    $\langle W_{\mathcal{L}}^{\Gamma}, R_{\mathcal{L}}^{\Gamma} \rangle$ is an $\mathcal{L}$-frame
     and
@@
     $\forall \Box \psi \in \operatorname{Sub}(\varphi) \:\:
-    \forall \Delta_1 \in W_{\mathcal{L}}^{\varphi} \:\: (\Box \psi \in \Delta_1 \Leftrightarrow
-    \forall \Delta_2 \in R_{\mathcal{L}}^{\varphi}(\Delta_1) \:\: \psi \in \Delta_2)$
+    \forall \Delta_1 \in W_{\mathcal{L}}^{\Gamma} \:\: (\Box \psi \in \Delta_1 \Leftrightarrow
+    \forall \Delta_2 \in R_{\mathcal{L}}^{\Gamma}(\Delta_1) \:\: \psi \in \Delta_2)$
     \end{center}
-    \item $\vartheta_{\mathcal{L}}^{\varphi}(p) = \{ \Delta \in W_{\mathcal{L}}^{\varphi} \: | \: p \in \Delta \}$.
+    \item $\vartheta_{\mathcal{L}}^{\Gamma}(p) = \{ \Delta \in W_{\mathcal{L}}^{\varphi} \: | \: p \in \Delta \}$ for every variable $p \in \Gamma$.
   \end{enumerate}
 \end{defin}
+
+\begin{lemma}
+  Let $\mathcal{L}$ be a modal logic and $\varphi \not\in \mathcal{L}$, then $\mathcal{M}_{\mathcal{L}}^{\operatorname{Sub}(\varphi)} \nvDash \varphi$.
+\end{lemma}
 
@@
     \item $\mathcal{L}$ admits strict filtrations
-    \item If $\varphi \not\in \mathcal{L}$, there exists a finite ``canonical'' model $\mathcal{M}_{\mathcal{L}}^{\varphi}$ such that
-    $\mathcal{M}_{\mathcal{L}}^{\varphi} \models \mathcal{L}$ and
-    $\mathcal{M}_{\mathcal{L}}^{\varphi}$ refutes $\varphi$.
+    \item If $\Gamma$ is a finite $\operatorname{Sub}$-closed set of formulas , there exists a finite ``canonical'' model $\mathcal{M}_{\mathcal{L}}^{\Gamma}$ such that
+    $\mathcal{M}_{\mathcal{L}}^{\Gamma} \models \mathcal{L}$.
   \end{enumerate}
@@
     \item $(\Rightarrow)$
-    Let $\varphi \not\in \mathcal{L}$, let us put $\Gamma = \operatorname{Sub}(\varphi)$.
-    $\mathcal{L}$.
-    In particular, $\mathcal{M}_{\mathcal{L}} \nvDash \varphi$, where $\mathcal{M}_{\mathcal{L}}$ is the canonical model of $\mathcal{L}$.
-$\mathcal{L}$ admits strict filtrations, so the filtration of $\mathcal{M}_{\mathcal{L}}$ through $\Gamma$ is also an $\mathcal{L}$-model.
+  Let $\Gamma$ be a finite $\operatorname{Sub}$-closed of formulas.
+$\mathcal{L}$ admits strict filtrations, so the filtration of the canonical model $\mathcal{M}_{\mathcal{L}}$ through $\Gamma$ is also an $\mathcal{L}$-model.
 The underlying set of $\mathcal{M}_{\mathcal{L}} / \sim_{\Gamma}$ consists of maximal $\mathcal{L}$ theories up to $\Gamma$-equivalence and this quotient set is finite.
@@
 
-    Let $\varphi \not\in \mathcal{L}$ and $\mathcal{M}_{\mathcal{L}}^{\varphi} = \langle W_{\mathcal{L}}^{\varphi}, R_{\mathcal{L}}^{\varphi}, \vartheta_{\mathcal{L}}^{\varphi} \rangle$
-    be a finite ``canonical'' model that refutes $\varphi$.
-    Let $\mathcal{M}$ be an $\mathcal{L}$-model that also refutes $\varphi$.
+    Let $\Gamma$ be a finite $\operatorname{Sub}$-closed of formulas and $\mathcal{M}_{\mathcal{L}}^{\Gamma} = \langle W_{\mathcal{L}}^{\Gamma}, R_{\mathcal{L}}^{\Gamma}, \vartheta_{\mathcal{L}}^{\Gamma} \rangle$ a ``finite canonical'' model generated by $\Gamma$.
+
     Let us show that there exists a model $\widehat{\mathcal{M}}$ such that
-    $\widehat{\mathcal{M}}$ is a strict filtration of $\mathcal{M}$ through $\operatorname{Sub}(\varphi)$.
+    $\widehat{\mathcal{M}}$ is a strict filtration of $\mathcal{M}$ through $\operatorname{Sub}(\Gamma)$.
 
-    Let $\sim_{\operatorname{Sub}(\varphi)}$ be an equivalence relation on $\underline{\mathcal{M}}$ defined as usual:
+    Let $\sim$ be an equivalence relation on $\underline{\mathcal{M}}$ defined as usual:
 
     \begin{center}
-      $u \sim_{\operatorname{Sub}(\varphi)} v$ iff $\forall \psi \in \operatorname{Sub}(\varphi)$
+      $u \sim v$ iff $\forall \psi \in \Gamma$
       $\mathcal{M}, u \models \psi \Leftrightarrow \mathcal{M}, v \models \psi$
@@
 
-    Let $[u] \in W/\sim_{\operatorname{Sub}(\varphi)}$, where $W = \underline{\mathcal{M}}$,
-    consider its truth set $||[u]|| = \{ \psi \in \operatorname{Sub}(\varphi) \: | \: \mathcal{M}, u \models \psi \}$.
+    Let $[u] \in \mathcal{M}/\sim$, consider its truth set $||[u]|| = \{ \psi \in \Gamma \: | \: \mathcal{M}, u \models \psi \}$.
     $||[u]||$ is clearly a finite $\mathcal{L}$-consistent set and it is maximal.
@@
     \begin{center}
-    $\iota : W/\sim_{\operatorname{Sub}(\varphi)} \to W_{\mathcal{L}}^{\varphi}$ with
-    $\iota : [u] \mapsto ||[u]||$
+    $\iota : W/\sim \to W_{\mathcal{L}}^{\varphi}$ with $\iota : [u] \mapsto ||[u]||$
     \end{center}
 
-    Consider the set $\iota(W/\sim_{\operatorname{Sub}(\varphi)})$ that we denote as $\operatorname{Im}(\iota)$. Let us show that $\operatorname{Im}(\iota)$ is $R_{\mathcal{L}}^{\varphi}$-closed.
+    Consider the set $\iota(W/\sim)$ that we denote as $\operatorname{Im}(\iota)$. Let us show that $\operatorname{Im}(\iota)$ is $R_{\mathcal{L}}^{\varphi}$-closed.
 
-    Let $\iota([u]) R_{\mathcal{L}}^{\varphi} \Delta$, let us show that $\Delta \in \iota(W/\sim_{\operatorname{Sub}(\varphi)})$, that is, $\Delta = \iota([v])$ for some $v \in W$.
+    Let $\iota([u]) R_{\mathcal{L}}^{\varphi} \Delta$, let us show that $\Delta \in \iota(W/\sim)$, that is, $\Delta = \iota([v])$ for some $v \in W$.
 
     Note that $\bigwedge \limits_{\psi \in \Delta} \psi \in \Delta$ since $\Delta$ is finite.
-    This conjuction has the form $\bigwedge \limits_{\psi \in \Delta} \psi \in \Delta = \psi_1 \land \dots \land \psi_n$ for some $n < \omega$ and for some $\psi_i \in \operatorname{Sub}(\varphi) \cup \neg \operatorname{Sub}(\varphi)$, where $0 < i < \omega$.
+    This conjuction has the form $\bigwedge \limits_{\psi \in \Delta} \psi \in \Delta = \psi_1 \land \dots \land \psi_n$ for some $n < \omega$ and for some $\psi_i \in \Gamma \cup \neg \Gamma$, where $0 < i < \omega$.
 
-    Therefore $\mathcal{M}^{\varphi}_{\mathcal{L}}, \iota([u]) \models \Diamond (\psi_1 \land \dots \land \psi_n)$.
+    Therefore $\mathcal{M}^{\Gamma}_{\mathcal{L}}, \iota([u]) \models \Diamond (\psi_1 \land \dots \land \psi_n)$.
     Then $\mathcal{M}, u \models \Diamond (\psi_1 \land \dots \land \psi_n)$. That implies $\mathcal{M}, v \models \psi_1 \land \dots \land \psi_n$ for some $v \in R(u)$.
@@
 
-    Thus, $\langle \operatorname{Im}(\iota), R_{\mathcal{L}}^{\varphi}, {\vartheta'}_{\mathcal{L}}^{\varphi} \rangle$, where ${\vartheta'}_{\mathcal{L}}^{\varphi}(p) = \vartheta_{\mathcal{L}}^{\varphi}(p) \cap \operatorname{Im}(\iota)$
-    is a generated submodel of $\mathcal{M}_{\mathcal{L}}^{\varphi}$.
+    Thus, $\langle \operatorname{Im}(\iota), R_{\mathcal{L}}^{\Gamma}, {\vartheta'}_{\mathcal{L}}^{\Gamma} \rangle$, where ${\vartheta'}_{\mathcal{L}}^{\Gamma}(p) = \vartheta_{\mathcal{L}}^{\Gamma}(p) \cap \operatorname{Im}(\iota)$
+    is a generated submodel of $\mathcal{M}_{\mathcal{L}}^{\Gamma}$.
 
-    Consider a model $\widehat{\mathcal{M}} = \langle W/\sim_{\operatorname{Sub}(\varphi)}, \widehat{R}, \widehat{\vartheta} \rangle$, where $\widehat{R}$ is defined as
+    Consider a model $\widehat{\mathcal{M}} = \langle W/\sim, \widehat{R}, \widehat{\vartheta} \rangle$, where $\widehat{R}$ is defined as
     \begin{center}
-      $[u] \widehat{R} [v]$ iff $||[u]|| {R}_{\mathcal{L}}^{\varphi} ||[v]||$
+      $[u] \widehat{R} [v]$ iff $||[u]|| {R}_{\mathcal{L}}^{\Gamma} ||[v]||$
     \end{center}
@@
     $[u] \widehat{R} [v]$. That is, let us show that $\Diamond \iota(v) \subseteq \iota(v)$.
-    But if $\mathcal{M}, v' \models \psi$ for $\Diamond \psi \in \operatorname{Sub}(\varphi)$, then
-    $\mathcal{M}, u' \models \Diamond \psi$. Recall that $u' \sim_{\operatorname{Sub}(\varphi)} u$ and $v' \sim_{\operatorname{Sub}(\varphi)} v$, so we have $\Diamond ||[v]|| \subseteq ||[u]||$.
+    But if $\mathcal{M}, v' \models \psi$ for $\Diamond \psi \in \Gamma$, then
+    $\mathcal{M}, u' \models \Diamond \psi$. Recall that $u' \sim u$ and $v' \sim v$, so we have $\Diamond ||[v]|| \subseteq ||[u]||$.
 

### 2021-03-04 frp-paper/Paper.tex
some thoughts
the finite representation property for join semilattice-ordered semigroups

@@
 
+\newcommand{\down}[1]{\ensuremath{{\downarrow}\,#1}}
+\newcommand{\up}[1]{\ensuremath{{\uparrow}\,#1}}
+
 \author{Daniel Rogozin \\ Institute for Information Transmission Problems, Russian Academy of Sciences \\ daniel.rogozin@serokell.io}
@@
 \title{The finite representation property for some reducts of relation algebras}
+\providecommand{\keywords}[1]{\textbf{\textit{Keywords---}} #1}
 
@@
 This result gives a positive solution to Problem 19.17 from the monograph by Hirsch and Hodkinson \cite{hirsch2002relation}.
+
+We also show that the class of representable join semilattice-ordered semigroups has a recursively
+enumerable recursive axiomatisation and the finite representation property.
 \end{abstract}
+
+\keywords{relation algebras, finite representation property, residuated semigroups, join semilattice-ordered semigroups}
 
@@
 
-Relation algebras are the kind of Boolean algebras with operators representing algebras of binary relations \cite{jonsson1951boolean}.
-One often emphasise the following two classes of relation algebras.
-The first class called ${\bf RA}$ consists of algebras the signature of which is $\{ 0, 1, +, -, ;, {}^{\smile}, {\bf 1}' \}$ obeying the certain axioms that we define precisely below.
-The second class called ${\bf RRA}$, the class of representable relation algebras, consists of algebras isomorphic to set relation algebras.
-${\bf RRA}$ is a subclass of ${\bf RA}$, but the converse inclusion does not hold.
-That is, there exist relation algebras having no representation as set relation algebras \cite{lyndon1950representation}.
-Moreover, the class ${\bf RRA}$ is not finitely axiomatisable in contrast to ${\bf RA}$ \cite{monk1964representable}.
-The problem of determining whether a given finite relation algebra is representable is undecidable, see \cite{hirsch2001representability}.
+Relation algebras are the kind of Boolean algebras with operators that provide algebraisation
+of binary relations \cite{jonsson1951boolean}. The class of all relation algebras,
+denoted as ${\bf RA}$, consists of algebras the signature of which is $\{ 0, 1, +, -, ;, {}^{\smile}, {\bf 1}' \}$ obeying certain axioms.
+The class of representable relation algebras, ${\bf RRA}$, consists of algebras isomorphic to set relation algebras. ${\bf RRA}$ is a subclass of ${\bf RA}$, but the converse inclusion does not hold.
+That is, there exist relation algebras with no representation as set relation algebras
+\cite{lyndon1950representation}. Moreover, the class ${\bf RRA}$ is not finitely axiomatisable \cite{monk1964representable}. The problem of determining whether a given finite relation algebra
+is representable is undecidable, see \cite{hirsch2001representability}.
 
-Under these circumstances, one may conclude that relation algebras are quite badly behaved.
+In other words, relation algebras are quite badly behaved.
 The study of such reducts is mostly motivated by such ``bad behaviour'' in order to avoid these restrictions and determine the possible reasons for them.
@@
 A finite algebra of relations has the finite representation property if it is isomorphic to some algebra of relations on a finite base.
-The investigation is of interest to study such aspects as the decidability of membership of ${\bf R}(\tau)$.
+The investigation of this problem is of interest to study such aspects as the decidability of membership of ${\bf R}(\tau)$.
 The finite representation property also implies recursivity of the class of all finite representable $\tau$-structures \cite{hirsch2004finite}, if the whole class is finitely axiomatisable.
@@
 
-Other algebras of relations such as weakly associative or relativised cylindric set algebras of a finite dimension have a similar property called the finite base property \cite{andreka1999finite}.
-
-There are subsignatures of relation algebras $\tau$ such that
-the class ${\bf R}(\tau)$ of representable reducts fails to have the finite representation property.
+There are subsignatures $\tau$ such that the class ${\bf R}(\tau)$ of representable reducts fails to have the finite representation property.
 For instance, $\{;, \cdot\}$, see \cite{maddux2016finite}.
@@
 
+In the last section, we study the class of representable join semilattice-ordered semigroups, denoted as ${\bf R}(;,+)$. It is
+already known that this class is not finitely axiomatisable \cite{andreka2011axiomatizability}.
+First, we show that ${\bf R}(;,+)$ has a recursively enumerable axiomatisation. For that,
+we define networks and representability games for them. This class is axiomatised with the axioms of join semilattice-ordered semigroups plus the countable set of universal formulas claiming that
+$\exists$ has a winning strategy on every finitary step.
+
+After that, we show that ${\bf R}(;,+)$ has the finite representation property using the same relational representation of quantales.
+In particular, we prove that, if $\mathcal{A} \in {\bf R}(;,+)$ is finite and representable, then the image $h(\mathcal{A})$ is isomorphic to its quantale representation.
+Here $h$ is a representation of $\mathcal{A}$.
+As a consequence, the representability of finite join-semilattice ordered semigroups is decidable.
+
 \section{Preliminaries}
@@
 Let us the basic definitions related to relation algebras \cite{hirsch2002relation}.
-\begin{definition} A relation algebra is an algebra $\mathcal{R} = \langle R, 0, 1, +, -, ;, {}^{\smile}, {\bf 1 }\rangle$ such that $\langle R, 0, 1, +, - \rangle$ is a Boolean algebra and the following
-    equations hold, for each $a, b, c \in R$:
+\begin{definition} A relation algebra is an algebra $\mathcal{R} = \langle R, 0, 1, +, -, ;, {}^{\smile}, {\bf 1 }\rangle$ such that $\langle R, 0, 1, +, - \rangle$ is a Boolean algebra and the following equations hold, for each $a, b, c \in R$:
     \begin{enumerate}
@@
 
-\section{The case of residuated semigroups}
+A \emph{residuated semigroup} is a structure $\mathcal{A} = \langle A, ;, \leq, \setminus, / \rangle$ such that $\langle A, ;, \leq, \rangle$ is a partially ordered residuated semigroup and $\setminus, /$ are binary operations satisfying the residuation property:
+
+  \begin{center}
+    $b \leq a \setminus c \Leftrightarrow a ; b \leq c \Leftrightarrow a \leq c \: / \: b$
+  \end{center}
+  ${\bf RS}$ is the class of all residuated semigroups.
+
+Let $\mathcal{A}$ be a semigroup and $X, Y \subseteq \mathcal{A}$, then $X \: ; \: Y = \{ x \: ; \: y \: | \: x \in X, y \in Y \}$.
+
+A \emph{join semilattice-ordered semigroup} is an algebra $\mathcal{A} = \langle A, ;, + \rangle$ such that $\langle A, ; \rangle$ is a semigroup, $\langle A, + \rangle$ is a join semilattice, and the following identity holds for all $a, b, c \in A$ the identity $a ; (b + c) = a ; b + a ; c$ holds.
+Every join semilattice-ordered semigroup is a poset and ordering is defined as $a \leq b$ iff $a + b = b$.
+
+\begin{definition}\label{joinrep}
+  A \emph{representation} $R$ of a join semilattice-ordered semigroup $\mathcal{A}$
+  is an injection $R : \mathcal{A} \to 2^{D \times D}$ (where $D$ is a non-empty base set) such that
+  \begin{enumerate}
+    \item $(a + b)^R = a^R \cup b^R$
+    \item $(a ; b)^R = a^R ; b^R$
+  \end{enumerate}
+\end{definition}
+
+A join semilattice-ordered semigroup $\mathcal{A}$ is \emph{representable}, if there exists a non-empty set $D$ and a function $f : \mathcal{A} \to 2^{D \times D}$ such that $f$ is a representation of $\mathcal{A}$.
+
+Let us also remind the reader several order-theoretic notions.
+Let $\langle P, \leq \rangle$ be a partial order.
+An upper cone generated by $x$ is the set $\up{x} = \{ a \in P \: | \: x \leq a \}$.
+Let $A \subseteq P$, then $\up{A} = \bigcup \limits_{x \in A} \up{x} = \{ a \in P \: | \: \exists x \in P \: x \leq a \}$.
+The set of all upper cones of a poset $\langle P, \leq \rangle$ is denoted as $\operatorname{Up}(P)$.
+Given $a \in P$, a \emph{lower cone} generated by $a$ is a subset $\down{a} = \{ x \in P \: | \: x \leq a \}$.
+
+A \emph{closure operator} on a poset $\langle P, \leq \rangle$ is a monotone map $j : P \to P$
+such that for each for each $a \in P$:
+\begin{enumerate}
+  \item $a \leq j a$
+  \item $j j a \leq j a$
+\end{enumerate}
+
+\section{The finite representation property for residuated semigroups}
 
@@
 
-Let us introduce the notion of a residuated semigroup.
-Historically, residuated structures were introduced by Krull to study ideals of rings \cite{krull1968idealtheorie}.
-Structures of this kind has been considered further within semantic aspects of substructural logics, see, for example, \cite{jipsen2002survey}.
-
-\begin{definition}
-  A residuated semigroup is an algebra $\mathcal{A} = \langle A, ;, \leq, \setminus, / \rangle$ such that $\langle A, ;, \leq, \rangle$ is a partially ordered residuated semigroup and $\setminus, /$ are binary operations satisfying the residuation property:
-
-  \begin{center}
-    $b \leq a \setminus c \Leftrightarrow a ; b \leq c \Leftrightarrow a \leq c \: / \: b$
-  \end{center}
-  ${\bf RS}$ is the class of all residuated semigroups.
-\end{definition}
-
-The class ${\bf R}(;, \setminus, /, \leq)$ consists of the following structures:
+The class ${\bf R}(;, \setminus, /, \leq)$ consists of the following structures, here is the explicit definition:
 
@@
 \end{definition}
+
 A residuated semigroup is called \emph{representable} if it is isomorphic to some algebra belonging to ${\bf R}(;, \setminus, /, \leq)$.
@@
 
- Andr\'{e}ka and Mikul\'{a}s proved the representation theorem for ${\bf RS}$ (\cite{andreka1994lambek}) in the fashion of step-by-step representation.
+ Andr\'{e}ka and Mikul\'{a}s proved the representation theorem for ${\bf RS}$ (\cite{andreka1994lambek}) in the step-by-step fashion.
 See this paper to learn more about step-by-step representations in general \cite{hirsch1997step}.
@@
 Thus, it is sufficient to show that any finite residuated semigroup has a representation over a finite base in order to show that ${\bf R}(;, \setminus, /, \leq)$ has the finite representation property.
-
-Let us start with preliminary order-theoretic definitions.
-Recall that a \emph{closure operator} on a poset $\langle P, \leq \rangle$ is a monotone map $j : P \to P$ satisfying $a \leq j a = j j a$ for each $a \in P$.
-Given $a \in P$, a \emph{lower cone} generated by $a$ is a subset $\downarrow a = \{ x \in P \: | \: x \leq a \}$.
+For that, we will use the representation of residuated
+semigroups as subalgebras of quantales and the relational representation of quantales.
 
 A quantale is a complete lattice-ordered semigroup.
-Quantales has been initially introduced by Mulvey to provide a noncommutative generalisation of locales, study the spectra of $C^{*}$-algebras, and classify Penrose tilings, see \cite{mulvey1986suppl} \cite{mulvey2005noncommutative}.
+Quantales has been initially introduced by Mulvey to provide a noncommutative generalisation of locales, see \cite{mulvey1986suppl}.
+The rigorous definition is the following:
 
@@
 
+Given a quantale $\mathcal{Q} = \langle Q, ;, \Sigma \rangle$, a set of \emph{generators} is a subset $\mathcal{G}(\mathcal{Q}) \subseteq \mathcal{Q}$, if
+\begin{enumerate}
+  \item For every $q \in Q$ one has $q \leq \Sigma \{ g \in \mathcal{G}(\mathcal{Q}) \: | \: g \leq q \}$.
+  \item For every $g \in \mathcal{G}(\mathcal{Q})$ and $q_1, q_2 \in \mathcal{Q}$, $g \leq q_1 ; q_2$ implies $g \leq q_1 ; r$ for some $r \in \mathcal{G}(\mathcal{Q})$ with $r \leq q_2$.
+\end{enumerate}
+
+Every generator is completely join-irreducible, this follows from the definition immediatly.
+
 Note that any quantale is a residuated semigroup as well.
-Given a quantale $\mathcal{Q} = \langle Q, ;, \Sigma \rangle$, One may express residuals uniquely with supremum and product as follows for each $a, b \in Q$:
+Given a quantale $\mathcal{Q} = \langle Q, ;, \Sigma \rangle$, One may express residuals with supremum and product as follows for each $a, b \in Q$:
 \begin{enumerate}
@@
 \end{enumerate}
+Note that, if a residual exists, then it is unique.
 
-A quantic nucleus is a closure operator on a quantale allowing one to define subalgebras.
-Such an operator is a generalisation of a well-known nucleus operator in locale theory and pointfree topology, see, e.
+A quantic nucleus is a closure operator on a quantale allowing one to define subquantales.
+Such an operator is a generalisation of a well-known nucleus operator from locale theory, see, e.
 g., \cite{bezhanishvili2016locales}.
-The following definition and the proposition below are due to \cite[Theorem 3.1.1]{rosenthal1990quantales}.
+The following definition and the proposition below are due to \cite[Definition 3.1.1, Theorem 3.1.1]{rosenthal1990quantales} respectively.
 \begin{definition}
@@
 \begin{proposition} \label{subsemi}
-  Let $\mathcal{A} = \langle A, ;, \Sigma \rangle$ be a quantale and $j$ a quantic nucleus, then
+  Let $\mathcal{A} = \langle A, ;, \Sigma \rangle$ be a quantale and $j$ a quantic nucleus, the set
   $\mathcal{A}_j = \{ a \in A \: | \: j a = a \}$ forms a quantale, where $a ;_j b = j(a ; b)$ and $\Sigma_j A = j (\Sigma A)$ for each $a, b \in {A}_j$ and $A \subseteq \mathcal{A}_j$.
@@
 
-We need this construction to solve the problem, let us discuss it briefly. See Goldblatt's paper to have a complete argument \cite{goldblatt2006kripke}.
+We need this construction to solve the problem, let us discuss it briefly. See Goldblatt's paper to have a complete argument in more detail \cite{goldblatt2006kripke}.
 
@@
 Note that the lower cone of an arbitrary $x$ is $m$-closed, that is,
-$m (\downarrow x) = \downarrow x$.
+$m (\down{x}) = \down{x}$.
 
@@
 \end{center}
-forms a complete lattice with $\Sigma_{m} \mathcal{X} = m ( \bigcup \mathcal{X})$ and $\Pi_{m} = \bigcap \mathcal{X}$ \cite{davey2002introduction}.
+forms a complete lattice with $\Sigma_{m} \mathcal{X} = m ( \bigcup \mathcal{X})$ and $\Pi_{m} = \bigcap \mathcal{X}$, see \cite[Theorem 7.3]{davey2002introduction}.
 
@@
 
-Let us define a mapping $f_m : \mathcal{A} \to (\mathcal{P}(A))_m$ such that $f_m : a \mapsto \downarrow a$.
+Let us define a mapping $f_m : \mathcal{A} \to (\mathcal{P}(A))_m$ such that $f_m : a \mapsto \down{a}$.
 This map is well-defined since any lower cone generated by a point is $m$-closed.
@@
 
-In their turn, quantales have a relational representation. First of all, let us define a relational quantale.
-The notion of a relational quantale was introduced by Brown and Gurr to represent quantales as algebras of relations and study relational semantics of the full Lambek calculus, the logic of bounded residuated lattices, see, e.g., \cite{brown1993representation}.
+Quantales, in turn, are representable with binary relations.
+The notion of a relational quantale was introduced by Brown and Gurr to represent quantales as algebras of relations \cite{brown1993representation}.
 \begin{definition}
@@
 
-The uniqueness of residuals in any quantale implies the following quite obvious fact.
+The uniqueness of residuals in any quantale implies the following fact.
 \begin{proposition}\label{ok}
@@
 \end{proposition}
-
-Let us discuss the relational representation of quantales. This construction is due to Brown and Gurr \cite{brown1993representation}.
-
-Let $\mathcal{Q}$ be a quantale and $\mathcal{G}(\mathcal{Q})$ a set of its generators. We define:
-
-\begin{center}
-  $\hat{a} = \{ \langle g, q \rangle \: | \: g \in \mathcal{G}(\mathcal{Q}), q \in \mathcal{Q}, g \leq a ; q \} \:\:\:\: \widehat{\mathcal{Q}} = \{ \hat{a} \: | \: a \in \mathcal{Q} \}$
-\end{center}
 
@@
 
-\item $\mathcal{Q}$ is isomorphic to $\langle \widehat{\mathcal{Q}}, \subseteq, ; \rangle$,
+\item $\mathcal{Q}$ is isomorphic to $\langle \widehat{\mathcal{Q}}, \subseteq, ; \rangle$ and $a \mapsto \hat{a}$ is a quantale isomorphism.
+\end{enumerate}
 
-\item $a \mapsto \hat{a}$ is a quantale isomorphism.
-\end{enumerate}
+We summarise the construction above with the following theorem proved by Brown and Gurr, see \cite[Theorem 3.11]{brown1993representation}.
 
@@
 $\widehat{\mathcal{Q}_{\mathcal{A}}}$ is the corresponding relational quantale.
-Let us define an interpretation $R : \mathcal{A} \to \widehat{\mathcal{Q}_{\mathcal{A}}}$ such that
+Let us define an interpretation $R : \mathcal{A} \to \widehat{\mathcal{Q}_{\mathcal{A}}}$ such that:
 
 \begin{center}
-  $R : a \mapsto a^{R} = \widehat{\downarrow a}$
+  $R : a \mapsto a^{R} = \widehat{\down{a}}$
 \end{center}
@@
 According to the lemma below, such an interpretation is a representation.
-As we have already said above, the mapping $a \mapsto \downarrow a$ is order-preserving.
-Moreover, this mapping commutes with products and residuals.
+As we have already said above, the function $a \mapsto \down{a}$ is order-preserving and it commutes with products and residuals.
 
@@
   Let $\tau$ be a signature of residuated semigroups.
-An interpretation $R : \mathcal{A} \to \widehat{\mathcal{Q}_{\mathcal{A}}}$ such that $R : a \mapsto a^{R} = \widehat{\downarrow a}$ is a $\tau$-representation.
+An interpretation $R : \mathcal{A} \to \widehat{\mathcal{Q}_{\mathcal{A}}}$ such that $R : a \mapsto a^{R} = \widehat{\down{a}}$ is a $\tau$-representation.
 \end{lemma}
@@
   By Theorem~\ref{orsRep}, $\mathcal{Q}_{\mathcal{A}}$ is isomorphic to $\widehat{\mathcal{Q}_{\mathcal{A}}}$.
-The isomorphism is established with the mapping $\downarrow{a} \mapsto \widehat{\downarrow a}$ according to Theorem~\ref{quantaleRep}.
+The isomorphism is established with the map $\downarrow{a} \mapsto \widehat{\down{a}}$ according to Theorem~\ref{quantaleRep}.
   Residuals in $\widehat{\mathcal{Q}_{\mathcal{A}}}$ are well-defined by Proposition~\ref{ok}.
@@
 \begin{center}
-  $\widehat{\mathcal{A}} = \langle \{ \widehat{\downarrow a} \}_{a \in \mathcal{A}}, ;, \setminus, /, \subseteq \rangle$.
+  $\widehat{\mathcal{A}} = \langle \{ \widehat{\down{a}} \}_{a \in \mathcal{A}}, ;, \setminus, /, \subseteq \rangle$.
 \end{center}
@@
 
-The main corollary of Theorem~\ref{solution} is that the Lambek calculus has the finite model property.
-Thus, we have a semantical proof of decidability of the Lambek calculus.
-Before that, there were several algebraic proofs that the Lambek caclulus has the FMP \cite{buszkowski2008infinitary}, but the authors used to consider arbitrary algebras, not representable ones.
-Alternatively, one may show that the Lambek calculus is decidable syntactically, that is, via cut elimination and the subformula property \cite{lambek1958mathematics}.
+As a corollary, of Theorem~\ref{solution} is that the Lambek calculus is complete with respect to finite representable residuated semigroups.
 
@@
 
-\section{The case of join semilattice-ordered semigroups}
-
-A \emph{join semilattice-ordered semigroup} is an algebra $\mathcal{A} = \langle A, ;, + \rangle$ such that $\langle A, ; \rangle$ is a semigroup, $\langle A, + \rangle$ is a join semilattice, and the following identity holds for all $a, b, c \in A$ the identity $a ; (b + c) = a ; b + a ; c$ holds.
-
-Let $\mathcal{A}$ be a join semilattice-ordered semigroup and $X, Y \subseteq \mathcal{A}$, then $X \: ; \: Y = \{ x \: ; \: y \: | \: x \in X, y \in Y \}$.
-
-Let $\langle P, \leq \rangle$ be a partial order.
-An upper cone generated by $x$ is the set $\uparrow x = \{ a \in P \: | \: x \leq a \}$.
-Let $A \subseteq P$, then $\uparrow A = \bigcup \limits_{x \in A} \uparrow x = \{ a \in P \: | \: \exists x \in P \: x \leq a \}$.
-The set of all upper cones of a poset $\langle P, \leq \rangle$ is denoted as $\operatorname{Up}(P)$.
+\section{Join semilattice-ordered semigroups: the explicit axiomatisation and the finite representation property}
 
@@
 
-\begin{definition}
-  A \emph{representation} $R$ of a join semilattice-ordered semigroup $\mathcal{A}$
-  is an injection $R : \mathcal{A} \to 2^{D \times D}$ (where $D$ is a non-empty base set) such that
+In this subsection, we show that ${\bf R}(+, ;)$ has a recursively enumerable universal axiomatisation.
+For that, we characterise representability using Ehrenfeucht-Fra\"{i}ss\'{e} games on prenetworks.
+
+\begin{definition} Let $\mathcal{A}$ be a join-semilattice ordered semigroup.
+A \emph{prenetwork} over $\mathcal{A}$ is a tuple $(V, E, l)$, where $V$ is a set of vertices, $E$ is a set of edges such that $\langle V, E \rangle$ is a directed graph, and $l$ is a labelling function $l : E \to \operatorname{Up}(\mathcal{A})$.
+
+  A prenetwork over $\mathcal{A} = (V, E, l)$ is a \emph{network} if the following hold:
   \begin{enumerate}
-    \item $(a + b)^R = a^R \cup b^R$
-    \item $(a ; b)^R = a^R ; b^R$
+    \item For every $u, v \in V$ and for every $x,y,z \in \mathcal{A}$ $z \in l(u, v)$ and $z \leq x \: ; \: y$ implies $x \in N(u, w)$ and $y \in l(w, v)$ for some $w \in V$.
+    \item For every $u, v, w \in V$ one has $l(u, v) ; l(v, w) \subseteq l(u, w)$.
+    \item For every $u, v \in V$ $l(u,v)$ is join-prime. That is, for each $a, b \in \mathcal{A}$ if $a + b \in l(u,v)$, then either $a \in l(u,v)$ or $b \in l(u,v)$.
   \end{enumerate}
@@
 
-It is known that the class of all representable join-semilattice ordered semigroups has no finite axiomatisation \cite{andreka1989union}.
-This proof has been reproduced in \cite{andreka2011axiomatizability} as well.
+If $\mathcal{A}$ is a prenetwork, then we will denote its sets of nodes as $\operatorname{Nodes}(\mathcal{A})$ occasionally.
 
-\begin{definition} Let $\mathcal{A}$ be a join-semilattice ordered semigroup.
-A prenetwork over $\mathcal{A}$ is a pair $(E, V, N)$, where $E$ is a set of nodes, $V$ is a set of nodes such that $\langle E, V \rangle$ is a complete directed graph, and $N$ is a function $l : V \to \operatorname{Up}(\mathcal{A})$.
+Let $\mathcal{N}_1 = (V_1, E_1, l_1)$ and $\mathcal{N}_2 = (V_2, E_2, l_2)$ be (pre)networks, then $\mathcal{N}_1$ is a \emph{subnetwork} of $\mathcal{N}_2$, written as $\mathcal{N}_1 \subseteq \mathcal{N}_2$,
+if $V_1 \subseteq V_2$, $E_1 \subseteq E_2$, and for all $x, y \in V_1$ if $a \in l_1(x, y)$ then there is $b \in l_2(x, y)$ with $b \leq a$.
+A $\mathcal{N}_1$ is an \emph{induced} sub(pre)network of $\mathcal{N}_2$, if $V_1 \subseteq V_2$ and for all $x, y \in V_1$ one has $l_1(x,y) = l_2(x,y)$.
 
-  A prenetwork over $\mathcal{A}$ is a network if the following hold:
-  \begin{enumerate}
-    \item For every $u, v \in E$ and for every $x,y,z \in \mathcal{A}$ $z \in l(u, v)$ and $z \leq x \: ; \: y$ implies $x \in N(u, w)$ and $y \in l(w, v)$ for some $w \in E$.
-    \item For every $u, v, w \in E$ one has $l(u, v) ; l(v, w) \subseteq l(u, w)$.
-    \item For every $u, v \in E$ $l(u,v)$ is join-prime. That is, for each $a, b \in \mathcal{A}$ if $a + b \in l(u,v)$, then either $a \in l(u,v)$ or $b \in l(u,v)$.
-  \end{enumerate}
-\end{definition}
-
-Let $\mathcal{N}_1 = (E_1, V_1, l_1)$ and $\mathcal{N}_2 = (E_2, V_2, l_2)$ be (pre)networks, then $\mathcal{N}_1$ is a \emph{subnetwork} of $\mathcal{N}_2$, written as $\mathcal{N}_1 \subseteq \mathcal{N}_2$,
-if $E_1 \subseteq E_2$, $V_1 \subseteq V_2$, and for all $x, y \in E_1$ if $a \in l_1(x, y)$ then there is $b \in l_2(x, y)$ with $b \leq a$.
-
-A $\mathcal{N}_1$ is an \emph{induced} sub(pre)network of $\mathcal{N}_2$, if $E_1 \subseteq E_2$ and for all $x, y \in E_1$ one has $l_1(x,y) = l_2(x,y)$.
-
-Let $\{ \mathcal{N}_i \}_{i \in I}$ be an indexed set of prenetworks (where each $\mathcal{N}_i = (E_i, V_i, l_i)$), then $\mathcal{N} = \bigcup \limits_{i \in I} \mathcal{N}_i$ defined as $(E, V, l)$
+Let $I$ be a non-empty index set and let $\{ \mathcal{N}_i \}_{i \in I}$ be an indexed set of prenetworks (where each $\mathcal{N}_i = (V_i, E_i, l_i)$), then $\mathcal{N} = \bigcup \limits_{i \in I} \mathcal{N}_i$ defined as $(V, E, l)$, where
 
 \begin{enumerate}
-  \item $E = \bigcup \limits_{i \in I} E_i$ and $V = \bigcup \limits_{i \in I} V_i$
-  \item $l(x, y) = \bigcup \limits_{ i \in I } l_i(x, y)$ for each $x, y \in E$
+  \item $V = \bigcup \limits_{i \in I} V_i$ and $E = \bigcup \limits_{i \in I} E_i$.
+  \item $l(x, y) = \bigcup \limits_{ i \in I } l_i(x, y)$ for each $x, y \in V$.
 \end{enumerate}
@@
 \begin{definition}
-  Let $n \leq \omega$ and $\mathcal{A}$ a join semilattice-ordered semigroup. A game $\mathcal{G}_n(\mathcal{A})$ has $n$ rounds and consists of $n$ finitary networks.
+  Let $n \leq \omega$ and $\mathcal{A}$ a join semilattice-ordered semigroup.
+A play of the game $\mathcal{G}_n(\mathcal{A})$ has $n$ rounds and consists of $n$ finitary networks.
+As usual, we have two players, $\forall$ (Abelard) and $\exists$ (H\'{e}lo\"{i}se).
 
@@
     \item Round $0$: $\forall$ picks elements $a, b \in \mathcal{A}$ with $a \not\leq b$.
-$\exists$ responds with a prenetwork $\mathcal{N}_0 = (E_0 = \{ x_0, x_1 \}, V_0 = \{ (x_0, x_1)\}, l_0)$ such that $l_0(x_0, x_1) = \uparrow b$.
-    \item Round $n + 1$. Suppose, the prenetwork $\mathcal{N}_n = (E_n, V_n, l_n)$ has been played.
+$\exists$ responds with a prenetwork $\mathcal{N}_0 = (V_0 = \{ x_0, x_1 \}, E_0 = \{ (x_0, x_1)\}, l_0)$ such that $l_0(x_0, x_1) = \up{b}$.
+    \item Round $n + 1$. Suppose, the prenetwork $\mathcal{N}_n = (V_n, E_n, l_n)$ has been played.
 
@@
     \begin{enumerate}
-      \item ({\bf Composition move}): $\forall$ can pick $x, y, z \in E_n$ with $a \in l_n(x, y)$ and $b \in l_n(y, z)$.
+      \item ({\bf Composition move}): $\forall$ can pick $x, y, z \in V_n$ with $a \in l_n(x, y)$ and $b \in l_n(y, z)$.
 We denote such a move as $N(x,y,z,a,b)$.
-Then $\exists$ responds with $\mathcal{N}_{n + 1} = (E_{n + 1}, V_{n + 1}, l_{n + 1})$ such that $\mathcal{N}_{n + 1}$
-      is the same as $\mathcal{N}_n$, but $l_{n + 1}(x, z) = \uparrow (l_{n}(x, z) \cup \{ a \: ; \: b\})$.
+Then $\exists$ responds with $\mathcal{N}_{n + 1} = (V_{n + 1}, E_{n + 1}, l_{n + 1})$ such that $\mathcal{N}_{n + 1}$
+      is the same as $\mathcal{N}_n$, but $l_{n + 1}(x, z) = \up{(l_{n}(x, z)} \cup \{ a \: ; \: b\})$.
       \item ({\bf Witness move}):
 
-      $\forall$ picks nodes $x, y \in E_n$ and $a, b \in \mathcal{A}$ such that $c \leq a ; b$ for $c \in l_n(x, y)$.
+      $\forall$ picks nodes $x, y \in V_n$ and $a, b \in \mathcal{A}$ such that $c \leq a ; b$ for $c \in l_n(x, y)$.
 $\exists$ has to find a witness.
 She responds with a fresh node $z$.
-We define the prenetwork $T$ , the edges of which are $x, y, z$ with the following labelling:
+We define the prenetwork $T$, the edges of which are $x, y, z$ with labelling:
       \begin{enumerate}
-        \item $l_T(x, z) = \uparrow a$
-        \item $l_T(z, y) = \uparrow b$
+        \item $l_T(x, z) = \up{a}$
+        \item $l_T(z, y) = \up{b}$
       \end{enumerate}
@@
 
-      $\forall$ picks nodes $x, y \in E_n$ and $a + b$ for $a, b \in \mathcal{A}$. $\exists$ has the following two alternatives for her response:
+      $\forall$ picks nodes $x, y \in V_n$ and $a + b$ for $a, b \in \mathcal{A}$. $\exists$ has the following two alternatives for her response:
       \begin{enumerate}
-        \item $\exists$ chooses $a$ and responds with the prenetwork $\mathcal{N}_{n + 1} = \langle E_n, V_{n}, l_{n + 1} \rangle$, where $l_{n+1}(x, y) = \uparrow (l_n(x,y) \cup \{ a \})$.
-        \item $\exists$ chooses $b$. The response is similar but $l_{n + 1}(x, y) = \uparrow (l_n(x,y) \cup \{ b \})$.
+        \item $\exists$ chooses $a$ and responds with the prenetwork $\mathcal{N}_{n + 1} = \langle V_{n + 1}, E_{n + 1}, l_{n + 1} \rangle$, where $l_{n+1}(x, y) = \up{(l_n(x,y) \cup \{ a \})}$.
+        \item $\exists$ chooses $b$. The response is similar but $l_{n + 1}(x, y) = \up{(l_n(x,y) \cup \{ b \})}$.
       \end{enumerate}
@@
 
-  $\exists$ wins for the match of the game $\mathcal{G}_n(\mathcal{A})$ if for each $0 \leq i \leq n$ $\mathcal{N}_i$ is a network.
+  $\exists$ wins for the match of the game $\mathcal{G}_n(\mathcal{A})$ if for each $0 \leq i \leq n \leq \omega$ $\mathcal{N}_i$ is a network.
 $\exists$ has a winning strategy, if she can win all the matches.
@@
 
-\begin{lemma} Let $\mathcal{A}$ be a join semilattice-ordered semigroup,
+\begin{lemma} Let $\mathcal{A} = \langle A, ;, + \rangle$ be a join semilattice-ordered semigroup,
 
@@
     \item If $\mathcal{A}$ is infinite and $\exists$ has a winning strategy in $\mathcal{G}_{\omega}(\mathcal{A})$ then $\mathcal{A}$ is representable.
-    \item If $\mathcal{A}$ is infinite and $\forall$ has a winning strategy in $\mathcal{G}_{\omega}(\mathcal{A})$ then $\mathcal{A}$ is not representable.
   \end{enumerate}
@@
   \begin{enumerate}
-    \item Let $h$ be a representation of $\mathcal{A}$ over some domain $D$.
-$\forall$ picks $(a, b)$ with $a \not\leq b$ for the initial round.
-Let $\mathcal{N} = \langle E, V, l \rangle$ be a network being played.
-To win the game $\exists$ maintains a map $\iota : E \to D$ such that $(\iota(x_0), \iota(x_1)) \in h(b) \setminus h(a)$ and for each $x, y \in E$ and
-    $a \in l_{\mathcal{N}}(x, y)$ one has $(\iota(x), \iota(y)) \in h(a)$. The argument repeats the proof of the first item of \cite[Proposition 5]{hirsch2005class}.
+    \item This item has the proof similar to the first item of \cite[Proposition 5]{hirsch2005class}.
 
@@
     \begin{center}
-      $rep(a) = \{ (x, y) \: | \: \exists b \in \downarrow a \:\: b \in l_{\mathcal{N}}(x, y)\}$
+      $rep(a) = \{ (x, y) \: | \: \exists b \in \down{a} \:\: b \in l_{\mathcal{N}}(x, y)\}$
     \end{center}
@@
 By the condition, $\forall$ makes the witness moves and $\exists$ responds with a witness.
-Her response is a fresh node $z$ such that $l_{\mathcal{N}_{i + 1}}(x, z) =  \uparrow (l_{\mathcal{N}_i}(x, z) \cup \{ a \})$ and $l_{\mathcal{N}_{i + 1}}(z, y) = \uparrow (l_{\mathcal{N}_i}(z, y) \cup \{ b\})$.
+Her response is a fresh node $z$ such that $l_{\mathcal{N}_{i + 1}}(x, z) =  \up{(l_{\mathcal{N}_i}(x, z) \cup \{ a \})}$ and $l_{\mathcal{N}_{i + 1}}(z, y) = \up{(l_{\mathcal{N}_i}(z, y) \cup \{ b\})}$.
 The inclusion $rep(a ; b) \subseteq rep(a) ; rep(b)$ holds since all witness moves have been played and for every such a move there exists a witness.
@@
 
-\item Injectivity: TODO
-
 \end{itemize}
-Now let us assume that $|\mathcal{A}| \geq \aleph_1$. By Tarksi-Vaught theorem on elementary chain union (see \cite[Theorem 2.5.2]{hodges1997shorter})
-
-\item
+For $|\mathcal{A}| \geq \aleph_1$ we apply the previous item and Tarksi-Vaught theorem on elementary chain union (see \cite[Theorem 2.5.2]{hodges1997shorter}).
   \end{enumerate}
@@
 Our purpose is to provide the recursively enumerable axiomatisation of ${\bf R}(;, +)$.
-Thus, according to famous Craig's trick \cite{craig1953axiomatizability}, ${\bf R}(;, +)$ will have the decidable axiomatisation.
-According to the proposition above, it is sufficient to extract the axioms claiming that $\exists$ has a winning strategy in all finitary games.
 
 \begin{definition}
-  Let $V = \{ v_0, v_1, ... \}$ be a countable set of variables. The set of terms $T$ is generated by the following grammar:
+  Let $\operatorname{Var} = \{ v_0, v_1, ... \}$ be a countable set of variables. The set of terms $T$ is generated by the following grammar:
 
   \begin{center}
-    $T ::= V \: | \: (T + T) \: | \: (T ; T)$
+    $T ::= \operatorname{Var} \: | \: (T + T) \: | \: (T ; T)$
   \end{center}
@@
 \begin{center}
-$t_1 \approx t_2$ iff $\theta(t_1) = \theta(t_2)$ for each interpretation of variables in $\theta : V \to F(\omega)$.
+$t_1 \approx t_2$ iff $F(\omega) \models t_1 = t_2$.
 \end{center}
-where $F(\omega)$ is the free countably generated algebra of the variety of all join semilattice-ordered group semigroup. We have the following partial order on the $T$ defined as
+where $F(\omega)$ is the free countably generated algebra join semilattice-ordered semigroup.
+We have the following partial order on $T$ defined as
 \begin{center}
@@
 \end{center}
-A subset $A \subseteq T$ of terms is called upward closed if for every term $t \in A$ one has $\uparrow t \subseteq A$, where $\uparrow t = \{ t_1 \in T \: | \: t \preceq t_1 \}$.
+
+We call subset $A \subseteq T$ of terms called upward closed if for every term
+$t \in A$ one has $\up{t} \subseteq A$, where $\up{t} = \{ t_1 \in T \: | \: t \preceq t_1 \}$.
 
 \begin{definition}
-  A term network is a graph $\langle E, V, l \rangle$, where $\langle E, V \rangle$ is a complete directed graph and $l : V \to \operatorname{Up}(F(\omega), \preceq)$ is a labelling function.
+  A \emph{term network} is a graph $\langle V, E, l \rangle$, where $\langle V, E \rangle$ is a directed graph and $l : E \to \operatorname{Up}(F(\omega), \preceq)$ is a labelling function.
 \end{definition}
 
-Let $\mathcal{A}$ be a join semilattice-ordered semigroup and $\vartheta : V \to \underline{{\mathcal{A}}}$ a valuation. The value of complex terms is defined inductively for $a, b \in T$:
+Let $\mathcal{A}$ be a join semilattice-ordered semigroup and $\vartheta : \operatorname{Var} \to \mathcal{A}$ a valuation. The value of complex terms is defined inductively for $a, b \in T$:
 
@@
 
-Let $\mathcal{N} = \langle E, V, l \rangle$ a term network, $\mathcal{A}$ be a join-semilattice ordered semigroup and $\vartheta : V \to \underline{{\mathcal{A}}}$ a valuation.
+As usual, one may think of $\vartheta$ as a homomorphism $\vartheta : F(\omega) \to \mathcal{A}$.
+
+Let $\mathcal{N} = \langle V, E, l \rangle$ a term network, $\mathcal{A}$ be a join-semilattice ordered semigroup and $\vartheta : \operatorname{Var} \to {\mathcal{A}}$ a valuation.
 Let us define the prenetwork $\mathcal{N}^{\vartheta}$ with the same edges and vertices with labelling $l^{\vartheta}(x, y) = \{ t^{\vartheta} \: | \: t \in l_{\mathcal{N}}(x, y)\}$.
@@
 Let us consider $\vartheta$ as a homomorphism $F(\omega) \to \mathcal{A}$.
-We would like to know whether ${\uparrow t^{\vartheta}} \subseteq l^{\vartheta}(x,y)$ for $t^{\vartheta} \in l^{\vartheta}(x,y)$.
+We would like to know whether $\up{t^{\vartheta}} \subseteq l^{\vartheta}(x,y)$ for $t^{\vartheta} \in l^{\vartheta}(x,y)$.
 So $t \in l(x,y)$.
@@
   \item Let $x, y \in \operatorname{Nodes}(\mathcal{N})$ and $t$ a term. $\mathcal{N}_c$ is the extension of $\mathcal{N}$,
-  where $\operatorname{Nodes}(\mathcal{N}_c) = \operatorname{Nodes}(\mathcal{N})$ and $l_{\mathcal{N}}(x, y) \cup {\uparrow t}$ and $l_{\mathcal{N}_c}(u, v) = l_{\mathcal{N}}(u, v)$ for each $u \neq x$ and $v \neq y$.
+  where $\operatorname{Nodes}(\mathcal{N}_c) = \operatorname{Nodes}(\mathcal{N})$ and $\up{l_{\mathcal{N}}(x, y) \cup t}$ and $l_{\mathcal{N}_c}(u, v) = l_{\mathcal{N}}(u, v)$ for each $u \neq x$ and $v \neq y$.
 We denote this network as $\mathcal{N}_c(\mathcal{N}, x, y, t)$.
@@
 Let us define a network $T$ such that $\operatorname{Nodes}(T) = \{ x, y, z\}$.
-We define labelling as $l_{T}(x, y) = {\uparrow t_1}$ and $l_{T}(y, z) = {\uparrow t_2}$.
+We define labelling as $l_{T}(x, y) = \up{t_1}$ and $l_{T}(y, z) = \up{t_2}$.
 So we put $\mathcal{N}_w = \mathcal{N} \cup T$.
@@
   \item Let $x, y \in \operatorname{Nodes}(\mathcal{N})$ and let $t_1, t_2$ be terms
-  Let us define $T_i = \langle \{ x, y\}, \{ (x, y) \}, l_{T_i} \rangle$, where $l_{T_i}(x, y) = \uparrow (l_{\mathcal{N}}(x, y) \cup t_i)$ for $i = 1,2$.
+  Let us define $T_i = \langle \{ x, y\}, \{ (x, y) \}, l_{T_i} \rangle$, where $l_{T_i}(x, y) = \up{(l_{\mathcal{N}}(x, y) \cup t_i)}$ for $i = 1,2$.
 So $\mathcal{N}_{j_1} = \mathcal{N} \cup T_1$ and $\mathcal{N}_{j_2} = \mathcal{N} \cup T_2$.
@@
 \end{enumerate}
-
-\newpage
 
@@
   \end{center}
-  where $\mathcal{A}$ is a join semilattice-ordered semigroup and $\vartheta : V \to \mathcal{A}$ is a variable assignment.
+  where $\mathcal{A}$ is a join semilattice-ordered semigroup and $\vartheta : \operatorname{Var} \to \mathcal{A}$ is a variable assignment.
 
-  We define the sequence of formulas $\{ \sigma_n \}_{n < \omega}$ recursively:
+  We define the following sequence of formulas $\{ \sigma_n \}_{n < \omega}$ inductively:
   \begin{enumerate}
@@
       \item ${\sigma_{n + 1}}_c(\mathcal{N}, v) = \bigwedge \limits_{x, y, z \in \operatorname{Nodes}(\mathcal{N}), t_1 \in l_{\mathcal{N}}(x, y), t_2 \in l_{\mathcal{N}}(y, z)} \sigma_{n}(\mathcal{N}_c(x, z, t_1, t_2), v)$
-      \item ${\sigma_{n + 1}}_w(\mathcal{N}, v) = \bigwedge \limits_{x,y \in \operatorname{Nodes}(\mathcal{N}), t \in l_{\mathcal{N}}(x,y)} \forall u_1, u_2 (t \leq u_1 ; u_2 \rightarrow \bigvee \mathcal{N}_c(x,y,z, t_1, t_2))$, where $z$ is a fresh node.
+      \item ${\sigma_{n + 1}}_w(\mathcal{N}, v) = \bigwedge \limits_{x,y \in \operatorname{Nodes}(\mathcal{N}), t \in l_{\mathcal{N}}(x,y)} \forall u_1, u_2 (t \leq u_1 ; u_2 \rightarrow \bigvee \limits_{w \in \operatorname{Nodes} \cup \{ z \}} \mathcal{N}_c(x,y,w, t_1, t_2))$, where $z$ is a fresh node.
       \item ${\sigma_{n + 1}}_j(\mathcal{N}, v) = \bigwedge \limits_{x, y \in \operatorname{Nodes}(\mathcal{N})} \sigma_{n}(\mathcal{N}_{j_1}(\mathcal{N}, x, y, t_1, t_2), v) \lor \sigma_{n}(\mathcal{N}_{j_2}(\mathcal{N}, x, y, t_1, t_2), v)$
@@
 
-  So, $\exists$ has a winning strategy iff these formulas are true under the valuation $\vartheta$ since the formulas $\{ \sigma_n \}_{n < \omega}$ encode the presence of a winning strategy for $\exists$ on every finite round.
+  So, $\exists$ has a winning strategy iff these formulas are true under the valuation $\vartheta$ since the formulas $\{ \sigma_n \}_{n < \omega}$ encode the presence of a winning strategy for $\exists$ on every finitary round.
 \end{proof}
@@
 
+One may consider $\mathcal{A}$, a finite join semilattice-ordered semigroup, as a quantale since the underlying semilattice is complete being finite.
+By Theorem~\label{quantaleRep}, $\mathcal{A}$ is isomorphic to the relational quantale $\widehat{A}$.
+From the other hand, if $\mathcal{Q}$ is a relational quantale, then $\Sigma A \subseteq \bigcup A$ for $A \subseteq \mathcal{Q}$.
+The converse inclusion does not hold in general, see \cite[Remark 3.2 and Example 3.6]{brown1993representation}.
+
+Let $\mathcal{Q}$ be a quantale and $q \in \mathcal{Q}$ is called \emph{completely coprime}, if
+\begin{center}
+  $q \leq \sum \limits_{i \in I} q_i$ implies $q \leq q_i$ for some $i \in I$.
+\end{center}
+We denote the set of completely coprimes as $\operatorname{CCP}(\mathcal{Q})$.
+In particular, if $|I| = 2$, then $q$ is called join-prime.
+If a quantale is finite, then $\operatorname{CCP}(\mathcal{Q})$ is clearly the set of all join-prime elements that we denote as $\operatorname{JP}(\mathcal{Q})$.
+
+A quantale $\mathcal{Q}$ is called \emph{completely coprime algebraic} (CCPA), if every $q \in \mathcal{Q}$ has the form:
+\begin{center}
+  $q = \sum \limits \{ p \in \mathcal{Q} \: | \: p \in \operatorname{CCP}(\mathcal{Q}), p \leq q \}$
+\end{center}
+
+A quantale $\mathcal{Q}$ is called \emph{simple}, if
+$\sum \limits_{i \in I} \widehat{A} = \bigcup \limits_{i \in A} \widehat{A}$ for $A \subseteq \mathcal{Q}$.
+
+It is known that if a quantale $\mathcal{Q}$ is CCPA, then $\operatorname{CCP}(\mathcal{Q})$ is a set of generators, and every CCPA quantale is simple, see \cite[Lemma 5.3, Lemma 5.5, and Proposition 5.6]{brown1993representation}.
+As a corollary, any quantale CCPA is isomorphic to a relational quantale where all joins are unions.
+
+Let $\mathcal{A}$ be a join semilattice, a subset $I \subseteq \mathcal{A}$ is called \emph{ideal}, if the following hold:
+\begin{enumerate}
+  \item If $x \in I$, then $\down{x} \subseteq I$,
+  \item $x, y \in I$ implies $x + y \in I$.
+\end{enumerate}
+The poset $\langle \operatorname{Id}(\mathcal{A}), \subseteq \rangle$ consists of all ideals of $\mathcal{A}$ ordered by inclusion.
+$\mathcal{A}$ a join semilattice is called \emph{distributive}, if for each $x, a, b \in \mathcal{A}$ such that $x \leq a + b$, so there are $a' \leq a$ and $b' \leq b$ such that $x = a' + b'$.
+A finite $\mathcal{A}$ is distributive is the lattice $\operatorname{Id}(\mathcal{A})$ is distributive in the usual sense, see \cite[Lemma 184]{gratzer2011lattice}.
+
+\begin{proposition} \label{distr}
+  Let $\mathcal{A} \in {\bf R}(;,+)$ and $\mathcal{A}$ is finite, then
+  \begin{enumerate}
+    \item $\mathcal{A}$ is distributive,
+    \item Every generator of $\mathcal{A}$ as a quantale is join-prime.
+  \end{enumerate}
+\end{proposition}
+
+\begin{proof}
+$ $
+
+  \begin{enumerate}
+    \item
+    $\mathcal{A}$ is finite, so the $\operatorname{Id}(\mathcal{A})$ consists of principal ideals, that is, lower cones generated by elements of $\mathcal{A}$.
+
+    \begin{center}
+    $\operatorname{Id}(\mathcal{A}) = \{ \down{a} \: | \: a \in \mathcal{A} \}$
+    \end{center}
+
+    Let $h$ be a representation of $\mathcal{A}$, consider $h(\mathcal{A})$ and its set of ideals $\operatorname{Id}(h(\mathcal{A}))$.
+Intersection of ideals is defined is usual since intersection of ideals is an ideal, join is the defined as follows:
+
+    \begin{center}
+    $\down{h(a)} \vee \down{h(b)} = \down{(h(a) \cup h(b))}$
+    \end{center}
+    Note that $\down{(h(a) \cup h(b))} = \down{h(a + b)}$ since $h$ is a representation of $\mathcal{A}$.
+
+    Let $c \in \mathcal{A}$, consider $h(c)$.
+
+    Suppose $h(d) \in \down{h(c)} \cap \down{h(a + b)}$.
+    $h(d) \in \down{h(a + b}$ implies $h(d) \subseteq h(a + b)$.
+
+    Then if $(x, y) \in h(d)$, then either $(x, y) \in h(a)$ or $(x, y) \in h(b)$.
+    If $(x, y) \in h(a)$ (or $h(b)$), then we have $h(d) \in \down{(h(c) \cap h(a))}$ (or $\down{(h(c) \cap h(b))}$).
+
+    Suppose $h(d) \in (\down{h(c)} \cap \down{h(a)}) \vee (\down{h(c)} \cap \down{h(b)})$.
+So, we have either $h(d) \in \down{h(c)} \cap \down{h(a)}$ or $h(d) \in \down{h(c)} \cap \down{h(b)}$.
+So $h(d) \in \down{h(c)}$ anyway.
+If $h(d) \in \down{h(a)}$, then $d \leq a \leq a + b$, so $h(d) \in \down{h(a + b)}$.
+    The second alternative is the same as the first one.
+
+    So, $\operatorname{Id}(h(\mathcal{A}))$ is distributive, and, therefore, $\mathcal{A}$ is distibutive as a join-semilattice by \cite[Lemma 184]{gratzer2011lattice}.
+
+    \item Let $g \in \mathcal{A}$ be a generator, suppose $g \leq a + b$, so there are $a' \leq a$ and $b' \leq b$ with $g = a' + b'$.
+$g$ is join-irreducible, thus, one has either $g = a'$ or $g = b'$.
+In fact, this statement generalises the fact that an element is join-prime iff it is join-irreducible in distributive lattices, see \cite[Lemma 5.11]{davey2002introduction}.
+  \end{enumerate}
+\end{proof}
+
+The following lemma allows one to consider finite elements of ${\bf R}(;,+)$ as simple quantales.
+If $\mathcal{A}$ is finite and representable, then every $a$ has the form $\Sigma \{ g \in \mathcal{G}(\mathcal{A}) \: | \: g \leq a \}$, but every generator is join-prime by Proposition~\ref{distr}.
+So we have:
+
+\begin{proposition}\label{quantalerepisok}
+  Let $\mathcal{A}$ be a finite representable join-semilattice semigroup, $\mathcal{A}$ is completely coprime algebraic as a quantale and, therefore, is simple.
+\end{proposition}
+
+As a corollary, we have the following theorem:
+
 \begin{theorem}
@@
 \begin{proof}
-  Let $\mathcal{A} \in {\bf R}(+, ;)$ such that $|\mathcal{A}| < \omega$ and $h$ a representation of $\mathcal{A}$. monkahmm
+  Let us show that if $\mathcal{A} \in {\bf R}(+, ;)$ such that $\mathcal{A}$ is finite, then
+  $h(\mathcal{A})$ is isomorphic to some subalgebra of $\widehat{A}$. Consider the map
+  $f : h(a) \mapsto \hat{a}$.
+$f$ preserves composition since $h$ and $\hat{.}$ are semigroup homomorphisms.
+The equality $f(h(a) \cup h(b)) = \hat{a} \cup \hat{b}$ holds since $\mathcal{A}$ is simple as a quantale by Proposition~\ref{quantalerepisok}.
+
+  $\mathcal{A}$ is a relational quantale on the underlying of $\mathcal{A}$, but $\mathcal{A}$ is finite, so $\mathcal{A}$ has a representation over a finite base.
 \end{proof}
 
-The following corollary follows from finite representability and finite axiomatisability.
+In turn, the finite representation property and the presence of a recursively enumerable axiomatisation imply the following corollary.
 
 \begin{corollary}
-  The class ${\bf R}_{fin}(;, +) = \{ \mathcal{A} \in {\bf R}(;, +) \: | \: |\mathcal{A}| < \omega \}$ is recursive. That is, the problem of membership for this class is decidable.
+  The problem of membership to ${\bf R}(;, +)$ is decidable for finite structures.
 \end{corollary}
@@
 \begin{proof}
-  Let $\mathcal{A}$ be a finite semilattice-ordered semigroup. We check axioms with the algorithm that yields ``no'', if $\mathcal{A}$ is not representable. So,
+  Let $\mathcal{A}$ be a finite semilattice-ordered semigroup. We check axioms defined in Lemma~\ref{ax} with the algorithm $S$ that yields ``no'' and halts, if $\mathcal{A}$ is not representable.
+
   In parallel, we check whether $\mathcal{A}$ has a finite representation.
-If there exists such $n < \omega$ such that $\mathcal{A}$ is representable over size $n$.
-Both sets are domains of corresponding computable functions, so they are enumerable.
-That makes ${\bf R}_{fin}(;, +)$ decidable by Post's theorem.
+If there exists such $0 < n < \omega$ such that $\mathcal{A}$ is representable over size $n$, then the algorithm $T$ does output ``yes'' and terminates.
+
+  So, one has either $S(\mathcal{A}) = \text{no}$ and $T(\mathcal{A}) = \text{yes}$.
 \end{proof}
 
-\begin{comment}
 \section{Acknowledgements}
 
-The author is sincerely grateful to Robin Hirsch, Ian Hodkinson, Stepan Kuznetsov, Valentin Shehtman, Eugeny Zolin, and his supervisor Ilya Shapirovsky for valuable conversations.\\The previous version of this paper contained an error.
+The author is sincerely grateful to Robin Hirsch, Ian Hodkinson, and his supervisor Ilya Shapirovsky for valuable comments.
+The previous version of this paper contained an error.
 Many thanks to Jas Semrl for noticing that.
-\end{comment}
 

### 2021-03-12 problem-5/Text.tex
the finite representation property for join semilattice-ordered semigroups
[Problem-5.18] the definition of a network and a game

--- 2020-11-29 problem-9.4/Text.tex
@@
 
-\begin{defin}
-  $ $
-
-    A relation algebra is an algebra $\mathcal{R} = \langle R, 0, 1, +, -, ;, {}^{\smile}, {\bf 1 }' \rangle$ such that $\langle R, 0, 1, +, - \rangle$ is a Boolean algebra and the following
-    equations hold, for each $a, b, c \in R$:
-    \begin{enumerate}
-      \item $a ; (b ; c) = (a ; b) ; c$
-      \item $(a + b) ; c = (a ; c) + (b ; c)$
-      \item $a ; {\bf 1}' = a$
-      \item $a^{\smile \smile} = a$
-      \item $(a + b)^{\smile} = a^{\smile} + b^{\smile}$
-      \item $(a ; b)^{\smile} = b^{\smile} ; a^{\smile}$
-      \item $a^{\smile} ; (- (a ; b)) \leq - b$
-    \end{enumerate}
-    where $a \leq b$ iff $a + b = b$. ${\bf RA}$ denotes the class of all relation
-    algebras.
-\end{defin}
-
 Cylindric algebras provide a generalisation of relation algebras for relations of an arbitrary arity.
@@
 
+One may define a representation explicitly in the following way:
+
+\begin{defin}
+  Let $\mathcal{A}$ be a cylindric algebra of dimension $\alpha$.
+A representation of $\mathcal{A}$ over the non-empty domain $X$ is a map $f : \mathcal{A} \hookrightarrow 2^{\prescript{\alpha}{}U}$ such that:
+  \begin{enumerate}
+    \item $f(1) = \bigcup \limits_{i \in I} \prescript{\alpha}{}X_i$ for some disjoint family $\{X_i\}_{i \in I}$ where each $X_i \subseteq X$
+    \item $h : \mathcal{A} \to 2^{f(1)}$ is a representation of a Boolean reduct
+    \item for all $\lambda, \eta < \alpha$, $x \in h(d_{\lambda \eta})$ iff $x_{\lambda} = x_{\eta}$
+    \item for all $\lambda < \alpha$ and $a \in \mathcal{A}$, $x \in h(c_{\lambda}(a))$ iff there is $y \in X$ such that $x[\lambda \mapsto y] \in h(a)$
+  \end{enumerate}
+\end{defin}
+
 \begin{defin}
@@
 Let $\mathcal{A} \in {\bf C}_{\omega}$, then $\mathcal{A}$ has a \emph{complete representation}, if this representation preserves all existing suprema.
+In other words, $\mathcal{A}$ is completely representable.
+
+\begin{defin} Atom
+\end{defin}
 
@@
 
+\section{The result}
+
+\begin{lemma}\label{Neat}
+  Let $\mathcal{A}$ be a BAO of type ${\bf CA}_{\alpha}$ and $\mathcal{B}$ be a $\beta$-dimensional cylindric algebra such that $\beta \leq \alpha$ and $\mathcal{A}$ neatly embeds to $\mathcal{B}$ by a complete embedding.
+
+  \begin{enumerate}
+  \item $\mathcal{A}^{+}$ neatly embeds to $\mathcal{B}^{+}$ by a complete embedding.
+  \item $\mathcal{A}$ is atomic.
+  \end{enumerate}
+\end{lemma}
+
+\begin{proof}
+$ $
+
+\begin{enumerate}
+  \item See \cite[Remark 2.7.25]{henkin1971cylindric}.
+  \item Is it true?
+\end{enumerate}
+\end{proof}
+
+\begin{lemma}[See \cite{hirsch1997complete}]
+  A Boolean algebra is compeletely representable iff it has an atomic representation.
+\end{lemma}
+
+\begin{defin}
+  Let $\mathcal{A}$ be a BAO of type ${\bf CA}_{\omega}$, an $\mathcal{A}$-pre-network is a pair
+  $\mathcal{N} = \langle N, l \rangle$, where $N$ is a set of nodes and $l : \prescript{\omega}{}N \to \operatorname{At}(\mathcal{A})$.
+
+  $\mathcal{N}$ is a network, if the following conditions hold, for all $x, y \in \prescript{\omega}{}N$ and $i, j < \omega$:
+  \begin{enumerate}
+    \item $l(x) \leq d_{ij}$ iff $x(i) = x(j)$
+    \item $x \equiv_i y$ implies $l(x) \leq c_i l(y)$
+  \end{enumerate}
+\end{defin}
+
+Let $\mathcal{N}_1 = \langle N_1, l_1 \rangle$ and $\mathcal{N}_2 = \langle N_2, l_2 \rangle$ be networks, then $\mathcal{N}_1 \subseteq \mathcal{N}_2$ if $N_1 \subseteq N_2$ and $l_1 = l_2 	\upharpoonright_{N_1}$.
+
+Let $\Lambda \in \operatorname{Lim}$ and $\{ \mathcal{N}_{\lambda}\}_{\lambda < \Lambda}$ a sequence of networks such that
+\begin{center}
+  $\langle N_0, l_0 \rangle \subseteq \langle N_1, l_1 \rangle \subseteq \dots \langle N_{\lambda}, l_{\lambda} \rangle \subseteq \dots $ for $\lambda < \Lambda$
+\end{center}
+then \emph{the limit} of the sequence $\{ \mathcal{N}_{\lambda}\}_{\lambda < \Lambda}$ is the network
+\begin{center}
+  $\mathcal{N} = \langle N, l \rangle = \bigcup \limits_{\lambda < \Lambda} \langle N_{\lambda}, l_{\lambda} \rangle$
+\end{center}
+with nodes $N = \bigcup \limits_{\lambda < \Lambda} N_{\lambda}$ and labelling $l = \bigcup \limits_{\lambda < \Lambda} l_{\lambda}$, that is,
+  for any $\lambda \in \Lambda$ and $x \in \prescript{\omega}{}N$ one has $l(x) = l_{\lambda}(x)$.
+
+The elements of $\prescript{\omega}{}N$ are called \emph{$\omega$-dimensional hyperedges} of a network. One may identify a complete representation of an atomic cylindric-type algebra $\mathcal{A}$
+with a set $\{ \mathcal{N}_a \: | \: a \in \operatorname{At}(\mathcal{A}) \}$ of $\mathcal{A}$-networks with the following additional condition:
+\begin{itemize}
+  \item For each $a \in \operatorname{At}(\mathcal{A})$ there exists $x \in \prescript{\omega}{}N_a$ such that $l_a(x) = a$ and for each $z \in \prescript{\omega}{}N_a$ and $b \in \operatorname{At}(\mathcal{A})$, $i < \omega$ with $l_a(z) \leq c_i b$
+  there exists $y \in \prescript{\omega}{}N_a$ such that $z \equiv_i y$ and $l_a(y) = b$.
+\end{itemize}
+
+We define a complete representation $h$ of a cylindric-type algebra $\mathcal{A}$ as follows, for any $b \in \mathcal{A}$:
+\begin{center}
+$h(b) = \{ x \: | \: \exists a \in \operatorname{At}(\mathcal{A}), x \in \prescript{\omega}{}N_a, l_a(x) \leq b \}$
+\end{center}
+
+Let us define an atomic game.
+
+\begin{defin}
+  Let $\mathcal{A}$ be an atomic BAO of type ${\bf CA}_{\omega}$ and $\kappa > 0$ a cardinal.
+The game $\mathcal{G}^{\kappa}(\mathcal{A})$ is defined as follows.
+The game has two players: $\forall$ (Abelard) and $\exists$ (H\'{e}lo\"{i}se).
+A play of the game is the sequence
+  \begin{center}
+    $\mathcal{N}_0 \subseteq \mathcal{N}_1 \subseteq \mathcal{N}_2 \subseteq \dots \subseteq \mathcal{N}_{\lambda} \subseteq \dots$ for $\lambda < \kappa$
+  \end{center}
+  The game consists of the following stages:
+  \begin{enumerate}
+    \item ({\bf Zero round})
+
+    $\forall$ picks an atom $a \in \operatorname{At}(\mathcal{A})$ and $\exists$ plays a network $\mathcal{N}_0$.
+If there is no $x \in \prescript{\omega}{}N_0$ such that $l_0(x) = a$, then $\forall$ wins the play.
+    \item ({\bf Successor round})
+
+    Let $0 < \lambda$ be a cardinal such that $\lambda + 1 < \kappa$ and $\mathcal{N}_{\lambda} = \langle N_{\lambda}, l_{\lambda} \rangle$ is a network that has been played.
+
+    $\forall$ picks $i < \omega$, $x \in \prescript{\omega}{}N_{\lambda}$, $a \in \operatorname{At}$ such that $l_{\lambda}(x) \leq c_i a$.
+    We denote this move as $(i, x, a)$.
+$\exists$ responds with a network $\mathcal{N}_{\lambda + 1} \supseteq \mathcal{N}_{\lambda}$.
+$\forall$ wins, if there is no node $c \in N$ such that $l_{\lambda+1}(x[i/c]) = a$, then $\forall$ wins
+    \item The limit of the play is defined as $\bigcup \limits_{\lambda < \kappa} \mathcal{N}_{\lambda}$.
+$\forall$ wins the play, if there exists $\kappa_1 < \kappa$ such that $\exists$ does not win the $\kappa_1$th-round.
+Otherwise, $\exists$ wins the play.
+  \end{enumerate}
+\end{defin}
+
+\begin{theorem}
+  Let $\mathcal{A}$ be an atomic $\omega$-dimensional cylindric-type algebra and $\kappa$ a cardinal such that $|\operatorname{At}(\mathcal{A})| = \kappa$, then the following are equivalent:
+
+  \begin{enumerate}
+    \item $\mathcal{A}$ is completely representable.
+    \item $\exists$ has a winning strategy in $\mathcal{G}^{\kappa + \omega}$.
+  \end{enumerate}
+\end{theorem}
+
+\begin{proof}
+  $ $
+
+  \begin{enumerate}
+    \item $\Rightarrow$
+    If $\mathcal{A}$ is completely representable, then its Boolean reduct is representable as well. $\exists$ maintains that embedding to win the play.
+
+    \item $\Leftarrow$
+
+    Suppose $\exists$ has a winning strategy in $\mathcal{G}^{\kappa + \omega}(\mathcal{A})$.
+    In every round $\forall$ picks all possible $i < \omega$, $a \in \operatorname{At}(\mathcal{A})$, all possible hyperedges and all appropriate atoms.
+    ...
+    Question: Is this claim correct?
+  \end{enumerate}
+\end{proof}
+
+\begin{theorem}[This assumption is by Ian Hodkinson]\label{Neat2}
+  $ $
+
+  Let $\mathcal{A}$ be a BAO of type ${\bf CA}_{\omega}$ such that $\mathcal{A}$ neatly embeds into ${\bf CA}_{\omega + \omega}$ by a complete embedding.
+Then $\mathcal{A}$ is completely representable as ${\bf CA}_{\omega}$.
+\end{theorem}
+
+\begin{proof}
+  Suppose $\mathcal{A} \subseteq \mathfrak{Nr}_{\omega} \mathcal{B}$, where $\mathcal{B} \in {\bf RCA}_{\omega + \omega}$ and the inclusion map $\rho : \mathcal{A} \hookrightarrow \mathfrak{Nr}_{\omega} \mathcal{B}$ is a complete embedding, that is:
+  \begin{center}
+    $\rho (\sum \limits_{i \in I} a_i) = \sum \limits_{i \in I} (\rho a_i)$, if $\sum \limits_{i \in I} a_i$ exists.
+  \end{center}
+
+  Let us show that $\mathcal{A}$ is atomic.
+
+  Consider $\rho(\mathcal{A})$. Let us show that $\exists$ has a winning strategy on $\mathcal{G}^{\kappa + \omega}(\rho(\mathcal{A}))$
+\end{proof}
+
+Lemma~\ref{Neat} and Theorem~\ref{Neat2} imply the following theorem.
+
+\begin{theorem}
+  Let $\mathcal{C} \in {\bf RCA}_{\omega}$, then $\mathcal{C}^{+} \in {\bf RCA}_{\omega}$. That is, ${\bf RCA}_{\omega}$ is closed under canonical extensions.
+\end{theorem}
+
+\begin{proof}
+
+\end{proof}
+
+\section{Sahlqvist and canonical axiomatisation of ${\bf CA}_{\omega}$}
+
+\section{Notes on the canonicity of ${\bf RRA}$}
+
+\begin{defin}
+  $ $
+
+    A relation algebra is an algebra $\mathcal{R} = \langle R, 0, 1, +, -, ;, {}^{\smile}, {\bf 1 }' \rangle$ such that $\langle R, 0, 1, +, - \rangle$ is a Boolean algebra and the following
+    equations hold, for each $a, b, c \in R$:
+    \begin{enumerate}
+      \item $a ; (b ; c) = (a ; b) ; c$
+      \item $(a + b) ; c = (a ; c) + (b ; c)$
+      \item $a ; {\bf 1}' = a$
+      \item $a^{\smile \smile} = a$
+      \item $(a + b)^{\smile} = a^{\smile} + b^{\smile}$
+      \item $(a ; b)^{\smile} = b^{\smile} ; a^{\smile}$
+      \item $a^{\smile} ; (- (a ; b)) \leq - b$
+    \end{enumerate}
+    where $a \leq b$ iff $a + b = b$. ${\bf RA}$ denotes the class of all relation
+    algebras.
+\end{defin}
+
 We will adapting the following proof of the fact that ${\bf RRA}$ is canonical \footnote{This idea is by Ian Hodkinson} to our case.
@@
 \begin{enumerate}
-  \item A relation algebra $\mathcal{A}$ is representable iff $\mathcal{A}$ neatly embeds to some omege-dimensional cylinric algebra,
+  \item A relation algebra $\mathcal{A}$ is representable iff $\mathcal{A}$ neatly embeds to some $\omega$-dimensional cylinric algebra,
   \item If $\mathcal{A}$ neatly embeds in $\mathcal{A}$ then $\mathcal{A}^{+}$ neatly embeds in $\mathcal{B}^{+}$,
@@
 
-The following characterisation results are by Henkin, Monk, and Tarski \cite{Henkin1988-HENCAP-4} as well:
+The following characterisation results are by Henkin, Monk, and Tarski \cite[5.3.13, 5.3.16]{Henkin1988-HENCAP-4} as well:
 \begin{theorem} \label{char}
@@
 
-\begin{lemma}[Henkin, Monk, Tarski]\label{Neat}
-  $ $
-
-  Let $\mathcal{A}$ be a BAO of type ${\bf CA}_{\alpha}$ and $\mathcal{B}$ be a $\beta$-dimensional cylindric algebra such that $\beta \leq \alpha$ and $\mathcal{A}$ neatly embeds to $\mathcal{B}$ by a complete embedding.
-Then $\mathcal{A}^{+}$ neatly embeds to $\mathcal{B}^{+}$ by a complete embedding.
-\end{lemma}
-
-\begin{proof}
-\end{proof}
-
-\begin{theorem}[This assumption is by Ian Hodkinson]\label{Neat2}
-  $ $
-
-  Let $\mathcal{A}$ be a BAO of type ${\bf CA}_{\omega}$ such that $\mathcal{A}$ neatly embeds into ${\bf CA}_{\omega + \omega}$ by a complete embedding.
-Then $\mathcal{A}$ is completely representable as ${\bf CA}_{\omega}$.
-\end{theorem}
-
-\begin{proof}
-  Hmmmmm, I believe so.
-\end{proof}
-
-Lemma~\ref{Neat} and Theorem~\ref{Neat2} imply the following theorem.
-
-\begin{theorem}
-  Let $\mathcal{C} \in {\bf RCA}_{\omega}$, then $\mathcal{C}^{+} \in {\bf RCA}_{\omega}$. That is, ${\bf RCA}_{\omega}$ is closed under canonical extensions.
-\end{theorem}
-
-\begin{proof}
-  :monkahmm:
-\end{proof}
-
-\begin{comment}
-\subsection{Substitution in cylindric algebras}
-
-\begin{prop} Let $\alpha$ be an ordinal and let $i, j, k, l < \alpha$. The following facts hold in ${\bf CA}_{\alpha}$
-
-  \begin{enumerate}
-    \item $s^{i}_j x \leq c_i x$.
-    \item $s^{i}_j (x \land y) = s^{i}_j x \land s^{i}_j y$, $s^{i}_j (x \lor y) = s^{i}_j x \lor s^{i}_j y$, $- s^{i}_j x = s^{i}_j (- x)$.
-    Moreover, $s^{i}_j$ is completely additive.
-    \item $i \neq k, l$ implies $s^i_j d_{ik} = d_{jk}$ and $s^i_j d_{kl} = d_{kl}$.
-    \item $d_{jk} \land s^{i}_j = d_{jk} \land s^{i}_k$.
-    \item $s^i_j c_i x = c_i x$.
-    \item $k \neq i,j$ implies $s^i_j c_i x = c_i s^i_j x$.
-    \item $c_j s^i_j x = c_i s^j_i x$.
-    \item $i \neq j$ implies $c_i s^i_j x = s^i_j x$.
-    \item $i \neq k$ implies $s^i_j s^i_k = s^i_k x$.
-    \item If either $i \notin \{ k, l\}$ and $k \notin \{ i, j\}$, or $j = l$, then $s^i_j s^k_l x = s^k_l s^i_j x$.
-    \item $s^i_j s^j_i x = s^i_j x$.
-    \item $s^i_k s^j_i x = s^i_k s^j_k x = s^j_k s^i_j x$
-  \end{enumerate}
-\end{prop}
-
-\section{Model-theoretic and universal algebraic preliminaries}
-
-\subsection{Ultraproducts}
-
-Here are the required notions and facts from model theory and universal algebra \cite{hirsch2002relation} \cite{hodges1997shorter} \cite{sankappanavar1981course}.
-
-Let $\Lambda$ be an index set and $D$ an ultrafilter on the Boolean algebra $\langle \mathcal{P}(\Lambda), \cup, -, \Lambda, \emptyset \rangle$.
-Consider the product $M = \prod \limits_{\lambda \in \Lambda} M_{\lambda}$ of the $\Omega$-structures $\{ M_{\lambda}\}_{\lambda \in \Lambda}$ and the equivalence relation on $\operatorname{dom}(M)$ defined as
-\begin{center}
-  $a_1 \sim a_2 \Leftrightarrow \{ \lambda \in \Lambda \: | \: a_1(\lambda) = a_2(\lambda)\} \in D$
-\end{center}
-Let us denote $\operatorname{dom}(M) / \sim$ as $U$ and $[a]_{\sim}$ as $a / D$, where $a \in \operatorname{dom}(M)$.
-We also denote the \emph{ultraproduct} of $\{ M_{\lambda} \}_{\lambda}$ as $\prod \limits_{\lambda \in \Lambda} M_{\lambda} / D$, or,
-for brevity, as $\prod_{D} M_{\lambda}$. The $\Omega$-symbols have the following interpretation
-\begin{enumerate}
-  \item If $c \in \operatorname{Cnst}$, then $c^{U} = c^{M} / D$
-  \item If $f \in \operatorname{Fn}$ is an $n$-ary function symbol and $\overline{a} \in M^{n}$, then $f^{U}(\overline{a}) = f^{M}(x) = f^{M}(\overline{a}) / D$
-  \item If $R \in \operatorname{Fn}$ is an $n$-ary relation symbol and $\overline{a} \in M^{n}$, then $U \models R(\overline{a} / D)$ iff
-  $\{ \lambda \in \Lambda \: | \: M_{\lambda} \models R(\overline{a} (\lambda))\} \in D$
-\end{enumerate}
-The ultraproduct is principal if $D$ is a principal filter.
-
-\begin{defin}
-$ $
-
-  \begin{enumerate}
-    \item Let $\{ M_{\lambda}\}_{\lambda \in \Lambda}$ be a set of $\Omega$-structures such that every
-    $M_{\lambda}$ is isomorphic to the single structure $M$, then their ultraproduct over $D$ is called the ultrapower over $D$.
-    The denotation is $\prod \limits_{D} M$ or $M^{\Lambda} / D$.
-    \item If $\prod \limits_{D} M \cong N$ for some structure $N$, then $M$ is an ultraroot of $N$.
-  \end{enumerate}
-\end{defin}
-
-\begin{theorem}[\L{}os]
-
-  Let $\{ M_{\lambda}\}_{\lambda \in \Lambda}$ be $\Omega$-structures and $D$ an ultrafilter on $\Lambda$, and let
-  $U = \prod \limits_{D} M_{\lambda}$ be an ultraproduct of $\{ M_{\lambda}\}_{\lambda \in \Lambda}$ over $D$. For each first-order formula
-  $\varphi(x_1, \dots, x_n)$ and for each $a_1 / D, \dots, a_n / D \in U$:
-
-\begin{center}
-  $U \models \varphi(a_1 / D, \dots, a_n / D)$ iff $\{ \lambda \in \Lambda \: | \: \varphi(a_1(\lambda), \dots, a_n(\lambda))\} \in D$
-\end{center}
-\end{theorem}
-
-The \L{}os has the following helpful corollary:
-
-\begin{col}
-  Let $\prod \limits_{D} M$ be an ultrapower of $M$. For $a \in M$, let us define a function $\overline{a} : a \mapsto a / D$.
-  Then such a map is an elementary embedding of $M$ into $\prod \limits_{D} M$.
-\end{col}
-
-Moreover, any elementary equivalent structures have isomorphic ultrapowers.
-
-Recall that a class of $\Omega$-structures ${\bf K}$ is called \emph{elementary}, if ${\bf K} = \operatorname{Mod}(\operatorname{T})$ for some first-order theory ${\bf T}$.
-In that case, $\operatorname{T}$ is an axiomatisation of ${\bf K}$.
-
-\begin{theorem}
-  Let ${\bf K}$ be a class $\Omega$-structures,
-  ${\bf K}$ is elementary iff ${\bf K}$ is closed under isomorphic copies, ultraroots, and ultrapowers.
-\end{theorem}
-
-\section{${\bf IG}_{\omega}$ and ultraproducts}
-
-\section{${\bf IG}_{\omega}$ is (not) (pseudo-)elementary; is (not) a variety}
-\end{comment}
-
 \bibliographystyle{plain}

### 2021-03-13 frp-paper/Paper.tex
[Problem-5.18] the definition of a network and a game
important fixes in frp-paper/Paper.tex

@@
 We also show that the class of representable join semilattice-ordered semigroups has a recursively
-enumerable recursive axiomatisation and the finite representation property.
+enumerable axiomatisation and the finite representation property.
 \end{abstract}
@@
 
-We will use the following notation due to, for example, \cite{hirsch2011positive}.
 Let $\tau$ be a subset of operations and predicates definable in ${\bf RA}$.
@@
 
-A quantale is a complete lattice-ordered semigroup.
-Quantales has been initially introduced by Mulvey to provide a noncommutative generalisation of locales, see \cite{mulvey1986suppl}.
-The rigorous definition is the following:
+A quantale is a complete lattice-ordered semigroup. Quantales has been initially introduced by Mulvey to provide a noncommutative generalisation of locales, see \cite{mulvey1986suppl}.
 
@@
 
-As a result, we have the following representation theorem.
+As a result, we have the following representation theorem \cite[Corollary 2]{goldblatt2006kripke}.
 
@@
   \begin{enumerate}
-    \item For every $u, v \in V$ and for every $x,y,z \in \mathcal{A}$ $z \in l(u, v)$ and $z \leq x \: ; \: y$ implies $x \in N(u, w)$ and $y \in l(w, v)$ for some $w \in V$.
-    \item For every $u, v, w \in V$ one has $l(u, v) ; l(v, w) \subseteq l(u, w)$.
-    \item For every $u, v \in V$ $l(u,v)$ is join-prime. That is, for each $a, b \in \mathcal{A}$ if $a + b \in l(u,v)$, then either $a \in l(u,v)$ or $b \in l(u,v)$.
+    \item ({\bf Saturation condition}) For every $u, v \in V$ and for every $x,y,z \in \mathcal{A}$ $z \in l(u, v)$ and $z \leq x \: ; \: y$ implies $x \in N(u, w)$ and $y \in l(w, v)$ for some $w \in V$.
+    \item ({\bf Coherence conidition}) For every $u, v, w \in V$ one has $l(u, v) ; l(v, w) \subseteq l(u, w)$.
+    \item ({\bf Join-primeness}) For every $u, v \in V$ $l(u,v)$ is join-prime. That is, for each $a, b \in \mathcal{A}$ if $a + b \in l(u,v)$, then either $a \in l(u,v)$ or $b \in l(u,v)$.
   \end{enumerate}
@@
 
-    Let us define the $rep : \mathcal{A} \to \operatorname{Nodes}(\mathcal{N})$:
+    Let us define the $\operatorname{rep} : \mathcal{A} \to \operatorname{Nodes}(\mathcal{N})$:
     \begin{center}
-      $rep(a) = \{ (x, y) \: | \: \exists b \in \down{a} \:\: b \in l_{\mathcal{N}}(x, y)\}$
+      $\operatorname{rep}(a) = \{ (x, y) \: | \: \exists b \in \down{a} \:\: b \in l_{\mathcal{N}}(x, y)\}$
     \end{center}
 
-    Let us ensure that $rep$ is a representation.
+    Let us ensure that $\operatorname{rep}$ is a representation.
 
     \begin{itemize}
-  \item Let us show that $rep(a + b) = rep(a) \cup rep(b)$
+  \item Let us show that $\operatorname{rep}(a + b) = \operatorname{rep}(a) \cup \operatorname{rep}(b)$
 
-  Suppose $(x, y) \in rep(a + b)$. That is, there exists $c \leq a + b$ with $c \in l_{\mathcal{N}}(x, y)$, so does $a + b$ since $l_{\mathcal{N}}$ is an upper cone.
+  Suppose $(x, y) \in \operatorname{rep}(a + b)$. That is, there exists $c \leq a + b$ with $c \in l_{\mathcal{N}}(x, y)$, so does $a + b$ since $l_{\mathcal{N}}$ is an upper cone.
 
@@
 $\forall$ makes all possible join moves, so either $a \in l_{\mathcal{N}_{i + 1}}(x, y)$ or $b \in l_{\mathcal{N}_{i + 1}}(x, y)$ depending on what $\exists$ chooses.
-Thus, $(x, y) \in rep(a) \cup rep(b)$.
+Thus, $(x, y) \in \operatorname{rep}(a) \cup \operatorname{rep}(b)$.
 
-Suppose for the converse, $(x, y) \in rep(a)$. We need $(x, y) \in rep(a + b)$.
+Suppose for the converse, $(x, y) \in \operatorname{rep}(a)$. We need $(x, y) \in \operatorname{rep}(a + b)$.
 In other words, we have some $c \in \mathcal{A}$ with $c \leq a$ and $c \in l_{\mathcal{N}}(x, y)$.
-We have $c \leq a \leq a + b$, so $(x, y) \in rep(a + b)$.
+We have $c \leq a \leq a + b$, so $(x, y) \in \operatorname{rep}(a + b)$.
 
-\item Let us show that $rep(a ; b) = rep(a) ; rep(b)$.
+\item Let us show that $\operatorname{rep}(a ; b) = \operatorname{rep}(a) ; \operatorname{rep}(b)$.
 
-Suppose $(x, y) \in rep(a ; b)$. We need some $z$ with $(x, z) \in rep(a)$ and $(z, y) \in rep(b)$.
+Suppose $(x, y) \in \operatorname{rep}(a ; b)$. We need some $z$ with $(x, z) \in \operatorname{rep}(a)$ and $(z, y) \in \operatorname{rep}(b)$.
 
@@
 Her response is a fresh node $z$ such that $l_{\mathcal{N}_{i + 1}}(x, z) =  \up{(l_{\mathcal{N}_i}(x, z) \cup \{ a \})}$ and $l_{\mathcal{N}_{i + 1}}(z, y) = \up{(l_{\mathcal{N}_i}(z, y) \cup \{ b\})}$.
-The inclusion $rep(a ; b) \subseteq rep(a) ; rep(b)$ holds since all witness moves have been played and for every such a move there exists a witness.
+The inclusion $\operatorname{rep}(a ; b) \subseteq \operatorname{rep}(a) ; \operatorname{rep}(b)$ holds since all witness moves have been played and for every such a move there exists a witness.
 
-\item Suppose $(x, y) \in rep(a) ; rep(b)$.
-We need $(x, y) \in rep(a;b)$.
-That is, there exists $z \in \operatorname{Nodes}(\mathcal{N})$ with $(x, z) \in rep(a)$ and $(z, y) \in rep(b)$.
+\item Suppose $(x, y) \in \operatorname{rep}(a) ; \operatorname{rep}(b)$.
+We need $(x, y) \in \operatorname{rep}(a;b)$.
+That is, there exists $z \in \operatorname{Nodes}(\mathcal{N})$ with $(x, z) \in \operatorname{rep}(a)$ and $(z, y) \in \operatorname{rep}(b)$.
 So, there are $c, d$ such that $c \leq a$ with $c \in l_{\mathcal{N}}(x, z)$ and $d \leq b$ with $d \in l_{\mathcal{N}}(z, y)$.
 We also know that $l_{\mathcal{N}}(x, z) ; l_{\mathcal{N}}(z, y) \subseteq l_{\mathcal{N}}(x, y)$ because all composition moves have been played.
-So $c ; d \in l_{\mathcal{N}}(x, y)$. That makes $(x, y) \in rep(a;b)$ since $c ; d \leq a ; b$.
+So $c ; d \in l_{\mathcal{N}}(x, y)$. That makes $(x, y) \in \operatorname{rep}(a;b)$ since $c ; d \leq a ; b$.
 
@@
 
-Let us define the equivalence relation $\approx$ on $T$ as follows for $t_1$ and $t_2$ are terms:
+Let us define the equivalence relation $\approx$ on $T$ as follows for terms $t_1, t_2$:
 \begin{center}
@@
 \begin{enumerate}
-  \item Let $x, y \in \operatorname{Nodes}(\mathcal{N})$ and $t$ a term. $\mathcal{N}_c$ is the extension of $\mathcal{N}$,
+  \item Let $x, y \in \operatorname{Nodes}(\mathcal{N})$ and let $t$ be a term. $\mathcal{N}_c$ is the extension of $\mathcal{N}$,
   where $\operatorname{Nodes}(\mathcal{N}_c) = \operatorname{Nodes}(\mathcal{N})$ and $\up{l_{\mathcal{N}}(x, y) \cup t}$ and $l_{\mathcal{N}_c}(u, v) = l_{\mathcal{N}}(u, v)$ for each $u \neq x$ and $v \neq y$.
@@
   \item
-  Let $x, y \in \operatorname{Nodes}(\mathcal{N})$, $z$ a fresh node, and $t_1$, $t_2$ any terms.
+  Let $x, y \in \operatorname{Nodes}(\mathcal{N})$, let $z$ be a fresh node, and $t_1$, $t_2$ any terms.
 Let us define a network $T$ such that $\operatorname{Nodes}(T) = \{ x, y, z\}$.
@@
 
-One may consider $\mathcal{A}$, a finite join semilattice-ordered semigroup, as a quantale since the underlying semilattice is complete being finite.
+If we extend $\mathcal{A}$, a finite join semilattice-ordered semigroup, with the bottom, then we may consided the obtained algebra as a quantale.
 By Theorem~\label{quantaleRep}, $\mathcal{A}$ is isomorphic to the relational quantale $\widehat{A}$.
@@
 A quantale $\mathcal{Q}$ is called \emph{simple}, if
-$\sum \limits_{i \in I} \widehat{A} = \bigcup \limits_{i \in A} \widehat{A}$ for $A \subseteq \mathcal{Q}$.
+$\sum \limits_{i \in I} \widehat{A} = \bigcup \limits_{i \in A} \widehat{A}$ for any $A \subseteq \mathcal{Q}$.
 
@@
 
+Note that an arbitrary finite $\mathcal{A} \in {\bf R}(;,+)$ does not have to be a quantale since a finite join semilattice may not have the bottom.
+Let us extend $\mathcal{A}$ with the bottom defined as $0 = \Sigma \emptyset$, we denote this algebra as $\mathcal{A}_0$.
+Note that $a_1 ; a_2 = 0$, where $a_1, a_2 \in \mathcal{A}_0$ and one has either $a_1 = 0$ or $a_2 = 0$.
+
+\begin{proposition}
+  If $\mathcal{A} \in {\bf R}(;,+)$ is finite, then $\mathcal{A}_0$ is representable.
+\end{proposition}
+
+\begin{proof}
+  Let $h$ be a representation of $\mathcal{A}$ over some base set $D$.
+Let us define a map $h_0 : \mathcal{A}_0 \to 2^{D \times D}$ such that $h_0(a) = h(a)$, if $a \neq 0$, and $h_0(0) = \emptyset$.
+$h_0$ is a representation since $h$ is a representation.
+\end{proof}
+
 \begin{proposition} \label{distr}
@@
   \begin{enumerate}
-    \item $\mathcal{A}$ is distributive,
-    \item Every generator of $\mathcal{A}$ as a quantale is join-prime.
+    \item $\mathcal{A}_0$ is distributive,
+    \item Every generator of $\mathcal{A}_0$ as a quantale is join-prime.
   \end{enumerate}
@@
     \item
-    $\mathcal{A}$ is finite, so the $\operatorname{Id}(\mathcal{A})$ consists of principal ideals, that is, lower cones generated by elements of $\mathcal{A}$.
+    $\mathcal{A}$ is finite, so the $\operatorname{Id}(\mathcal{A}_0)$ consists of principal ideals, that is, lower cones generated by elements of $\mathcal{A}$.
 
     \begin{center}
-    $\operatorname{Id}(\mathcal{A}) = \{ \down{a} \: | \: a \in \mathcal{A} \}$
+    $\operatorname{Id}(\mathcal{A}_0) = \{ \down{a} \: | \: a \in \mathcal{A} \}$
     \end{center}
 
-    Let $h$ be a representation of $\mathcal{A}$, consider $h(\mathcal{A})$ and its set of ideals $\operatorname{Id}(h(\mathcal{A}))$.
-Intersection of ideals is defined is usual since intersection of ideals is an ideal, join is the defined as follows:
+    Let $h$ be a representation of $\mathcal{A}$, consider $h_0(\mathcal{A}_0)$ and its set of ideals $\operatorname{Id}(h_0(\mathcal{A}_0))$.
+Intersection of ideals is defined is usual since intersection of ideals is an ideal, join is defined as follows:
 
     \begin{center}
-    $\down{h(a)} \vee \down{h(b)} = \down{(h(a) \cup h(b))}$
+    $\down{h_0(a)} \vee \down{h_0(b)} = \down{(h_0(a) \cup h_0(b))}$
     \end{center}
-    Note that $\down{(h(a) \cup h(b))} = \down{h(a + b)}$ since $h$ is a representation of $\mathcal{A}$.
+    Note that $\down{(h_0(a) \cup h_0(b))} = \down{h_0(a + b)}$ since $h$ is a representation of $\mathcal{A}$.
 
@@
 
-    Suppose $h(d) \in \down{h(c)} \cap \down{h(a + b)}$.
-    $h(d) \in \down{h(a + b}$ implies $h(d) \subseteq h(a + b)$.
+    Suppose $h(d) \in \down{h_0(c)} \cap \down{h_0(a + b)}$. We need $h_0(d) \in \down{(h_0(c) \cap h_0(a))} \vee \down{(h_0(c) \cap h_0(b))}$.
 
-    Then if $(x, y) \in h(d)$, then either $(x, y) \in h(a)$ or $(x, y) \in h(b)$.
-    If $(x, y) \in h(a)$ (or $h(b)$), then we have $h(d) \in \down{(h(c) \cap h(a))}$ (or $\down{(h(c) \cap h(b))}$).
+    Let $(x, y) \in h_0(d)$, then either $(x, y) \in h_0(a)$ or $(x, y) \in h_0(b)$ since $d \leq a + b$.
+    Suppose $(x, y) \in h_0(a)$.
+We need $h_0(d) \in \down{(h_0(c) \cap h_0(a))}$, that is, $h_0(d) \subseteq h_0(c) \cap h_0(a)$, but if $(x, y) \in h_0(d)$, then $(x, y) \in h_0(c)$ since $h_0(d) \subseteq h_0(c)$.
+Therefore, every $(x, y) \in h_0(d)$ belongs to $h_0(c) \cap h_0(a)$, if $h_0(d)$ is a subset of $h_0(a)$.
 
-    Suppose $h(d) \in (\down{h(c)} \cap \down{h(a)}) \vee (\down{h(c)} \cap \down{h(b)})$.
-So, we have either $h(d) \in \down{h(c)} \cap \down{h(a)}$ or $h(d) \in \down{h(c)} \cap \down{h(b)}$.
-So $h(d) \in \down{h(c)}$ anyway.
-If $h(d) \in \down{h(a)}$, then $d \leq a \leq a + b$, so $h(d) \in \down{h(a + b)}$.
+    The second alternative with the assumption that $(x, y) \in h_0(a)$ is similar to the previous one.
+
+    Suppose $h_0(d) \in (\down{h_0(c)} \cap \down{h_0(a)}) \vee (\down{h_0(c)} \cap \down{h_0(b)})$.
+So, we have either $h_0(d) \in \down{h_0(c)} \cap \down{h_0(a)}$ or $h_0(d) \in \down{h_0(c)} \cap \down{h_0(b)}$.
+So $h_0(d) \in \down{h_0(c)}$ anyway.
+If $h_0(d) \in \down{h_0(a)}$, then $d \leq a \leq a + b$, so $h_0(d) \in \down{h_0(a + b)}$.
     The second alternative is the same as the first one.
 
-    So, $\operatorname{Id}(h(\mathcal{A}))$ is distributive, and, therefore, $\mathcal{A}$ is distibutive as a join-semilattice by \cite[Lemma 184]{gratzer2011lattice}.
+    So, $\operatorname{Id}(h_0(\mathcal{A}))$ is distributive, and, therefore, $\mathcal{A}_0$ is distibutive as a join-semilattice by \cite[Lemma 184]{gratzer2011lattice}.
 
-    \item Let $g \in \mathcal{A}$ be a generator, suppose $g \leq a + b$, so there are $a' \leq a$ and $b' \leq b$ with $g = a' + b'$.
+    \item Let $g \in \mathcal{A}_0$ be a generator, suppose $g \leq a + b$, so there are $a' \leq a$ and $b' \leq b$ with $g = a' + b'$.
 $g$ is join-irreducible, thus, one has either $g = a'$ or $g = b'$.
@@
 
-The following lemma allows one to consider finite elements of ${\bf R}(;,+)$ as simple quantales.
-If $\mathcal{A}$ is finite and representable, then every $a$ has the form $\Sigma \{ g \in \mathcal{G}(\mathcal{A}) \: | \: g \leq a \}$, but every generator is join-prime by Proposition~\ref{distr}.
+The following lemma allows one to consider extensions finite elements of ${\bf R}(;,+)$ with bottoms as simple quantales.
+If $\mathcal{A}$ is finite and representable, then every $a \in \mathcal{A}_0$ has the form $\Sigma \{ g \in \mathcal{G}(\mathcal{A}) \: | \: g \leq a \}$, but every generator is join-prime by Proposition~\ref{distr}.
 So we have:
@@
 \begin{proposition}\label{quantalerepisok}
-  Let $\mathcal{A}$ be a finite representable join-semilattice semigroup, $\mathcal{A}$ is completely coprime algebraic as a quantale and, therefore, is simple.
+  Let $\mathcal{A}$ be a finite representable join-semilattice semigroup, then $\mathcal{A}_0$ is completely coprime algebraic as a quantale and, therefore, is simple.
 \end{proposition}
@@
 \begin{proof}
-  Let us show that if $\mathcal{A} \in {\bf R}(+, ;)$ such that $\mathcal{A}$ is finite, then
-  $h(\mathcal{A})$ is isomorphic to some subalgebra of $\widehat{A}$. Consider the map
-  $f : h(a) \mapsto \hat{a}$.
+  Let us show that if $\mathcal{A} \in {\bf R}(+, ;)$ such that $\mathcal{A}$ is finite. Let $h$ be a representation of $\mathcal{A}$, so $h_0$ is a representation of $\mathcal{A}_0$.
+  $h(\mathcal{A})$ is a subalgebra of $h_0(\mathcal{A}_0)$, but
+  $h_0(\mathcal{A}_0)$ is isomorphic to $\widehat{A_0}$, the relational quantale representation of $\widehat{A_0}$ for the following reason.
+
+  Consider the map
+  $f : h_0(a) \mapsto \hat{a}$.
 $f$ preserves composition since $h$ and $\hat{.}$ are semigroup homomorphisms.
-The equality $f(h(a) \cup h(b)) = \hat{a} \cup \hat{b}$ holds since $\mathcal{A}$ is simple as a quantale by Proposition~\ref{quantalerepisok}.
+The equality $f(h_0(a) \cup h_0(b)) = \hat{a} \cup \hat{b}$ holds since $\mathcal{A}_0$ is simple as a quantale by Proposition~\ref{quantalerepisok}.
+$f$ is also obviously injective and surjective.
 
-  $\mathcal{A}$ is a relational quantale on the underlying of $\mathcal{A}$, but $\mathcal{A}$ is finite, so $\mathcal{A}$ has a representation over a finite base.
+  $\widehat{\mathcal{A}_0}$ is a relational quantale on the underlying set of $\mathcal{A}_0$, but $\mathcal{A}$ is finite, so $\mathcal{A}$ has a representation over a finite base since $h(\mathcal{A})$ is isomorphic to $h_0(\mathcal{A})$, a subalgebra of $\widehat{A_0}$.
 \end{proof}
@@
 The previous version of this paper contained an error.
-Many thanks to Jas Semrl for noticing that.
+Many thanks to Ja\v{s} \v{S}emrl for noticing that.
 

### 2021-04-17 filtration-stuff/NewText.tex
New file with filtrations stuff

--- 2021-03-04 filtration-stuff/Text.tex
@@
 \title{Notes on filtration of logics containing {\bf K5}}
-
 \begin{document}
@@
 \maketitle
-
-\section{Preliminaries}
-
-\begin{defin}
-  A normal modal logic is a set of formulas that contains all Boolean tautologies,
-  formulas $\Diamond p \lor \Diamond q \leftrightarrow \Diamond (p \lor q)$ and
-  $\Diamond \bot \leftrightarrow \bot$, and is closed under Modus Ponens, Substitution, and
-  Monotonicity:
-  from $\varphi \rightarrow \psi$ infer $\Diamond \varphi \rightarrow \Diamond \psi$.
-\end{defin}
-
-\begin{defin} An Kripke model is a triple $\mathcal{M} = \langle W, R, \vartheta \rangle$,
-  where $R \subseteq W \times W$, $\vartheta : \operatorname{PV} \to 2^W$, and
-  the connectives have the following semantics:
-
-  \begin{enumerate}
-    \item $\mathcal{M}, w \models p \Leftrightarrow w \in \vartheta(p)$
-    \item $\mathcal{M}, w \models \neg \varphi \Leftrightarrow \mathcal{M}, w \nvDash \varphi$
-    \item $\mathcal{M}, w \models \varphi \lor \psi \Leftrightarrow \mathcal{M}, w \models \varphi \text{ or } \mathcal{M}, w \models \psi$
-    \item $\mathcal{M}, w \models \Diamond \varphi \Leftrightarrow \exists v \in R(w) \: \mathcal{M}, v \models \varphi$
-  \end{enumerate}
-\end{defin}
-
-\begin{defin}
-  Let $\mathcal{M} = \langle W, R, \vartheta \rangle$ be a Kripke model.
-  A Kripke model $\mathcal{M}' = \langle W', R', \vartheta' \rangle$ is a generated submodel of
-  $\mathcal{M}$, where:
-  \begin{enumerate}
-    \item $\emptyset \neq W' \subseteq W$ and $W'$ is $R$-closed, i.e., for each $u \in W'$ and $v \in W$,
-    $u R v$ implies $v \in W'$
-    \item $R' = R \cap W' \times W'$
-    \item for every propositional variable $p$, $\vartheta(p)' = \vartheta(p) \cap W'$
-  \end{enumerate}
-\end{defin}
-
-\begin{fact}
-  Let $\mathcal{M} = \langle W, R, \vartheta \rangle$ be a Kripke model and
-  $\mathcal{M}' = \langle W', R', \vartheta' \rangle$ its generated submodel,
-  then for each $w \in W'$
-  \begin{center}
-    $\mathcal{M}, w \models \varphi$ iff $\mathcal{M}', w \models \varphi$
-  \end{center}
-\end{fact}
-
-By ${\bf K5}$ we mean the logic ${\bf K} \oplus {A5}$, where ${A5} = \Diamond p \to \Box \Diamond p$.
-It is known that ${\bf K5}$ is the modal logic of all Euclidean frames.
-A frame is called Euclidean if for each $x, y, z$, $x R y$ and $x R z$ implies $y R z$.
-
-\begin{prop} \label{K5proves}
-  ${\bf K5}$ proves
-  \begin{enumerate}
-    \item $\Box^3 p \leftrightarrow \Box^2 p$
-    \item $\Box^2 \Diamond p \leftrightarrow \Box \Diamond p$
-    \item $\Box \Diamond \Box p \leftrightarrow \Box \Box p$
-    \item $\Box \Diamond^2 p \leftrightarrow \Box \Diamond p$
-  \end{enumerate}
-\end{prop}
-
-\begin{prop} \label{k5useful}
-  Let $\mathcal{M}$ be a ${\bf K}5$ model, $x R y$ for $x, y \in W$ then one has
-  \begin{center}
-    $\mathcal{M}, x \models \Box \Diamond \varphi$ iff $\mathcal{M}, y \models \Box \Diamond \varphi$.
-  \end{center}
-\end{prop}
-
-\begin{proof}
-$ $
-
-  \begin{enumerate}
-    \item Suppose $\mathcal{M}, x \models \Box \Diamond \varphi$. Then $\mathcal{M}, y \models \Diamond \varphi$ and $\mathcal{M}, y \models \Box \Diamond \varphi$
-    \item Suppose $\mathcal{M}, y \models \Box \Diamond \varphi$, then $\mathcal{M}, x \models \Diamond \Box \Diamond \varphi$, so $\mathcal{M}, x \models \Box \Diamond \varphi$.
-  \end{enumerate}
-\end{proof}
-
-\subsection{Filtrations: general definitions}
 
@@
   \item Let $\mathcal{L}$ be a complete normal modal logic. If $\operatorname{Frames}(\mathcal{L})$ admits filtration, then $\mathcal{L}$ has the finite model property.
-  \item If the class of models $\operatorname{Mod}(\mathcal{L})$ admits filtration, then $\mathcal{L}$ has the finite model property and Kripke complete as well.
+  \item If the class of models $\operatorname{Mod}(\mathcal{L})$ admits filtration, then $\mathcal{L}$ has the finite model property and it is Kripke complete as well.
 \end{enumerate}
 \end{lemma}
-
-\section{Filtration of Euclidean logics}
-
-First of all, let us ensure that a minimal filtration of an Euclidean frame is not necessary Euclidean.
-Let $[x] \sim_{\Gamma} [y]$ and $[x] \sim_{\Gamma} [z]$. Then for some $x' \in [x]$ $y' \in [y]$, one has
-$x' R y'$ and $x'' R z'$ for some $x'' \in [x]$ and $z' \in [z]$. Clearly, we cannot claim that $x' = x''$ in general. Thus, minimal filtration does not preserve the required property.
-
-\begin{lemma}
-  ${\bf K5}$ admit filtration.
-\end{lemma}
-
-\begin{proof}
-  Let $\mathcal{M}$ be a ${\bf K5}$-model and $\Gamma_0$ a finite set of formulas closed under subformulas.
-  Let us put $\Gamma = \Gamma_0 \cup \operatorname{Sub}(\{ \Diamond \Box \psi \: | \: \Box \psi \in \Gamma_0 \}) \cup \Psi$, where $\Psi = \nabla_1 \nabla_2 \dots \nabla_n \Box \psi$ for $\Box \psi \in \Gamma_0$ and
-  $\nabla_i \in \{ \Diamond, \Box \}$. By Proposition~\ref{K5proves}, any element of $\Phi$ has one of the four forms. Thus, $W \sim_{\equiv_{\Gamma}}$ has a finite index.
-  We put $\widehat{R} = R^{\operatorname{max}}_{\Gamma}$.
-\end{proof}
 
@@
   &&&&&&& w_5\ar@(ul,ur) \ar@<-.5ex>[d] \\
-  &&&&&& w_1\ar@(ul,ur)  & w_4 \ar@<.5ex>[u] \\
+  &&&&&& w_1\ar@(ul,ur)  & w_4 \ar@<.5ex>[u] \ar@(dl,ul)^{} \\
   &&&&&& w_2 \ar[u] & w_3 \ar[u]
@@
 On the other hand, $R_{min}^E \not\subseteq R_{max}$, so is not $\widehat{R}^{\mathcal{E}}$.
-\section{Filtration for {\bf K4}}
-
-\begin{prop}
-  Let $R$ be a binary relation on $W \neq \emptyset$. Define $R^{+} = \bigcup \limits_{i < \omega} R_i$
-  \begin{enumerate}
-    \item $R_0 = R$
-    \item $R_{n + 1} = R_n \circ R$
-  \end{enumerate}
-  Then $R^{+}$ is transitive
-\end{prop}
-
-\begin{lemma}
-  Let $\mathcal{M} = \langle W, R, \vartheta \rangle$ be a transitive model and $\overline{\mathcal{M}} = \langle \overline{W}, \overline{R}, \overline{\vartheta} \rangle$ its minimal filtration through a finite $\operatorname{Sub}$-closed set of formulas $\Theta$.
-
-  Then $\overline{\mathcal{M}}^{+} = \langle \overline{W}, (\overline{R})^{+}, \overline{\vartheta} \rangle$ is a $\Theta$-filtration of $\mathcal{M}$.
-\end{lemma}
-
-\begin{proof}
-  $(\overline{R})^{+}$ obviously contains $R$. By the previous proposition, $(\overline{R})^{+}$ is transitive.
-  Let us show that $(\overline{R})^{+} \subseteq R^{max}_{\Theta}$.
-
-  Let $\hat{x}, \hat{y} \in \overline{W}$ with $\hat{x} (\overline{R})^{+} \hat{y}$ and
-  $\Box \varphi \in \Theta$ with $\mathcal{M}, x \models \Box \varphi$. Let us show that $\mathcal{M}, y \models \varphi$.
-
-  If $\hat{x} (\overline{R})^{+} \hat{y}$, then there exist equivalence classes $\hat{x}_1, \dots, \hat{x}_n$ such that
-  \begin{center}
-    $\hat{x} \overline{R} \hat{x}_1 \overline{R} \dots \overline{R} \hat{x}_n \overline{R} \hat{y}$
-  \end{center}
-
-  $\mathcal{M}, x \models \Box \varphi$ implies $\mathcal{M}, x \models \Box \Box \varphi$. Thus, $\overline{M}, \hat{x} \models \Box \Box \varphi$.
-
-  $\hat{x} \overline{R} \hat{x}_1$, so there are $x_1 \in \hat{x}$ and $x_2 \in \hat{x}_1$ with $x_1 R x_2$. In particular, $\mathcal{M}, x_2 \models \Box \varphi$,
-  so $\overline{\mathcal{M}}, \hat{x}_2 \models \Box \varphi$, and et cetera.
-
-  For each $i \in \{1, \dots, n\}$ we have $\mathcal{M}, x_i \models \Box \varphi$ which is shown inductively:
-
-  If $\mathcal{M}, x_i \models \Box \varphi$ for $x_i \in \hat{x}_i$, so $\mathcal{M}, x_i \models \Box \Box \varphi$, but there exist ${x_i}' \in \hat{x}_i$ and $x_{i + 1} \in \hat{x}_{i + 1}$, so
-  $\mathcal{M}, x_{i + 1} \models \Box \varphi$.
-
-  Finally, we have $\mathcal{M}, x_n \models \Box \varphi$ for $x_n \in \hat{x}_n$, but $\hat{x}_n \overline{R} \hat{y}$, so $\mathcal{M}, y' \models \varphi$ for each $y' \in \hat{y}$.
-Thus, $\varphi$ is true at $y$ as well.
-\end{proof}
-
-\begin{proof}
-  Let $\hat{x}, \hat{y} \in \overline{W}$ with $\hat{x} (\overline{R})^{+} \hat{y}$ and
-  $\Box \varphi \in \Theta$ with $\mathcal{M}, x \models \Box \varphi$. Let us show that $\mathcal{M}, y \models \varphi$.
-
-  If $\hat{x} (\overline{R})^{+} \hat{y}$, then there exist equivalence classes $\hat{x}_1, \dots, \hat{x}_n$ such that
-  \begin{center}
-    $\hat{x} \overline{R} \hat{x}_1 \overline{R} \dots \overline{R} \hat{x}_n \overline{R} \hat{y}$
-  \end{center}
-
-  Let us show that $\mathcal{M}, \hat{x}_i \models \Box \varphi$ inductively:
-  \begin{enumerate}
-    \item $n = 1$
-    We have the following sequence:
-    \begin{center}
-      $\hat{x} \overline{R} \hat{x}_1 \overline{R} \hat{y}$
-    \end{center}
-    $\hat{x} \overline{R} \hat{x}_1$, so there are $x' \in \hat{x}$ and $x_1' \in \hat{x}_1$ such that $x' R x_1'$. $\Box \varphi$ is true at $x'$, so is $\Box \Box \varphi$. Then
-    $\mathcal{M}, x_1' \models \Box \varphi$ since $x_1' \in R(x')$.
-    So $\overline{\mathcal{M}}, \hat{x}_1 \models \Box \varphi$.
-
-    \item $n = i + 1$
-    The case is the following:
-    \begin{center}
-      $\hat{x} \overline{R} \hat{x}_1 \overline{R} \dots \overline{R} \hat{x}_i \overline{R} \hat{x}_{i + 1} \overline{R} \hat{y}$
-    \end{center}
-
-    By IH, $\Box \varphi$ is true at $\hat{x}_i$, so is $\Box \Box \varphi$. Hence, we have $\overline{\mathcal{M}}, \hat{x}_{i + 1} \models \Box \varphi$ since
-    $\hat{x}_i \overline{R} \hat{x}_{i + 1}$.
-  \end{enumerate}
-
-  That is, for each $0 < n < \omega$, if we have a sequence of equivalence classes with $\hat{x} \overline{R} \hat{x}_1 \overline{R} \dots \overline{R} \hat{x}_n \overline{R} \hat{y}$ where $\overline{\mathcal{M}}, \hat{x} \models \Box \varphi$, then
-  $\overline{\mathcal{M}}, \hat{x}_n \models \Box \varphi$.
-
-  If $\hat{x}_n \overline{R} \hat{y}$,
-  then there are $x_n' \in \hat{x}_n$ and  $y' \in \hat{y}$ with $x_n' R y'$.
-$\mathcal{M}, x_n' \models \Box \varphi$ implies $\mathcal{M}, y' \models \varphi$, but $y'$ and $y$ are $\Gamma$-equivalent and $\varphi \in \Gamma$, so $\mathcal{M}, y \models \varphi$.
-\end{proof}
 
@@
 
-\begin{lemma}
+\begin{lemma}\label{good}
   Let $\mathcal{L}$ be a modal logic and $\varphi \not\in \mathcal{L}$, then $\mathcal{M}_{\mathcal{L}}^{\operatorname{Sub}(\varphi)} \nvDash \varphi$.
@@
 
-\begin{theorem} $\mathcal{L}$ be a modal logic, then the following are equivalent:
-  \begin{enumerate}
-    \item $\mathcal{L}$ admits strict filtrations
-    \item If $\Gamma$ is a finite $\operatorname{Sub}$-closed set of formulas , there exists a finite ``canonical'' model $\mathcal{M}_{\mathcal{L}}^{\Gamma}$ such that
-    $\mathcal{M}_{\mathcal{L}}^{\Gamma} \models \mathcal{L}$.
-  \end{enumerate}
-\end{theorem}
+\begin{lemma} Let $\mathcal{L}$ be a modal logic and $\Gamma$ a finite $\operatorname{Sub}$-closed set of formulas, then if $\mathcal{L}$ admits strict filtrations, then there exists a finite ``canonical'' model $\mathcal{M}_{\mathcal{L}}^{\Gamma}$ such that $\mathcal{M}_{\mathcal{L}}^{\Gamma} \models \mathcal{L}$.
+\end{lemma}
 
@@
   $ $
-
-  \begin{enumerate}
-    \item $(\Rightarrow)$
+$(\Rightarrow)$
   Let $\Gamma$ be a finite $\operatorname{Sub}$-closed of formulas.
@@
   It is readily checked that the quotient model $\mathcal{M}_{\mathcal{L}} / \sim_{\Gamma}$ satifsfies Definition~\ref{fin}.
+\end{proof}
 
-    \item $(\Leftarrow)$
+The converse implication does not have to true generally.
+We prove a bit more weak version of this implication.
+A logic admits $\mathcal{L}$ admits $\Diamond$-filtrations, if for each model $\mathcal{M} \models \mathcal{L}$ and for each $\Gamma$, a $\operatorname{Sub}$-closed set of formulas, there exists a model $\widehat{\mathcal{M}}$, a filtration of $\mathcal{L}$ through $\operatorname{Sub}(\Gamma \cup \Diamond \Gamma)$.
 
+${\bf K}5$ is an example of such a logic since ${\bf K}5$ admits filtrations through sets having the form $\Gamma \cup \{ \Diamond \Box \psi \: | \: \Box \psi \in \Gamma \}$.
+
+Note that, if $\mathcal{L}$ admits $\Diamond$-filtrations, so $\mathcal{L}$ admits the ``finite canonical'' model property, the proof is the same as in Lemma~\ref{good}.
+
+\begin{lemma}
+  Let $\mathcal{L}$ be a logic that admits $\Diamond$-filtrations and $\Gamma$ a $\operatorname{Sub}$-closed set of formulas, then there exists a ``finite canonical'' model $\mathcal{M}^{\Gamma}_{\mathcal{L}} \models \mathcal{L}$
+\end{lemma}
+
+\begin{proof}
 Let $\Gamma$ be a finite $\operatorname{Sub}$-closed of formulas and $\mathcal{M}_{\mathcal{L}}^{\Gamma} = \langle W_{\mathcal{L}}^{\Gamma}, R_{\mathcal{L}}^{\Gamma}, \vartheta_{\mathcal{L}}^{\Gamma} \rangle$ a ``finite canonical'' model generated by $\Gamma$.
 
-    Let us show that there exists a model $\widehat{\mathcal{M}}$ such that
-    $\widehat{\mathcal{M}}$ is a strict filtration of $\mathcal{M}$ through $\operatorname{Sub}(\Gamma)$.
+Let us show that there exists a model $\widehat{\mathcal{M}}$, a strict filtration of $\mathcal{M}$ through $\operatorname{Sub}(\Gamma \cup \Diamond \Gamma) := \Psi$.
 
-    Let $\sim$ be an equivalence relation on $\underline{\mathcal{M}}$ defined as usual:
+Let $\sim$ be an equivalence relation on $\mathcal{M}$ defined as usual:
 
     \begin{center}
-      $u \sim v$ iff $\forall \psi \in \Gamma$
+      $u \sim v$ iff $\forall \psi \in \Psi$
       $\mathcal{M}, u \models \psi \Leftrightarrow \mathcal{M}, v \models \psi$
@@
 
-    Let $[u] \in \mathcal{M}/\sim$, consider its truth set $||[u]|| = \{ \psi \in \Gamma \: | \: \mathcal{M}, u \models \psi \}$.
+    Let $[u] \in \mathcal{M}/\sim$, consider its truth set $||[u]|| = \{ \psi \in \Psi \: | \: \mathcal{M}, u \models \psi \}$.
     $||[u]||$ is clearly a finite $\mathcal{L}$-consistent set and it is maximal.
-    Consider a map
+    Consider a map:
     \begin{center}
-    $\iota : W/\sim \to W_{\mathcal{L}}^{\varphi}$ with $\iota : [u] \mapsto ||[u]||$
+    $\iota : W/\sim \to W_{\mathcal{L}}^{\Gamma}$ with $\iota : [u] \mapsto ||[u]||$
     \end{center}
 
-    Consider the set $\iota(W/\sim)$ that we denote as $\operatorname{Im}(\iota)$. Let us show that $\operatorname{Im}(\iota)$ is $R_{\mathcal{L}}^{\varphi}$-closed.
+    Consider the set $\iota(W/\sim)$ that we denote as $\operatorname{Im}(\iota)$. Let us show that $\operatorname{Im}(\iota)$ is $R_{\mathcal{L}}^{\Gamma}$-closed.
 
-    Let $\iota([u]) R_{\mathcal{L}}^{\varphi} \Delta$, let us show that $\Delta \in \iota(W/\sim)$, that is, $\Delta = \iota([v])$ for some $v \in W$.
+    Let $\iota([u]) R_{\mathcal{L}}^{\Gamma} \Delta$, let us show that $\Delta \in \iota(W/\sim)$, that is, $\Delta = \iota([v])$ for some $v \in W$.
 
@@
     Suppose that there are $u' \in [u]$ and $v' \in [v]$ with $u' R v'$. Let us show that
-    $[u] \widehat{R} [v]$. That is, let us show that $\Diamond \iota(v) \subseteq \iota(v)$.
-    But if $\mathcal{M}, v' \models \psi$ for $\Diamond \psi \in \Gamma$, then
-    $\mathcal{M}, u' \models \Diamond \psi$. Recall that $u' \sim u$ and $v' \sim v$, so we have $\Diamond ||[v]|| \subseteq ||[u]||$.
+    $[u] \widehat{R} [v]$. That is, let us show that $\Diamond \iota(v) \subseteq \iota(u)$.
+
+    Suppose $\psi \in \iota(v)$, that is, $\psi \in ||v||$, that denotes $\psi \in \Psi$ and $\mathcal{M}, v \models \psi$.
+Therefore, $\mathcal{M}, v' \models \psi$.
+Then $\mathcal{M}, u' \models \Diamond \psi$.
+$\Diamond \psi \in \Psi$ and $u \sim_{\Gamma} u'$, so
+    $\mathcal{M}, u \models \Diamond \psi$.
 
     The second inclusion follows from the definition of a relation in a finite ``canonical'' model.
-  \end{enumerate}
 \end{proof}
 
-In particular, ${\bf K}5$ fails to have the finite ``canonical'' model property that
-follows from the contraposition of the theorem above and Theorem~\ref{sad}.
+In particular, ${\bf K}5$ has the finite ``canonical'' model property.
 

### 2021-04-17 frp-paper/Paper.tex
New file with filtrations stuff

@@
 The poset $\langle \operatorname{Id}(\mathcal{A}), \subseteq \rangle$ consists of all ideals of $\mathcal{A}$ ordered by inclusion.
-$\mathcal{A}$ a join semilattice is called \emph{distributive}, if for each $x, a, b \in \mathcal{A}$ such that $x \leq a + b$, so there are $a' \leq a$ and $b' \leq b$ such that $x = a' + b'$.
+A join semilattice $\mathcal{A}$ is called \emph{distributive}, if for each $x, a, b \in \mathcal{A}$ such that $x \leq a + b$, so there are $a' \leq a$ and $b' \leq b$ such that $x = a' + b'$.
 A finite $\mathcal{A}$ is distributive is the lattice $\operatorname{Id}(\mathcal{A})$ is distributive in the usual sense, see \cite[Lemma 184]{gratzer2011lattice}.

### 2021-04-17 frp-paper/Paper.tex
elementary chain argument for uncountable join semilattice-ordered semigroups

@@
 \newtheorem{claim}{Claim}
+
+\let\strokeL\L
+\renewcommand\L{\mathbf{L}}
 
@@
 
-\begin{lemma} Let $\mathcal{A} = \langle A, ;, + \rangle$ be a join semilattice-ordered semigroup,
+\begin{lemma} \label{rep} Let $\mathcal{A} = \langle A, ;, + \rangle$ be a join semilattice-ordered semigroup,
 
@@
     \item If $\mathcal{A}$ is representable then $\exists$ has a winning strategy in $\mathcal{G}_{\omega}(\mathcal{A})$.
-    \item If $\mathcal{A}$ is infinite and $\exists$ has a winning strategy in $\mathcal{G}_{\omega}(\mathcal{A})$ then $\mathcal{A}$ is representable.
+    \item If $|\mathcal{A}| = \omega$ and $\exists$ has a winning strategy in $\mathcal{G}_{\omega}(\mathcal{A})$ then $\mathcal{A}$ is representable.
   \end{enumerate}
@@
     \item
-    Firstly, let us suppose that $|\mathcal{A}| = \omega$.
-
     Let $a, b \in \mathcal{A}$ such that $a \not\leq b$.
@@
 \end{itemize}
-For $|\mathcal{A}| \geq \aleph_1$ we apply the previous item and Tarksi-Vaught theorem on elementary chain union (see \cite[Theorem 2.5.2]{hodges1997shorter}).
   \end{enumerate}
@@
 
-\begin{proposition}
+\begin{proposition} \label{fin}
   Let $\mathcal{A}$ be a join semilattice-ordered semigroup, then there exists a game such that $\mathcal{A}$ is representable iff $\exists$ has a winning strategy in the all finite lenght games.
@@
 
+We have already considered finite and countable algebras. Let us consider also arbitrary to complete the axiomatisation of ${\bf R}(;,+)$. We have the following statement.
+
+\begin{lemma}
+  Let $\mathcal{A}$ be a join-semilattice semigroup, then $\exists$ has a winning strategy has a winning strategy in $\mathcal{G}_m(\mathcal{A})$ iff $\mathcal{A}$ is elementary equivalent to some countable representable join-semilattice semigroups.
+\end{lemma}
+
+\begin{proof}
+  The $\Leftarrow$ implication is obvious.
+If $\mathcal{A} \equiv \mathcal{A}'$, where $|\mathcal{A}'| = \omega$ and $\mathcal{A}' \in {\bf R}(;,+)$, then $\exists$ has a winning strategy in $\mathcal{G}_{\omega}(\mathcal{A}')$ (Lemma~\ref{rep}), so $\exists$ has a winning strategy in
+  every $\mathcal{G}_{m}(\mathcal{A}')$ for $m < \omega$ (Proposition~\ref{fin}). Thus, so $\exists$ has a winning strategy in
+  every $\mathcal{G}_{m}(\mathcal{A})$ since $\mathcal{A}$ and $\mathcal{A}'$ are elementary equivalent.
+
+  Suppose $\exists$ has a winning strategy has a winning strategy in $\mathcal{G}_m(\mathcal{A})$ for $m < \omega$.
+The proof scheme is the same as for relation or cylindric algebras, so we provide the following sketch.
+In particular, we have the following proof that may be proved using \strokeL o\'s's Theorem similarly to.
+
+  \begin{claim}
+    If $\exists$ has a winning strategy has a winning strategy in $\mathcal{G}_m(\mathcal{A})$ for $m < \omega$, then $\exists$ has a winning strategy for $\mathcal{G}_{\omega}(\Pi_{U} \mathcal{A})$, where $\Pi_{U} \mathcal{A}$ is a non-principal ultrapower of $\mathcal{A}$ for $U$, an ultrafilter over $\omega$.
+  \end{claim}
+
+  Consider $\mathcal{B}$, a countable elementary equivalent subalgbera of $\Pi_{U} \mathcal{A}$ that exists by the downward L\"{o}wenheim-Skolem theorem.
+  We construct the following elementary chain inxeded over $\omega$:
+  \begin{center}
+    $\mathcal{B}_0 \preceq \mathcal{B}_1 \preceq \dots \mathcal{B}_i \preceq\mathcal{B}_{i + 1} \preceq \dots (\preceq \Pi_{U} \mathcal{A})$.
+  \end{center}
+  Let us put $\mathcal{B}_0 := \mathcal{B}$.
+  Suppose the initial segment $\mathcal{B}_0 \preceq \mathcal{B}_1 \preceq \dots \mathcal{B}_i$ has been already constructed. Let us constuct $\mathcal{B}_{i + 1}$.
+
+  Consider a function $\sigma$ that takes an initial game of the play $\mathcal{G}_{\omega}(\Pi_{U} \mathcal{A})$, say, $\mathcal{N}_0 \subseteq \mathcal{N}_1 \subseteq \dots \mathcal{N}_k$ (all $k$) and Abelard's move and returns $\exists$'s repsonse, $\mathcal{N}_{k + 1}$, an extension of $\mathcal{N}_k$.
+
+  Consider $\sigma \upharpoonright_{\mathcal{B}_i}$, a restriction of $\sigma$, where $\forall$ takes elements only from $\mathcal{B}_i$ for his moves.
+The range of this restriction consists of $\Pi_{U} \mathcal{A}$-networks, and $E$, the set of elements that occur in those networks is also countable.
+We define $\mathcal{B}_{i + 1}$ as a countable elementary subalgebra of $\Pi_{U} \mathcal{A}$ that contains $\mathcal{B}_{i}$ and $E$.
+The existence follows from the downward L\"{o}wenheim-Skolem theorem.
+
+  Let us put $\mathfrak{B} = \bigcup \limits_{i < \omega} \mathcal{B}_i$.
+Note that $\mathfrak{B}$ is countable.
+By Tarksi-Vaught elementary chain union theorem, for each $i < \omega$ $\mathcal{B}_i \preceq \mathfrak{B}$.
+Therefore, $\mathfrak{B} \equiv \Pi_{U} \mathcal{A}$.
+On the other hand, $\mathcal{A} \equiv \Pi_{U} \mathcal{A}$.
+  Moreover, $\mathfrak{B}$ is representable and $\mathfrak{B} \equiv \mathcal{A}$.
+\end{proof}
+
 This inductive sequence of formulas provides us the explicit axiomatisation of the class of representable join semilattice-ordered semigroups.
@@
 
-Let $\mathcal{A}$ be a join semilattice, a subset $I \subseteq \mathcal{A}$ is called \emph{ideal}, if the following hold:
-\begin{enumerate}
-  \item If $x \in I$, then $\down{x} \subseteq I$,
-  \item $x, y \in I$ implies $x + y \in I$.
-\end{enumerate}
-The poset $\langle \operatorname{Id}(\mathcal{A}), \subseteq \rangle$ consists of all ideals of $\mathcal{A}$ ordered by inclusion.
-A join semilattice $\mathcal{A}$ is called \emph{distributive}, if for each $x, a, b \in \mathcal{A}$ such that $x \leq a + b$, so there are $a' \leq a$ and $b' \leq b$ such that $x = a' + b'$.
-A finite $\mathcal{A}$ is distributive is the lattice $\operatorname{Id}(\mathcal{A})$ is distributive in the usual sense, see \cite[Lemma 184]{gratzer2011lattice}.
-
-Note that an arbitrary finite $\mathcal{A} \in {\bf R}(;,+)$ does not have to be a quantale since a finite join semilattice may not have the bottom.
-Let us extend $\mathcal{A}$ with the bottom defined as $0 = \Sigma \emptyset$, we denote this algebra as $\mathcal{A}_0$.
-Note that $a_1 ; a_2 = 0$, where $a_1, a_2 \in \mathcal{A}_0$ and one has either $a_1 = 0$ or $a_2 = 0$.
-
-\begin{proposition}
-  If $\mathcal{A} \in {\bf R}(;,+)$ is finite, then $\mathcal{A}_0$ is representable.
-\end{proposition}
-
-\begin{proof}
-  Let $h$ be a representation of $\mathcal{A}$ over some base set $D$.
-Let us define a map $h_0 : \mathcal{A}_0 \to 2^{D \times D}$ such that $h_0(a) = h(a)$, if $a \neq 0$, and $h_0(0) = \emptyset$.
-$h_0$ is a representation since $h$ is a representation.
-\end{proof}
-
-\begin{proposition} \label{distr}
-  Let $\mathcal{A} \in {\bf R}(;,+)$ and $\mathcal{A}$ is finite, then
-  \begin{enumerate}
-    \item $\mathcal{A}_0$ is distributive,
-    \item Every generator of $\mathcal{A}_0$ as a quantale is join-prime.
-  \end{enumerate}
-\end{proposition}
-
-\begin{proof}
-$ $
-
-  \begin{enumerate}
-    \item
-    $\mathcal{A}$ is finite, so the $\operatorname{Id}(\mathcal{A}_0)$ consists of principal ideals, that is, lower cones generated by elements of $\mathcal{A}$.
-
-    \begin{center}
-    $\operatorname{Id}(\mathcal{A}_0) = \{ \down{a} \: | \: a \in \mathcal{A} \}$
-    \end{center}
-
-    Let $h$ be a representation of $\mathcal{A}$, consider $h_0(\mathcal{A}_0)$ and its set of ideals $\operatorname{Id}(h_0(\mathcal{A}_0))$.
-Intersection of ideals is defined is usual since intersection of ideals is an ideal, join is defined as follows:
-
-    \begin{center}
-    $\down{h_0(a)} \vee \down{h_0(b)} = \down{(h_0(a) \cup h_0(b))}$
-    \end{center}
-    Note that $\down{(h_0(a) \cup h_0(b))} = \down{h_0(a + b)}$ since $h$ is a representation of $\mathcal{A}$.
-
-    Let $c \in \mathcal{A}$, consider $h(c)$.
-
-    Suppose $h(d) \in \down{h_0(c)} \cap \down{h_0(a + b)}$. We need $h_0(d) \in \down{(h_0(c) \cap h_0(a))} \vee \down{(h_0(c) \cap h_0(b))}$.
-
-    Let $(x, y) \in h_0(d)$, then either $(x, y) \in h_0(a)$ or $(x, y) \in h_0(b)$ since $d \leq a + b$.
-    Suppose $(x, y) \in h_0(a)$.
-We need $h_0(d) \in \down{(h_0(c) \cap h_0(a))}$, that is, $h_0(d) \subseteq h_0(c) \cap h_0(a)$, but if $(x, y) \in h_0(d)$, then $(x, y) \in h_0(c)$ since $h_0(d) \subseteq h_0(c)$.
-Therefore, every $(x, y) \in h_0(d)$ belongs to $h_0(c) \cap h_0(a)$, if $h_0(d)$ is a subset of $h_0(a)$.
-
-    The second alternative with the assumption that $(x, y) \in h_0(a)$ is similar to the previous one.
-
-    Suppose $h_0(d) \in (\down{h_0(c)} \cap \down{h_0(a)}) \vee (\down{h_0(c)} \cap \down{h_0(b)})$.
-So, we have either $h_0(d) \in \down{h_0(c)} \cap \down{h_0(a)}$ or $h_0(d) \in \down{h_0(c)} \cap \down{h_0(b)}$.
-So $h_0(d) \in \down{h_0(c)}$ anyway.
-If $h_0(d) \in \down{h_0(a)}$, then $d \leq a \leq a + b$, so $h_0(d) \in \down{h_0(a + b)}$.
-    The second alternative is the same as the first one.
-
-    So, $\operatorname{Id}(h_0(\mathcal{A}))$ is distributive, and, therefore, $\mathcal{A}_0$ is distibutive as a join-semilattice by \cite[Lemma 184]{gratzer2011lattice}.
-
-    \item Let $g \in \mathcal{A}_0$ be a generator, suppose $g \leq a + b$, so there are $a' \leq a$ and $b' \leq b$ with $g = a' + b'$.
-$g$ is join-irreducible, thus, one has either $g = a'$ or $g = b'$.
-In fact, this statement generalises the fact that an element is join-prime iff it is join-irreducible in distributive lattices, see \cite[Lemma 5.11]{davey2002introduction}.
-  \end{enumerate}
-\end{proof}
-
-The following lemma allows one to consider extensions finite elements of ${\bf R}(;,+)$ with bottoms as simple quantales.
-If $\mathcal{A}$ is finite and representable, then every $a \in \mathcal{A}_0$ has the form $\Sigma \{ g \in \mathcal{G}(\mathcal{A}) \: | \: g \leq a \}$, but every generator is join-prime by Proposition~\ref{distr}.
-So we have:
+TODO: complete the argument
 

### 2021-04-17 frp-paper/Paper.tex
update refs

@@
 
-The following proposition is a version of \cite[Proposition 7.24]{hirsch2002relation} and the right-to-left part is proved using K\"{o}nig's lemma \cite[Exercise 5.6.3]{hodges1997shorter}.
+The following proposition is a version of \cite[Proposition 7.24]{hirsch2002relation} and the right-to-left part is proved using K\"{o}nig's lemma \cite[Exercise 5.6.5]{hodges1993model}.
 
@@
   Suppose $\exists$ has a winning strategy has a winning strategy in $\mathcal{G}_m(\mathcal{A})$ for $m < \omega$.
-The proof scheme is the same as for relation or cylindric algebras, so we provide the following sketch.
-In particular, we have the following proof that may be proved using \strokeL o\'s's Theorem similarly to.
+The proof scheme is the same as for relation (see \cite[Lemma 19]{hirsch1997complete}) or cylindric algebras (see \cite[Corollary 3.3.5]{hirsch2013completions}), so we provide the following sketch.
+In particular, we have the following proof that may be proved using \strokeL o\'s's theorem (see \cite[Theorem 9.5.1]{hodges1993model}) similarly to.
 
@@
 
-  Consider $\mathcal{B}$, a countable elementary equivalent subalgbera of $\Pi_{U} \mathcal{A}$ that exists by the downward L\"{o}wenheim-Skolem theorem.
+  Consider $\mathcal{B}$, a countable elementary equivalent subalgbera of $\Pi_{U} \mathcal{A}$ that exists by the downward L\"{o}wenheim-Skolem theorem, see \cite[Corollary 3.1.5]{hodges1993model}.
   We construct the following elementary chain inxeded over $\omega$:
@@
 Note that $\mathfrak{B}$ is countable.
-By Tarksi-Vaught elementary chain union theorem, for each $i < \omega$ $\mathcal{B}_i \preceq \mathfrak{B}$.
+By Tarksi-Vaught elementary chain union theorem (see \cite[Theorem 2.5.2]{hodges1993model}), for each $i < \omega$ $\mathcal{B}_i \preceq \mathfrak{B}$.
 Therefore, $\mathfrak{B} \equiv \Pi_{U} \mathcal{A}$.
-On the other hand, $\mathcal{A} \equiv \Pi_{U} \mathcal{A}$.
+  On the other hand, $\mathcal{A} \equiv \Pi_{U} \mathcal{A}$ (see \cite[Corollary 9.5.2]{hodges1993model}).
   Moreover, $\mathfrak{B}$ is representable and $\mathfrak{B} \equiv \mathcal{A}$.

### 2021-04-20 filtration-stuff/NewText.tex
GL and finite canonical models

@@
 The converse implication does not have to true generally.
-We prove a bit more weak version of this implication.
-A logic admits $\mathcal{L}$ admits $\Diamond$-filtrations, if for each model $\mathcal{M} \models \mathcal{L}$ and for each $\Gamma$, a $\operatorname{Sub}$-closed set of formulas, there exists a model $\widehat{\mathcal{M}}$, a filtration of $\mathcal{L}$ through $\operatorname{Sub}(\Gamma \cup \Diamond \Gamma)$.
 
-${\bf K}5$ is an example of such a logic since ${\bf K}5$ admits filtrations through sets having the form $\Gamma \cup \{ \Diamond \Box \psi \: | \: \Box \psi \in \Gamma \}$.
+Recall that ${\bf GL} = {\bf K} \oplus \Diamond (p \land \neg \Diamond p) \to \Diamond p$.
+${\bf GL}$ does not admit filtrations, but one may prove that ${\bf GL}$ has the FMP using the selective filtation technique.
 
-Note that, if $\mathcal{L}$ admits $\Diamond$-filtrations, so $\mathcal{L}$ admits the ``finite canonical'' model property, the proof is the same as in Lemma~\ref{good}.
+\begin{defin}
+  Let $\mathcal{M} = \langle W, R, \vartheta \rangle$ be a model and $\Gamma$ a set closed under subformulas, then a weak submodel $\mathcal{M} = \langle W', R', \eta \rangle$ (that is, $W' \subseteq W$, $R' \subseteq R$, $\eta = \vartheta \upharpoonright_{W'}$) is called a selective filtration of $\mathcal{M}$ through $\Gamma$ is the following holds, for $\Diamond \psi \in \Gamma$, $x \in W'$:
+  \begin{center}
+    $\mathcal{M}, x \models \Diamond \psi$ iff $\mathcal{M}, y \models \psi$ for some $y \in R'(x)$
+  \end{center}
+\end{defin}
 
-\begin{lemma}
-  Let $\mathcal{L}$ be a logic that admits $\Diamond$-filtrations and $\Gamma$ a $\operatorname{Sub}$-closed set of formulas, then there exists a ``finite canonical'' model $\mathcal{M}^{\Gamma}_{\mathcal{L}} \models \mathcal{L}$
-\end{lemma}
+\begin{theorem}
+  ${\bf GL}$ admits the ``finite canonical'' model property.
+\end{theorem}
 
 \begin{proof}
-Let $\Gamma$ be a finite $\operatorname{Sub}$-closed of formulas and $\mathcal{M}_{\mathcal{L}}^{\Gamma} = \langle W_{\mathcal{L}}^{\Gamma}, R_{\mathcal{L}}^{\Gamma}, \vartheta_{\mathcal{L}}^{\Gamma} \rangle$ a ``finite canonical'' model generated by $\Gamma$.
+  Let $\Gamma$ be a finite $\operatorname{\bf Sub}$-closed set.
+So, $\Gamma = \operatorname{\bf Sub}(\gamma_1) \cup \dots \operatorname{\bf Sub}(\gamma_n)$, for some $\gamma_1, \dots, \gamma_n \in \Gamma$ and every $\gamma_i$ is ${\bf GL}$-consistent.
 
-Let us show that there exists a model $\widehat{\mathcal{M}}$, a strict filtration of $\mathcal{M}$ through $\operatorname{Sub}(\Gamma \cup \Diamond \Gamma) := \Psi$.
+  $\gamma$ is true in $\mathcal{M}_{\Gamma} = \langle V, R^{\bf GL}, \vartheta \rangle$, where
+  \begin{center}
+  $V_{\Gamma} = \bigcup \limits_{\psi \in \Gamma} \{ x \in W_{\bf GL} \: | \: \mathcal{M}_{\bf GL}, x \models \psi \land \neg \Diamond \psi \}$
+  \end{center}
+  since $\mathcal{M}_{\Gamma}$ is a selective filtration of $\mathcal{M}_{\bf M}$ through $\Gamma$.
 
-Let $\sim$ be an equivalence relation on $\mathcal{M}$ defined as usual:
+  Let $x \in V$, consider $||x|| = \{ \psi \in \Gamma \: | \: \mathcal{M}_{\bf GL}, x \models \psi \}$
+  and define $x R^{\bf GL} y$ implies $||x|| R_{\Gamma} ||y||$. Every $||x||$ is clearly finite.
+  The valuation map is usual, $\vartheta_{\Gamma}(p) = \{ ||x|| \: | \: p \in ||x|| \}$ for $p \in \operatorname{PV} \cap \Gamma$. We put $W_{\Gamma}$
 
-    \begin{center}
-      $u \sim v$ iff $\forall \psi \in \Psi$
-      $\mathcal{M}, u \models \psi \Leftrightarrow \mathcal{M}, v \models \psi$
-    \end{center}
+  Let us check that $\langle W_{\Gamma}, R_{\Gamma}, \vartheta_{\Gamma} \rangle$ is a ``finite canonical`` ${\bf GL}$-model.
+  Let us check that $R_{\Gamma}$ is transitive and Noetherian and it has no relflexive points.
+  If $W_{\Gamma}$ had a relfexive point, say $||y||$, then we would have $\mathcal{M}, y \models \psi \to \Diamond \psi$, but that is not impossible. Transitivity is obvious.
 
-    Let $[u] \in \mathcal{M}/\sim$, consider its truth set $||[u]|| = \{ \psi \in \Psi \: | \: \mathcal{M}, u \models \psi \}$.
-    $||[u]||$ is clearly a finite $\mathcal{L}$-consistent set and it is maximal.
-    Consider a map:
-    \begin{center}
-    $\iota : W/\sim \to W_{\mathcal{L}}^{\Gamma}$ with $\iota : [u] \mapsto ||[u]||$
-    \end{center}
-
-    Consider the set $\iota(W/\sim)$ that we denote as $\operatorname{Im}(\iota)$. Let us show that $\operatorname{Im}(\iota)$ is $R_{\mathcal{L}}^{\Gamma}$-closed.
-
-    Let $\iota([u]) R_{\mathcal{L}}^{\Gamma} \Delta$, let us show that $\Delta \in \iota(W/\sim)$, that is, $\Delta = \iota([v])$ for some $v \in W$.
-
-    Note that $\bigwedge \limits_{\psi \in \Delta} \psi \in \Delta$ since $\Delta$ is finite.
-    This conjuction has the form $\bigwedge \limits_{\psi \in \Delta} \psi \in \Delta = \psi_1 \land \dots \land \psi_n$ for some $n < \omega$ and for some $\psi_i \in \Gamma \cup \neg \Gamma$, where $0 < i < \omega$.
-
-    Therefore $\mathcal{M}^{\Gamma}_{\mathcal{L}}, \iota([u]) \models \Diamond (\psi_1 \land \dots \land \psi_n)$.
-    Then $\mathcal{M}, u \models \Diamond (\psi_1 \land \dots \land \psi_n)$. That implies $\mathcal{M}, v \models \psi_1 \land \dots \land \psi_n$ for some $v \in R(u)$.
-    This is clearly a desired $v$ since $||[v]|| \subseteq \Delta$ and such a $\Delta$ is unique.
-
-    Thus, $\langle \operatorname{Im}(\iota), R_{\mathcal{L}}^{\Gamma}, {\vartheta'}_{\mathcal{L}}^{\Gamma} \rangle$, where ${\vartheta'}_{\mathcal{L}}^{\Gamma}(p) = \vartheta_{\mathcal{L}}^{\Gamma}(p) \cap \operatorname{Im}(\iota)$
-    is a generated submodel of $\mathcal{M}_{\mathcal{L}}^{\Gamma}$.
-
-    Consider a model $\widehat{\mathcal{M}} = \langle W/\sim, \widehat{R}, \widehat{\vartheta} \rangle$, where $\widehat{R}$ is defined as
-    \begin{center}
-      $[u] \widehat{R} [v]$ iff $||[u]|| {R}_{\mathcal{L}}^{\Gamma} ||[v]||$
-    \end{center}
-
-    Let us show that $R_{min} \subseteq \widehat{R} \subseteq R_{max}$.
-
-    Suppose that there are $u' \in [u]$ and $v' \in [v]$ with $u' R v'$. Let us show that
-    $[u] \widehat{R} [v]$. That is, let us show that $\Diamond \iota(v) \subseteq \iota(u)$.
-
-    Suppose $\psi \in \iota(v)$, that is, $\psi \in ||v||$, that denotes $\psi \in \Psi$ and $\mathcal{M}, v \models \psi$.
-Therefore, $\mathcal{M}, v' \models \psi$.
-Then $\mathcal{M}, u' \models \Diamond \psi$.
-$\Diamond \psi \in \Psi$ and $u \sim_{\Gamma} u'$, so
-    $\mathcal{M}, u \models \Diamond \psi$.
-
-    The second inclusion follows from the definition of a relation in a finite ``canonical'' model.
+  Clearly that $x R^{\bf GL} y$ implies $||x|| R_{\Gamma} ||y||$ and vice versa.
+So, if we had an infinite decreasing $R_{\Gamma}$-chain, then we would have an infinite decreasing $R^{\bf GL}$-chain, but that is not possible.
+  So $\langle W_{\Gamma}, R_{\Gamma}, \vartheta_{\Gamma} \rangle$ is a ``finite canonical`` ${\bf GL}$-model.
 \end{proof}
 
-In particular, ${\bf K}5$ has the finite ``canonical'' model property.
+So we have an example of a logic that has the ``finite canonical'' model property with no filtrations.
 

### 2021-04-24 frp-paper/Paper.tex
fixing plenty of bugs

@@
 The examples of the class having the finite representation property are some classes of algebras \cite{hirsch2004finite} \cite{hirsch2013ordered} \cite{mclean2016finite}, the subsignature of which contains the domain and range operators.
-The other kind of algebras of binary relations having the finite representation property is semigroups with so-called demonic refinement has been recently studied by Hirsch and \v{S}emrl \cite{hirsch2020finite}.
+The other kind of algebras of binary relations having the finite representation property is semigroups with so-called demonic refinement has been recently studied by Hirsch and \v{S}emrl \cite{hirsch2021finite}.
 
-There are subsignatures $\tau$ such that the class ${\bf R}(\tau)$ of representable reducts fails to have the finite representation property.
-For instance, $\{;, \cdot\}$, see \cite{maddux2016finite}.
+There are subsignatures $\tau$ such that the class ${\bf R}(\tau)$ of representable reducts fails to have the finite representation property, for example, $\{;, \cdot\}$, see \cite[Theorem 4.1]{hirsch2021finite}.
 In general, (un)decidability of determining whether a finite relation algebra has a finite representation is an open question \cite[Problem 18.18]{hirsch2002relation}.
@@
 
-After that, we show that ${\bf R}(;,+)$ has the finite representation property using the same relational representation of quantales.
-In particular, we prove that, if $\mathcal{A} \in {\bf R}(;,+)$ is finite and representable, then the image $h(\mathcal{A})$ is isomorphic to its quantale representation.
-Here $h$ is a representation of $\mathcal{A}$.
-As a consequence, the representability of finite join-semilattice ordered semigroups is decidable.
+After that, we show that ${\bf R}(;,+)$ has the finite representation property.
+Moreover, we extract the explicit recursively enumerable axiomatisation of ${\bf R}(;,+)$.
+As a corollary, the representability problem for finite join semilattice-ordered semigroup is decidable.
 
@@
   \end{center}
-  ${\bf RS}$ is the class of all residuated semigroups.
 
@@
 
- Andr\'{e}ka and Mikul\'{a}s proved the representation theorem for ${\bf RS}$ (\cite{andreka1994lambek}) in the step-by-step fashion.
+ Andr\'{e}ka and Mikul\'{a}s proved the representation theorem for residuated semigroups (\cite{andreka1994lambek}) in the step-by-step fashion.
 See this paper to learn more about step-by-step representations in general \cite{hirsch1997step}.
@@
 
-Given a quantale $\mathcal{Q} = \langle Q, ;, \Sigma \rangle$, a set of \emph{generators} is a subset $\mathcal{G}(\mathcal{Q}) \subseteq \mathcal{Q}$, if
+\begin{definition}\label{gen}
+Given a quantale $\mathcal{Q} = \langle Q, ;, \Sigma \rangle$, a set of \emph{generators} is a subset $G \subseteq \mathcal{Q}$ (denotes as $\langle G \rangle$), if
 \begin{enumerate}
-  \item For every $q \in Q$ one has $q \leq \Sigma \{ g \in \mathcal{G}(\mathcal{Q}) \: | \: g \leq q \}$.
-  \item For every $g \in \mathcal{G}(\mathcal{Q})$ and $q_1, q_2 \in \mathcal{Q}$, $g \leq q_1 ; q_2$ implies $g \leq q_1 ; r$ for some $r \in \mathcal{G}(\mathcal{Q})$ with $r \leq q_2$.
+  \item For every $q \in Q$ one has $q \leq \Sigma \{ g \in \langle G \rangle \: | \: g \leq q \}$.
+  \item For every $g \in \langle G \rangle$ and $q_1, q_2 \in \mathcal{Q}$, $g \leq q_1 ; q_2$ implies $g \leq q_1 ; r$ for some $r \in \langle G \rangle$ with $r \leq q_2$.
 \end{enumerate}
+\end{definition}
 
-Every generator is completely join-irreducible, this follows from the definition immediatly.
+As a corollary, every $q \in \mathcal{Q}$ has the form $q = \Sigma \{ g \in \langle G \rangle \: | \: g \leq q \}$.
+Every generator is also an atom, that is, a minimal non-zero element.
+The existence of a set of generators for an arbitrary quantale was shown here \cite[Lemma 3.12]{brown1993representation}.
 
@@
 
-One may embed any residuated semigroup into some quantale with the Dedekind-MacNeille completion (see, for example, \cite{theunissen2007macneille}) as follows.
+One may embed any residuated semigroup into some quantale with using Dedekind-MacNeille completion (see, for example, \cite{theunissen2007macneille}) as follows.
 According to Goldblatt \cite{goldblatt2006kripke}, residuated semigroups have the following representation based on quantic nuclei and the Galois connection.
@@
 \end{proposition}
+
+Now let us discuss the representation theorem for quantales.
+Let $\mathcal{Q}$ be a quantale, $Q$ its underlying set, and $\langle G \rangle$ a set of its generators.
+Given $a \in \mathcal{Q}$, define the binary relation $\hat{a} \subseteq Q \times Q$ as:
+
+\begin{center}
+  $\hat{a} = \{ (g,p) \: | \: g \in \langle G \rangle, p \in Q \:\: g \leq a ; p \}$
+\end{center}
+Denote $\widehat{Q}$ as $\{ \hat{a} \: | \: a \in \mathcal{Q} \}$.
 
@@
 
+We  note  that  a  similar  construction  does  not  work  for  finite  representableupper semilattice-ordered semigroups.
+From the one hand, the notions of afinite upper semilattice-ordered semigroup and finite quantale are quite close to each other.
+From the other hand, there are examples of non-representable upper semilattice-ordered semigroups whose any non-principal ultraproduct is representable, see \cite[Theorem 3.1]{andreka2011axiomatizability}.
+That makes ${\bf R}(;,+)$ not finitely axiomatisable.
+Thus, one needs to show the decidability of membership in a different way.
+We have the following fact:
+
+\begin{proposition}\label{dec}
+  Let $\tau$ be a signature of operations and predicates definable in ${\bf RA}$, then, if ${\bf R}(\tau)$ has a recursively enumerable axiomatisation and the finite representation property, then the membership problem for finite $\tau$-structures is decidable.
+\end{proposition}
+
+\begin{proof}
+  Suppose ${\bf R}(\tau)$ has a recursively enumerable axiomatisation and every finite
+  $\mathcal{A} \in {\bf R}(\tau)$ is representable over a finite base.
+
+  Let $\mathcal{A}$ be a $\tau$-structure. We check axioms of $\mathcal{A} \in {\bf R}(\tau)$  with the algorithm $S$ that yields ``no'' and halts, if $\mathcal{A}$ is not representable.
+
+  In parallel, we check whether $\mathcal{A}$ has a finite representation.
+If there exists such $0 < n < \omega$ such that $\mathcal{A}$ is representable over size $n$, then the algorithm $T$ does output ``yes'' and terminates.
+
+  So, one has either $S(\mathcal{A}) = \text{no}$ and $T(\mathcal{A}) = \text{yes}$.
+\end{proof}
+
+In this section, we show that ${\bf R}(+, ;)$ has a recursively enumerable universal axiomatisation.
+For that, we characterise representability using Ehrenfeucht-Fra\"{i}ss\'{e} games on networks.
+After that, we show that every finite $\mathcal{A} \in {\bf R}(+, ;)$ is representable by factorising its network through an equivalence relation having a finite index.
+That is, an equivalence relation whose quotient is finite.
+
 \subsection{Axiomatisating ${\bf R}(+, ;)$}
-
-In this subsection, we show that ${\bf R}(+, ;)$ has a recursively enumerable universal axiomatisation.
-For that, we characterise representability using Ehrenfeucht-Fra\"{i}ss\'{e} games on prenetworks.
 
@@
     \item ({\bf Saturation condition}) For every $u, v \in V$ and for every $x,y,z \in \mathcal{A}$ $z \in l(u, v)$ and $z \leq x \: ; \: y$ implies $x \in N(u, w)$ and $y \in l(w, v)$ for some $w \in V$.
-    \item ({\bf Coherence conidition}) For every $u, v, w \in V$ one has $l(u, v) ; l(v, w) \subseteq l(u, w)$.
+    \item ({\bf Coherence condition}) For every $u, v, w \in V$ one has $l(u, v) ; l(v, w) \subseteq l(u, w)$.
     \item ({\bf Join-primeness}) For every $u, v \in V$ $l(u,v)$ is join-prime. That is, for each $a, b \in \mathcal{A}$ if $a + b \in l(u,v)$, then either $a \in l(u,v)$ or $b \in l(u,v)$.
@@
   \begin{enumerate}
-    \item Round $0$: $\forall$ picks elements $a, b \in \mathcal{A}$ with $a \not\leq b$.
-$\exists$ responds with a prenetwork $\mathcal{N}_0 = (V_0 = \{ x_0, x_1 \}, E_0 = \{ (x_0, x_1)\}, l_0)$ such that $l_0(x_0, x_1) = \up{b}$.
+    \item Round $0$: $\forall$ picks $a, b \in \mathcal{A}$ such that $a \not\leq b$.
+$\exists$ responds with a prenetwork $\mathcal{N}_0 = (V_0 = \{ x_0, x_1 \}, E_0 = \{ (x_0, x_1)\}, l_0)$ such that $l_0(x_0, x_1) = A$, where $A$ is an upward closed set such that $a \in A$ and $b \notin A$.
     \item Round $n + 1$. Suppose, the prenetwork $\mathcal{N}_n = (V_n, E_n, l_n)$ has been played.
@@
     \begin{enumerate}
-      \item ({\bf Composition move}): $\forall$ can pick $x, y, z \in V_n$ with $a \in l_n(x, y)$ and $b \in l_n(y, z)$.
-We denote such a move as $N(x,y,z,a,b)$.
+      \item ({\bf Composition move}): $\forall$ can pick $x, y, z \in V_n$ with $b \in l_n(x, y)$ and $c \in l_n(y, z)$.
+We denote such a move as $N(x,y,z,b,c)$.
 Then $\exists$ responds with $\mathcal{N}_{n + 1} = (V_{n + 1}, E_{n + 1}, l_{n + 1})$ such that $\mathcal{N}_{n + 1}$
-      is the same as $\mathcal{N}_n$, but $l_{n + 1}(x, z) = \up{(l_{n}(x, z)} \cup \{ a \: ; \: b\})$.
+      is the same as $\mathcal{N}_n$, but $l_{n + 1}(x, z) = \up{(l_{n}(x, z)} \cup \{ b \: ; \: c \})$.
       \item ({\bf Witness move}):
 
-      $\forall$ picks nodes $x, y \in V_n$ and $a, b \in \mathcal{A}$ such that $c \leq a ; b$ for $c \in l_n(x, y)$.
+      $\forall$ picks nodes $x, y \in V_n$ and $d, e \in \mathcal{A}$ such that $c \leq d ; e$ for $c \in l_n(x, y)$.
 $\exists$ has to find a witness.
-She responds with a fresh node $z$.
-We define the prenetwork $T$, the edges of which are $x, y, z$ with labelling:
+She has to find a $z$ which is either a fresh node or an old one.
+If $z$ is fresh, then she defines the prenetwork $T$, the edges of which are $x, y, z$ with the labelling:
       \begin{enumerate}
-        \item $l_T(x, z) = \up{a}$
-        \item $l_T(z, y) = \up{b}$
+        \item $l_T(x, z) = \up{d}$
+        \item $l_T(z, y) = \up{e}$
       \end{enumerate}
+      If $z$ is already an element of $\mathcal{A}$, then her response is similar.
       For her response, $\exists$ plays $N_{n + 1} = N_n \cup T$
@@
 
-      $\forall$ picks nodes $x, y \in V_n$ and $a + b$ for $a, b \in \mathcal{A}$. $\exists$ has the following two alternatives for her response:
+      $\forall$ picks nodes $x, y \in V_n$ and $c + d$ for $c, d \in \mathcal{A}$. $\exists$ has the following two alternatives for her response:
       \begin{enumerate}
-        \item $\exists$ chooses $a$ and responds with the prenetwork $\mathcal{N}_{n + 1} = \langle V_{n + 1}, E_{n + 1}, l_{n + 1} \rangle$, where $l_{n+1}(x, y) = \up{(l_n(x,y) \cup \{ a \})}$.
-        \item $\exists$ chooses $b$. The response is similar but $l_{n + 1}(x, y) = \up{(l_n(x,y) \cup \{ b \})}$.
+        \item $\exists$ chooses $c$ and responds with the prenetwork $\mathcal{N}_{n + 1} = \langle V_{n + 1}, E_{n + 1}, l_{n + 1} \rangle$, where $l_{n+1}(x, y) = \up{(l_n(x,y) \cup \{ c \})}$.
+        \item $\exists$ chooses $b$. The response is similar but $l_{n + 1}(x, y) = \up{(l_n(x,y) \cup \{ d \})}$.
       \end{enumerate}
@@
 
-  $\exists$ wins for the match of the game $\mathcal{G}_n(\mathcal{A})$ if for each $0 \leq i \leq n \leq \omega$ $\mathcal{N}_i$ is a network.
-$\exists$ has a winning strategy, if she can win all the matches.
-Otherwise, $\forall$ wins the game.
+  $\forall$ wins the play if there exists $b \leq a$ such that $b \in \mathcal{A}$ such that $b \in l_{\mathcal{N}_i}(x,y)$ for some $i < n$. Otherwise, $\exists$ wins the play.
+
+  Let $a \in \mathcal{A}$ and $\mathcal{N}$ a network, define a game $\mathcal{G}(\mathcal{N}, \mathcal{A}, a)$ such that $\forall$ picks $a$ in the initial round and $\mathcal{N}_0 = \mathcal{N}$.
+The rules of the game are the same as previously.
 \end{definition}
@@
   \begin{enumerate}
-    \item This item has the proof similar to the first item of \cite[Proposition 5]{hirsch2005class}.
+    \item Let $h : \mathcal{A} \to 2^{D \times D}$ is a representation of the base set $D \neq \emptyset$.
+$\exists$ maintains a map ${}^{'} : \operatorname{Nodes}(\mathcal{N}) \to D$, where $\mathcal{N}$ is a network being played, such that $(x',y') \in h(a)$ iff $\up a \subseteq l_{\mathcal{N}}(x,y)$.
 
     \item
-    Let $a, b \in \mathcal{A}$ such that $a \not\leq b$.
-We consider a play of the game where $\forall$ pick the pair $(a, b)$ in the initial round and plays $(\mathcal{N}, x, y, z, c, d)$ in the further rounds for each $x, y, z \in \operatorname{Nodes}(\mathcal{N})$ and $c,d \in \mathcal{A}$.
+    Let $a \in \mathcal{A}$.
+We consider a play of the game where $\forall$ picks $a$ and $b$ with $a \not\leq b$ in the initial round and plays $(\mathcal{N}, x, y, z, c, d)$ in the further rounds for each $x, y, z \in \operatorname{Nodes}(\mathcal{N})$ and $c,d \in \mathcal{A}$.
 Here, $c \in l_N(x, y)$ and $d \in l_N(y, z)$.
@@
 
-    Let us put $\mathcal{N}^{*}(a, b) = \bigcup \limits_{i < \omega} \mathcal{N}_i$. It is clear that $\mathcal{N}^{*}$ is a network.
+    Let us put $\mathcal{N}^{*}(a) = \bigcup \limits_{i < \omega} \mathcal{N}_i$. It is clear that $\mathcal{N}^{*}$ is a network.
 
-    Let us put the following network:
+    Let us put the following network assuming that $\mathcal{N}^{*}(a_1, a_2)$ and $\mathcal{N}^{*}(b_1, b_2)$ are disjoint for $a_1 \neq 2$ and $b_1 \neq b_2$:
 
     \begin{center}
-      $\mathcal{N} = \bigcup \limits_{\substack{a, b \in \mathcal{A} \\ a \not\leq b}} \mathcal{N}^{*}(a,b)$
+      $\mathcal{N} = \bigcup \limits_{a,b \in \mathcal{A}, a \not\leq b} \mathcal{N}^{*}(a,b)$
     \end{center}
@@
 
-    The labelling function $l_{\mathcal{N}}$ respects $;$. Here, the argument is similar to the proof of \cite[Proposition 5]{hirsch2005class}.
-
     Let us define the $\operatorname{rep} : \mathcal{A} \to \operatorname{Nodes}(\mathcal{N})$:
     \begin{center}
-      $\operatorname{rep}(a) = \{ (x, y) \: | \: \exists b \in \down{a} \:\: b \in l_{\mathcal{N}}(x, y)\}$
+      $\operatorname{rep}(a) = \{ (x, y) \: | \: \exists b \leq a \:\: b \in l_{\mathcal{N}}(x, y)\}$
     \end{center}
@@
 
-    \begin{itemize}
-  \item Let us show that $\operatorname{rep}(a + b) = \operatorname{rep}(a) \cup \operatorname{rep}(b)$
+Let us show that $\operatorname{rep}(a + b) = \operatorname{rep}(a) \cup \operatorname{rep}(b)$
 
@@
 \begin{center}
-$a + b \in \bigcup \limits_{ \substack{a_1, a_0 \in \mathcal{A} \\ a_1 \not\leq a_0}} l_{\mathcal{N}^{*}(a_1, a_0)}(x,y)$
+$a + b \in \bigcup \limits_{ c_1, c_2 \in \mathcal{A}, c_1 \not\leq c_2} l_{\mathcal{N}^{*}(c_1, c_2)}(x,y)$
 \end{center}
 
-That is, there are $a_1, a_0 \in \mathcal{A}$ with $a_1 \not\leq a_0$ such that $a + b \in l_{\mathcal{N}^{*}(a_1, a_0)}(x,y)$.
-This means $a + b \in \bigcup \limits_{i < \omega} \mathcal{N}_i$.
-$\forall$ makes all possible join moves, so either $a \in l_{\mathcal{N}_{i + 1}}(x, y)$ or $b \in l_{\mathcal{N}_{i + 1}}(x, y)$ depending on what $\exists$ chooses.
-Thus, $(x, y) \in \operatorname{rep}(a) \cup \operatorname{rep}(b)$.
+That is, there is $c \in \mathcal{A}$ with such that $a + b \in l_{\mathcal{N}^{*}(c_1, c_2)}(x,y)$, but $l_{\mathcal{N}^{*}(c_1, c_2)}(x,y)$ is join-prime, so we have either $a \in l_{\mathcal{N}^{*}(c_1, c_2)}(x,y)$ or $b \in l_{\mathcal{N}^{*}(c_1, c_2)}(x,y)$.
+Thus, $\operatorname{rep}(a + b) \subseteq \operatorname{rep}(a) \cup \operatorname{rep}(b)$.
 
@@
 
-\item Let us show that $\operatorname{rep}(a ; b) = \operatorname{rep}(a) ; \operatorname{rep}(b)$.
+Let us show that $\operatorname{rep}(a ; b) = \operatorname{rep}(a) ; \operatorname{rep}(b)$.
 
@@
 By the condition, $\forall$ makes the witness moves and $\exists$ responds with a witness.
-Her response is a fresh node $z$ such that $l_{\mathcal{N}_{i + 1}}(x, z) =  \up{(l_{\mathcal{N}_i}(x, z) \cup \{ a \})}$ and $l_{\mathcal{N}_{i + 1}}(z, y) = \up{(l_{\mathcal{N}_i}(z, y) \cup \{ b\})}$.
+Her response is a node $z$ such that $l_{\mathcal{N}_{i + 1}}(x, z) =  \up{(l_{\mathcal{N}_i}(x, z) \cup \{ a \})}$ and $l_{\mathcal{N}_{i + 1}}(z, y) = \up{(l_{\mathcal{N}_i}(z, y) \cup \{ b\})}$.
 The inclusion $\operatorname{rep}(a ; b) \subseteq \operatorname{rep}(a) ; \operatorname{rep}(b)$ holds since all witness moves have been played and for every such a move there exists a witness.
 
-\item Suppose $(x, y) \in \operatorname{rep}(a) ; \operatorname{rep}(b)$.
+Suppose $(x, y) \in \operatorname{rep}(a) ; \operatorname{rep}(b)$.
 We need $(x, y) \in \operatorname{rep}(a;b)$.
@@
 
-\end{itemize}
+For injectivity, suppose $a \leq b$ and $(x, y) \in \operatorname{rep}(a)$, that is, there is $c \leq a$ such that $c \in l_{\mathcal{N}}(x, y)$, but $c \leq a \leq b$, so $(x, y) \in \operatorname{rep}(b)$.
+
+Suppose $a \not\leq b$, then there are $x, y \in \operatorname{Nodes}(\mathcal{N}(a, b))$ such that $a \in l_{\mathcal{N}}(x, y)$.
+These elements are $x_0, x_1$ that $\exists$ picks as her response in the zero round.
+$\exists$ has a winning strategy, so $b \notin l(x_0, x_1)$, but $(x,y) \in \operatorname{rep}(a)$, but $(x,y) \notin \operatorname{rep}(b)$.
   \end{enumerate}
@@
 \begin{definition}
-  Let $\operatorname{Var} = \{ v_0, v_1, ... \}$ be a countable set of variables. The set of terms $T$ is generated by the following grammar:
+  Let $\operatorname{Var}$ be a set of variables. The set of terms $T$ is generated by the following grammar:
 
   \begin{center}
-    $T ::= \operatorname{Var} \: | \: (T + T) \: | \: (T ; T)$
+    $\operatorname{Term} ::= \operatorname{Var} \: | \: (T + T) \: | \: (T ; T)$
   \end{center}
@@
 
-Let us define the equivalence relation $\approx$ on $T$ as follows for terms $t_1, t_2$:
-\begin{center}
-$t_1 \approx t_2$ iff $F(\omega) \models t_1 = t_2$.
-\end{center}
-where $F(\omega)$ is the free countably generated algebra join semilattice-ordered semigroup.
-We have the following partial order on $T$ defined as
-\begin{center}
-  $t_1 \preceq t_2$ iff $t_1 + t_2 \approx t_2$, that is, $F(\omega) \models t_1 + t_2 = t_2$
-\end{center}
-
-We call subset $A \subseteq T$ of terms called upward closed if for every term
-$t \in A$ one has $\up{t} \subseteq A$, where $\up{t} = \{ t_1 \in T \: | \: t \preceq t_1 \}$.
-
 \begin{definition}
-  A \emph{term network} is a graph $\langle V, E, l \rangle$, where $\langle V, E \rangle$ is a directed graph and $l : E \to \operatorname{Up}(F(\omega), \preceq)$ is a labelling function.
+  A \emph{term network} is a finite network $\langle V, E, l \rangle$, where $\langle V, E \rangle$ is a directed graph and $l : E \to 2^{Term}$ is a labelling function such that every $l_(x,y)$ is finite for each $(x, y) \in E$.
 \end{definition}
@@
 
-As usual, one may think of $\vartheta$ as a homomorphism $\vartheta : F(\omega) \to \mathcal{A}$.
-
 Let $\mathcal{N} = \langle V, E, l \rangle$ a term network, $\mathcal{A}$ be a join-semilattice ordered semigroup and $\vartheta : \operatorname{Var} \to {\mathcal{A}}$ a valuation.
-Let us define the prenetwork $\mathcal{N}^{\vartheta}$ with the same edges and vertices with labelling $l^{\vartheta}(x, y) = \{ t^{\vartheta} \: | \: t \in l_{\mathcal{N}}(x, y)\}$.
-
-Let us ensure that $\mathcal{N}^{\vartheta}$ is well-defined.
-Let us consider $\vartheta$ as a homomorphism $F(\omega) \to \mathcal{A}$.
-We would like to know whether $\up{t^{\vartheta}} \subseteq l^{\vartheta}(x,y)$ for $t^{\vartheta} \in l^{\vartheta}(x,y)$.
-So $t \in l(x,y)$.
-Suppose $t \preceq t_1$, then $t_1$ clearly belongs to $l(x,y)$.
-So $l^{\vartheta}(x,y)$ is upward closed by the construction.
+Let us define the prenetwork $\mathcal{N}^{\vartheta}$ with the same edges and vertices with labelling $l^{\vartheta}(x, y) = \bigcup \{ F \in \operatorname{Up}(\mathcal{A}, \leq) \: | \: l_{\mathcal{N}}(x, y) \subseteq F \}$.
 
@@
   \item Let $x, y \in \operatorname{Nodes}(\mathcal{N})$ and let $t$ be a term. $\mathcal{N}_c$ is the extension of $\mathcal{N}$,
-  where $\operatorname{Nodes}(\mathcal{N}_c) = \operatorname{Nodes}(\mathcal{N})$ and $\up{l_{\mathcal{N}}(x, y) \cup t}$ and $l_{\mathcal{N}_c}(u, v) = l_{\mathcal{N}}(u, v)$ for each $u \neq x$ and $v \neq y$.
+  where $\operatorname{Nodes}(\mathcal{N}_c) = \operatorname{Nodes}(\mathcal{N})$ and $l_{\mathcal{N}_c}(x, y) = l_{\mathcal{N}}(x, y) \cup \{ t\}$ and $l_{\mathcal{N}_c}(u, v) = l_{\mathcal{N}}(u, v)$ for each $u \neq x$ and $v \neq y$.
 We denote this network as $\mathcal{N}_c(\mathcal{N}, x, y, t)$.
   \item
-  Let $x, y \in \operatorname{Nodes}(\mathcal{N})$, let $z$ be a fresh node, and $t_1$, $t_2$ any terms.
+  Let $x, y \in \operatorname{Nodes}(\mathcal{N})$, let $z$ be a node (regardless of whether $z \in \operatorname{Nodes}(\mathcal{N})$), and $t_1$, $t_2$ any terms.
 Let us define a network $T$ such that $\operatorname{Nodes}(T) = \{ x, y, z\}$.
-We define labelling as $l_{T}(x, y) = \up{t_1}$ and $l_{T}(y, z) = \up{t_2}$.
+We define labelling as $l_{T}(x, y) = \{ t_1 \}$ and $l_{T}(y, z) = \{ t_2 \}$.
 So we put $\mathcal{N}_w = \mathcal{N} \cup T$.
@@
   \item Let $x, y \in \operatorname{Nodes}(\mathcal{N})$ and let $t_1, t_2$ be terms
-  Let us define $T_i = \langle \{ x, y\}, \{ (x, y) \}, l_{T_i} \rangle$, where $l_{T_i}(x, y) = \up{(l_{\mathcal{N}}(x, y) \cup t_i)}$ for $i = 1,2$.
+  Let us define $T_i = \langle \{ x, y\}, \{ (x, y) \}, l_{T_i} \rangle$, where $l_{T_i}(x, y) = l_{\mathcal{N}}(x, y) \cup \{ t_i \}$ for $i = 1,2$.
 So $\mathcal{N}_{j_1} = \mathcal{N} \cup T_1$ and $\mathcal{N}_{j_2} = \mathcal{N} \cup T_2$.
-  We denote this as $\mathcal{N}_j(\mathcal{N}, x, y, t_1, t_2)$.
 \end{enumerate}
@@
 \begin{lemma}\label{ax}
-  For each $n < \omega$ there exists a first-order sentence $\sigma_n$ such that $\exists$ has a winning strategy in $\mathcal{G}_n(\mathcal{A})$ iff $\mathcal{A} \models \sigma_n$.
+  For each $n < \omega$ there exists a first-order sentence $\rho_n$ such that $\exists$ has a winning strategy in $\mathcal{G}_n(\mathcal{A})$ iff $\mathcal{A} \models \rho_n$.
 \end{lemma}
@@
   \end{center}
-  where $\mathcal{A}$ is a join semilattice-ordered semigroup and $\vartheta : \operatorname{Var} \to \mathcal{A}$ is a variable assignment.
+  where $\mathcal{A}$ is a join semilattice-ordered semigroup, $\vartheta : \operatorname{Var} \to \mathcal{A}$ is a variable assignment, and $\mathcal{N}$ is a term network.
 
@@
 
-This formula merely claims that $\exists$ has a winning strategy in the zero length game.
-
-    \item Suppose $\sigma_{n}(\mathcal{N}, v)$ for $n < \omega$.
-Let us define a formula $\sigma_{n + 1}$ claiming that $\exists$ always has a proper response for a network $\mathcal{N}$ being played (which is defined by $\sigma_{n}(\mathcal{N}, v)$).
+$\sigma_0(\mathcal{N}, v)$ merely claims that $\exists$ has a winning strategy in the zero length game.
+    \item Suppose $\sigma_{n}(\mathcal{N}, v)$ are already constructed for some $n < \omega$.
+Let us define a formula $\sigma_{n + 1}$ claiming that $\exists$ always has a proper response for a network $\mathcal{M}$ being played.
 
@@
       \item ${\sigma_{n + 1}}_c(\mathcal{N}, v) = \bigwedge \limits_{x, y, z \in \operatorname{Nodes}(\mathcal{N}), t_1 \in l_{\mathcal{N}}(x, y), t_2 \in l_{\mathcal{N}}(y, z)} \sigma_{n}(\mathcal{N}_c(x, z, t_1, t_2), v)$
-      \item ${\sigma_{n + 1}}_w(\mathcal{N}, v) = \bigwedge \limits_{x,y \in \operatorname{Nodes}(\mathcal{N}), t \in l_{\mathcal{N}}(x,y)} \forall u_1, u_2 (t \leq u_1 ; u_2 \rightarrow \bigvee \limits_{w \in \operatorname{Nodes} \cup \{ z \}} \mathcal{N}_c(x,y,w, t_1, t_2))$, where $z$ is a fresh node.
-      \item ${\sigma_{n + 1}}_j(\mathcal{N}, v) = \bigwedge \limits_{x, y \in \operatorname{Nodes}(\mathcal{N})} \sigma_{n}(\mathcal{N}_{j_1}(\mathcal{N}, x, y, t_1, t_2), v) \lor \sigma_{n}(\mathcal{N}_{j_2}(\mathcal{N}, x, y, t_1, t_2), v)$
+      \item ${\sigma_{n + 1}}_w(\mathcal{N}, v) = \bigwedge \limits_{x,y \in \operatorname{Nodes}(\mathcal{N}), t \in l_{\mathcal{N}}(x,y)} \forall u_1, u_2 (t \leq u_1 ; u_2 \rightarrow \bigvee \limits_{w \in \operatorname{Nodes}(\mathcal{N}) \cup \{ z \}} \mathcal{N}_c(x,y,w, u_1, u_2))$, where $z \notin \operatorname{Nodes}(\mathcal{N})$.
+      \item ${\sigma_{n + 1}}_j(\mathcal{N}, v) = \forall a \: \forall b (v = a + b \rightarrow \bigwedge \limits_{x, y \in \operatorname{Nodes}(\mathcal{N})} \sigma_{n}(\mathcal{N}_{j_1}(\mathcal{N}, x, y, a), v) \lor \sigma_{n}(\mathcal{N}_{j_2}(\mathcal{N}, x, y, b), v))$
     \end{itemize}
@@
   So, $\exists$ has a winning strategy iff these formulas are true under the valuation $\vartheta$ since the formulas $\{ \sigma_n \}_{n < \omega}$ encode the presence of a winning strategy for $\exists$ on every finitary round.
-\end{proof}
 
-We have already considered finite and countable algebras. Let us consider also arbitrary to complete the axiomatisation of ${\bf R}(;,+)$. We have the following statement.
-
-\begin{lemma}
-  Let $\mathcal{A}$ be a join-semilattice semigroup, then $\exists$ has a winning strategy has a winning strategy in $\mathcal{G}_m(\mathcal{A})$ iff $\mathcal{A}$ is elementary equivalent to some countable representable join-semilattice semigroups.
-\end{lemma}
-
-\begin{proof}
-  The $\Leftarrow$ implication is obvious.
-If $\mathcal{A} \equiv \mathcal{A}'$, where $|\mathcal{A}'| = \omega$ and $\mathcal{A}' \in {\bf R}(;,+)$, then $\exists$ has a winning strategy in $\mathcal{G}_{\omega}(\mathcal{A}')$ (Lemma~\ref{rep}), so $\exists$ has a winning strategy in
-  every $\mathcal{G}_{m}(\mathcal{A}')$ for $m < \omega$ (Proposition~\ref{fin}). Thus, so $\exists$ has a winning strategy in
-  every $\mathcal{G}_{m}(\mathcal{A})$ since $\mathcal{A}$ and $\mathcal{A}'$ are elementary equivalent.
-
-  Suppose $\exists$ has a winning strategy has a winning strategy in $\mathcal{G}_m(\mathcal{A})$ for $m < \omega$.
-The proof scheme is the same as for relation (see \cite[Lemma 19]{hirsch1997complete}) or cylindric algebras (see \cite[Corollary 3.3.5]{hirsch2013completions}), so we provide the following sketch.
-In particular, we have the following proof that may be proved using \strokeL o\'s's theorem (see \cite[Theorem 9.5.1]{hodges1993model}) similarly to.
-
-  \begin{claim}
-    If $\exists$ has a winning strategy has a winning strategy in $\mathcal{G}_m(\mathcal{A})$ for $m < \omega$, then $\exists$ has a winning strategy for $\mathcal{G}_{\omega}(\Pi_{U} \mathcal{A})$, where $\Pi_{U} \mathcal{A}$ is a non-principal ultrapower of $\mathcal{A}$ for $U$, an ultrafilter over $\omega$.
-  \end{claim}
-
-  Consider $\mathcal{B}$, a countable elementary equivalent subalgbera of $\Pi_{U} \mathcal{A}$ that exists by the downward L\"{o}wenheim-Skolem theorem, see \cite[Corollary 3.1.5]{hodges1993model}.
-  We construct the following elementary chain inxeded over $\omega$:
+  Let $v_0$ be any variable, $\mathcal{N}_{v_0}$ denotes the term network having the form
+  $\langle \{ \{ x_0, x_1 \}, \{ (x_0, x_1) \}, l \} \rangle$, where $l(x,y) = \{ v_0 \}$.
+  We define the following sequence of formulas $(\rho_n)_{n < \omega}$:
   \begin{center}
-    $\mathcal{B}_0 \preceq \mathcal{B}_1 \preceq \dots \mathcal{B}_i \preceq\mathcal{B}_{i + 1} \preceq \dots (\preceq \Pi_{U} \mathcal{A})$.
+    $\rho_n = \forall v_0 \forall v_1 (\neg (v_0 \leq v_1) \to \sigma(\mathcal{N}_{v_0}, v_0))$
   \end{center}
-  Let us put $\mathcal{B}_0 := \mathcal{B}$.
-  Suppose the initial segment $\mathcal{B}_0 \preceq \mathcal{B}_1 \preceq \dots \mathcal{B}_i$ has been already constructed. Let us constuct $\mathcal{B}_{i + 1}$.
-
-  Consider a function $\sigma$ that takes an initial game of the play $\mathcal{G}_{\omega}(\Pi_{U} \mathcal{A})$, say, $\mathcal{N}_0 \subseteq \mathcal{N}_1 \subseteq \dots \mathcal{N}_k$ (all $k$) and Abelard's move and returns $\exists$'s repsonse, $\mathcal{N}_{k + 1}$, an extension of $\mathcal{N}_k$.
-
-  Consider $\sigma \upharpoonright_{\mathcal{B}_i}$, a restriction of $\sigma$, where $\forall$ takes elements only from $\mathcal{B}_i$ for his moves.
-The range of this restriction consists of $\Pi_{U} \mathcal{A}$-networks, and $E$, the set of elements that occur in those networks is also countable.
-We define $\mathcal{B}_{i + 1}$ as a countable elementary subalgebra of $\Pi_{U} \mathcal{A}$ that contains $\mathcal{B}_{i}$ and $E$.
-The existence follows from the downward L\"{o}wenheim-Skolem theorem.
-
-  Let us put $\mathfrak{B} = \bigcup \limits_{i < \omega} \mathcal{B}_i$.
-Note that $\mathfrak{B}$ is countable.
-By Tarksi-Vaught elementary chain union theorem (see \cite[Theorem 2.5.2]{hodges1993model}), for each $i < \omega$ $\mathcal{B}_i \preceq \mathfrak{B}$.
-Therefore, $\mathfrak{B} \equiv \Pi_{U} \mathcal{A}$.
-  On the other hand, $\mathcal{A} \equiv \Pi_{U} \mathcal{A}$ (see \cite[Corollary 9.5.2]{hodges1993model}).
-  Moreover, $\mathfrak{B}$ is representable and $\mathfrak{B} \equiv \mathcal{A}$.
 \end{proof}
@@
 
-\begin{theorem}
+\begin{theorem}\label{axiomatisation}
   ${\bf R}(+, ;)$ is axiomatised with the axioms of join semilattice-ordered semigroups and the coutable sequences of universal formulas $\{ \rho_n \}_{n < \omega}$.
@@
 \begin{proof}
-  Being the class of ${\bf RRA}$-reducts, ${\bf R}(+, ;)$ forms a pseudo-elementary class (see \cite{eklof1977ultraproducts} to have a rigorous definition).
+  ${\bf R}(+, ;)$ forms a pseudo-elementary class (see \cite{eklof1977ultraproducts} to have a rigorous definition).
+Let us define a two sorted language with sorts ${\bf a}$ (algebra) and ${\bf r}$ (representation).
+We have ${\bf R}(+, ;) = \{ \mathcal{M} \upharpoonright_{\bf a} \: | \: \mathcal{M} \models (\rho_m)_{n < \omega} \}$.
+  Every $\rho_n$ may be equivalently transformed into a universal formula by bringing all universal quantifiers into the prefix.
   So we apply \cite[Theorem 9.28]{hirsch2002relation}.
-Therefore, a join semilattice-ordered semigroup $\mathcal{A}$ is representable iff $\mathcal{A} \models \{ \sigma_n \}_{n < \omega}$.
-  Every $\sigma_n$ for $n < \omega$ defined in the proof of Lemma~\ref{ax} is equivalent to the universal formula $\rho_n$ by bringing all universal quantifiers into the prefix of $\sigma_n$.
 \end{proof}
@@
 
-If we extend $\mathcal{A}$, a finite join semilattice-ordered semigroup, with the bottom, then we may consided the obtained algebra as a quantale.
-By Theorem~\label{quantaleRep}, $\mathcal{A}$ is isomorphic to the relational quantale $\widehat{A}$.
-From the other hand, if $\mathcal{Q}$ is a relational quantale, then $\Sigma A \subseteq \bigcup A$ for $A \subseteq \mathcal{Q}$.
-The converse inclusion does not hold in general, see \cite[Remark 3.2 and Example 3.6]{brown1993representation}.
+Let us show finite representability for ${\bf R}(+, ;)$.
+Suppose that a finite $\mathcal{A}$ is representable, so $\exists$ has a winning strategy on $\mathcal{G}^{\omega}(\mathcal{A})$.
+Let us consider $\mathcal{N} = \langle V, E, l \rangle$, the network constructed during the $\mathcal{G}^{\omega}(\mathcal{A})$ game.
 
-Let $\mathcal{Q}$ be a quantale and $q \in \mathcal{Q}$ is called \emph{completely coprime}, if
+$\mathcal{N} = \langle V, E, l \rangle$ be a $\mathcal{A}$-network and $\mathcal{A}$ is finite. Define the following equivalence relation on $E \times E$ for $(x_1, y_1), (x_2, y_2) \in E$:
 \begin{center}
-  $q \leq \sum \limits_{i \in I} q_i$ implies $q \leq q_i$ for some $i \in I$.
+  $(x_1, y_1) \equiv (x_2, y_2)$ iff $l_{\mathcal{N}}(x_1, y_1) = l_{\mathcal{N}}(x_2, y_2)$
 \end{center}
-We denote the set of completely coprimes as $\operatorname{CCP}(\mathcal{Q})$.
-In particular, if $|I| = 2$, then $q$ is called join-prime.
-If a quantale is finite, then $\operatorname{CCP}(\mathcal{Q})$ is clearly the set of all join-prime elements that we denote as $\operatorname{JP}(\mathcal{Q})$.
+If $A$ is finite, then $E /_{\equiv}$ is finite as well since the set $\operatorname{Up}(\mathcal{A}, \leq)$ is finite.
 
-A quantale $\mathcal{Q}$ is called \emph{completely coprime algebraic} (CCPA), if every $q \in \mathcal{Q}$ has the form:
-\begin{center}
-  $q = \sum \limits \{ p \in \mathcal{Q} \: | \: p \in \operatorname{CCP}(\mathcal{Q}), p \leq q \}$
-\end{center}
+Let us construct the following $\mathcal{A}$-prenetwork $\mathcal{N}_{\equiv} = \langle V_{\equiv}, E /_{\equiv}, l \rangle$ such that:
+\begin{enumerate}
+  \item $V_{\equiv} = \bigcup \{ \{ x, y \} \: | \: [x,y]_{\equiv} \in E /_{\equiv} \}$, that is, the union of two element sets $\{ x, y \}$ such that there exists an edge $(x,y)$ up to $\equiv$.
+  \item $l([x,y]_{\equiv}) = l_{\mathcal{N}}(x,y)$
+\end{enumerate}
 
-A quantale $\mathcal{Q}$ is called \emph{simple}, if
-$\sum \limits_{i \in I} \widehat{A} = \bigcup \limits_{i \in A} \widehat{A}$ for any $A \subseteq \mathcal{Q}$.
+\begin{lemma} Let $\mathcal{A}$ be a finite join semilattice-ordered semigroup and $\mathcal{N} = \langle V, E, l \rangle$ an $\mathcal{A}$-prenetwork.
+  \begin{enumerate}
+  \item Let $a \in \mathcal{A}$, then $a \in l_{\mathcal{N}}(x,y)$ iff $a \in l([x,y]_{\equiv})$.
 
-It is known that if a quantale $\mathcal{Q}$ is CCPA, then $\operatorname{CCP}(\mathcal{Q})$ is a set of generators, and every CCPA quantale is simple, see \cite[Lemma 5.3, Lemma 5.5, and Proposition 5.6]{brown1993representation}.
-As a corollary, any quantale CCPA is isomorphic to a relational quantale where all joins are unions.
+  \item Let $\mathcal{A}$ be a finite join semilattice-ordered semigroup and $\mathcal{N}$ an $\mathcal{A}$-prenetwork.
+If $\mathcal{A}$ is a network, so is $\mathcal{N}_{\equiv}$.
+Moreover, $\mathcal{N}_{\equiv}$ is finite.
+\end{enumerate}
+\end{lemma}
 
-TODO: complete the argument
+\begin{proof}
+$ $
+  \begin{enumerate}
+  \item If $a \in l_{\mathcal{N}}(x,y)$, suppose $(x_1,y_1) \equiv (x,y)$, so $l_{\mathcal{N}}(x,y) = l_{\mathcal{N}}(x_1,y_1)$, then $a \in l_{\mathcal{N}}(x_1,y_1)$.
+Therefore, $a \in l([x,y]_{\equiv})$.
+The converse implication is obvious.
+  \item Follows from the definition of a network and the fact that the proof does not depend on the choice of equivalence classes representatives.
+  Let us check join-primeness as an example. The rest of the cases are considered in a similar fashion:
+  Let $a, b \in \mathcal{A}$ and $x, y \in V$ such that $a + b \in l([x,y]_{\equiv})$.
+  Then $a + b \in l(x,y)$ in the original network.
+Then we have either $a \in l(x,y)$ or $b \in l(x,y)$ by the join-primeness condition.
+Therefore, either $a \in l([x,y]_{\equiv})$ or $b \in l([x,y]_{\equiv})$
 
-\begin{proposition}\label{quantalerepisok}
-  Let $\mathcal{A}$ be a finite representable join-semilattice semigroup, then $\mathcal{A}_0$ is completely coprime algebraic as a quantale and, therefore, is simple.
-\end{proposition}
+  $\mathcal{N}_{\equiv}$ is finite since $E_{\equiv}$ is finite. That makes $V_{\equiv}$ finite as well.
+  \end{enumerate}
+\end{proof}
 
-As a corollary, we have the following theorem:
-
-\begin{theorem}
+\begin{theorem}\label{frpjoin}
   ${\bf R}(+, ;)$ has the finite representation property.
@@
 \begin{proof}
-  Let us show that if $\mathcal{A} \in {\bf R}(+, ;)$ such that $\mathcal{A}$ is finite. Let $h$ be a representation of $\mathcal{A}$, so $h_0$ is a representation of $\mathcal{A}_0$.
-  $h(\mathcal{A})$ is a subalgebra of $h_0(\mathcal{A}_0)$, but
-  $h_0(\mathcal{A}_0)$ is isomorphic to $\widehat{A_0}$, the relational quantale representation of $\widehat{A_0}$ for the following reason.
+  Let $\mathcal{A} \in {\bf R}(+, ;)$ and $\mathcal{A}$ is finite.
+  Let us show that $\mathcal{A}$ has a representation over a finite base.
+  We already know that $\exists$ has a winning strategy on $\mathcal{G}^{\omega}(\mathcal{A})$, so we have $\mathcal{N}$, the resulting network.
+So $\mathcal{N}_{\equiv}$ is a finite network by the previous lemma.
+  Define the following map $f : \mathcal{A} \to 2^{E_{\equiv}}$ as follows:
+\begin{center}
+  $f : a \mapsto \{ (x,y) \: | \: \exists b \leq a \: \& \: b \in l([x,y]_{\equiv}) \}$.
+\end{center}
+$f$ preserves all operations and it is injective for the same reason as $\operatorname{rep}$ constructed above. Let us check that.
 
-  Consider the map
-  $f : h_0(a) \mapsto \hat{a}$.
-$f$ preserves composition since $h$ and $\hat{.}$ are semigroup homomorphisms.
-The equality $f(h_0(a) \cup h_0(b)) = \hat{a} \cup \hat{b}$ holds since $\mathcal{A}_0$ is simple as a quantale by Proposition~\ref{quantalerepisok}.
-$f$ is also obviously injective and surjective.
+If $(x,y) \in f(a + b)$, then there is $c \leq a + b$ such that $c \in l([x,y]_{\equiv})$, then $c \in l_{\mathcal{N}}(x,y)$, so we have either $a \in l_{\mathcal{N}}(x,y)$ or $b \in l_{\mathcal{N}}(x,y)$.
+Thus, either $a \in l([x,y]_{\equiv})$ or $b \in l([x,y]_{\equiv})$.
+If $(x, y) \in f(a)$, then there exists $c \leq a$ such that $c \in l([x,y]_{\equiv})$, but $a + b \in \in l([x,y]_{\equiv})$. Therefore, $f(a + b) = f(a) \cup f(b)$.
 
-  $\widehat{\mathcal{A}_0}$ is a relational quantale on the underlying set of $\mathcal{A}_0$, but $\mathcal{A}$ is finite, so $\mathcal{A}$ has a representation over a finite base since $h(\mathcal{A})$ is isomorphic to $h_0(\mathcal{A})$, a subalgebra of $\widehat{A_0}$.
+If $(x,y) \in f(a ; b)$, then there exists $c \leq a ; b$ with $c \in l([x,y]_{\equiv})$, then $c \in l_{\mathcal{N}}(x,y)$.
+By the saturation condition, there is $z \in \operatorname{Nodes}(\mathcal{N})$ with $a \in l_{\mathcal{N}}(x,z)$ and $b \in l_{\mathcal{N}}(z,y)$.
+Then $a \in l([x,z]_{\equiv})$ and $b \in l([z,y]_{\equiv})$.
+If $(x,y) \in f(a);f(b)$, then there is $z \in V_{\equiv}$ with $(x, z) \in f(a)$ and $(z, y) \in f(b)$, then there are $c$ and $d$ with $c \leq a$ and $c \in l([x,z]_{\equiv})$ and $d \leq b \in l([z,y]_{\equiv})$ respectively.
+Then $c ; d \in l([x,y]_{\equiv})$ by the saturation condition.
+
+Suppose $a \neq b$, so we have either $a \not\leq b$ or $b \not\leq a$.
+Suppose $a \not\leq b$. Let $(x, y) \in f(a)$, then there is $c \leq a$ with $c \in l([x,y]_{\equiv})$, then $c \in l_{\mathcal{N}}(x,y))$ and, moreover, $c \in l_{\mathcal{N}^{*}(a,b)}(x,y))$.
+We know that $\exists$ has a winning strategy in $\mathcal{G}_{\omega}(\mathcal{A})$, so
+$b \notin l_{\mathcal{N}^{*}(a,b)}(x,y))$. So $b \notin l([x,y]_{\equiv})$ and $f(a) \neq f(b)$.
+The second alternative is similar to the previous one.
+
+The base set of a representation, $V /_{\equiv}$, is finite, see the previous lemma.
 \end{proof}
@@
 \begin{proof}
-  Let $\mathcal{A}$ be a finite semilattice-ordered semigroup. We check axioms defined in Lemma~\ref{ax} with the algorithm $S$ that yields ``no'' and halts, if $\mathcal{A}$ is not representable.
-
-  In parallel, we check whether $\mathcal{A}$ has a finite representation.
-If there exists such $0 < n < \omega$ such that $\mathcal{A}$ is representable over size $n$, then the algorithm $T$ does output ``yes'' and terminates.
-
-  So, one has either $S(\mathcal{A}) = \text{no}$ and $T(\mathcal{A}) = \text{yes}$.
+  Follows from Proposition~\ref{dec}, Theorem~\ref{axiomatisation}, and Theorem~\ref{frpjoin}.
 \end{proof}

### 2021-05-30 filtration-stuff/NewText.tex
filtrations for K5 * K5

@@
 \usepackage{graphicx}
+\usepackage{wasysym}
 
@@
   }
-\end{proof}
 
@@
 On the other hand, $R_{min}^E \not\subseteq R_{max}$, so is not $\widehat{R}^{\mathcal{E}}$.
+
+\end{proof}
+
+Let us define the logic $\mathcal{L}$ as ${\bf K} \oplus \Diamond \Diamond \Diamond p \to \Diamond p$.
+Let $R$ be a binary relation, the $\mathcal{L}$-closure of $R$ is defined (denoted as $R^{\davidsstar}$) as the following union:
+\begin{center}
+  $R^{\davidsstar} = R \cup R^3 \cup R^5 \cup \dots \cup R^{2k+1} \cup \dots$
+\end{center}
+
+\begin{theorem}
+  $\mathcal{L}$ does not admit strict filtrations.
+\end{theorem}
+
+\begin{proof}
+  Consider the following frame $\mathcal{F} = \langle W, R \rangle$:
+
+\vspace{\baselineskip}
+
+  \xymatrix{
+  &&&&& x \ar@(ur,ul) \ar[r] & y \\
+  &&&&&& y_1 \ar[r] \ar@/_/[rr] & z_1 \ar[r] & z \ar@(ur,ul)
+  }
+
+\vspace{\baselineskip}
+
+  Clearly that $\mathcal{F}$ is an $\mathcal{L}$-frame. We define the valuation $\vartheta$ as follows:
+
+  \begin{center}
+    $\vartheta(p) = \{ x \}$
+
+    $\vartheta(q) = \{ y, y_1 \}$
+
+    $\vartheta(r) = \{ z \}$
+  \end{center}
+  Let us put $\Gamma = \operatorname{Sub} \{ p, q, \Diamond r \}$.
+We factorise $W$ through $\sim_{\Gamma}$ and consider a model $\widehat{\mathcal{M}} = \langle W / \sim_{\Gamma}, \widehat{R}, \widehat{\vartheta} \rangle$, where $\widehat{R} = (\widehat{R}_{min})^{\davidsstar}$.
+  We have $(\hat{x}, \hat{z}) \in \widehat{R} \circ \widehat{R} \circ \widehat{R}$, but $\Diamond r$ is not true at $x$.
+\end{proof}
 
@@
 
-The converse implication does not have to true generally.
+The converse implication does not have to true generally. ${\bf GL}$ might be an example of a logic that has the ``finite canonical'' model property with no filtrations.
 
-Recall that ${\bf GL} = {\bf K} \oplus \Diamond (p \land \neg \Diamond p) \to \Diamond p$.
-${\bf GL}$ does not admit filtrations, but one may prove that ${\bf GL}$ has the FMP using the selective filtation technique.
+\section{Fusion stuff}
 
 \begin{defin}
-  Let $\mathcal{M} = \langle W, R, \vartheta \rangle$ be a model and $\Gamma$ a set closed under subformulas, then a weak submodel $\mathcal{M} = \langle W', R', \eta \rangle$ (that is, $W' \subseteq W$, $R' \subseteq R$, $\eta = \vartheta \upharpoonright_{W'}$) is called a selective filtration of $\mathcal{M}$ through $\Gamma$ is the following holds, for $\Diamond \psi \in \Gamma$, $x \in W'$:
-  \begin{center}
-    $\mathcal{M}, x \models \Diamond \psi$ iff $\mathcal{M}, y \models \psi$ for some $y \in R'(x)$
-  \end{center}
+  Let $\mathcal{L}_1$ and $\mathcal{L}_2$ be modal logics, then the fusion $\mathcal{L}_1 * \mathcal{L}_2$ is the minimal bimodal logic that contains $\mathcal{L}_1$ and $\mathcal{L}_2$ \cite{kurucz200715}.
 \end{defin}
@@
 \begin{theorem}
-  ${\bf GL}$ admits the ``finite canonical'' model property.
+  $ $
+
+  \begin{enumerate}
+  \item ${\bf K}5 * {\bf K}5$ admits definable filtrations.
+  \item ${\bf K}5 * \dots * {\bf K}5$ admits definable filtrations.
+  \item If $\mathcal{L}_1, \dots, \mathcal{L}_n$ admit strict filtrations, then ${\bf K}5 * \dots * \dots * {\bf K}5 * \mathcal{L}_1 * \dots * \mathcal{L}_n$
+  \end{enumerate}
 \end{theorem}
@@
 \begin{proof}
-  Let $\Gamma$ be a finite $\operatorname{\bf Sub}$-closed set.
-So, $\Gamma = \operatorname{\bf Sub}(\gamma_1) \cup \dots \operatorname{\bf Sub}(\gamma_n)$, for some $\gamma_1, \dots, \gamma_n \in \Gamma$ and every $\gamma_i$ is ${\bf GL}$-consistent.
+$ $
 
-  $\gamma$ is true in $\mathcal{M}_{\Gamma} = \langle V, R^{\bf GL}, \vartheta \rangle$, where
+  \begin{enumerate}
+  \item Let $\Gamma$ be a finite $\operatorname{Sub}$-closed set of bimodal formulas,
+  $\mathcal{F} = \langle W, R_1, R_2 \rangle$ a ${\bf K}5 * {\bf K}5$-frame, and $\vartheta$ a valuation on $\mathcal{F}$. Denote $\langle \mathcal{F}, \vartheta \rangle$ as $\mathcal{M}$.
+
+  We introduce the set of fresh variables $V = \{ p_{\psi} \: | \: \psi \in \Gamma \}$ and define a new model $\mathcal{M}' = \langle \mathcal{F}, \vartheta' \rangle$ as follows:
   \begin{center}
-  $V_{\Gamma} = \bigcup \limits_{\psi \in \Gamma} \{ x \in W_{\bf GL} \: | \: \mathcal{M}_{\bf GL}, x \models \psi \land \neg \Diamond \psi \}$
+    For all $\psi \in \Gamma$, $\mathcal{M}, x \models \psi \Leftrightarrow \mathcal{M}', x \models \psi \Leftrightarrow \mathcal{M}', x \models p_{\psi}$.
   \end{center}
-  since $\mathcal{M}_{\Gamma}$ is a selective filtration of $\mathcal{M}_{\bf M}$ through $\Gamma$.
 
-  Let $x \in V$, consider $||x|| = \{ \psi \in \Gamma \: | \: \mathcal{M}_{\bf GL}, x \models \psi \}$
-  and define $x R^{\bf GL} y$ implies $||x|| R_{\Gamma} ||y||$. Every $||x||$ is clearly finite.
-  The valuation map is usual, $\vartheta_{\Gamma}(p) = \{ ||x|| \: | \: p \in ||x|| \}$ for $p \in \operatorname{PV} \cap \Gamma$. We put $W_{\Gamma}$
+  Consider these modifications of $\Gamma$ and $V$:
 
-  Let us check that $\langle W_{\Gamma}, R_{\Gamma}, \vartheta_{\Gamma} \rangle$ is a ``finite canonical`` ${\bf GL}$-model.
-  Let us check that $R_{\Gamma}$ is transitive and Noetherian and it has no relflexive points.
-  If $W_{\Gamma}$ had a relfexive point, say $||y||$, then we would have $\mathcal{M}, y \models \psi \to \Diamond \psi$, but that is not impossible. Transitivity is obvious.
+  \begin{center}
+    $\Gamma' = \Gamma \cup \{ \Diamond_1 \Box_1 \psi \: | \: \Box_1 \psi \in \Gamma \} \cup \{ \Diamond_2 \Box_2 \psi \: | \: \Box_2 \psi \in \Gamma \}$
 
-  Clearly that $x R^{\bf GL} y$ implies $||x|| R_{\Gamma} ||y||$ and vice versa.
-So, if we had an infinite decreasing $R_{\Gamma}$-chain, then we would have an infinite decreasing $R^{\bf GL}$-chain, but that is not possible.
-  So $\langle W_{\Gamma}, R_{\Gamma}, \vartheta_{\Gamma} \rangle$ is a ``finite canonical`` ${\bf GL}$-model.
+    $\Delta_1 = V \cup \operatorname{Sub}(\{ \Diamond \Box p_{\psi} \: | \: \Box_1 \psi \in \Gamma \})$
+
+    $\Delta_2 = V \cup \operatorname{Sub}(\{ \Diamond \Box p_{\psi} \: | \: \Box_2 \psi \in \Gamma \})$
+  \end{center}
+  Let us define equivalence relations $\sim_1$ and $\sim_2$ induced by $\Delta_1$ and $\Delta_2$ respectively.
+
+  Consider $\mathcal{M}_i = \langle W, R_i, \vartheta' \rangle$, a reduct of $\mathcal{M}'$, we have:
+  \begin{enumerate}
+    \item $\mathcal{M}_i, x \models \Box p_{\psi}$ iff $\mathcal{M}, x \models \Box_i \psi$
+    \item $\mathcal{M}_i, x \models \Diamond \Box p_{\psi}$ iff $\mathcal{M}, x \models \Diamond_i \Box_i \psi$
+  \end{enumerate}
+  So $\sim_i = \sim_{\Gamma'}$ by the construction. Let us put $\widehat{W} = W / \sim_{\Gamma'}$.
+
+  \begin{claim}
+    Let $\widehat{R}_i = (R^{min}_{\Delta_i})^{E}$ and
+    $\widehat{\vartheta(p)} = \{ [x]_{\sim_i} \: | \: \mathcal{M}_i, x \models p \}$ for $p \in \Delta_1$, define $\widehat{\mathcal{M}}_i = \widehat{W}, \widehat{R}_i, \widehat{\vartheta}$.
+Then $\widehat{\mathcal{M}}_i \models {\bf K}5$ and $\widehat{\mathcal{M}}_i$ is a filtration of $\mathcal{M}_i$ through $\Delta_i$.
+  \end{claim}
+  \begin{proof}
+    Let us show that $\widehat{R}_i \subseteq R^{max}_{\Delta_i}$ by induction.
+    Let $\hat{x} \widehat{R}_i \hat{y}$ such that $\mathcal{M}_i, x \models \Box p_{\psi}$.
+    We need $\mathcal{M}_i, y \models p_{\psi}$.
+    Recall that $\widehat{R}_i$ is defined inductively as:
+    \begin{enumerate}
+      \item $R_{\Delta_i}^{0} = R_{\Delta_i}^{min}$
+      \item $R_{\Delta_i}^{n + 1} = R_{\Delta_i}^{n} \cup ({R_{\Delta_i}^{n}}^{-1} \circ R_{\Delta_i}^{n})$
+      \item $(R^{min}_{\Delta_i})^{E} = \bigcup_{k < \omega} R_{\Delta_i}^{k}$
+    \end{enumerate}
+
+    \begin{enumerate}
+      \item $n = 0$. This is obvious.
+      \item $m = n + 1$. Suppose the statement holds for $R_{\Delta_i}^{n}$, we also have $(\hat{x}, \hat{y}) \in R_{\Delta_i}^{n} \cup ({R_{\Delta_i}^{n}}^{-1} \circ R_{\Delta_i}^{n})$.
+      If $(\hat{x}, \hat{y}) \in R_{\Delta_i}^{n}$, then we just apply IH.
+
+      $(\hat{x}, \hat{y}) \in {R_{\Delta_i}^{n}}^{-1} \circ R_{\Delta_i}^{n} \in R_{\Delta_i}^{n}$, then there exists $\hat{z}$ such that $(\hat{z}, \hat{x}) \in R_{\Delta_i}^{n}$ and $(\hat{z}, \hat{y}) \in R_{\Delta_i}^{n}$.
+
+      We rewrite $(\hat{z}, \hat{x}) \in R_{\Delta_i}^{n}$ as the following sausage (for some $\widehat{u_1},  \widehat{u_2}, \dots, \widehat{u_{n - 1}}, \widehat{u_{n}}$):
+
+      \vspace{\baselineskip}
+
+\xymatrix{
+&&& \hat{z} & \ar[l]_{R_{\Delta_i}^{min}} \ar[r]^{R'} \widehat{u_1} & \ar[r]^{R'} \widehat{u_2} \ar[r]^{R'} & \dots & \ar[r]^{R'} \widehat{u_{n - 1}} \ar[r]^{R'} & \widehat{u_{n}} \ar[r]^{R_{\Delta_i}^{min}} & \hat{x}
+}
+\vspace{\baselineskip}
+
+where $R'$ is either $R_{\Delta_i}^{min}$ or its converse.
+We clearly have $\mathcal{M}, z \models \Box p_{\psi}$, since $\mathcal{M}, x \models \Box p_{\psi}$ implies $\mathcal{M}, u_{n} \models \Diamond \Box p_{\psi}$.
+After that we apply the following property of ${\bf K}5$-models:
+\begin{center}
+  For each $a, b \in M_i$ such that $a R_i b$ we have $\mathcal{M}_i, a \models \Diamond \Box p_{\psi}$ iff $\mathcal{M}_i, b \models \Diamond \Box p_{\psi}$
+\end{center}
+Then $\mathcal{M}_i, y \models p_{\psi}$ by IH.
+    \end{enumerate}
   \end{proof}
 
-So we have an example of a logic that has the ``finite canonical'' model property with no filtrations.
+  Finally, we consider a model $\widehat{\mathcal{M}} = \langle \widehat{W}, \widehat{R}_1, \widehat{R}_2, \vartheta \rangle$, where $\widehat{R_{\Gamma'}}_i = {{R_i}^{min}_{\Gamma'}}^E$ and $\vartheta(p)$ is defined as usual for $p \in \Gamma$.
+$\widehat{\mathcal{M}}$ is a filtration of $\mathcal{M}$ through $\Gamma'$.
+
+  Let $\hat{x} \widehat{R_{\Gamma'}}_i \hat{y}$ and $\mathcal{M}, x \models \Box_i \psi$ for $\Box_i \psi \in \Gamma$.
+Then $\mathcal{M}_i, x \models \Box p_{\psi}$, so $\widehat{\mathcal{M}_i}, \hat{x} \models \Box p_{\psi}$.
+  By the claim above, $\widehat{\mathcal{M}_i}$ is a filtration of $\mathcal{M}_i$ through $\Delta_i$, so $\mathcal{M}_i, y \models p_{\psi}$. Then $\mathcal{M}, y \models \psi$.
+
+  \item Likewise
+  \item
+  \end{enumerate}
+\end{proof}
+
+\begin{theorem}
+  Let $\mathcal{L}_1$ and $\mathcal{L}_2$ be modal logics such that $\operatorname{Fr}(\mathcal{L}_1)$ and $\operatorname{Fr}(\mathcal{L}_2)$ admit filtrations, so does $\operatorname{Fr}(\mathcal{L}_1 * \mathcal{L}_2$).
+\end{theorem}
+
+\begin{proof}
+
+\end{proof}
 

### 2021-05-30 frp-paper/Paper.tex
filtrations for K5 * K5

@@
 
-\subsection{Axiomatisating ${\bf R}(+, ;)$}
+\subsection{Axiomatising ${\bf R}(+, ;)$}
 
@@
 
-      $\forall$ picks nodes $x, y \in V_n$ and $d, e \in \mathcal{A}$ such that $c \leq d ; e$ for $c \in l_n(x, y)$.
+      $\forall$ picks an edge $(x, y) \in E_n$ and $d, e \in \mathcal{A}$ such that $c \leq d ; e$ for $c \in l_n(x, y)$.
 $\exists$ has to find a witness.
 She has to find a $z$ which is either a fresh node or an old one.
-If $z$ is fresh, then she defines the prenetwork $T$, the edges of which are $x, y, z$ with the labelling:
+If $z$ is fresh, then she defines the prenetwork $T$, the edges of which are $x, y, z$ with labelling:
       \begin{enumerate}
@@
 
-      $\forall$ picks nodes $x, y \in V_n$ and $c + d$ for $c, d \in \mathcal{A}$. $\exists$ has the following two alternatives for her response:
+      $\forall$ picks an edge $(x, y) \in E_n$ and $c + d$ for $c, d \in \mathcal{A}$. $\exists$ has the following two alternatives for her response:
       \begin{enumerate}
@@
 
-  $\forall$ wins the play if there exists $b \leq a$ such that $b \in \mathcal{A}$ such that $b \in l_{\mathcal{N}_i}(x,y)$ for some $i < n$. Otherwise, $\exists$ wins the play.
+  $\forall$ wins the play if $b \notin l_{\mathcal{N}_i}(x,y)$ for some $i < n$. Otherwise, $\exists$ wins the play.
 
@@
     \item Let $h : \mathcal{A} \to 2^{D \times D}$ is a representation of the base set $D \neq \emptyset$.
-$\exists$ maintains a map ${}^{'} : \operatorname{Nodes}(\mathcal{N}) \to D$, where $\mathcal{N}$ is a network being played, such that $(x',y') \in h(a)$ iff $\up a \subseteq l_{\mathcal{N}}(x,y)$.
+$\exists$ maintains a map ${}^{'} : \operatorname{Nodes}(\mathcal{N}) \to D$, where $\mathcal{N}$ is a network being played, such that $a \in l_{\mathcal{N}}(x,y)$ implies $(x',y') \in h(a)$.
 
@@
 \begin{definition}
-  A \emph{term network} is a finite network $\langle V, E, l \rangle$, where $\langle V, E \rangle$ is a directed graph and $l : E \to 2^{Term}$ is a labelling function such that every $l_(x,y)$ is finite for each $(x, y) \in E$.
+  A \emph{term network} is a finite network $\langle V, E, l \rangle$, where $\langle V, E \rangle$ is a directed graph and $l : E \to 2^{Term}$ is a labelling function such that every $l(x,y)$ is finite for each $(x, y) \in E$.
 \end{definition}
@@
 Let $\mathcal{N} = \langle V, E, l \rangle$ a term network, $\mathcal{A}$ be a join-semilattice ordered semigroup and $\vartheta : \operatorname{Var} \to {\mathcal{A}}$ a valuation.
-Let us define the prenetwork $\mathcal{N}^{\vartheta}$ with the same edges and vertices with labelling $l^{\vartheta}(x, y) = \bigcup \{ F \in \operatorname{Up}(\mathcal{A}, \leq) \: | \: l_{\mathcal{N}}(x, y) \subseteq F \}$.
+Let us define the prenetwork $\mathcal{N}^{\vartheta}$ with the same edges and vertices with labelling $l^{\vartheta}(x, y) = \up \vartheta[l_{\mathcal{N}}(x, y)]$.
 
@@
 
-$\mathcal{N} = \langle V, E, l \rangle$ be a $\mathcal{A}$-network and $\mathcal{A}$ is finite. Define the following equivalence relation on $E \times E$ for $(x_1, y_1), (x_2, y_2) \in E$:
-\begin{center}
-  $(x_1, y_1) \equiv (x_2, y_2)$ iff $l_{\mathcal{N}}(x_1, y_1) = l_{\mathcal{N}}(x_2, y_2)$
-\end{center}
-If $A$ is finite, then $E /_{\equiv}$ is finite as well since the set $\operatorname{Up}(\mathcal{A}, \leq)$ is finite.
+\begin{definition}
+  Let $\mathcal{A}$ be a join semilattice-ordered semigroup and let $\mathcal{N}_1$ and $\mathcal{N}_2$ be $\mathcal{A}$-networks.
+A strong homomorphism from $\mathcal{N}_1$ onto $\mathcal{N}_2$ is a map $f : \mathcal{N}_1 \twoheadrightarrow \mathcal{N}_2$ such that:
+  \begin{enumerate}
+    \item $(x, y) \in E_2$ implies $(f(x), f(y)) \in E_1$,
+    \item If $(x', y') \in E_2$, then there exists $x, y \in E_1$ such that $f(x) = x'$ and $f(y) = y'$ and $(x, y) \in E_1$,
+    \item $l_{\mathcal{N}_1}(x, y) = l_{\mathcal{N}_2}(f(x),f(y))$
+  \end{enumerate}
+  By $\mathcal{N}_1 \twoheadrightarrow \mathcal{N}_2$ we mean that there exists a strong homomorphism of $\mathcal{N}_1$ onto $\mathcal{N}_2$.
+\end{definition}
 
-Let us construct the following $\mathcal{A}$-prenetwork $\mathcal{N}_{\equiv} = \langle V_{\equiv}, E /_{\equiv}, l \rangle$ such that:
-\begin{enumerate}
-  \item $V_{\equiv} = \bigcup \{ \{ x, y \} \: | \: [x,y]_{\equiv} \in E /_{\equiv} \}$, that is, the union of two element sets $\{ x, y \}$ such that there exists an edge $(x,y)$ up to $\equiv$.
-  \item $l([x,y]_{\equiv}) = l_{\mathcal{N}}(x,y)$
-\end{enumerate}
+\begin{lemma}
+  Let $\mathcal{A}$ be a representable join semilattice-ordered semigroup and $\mathcal{N} = \langle V, E, l \rangle$ and $\mathcal{N}' = \langle V', E', l' \rangle$ $\mathcal{A}$-networks.
 
-\begin{lemma} Let $\mathcal{A}$ be a finite join semilattice-ordered semigroup and $\mathcal{N} = \langle V, E, l \rangle$ an $\mathcal{A}$-prenetwork.
-  \begin{enumerate}
-  \item Let $a \in \mathcal{A}$, then $a \in l_{\mathcal{N}}(x,y)$ iff $a \in l([x,y]_{\equiv})$.
-
-  \item Let $\mathcal{A}$ be a finite join semilattice-ordered semigroup and $\mathcal{N}$ an $\mathcal{A}$-prenetwork.
-If $\mathcal{A}$ is a network, so is $\mathcal{N}_{\equiv}$.
-Moreover, $\mathcal{N}_{\equiv}$ is finite.
-\end{enumerate}
+  Suppose $\mathcal{N} \twoheadrightarrow \mathcal{N}'$. If $\mathcal{N}^{*}$ is a representation of $\mathcal{A}$, so is $ $
 \end{lemma}
 
-\begin{proof}
-$ $
-  \begin{enumerate}
-  \item If $a \in l_{\mathcal{N}}(x,y)$, suppose $(x_1,y_1) \equiv (x,y)$, so $l_{\mathcal{N}}(x,y) = l_{\mathcal{N}}(x_1,y_1)$, then $a \in l_{\mathcal{N}}(x_1,y_1)$.
-Therefore, $a \in l([x,y]_{\equiv})$.
-The converse implication is obvious.
-  \item Follows from the definition of a network and the fact that the proof does not depend on the choice of equivalence classes representatives.
-  Let us check join-primeness as an example. The rest of the cases are considered in a similar fashion:
-  Let $a, b \in \mathcal{A}$ and $x, y \in V$ such that $a + b \in l([x,y]_{\equiv})$.
-  Then $a + b \in l(x,y)$ in the original network.
-Then we have either $a \in l(x,y)$ or $b \in l(x,y)$ by the join-primeness condition.
-Therefore, either $a \in l([x,y]_{\equiv})$ or $b \in l([x,y]_{\equiv})$
-
-  $\mathcal{N}_{\equiv}$ is finite since $E_{\equiv}$ is finite. That makes $V_{\equiv}$ finite as well.
-  \end{enumerate}
-\end{proof}
-
-\begin{theorem}\label{frpjoin}
-  ${\bf R}(+, ;)$ has the finite representation property.
-\end{theorem}
-
-\begin{proof}
-  Let $\mathcal{A} \in {\bf R}(+, ;)$ and $\mathcal{A}$ is finite.
-  Let us show that $\mathcal{A}$ has a representation over a finite base.
-  We already know that $\exists$ has a winning strategy on $\mathcal{G}^{\omega}(\mathcal{A})$, so we have $\mathcal{N}$, the resulting network.
-So $\mathcal{N}_{\equiv}$ is a finite network by the previous lemma.
-  Define the following map $f : \mathcal{A} \to 2^{E_{\equiv}}$ as follows:
-\begin{center}
-  $f : a \mapsto \{ (x,y) \: | \: \exists b \leq a \: \& \: b \in l([x,y]_{\equiv}) \}$.
-\end{center}
-$f$ preserves all operations and it is injective for the same reason as $\operatorname{rep}$ constructed above. Let us check that.
-
-If $(x,y) \in f(a + b)$, then there is $c \leq a + b$ such that $c \in l([x,y]_{\equiv})$, then $c \in l_{\mathcal{N}}(x,y)$, so we have either $a \in l_{\mathcal{N}}(x,y)$ or $b \in l_{\mathcal{N}}(x,y)$.
-Thus, either $a \in l([x,y]_{\equiv})$ or $b \in l([x,y]_{\equiv})$.
-If $(x, y) \in f(a)$, then there exists $c \leq a$ such that $c \in l([x,y]_{\equiv})$, but $a + b \in \in l([x,y]_{\equiv})$. Therefore, $f(a + b) = f(a) \cup f(b)$.
-
-If $(x,y) \in f(a ; b)$, then there exists $c \leq a ; b$ with $c \in l([x,y]_{\equiv})$, then $c \in l_{\mathcal{N}}(x,y)$.
-By the saturation condition, there is $z \in \operatorname{Nodes}(\mathcal{N})$ with $a \in l_{\mathcal{N}}(x,z)$ and $b \in l_{\mathcal{N}}(z,y)$.
-Then $a \in l([x,z]_{\equiv})$ and $b \in l([z,y]_{\equiv})$.
-If $(x,y) \in f(a);f(b)$, then there is $z \in V_{\equiv}$ with $(x, z) \in f(a)$ and $(z, y) \in f(b)$, then there are $c$ and $d$ with $c \leq a$ and $c \in l([x,z]_{\equiv})$ and $d \leq b \in l([z,y]_{\equiv})$ respectively.
-Then $c ; d \in l([x,y]_{\equiv})$ by the saturation condition.
-
-Suppose $a \neq b$, so we have either $a \not\leq b$ or $b \not\leq a$.
-Suppose $a \not\leq b$. Let $(x, y) \in f(a)$, then there is $c \leq a$ with $c \in l([x,y]_{\equiv})$, then $c \in l_{\mathcal{N}}(x,y))$ and, moreover, $c \in l_{\mathcal{N}^{*}(a,b)}(x,y))$.
-We know that $\exists$ has a winning strategy in $\mathcal{G}_{\omega}(\mathcal{A})$, so
-$b \notin l_{\mathcal{N}^{*}(a,b)}(x,y))$. So $b \notin l([x,y]_{\equiv})$ and $f(a) \neq f(b)$.
-The second alternative is similar to the previous one.
-
-The base set of a representation, $V /_{\equiv}$, is finite, see the previous lemma.
-\end{proof}
-
-In turn, the finite representation property and the presence of a recursively enumerable axiomatisation imply the following corollary.
-
-\begin{corollary}
-  The problem of membership to ${\bf R}(;, +)$ is decidable for finite structures.
-\end{corollary}
-
-\begin{proof}
-  Follows from Proposition~\ref{dec}, Theorem~\ref{axiomatisation}, and Theorem~\ref{frpjoin}.
-\end{proof}
+\begin{lemma}
+\end{lemma}
 

### 2021-06-04 filtration-stuff/NewText.tex
proof for K5 * K5

@@
 
+\begin{lemma}\label{extension}
+  Let $\Gamma$ be a finite and $\operatorname{Sub}$-closed set of formulas and $\mathcal{M} = \langle W, R, \vartheta \rangle$.
+Consider $\Gamma' = \Gamma \cup \{ \Diamond \Box \psi \: | \: \Box \psi \in \Gamma \}$.
+Let $\Delta$ be any finite and $\operatorname{Sub}$-closed extension of $\Gamma'$.
+Then a model $\widehat{M} = \langle W / \sim_{\Delta}, (R^{min}_{\Delta})^{E}, \widehat{\vartheta} \rangle$ is a filtration of $\mathcal{M}$ through $\Delta$.
+\end{lemma}
+
+\begin{proof}
+  Recall that $(R^{min}_{\Delta})^{E}$ is defined inductively as:
+  \begin{enumerate}
+    \item $R_{\Delta}^{0} = R_{\Delta}^{min}$
+    \item $R_{\Delta}^{n + 1} = R_{\Delta}^{n} \cup ({R_{\Delta}^{n}}^{-1} \circ R_{\Delta}^{n})$
+    \item $(R^{min}_{\Delta})^{E} = \bigcup_{k < \omega} R_{\Delta}^{k}$
+  \end{enumerate}
+
+  If $R^{min}_{\Delta}$ is already a subrelation of $R^{max}$, so the base case is self-evident.
+
+  Suppose the statement holds for $R_{\Delta}^{n}$, $(\hat{x},\hat{y}) \in R_{\Delta}^{n + 1}$, that is, there exists $\hat{z}$ such that $(\hat{z}, \hat{x}), (\hat{z}, \hat{y}) \in R_{\Delta}^{n}$.
+  Let $\mathcal{M}, x \models \Box \psi$ (for $\Box \psi \in \Gamma$).
+
+  We rewrite $(\hat{z}, \hat{x}) \in R_{\Delta}^{n}$ as the following sausage (for some $\widehat{u_1},  \widehat{u_2}, \dots, \widehat{u_{n - 1}}, \widehat{u_{n}}$):
+
+  \xymatrix{
+  &&& \hat{z} & \ar[l]_{R_{\Delta}^{min}} \ar[r]^{R'} \widehat{u_1} & \ar[r]^{R'} \widehat{u_2} \ar[r]^{R'} & \dots \ar[r]^{R'} & \widehat{u_{n - 1}} \ar[r]^{R'} & \widehat{u_{n}} \ar[r]^{R_{\Delta}^{min}} & \hat{x}
+  }
+  where $R'$ is either $R_{\Delta}^{min}$ or its converse.
+We have $\mathcal{M}, z \models \Box \psi$, since $\mathcal{M}, x \models \Box \psi$ implies $\mathcal{M}, u_{n} \models \Diamond \Box \psi$.
+After that we apply the following property of ${\bf K}5$-models:
+  \begin{center}
+    For each $a, b \in M_i$ such that $a R_i b$ we have $\mathcal{M}_i, a \models \Diamond \Box p_{\psi}$ iff $\mathcal{M}_i, b \models \Diamond \Box p_{\psi}$
+  \end{center}
+  Note that we always stay within $\Delta$ since $\Gamma' \subseteq \Delta$.
+  Finally, we conclude $\mathcal{M}, x \models \psi$ from IH and $\mathcal{M}, z \models \Box \psi$.
+\end{proof}
+
 \begin{theorem}
@@
 
-    $\Delta_1 = V \cup \operatorname{Sub}(\{ \Diamond \Box p_{\psi} \: | \: \Box_1 \psi \in \Gamma \})$
-
-    $\Delta_2 = V \cup \operatorname{Sub}(\{ \Diamond \Box p_{\psi} \: | \: \Box_2 \psi \in \Gamma \})$
+    $\Delta = V \cup \operatorname{Sub}(\{ \Diamond \Box p_{\psi} \: | \: \Box_i \psi \in \Gamma, i n= 1,2 \})$
   \end{center}
-  Let us define equivalence relations $\sim_1$ and $\sim_2$ induced by $\Delta_1$ and $\Delta_2$ respectively.
+  Let us define an equivalence relation $\sim_{\Delta}$ induced by $\Delta$.
 
@@
   \end{enumerate}
-  So $\sim_i = \sim_{\Gamma'}$ by the construction. Let us put $\widehat{W} = W / \sim_{\Gamma'}$.
+  So $\sim = \sim_{\Gamma'}$ by the construction. Let us put $\widehat{W} = W / \sim_{\Gamma'}$. Lemma~\ref{extension} implies the following claim:
 
   \begin{claim}
-    Let $\widehat{R}_i = (R^{min}_{\Delta_i})^{E}$ and
+    Let $\widehat{R}_i = (R^{min}_{\Delta})^{E}$ and
     $\widehat{\vartheta(p)} = \{ [x]_{\sim_i} \: | \: \mathcal{M}_i, x \models p \}$ for $p \in \Delta_1$, define $\widehat{\mathcal{M}}_i = \widehat{W}, \widehat{R}_i, \widehat{\vartheta}$.
-Then $\widehat{\mathcal{M}}_i \models {\bf K}5$ and $\widehat{\mathcal{M}}_i$ is a filtration of $\mathcal{M}_i$ through $\Delta_i$.
+Then $\widehat{\mathcal{M}}_i \models {\bf K}5$ and $\widehat{\mathcal{M}}_i$ is a filtration of $\mathcal{M}_i$ through $\Delta$.
   \end{claim}
-  \begin{proof}
-    Let us show that $\widehat{R}_i \subseteq R^{max}_{\Delta_i}$ by induction.
-    Let $\hat{x} \widehat{R}_i \hat{y}$ such that $\mathcal{M}_i, x \models \Box p_{\psi}$.
-    We need $\mathcal{M}_i, y \models p_{\psi}$.
-    Recall that $\widehat{R}_i$ is defined inductively as:
-    \begin{enumerate}
-      \item $R_{\Delta_i}^{0} = R_{\Delta_i}^{min}$
-      \item $R_{\Delta_i}^{n + 1} = R_{\Delta_i}^{n} \cup ({R_{\Delta_i}^{n}}^{-1} \circ R_{\Delta_i}^{n})$
-      \item $(R^{min}_{\Delta_i})^{E} = \bigcup_{k < \omega} R_{\Delta_i}^{k}$
-    \end{enumerate}
-
-    \begin{enumerate}
-      \item $n = 0$. This is obvious.
-      \item $m = n + 1$. Suppose the statement holds for $R_{\Delta_i}^{n}$, we also have $(\hat{x}, \hat{y}) \in R_{\Delta_i}^{n} \cup ({R_{\Delta_i}^{n}}^{-1} \circ R_{\Delta_i}^{n})$.
-      If $(\hat{x}, \hat{y}) \in R_{\Delta_i}^{n}$, then we just apply IH.
-
-      $(\hat{x}, \hat{y}) \in {R_{\Delta_i}^{n}}^{-1} \circ R_{\Delta_i}^{n} \in R_{\Delta_i}^{n}$, then there exists $\hat{z}$ such that $(\hat{z}, \hat{x}) \in R_{\Delta_i}^{n}$ and $(\hat{z}, \hat{y}) \in R_{\Delta_i}^{n}$.
-
-      We rewrite $(\hat{z}, \hat{x}) \in R_{\Delta_i}^{n}$ as the following sausage (for some $\widehat{u_1},  \widehat{u_2}, \dots, \widehat{u_{n - 1}}, \widehat{u_{n}}$):
-
-      \vspace{\baselineskip}
-
-\xymatrix{
-&&& \hat{z} & \ar[l]_{R_{\Delta_i}^{min}} \ar[r]^{R'} \widehat{u_1} & \ar[r]^{R'} \widehat{u_2} \ar[r]^{R'} & \dots & \ar[r]^{R'} \widehat{u_{n - 1}} \ar[r]^{R'} & \widehat{u_{n}} \ar[r]^{R_{\Delta_i}^{min}} & \hat{x}
-}
-\vspace{\baselineskip}
-
-where $R'$ is either $R_{\Delta_i}^{min}$ or its converse.
-We clearly have $\mathcal{M}, z \models \Box p_{\psi}$, since $\mathcal{M}, x \models \Box p_{\psi}$ implies $\mathcal{M}, u_{n} \models \Diamond \Box p_{\psi}$.
-After that we apply the following property of ${\bf K}5$-models:
-\begin{center}
-  For each $a, b \in M_i$ such that $a R_i b$ we have $\mathcal{M}_i, a \models \Diamond \Box p_{\psi}$ iff $\mathcal{M}_i, b \models \Diamond \Box p_{\psi}$
-\end{center}
-Then $\mathcal{M}_i, y \models p_{\psi}$ by IH.
-    \end{enumerate}
-  \end{proof}
 
@@
 Then $\mathcal{M}_i, x \models \Box p_{\psi}$, so $\widehat{\mathcal{M}_i}, \hat{x} \models \Box p_{\psi}$.
-  By the claim above, $\widehat{\mathcal{M}_i}$ is a filtration of $\mathcal{M}_i$ through $\Delta_i$, so $\mathcal{M}_i, y \models p_{\psi}$. Then $\mathcal{M}, y \models \psi$.
+  By the claim above, $\widehat{\mathcal{M}_i}$ is a filtration of $\mathcal{M}_i$ through $\Delta$, so $\mathcal{M}_i, y \models p_{\psi}$. Then $\mathcal{M}, y \models \psi$.
 
   \item Likewise
-  \item
   \end{enumerate}

### 2021-06-15 filtration-stuff/NewText.tex
more facts about definable filtrations for fusions

@@
 \date{}
-\title{Notes on filtration of logics containing {\bf K5}}
+\title{Notes on filtrations for logics that contain {\bf K5}}
 \begin{document}
@@
 \maketitle
+
+\section{Preliminaries}
 
@@
 \end{lemma}
+
+\section{Filtrations for {\bf K5}}
 
@@
 \begin{lemma}\label{extension}
-  Let $\Gamma$ be a finite and $\operatorname{Sub}$-closed set of formulas and $\mathcal{M} = \langle W, R, \vartheta \rangle$.
+  $ $
+
+  \begin{itemize}
+  \item Let $\Gamma$ be a finite and $\operatorname{Sub}$-closed set of formulas and $\mathcal{M} = \langle W, R, \vartheta \rangle$.
 Consider $\Gamma' = \Gamma \cup \{ \Diamond \Box \psi \: | \: \Box \psi \in \Gamma \}$.
@@
 Then a model $\widehat{M} = \langle W / \sim_{\Delta}, (R^{min}_{\Delta})^{E}, \widehat{\vartheta} \rangle$ is a filtration of $\mathcal{M}$ through $\Delta$.
+  \item The same for ${\bf K} \oplus \Diamond \Diamond \Diamond p \to \Diamond p$
+  \end{itemize}
 \end{lemma}
@@
   Finally, we conclude $\mathcal{M}, x \models \psi$ from IH and $\mathcal{M}, z \models \Box \psi$.
+
+  The second item is shown similarly.
 \end{proof}
@@
   \item ${\bf K}5 * \dots * {\bf K}5$ admits definable filtrations.
+  \item If $\mathcal{L}$ admits strict filtrations, then ${\bf K}5 * \mathcal{L}$ admits definable filtrations
   \item If $\mathcal{L}_1, \dots, \mathcal{L}_n$ admit strict filtrations, then ${\bf K}5 * \dots * \dots * {\bf K}5 * \mathcal{L}_1 * \dots * \mathcal{L}_n$
+  \item Let $\mathcal{L} = {\bf K} \oplus \Diamond \Diamond \Diamond p \to \Diamond p$  (here and below), then $\mathcal{L} * \mathcal{L}$ admits definable filtrations.
+  \item Let $\mathcal{L} = {\bf K} \oplus \Diamond \Diamond \Diamond p \to \Diamond p$ and $\mathcal{L}_1$ a logic that admits strict filtrations, then $\mathcal{L} * \mathcal{L}_1$
   \end{enumerate}
@@
   \item Likewise
+  \item The argument is the same as in the proof of the first item, except for {\bf Claim} 1 that has the following formulation:
+  Let $\widehat{R}_1 = (R^{min}_{\Delta})^{E}$ and $\widehat{R}_2 = (R^{min}_{\Delta})^{\mathcal{L}_1}$
+  Define a valuation as usual as $\widehat{\vartheta(p)} = \{ [x]_{\sim_i} \: | \: \mathcal{M}_i, x \models p \}$ for $p \in \Delta_1$, define $\widehat{\mathcal{M}}_1 = \langle \widehat{W}, \widehat{R}_1, \widehat{\vartheta} \rangle$ and $\widehat{\mathcal{M}}_2 = \langle \widehat{W}, \widehat{R}_2, \widehat{\vartheta} \rangle$.
+
+  Then $\widehat{\mathcal{M}}_1 \models {\bf K}5$ and $\widehat{\mathcal{M}}_1 \models {\bf L}$ and $\widehat{\mathcal{M}}_i$ is a filtration of $\mathcal{M}_i$ through $\Delta$.
+
+  \item Likewise
+  \item The argument is similar to the proof of first item, but filtrations are slightly different.
+  Let $\mathcal{M} = \langle W, R_1, R_2, \vartheta \rangle$ be a $\mathcal{L} * \mathcal{L}$ model and $\Gamma$ a $\operatorname{Sub}$-closed set of formulas.
+As above, we define a set $V$ and a model $\mathcal{M}'$.
+Define extensions of $\Gamma$ and $V$:
+
+  \begin{center}
+    $\Gamma' = \Gamma \cup \{ \Diamond_1 \Diamond_1 \psi \: | \: \Diamond_1 \psi \in \Gamma \} \cup \{ \Diamond_2 \Diamond_2 \psi \: | \: \Diamond_2 \psi \in \Gamma \}$
+
+    $\Delta = V \cup \operatorname{Sub}(\{ \Diamond \Diamond p_{\psi} \: | \: \Diamond \psi \in \Gamma, i = 1,2 \})$
+  \end{center}
+  As above $\sim_{\Delta} = \Gamma'$ and $\widehat{\mathcal{M}'} = \langle W /\sim_{\Delta}, \widehat{R_i}, \widehat{\vartheta} \rangle$ are filtrations of reducts of $\mathcal{M}'$ through $\Delta$.
+Then $\widehat{\mathcal{M}} = \langle W /\sim_{\Delta}, \widehat{R_1}, \widehat{R_2}, \widehat{\vartheta} \rangle$ is a required filtration of the original $\mathcal{M}$.
+  \item Extend $\Gamma$ with $\{ \Diamond_i \Diamond_i \psi \: | \: \Diamond_i \psi \in \Gamma, i = 1,2 \}$ and
+  $V$ with $\{\Diamond \Diamond p_{\psi} \: | \: \Diamond_i {\psi}, i = 1, 2\}$
   \end{enumerate}
@@
 \begin{theorem}
-  Let $\mathcal{L}_1$ and $\mathcal{L}_2$ be modal logics such that $\operatorname{Fr}(\mathcal{L}_1)$ and $\operatorname{Fr}(\mathcal{L}_2)$ admit filtrations, so does $\operatorname{Fr}(\mathcal{L}_1 * \mathcal{L}_2$).
+  Let $\mathcal{L}_1$ and $\mathcal{L}_2$ be modal logics that admit definable filtrations.
+If $\operatorname{Mod}(\mathcal{L}_1)$ and $\operatorname{Mod}(\mathcal{L}_2)$ admit definable filtrations, so does $\operatorname{Mod}(\mathcal{L}_1 * \mathcal{L}_2$).
 \end{theorem}
@@
 \begin{proof}
+  Let $\mathcal{M} = \langle W, R_1, R_2, \vartheta \rangle$ be an $\mathcal{L}_1 * \mathcal{L}_2$.
+  We define a notation $\nabla = \{ \neg \Diamond, \Diamond \neg, \Diamond \}$.
 
+  Both logics admit definable filtrations, so for every finite $\operatorname{Sub}$-closed set $\Gamma$ and for every $\mathfrak{M}$, an $\mathcal{L}_1$-model (or an $\mathcal{L}_2$ one) there exists there exists $\Delta$, a extension of $\Gamma$ having the form:
+  \begin{center}
+    $\Delta_1 = \Gamma \cup \operatorname{Sub}\{ \nabla_1 \nabla_2 \dots \nabla_n \Diamond \psi \: | \: \Diamond \psi \in \Gamma \}$ (for $\mathcal{L}_1$)
+
+    $\Delta_2 = \Gamma \cup \operatorname{Sub}\{ \nabla_1 \nabla_2 \dots \nabla_k \Diamond \psi \: | \: \Diamond \psi \in \Gamma \}$ (for $\mathcal{L}_2$)
+  \end{center}
+  such that $\widehat{\mathfrak{M}} = \langle W / \sim_{\Delta_i}, \widehat{R}, \vartheta \rangle$ is a filtration of $\mathfrak{M}$ through the corresponding $\Delta_i$.
+
+  Let $V$ be a set of fresh variables indexed over $\Gamma$ as in the proof for a fusion of ${\bf K5}$ with something else.
+Let $\mathcal{M}'$ be a model defined as previously.
+We extend $V$ and $\Gamma$ in the following way:
+
+\begin{center}
+  $\Gamma' = \Gamma \cup \operatorname{Sub}\{ {\nabla_1}_1 {\nabla_2}_1 \dots {\nabla_n}_1 \Diamond_1 \psi \: | \: \Diamond_1 \psi \in \Gamma \} \cup \operatorname{Sub}\{ {\nabla_1}_2 {\nabla_2}_2 \dots {\nabla_n}_2 \Diamond_2 \psi \: | \: \Diamond_2 \psi \in \Gamma \}$
+
+  $\Delta = V \cup \operatorname{Sub}\{ \nabla_1 \nabla_2 \dots \nabla_n \Diamond p_{\psi} \: | \: {\nabla_{n + 1}}_1 \psi \in \Gamma' \} \cup \operatorname{Sub}\{ \nabla_1 \nabla_2 \dots \nabla_k \Diamond p_{\psi} \: | \: \Diamond_2 \psi \in \Gamma \}$.
+\end{center}
+By the construction, $\sim_{\Gamma'} = \sim_{\Delta}$. So we have filtrations for the corresponding reducts of $\mathcal{M}'$ through $\Delta$ as well as for the original $\mathcal{M}$.
 \end{proof}

### 2021-06-15 frp-paper/Paper.tex
more facts about definable filtrations for fusions

@@
 
-Let $\mathcal{A}$ be a residuated semigroup and $\mathcal{Q}_{\mathcal{A}}$ is a quantale of Galois closed subsets of $\mathcal{A}$.
+Let $\mathcal{A}$ be a residuated semigroup and $\mathcal{Q}_{\mathcal{A}}$ a quantale of Galois closed subsets of $\mathcal{A}$.
 $\widehat{\mathcal{Q}_{\mathcal{A}}}$ is the corresponding relational quantale.
@@
     \item If $\mathcal{A}$ is representable then $\exists$ has a winning strategy in $\mathcal{G}_{\omega}(\mathcal{A})$.
-    \item If $|\mathcal{A}| = \omega$ and $\exists$ has a winning strategy in $\mathcal{G}_{\omega}(\mathcal{A})$ then $\mathcal{A}$ is representable.
+    \item If $|\mathcal{A}| \leq \omega$ and $\exists$ has a winning strategy in $\mathcal{G}_{\omega}(\mathcal{A})$ then $\mathcal{A}$ is representable.
   \end{enumerate}

### 2021-06-16 filtration-stuff/NewText.tex
K5-i,j stuff (wip)

@@
 
+\section{i,j-Euclideaness}
+
+Let $R$ be a binary relation on $W \neq \emptyset$, then the $i,j$-Euclidean closure of $R$ (where $i, j < \omega$), denoted as $R^{E_{i,j}}$, is a binary relation defined recursively as follows:
+\begin{enumerate}
+  \item $R_0 = R$
+  \item $R_{n + 1} = R_n \cup (((R_n)^{-1})^i \circ {R_n}^j)$
+  \item $R^{E_{i,j}} = \bigcup \limits_{k < \omega} R_k$
+\end{enumerate}
+
+\begin{prop}
+  If $R$ be a binary relation on $W \neq \emptyset$, $R^{E_{i,j}}$ is $i,j$-Euclidean.
+\end{prop}
+\begin{proof}
+  By induction. Let $(x,y_1), (y_1,y_2), (y_2,y_3), \dots, (y_{i - 1},y_i), (x,z_1), (z_1,z_2), (z_2,z_3), \dots, (z_{j - 1},z_j) \in R^{E_{i,j}}$.
+  That is, each of these pairs belong to some $R_k$ and every $k < \omega$ for every pair is its own.
+  If $k = 0$ for each pair, then $(y_i, z_j) \in R_1$.
+  If some those $k$ is non-zero, then we take the greatest amongst them, say $e$ and all given pairs will be elements of $R_{e + 1}$.
+\end{proof}
+
+\begin{theorem}
+  ${\bf K5}^{i,j} = {\bf K} \oplus \Diamond^{i} p \to \Box^{j} \Diamond p$ admits definable filtrations.
+\end{theorem}
+
+\begin{proof}
+  Let $\mathcal{M} = \langle W, R, \vartheta \rangle$ and $\Gamma$ a finite $\operatorname{Sub}$-closed of formulas. We extend $\Gamma$ as
+  \begin{center}
+    $\Delta = \Gamma \cup \operatorname{Sub}\{  \Diamond^{i} \psi \: | \: \Box \psi \in \Gamma \} \cup \operatorname{Sub}\{ \Box^{j} \Diamond \psi \: | \: \Diamond \psi \in \Gamma \}$
+  \end{center}
+  Let $(\hat{x}, \hat{y}) \in R^{E_{i,j}}$, $\mathcal{M}, x \models \Box \psi$ for $\Box \psi$ in $\Delta$.
+  If $n = 0$, then the statement is obvious. Suppose $n = 1$.
+  Then $(\hat{x}, \hat{y}) \in R_{min} \cup (((R_{min})^{-1})^i \circ {R_{min}}^j)$.
+Consider the second alternative.
+Then there exists $\hat{z}$ such that $(\hat{x}, \hat{z}) \in ((R_{min})^{-1})^i $ and $(\hat{z}, \hat{y}) \in {R_{min}}^j$.
+
+  That is, there are $\hat{x_1}, \hat{x_2}, \dots, \hat{x_i}$ and $\hat{y_1}, \hat{y_2}, \dots, \hat{y_j}$ such that
+
+  \xymatrix{
+  &&&& \hat{z} \ar[r]^{R_{min}} & \ar[r]^{R_{min}} \hat{x_1} \ar[r]^{R_{min}} & \hat{x_2} \ar[r]^{R_{min}} & \dots \ar[r]^{R_{min}} & \ar[r]^{R_{min}} \hat{x_i} \ar[r]^{R_{min}} & \hat{x} \\
+  &&&& \hat{z} \ar[r]^{R_{min}} & \ar[r]^{R_{min}} \hat{y_1} \ar[r]^{R_{min}} & \hat{y_2} \ar[r]^{R_{min}} & \dots \ar[r]^{R_{min}} & \ar[r]^{R_{min}} \hat{y_j} \ar[r]^{R_{min}} & \hat{y} \\
+  }
+
+  ???
+\end{proof}
+
+\section{The case of 2-transitivity}
+
 Let us define the logic $\mathcal{L}$ as ${\bf K} \oplus \Diamond \Diamond \Diamond p \to \Diamond p$.

### 2021-06-24 filtration-stuff/NewText.tex
wip

@@
 
+The following definition of a filtration is due to, e.g., \cite{shehtman2004filtration}.
+
 \begin{defin}
@@
   \begin{enumerate}
-    \item $\widehat{W} = W / \sim$, where $\sim$ is an equivalence relation having a finite index that respects $\Gamma$
+    \item $\widehat{W} = W / \sim$, where $\sim$ is an equivalence relation that respects $\Gamma$
     \item $\widehat{\vartheta}(p) = \{ [x]_{\sim} \: | \: x \in W \: \& \: x \in \vartheta(p)\}$
@@
 
-If $\Phi$ is finite subset of $\Gamma$ and $\sim = \sim_{\Phi}$, then $\widehat{\mathcal{M}}$ is a definable $\Gamma$-filtration of $\mathcal{M}$ through $\Phi$.
+Alternatively, one may reformulate the condition of the maximal filtered relations using $\Diamond$'s as follows. We will use this formulation occasionally:
+
+\begin{center}
+  $\hat{x} \widehat{R}_{\Gamma,i}^{\text{max}} \hat{y} \Leftrightarrow \forall \Diamond_i \varphi \in \Gamma \: (\mathcal{M}, y \models \varphi \Rightarrow \mathcal{M}, x \models \Diamond_i \varphi)$
+\end{center}
+
+If $\Phi$ is an extension of $\Gamma$ and $\sim = \sim_{\Phi}$, then $\widehat{\mathcal{M}}$ is a definable $\Gamma$-filtration of $\mathcal{M}$ through $\Phi$.
 If $\sim = \sim_{\Gamma}$, then such a filtration by means of the definiton above is called \emph{strict}.
-A class of models $\mathbb{M}$ admits strict filtrations for models (ASF), if for every $\operatorname{Sub}$-closed set $\Gamma$ and for every $\mathcal{M} \in \mathbb{M}$ there exists a $\Gamma$ filtration of $\mathcal{M}$.
-A class of frames $\mathbb{F}$ admits strict filtrations for frames, if for every $\operatorname{Sub}$-closed set $\Gamma$ and for every frame $\mathcal{F} \in \mathbb{F}$ and every model $\mathcal{M}$ over $\mathcal{F}$ there exists a $\Gamma$ filtration of $\mathcal{M}$.
-If $\mathcal{L}$ is canonical, then the ASF property for frames and ASF property for models are equivalent \cite[Theorem 2.10]{kikot2020completeness}.
+
+A class of models $\mathbb{M}$ admits strict filtrations for models (ASF), if for every $\operatorname{Sub}$-closed set $\Gamma$ and for every $\mathcal{M} \in \mathbb{M}$ there exists a model $\widehat{\mathcal{M}}$ such that $\widehat{\mathcal{M}} \in \mathbb{M}$ and $\widehat{\mathcal{M}}$ is a filtration of $\mathcal{M}$ through $\Gamma$.
+
+A class of frames $\mathbb{F}$ admits strict filtrations for frames, if for every $\operatorname{Sub}$-closed set $\Gamma$ and for every frame $\mathcal{F} \in \mathbb{F}$ and every model $\mathcal{M}$ over $\mathcal{F}$ there exists a $\Gamma$ filtration of $\mathcal{M}$, and the underlying frame of this filtration belongs to $\mathbb{F}$.
+
+If $\mathcal{L}$ is canonical, then the ASF property for frames and ASF property for models are equivalent, see \cite[Theorem 2.10]{kikot2020completeness}.
+
+The key lemma about filtrations is the following, see \cite[Theorem 2.39]{blackburn_rijke_venema_2001}:
 
@@
 
-\begin{defin} Let $\mathbb{F}$ be a class of Kripke frames and $\Gamma$ a finite set of formulas closed under subformulas.
+\begin{defin}
+$ $
+
+  \begin{enumerate}
+    \item Let $\mathbb{F}$ be a class of Kripke frames and $\Gamma$ a finite set of formulas closed under subformulas.
 If for every model $\mathcal{M}$ over $\mathcal{F} \in \mathbb{F}$ there exists a model that is a $\Gamma$-definable filtration of $\mathcal{M}$, then $\mathbb{F}$ admits definable filtration.
-A class of models $\mathbb{M}$ admits definable filtration if for every $\mathcal{M} \in \mathbb{M}$ there exists a model belonging to the same class that is a definable $\Gamma$-filtration of $\mathcal{M}$.
+
+    \item A class of models $\mathbb{M}$ admits definable filtration if for every $\mathcal{M} \in \mathbb{M}$ there exists a model belonging to the same class that is a definable $\Gamma$-filtration of $\mathcal{M}$.
+  \end{enumerate}
 \end{defin}
@@
 
-\begin{defin} A first-order formula is called Horn if it has the following form:
+\begin{defin} A first-order formula is called Horn if it has the following form (see \cite{chang1990model}):
 
   \begin{center}
-    $\forall x_1, \dots, x_n (x_{i_1} R x_{j_1} \land \dots \land x_{i_s} R x_{j_s} \rightarrow x_k R x_l)$
+    $\forall x_1, \dots, x_n (x_{i_1} R x_{j_1} \land \dots \land x_{i_s} R x_{j_s} \rightarrow A)$, where $A$ is either $x_k R x_l$ or $\bot$.
   \end{center}
@@
 
-$E$-closure (an Euclidean Horn closure of a binary relation) has the following equivalent definitions:
+The ${\bf K5}$-closure (an Euclidean Horn closure of a binary relation) has the following equivalent definitions:
 \begin{lemma} \label{equivHorn}
-  Let $\mathcal{F} = \langle W, R \rangle$ be a Kripke frame.
+  Let $\mathcal{F} = \langle W, R \rangle$ be a Kripke frame and $R \subseteq R^{\bf K5}$.
   The following conditions are equivalent:
@@
   \begin{enumerate}
-    \item $R^{E}$ is the smallest Euclidean relation containing $R$.
-    \item $R^{E} = \bigcup \limits_{i < \omega} R_i$, where
+    \item $R^{\bf K5}$ is the smallest Euclidean relation containing $R$, that is, the Horn closure of $R$.
+    \item $R^{\bf K5} = \bigcup \limits_{i < \omega} R_i$, where
     \begin{itemize}
@@
     \end{itemize}
-    \item $x R^E y$ iff there exists $n < \omega$ such that
+    \item $x R^{\bf K5} y$ iff there exists $n < \omega$ such that
     either $x R y$ or $\exists z_1, \dots, z_n$ with $z_1 R x$ and $z_{n - 1} R y$ and for each $1 < i \leq n$ one has
     either $z_{i - 1} R z_i$ or $z_i R z_{i - 1}$.
-    \item $R^{E} = R \cup \bigcup \limits_{i < \omega} (R^{-1} \circ (R \circ R^{-1})^n \circ R)$.
+    \item $R^{\bf K5} = R \cup \bigcup \limits_{i < \omega} (R^{-1} \circ (R \cup R^{-1})^n \circ R)$.
   \end{enumerate}
@@
 
-\begin{lemma}
- Let $\mathcal{F} = \langle W, R \rangle$ be a Kripke frame. Let us define $R^{E} = \bigcup \limits_{i < \omega} R_i$ where:
-
- \begin{enumerate}
-   \item $R_0 = R$
-   \item $R_{n + 1} = R_n \cup ({R_n}^{-1} \circ R_n)$
- \end{enumerate}
- Then $R^{E}$ is Euclidean.
-\end{lemma}
+\begin{theorem}\label{sad}
+  ${\bf K}5$ does not admit strict filtrations.
+\end{theorem}
 
 \begin{proof}
-  Let $(x, y), (x, z) \in R^{E}$, one needs to show that $(y, z) \in R^{E}$.
-  Clearly that $(x, y) \in R_i$ and $(x, z) \in R_j$ for some $i, j < \omega$. Thus, we need $(y, z) \in R_m$ for some $m$ depending on $i$ and $j$.
+  Let us consider a ${\bf K}5$ model whose Euclidean closure of the minimal filtration does not give us a filtration.
 
-  Let us consider the following cases:
-  \begin{enumerate}
-    \item $i = 0$ and $j = 0$
+  Let us consider a frame called $\mathcal{F}_{bad}$. We define this frame with the following graph:
 
-    Suppose $(x, y), (x, z) \in R_0 = R$, then $(y, z) \in R^{-1} \circ R$. Thus, $(y, z) \in R_1$
-    \item $i = 0$ and $j = k + 1$
+\vspace{\baselineskip}
 
-    Suppose $(x, y) \in R$ and $(x, z) \in R_{k + 1} = R_{k} \cup ({R_k}^{-1} \circ R_k)$.
-    Clearly that $(x, y) \in R_{k + 1}$ as well. It is obviously that $(y, z) \in R_{k + 2}$ since
-    $(y, x) \in R_{k + 1}^{-1}$ and $(x, z) \in R_{k + 1}$.
+  \xymatrix{
+  &&&&&&& w_5\ar@(ul,ur) \ar@<-.5ex>[d] \\
+  &&&&&& w_1\ar@(ul,ur)  & w_4 \ar@<.5ex>[u] \ar@(dl,ul)^{} \\
+  &&&&&& w_2 \ar[u] & w_3 \ar[u]
+  }
 
-    \item The case with $i = k + 1$ and $j = 0$ is similar to the previous one.
+Let us define a valuation $\vartheta$ such that $\vartheta(p) = \{ w_5\}$ and $\vartheta(q) = \{ w_1\}$.
+Let us consider a minimal filtration of $\mathcal{M}_{bad}$ through the Sub-closure of $\Gamma = \{ \neg p, \neg \Diamond p\}$.
 
-    \item Suppose $i = m + 1$ and $j = k + 1$. That is, $(x, y) \in R_{m + 1} = R_m \cup ({R_m}^{-1} \circ R_m)$ and $(x, z) \in R_{k + 1} = R_k \cup ({R_k}^{-1} \circ R_k)$.
-    Consider the following four subcases:
-    \begin{enumerate}
-      \item Suppose $(x, y) \in R_m$ and $(x, z) \in R_k$ and $m \leq k$ without loss of generality.
-      $m \leq k$ implies $R_m \subseteq R_k$ and $(x, y) \in R_k$ in particular. Thus, $(y, z) \in {R_k}^{-1} \circ R_k$, so $(y, z) \in R_{k + 1}$.
-      \item The rest of the cases are similar to the first one.
-    \end{enumerate}
-    \end{enumerate}
+Clearly that $w_2 \sim_{\Gamma} w_3$, since $\neg p$ and $\neg \Diamond p$ are true both at $w_2$ and $w_3$.
+
+Moreover, $R_{min} \cup (R_{min}^{-1} \circ R_{min})$ is not a subset of $R_{max}$ since $(\hat{w_1}, \hat{w_5}) \in (R_{min}^{-1} \circ R_{min})$, but $\Diamond p$ is not true at $w_5$.
+
+Let us also note that strict filtrations of this model is not Euclidean.
+Suppose by contrary that $\widehat{R}^{\mathcal{E}}$ is a strict filtraction of that model.
+So $R_{min}^E \subseteq \widehat{R}^{\mathcal{E}}$, since $R_{min}^E$ is the minimal Euclidean relation containing $R_{min}$.
+On the other hand, $R_{min}^E \not\subseteq R_{max}$, so is not $\widehat{R}^{\mathcal{E}}$.
 \end{proof}
+
+\begin{theorem}
+  ${\bf K5}$ admits definable filtrations.
+\end{theorem}
 
@@
   Let $\mathcal{M} = \langle W, R, \vartheta \rangle$ be a transitive Euclidean model and $\overline{\mathcal{M}} = \langle \overline{W}, \overline{R}, \overline{\vartheta} \rangle$ its minimal filtration through $\Gamma$, where $\Gamma$ is finite and Sub-closed.
-Let us put $\widehat{R} = \overline{R}^{+} \cup \overline{R}^{E}$.
-Let us show that $\overline{R}^{+} \cup \overline{R}^{E} \subseteq \overline{R}^{max}$.
+Let us put $\widehat{R} = \overline{R}^{+} \cup \overline{R}^{\bf K5}$.
+Let us show that $\overline{R}^{+} \cup \overline{R}^{{\bf K}45} \subseteq \overline{R}^{max}$.
 
@@
 
-  Let $\hat{x} \widehat{R} \hat{y}$. Let us consider the case when $(\hat{x}, \hat{y}) \in \overline{R}^{E}$
+  Let $\hat{x} \widehat{R} \hat{y}$.
 
@@
     \item Suppose $(\hat{x}, \hat{y}) \in \overline{R}$, then $\mathcal{M}, x \models \Diamond \varphi$ holds trivially by the definition of the minimal filtration.
-    \item Suppose the statement holds $\overline{R}_n$ and $(\hat{x}, \hat{y}) \in \overline{R}_{n + 1} = \overline{R}_n \cup (\overline{R}_n^{-1} \circ \overline{R}_n)$.
+    \item Let us consider the case when $(\hat{x}, \hat{y}) \in \overline{R}^{{\bf K}5}$. The second alternative is the same as for the ${\bf K}4$-case, see \cite[p. 141]{chagrov}.
+
+    Suppose the statement holds $\overline{R}_n$ and $(\hat{x}, \hat{y}) \in \overline{R}_{n + 1} = \overline{R}_n \cup (\overline{R}_n^{-1} \circ \overline{R}_n)$.
 We consider the case of $(\hat{x}, \hat{y}) \in (\overline{R}_n^{-1} \circ \overline{R}_n)$.
@@
 
-\begin{theorem}\label{sad}
-  ${\bf K}5$ does not admit strict filtrations.
-\end{theorem}
-
-\begin{proof}
-  Let us consider a ${\bf K}5$ model whose Euclidean closure of the minimal filtration does not give us a filtration.
-
-  Let us consider a frame called $\mathcal{F}_{bad}$. We define this frame with the following graph:
-
-\vspace{\baselineskip}
-
-  \xymatrix{
-  &&&&&&& w_5\ar@(ul,ur) \ar@<-.5ex>[d] \\
-  &&&&&& w_1\ar@(ul,ur)  & w_4 \ar@<.5ex>[u] \ar@(dl,ul)^{} \\
-  &&&&&& w_2 \ar[u] & w_3 \ar[u]
-  }
-
-Let us define a valuation $\vartheta$ such that $\vartheta(p) = \{ w_5\}$ and $\vartheta(q) = \{ w_1\}$.
-Let us consider a minimal filtration of $\mathcal{M}_{bad}$ through the Sub-closure of $\Gamma = \{ \neg p, \neg \Diamond p\}$.
-
-Clearly that $w_2 \sim_{\Gamma} w_3$, since $\neg p$ and $\neg \Diamond p$ are true both at $w_2$ and $w_3$.
-
-Moreover, $R_{min} \cup (R_{min}^{-1} \circ R_{min})$ is not a subset of $R_{max}$ since $(\hat{w_1}, \hat{w_5}) \in (R_{min}^{-1} \circ R_{min})$, but $\Diamond p$ is not true at $w_5$.
-
-Let us also note that strict filtrations of this model is not Euclidean.
-Suppose by contrary that $\widehat{R}^{\mathcal{E}}$ is a strict filtraction of that model.
-So $R_{min}^E \subseteq \widehat{R}^{\mathcal{E}}$, since $R_{min}^E$ is the minimal Euclidean relation containing $R_{min}$.
-On the other hand, $R_{min}^E \not\subseteq R_{max}$, so is not $\widehat{R}^{\mathcal{E}}$.
-
-\end{proof}
-
 \section{i,j-Euclideaness}
 
-Let $R$ be a binary relation on $W \neq \emptyset$, then the $i,j$-Euclidean closure of $R$ (where $i, j < \omega$), denoted as $R^{E_{i,j}}$, is a binary relation defined recursively as follows:
+A binary relation $R \subseteq W \times W$ is called $i,j$-Euclidean for $i, j < \omega$, if for each $x, y, z$ such that $x R^{i} y$ and $x R^{j} z$ implies $x R z$.
+
+\begin{prop}
+  Let $\mathcal{F} = \langle W, R \rangle$, then $\mathcal{F}$ is $i,j$-Euclidean iff $\mathcal{F} \models \Diamond^i p \to \Box^j \Diamond p$.
+As a corollary, the logic ${\bf K}5^{i,j} = {\bf K} \oplus \Diamond^i p \to \Box^j \Diamond p$ is Kripke-complete.
+\end{prop}
+
+Let $R$ be a binary relation on $W \neq \emptyset$, then the $i,j$-Euclidean closure of $R$ (where $i, j < \omega$), denoted as $R^{{\bf K}5_{i,j}}$, is a binary relation defined recursively as follows:
 \begin{enumerate}
@@
 
-\begin{prop}
-  If $R$ be a binary relation on $W \neq \emptyset$, $R^{E_{i,j}}$ is $i,j$-Euclidean.
-\end{prop}
-\begin{proof}
-  By induction. Let $(x,y_1), (y_1,y_2), (y_2,y_3), \dots, (y_{i - 1},y_i), (x,z_1), (z_1,z_2), (z_2,z_3), \dots, (z_{j - 1},z_j) \in R^{E_{i,j}}$.
-  That is, each of these pairs belong to some $R_k$ and every $k < \omega$ for every pair is its own.
-  If $k = 0$ for each pair, then $(y_i, z_j) \in R_1$.
-  If some those $k$ is non-zero, then we take the greatest amongst them, say $e$ and all given pairs will be elements of $R_{e + 1}$.
-\end{proof}
+Note that the following property holds not for every ${\bf K5}^{i,j}$, consider ${\bf KB}$ as an example (clearly, ${\bf KB} = {\bf K5}^{0,1}$).
+
+\begin{lemma}
+${\bf K5}^{i,j}$ has $n$ non-equivalent sequences of modalities, if . TODO: define $n$.
+\end{lemma}
 
 \begin{theorem}
-  ${\bf K5}^{i,j} = {\bf K} \oplus \Diamond^{i} p \to \Box^{j} \Diamond p$ admits definable filtrations.
+  ${\bf K5}^{i,j}$ admits definable filtrations.
 \end{theorem}

### 2021-06-24 filtration-stuff/NewText.tex
wip fixes related to filtrations for fusions

@@
 
-\section{Finite ``canonical'' models}
-
-Let $\mathcal{L}$ be a normal modal logic, $\mathcal{M}_{\mathcal{L}}$ its canonical model, and $\Gamma$ a finite $\operatorname{Sub}$-closed set of formulas.
-Let us put $\Gamma' = \operatorname{Sub}(\varphi) \cup \{ \neg \psi \: | \: \psi \in \operatorname{Sub}(\varphi) \}$.
-
-A subset $\Delta \subseteq '\Gamma$ is a \emph{finite $\mathcal{L}$-consistent set} if $\neg \bigwedge \Delta \notin \mathcal{L}$.
-A subset $\Delta$ is maximal, if (the following are obviously equivalent):
-\begin{enumerate}
-  \item $\Delta$ is maximal amongst finite $\mathcal{L}$-consistent sets,
-  \item For each $\psi \in \operatorname{Sub}(\varphi)$ either $\psi \in \Delta$ or
-  $\neg \psi \in \Delta$.
-\end{enumerate}
-
-Every finite $\mathcal{L}$-theory is clearly can be extended to some maximal one. It is the finite version of Lindenbaum's lemma.
-
-\begin{defin}\label{fin}
-  Let $\mathcal{L}$ be a modal logic and $\Gamma$ be a finite $\operatorname{Sub}$-closed set of formulas.
-A finite ``canonical'' model is a triple $\mathcal{M}_{\mathcal{L}}^{\Gamma} = \langle W_{\mathcal{L}}^{\Gamma}, R_{\mathcal{L}}^{\Gamma}, \vartheta_{\mathcal{L}}^{\Gamma} \rangle$, where
-  \begin{enumerate}
-    \item $W_{\mathcal{L}}^{\Gamma}$ is the set all maximal theories that extend finite $\mathcal{L}$-theories
-    \item $R_{\mathcal{L}}^{\Gamma}$ is a relation such that
-    $\langle W_{\mathcal{L}}^{\Gamma}, R_{\mathcal{L}}^{\Gamma} \rangle$ is an $\mathcal{L}$-frame
-    and
-    \begin{center}
-    $\forall \Box \psi \in \operatorname{Sub}(\varphi) \:\:
-    \forall \Delta_1 \in W_{\mathcal{L}}^{\Gamma} \:\: (\Box \psi \in \Delta_1 \Leftrightarrow
-    \forall \Delta_2 \in R_{\mathcal{L}}^{\Gamma}(\Delta_1) \:\: \psi \in \Delta_2)$
-    \end{center}
-    \item $\vartheta_{\mathcal{L}}^{\Gamma}(p) = \{ \Delta \in W_{\mathcal{L}}^{\varphi} \: | \: p \in \Delta \}$ for every variable $p \in \Gamma$.
-  \end{enumerate}
-\end{defin}
-
-\begin{lemma}\label{good}
-  Let $\mathcal{L}$ be a modal logic and $\varphi \not\in \mathcal{L}$, then $\mathcal{M}_{\mathcal{L}}^{\operatorname{Sub}(\varphi)} \nvDash \varphi$.
-\end{lemma}
-
-\begin{lemma} Let $\mathcal{L}$ be a modal logic and $\Gamma$ a finite $\operatorname{Sub}$-closed set of formulas, then if $\mathcal{L}$ admits strict filtrations, then there exists a finite ``canonical'' model $\mathcal{M}_{\mathcal{L}}^{\Gamma}$ such that $\mathcal{M}_{\mathcal{L}}^{\Gamma} \models \mathcal{L}$.
-\end{lemma}
-
-\begin{proof}
-  $ $
-$(\Rightarrow)$
-  Let $\Gamma$ be a finite $\operatorname{Sub}$-closed of formulas.
-$\mathcal{L}$ admits strict filtrations, so the filtration of the canonical model $\mathcal{M}_{\mathcal{L}}$ through $\Gamma$ is also an $\mathcal{L}$-model.
-The underlying set of $\mathcal{M}_{\mathcal{L}} / \sim_{\Gamma}$ consists of maximal $\mathcal{L}$ theories up to $\Gamma$-equivalence and this quotient set is finite.
-
-  It is readily checked that the quotient model $\mathcal{M}_{\mathcal{L}} / \sim_{\Gamma}$ satifsfies Definition~\ref{fin}.
-\end{proof}
-
-The converse implication does not have to true generally. ${\bf GL}$ might be an example of a logic that has the ``finite canonical'' model property with no filtrations.
-
 \section{Fusion stuff}
@@
 \begin{lemma}\label{extension}
-  $ $
+  Let $\Gamma$ be a finite and $\operatorname{Sub}$-closed set of formulas.
 
-  \begin{itemize}
-  \item Let $\Gamma$ be a finite and $\operatorname{Sub}$-closed set of formulas and $\mathcal{M} = \langle W, R, \vartheta \rangle$.
+  \begin{enumerate}
+  \item  $\mathcal{M} = \langle W, R, \vartheta \rangle$ be a ${\bf K}5$-model.
 Consider $\Gamma' = \Gamma \cup \{ \Diamond \Box \psi \: | \: \Box \psi \in \Gamma \}$.
@@
 Then a model $\widehat{M} = \langle W / \sim_{\Delta}, (R^{min}_{\Delta})^{E}, \widehat{\vartheta} \rangle$ is a filtration of $\mathcal{M}$ through $\Delta$.
-  \item The same for ${\bf K} \oplus \Diamond \Diamond \Diamond p \to \Diamond p$
-  \end{itemize}
+  \item Let $\mathcal{L} = {\bf K} \oplus \Diamond \Diamond \Diamond p \to \Diamond p$, then we have the similar statement for $\mathcal{L}$, where $\Delta$ is an extension of $\Gamma' = \{ \Diamond \Diamond \psi \: | \: \Diamond \psi \}$.
+  \end{enumerate}
 \end{lemma}
@@
 \begin{proof}
-  Recall that $(R^{min}_{\Delta})^{E}$ is defined inductively as:
+\begin{enumerate}
+  $ $
+
+  \item Recall that $(R^{min}_{\Delta})^{E}$ is defined inductively as:
   \begin{enumerate}
@@
 
-  Suppose the statement holds for $R_{\Delta}^{n}$, $(\hat{x},\hat{y}) \in R_{\Delta}^{n + 1}$, that is, there exists $\hat{z}$ such that $(\hat{z}, \hat{x}), (\hat{z}, \hat{y}) \in R_{\Delta}^{n}$.
-  Let $\mathcal{M}, x \models \Box \psi$ (for $\Box \psi \in \Gamma$).
+  Suppose the statement holds for $R_{\Delta}^{n}$, $(\hat{x},\hat{y}) \in R_{\Delta}^{n + 1}$ such that $\mathcal{M}, y \models \psi$ for $\Diamond \psi \in \Delta$.
 
-  We rewrite $(\hat{z}, \hat{x}) \in R_{\Delta}^{n}$ as the following sausage (for some $\widehat{u_1},  \widehat{u_2}, \dots, \widehat{u_{n - 1}}, \widehat{u_{n}}$):
+  According to the third item of Lemma~\ref{equivHorn}, this is the same as there exist $\widehat{z_0}, \widehat{z_1}, \dots, \widehat{z_{n-1}}, \widehat{z_n}$ such that
+  $\widehat{z_1} R^{min}_{\Delta} \hat{x}$, $\widehat{z_n} R^{min}_{\Delta} \hat{y}$, and for each $i \in n + 1$ we have either $\widehat{z_i} R^{min}_{\Delta} \widehat{z_{i + 1}}$ or $\widehat{z_{i + 1}} R^{min}_{\Delta} \widehat{z_i}$.
+
+  We visualise this with the following graph:
 
   \xymatrix{
-  &&& \hat{z} & \ar[l]_{R_{\Delta}^{min}} \ar[r]^{R'} \widehat{u_1} & \ar[r]^{R'} \widehat{u_2} \ar[r]^{R'} & \dots \ar[r]^{R'} & \widehat{u_{n - 1}} \ar[r]^{R'} & \widehat{u_{n}} \ar[r]^{R_{\Delta}^{min}} & \hat{x}
+  &&& \hat{x} & \ar[l]_{R_{\Delta}^{min}} \ar@{<->}[r]^{R'} \widehat{z_0} & \ar@{<->}[r]^{R'} \widehat{z_1} \ar@{<->}[r]^{R'} & \dots \ar@{<->}[r]^{R'} & \widehat{z_{n - 1}} \ar[r]^{R'} & \widehat{z_n} \ar[r]^{R_{\Delta}^{min}} & \hat{y}
   }
   where $R'$ is either $R_{\Delta}^{min}$ or its converse.
-We have $\mathcal{M}, z \models \Box \psi$, since $\mathcal{M}, x \models \Box \psi$ implies $\mathcal{M}, u_{n} \models \Diamond \Box \psi$.
+We have $\mathcal{M}, z \models \Box \psi$, $\mathcal{M}, y \models \psi$, so $\mathcal{M}, z_{n} \models \Diamond \psi$.
+Since $\mathcal{M}$ is a ${\bf K}5$-model, we have $\mathcal{M}, z_{n} \models \Box \Diamond \psi$.
+
   After that we apply the following property of ${\bf K}5$-models:
   \begin{center}
-    For each $a, b \in M_i$ such that $a R_i b$ we have $\mathcal{M}_i, a \models \Diamond \Box p_{\psi}$ iff $\mathcal{M}_i, b \models \Diamond \Box p_{\psi}$
+    Let $\mathcal{M} \models {\bf K}5$ and $\varphi$ a formula, then for each $a, b \in \mathcal{M}$ such that $a R b$ we have $\mathcal{M}, a \models \Box \Diamond \varphi$ iff $\mathcal{M}, b \models \Box \Diamond \varphi$
   \end{center}
-  Note that we always stay within $\Delta$ since $\Gamma' \subseteq \Delta$.
-  Finally, we conclude $\mathcal{M}, x \models \psi$ from IH and $\mathcal{M}, z \models \Box \psi$.
+  So we have $\mathcal{M}, z_0 \models \Box \Diamond \varphi$. Note that we always stay within $\Delta$. Thus, $\mathcal{M}, x \models \Diamond \varphi$.
 
-  The second item is shown similarly.
+  \item Let us prove the second item. Let $\mathcal{M}$ be an $\mathcal{L}$-model.
+  The $\mathcal{L}$-closure of the minimal filtered relation modulo $\Delta$, namely ${R_{\Delta}^{min}}^{\mathcal{L}}$ has the following form:
+  \begin{center}
+    ${R_{\Delta}^{min}}^{\mathcal{L}} = \bigcup \limits_{i < \omega} {R_{\Delta}^{min}}^{2i + 1}$
+  \end{center}
+  Let $\Diamond \psi \in \Delta$, $(\hat{x}, \hat{y}) \in {R_{\Delta}^{min}}^{\mathcal{L}}$, and $\mathcal{M}, y \models \psi$. The base case is self-evident.
+  Let $k < \omega$ and $(\hat{x}, \hat{y}) \in {R_{\Delta}^{min}}^{2k + 3}$ and the statement holds for ${R_{\Delta}^{min}}^{2k + 1}$.
+
+\end{enumerate}
 \end{proof}

### 2021-06-24 filtration-stuff/Text.tex
more facts about definable filtrations for fusions
wip fixes related to filtrations for fusions

@@
 \date{}
-\title{Notes on filtration of logics containing {\bf K5}}
+\title{Notes on filtrations for logics that contain {\bf K5}}
 
@@
 \begin{lemma}
-  ${\bf K5}$ admit filtration.
+  ${\bf K5}$ admit filtrations.
 \end{lemma}
@@
   It is readily checked that the quotient model $\mathcal{M}_{\mathcal{L}} / \sim_{\Gamma}$ satifsfies Definition~\ref{fin}.
-
-    \item $(\Leftarrow)$
-
-    Let $\Gamma$ be a finite $\operatorname{Sub}$-closed of formulas and $\mathcal{M}_{\mathcal{L}}^{\Gamma} = \langle W_{\mathcal{L}}^{\Gamma}, R_{\mathcal{L}}^{\Gamma}, \vartheta_{\mathcal{L}}^{\Gamma} \rangle$ a ``finite canonical'' model generated by $\Gamma$.
-
-    Let us show that there exists a model $\widehat{\mathcal{M}}$ such that
-    $\widehat{\mathcal{M}}$ is a strict filtration of $\mathcal{M}$ through $\operatorname{Sub}(\Gamma)$.
-
-    Let $\sim$ be an equivalence relation on $\underline{\mathcal{M}}$ defined as usual:
-
-    \begin{center}
-      $u \sim v$ iff $\forall \psi \in \Gamma$
-      $\mathcal{M}, u \models \psi \Leftrightarrow \mathcal{M}, v \models \psi$
-    \end{center}
-
-    Let $[u] \in \mathcal{M}/\sim$, consider its truth set $||[u]|| = \{ \psi \in \Gamma \: | \: \mathcal{M}, u \models \psi \}$.
-    $||[u]||$ is clearly a finite $\mathcal{L}$-consistent set and it is maximal.
-    Consider a map
-    \begin{center}
-    $\iota : W/\sim \to W_{\mathcal{L}}^{\varphi}$ with $\iota : [u] \mapsto ||[u]||$
-    \end{center}
-
-    Consider the set $\iota(W/\sim)$ that we denote as $\operatorname{Im}(\iota)$. Let us show that $\operatorname{Im}(\iota)$ is $R_{\mathcal{L}}^{\varphi}$-closed.
-
-    Let $\iota([u]) R_{\mathcal{L}}^{\varphi} \Delta$, let us show that $\Delta \in \iota(W/\sim)$, that is, $\Delta = \iota([v])$ for some $v \in W$.
-
-    Note that $\bigwedge \limits_{\psi \in \Delta} \psi \in \Delta$ since $\Delta$ is finite.
-    This conjuction has the form $\bigwedge \limits_{\psi \in \Delta} \psi \in \Delta = \psi_1 \land \dots \land \psi_n$ for some $n < \omega$ and for some $\psi_i \in \Gamma \cup \neg \Gamma$, where $0 < i < \omega$.
-
-    Therefore $\mathcal{M}^{\Gamma}_{\mathcal{L}}, \iota([u]) \models \Diamond (\psi_1 \land \dots \land \psi_n)$.
-    Then $\mathcal{M}, u \models \Diamond (\psi_1 \land \dots \land \psi_n)$. That implies $\mathcal{M}, v \models \psi_1 \land \dots \land \psi_n$ for some $v \in R(u)$.
-    This is clearly a desired $v$ since $||[v]|| \subseteq \Delta$ and such a $\Delta$ is unique.
-
-    Thus, $\langle \operatorname{Im}(\iota), R_{\mathcal{L}}^{\Gamma}, {\vartheta'}_{\mathcal{L}}^{\Gamma} \rangle$, where ${\vartheta'}_{\mathcal{L}}^{\Gamma}(p) = \vartheta_{\mathcal{L}}^{\Gamma}(p) \cap \operatorname{Im}(\iota)$
-    is a generated submodel of $\mathcal{M}_{\mathcal{L}}^{\Gamma}$.
-
-    Consider a model $\widehat{\mathcal{M}} = \langle W/\sim, \widehat{R}, \widehat{\vartheta} \rangle$, where $\widehat{R}$ is defined as
-    \begin{center}
-      $[u] \widehat{R} [v]$ iff $||[u]|| {R}_{\mathcal{L}}^{\Gamma} ||[v]||$
-    \end{center}
-
-    Let us show that $R_{min} \subseteq \widehat{R} \subseteq R_{max}$.
-
-    Suppose that there are $u' \in [u]$ and $v' \in [v]$ with $u' R v'$. Let us show that
-    $[u] \widehat{R} [v]$. That is, let us show that $\Diamond \iota(v) \subseteq \iota(v)$.
-    But if $\mathcal{M}, v' \models \psi$ for $\Diamond \psi \in \Gamma$, then
-    $\mathcal{M}, u' \models \Diamond \psi$. Recall that $u' \sim u$ and $v' \sim v$, so we have $\Diamond ||[v]|| \subseteq ||[u]||$.
-
-    The second inclusion follows from the definition of a relation in a finite ``canonical'' model.
-  \end{enumerate}
 \end{proof}
 
-In particular, ${\bf K}5$ fails to have the finite ``canonical'' model property that
-follows from the contraposition of the theorem above and Theorem~\ref{sad}.
+The converse implication does not have to true generally. ${\bf GL}$ might be an example of a logic that has the ``finite canonical'' model property with no filtrations.
 

### 2021-06-26 problem-5/Text.tex
New file with filtrations stuff
updates in cylindric notes

@@
 
+\let\strokeL\L
+\DeclareRobustCommand{\L}{\ifmmode\mathbf{L}\else\strokeL\fi}
+
 \author{Daniel Rogozin}
@@
 
-\section{Boolean algebras with operators and cylindric algebras}
+\section{Boolean algebras with operators and cylindric algebras, a bit of the backgroud}
+
+Let $a \in \mathcal{B}$ be an element of a Boolean algebra $\mathcal{B}$, $a$ is called an atom, if for every $b \in \mathcal{B}$
+$b < a$ implies $b = 0$. That is, an atom is a minimal non-zero element. $\operatorname{At}(\mathcal{B})$ is the set of all atoms
+of $\mathcal{B}$.
+
+Let $\mathcal{B}$ be a Boolean algebra and $\mathcal{F}$ a field of sets such that $h : \mathcal{B} \to \mathcal{F}$ is a
+representation of $\mathcal{B}$, then $\mathcal{B}$ is a complete representation of $\mathcal{B}$, if for every
+$A \subseteq \mathcal{B}$ whenever $\Sigma A$ we have the following:
+\begin{center}
+  $h(\Sigma A) = \bigcup h[A]$
+\end{center}
+
+\begin{theorem} \label{completeboolean}
+  Let $\mathcal{B}$ be a Boolean algebra, then $\mathcal{B}$ is atomic iff $\mathcal{B}$ is completely representable. See \cite[Corollary 6]{hirsch1997complete}.
+\end{theorem}
 
@@
   $\mathcal{F} = \langle W, \{ R_{i} \}_{i \in I} \rangle$
-  such that $R_{i}$ is a $n+1$-ary relation symbol, if $\Omega_{i \in I}$ is an $n$-ary function symbol,
+  such that $R_{i}$ is a $n+1$-ary relation symbol, where $\Omega_{i \in I}$ is an $n$-ary function symbol,
   \item Let $\mathcal{B}$ be an atomic BAO of the signature $I$,
-the atom structure of $\mathcal{B}$, written as ${\bf At} \mathcal{B}$, is an atom structure $\langle \operatorname{At}(\mathcal{B}), \{ R_{i}\}_{i \in I} \rangle$ such that for each
+the atom structure of $\mathcal{B}$, written as $\mathfrak{At} \mathcal{B}$, is an atom structure $\langle \operatorname{At}(\mathcal{B}), \{ R_{i}\}_{i \in I} \rangle$ such that for each
 $a, b_0, \dots, b_{n + 1} \in \operatorname{At}(\mathcal{B})$ and for each $i \in I$
 \begin{center}
-  ${\bf At} \mathcal{B} \models R_{i}(a,  b_0, \dots, b_{n + 1})$ iff $\mathcal{B} \models a \leq \Omega_{i}(b_0, \dots, b_{n + 1})$
+  $\mathfrak{At} \mathcal{B} \models R_{i}(a,  b_0, \dots, b_{n + 1})$ iff $\mathcal{B} \models a \leq \Omega_{i}(b_0, \dots, b_{n + 1})$
 \end{center}
@@
   \begin{enumerate}
-    \item Let $\mathcal{B}$ be a complete atomic BAO, then $\mathcal{B} \cong \bf{Cm} (\bf{At}(\mathcal{B}))$,
-    \item Let $\mathcal{F}$ be an atom structure, then $\mathcal{F} \cong {\bf At}({\bf Cm}(\mathcal{B}))$.
+    \item Let $\mathcal{B}$ be a complete atomic BAO, then $\mathcal{B} \cong \mathfrak{Cm} (\bf{At}(\mathcal{B}))$,
+    \item Let $\mathcal{F}$ be an atom structure, then $\mathcal{F} \cong \mathfrak{At}(\mathfrak{Cm}(\mathcal{B}))$.
   \end{enumerate}
@@
 
-Let $A$ be a non-empty subset of a Boolean algebra $\mathcal{B}$, $A$ is a \emph{filter}, if $A$ is closed under finite infima and upwardly closed.
+Let $A$ be a non-empty subset of a Boolean algebra $\mathcal{B}$, $A$ is a \emph{filter}, if $A$ is closed under finite infima and upward closed.
 $A$ is an ultrafilter, if it has no non-trivial extensions.
 That is, if $A \subseteq A'$, then $A' = \mathcal{B}$.
+This is a well-known fact that every filter can be extended to a maximal one using Zorn's lemma.
+
+The following definition is due to, for example, \cite[Definition 5.40]{venema2010}.
 
@@
   \begin{enumerate}
-    \item The canonical extension of $\mathcal{B}$ is a complex algebra of the canonical frame ${\bf Cm}(\mathcal{F}_{\mathcal{B}})$ denoted as $\mathcal{B}^{+}$,
+    \item The canonical extension of $\mathcal{B}$ is a complex algebra of the canonical frame $\mathfrak{Cm}(\mathcal{F}_{\mathcal{B}})$ denoted as $\mathcal{B}^{+}$,
     \item The class of BAOs is canonical, if it is closed under canonical extensions.
@@
 
-\begin{defin}
-$ $
-
-  \begin{enumerate}
-    \item A subset of $\prescript{\alpha}{}U$ is an $\alpha$-ry relation on $U$. For $i, j < \alpha$, the $i,j$-diagonal $D_{ij}$ is the set of all
-    elements of $U$ such that $y_i = y_j$.
+A subset of $\prescript{\alpha}{}U$ is an $\alpha$-ry relation on $U$.
+For $i, j < \alpha$, the \emph{$i,j$-diagonal} $D_{ij}$ is the set of all elements of $\prescript{\alpha}{}U$ such that $y_i = y_j$.
 
 If $i < \alpha$ and $X$ is an $\alpha$-ry relation on $U$, then
-    the $i$-th cylindrification $C_i X$ is the set of all elements of $U$ that agree with some element of $X$ on each coordinate except the
-    $i$-th one. To be more precise,
-    $C_i X = \{ y \in \prescript{\alpha}{}U \: | \: \exists x \in X \forall i < \alpha \: (i \neq j \Rightarrow y_j = x_j)\}$.
-    \item A cylindic set algebra of dimension $\alpha$ is an algebra consisting of a set $S$ of $\alpha$-ry relation on some base set $U$
-    with the constants and operations $0 = \emptyset$, $1 = \prescript{\alpha}{}U$, $\cap$, $-$, the diagonal elements $\{ D_{ij} \}_{i, j < \alpha}$,
-    the cylindrifications $\{ C \}_{i < \alpha}$.
+the $i$-th cylindrification $C_i X$ is the set of all elements of $U$ that agree with some element of $X$ on each coordinate except, perhaps, the $i$-th one. To be more precise,
+\begin{center}
+$C_i X = \{ y \in \prescript{\alpha}{}U \: |
+\: \exists x \in X \forall i < \alpha \: (i \neq j \Rightarrow y_j = x_j)\}$.
+\end{center}
+We define the following equivalence relation for $i < \alpha$ and $x, y \in \prescript{\alpha}{}U$:
+\begin{center}
+  $x \equiv_i y \Leftrightarrow \forall j \in \alpha \: (i \neq j \Rightarrow x(i) = y(j))$
+\end{center}
+Then one may reformulate the definition of the $i$-th cylindrification in the following way:
 
-    A generalised cylindric set algebra of dimension $\alpha$ is a subdirect of
-    cylindric algebras that have dimension $\alpha$
-    \item A cylindric algebra of dimension $\alpha$ is an algebra $\mathcal{C} = \langle \mathcal{B}, \{ c_i \}_{i < \alpha}, \{ d_{ij} \}_{i, j < \alpha} \rangle$ such that
+\begin{center}
+  $C_i X = \{ y \in \prescript{\alpha}{}U \: | \: \exists x \in X \:\: x \equiv_i y \}$
+\end{center}
+
+According to this version of the definiton, one may think of the cylindrification as an ${\bf S}5$ modal operator.
+
+The following definition is due to \cite{sankappanavar1981course}:
+
+\begin{defin}
+  Let $(\mathcal{A}_i)_{i \in I}$ be a family of algebras (of an abstract signature) and $\mathcal{A}$ is a subalgebra of $\prod \limits_{i \in I} \mathcal{A}_i$, then $\mathcal{A}$ is a subdirect product, if every projection is onto.
+That is, for every $i \in I$, $\pi_i[\mathcal{A}] = \mathcal{A}_i$.
+\end{defin}
+
+\begin{defin}
+  A cylindic set algebra of dimension $\alpha$ is an algebra consisting of a set $S$ of $\alpha$-ry relation on some base set $U$
+    with the constants and operations $0 = \emptyset$, $1 = \prescript{\alpha}{}U$, $\cap$, $-$, the diagonal elements $\{ D_{ij} \}_{i, j < \alpha}$, the cylindrifications $\{ C_i \}_{i < \alpha}$.
+A generalised cylindric set algebra of dimension $\alpha$ is a subdirect of cylindric algebras that have dimension $\alpha$
+\end{defin}
+
+\begin{defin}
+    A cylindric algebra of dimension $\alpha$ is an algebra $\mathcal{C} = \langle \mathcal{B}, \{ c_i \}_{i < \alpha}, \{ d_{ij} \}_{i, j < \alpha} \rangle$ such that
     \begin{itemize}
@@
     ${\bf CA}_{\alpha}$ is the class of all cylindric algebras of dimension $\alpha$
-    \item An $\alpha$-dimensional cylindric algebra $C$ is representable, if it is isomorphic to a generalised cylindric set algebra
-    of dimension $\alpha$. Such is isomorphism is a representation of $C$.
-
-    ${\bf RCA}_{\alpha}$ is the class of all representable cylindric algebras that have dimension $\alpha$. In particular, we are interested in the case when $\alpha = \omega$.
-  \end{enumerate}
 \end{defin}
 
-One may define a representation explicitly in the following way:
+One may define a representation of a cylindric algebra explicitly in the following way:
 
@@
 \end{defin}
+
+An $\alpha$-dimensional cylindric algebra $C$ is representable, if it is isomorphic to a generalised cylindric set algebra
+of dimension $\alpha$.
+Such is isomorphism is a representation of $C$.
+${\bf RCA}_{\alpha}$ is the class of all representable cylindric algebras that have dimension $\alpha$.
+In particular, we are interested in the case when $\alpha = \omega$.
 
@@
 
-\begin{defin} Atom
-\end{defin}
-
-Let us concretise the definition of a canonical extension for $\alpha$-dimensional cylindric algebras:
+Let us concretise the definition of a canonical extension for ${\bf CA}_{\alpha}$-type BAOs.
 
@@
 
-\section{The result}
-
-\begin{lemma}\label{Neat}
-  Let $\mathcal{A}$ be a BAO of type ${\bf CA}_{\alpha}$ and $\mathcal{B}$ be a $\beta$-dimensional cylindric algebra such that $\beta \leq \alpha$ and $\mathcal{A}$ neatly embeds to $\mathcal{B}$ by a complete embedding.
-
-  \begin{enumerate}
-  \item $\mathcal{A}^{+}$ neatly embeds to $\mathcal{B}^{+}$ by a complete embedding.
-  \item $\mathcal{A}$ is atomic.
-  \end{enumerate}
-\end{lemma}
-
-\begin{proof}
-$ $
-
-\begin{enumerate}
-  \item See \cite[Remark 2.7.25]{henkin1971cylindric}.
-  \item Is it true?
-\end{enumerate}
-\end{proof}
-
-\begin{lemma}[See \cite{hirsch1997complete}]
-  A Boolean algebra is compeletely representable iff it has an atomic representation.
-\end{lemma}
+\section{Completely representable cylindric algebras of dimension $\omega$}
 
@@
   \begin{enumerate}
-    \item $l(x) \leq d_{ij}$ iff $x(i) = x(j)$
+    \item $l(x) \leq d_{ij}$ iff $x_i = x_j$
     \item $x \equiv_i y$ implies $l(x) \leq c_i l(y)$
@@
 The game $\mathcal{G}^{\kappa}(\mathcal{A})$ is defined as follows.
-The game has two players: $\forall$ (Abelard) and $\exists$ (H\'{e}lo\"{i}se).
-A play of the game is the sequence
+The game has two players: $\forall$ (Abelard, he/his) and $\exists$ (H\'{e}lo\"{i}se, she/her).
+A play of the game $\mathcal{G}^{\kappa}(\mathcal{A})$ is the sequence of networks
   \begin{center}
@@
 
-    Let $0 < \lambda$ be a cardinal such that $\lambda + 1 < \kappa$ and $\mathcal{N}_{\lambda} = \langle N_{\lambda}, l_{\lambda} \rangle$ is a network that has been played.
+    Let $0 < \lambda$ be a cardinal such that $\lambda + 1 < \kappa$ and a network $\mathcal{N}_{\lambda} = \langle N_{\lambda}, l_{\lambda} \rangle$ has been already played.
 
@@
     \item $\Rightarrow$
-    If $\mathcal{A}$ is completely representable, then its Boolean reduct is representable as well. $\exists$ maintains that embedding to win the play.
+    If $\mathcal{A}$ is completely representable, then its Boolean reduct is completely representable as well by Theorem~\ref{completeboolean}.
+$\exists$ maintains that embedding to win the play.
+TODO: write down this proof in more detail
 
@@
     Suppose $\exists$ has a winning strategy in $\mathcal{G}^{\kappa + \omega}(\mathcal{A})$.
-    In every round $\forall$ picks all possible $i < \omega$, $a \in \operatorname{At}(\mathcal{A})$, all possible hyperedges and all appropriate atoms.
-    ...
-    Question: Is this claim correct?
+    In every round $\forall$ picks all possible $i < \omega$, $a \in \operatorname{At}(\mathcal{A})$, all possible hyperedges and all appropriate atoms and $\exists$ has a proper response for every $\forall$'s move.
+
+    For each atom consider a play of the game with fewer than $\kappa + \omega$ nodes.
+For each $a \in \operatorname{At}(\mathcal{A})$ we associate a network $\mathcal{N}_a$, the resulting network of a corresponding game.
+Consider the set $\{ \mathcal{N}_a \: | \: a \in \operatorname{At}(\mathcal{A}) \}$.
+
+    Let $a$ be an atom, consider the network $\mathcal{N}_a = \langle V, l_a \rangle$.
+If there was not $x \in V$ such that $l_a(x) = a$, then $\forall$ would have a winning strategy, but that is not true, such an $x$ does exist.
+The second item of this criterion follows from the presence of a winning strategy for $\exists$ as well.
+
+    So we define a map $rep$:
+    \begin{center}
+    $rep(a) = \{ x \: | \: \exists b \in \operatorname{At}(\mathcal{A}) \: x \in \prescript{\omega}{}N_a, l_a(x) \leq b \}$.
+  \end{center}
+    We check that $rep$ preserves cylindrifications and diagonal elements. Let $i, j < \omega$ and $a \in \mathcal{A}$:
+    \begin{enumerate}
+    \item Suppose $x \in rep(c_i a)$, then there exists an atom $b$ such that $x \in \prescript{\omega}{}N_b$ with $l_b(x) \leq c_i a$.
+Then there exists $y \equiv_i x$ with $l_b(y) \leq a$, so $x \in {\bf C}_i(rep(a))$.
+
+    If $x \in {\bf C}_i(rep(a))$, then there exists $y \equiv_i x$ such that $y \in rep(a)$, that is, there exists an atom $b$ such that $y \in \prescript{\omega}{}N_b$ and $l_b(y) \leq a$.
+    \item If $x \in rep(d_{ij})$, so there exists an atom $b$ with $x \in \prescript{\omega}{}N_a$ and $l_b(x) \leq d_{ij}$, then $x_i = x_j$, then $x \in D_{ij}$.
+    \end{enumerate}
+  \end{enumerate}
+\end{proof}
+
+\begin{theorem} Let $\mathcal{A}$ be a BAO of type ${\bf CA}_{\omega}$:
+
+\begin{enumerate}
+  \item $\exists$ has a winning strategy in $\mathcal{G}_{m}(\mathcal{A})$ ($m < \omega$), then $\exists$ has a winning strategy in $\mathcal{G}_{\omega}(\Pi_U \mathcal{A})$, where $\Pi_U \mathcal{A}$ is the non-principal ultrapower of $\mathcal{A}$ modulo ultrafilter $U$ over $\omega$.
+  \item $\exists$ has a winning strategy in $\mathcal{G}_{m}(\mathcal{A})$ (for every $m < \omega$) iff $\mathcal{A}$ is elementarily equivalent to a completely representable cylindric algebra of dimension $\omega$.
+\end{enumerate}
+\end{theorem}
+
+\begin{proof}
+  $ $
+
+  The argument uses \L o\'s's Theorem, see \cite[Theorem 8.5.3]{hodges1997shorter}.
+
+  \begin{enumerate}
+    \item
+    \item
+  \end{enumerate}
+\end{proof}
+
+\section{The result itself}
+
+\begin{lemma}\label{Neat}
+  Let $\mathcal{A}$ be a BAO of type ${\bf CA}_{\alpha}$ and $\mathcal{B}$ be a $\beta$-dimensional cylindric algebra such that $\beta \leq \alpha$ and $\mathcal{A}$ neatly embeds to $\mathcal{B}$ by a complete embedding.
+
+  \begin{enumerate}
+  \item $\mathcal{A}^{+}$ neatly embeds to $\mathcal{B}^{+}$ by a complete embedding.
+  \item $\mathcal{A}$ is atomic.
+  \end{enumerate}
+\end{lemma}
+
+\begin{proof}
+$ $
+
+\begin{enumerate}
+  \item See \cite[Remark 2.7.25]{henkin1971cylindric}.
+  \item Is it true?
 \end{enumerate}
@@
 
-\section{Sahlqvist and canonical axiomatisation of ${\bf CA}_{\omega}$}
+\section{(Lack of) canonical axiomatisation of ${\bf CA}_{\omega}$}
+
+Here we are going to show that ${\bf CA}_{\omega}$ fails to have a canonical axiomatisation, the similar results for ${\bf RRA}$ and ${\bf RCA}_n$ for finite $n \geq 3$ have been shown by Hodkinson and Venema \cite{hodkinson2005canonical} and by Bulian and Hodkinson respectively \cite{bulian2013bare}.
 

### 2021-07-04 filtration-stuff/NewText.tex
proof for fusions

@@
 Consider $\Gamma' = \Gamma \cup \{ \Diamond \Box \psi \: | \: \Box \psi \in \Gamma \}$.
-Let $\Delta$ be any finite and $\operatorname{Sub}$-closed extension of $\Gamma'$.
+Let $\Delta$ be any finite and $\operatorname{Sub}$-closed extension of $\Gamma'$ that contains all $\Diamond \Box \psi$ for each $\Box \psi \in \Delta$.
 Then a model $\widehat{M} = \langle W / \sim_{\Delta}, (R^{min}_{\Delta})^{E}, \widehat{\vartheta} \rangle$ is a filtration of $\mathcal{M}$ through $\Delta$.
-  \item Let $\mathcal{L} = {\bf K} \oplus \Diamond \Diamond \Diamond p \to \Diamond p$, then we have the similar statement for $\mathcal{L}$, where $\Delta$ is an extension of $\Gamma' = \{ \Diamond \Diamond \psi \: | \: \Diamond \psi \}$.
+  \item Let $\mathcal{L} = {\bf K} \oplus \Diamond \Diamond \Diamond p \to \Diamond p$, then we have the similar statement for $\mathcal{L}$, where $\Delta$ is an extension of $\Gamma' = \{ \Diamond \Diamond \psi \: | \: \Diamond \psi \}$ that contains all $\Diamond \Diamond \varphi$ for $\Diamond \varphi \in \Delta$.
   \end{enumerate}
@@
   \begin{center}
-    ${R_{\Delta}^{min}}^{\mathcal{L}} = \bigcup \limits_{i < \omega} {R_{\Delta}^{min}}^{2i + 1}$
+    ${R_{\Delta}^{min}}^{\mathcal{L}} = \bigcup \limits_{k < \omega} {R_{\Delta}^{min}}^{2k + 1}$
   \end{center}
-  Let $\Diamond \psi \in \Delta$, $(\hat{x}, \hat{y}) \in {R_{\Delta}^{min}}^{\mathcal{L}}$, and $\mathcal{M}, y \models \psi$. The base case is self-evident.
-  Let $k < \omega$ and $(\hat{x}, \hat{y}) \in {R_{\Delta}^{min}}^{2k + 3}$ and the statement holds for ${R_{\Delta}^{min}}^{2k + 1}$.
+  We reformulate this closure equivalently as follows:
+  \begin{enumerate}
+    \item $R_0 = R_{\Delta}^{min}$
+    \item $R_{n + 1} = R_n \cup ((R_{\Delta}^{min})^2 \circ R_n)$
+    \item ${R_{\Delta}^{min}}^{\mathcal{L}} = \bigcup \limits_{k < \omega}$
+  \end{enumerate}
 
+  The base case is self-evident.
+Suppose the statement holds for $R_{n}$ and $(\hat{x}, \hat{y}) \in (R_{\Delta}^{min})^2 \circ R_n$, that is, there exists $\hat{z}$ such that $(\hat{x}, \hat{z}) \in (R_{\Delta}^{min})^2$ and $(\hat{z}, \hat{y}) \in R_n$.
+By IH, we have $\mathcal{M}, z \models \Diamond \varphi$.
+
+  We have the following:
+
+  \xymatrix{
+  &&&&&& \hat{x} \ar[r]^{R_{\Delta}^{min}} & \widehat{y_1} \ar[r]^{R_{\Delta}^{min}} & \hat{z}
+  }
 \end{enumerate}
+  The sequence of implications if the following:
+  \begin{center}
+    $\widehat{\mathcal{M}}, \hat{z} \models \Diamond \varphi \Rightarrow \widehat{\mathcal{M}}, \widehat{y_1} \models \Diamond \Diamond \varphi \Rightarrow \mathcal{M}, x \models \Diamond \Diamond \Diamond \varphi \Rightarrow \mathcal{M}, x \models \Diamond \varphi$
+  \end{center}
 \end{proof}

### 2021-07-04 problem-5/Text.tex
proof for fusions

@@
 
-    If $x \in {\bf C}_i(rep(a))$, then there exists $y \equiv_i x$ such that $y \in rep(a)$, that is, there exists an atom $b$ such that $y \in \prescript{\omega}{}N_b$ and $l_b(y) \leq a$.
+    Let $x \in {\bf C}_i(rep(a))$. We need $x \in (rep(c_i a))$, that is, one needs to find an atom $c$ such that $l_c(x) \leq c_i a$.
+
+    We already know that there exists $y \equiv_i x$ such that $y \in rep(a)$, that is, there exists an atom $b$ such that $y \in \prescript{\omega}{}N_b$ and $l_b(y) \leq a$.
+
     \item If $x \in rep(d_{ij})$, so there exists an atom $b$ with $x \in \prescript{\omega}{}N_a$ and $l_b(x) \leq d_{ij}$, then $x_i = x_j$, then $x \in D_{ij}$.
@@
 \begin{enumerate}
-  \item $\exists$ has a winning strategy in $\mathcal{G}_{m}(\mathcal{A})$ ($m < \omega$), then $\exists$ has a winning strategy in $\mathcal{G}_{\omega}(\Pi_U \mathcal{A})$, where $\Pi_U \mathcal{A}$ is the non-principal ultrapower of $\mathcal{A}$ modulo ultrafilter $U$ over $\omega$.
+  \item $\exists$ has a winning strategy in $\mathcal{G}_{m}(\mathcal{A})$ ($m < \omega$), then $\exists$ has a winning strategy in $\mathcal{G}_{\omega}(\Pi_U \mathcal{A})$, where $\Pi_U \mathcal{A}$ is the non-principal ultrapower of $\mathcal{A}$ modulo $U$, an ultrafilter over $\omega$.
   \item $\exists$ has a winning strategy in $\mathcal{G}_{m}(\mathcal{A})$ (for every $m < \omega$) iff $\mathcal{A}$ is elementarily equivalent to a completely representable cylindric algebra of dimension $\omega$.

### 2021-07-08 filtration-stuff/NewText.tex
bisimulation stuff

@@
 \section{Preliminaries}
+
+\begin{defin}
+  A normal modal logic is a set of formulas that contains all Boolean tautologies,
+  formulas $\Diamond p \lor \Diamond q \leftrightarrow \Diamond (p \lor q)$ and
+  $\Diamond \bot \leftrightarrow \bot$, and is closed under Modus Ponens, Substitution, and
+  Monotonicity:
+  from $\varphi \rightarrow \psi$ infer $\Diamond \varphi \rightarrow \Diamond \psi$.
+\end{defin}
+
+\begin{defin} An Kripke model is a triple $\mathcal{M} = \langle W, R, \vartheta \rangle$,
+  where $R \subseteq W \times W$, $\vartheta : \operatorname{PV} \to 2^W$, and
+  the connectives have the following semantics:
+
+  \begin{enumerate}
+    \item $\mathcal{M}, w \models p \Leftrightarrow w \in \vartheta(p)$
+    \item $\mathcal{M}, w \models \neg \varphi \Leftrightarrow \mathcal{M}, w \nvDash \varphi$
+    \item $\mathcal{M}, w \models \varphi \lor \psi \Leftrightarrow \mathcal{M}, w \models \varphi \text{ or } \mathcal{M}, w \models \psi$
+    \item $\mathcal{M}, w \models \Diamond \varphi \Leftrightarrow \exists v \in R(w) \: \mathcal{M}, v \models \varphi$
+  \end{enumerate}
+\end{defin}
+
+A modal logic $\mathcal{L}$ is called locally finite, if every fragment of $\mathcal{L}$ with a finite number of variables is finite.
+Equivalently, its weak canonical model generated by $k < \omega$ variables is finite.
+Moreover, if $\mathcal{L}$ is locally finite, then $\mathcal{L} = Log(\mathcal{F}_{\mathcal{L}} \upharpoonright k \: | \: k < \omega)$, where $\mathcal{F}_{\mathcal{L}} \upharpoonright k$ is a weak canonical frame generated by $k$ variables.
+
+\begin{defin}
+The modal depth of a formula is defined recursively by the following function:
+\begin{center}
+  $md(p_i) = 0$
+
+  $md(\neg \varphi) = md(\varphi)$
+
+  $md(\varphi \lor \psi) = \max(md(\varphi), md(\psi))$
+
+  $md(\Diamond \varphi) = 1 + md(\psi)$
+\end{center}
+
+Let $\mathcal{L}$ be a modal logic, then the modal depth of $\mathcal{L}$, denoted as $md(\mathcal{L})$, is the minimal $n$ such that any $\varphi \in \mathcal{L}$ is equivalent to some $\psi$ such that $md(\psi) = n$.
+If such an $n$ does not exists, then $md(\mathcal{L}) = \infty$.
+\end{defin}
+
+\subsection{Filtrations}
 
@@
 
+\subsection{Horn closure}
+
 \begin{defin} A first-order formula is called Horn if it has the following form (see \cite{chang1990model}):
@@
 
+\subsection{Bisimulations}
+
+A model $\mathcal{M}$ is called \emph{weak} model, if its valuation is restriced to some finite subset of variables, namely $PV \upharpoonright k$, if the cardinality of this subset equals $k$.
+
+\begin{defin}
+Let $\mathcal{M}_1 = \langle W_1, R_1, \vartheta_1 \rangle$ and $\mathcal{M}_2 = \langle W_2, R_2, \vartheta_2 \rangle$ be weak Kripke models (with the same variables), $x \in W_1$, and $y \in W_2$, then $x$ and $y$ are $0$-equivalent, if the following holds:
+
+\begin{center}
+  $x \equiv_0 y$ iff for each $p \in PV \upharpoonright k$ (or for each $p \in PV$) one has $\mathcal{M}_1, x \models p \Leftrightarrow \mathcal{M}_2, y \models p$
+\end{center}
+\end{defin}
+
+\begin{defin}
+  Let $\mathcal{M}_1 = \langle W_1, R_1, \vartheta_1 \rangle$ $\mathcal{M}_2 = \langle W_2, R_2, \vartheta_2 \rangle$ be Kripke models, then a bisimulation between $\mathcal{M}_1$ and $\mathcal{M}_2$ is a binary relation $E \subseteq W_1 \times W_2$ between two models such that:
+  \begin{itemize}
+    \item (0) $x_1 E x_2$ implies $x_1 \equiv_0 x_1$
+    \item ({\bf zig}) $x_1 E y_1$ and $x_1 R_1 z$ implies $y_1 R_2 y_2$ and $z E y_2$ for some $y_2 \in W_2$
+    \item ({\bf zag}) $x_1 E y_1$ and $y_1 R_2 y_2$ implies $x_1 R_1 x_2$ and $x_2 E y_2$ for some $x_2 \in W_1$
+  \end{itemize}
+\end{defin}
+
+Models with designated points $\langle \mathcal{M}_1, x \rangle$ and  $\langle \mathcal{M}_2, y \rangle$ are bisimilar, if there exists a bisimulation $E$ between them with $x E y$.
+We denote that as $\langle \mathcal{M}_1, x \rangle \: \underline{\leftrightarrow} \: \langle \mathcal{M}_2, y \rangle$.
+Moreover, we have
+\begin{center}
+  $\langle \mathcal{M}_1, x \rangle \: \underline{\leftrightarrow} \: \langle \mathcal{M}_2, y \rangle$ implies $Th(\mathcal{M}_1, x) = Th(\mathcal{M}_2, y)$.
+\end{center}
+
+\begin{defin}
+  Let $\langle \mathcal{M}_1, x \rangle$ and $\langle \mathcal{M}_2, y \rangle$ be models with designated points and $r < \omega$.
+The $r$-round bisimulation game $\mathcal{G}(\mathcal{M}_1, \mathcal{M}_2, x, y)$ is played by two players, $\forall$ (Abelard, man) and $\exists$ (Heloise, woman).
+The rules of the game are the following:
+  \begin{enumerate}
+    \item ({\bf Round 0}) $\exists$ wins, if $x \equiv_0 y$. Otherwise, $\forall$ wins.
+    \item ({\bf Round $n+1$}) Let $n + 1 < r$. Suppose rounds $0, \dots, n$ have been played.
+    The $n$-th position is a pair $(x_n, y_n)$.
+    $\forall$ chooses $x_{n + 1}$ (or $y_n \in W_2$) with $x_n R_1 x_{n + 1}$ (or $y_n R_1 y_{n + 1}$).
+$\exists$ chooses $y_{n + 1}$ ($x_{n + 1}$) with $y_n R_1 y_{n + 1}$ (or $y_n R_1 y_{n + 1}$) and $x_{n + 1} \equiv_0 y_{n + 1}$.
+    $\forall$ ($\exists$) wins, if $\exists$ ($\forall$) cannot move.
+  \end{enumerate}
+  $\exists$ wins the game after $r$ rounds. $\forall$ wins the game, if there exists $k < r$ such that $\forall$ wins in {\bf Round $k$}.
+\end{defin}
+
+There are two $n$-equivalences for formulas and games (the first one is due to Fine \cite{fine1974logics}). These definitions are equivalent, this is due to Stirling.
+
+\begin{defin}
+  $ $
+
+  \begin{itemize}
+  \item $(\mathcal{M}_1, x) \equiv_n (\mathcal{M}_2, y)$ iff for every $\varphi$ such that
+  $md(\varphi) \leq n$ we have $\mathcal{M}_1, x \models \varphi$ iff $\mathcal{M}_2, y \models \varphi$.
+  \item $(\mathcal{M}_1, x) \equiv_n (\mathcal{M}_2, y)$ iff $\exists$ has a winning strategy in $\mathcal{G}(\mathcal{M}_1, \mathcal{M}_2, x, y)$.
+  \end{itemize}
+\end{defin}
+
+\begin{lemma} (Stabilisation lemma)
+
+  Let $n$ be a natural number and $\mathcal{L}$ a modal logic.
+  If $\equiv_n = \equiv_{n + 1}$ in a weak canonical model $\mathcal{M}_{\mathcal{L}} \upharpoonright k$, then $md(\mathcal{L} \upharpoonright k) \leq n$.
+\end{lemma}
+
+\begin{theorem} [See \cite{gabbay1998products}]
+  Let $\mathcal{L}$ be a modal logic such that $md(\mathcal{L}) = n$ for some $n < \omega$, then $\mathcal{L}$ is locally finite.
+\end{theorem}
+
 \section{Filtrations for {\bf K5}}
 
-The ${\bf K5}$-closure (an Euclidean Horn closure of a binary relation) has the following equivalent definitions:
+The ${\bf K5}$-closure (the Euclidean Horn closure of a binary relation) has the following equivalent definitions:
 \begin{lemma} \label{equivHorn}
@@
 
-Note that the following property holds not for every ${\bf K5}^{i,j}$, consider ${\bf KB}$ as an example (clearly, ${\bf KB} = {\bf K5}^{0,1}$).
+Note that the following property holds not for every ${\bf K5}^{i,j}$, consider ${\bf KB}$ as an example (clearly, ${\bf KB} = {\bf K5}^{0,1}$), whose modal depth is $\infty$.
 
-\begin{lemma}
-${\bf K5}^{i,j}$ has $n$ non-equivalent sequences of modalities, if . TODO: define $n$.
-\end{lemma}
+\begin{theorem}
+$md({\bf K5}^{i,j}) = 1 + j + i$, if $i \neq 0$.
+\end{theorem}
+
+\begin{proof}
+
+\end{proof}
 
@@
 Let $\Delta$ be any finite and $\operatorname{Sub}$-closed extension of $\Gamma'$ that contains all $\Diamond \Box \psi$ for each $\Box \psi \in \Delta$.
-Then a model $\widehat{M} = \langle W / \sim_{\Delta}, (R^{min}_{\Delta})^{E}, \widehat{\vartheta} \rangle$ is a filtration of $\mathcal{M}$ through $\Delta$.
+Then a model $\widehat{M} = \langle W / \sim_{\Delta}, (R^{min}_{\Delta})^{\bf K5}, \widehat{\vartheta} \rangle$ is a filtration of $\mathcal{M}$ through $\Delta$.
   \item Let $\mathcal{L} = {\bf K} \oplus \Diamond \Diamond \Diamond p \to \Diamond p$, then we have the similar statement for $\mathcal{L}$, where $\Delta$ is an extension of $\Gamma' = \{ \Diamond \Diamond \psi \: | \: \Diamond \psi \}$ that contains all $\Diamond \Diamond \varphi$ for $\Diamond \varphi \in \Delta$.

### 2021-07-10 filtration-stuff/NewText.tex
cylindric background

@@
 \newtheorem{fact}{Fact}
+
+\theoremstyle{exercise}
+\newtheorem{exercise}{Exercise}
 
@@
   Let $\langle \mathcal{M}_1, x \rangle$ and $\langle \mathcal{M}_2, y \rangle$ be models with designated points and $r < \omega$.
-The $r$-round bisimulation game $\mathcal{G}(\mathcal{M}_1, \mathcal{M}_2, x, y)$ is played by two players, $\forall$ (Abelard, man) and $\exists$ (Heloise, woman).
+The $r$-round bisimulation game $\mathcal{G}_r(\mathcal{M}_1, \mathcal{M}_2, x, y)$ is played by two players, $\forall$ (Abelard, man) and $\exists$ (Heloise, woman).
 The rules of the game are the following:
@@
 
-There are two $n$-equivalences for formulas and games (the first one is due to Fine \cite{fine1974logics}). These definitions are equivalent, this is due to Stirling.
+There are two $n$-equivalences for formulas and games (the first one is due to Fine \cite{fine1974logics}).
 
@@
   $md(\varphi) \leq n$ we have $\mathcal{M}_1, x \models \varphi$ iff $\mathcal{M}_2, y \models \varphi$.
-  \item $(\mathcal{M}_1, x) \equiv_n (\mathcal{M}_2, y)$ iff $\exists$ has a winning strategy in $\mathcal{G}(\mathcal{M}_1, \mathcal{M}_2, x, y)$.
+  \item $(\mathcal{M}_1, x) \equiv_n (\mathcal{M}_2, y)$ iff $\exists$ has a winning strategy in $\mathcal{G}_n(\mathcal{M}_1, \mathcal{M}_2, x, y)$.
   \end{itemize}
 \end{defin}
+
+\begin{lemma} [See \cite{goranko20075}] \label{bisimequiv}
+
+  The definitions of $n$-equivalence are equivalent for weak Kripke models.
+\end{lemma}
 
@@
 \begin{lemma} \label{equivHorn}
-  Let $\mathcal{F} = \langle W, R \rangle$ be a Kripke frame and $R \subseteq R^{\bf K5}$.
-  The following conditions are equivalent:
+  Let $\mathcal{F} = \langle W, R \rangle$ be a Kripke frame. The following are equivalent:
 
@@
 \end{lemma}
-
-\begin{proof}
-$ $
-
-\begin{enumerate}
-  \item $(1) \Rightarrow (2)$
-  Let us show that if $R^{E}$ is the smallest Euclidean relation containing $R$, then $R^{E} = \bigcup \limits_{i < \omega} R_i$. There are two inclusions:
-  \begin{itemize}
-    \item $R^{E} \subseteq \bigcup \limits_{i < \omega} R_i$. Recall that $R^{E}$ has the form (?):
-    \begin{center}
-      $R^{E} = \bigcap \{ R' \: | \: R \subseteq R', \forall a, b \in W \: R'(a,b) \Rightarrow \exists x \in W \: R'(x, a) \: \& \: R'(x, b)\}$
-    \end{center}
-
-    \item $\bigcup \limits_{i < \omega} R_i \subseteq R^{E}$.
-    Let us show that $x R_n y$ for each $n < \omega$ implies $x R^E y$ by induction on $n$.
-    If $n = 0$, then $x R y$, thus, $x R^{E} y$, since $R$ is a subrelation of $R^{E}$.
-    Suppose $n = m + 1$ and $x R_{m + 1} y$. Let us show that $x R^{E} y$.
-    From $x R_{m + 1} y$, one has $(x, y) \in R^n \cup (R^{-1}_n \circ R_n)$. There are two cases:
-    \begin{itemize}
-      \item $x R^n y$, one needs to merely apply the IH.
-      \item $x R^{-1}_n \circ R_n y$. Then $\exists z \in W \: x R^{-1}_n z \: \& \: z R_n$. That is,
-      $z R_n x$ and $z R_n y$ for some $z$. $R_n$ is already a subrelation of $R^E$. Thus,
-      $z R^{E} x$ and $z R^{E} y$. That implies $x R^{E} y$.
-    \end{itemize}
-  \end{itemize}
-  \item $(2) \Rightarrow (3)$
-  Let $(x, y) \in R_m$, let us the statement by induction on $m$.
-  \begin{enumerate}
-    \item Suppose $m = 0$, then $x R y$, and the statement is shown putting $n = 0$.
-    \item Suppose $m = p + 1$ and $x R_{p + 1} y$.
-    Assume that either $x R y$ or $\exists z_1, \dots, z_p$ with $z_1 R x$ and $z_{p - 1} R y$ and for each $1 < i \leq p$ one has
-    either $z_{i - 1} R z_i$ or $z_i R z_{i - 1}$.
-
-    $x R_{p + 1} y$ implies $(x, y) \in R_{p} \cup (R^{-1}_p \circ R_p)$. If $(x, y) \in R_{p}$, then we merely apply the IH.
-    Suppose $(x, y) \in R^{-1}_p \circ R_p$, then $(z, x) \in R_p$ and $(z, y) \in R_p$
-  \end{enumerate}
-  \item $(3) \Rightarrow (4)$
-  Suppose either $x R y$ or there exist $n \geq 1$ and $z_1, \dots, z_n$ with $z_1 R x$ and $z_{n - 1} R y$ and for each $1 < i \leq n$ one has either $z_{i - 1} R z_i$ or $z_i R z_{i - 1}$.
-  If $x R y$, then we are done. Otherwise there exists $n \geq 1$ with the condition above. Then $(x, y) \in R_{n + 1}$ that follows from the condition.
-  \item $(4) \Rightarrow (1)$
-\end{enumerate}
-\end{proof}
 
@@
 
+\begin{defin}
 A binary relation $R \subseteq W \times W$ is called $i,j$-Euclidean for $i, j < \omega$, if for each $x, y, z$ such that $x R^{i} y$ and $x R^{j} z$ implies $x R z$.
+\end{defin}
 
@@
   \item $R_{n + 1} = R_n \cup (((R_n)^{-1})^i \circ {R_n}^j)$
-  \item $R^{E_{i,j}} = \bigcup \limits_{k < \omega} R_k$
+  \item $R^{{\bf K}5_{i,j}} = \bigcup \limits_{k < \omega} R_k$
 \end{enumerate}
@@
 
-\begin{theorem}
-$md({\bf K5}^{i,j}) = 1 + j + i$, if $i \neq 0$.
-\end{theorem}
+\begin{exercise}
+$md({\bf K5}) = 2$.
+\end{exercise}
+
+\begin{proof}
+  For the upper bound, let us show that $\equiv_2 = \equiv_3$ in every weak canonical model of ${\bf K5}$, namely $\mathcal{L}_{\bf K5} \upharpoonright k$ for $k < \omega$.
+  For that, we show that if $\exists$ can win $\mathcal{G}_2(\mathcal{L}_{\bf K5} \upharpoonright k, \Gamma, \Delta)$ (where $\Gamma$ and $\Delta$ are maximal consistent ${\bf K5} \upharpoonright k$-theories), then she can the game $\mathcal{G}_2(\mathcal{L}_{\bf K5} \upharpoonright k, \Gamma, \Delta)$ as well.
+
+  If $\exists$ has a winning strategy in $\mathcal{G}_2(\mathcal{L}_{\bf K5} \upharpoonright k, \Gamma_0, \Delta_0)$, then, by Lemma~\ref{bisimequiv}, $\langle \mathcal{L}_{\bf K5} \upharpoonright k, \Gamma_0 \rangle \equiv_2 \langle \mathcal{L}_{\bf K5} \upharpoonright k, \Delta_0 \rangle$.
+
+  Therefore, $\Gamma_0$ and $\Delta_0$ consist of the same formulas whose modal depth is less than or equal to $2$.
+
+  Suppose $(\Gamma_2, \Delta_2)$ is the second position. Clearly $\Gamma_2 \equiv_0 \Delta_2$
+
+  Let us assume that $\forall$ picks $\Gamma_3$ such that $\Gamma_2 R_{{\bf K5} \upharpoonright k} \Gamma_3$. $\exists$ has to find a proper response, namely $\Delta_3$, to win the game.
+
+\end{proof}
+
+\begin{exercise}
+  Let $i = j$, then $md({\bf K5}^{i,i}) = i + 1$. So ${\bf K5}^{i,i}$ is locally finite.
+\end{exercise}
 
@@
 \end{proof}
+
+\begin{theorem}
+$md({\bf K5}^{i,j}) = 1 + j + i$, where $i > 0$.
+\end{theorem}
 
@@
   \end{center}
-  Let $(\hat{x}, \hat{y}) \in R^{E_{i,j}}$, $\mathcal{M}, x \models \Box \psi$ for $\Box \psi$ in $\Delta$.
+  Let $(\hat{x}, \hat{y}) \in R^{{\bf K}5_{i,j}}$, $\mathcal{M}, x \models \Box \psi$ for $\Box \psi$ in $\Delta$.
   If $n = 0$, then the statement is obvious. Suppose $n = 1$.
@@
 \begin{proof}
-  Let $\mathcal{M} = \langle W, R_1, R_2, \vartheta \rangle$ be an $\mathcal{L}_1 * \mathcal{L}_2$.
-  We define a notation $\nabla = \{ \neg \Diamond, \Diamond \neg, \Diamond \}$.
+  Let $\mathcal{M} = \langle W, R_1, R_2, \vartheta \rangle$ be an $\mathcal{L}_1 * \mathcal{L}_2$-model. We define a notation $\nabla = \{ \neg \Diamond, \Diamond \neg, \Diamond \}$.
 

### 2021-07-10 problem-5/Text.tex
cylindric background

@@
 \date{}
-\title{Representable cylindric algebras of dimension $\omega$: the aspects of canonicity}
+\title{Representable cylindric algebras of dimension $\omega$: the aspects of canonicity and axiomatisability}
 
@@
   $\iota : a \mapsto \{ \gamma \in {\bf Uf}(\mathcal{A}) \: | \: a \in \gamma \}$.
-  \item If $i : \mathcal{A} \hookrightarrow \mathcal{B}$, then this embedding might be extented to the embedding
+  \item $i : \mathcal{A} \hookrightarrow \mathcal{B}$ implies
   $i^{+} : \mathcal{A}^{+} \hookrightarrow \mathcal{B}^{+}$
@@
 
-\subsection{(Representable) cylindric algebras and cylindric set algebras}
+\section{Cylindric algebras and cylindric set algebras}
 
-Cylindric algebras provide a generalisation of relation algebras for relations of an arbitrary arity.
-Let $\alpha$ be an ordinal.
-Let $\prescript{\alpha}{}U$ be the set of all functions mapping $\alpha$ to a non-empty set $U$.
-We denote $x(i) = x_i$ for
+ Let $\alpha$ be an ordinal. Let $\prescript{\alpha}{}U$ be the set of all functions mapping $\alpha$ to a non-empty set $U$. We denote $x(i) = x_i$ for
 $x \in \prescript{\alpha}{}U$ and $i < \alpha$.
@@
 
-The following definition is due to \cite{sankappanavar1981course}:
+The following definition is due to, e.g., \cite{sankappanavar1981course}:
 
@@
     with the constants and operations $0 = \emptyset$, $1 = \prescript{\alpha}{}U$, $\cap$, $-$, the diagonal elements $\{ D_{ij} \}_{i, j < \alpha}$, the cylindrifications $\{ C_i \}_{i < \alpha}$.
-A generalised cylindric set algebra of dimension $\alpha$ is a subdirect of cylindric algebras that have dimension $\alpha$
+A generalised cylindric set algebra of dimension $\alpha$ is a subdirect of cylindric algebras that have dimension $\alpha$.
 \end{defin}
@@
     \end{itemize}
-    ${\bf CA}_{\alpha}$ is the class of all cylindric algebras of dimension $\alpha$
+    ${\bf CA}_{\alpha}$ is the class of all cylindric algebras of dimension $\alpha$.
+\end{defin}
+
+\begin{defin}
+  Let $\mathcal{A} = \langle A, +, -, 0, 1, (c_i)_{i < \alpha}, (d_ij)_{i,j < \alpha} \rangle \in {\bf CA}_{\alpha}$ and $\beta < \alpha$, then the $\beta$-reduct of $\mathcal{A}$ is an algebra $\mathcal{A}_{\beta} = \langle A, +, -, 0, 1, (c_i)_{i < \beta}, (d_ij)_{i,j < \beta} \rangle$,
 \end{defin}
@@
 
-Let $\mathcal{A} \in {\bf C}_{\omega}$, then $\mathcal{A}$ has a \emph{complete representation}, if this representation preserves all existing suprema.
+Let $\mathcal{A} \in {\bf CA}_{\omega}$, then $\mathcal{A}$ has a \emph{complete representation}, if this representation preserves all existing suprema.
 In other words, $\mathcal{A}$ is completely representable.
@@
 
-Let us discuss the connection between representability and canonical extensions.
+\subsection{Closed elements}
 
@@
 In particular, an element $x$ is called \emph{zero-dimensional} if $\Delta x = 0$.
-Zero-dimensional elements reflect equivalence classes of sentences in the Lindenbaum-Tarski algebra of a given first-order theory.
+Zero-dimensional elements correspond to equivalence classes of sentences in the Lindenbaum-Tarski algebra of a given first-order theory.
 Thus, the set of zero-dimensional elements form a Boolean algebras of sentences associated with $\Theta$.
@@
 Metamathematically, $\Gamma$-closed elements reflect universal closures (is it correct?).
+
+\subsection{Neat reducts}
 
@@
 
-\begin{defin}
+\begin{defin} \label{network}
   Let $\mathcal{A}$ be a BAO of type ${\bf CA}_{\omega}$, an $\mathcal{A}$-pre-network is a pair
@@
 
-\begin{defin}
-  Let $\mathcal{A}$ be an atomic BAO of type ${\bf CA}_{\omega}$ and $\kappa > 0$ a cardinal.
-The game $\mathcal{G}^{\kappa}(\mathcal{A})$ is defined as follows.
+\begin{defin} Let $n \leq \omega$ an $\mathcal{A}$ be an atomic BAO of type ${\bf CA}_{\omega}$ that has countably atoms.
+The game $\mathcal{G}_{n}(\mathcal{A})$ is defined as follows.
 The game has two players: $\forall$ (Abelard, he/his) and $\exists$ (H\'{e}lo\"{i}se, she/her).
-A play of the game $\mathcal{G}^{\kappa}(\mathcal{A})$ is the sequence of networks
+A play of the game $\mathcal{G}_{n}(\mathcal{A})$ is the sequence of networks
   \begin{center}
-    $\mathcal{N}_0 \subseteq \mathcal{N}_1 \subseteq \mathcal{N}_2 \subseteq \dots \subseteq \mathcal{N}_{\lambda} \subseteq \dots$ for $\lambda < \kappa$
+    $\mathcal{N}_0 \subseteq \mathcal{N}_1 \subseteq \mathcal{N}_2 \subseteq \dots \subseteq \mathcal{N}_{k} \subseteq \dots$ for $k < n$
   \end{center}
@@
 
-    Let $0 < \lambda$ be a cardinal such that $\lambda + 1 < \kappa$ and a network $\mathcal{N}_{\lambda} = \langle N_{\lambda}, l_{\lambda} \rangle$ has been already played.
+    Let $k < n$ such that $k + 1 < \kappa$ and a network $\mathcal{N}_{\lambda} = \langle N_{k}, l_{k} \rangle$ has been already played.
 
-    $\forall$ picks $i < \omega$, $x \in \prescript{\omega}{}N_{\lambda}$, $a \in \operatorname{At}$ such that $l_{\lambda}(x) \leq c_i a$.
+    $\forall$ picks $i \leq k$, $x \in \prescript{\omega}{}N_{k}$, $a \in \operatorname{At}(\mathcal{A})$ such that $l_{k}(x) \leq c_i a$.
     We denote this move as $(i, x, a)$.
-$\exists$ responds with a network $\mathcal{N}_{\lambda + 1} \supseteq \mathcal{N}_{\lambda}$.
+$\exists$ responds with a network $\mathcal{N}_{k + 1} \supseteq \mathcal{N}_{k}$.
 $\forall$ wins, if there is no node $c \in N$ such that $l_{\lambda+1}(x[i/c]) = a$, then $\forall$ wins
-    \item The limit of the play is defined as $\bigcup \limits_{\lambda < \kappa} \mathcal{N}_{\lambda}$.
-$\forall$ wins the play, if there exists $\kappa_1 < \kappa$ such that $\exists$ does not win the $\kappa_1$th-round.
+    \item The limit of the play is defined as $\bigcup \limits_{k < n} \mathcal{N}_{k}$.
+$\forall$ wins the play, if there exists $m < n$ such that $\exists$ does not win the $m$-th round.
 Otherwise, $\exists$ wins the play.
@@
     \item $\mathcal{A}$ is completely representable.
-    \item $\exists$ has a winning strategy in $\mathcal{G}^{\kappa + \omega}$.
+    \item $\exists$ has a winning strategy in $\mathcal{G}_{\kappa + \omega}$.
   \end{enumerate}
@@
 
-    Suppose $\exists$ has a winning strategy in $\mathcal{G}^{\kappa + \omega}(\mathcal{A})$.
+    Suppose $\exists$ has a winning strategy in $\mathcal{G}_{\kappa + \omega}(\mathcal{A})$.
     In every round $\forall$ picks all possible $i < \omega$, $a \in \operatorname{At}(\mathcal{A})$, all possible hyperedges and all appropriate atoms and $\exists$ has a proper response for every $\forall$'s move.
@@
     \begin{center}
-    $rep(a) = \{ x \: | \: \exists b \in \operatorname{At}(\mathcal{A}) \: x \in \prescript{\omega}{}N_a, l_a(x) \leq b \}$.
+    $\operatorname{rep}(a) = \{ x \: | \: \exists b \in \operatorname{At}(\mathcal{A}) \: x \in \prescript{\omega}{}N_a, l_a(x) \leq b \}$.
   \end{center}
@@
     \begin{enumerate}
-    \item Suppose $x \in rep(c_i a)$, then there exists an atom $b$ such that $x \in \prescript{\omega}{}N_b$ with $l_b(x) \leq c_i a$.
-Then there exists $y \equiv_i x$ with $l_b(y) \leq a$, so $x \in {\bf C}_i(rep(a))$.
+    \item Suppose $x \in \operatorname{rep}(c_i a)$, then there exists an atom $b$ such that $x \in \prescript{\omega}{}N_b$ with $l_b(x) \leq c_i a$.
+Then there exists $y \equiv_i x$ with $l_b(y) \leq a$, so $x \in {\bf C}_i(\operatorname{rep}(a))$.
 
-    Let $x \in {\bf C}_i(rep(a))$. We need $x \in (rep(c_i a))$, that is, one needs to find an atom $c$ such that $l_c(x) \leq c_i a$.
+    Let $x \in {\bf C}_i(\operatorname{rep}(a))$. We need $x \in (\operatorname{rep}(c_i a))$, that is, one needs to find an atom $c$ such that $l_c(x) \leq c_i a$.
 
-    We already know that there exists $y \equiv_i x$ such that $y \in rep(a)$, that is, there exists an atom $b$ such that $y \in \prescript{\omega}{}N_b$ and $l_b(y) \leq a$.
+    We already know that there exists $y \equiv_i x$ such that $y \in \operatorname{rep}(a)$, that is, there exists an atom $b$ such that $y \in \prescript{\omega}{}N_b$ and $l_b(y) \leq a$.
 
-    \item If $x \in rep(d_{ij})$, so there exists an atom $b$ with $x \in \prescript{\omega}{}N_a$ and $l_b(x) \leq d_{ij}$, then $x_i = x_j$, then $x \in D_{ij}$.
+    \item If $x \in \operatorname{rep}(d_{ij})$, so there exists an atom $b$ with $x \in \prescript{\omega}{}N_a$ and $l_b(x) \leq d_{ij}$, then $x_i = x_j$, then $x \in D_{ij}$.
     \end{enumerate}
@@
 \end{proof}
+
+If $\mathcal{A} \in {\bf CA}_{n}$ is atomic, then we have the following criterion of complete representability formulated in terms of back-and-forth games, see \cite[Theorem 3.3.3 and Corollary 3.3.5]{hirsch2013completions}.
+Here, the definition of a network for the case of ${\bf CA}_n$ is the same as Definition~\ref{network}, but hyperedges have lenght $n$.
+
+\begin{theorem}
+  Let $\mathcal{A}$ be an atomic BAO of type ${\bf CA}_n$ such that $|\operatorname{At}(\mathcal{A})| = \kappa$, where $\kappa$ is an arbitrary cardinal. Then the following are equivalent:
+
+  \begin{enumerate}
+    \item $\mathcal{A}$ is completely representable iff $\exists$ has a winning strategy in $\mathcal{G}_{\kappa + \omega}(\mathcal{A})$.
+    \item $\exists$ has a winning strategy in $\mathcal{G}_{m}(\mathcal{A})$ for every $m < \omega$ iff $\mathcal{A}$ is elementarily equivalent to a completely representable cylindric algebra.
+  \end{enumerate}
+\end{theorem}
+
+Now we formulate the Lyndon conditions for BAOs of type ${\bf CA}_{\omega}$, but we need their analogues for the finite case which are due to \cite[Theorem 34]{hirsch1997step}.
+Let $\mathcal{A}$ be an atomic BAO of type ${\bf CA}_{n}$.
+$lc_k^n$ is a first-order sentence such that $\mathcal{A} \models lc_k^n$ iff $\exists$ has a winning strategy in the game $\mathcal{G}_k(\mathcal{A})$ of length $k < \omega$.
+
+Consider the following set of formulas:
+\begin{center}
+  $LC_{\omega} = \{ lc_n^n \: | \: n < \omega \}$
+\end{center}
+We shall think of $LC_{\omega}$ as the Lyndon conditions for atomic cylindric algebras of dimension $\omega$.
+Let us write down those $lc_n^n$s explicitly, this is due to \cite[Sections 11.3 and 11.4]{hirsch1997step}.
+For that, we need a bit of definitions.
+
+\begin{proposition}
+  Let $\mathcal{A}$ be a BAO of type ${\bf CA}_{\omega}$ and $m < \omega$, then $\exists$ has a winning strategy in $\mathcal{G}_m(\mathcal{A})$ iff $\mathcal{A}_m \models lc(m,m)$
+\end{proposition}
+
+\begin{proof}
+\end{proof}
+
+Let us recall the definition of an ultraproduct, see \cite[Section 9.5]{hodges1993model}.
+Let $I$ be an index set, $U$ an ultrafiler over $I$, an $(\mathcal{A}_i)_{i \in I}$ a family of structures of the same signature.
+The ultraproduct of $(\mathcal{A}_i)_{i \in I}$ modulo $U$, denoted as $\Pi_U \mathcal{A}_i$, is an algebra of the same signature whose elements are elements of the direct product
+$\Pi_{i \in I} \mathcal{A}_i$ factorised through an equivalence relation $\sim$ such that for every $a, b \in \Pi_{i \in I} \mathcal{A}_i$:
+\begin{center}
+  $a \sim b \Leftrightarrow \{ i \in I \: | \: a(i) = b(i)\} \in U$
+\end{center}
+
+\begin{lemma}
+  Let $\mathcal{A}$ be an atomic BAO of type ${\bf CA}_{\omega}$ such that for every $n < \omega$ $\mathcal{A}_n \models lc(n,n)$, then $\exists$ has a winning strategy in $\mathcal{G}_{\omega}(\Pi_U \mathcal{A})$, where $\Pi_U \mathcal{A}$ is a non-principal ultrapower of $\mathcal{A}$ and $U$ is a non-principal ultrafilter over $\omega$.
+\end{lemma}
+
+\begin{proof}
+  Suppose $\mathcal{A}_n \models lc(n,n)$ for each $n$.
+\end{proof}
+
+\begin{theorem}
+  What's going on? :(
+\end{theorem}
 
@@
 
-  The argument uses \L o\'s's Theorem, see \cite[Theorem 8.5.3]{hodges1997shorter}.
+  The argument uses \L o\'s's Theorem, see \cite[Theorem 9.5.1]{hodges1993model}.
 
@@
 \end{proof}
+
+\begin{theorem}
+  The elementary closure of the class of completely representable cylindric algebras of dimension $\omega$ is axiomatised with the axioms of ${\bf CA}_{\omega}$ and the Lyndon conditions.
+\end{theorem}
 
@@
 
-  Consider $\rho(\mathcal{A})$. Let us show that $\exists$ has a winning strategy on $\mathcal{G}^{\kappa + \omega}(\rho(\mathcal{A}))$
+  Consider $\rho(\mathcal{A})$. Let us show that $\exists$ has a winning strategy on $\mathcal{G}_{\kappa + \omega}(\rho(\mathcal{A}))$
 \end{proof}

### 2021-07-12 filtration-stuff/NewText.tex
fix lemma for fusions

@@
 Consider $\Gamma' = \Gamma \cup \{ \Diamond \Box \psi \: | \: \Box \psi \in \Gamma \}$.
-Let $\Delta$ be any finite and $\operatorname{Sub}$-closed extension of $\Gamma'$ that contains all $\Diamond \Box \psi$ for each $\Box \psi \in \Delta$.
-Then a model $\widehat{M} = \langle W / \sim_{\Delta}, (R^{min}_{\Delta})^{\bf K5}, \widehat{\vartheta} \rangle$ is a filtration of $\mathcal{M}$ through $\Delta$.
-  \item Let $\mathcal{L} = {\bf K} \oplus \Diamond \Diamond \Diamond p \to \Diamond p$, then we have the similar statement for $\mathcal{L}$, where $\Delta$ is an extension of $\Gamma' = \{ \Diamond \Diamond \psi \: | \: \Diamond \psi \}$ that contains all $\Diamond \Diamond \varphi$ for $\Diamond \varphi \in \Delta$.
+Let $\Delta$ be any finite and $\operatorname{Sub}$-closed extension of $\Gamma'$ tThen a model $\widehat{M} = \langle W / \sim_{\Delta}, (R^{min}_{\Gamma'})^{\bf K5}, \widehat{\vartheta} \rangle$ is a filtration of $\mathcal{M}$ through $\Gamma'$.
+  \item Let $\mathcal{L} = {\bf K} \oplus \Diamond \Diamond \Diamond p \to \Diamond p$, then we have the similar statement for $\mathcal{L}$, where $\Delta$ is a $\operatorname{Sub}$-closed extension extension of $\Gamma'$
   \end{enumerate}
@@
 
-  \item Recall that $(R^{min}_{\Delta})^{E}$ is defined inductively as:
+  \item Recall that $(R^{min}_{\Delta})^{\bf K5}$ is defined inductively as:
   \begin{enumerate}
-    \item $R_{\Delta}^{0} = R_{\Delta}^{min}$
-    \item $R_{\Delta}^{n + 1} = R_{\Delta}^{n} \cup ({R_{\Delta}^{n}}^{-1} \circ R_{\Delta}^{n})$
-    \item $(R^{min}_{\Delta})^{E} = \bigcup_{k < \omega} R_{\Delta}^{k}$
+    \item $R_{\Gamma'}^{0} = R_{\Gamma'}^{min}$
+    \item $R_{\Gamma'}^{n + 1} = R_{\Gamma'}^{n} \cup ({R_{\Gamma'}^{n}}^{-1} \circ R_{\Gamma'}^{n})$
+    \item $(R^{min}_{\Gamma'})^{E} = \bigcup_{k < \omega} R_{\Gamma'}^{k}$
   \end{enumerate}
 
-  If $R^{min}_{\Delta}$ is already a subrelation of $R^{max}$, so the base case is self-evident.
+  If $R^{min}_{\Gamma'}$ is already a subrelation of $R^{max}$, so the base case is self-evident.
 
-  Suppose the statement holds for $R_{\Delta}^{n}$, $(\hat{x},\hat{y}) \in R_{\Delta}^{n + 1}$ such that $\mathcal{M}, y \models \psi$ for $\Diamond \psi \in \Delta$.
+  Suppose the statement holds for $R_{\Gamma'}^{n}$, $(\hat{x},\hat{y}) \in R_{\Gamma'}^{n + 1}$ such that $\mathcal{M}, y \models \psi$ for $\Diamond \psi \in \Gamma'$.
 
   According to the third item of Lemma~\ref{equivHorn}, this is the same as there exist $\widehat{z_0}, \widehat{z_1}, \dots, \widehat{z_{n-1}}, \widehat{z_n}$ such that
-  $\widehat{z_1} R^{min}_{\Delta} \hat{x}$, $\widehat{z_n} R^{min}_{\Delta} \hat{y}$, and for each $i \in n + 1$ we have either $\widehat{z_i} R^{min}_{\Delta} \widehat{z_{i + 1}}$ or $\widehat{z_{i + 1}} R^{min}_{\Delta} \widehat{z_i}$.
+  $\widehat{z_1} R^{min}_{\Gamma'} \hat{x}$, $\widehat{z_n} R^{min}_{\Gamma'} \hat{y}$, and for each $i \in n + 1$ we have either $\widehat{z_i} R^{min}_{\Gamma'} \widehat{z_{i + 1}}$ or $\widehat{z_{i + 1}} R^{min}_{\Delta} \widehat{z_i}$.
 
@@
   \xymatrix{
-  &&& \hat{x} & \ar[l]_{R_{\Delta}^{min}} \ar@{<->}[r]^{R'} \widehat{z_0} & \ar@{<->}[r]^{R'} \widehat{z_1} \ar@{<->}[r]^{R'} & \dots \ar@{<->}[r]^{R'} & \widehat{z_{n - 1}} \ar[r]^{R'} & \widehat{z_n} \ar[r]^{R_{\Delta}^{min}} & \hat{y}
+  &&& \hat{x} & \ar[l]_{R_{\Gamma'}^{min}} \ar@{<->}[r]^{R'} \widehat{z_0} & \ar@{<->}[r]^{R'} \widehat{z_1} \ar@{<->}[r]^{R'} & \dots \ar@{<->}[r]^{R'} & \widehat{z_{n - 1}} \ar[r]^{R'} & \widehat{z_n} \ar[r]^{R_{\Gamma'}^{min}} & \hat{y}
   }
-  where $R'$ is either $R_{\Delta}^{min}$ or its converse.
+  where $R'$ is either $R_{\Gamma'}^{min}$ or its converse.
 We have $\mathcal{M}, z \models \Box \psi$, $\mathcal{M}, y \models \psi$, so $\mathcal{M}, z_{n} \models \Diamond \psi$.
@@
   \end{center}
-  So we have $\mathcal{M}, z_0 \models \Box \Diamond \varphi$. Note that we always stay within $\Delta$. Thus, $\mathcal{M}, x \models \Diamond \varphi$.
+  So we have $\mathcal{M}, z_0 \models \Box \Diamond \varphi$. Note that we always stay within $\Gamma'$. Thus, $\mathcal{M}, x \models \Diamond \varphi$.
 
   \item Let us prove the second item. Let $\mathcal{M}$ be an $\mathcal{L}$-model.
-  The $\mathcal{L}$-closure of the minimal filtered relation modulo $\Delta$, namely ${R_{\Delta}^{min}}^{\mathcal{L}}$ has the following form:
+  The $\mathcal{L}$-closure of the minimal filtered relation modulo $\Gamma'$, namely ${R_{\Gamma'}^{min}}^{\mathcal{L}}$ has the following form:
   \begin{center}
-    ${R_{\Delta}^{min}}^{\mathcal{L}} = \bigcup \limits_{k < \omega} {R_{\Delta}^{min}}^{2k + 1}$
+    ${R_{\Gamma'}^{min}}^{\mathcal{L}} = \bigcup \limits_{k < \omega} {R_{\Gamma'}^{min}}^{2k + 1}$
   \end{center}
@@
   \begin{enumerate}
-    \item $R_0 = R_{\Delta}^{min}$
+    \item $R_0 = R_{\Gamma'}^{min}$
     \item $R_{n + 1} = R_n \cup ((R_{\Delta}^{min})^2 \circ R_n)$
-    \item ${R_{\Delta}^{min}}^{\mathcal{L}} = \bigcup \limits_{k < \omega}$
+    \item ${R_{\Gamma'}^{min}}^{\mathcal{L}} = \bigcup \limits_{k < \omega}$
   \end{enumerate}
@@
   The base case is self-evident.
-Suppose the statement holds for $R_{n}$ and $(\hat{x}, \hat{y}) \in (R_{\Delta}^{min})^2 \circ R_n$, that is, there exists $\hat{z}$ such that $(\hat{x}, \hat{z}) \in (R_{\Delta}^{min})^2$ and $(\hat{z}, \hat{y}) \in R_n$.
+Suppose the statement holds for $R_{n}$ and $(\hat{x}, \hat{y}) \in (R_{\Gamma'}^{min})^2 \circ R_n$, that is, there exists $\hat{z}$ such that $(\hat{x}, \hat{z}) \in (R_{\Gamma'}^{min})^2$ and $(\hat{z}, \hat{y}) \in R_n$.
 By IH, we have $\mathcal{M}, z \models \Diamond \varphi$.
@@
   \xymatrix{
-  &&&&&& \hat{x} \ar[r]^{R_{\Delta}^{min}} & \widehat{y_1} \ar[r]^{R_{\Delta}^{min}} & \hat{z}
+  &&&&&& \hat{x} \ar[r]^{R_{\Delta}^{min}} & \widehat{y_1} \ar[r]^{R_{\Gamma'}^{min}} & \hat{z}
   }

### 2021-11-08 cylindric-notes/Text.tex
cylindric notes, wip

--- 2021-06-26 problem-5/Text.tex
@@
 \date{}
-\title{Representable cylindric algebras of dimension $\omega$: the aspects of canonicity}
+\title{Representable cylindric algebras of dimension $\omega$}
 
 \begin{document}
-
 \maketitle
-
-\section{Intro}
 
@@
 
-\section{Boolean algebras with operators and cylindric algebras, a bit of the backgroud}
+\section{Atomic Representations}
+
+A representation of a Boolean algebra $\mathcal{B}$ is an embedding $h$ of $\mathcal{B}$ to some field of sets.
 
@@
 representation of $\mathcal{B}$, then $\mathcal{B}$ is a complete representation of $\mathcal{B}$, if for every
-$A \subseteq \mathcal{B}$ whenever $\Sigma A$ we have the following:
+$A \subseteq \mathcal{B}$ we have the following whenever $\Sigma A$ is defined:
 \begin{center}
@@
 
+A representation $h$ is called atomic, if $x \in h(1)$ there exists $b \in \operatorname{At}(\mathcal{B})$ such that $x \in h(b)$.
+
 \begin{theorem} \label{completeboolean}
@@
 \end{theorem}
+
+\section{BAOs and Duality}
+
+By default, we assume that all operators are at most unary. Here is the rigorous definition:
 
@@
   \begin{enumerate}
-    \item Let $\mathcal{B} = \langle B, +, -, 0, 1 \rangle$ be a Boolean algebra. An operator is an $n$-ary function $\Omega : B^n \to B$ satisfying the following conditions:
+    \item Let $\mathcal{B} = \langle B, +, -, 0, 1 \rangle$ be a Boolean algebra. An operator is a function $\Omega : B \to B$ satisfying the following conditions:
     \begin{itemize}
-      \item Normality: for all $b_0, \dots, b_{n - 1} \in B$, if $b_1 = 0$ for some $i < n$, then
-
-\begin{center}
-      $\Omega(b_0, \dots, b_{i - 1}, 0, b_{i+1}, \dots, b_{n - 1}) = 0$
-\end{center}
-      \item Additivity: for all $b_0, \dots, b_{n - 1}, b, b' \in B$ we have
-      \begin{center}
-        $\Omega(b_0, \dots, b_{i - 1}, (b + b'), b_{i+1}, \dots, b_{n - 1}) = \Omega(b_0, \dots, b_{i - 1}, b, b_{i+1}, \dots, b_{n - 1}) + \Omega(b_0, \dots, b_{i - 1}, b', b_{i+1}, \dots, b_{n - 1})$
-      \end{center}
+      \item Normality: $\Omega(0) = 0$
+      \item Additivity: $\Omega(b + b') = \Omega(b) + \Omega(b')$
     \end{itemize}
-    \item Let $I$ be an index set, a Boolean algebra with operators (BAO) is an algebra $\langle B, +, -, 0, 1, \{ \Omega_{i} \}_{i \in I} \rangle$ such that $\langle B, +, -, 0, 1 \rangle$ is a Boolean algebra and for each $i \in I$ $\Omega_{i}$ is an operator.
+    \item Let $I$ be an index set, a Boolean algebra with operators (BAO) is an algebra $\langle B, +, -, 0, 1, (\Omega_{i})_{i \in I} \rangle$ such that $\langle B, +, -, 0, 1 \rangle$ is a Boolean algebra and for each $i \in I$ $\Omega_{i}$ is an operator.
   \end{enumerate}
@@
 
-\begin{defin} Let $\mathcal{B} = \langle B, +, -, 0, 1, \{ \Omega_{i} \}_{i \in I} \rangle$ be a BAO, then
+\begin{defin} Let $\mathcal{B} = \langle B, +, -, 0, 1, (\Omega_{i})_{i \in I} \rangle$ be a BAO, then
 
   \begin{enumerate}
-    \item An opetator $\Omega$ is completely additive, if for each $b_0, \dots, b_{n - 1} \in B$ and $X \subseteq B$, one has
+    \item An opetator $\Omega$ is completely additive, if for every $X \subseteq B$ such that $\Sigma X$ is defined, one has
 
     \begin{center}
-      $\Omega(b_0, \dots, b_{i-1}, \sum X, b_{i+1}, \dots, b_{n - 1}) = \sum \limits_{x \in X} \Omega(b_0, \dots, b_{i-1}, x, b_{i+1}, \dots, b_{n - 1})$
+      $\Omega(\sum X) = \sum \limits_{x \in X} \Omega(x)$
     \end{center}
@@
 
-\begin{defin}  Let $I$ be an index set and $\{ \Omega_i\}_{i \in I}$ a set of function symbols
+\begin{defin}  Let $I$ be an index set and $(\Omega_i)_{i \in I}$ a set of function symbols
 \begin{enumerate}
-  \item An atom structure is a relational structrure
-  $\mathcal{F} = \langle W, \{ R_{i} \}_{i \in I} \rangle$
-  such that $R_{i}$ is a $n+1$-ary relation symbol, where $\Omega_{i \in I}$ is an $n$-ary function symbol,
+  \item A structure is a relational structrure
+  $\mathcal{F} = \langle W, (R_{i})_{i \in I} \rangle$
+  such that $R_{i}$ is a binary relation symbol for a function symbol $\Omega_{i \in I}$ with the corresponding index,
   \item Let $\mathcal{B}$ be an atomic BAO of the signature $I$,
-the atom structure of $\mathcal{B}$, written as $\mathfrak{At} \mathcal{B}$, is an atom structure $\langle \operatorname{At}(\mathcal{B}), \{ R_{i}\}_{i \in I} \rangle$ such that for each
-$a, b_0, \dots, b_{n + 1} \in \operatorname{At}(\mathcal{B})$ and for each $i \in I$
+the atom structure of $\mathcal{B}$, written as $\mathfrak{At} \mathcal{B}$, is a structure $\langle \operatorname{At}(\mathcal{B}), (R_{i})_{i \in I} \rangle$ such that for all
+$a, b \in \operatorname{At}(\mathcal{B})$ and for all $i \in I$
 \begin{center}
-  $\mathfrak{At} \mathcal{B} \models R_{i}(a,  b_0, \dots, b_{n + 1})$ iff $\mathcal{B} \models a \leq \Omega_{i}(b_0, \dots, b_{n + 1})$
+  $\mathfrak{At} \mathcal{B} \models R_{i}(a,b)$ iff $\mathcal{B} \models a \leq \Omega_{i}(b)$
 \end{center}
-\item Let $\mathcal{F} = \langle W, \{ R_{i} \}_{i \in I} \rangle$ be an atom structure, the complex algebra of $\mathcal{F}$, written as $\bf{Cm} \mathcal{F}$, is a BAO
-$\langle \mathcal{P}(W), \cup, -, \emptyset, W, \{\Omega_{R_{i}}\}_{i \in I} \rangle$ such that
-for all $X_0, \dots, X_{n - 1} \subseteq W$ and for each $i \in I$
+\item Let $\mathcal{F} = \langle W, (R_{i})_{i \in I} \rangle$ be an atom structure, the complex algebra of $\mathcal{F}$, written as $\mathfrak{Cm} \mathcal{F}$, is a BAO
+$\langle \mathcal{P}(W), \cup, -, \emptyset, W, (\Omega_{R_{i}})_{i \in I} \rangle$ such that
+for all $X \subseteq W$ and for each $i \in I$:
 \begin{center}
-  $\Omega_{R_{i}}(X_0, \dots, X_{n - 1}) = \{ a \in W \: | \: \exists b_0 \in X_0 \dots \exists b_{n -1} \in X_{n - 1} \: \mathcal{F} \models R_{i}(a, b_0, \dots, b_{n - 1})\}$
+  $\Omega_{R_{i}}(X) = \{ a \in W \: | \: \exists b \in X \: \mathcal{F} \models R_{i}(a, b)\}$
 \end{center}
@@
 
-The following duality is due to Thomason \cite{thomason1975categories}.
+\begin{defin} Let $\mathcal{F} = \langle W, (R_{i})_{i \in I} \rangle$ and $\mathcal{F}' = \langle W', ({R'}_{i})_{i \in I} \rangle$, then a function $f : \mathcal{F} \to \mathcal{F}'$ is a bounded morphism, if the following holds:
+\begin{enumerate}
+\item $x R_i y$ implies $f(x) {R'}_i f(y)$;
+\item $f(x) {R'}_i z$, then there exists $y \in W$ such that $x R_i y$ and $f(y) = z$.
+\end{enumerate}
+A bounded morphism $f : \mathcal{F} \to \mathcal{F}'$ is a $p$-morphism, if $f$ is onto.
+$\mathcal{F} \twoheadrightarrow \mathcal{F}'$ iff there exists a $p$-morphism from $\mathcal{F}$ onto $\mathcal{F}'$, or $\mathcal{F}'$ is a $p$-morphic image of $\mathcal{F}$.
+\end{defin}
 
-\begin{fact}
-  $ $
+\begin{defin} Let $\mathcal{F} = \langle W, (R_{i})_{i \in I} \rangle$ is an inner substructure \footnote{Or alternatively, a generated subframe} of $\mathcal{F}' = \langle W', ({R'}_{i})_{i \in I} \rangle$, if $W \subseteq W'$ and the embedding $\mathcal{F} \hookrightarrow \mathcal{F}'$ is a bounded morphism.
+\end{defin}
 
+Let $\mathbb{F}$ be a class of structures, define:
 \begin{enumerate}
-    \item Let $\mathcal{B}$ be a complete atomic BAO, then $\mathcal{B} \cong \mathfrak{Cm} (\bf{At}(\mathcal{B}))$,
-    \item Let $\mathcal{F}$ be an atom structure, then $\mathcal{F} \cong \mathfrak{At}(\mathfrak{Cm}(\mathcal{B}))$.
+\item $\mathfrak{Cm}(\mathbb{F}) = \{ \mathcal{B} \: | \: \mathcal{B} \cong \mathfrak{Cm}(\mathcal{F}) \text{ for some $\mathcal{F} \in {\bf F}$}\}$.
+\item ${\bf Up}(\mathbb{F})$ is the class of structures isomorphic to disjoint unions of elements of $\mathbb{F}$.
+\item ${\bf S}(\mathbb{F})$ is the closure of $\mathbb{F}$ under inner substructures.
 \end{enumerate}
-\end{fact}
 
-Let $A$ be a non-empty subset of a Boolean algebra $\mathcal{B}$, $A$ is a \emph{filter}, if $A$ is closed under finite infima and upward closed.
+Let $A$ be a non-empty subset of a Boolean algebra $\mathcal{B}$, $A$ is a \emph{filter}, if $A$ is closed under finite infima and it is upward closed.
 $A$ is an ultrafilter, if it has no non-trivial extensions.
@@
 \begin{defin}
-  Let $\mathcal{B} = \langle B, +, -, 0, 1, \{ \Omega_i \}_{i \in I} \rangle$ be a BAO and ${\bf Uf}(\mathcal{B})$ the set of its ultrafilters.
-The ultrafilter frame of $\mathcal{B}$ (or canonical frame) is a relational structure $\mathcal{F}_{\mathcal{B}} = \langle {\bf Uf}(\mathcal{B}), R_{\Omega_i} \rangle$ such that for each ultrafilters $\beta_0, \dots, \beta_{n - 1}, \gamma$ one has
+  Let $\mathcal{B} = \langle B, +, -, 0, 1, (\Omega_i)_{i \in I} \rangle$ be a BAO and ${\bf Spec}(\mathcal{B})$ the set of its ultrafilters.
+The ultrafilter frame of $\mathcal{B}$ (or the canonical frame) is a relational structure $\mathcal{F}_{\mathcal{B}} = \langle {\bf Spec}(\mathcal{B}), R_{\Omega_i} \rangle$ such that for all ultrafilters $U_1, U_2$ one has
   \begin{center}
-    ${\bf Uf}(\mathcal{B}) \models R_{\Omega_i}(\beta_0, \dots, \beta_{n - 1}, \gamma)$ iff $\{ \Omega(b_0, \dots, b_{n - 1}) \: | \: b_0 \in \beta_0, \dots, b_{n - 1} \in \beta_{n - 1}\} \subseteq \gamma$.
+    ${\bf Spec}(\mathcal{B}) \models R_{\Omega_i}(U_1, U_2)$ iff $\{ \Omega_i(b) \: | \: b \in U_1 \} \subseteq U_2$.
   \end{center}
@@
 
-\begin{defin} Let $\mathcal{B}$ be a BAO, then
-
-  \begin{enumerate}
-    \item The canonical extension of $\mathcal{B}$ is a complex algebra of the canonical frame $\mathfrak{Cm}(\mathcal{F}_{\mathcal{B}})$ denoted as $\mathcal{B}^{+}$,
-    \item The class of BAOs is canonical, if it is closed under canonical extensions.
-  \end{enumerate}
-\end{defin}
+Given $\mathcal{B}$ be a BAO, we denoted as $\mathcal{B}^{+}$ as the complex algebra of the canonical frame $\mathfrak{Cm}(\mathcal{F}_{\mathcal{B}})$, that is, \emph{the canonical extension} of $\mathcal{B}$.
+A class of BAOs ${\bf K}$ is canonical, if it is closed under canonical extensions. That is, $\mathcal{B}^{+} \in {\bf K}$ whenever $\mathcal{B} \in {\bf K}$.
 
@@
   \item There exists $\iota : \mathcal{A} \hookrightarrow \mathcal{A}^{+}$ such that
-  $\iota : a \mapsto \{ \gamma \in {\bf Uf}(\mathcal{A}) \: | \: a \in \gamma \}$.
-  \item If $i : \mathcal{A} \hookrightarrow \mathcal{B}$, then this embedding might be extented to the embedding
+  $\iota : a \mapsto \{ \gamma \in {\bf Spec}(\mathcal{A}) \: | \: a \in \gamma \}$.
+  \item $i : \mathcal{A} \hookrightarrow \mathcal{B}$ implies
   $i^{+} : \mathcal{A}^{+} \hookrightarrow \mathcal{B}^{+}$
@@
 
-\begin{fact}
-\end{fact}
+\section{Representable cylindric algebras}
 
-\subsection{(Representable) cylindric algebras and cylindric set algebras}
-
-Cylindric algebras provide a generalisation of relation algebras for relations of an arbitrary arity.
-Let $\alpha$ be an ordinal.
-Let $\prescript{\alpha}{}U$ be the set of all functions mapping $\alpha$ to a non-empty set $U$.
-We denote $x(i) = x_i$ for
+Let $\alpha$ be an ordinal. Let $\prescript{\alpha}{}U$ be the set of all functions mapping $\alpha$ to a non-empty set $U$. We denote $x(i) = x_i$ for
 $x \in \prescript{\alpha}{}U$ and $i < \alpha$.
@@
 
-The following definition is due to \cite{sankappanavar1981course}:
-
-\begin{defin}
-  Let $(\mathcal{A}_i)_{i \in I}$ be a family of algebras (of an abstract signature) and $\mathcal{A}$ is a subalgebra of $\prod \limits_{i \in I} \mathcal{A}_i$, then $\mathcal{A}$ is a subdirect product, if every projection is onto.
-That is, for every $i \in I$, $\pi_i[\mathcal{A}] = \mathcal{A}_i$.
-\end{defin}
-
 \begin{defin}
  A cylindic set algebra of dimension $\alpha$ is an algebra consisting of a set $S$ of $\alpha$-ry relation on some base set $U$
-    with the constants and operations $0 = \emptyset$, $1 = \prescript{\alpha}{}U$, $\cap$, $-$, the diagonal elements $\{ D_{ij} \}_{i, j < \alpha}$, the cylindrifications $\{ C_i \}_{i < \alpha}$.
-A generalised cylindric set algebra of dimension $\alpha$ is a subdirect of cylindric algebras that have dimension $\alpha$
+   with the constants and operations $0 = \emptyset$, $1 = \prescript{\alpha}{}U$, $\cap$, $-$, the diagonal elements $(D_{ij})_{i, j < \alpha}$, the cylindrifications $(C_i)_{i < \alpha}$.
+A generalised cylindric set algebra of dimension $\alpha$ is a subdirect of cylindric algebras that have dimension $\alpha$.
 \end{defin}
@@
      \item $\mathcal{B}$ is a Boolean algebra, for each $i, j < \alpha$ $c_i$ is an operator and $d_{ij} \in \mathcal{B}$
-      \item For each $i < \alpha$, $a \leq c_i a$, $c_i (a \land c_i b) = c_i a \land c_i b$ and $d_{ii} = 1$
+     \item For each $i < \alpha$, $a \leq c_i a$, $c_i (a \cdot c_i b) = c_i a \cdot c_i b$ and $d_{ii} = 1$
      \item For every $i, j < \alpha$, $c_i c_j a = c_j c_i a$
-      \item If $k \neq i, j < \alpha$, then $d_{ij} = c_k (d_{ij} \land d_{jk})$
-      \item If $i \neq j$, then $c_i (d_{ij} \land a) \land c_i (d_{ij} \land - a) = 0$
+     \item If $k \neq i, j < \alpha$, then $d_{ij} = c_k (d_{ij} \cdot d_{jk})$
+     \item If $i \neq j$, then $c_i (d_{ij} \cdot a) \cdot c_i (d_{ij} \cdot - a) = 0$
    \end{itemize}
-    ${\bf CA}_{\alpha}$ is the class of all cylindric algebras of dimension $\alpha$
+   ${\bf CA}_{\alpha}$ is the class of all cylindric algebras of dimension $\alpha$.
 \end{defin}
@@
 
-An $\alpha$-dimensional cylindric algebra $C$ is representable, if it is isomorphic to a generalised cylindric set algebra
-of dimension $\alpha$.
-Such is isomorphism is a representation of $C$.
-${\bf RCA}_{\alpha}$ is the class of all representable cylindric algebras that have dimension $\alpha$.
-In particular, we are interested in the case when $\alpha = \omega$.
-
-\begin{defin}
-  Given a cylindric algebra of dimension $\alpha$ $C$, let $x$ be a term of its signature, the substitution operator $s^{i}_{j}$ have the following definition:
-  \begin{center}
-  $s^{i}_{j} x = \begin{cases} x, \text{if } i = j \\ c_i (d_{ij} \land x), \text{otherwise} \end{cases}$
-  \end{center}
-\end{defin}
+An $\alpha$-dimensional cylindric algebra $C$ is representable, if there exists a representation of $h$.
+${\bf RCA}_{\alpha}$ is the class of all representable cylindric algebras that have dimension $\alpha$. In particular, we are interested in the case $\alpha = \omega$.
 
@@
 
-Let $\mathcal{A} \in {\bf C}_{\omega}$, then $\mathcal{A}$ has a \emph{complete representation}, if this representation preserves all existing suprema.
+Let $\mathcal{A} \in {\bf CA}_{\omega}$, then $\mathcal{A}$ has a \emph{complete representation}, if its representation preserves all existing suprema.
 In other words, $\mathcal{A}$ is completely representable.
 
-Let us concretise the definition of a canonical extension for ${\bf CA}_{\alpha}$-type BAOs.
+\section{${\bf RCA}_{\omega}$ and canonicity}
 
+The following definition of an $\omega$-frame is due to \cite{Venema2013}.
 \begin{defin}
-  Let $\mathcal{C} = \langle C, +, -, 0, 1, \{ d_{ij} \}_{i, j < \alpha}, \{c_i\}_{i < \alpha} \rangle$ $\mathcal{A}$ be a BAO of type ${\bf CA}_{\alpha}$ Let ${\bf Uf}(\mathcal{C})$ be the set of all ultrafilters of $\mathfrak{B}\mathcal{C}$, the Boolean part of $\mathcal{C}$.
-
-  Let us define ${\bf C}_{i} : {\bf Uf}(\mathcal{C}) \to {\bf Uf}(\mathcal{C})$ for each $i, j < \alpha$ as
+  A cylindric $\omega$-frame is a structure $\mathcal{F} = \langle W, (R_i)_{i < \omega}, (E_{ij})_{i, j < \omega} \rangle$ where $(R_i)_{i < \omega}$ are binary relations and $(E_ij)_{i, j < \omega}$ are unary relations such that, for all $i, j, k < \omega$:
   \begin{enumerate}
-\item ${\bf C}_{i} \mathcal{X} = \{ \mathcal{F} \in {\bf Uf}(\mathcal{C}) \: | \: \exists \mathcal{F}' \in {\bf Uf}(\mathcal{C}) \: (a \in \mathcal{F} \Rightarrow c_i a \in \mathcal{F}' R)\}$,
-\item $D_{ij} = \{ \mathcal{F} \in {\bf Uf}(\mathcal{C}) \: | \: d_{ij} \in \mathcal{F} \}$.
-  \end{enumerate}
-  The structure $\mathcal{C}^{+} = \langle  {\bf Uf}(\mathcal{C}), \cup, -, \emptyset, C, {\bf C}_{i < \alpha}, \{D_{ij}\}_{i, j < \alpha} \rangle$ is called the canonical extension of $\mathcal{C}$.
-\end{defin}
-
-Let us discuss the connection between representability and canonical extensions.
-
-The following definitions and facts are due to Henkin, Monk, and Tarski \cite{henkin1971cylindric}.
-
-Let $\mathcal{A} \in {\bf CA}_{\alpha}$ and $x \in \mathcal{A}$. Recall that \emph{the dimension of $x$}
-is the set of all ordinals $\gamma < \alpha$ such that $c_\gamma x \neq x$. More formally,
-\begin{center}
-  $\Delta x = \{ \gamma \: | \: \gamma < \alpha \: \& \: c_\gamma x \neq x\}$
-\end{center}
-
-Let us discuss some metamathematical intuitions standing behind the notion of a dimension.
-Let $\Theta$ be a first-order theory and $\mathcal{C}/\equiv_{\Theta}$ its Lindenbaum-Tarski algebra.
-Let $\varphi$ be a formula in the signature of $\Theta$.
-Then $\Delta(\varphi/\Theta)$ consists of all $\kappa < \alpha$ such that $\exists x_{\kappa} \varphi \leftrightarrow \varphi$ is not valid in $\Theta$.
-That is, $\Delta(\varphi/\Theta)$ contains ordinals $\kappa$ for which $x_{\kappa}$ is free in $\varphi$.
-Moreover, $\Delta(\varphi/\Theta)$ consists only of those ordinals for which $x_{\kappa}$ is free in every $\psi \in \varphi/\Theta$.
-
-In particular, an element $x$ is called \emph{zero-dimensional} if $\Delta x = 0$.
-Zero-dimensional elements reflect equivalence classes of sentences in the Lindenbaum-Tarski algebra of a given first-order theory.
-Thus, the set of zero-dimensional elements form a Boolean algebras of sentences associated with $\Theta$.
-
-\begin{defin}
-  Let $\mathcal{A}$ be an $\alpha$-dimensional cylindric algebra.
-Let $\alpha$ be an ordinal and $\Gamma$ a subset $\alpha$, then an element $x \in \mathcal{A}$ is $\Gamma$-closed if $\Delta x \cdot \Gamma = \emptyset$.
-Alternatively, $x$ is a $\Gamma$-cylinder.
-
- $\operatorname{Cl}_{\Gamma}\mathcal{A}$ is the set of all $\Gamma$-closed elements.
-\end{defin}
-
-Metamathematically, $\Gamma$-closed elements reflect universal closures (is it correct?).
-
-Let $\mathcal{C} = \langle C, +, -, 0, 1, \{d_{ij}\}_{i, j < \beta}, \{ c \}_{c < \beta} \rangle$ be a $\beta$-dimensional cylindic algebra and $\alpha \leq \beta$ an ordinal.
-The \emph{$\alpha$-th reduct} of $\mathcal{C}$, denoted as $\mathfrak{R}\mathfrak{d}_{\alpha}\mathcal{C}$, is an algebra having the form
-\begin{center}
-$\mathfrak{R}\mathfrak{d}_{\alpha}\mathcal{C} = \langle C, +, -, 0, 1, \{d_{ij}\}_{i, j < \alpha}, \{ c \}_{c < \alpha} \rangle$
-\end{center}
-
-$\mathcal{B}$ is a subreduct of $\mathcal{C}$, denoted as $\mathcal{B} \subseteq^{r} \mathcal{C}$, if $\mathcal{B} \subseteq \mathfrak{R}\mathfrak{d}_{\gamma}\mathcal{C}$ for some $\gamma \leq \beta$.
-
-\begin{defin}
-  Let $\mathcal{C}$ be a $\beta$-dimensional cylindic algebra and $\alpha$ an ordinal such that $\alpha \leq \beta$.
-  The neat $\alpha$-reduct of $\mathcal{C}$, denoted as $\mathfrak{N}\mathfrak{r}_\alpha \mathcal{C}$,
-  is the subalgbera $\mathcal{A}$ of $\mathfrak{R}\mathfrak{d}_{\alpha}\mathcal{C}$ with $\mathcal{A} = \operatorname{Cl}_{\kappa} \mathcal{C}$ where $\alpha + \kappa = \beta$.
-
-  Let $\mathbb{K}$ be a class of $\beta$-dimensional cylindic algebras, then we put
-  \begin{center}
-    ${\bf Nr}_{\alpha}\mathbb{K} = \{ \mathfrak{N}\mathfrak{r}_\alpha \mathcal{C} \: | \: \mathcal{C} \in \mathbb{K} \}$
-  \end{center}
-
-  An algebra $\mathcal{B}$ is a neat subreduct of $\mathcal{C}$, or $\mathcal{B}$ is neatly embeddable to $\mathcal{C}$ if there exists an ordinal $\gamma \leq \alpha$ such that
-  $\mathcal{C} \subseteq \mathfrak{R}\mathfrak{d}_{\gamma}\mathcal{B}$.
-\end{defin}
-
-One may define neat reducts alternatively as follows.
-Let $\mathcal{C}$ be a $\beta$-dimensional cylindic algebra and $\alpha$ an ordinal such that $\alpha \leq \beta$.
-The neat $\alpha$-reduct of $\mathcal{C}$ is the $\alpha$-dimensional cylindric algebra having the form
-
-\begin{center}
-  $\mathfrak{N}\mathfrak{r}_\alpha \mathcal{C} = \langle \{ a \in \mathcal{C} \: | \: \forall j (\alpha \leq j \: \& \: j < \beta \Rightarrow c_j a = a) \}, +, -, 0, 1, \{d_{ij}\}_{i,j < \alpha}, \{c_{\gamma}\}_{\gamma} \rangle$
-\end{center}
-
-\section{Completely representable cylindric algebras of dimension $\omega$}
-
-\begin{defin}
-  Let $\mathcal{A}$ be a BAO of type ${\bf CA}_{\omega}$, an $\mathcal{A}$-pre-network is a pair
-  $\mathcal{N} = \langle N, l \rangle$, where $N$ is a set of nodes and $l : \prescript{\omega}{}N \to \operatorname{At}(\mathcal{A})$.
-
-  $\mathcal{N}$ is a network, if the following conditions hold, for all $x, y \in \prescript{\omega}{}N$ and $i, j < \omega$:
-  \begin{enumerate}
-    \item $l(x) \leq d_{ij}$ iff $x_i = x_j$
-    \item $x \equiv_i y$ implies $l(x) \leq c_i l(y)$
+  \item Every $R_i$ is an equivalence relation on $W$,
+  \item $R_i \circ R_j = R_j \circ R_i$, that is, the set $(R_i)_{i < \omega}$ forms a commutative semigroup under composition.
+  \item For all $x \in W$, $E_{ii}(x)$ holds.
+  \item For all $x, y, z \in W$, $x R_i y \: \& \: E_{ij}(y) \: \& \: x R_i z \: \& \: E_{ij}(y)$ implies $y = z$.
+  \item For all $x \in W$, $E_{ij}(x)$ iff there exists $y \in W$ such that $x R_k y$, $E_{ik}(y)$, and $E_{kj}(y)$.
   \end{enumerate}
@@
 
-Let $\mathcal{N}_1 = \langle N_1, l_1 \rangle$ and $\mathcal{N}_2 = \langle N_2, l_2 \rangle$ be networks, then $\mathcal{N}_1 \subseteq \mathcal{N}_2$ if $N_1 \subseteq N_2$ and $l_1 = l_2 	\upharpoonright_{N_1}$.
+The following fact is by Venema, see \cite[Proposition 2.1.5]{Venema2013}:
+\begin{prop}
+An $\omega$-frame $\mathcal{F}$ is cylindric iff $\mathfrak{Cm}(\mathcal{F})$ is a cylindric algebra of dimension $\omega$.
+\end{prop}
+A cylindric $\omega$-frame $\mathcal{F}$ is completely representable, if $\mathfrak{Cm}(\mathcal{F})$ is completely representable as a cylindric algebra of dimension $\omega$.
 
-Let $\Lambda \in \operatorname{Lim}$ and $\{ \mathcal{N}_{\lambda}\}_{\lambda < \Lambda}$ a sequence of networks such that
-\begin{center}
-  $\langle N_0, l_0 \rangle \subseteq \langle N_1, l_1 \rangle \subseteq \dots \langle N_{\lambda}, l_{\lambda} \rangle \subseteq \dots $ for $\lambda < \Lambda$
-\end{center}
-then \emph{the limit} of the sequence $\{ \mathcal{N}_{\lambda}\}_{\lambda < \Lambda}$ is the network
-\begin{center}
-  $\mathcal{N} = \langle N, l \rangle = \bigcup \limits_{\lambda < \Lambda} \langle N_{\lambda}, l_{\lambda} \rangle$
-\end{center}
-with nodes $N = \bigcup \limits_{\lambda < \Lambda} N_{\lambda}$ and labelling $l = \bigcup \limits_{\lambda < \Lambda} l_{\lambda}$, that is,
-  for any $\lambda \in \Lambda$ and $x \in \prescript{\omega}{}N$ one has $l(x) = l_{\lambda}(x)$.
+In this section, we reproduce the results related to characterisation ${\bf RCA}_{\omega}$. The following results are due to Goldblatt \cite{goldblatt1995elementary}.
 
-The elements of $\prescript{\omega}{}N$ are called \emph{$\omega$-dimensional hyperedges} of a network. One may identify a complete representation of an atomic cylindric-type algebra $\mathcal{A}$
-with a set $\{ \mathcal{N}_a \: | \: a \in \operatorname{At}(\mathcal{A}) \}$ of $\mathcal{A}$-networks with the following additional condition:
-\begin{itemize}
-  \item For each $a \in \operatorname{At}(\mathcal{A})$ there exists $x \in \prescript{\omega}{}N_a$ such that $l_a(x) = a$ and for each $z \in \prescript{\omega}{}N_a$ and $b \in \operatorname{At}(\mathcal{A})$, $i < \omega$ with $l_a(z) \leq c_i b$
-  there exists $y \in \prescript{\omega}{}N_a$ such that $z \equiv_i y$ and $l_a(y) = b$.
-\end{itemize}
-
-We define a complete representation $h$ of a cylindric-type algebra $\mathcal{A}$ as follows, for any $b \in \mathcal{A}$:
-\begin{center}
-$h(b) = \{ x \: | \: \exists a \in \operatorname{At}(\mathcal{A}), x \in \prescript{\omega}{}N_a, l_a(x) \leq b \}$
-\end{center}
-
-Let us define an atomic game.
-
-\begin{defin}
-  Let $\mathcal{A}$ be an atomic BAO of type ${\bf CA}_{\omega}$ and $\kappa > 0$ a cardinal.
-The game $\mathcal{G}^{\kappa}(\mathcal{A})$ is defined as follows.
-The game has two players: $\forall$ (Abelard, he/his) and $\exists$ (H\'{e}lo\"{i}se, she/her).
-A play of the game $\mathcal{G}^{\kappa}(\mathcal{A})$ is the sequence of networks
-  \begin{center}
-    $\mathcal{N}_0 \subseteq \mathcal{N}_1 \subseteq \mathcal{N}_2 \subseteq \dots \subseteq \mathcal{N}_{\lambda} \subseteq \dots$ for $\lambda < \kappa$
-  \end{center}
-  The game consists of the following stages:
-  \begin{enumerate}
-    \item ({\bf Zero round})
-
-    $\forall$ picks an atom $a \in \operatorname{At}(\mathcal{A})$ and $\exists$ plays a network $\mathcal{N}_0$.
-If there is no $x \in \prescript{\omega}{}N_0$ such that $l_0(x) = a$, then $\forall$ wins the play.
-    \item ({\bf Successor round})
-
-    Let $0 < \lambda$ be a cardinal such that $\lambda + 1 < \kappa$ and a network $\mathcal{N}_{\lambda} = \langle N_{\lambda}, l_{\lambda} \rangle$ has been already played.
-
-    $\forall$ picks $i < \omega$, $x \in \prescript{\omega}{}N_{\lambda}$, $a \in \operatorname{At}$ such that $l_{\lambda}(x) \leq c_i a$.
-    We denote this move as $(i, x, a)$.
-$\exists$ responds with a network $\mathcal{N}_{\lambda + 1} \supseteq \mathcal{N}_{\lambda}$.
-$\forall$ wins, if there is no node $c \in N$ such that $l_{\lambda+1}(x[i/c]) = a$, then $\forall$ wins
-    \item The limit of the play is defined as $\bigcup \limits_{\lambda < \kappa} \mathcal{N}_{\lambda}$.
-$\forall$ wins the play, if there exists $\kappa_1 < \kappa$ such that $\exists$ does not win the $\kappa_1$th-round.
-Otherwise, $\exists$ wins the play.
-  \end{enumerate}
-\end{defin}
+\begin{lemma}
+${\bf RCA}_{\omega} = {\bf S} \: \mathfrak{Cm} {\bf Ud} \mathcal{F} \mathfrak{ct}_{\alpha}$
+\end{lemma}
 
 \begin{theorem}
-  Let $\mathcal{A}$ be an atomic $\omega$-dimensional cylindric-type algebra and $\kappa$ a cardinal such that $|\operatorname{At}(\mathcal{A})| = \kappa$, then the following are equivalent:
-
-  \begin{enumerate}
-    \item $\mathcal{A}$ is completely representable.
-    \item $\exists$ has a winning strategy in $\mathcal{G}^{\kappa + \omega}$.
-  \end{enumerate}
+${\bf RCA}_{\omega}$ is a canonical variety.
 \end{theorem}
 
-\begin{proof}
-  $ $
-
-  \begin{enumerate}
-    \item $\Rightarrow$
-    If $\mathcal{A}$ is completely representable, then its Boolean reduct is completely representable as well by Theorem~\ref{completeboolean}.
-$\exists$ maintains that embedding to win the play.
-TODO: write down this proof in more detail
-
-    \item $\Leftarrow$
-
-    Suppose $\exists$ has a winning strategy in $\mathcal{G}^{\kappa + \omega}(\mathcal{A})$.
-    In every round $\forall$ picks all possible $i < \omega$, $a \in \operatorname{At}(\mathcal{A})$, all possible hyperedges and all appropriate atoms and $\exists$ has a proper response for every $\forall$'s move.
-
-    For each atom consider a play of the game with fewer than $\kappa + \omega$ nodes.
-For each $a \in \operatorname{At}(\mathcal{A})$ we associate a network $\mathcal{N}_a$, the resulting network of a corresponding game.
-Consider the set $\{ \mathcal{N}_a \: | \: a \in \operatorname{At}(\mathcal{A}) \}$.
-
-    Let $a$ be an atom, consider the network $\mathcal{N}_a = \langle V, l_a \rangle$.
-If there was not $x \in V$ such that $l_a(x) = a$, then $\forall$ would have a winning strategy, but that is not true, such an $x$ does exist.
-The second item of this criterion follows from the presence of a winning strategy for $\exists$ as well.
-
-    So we define a map $rep$:
-    \begin{center}
-    $rep(a) = \{ x \: | \: \exists b \in \operatorname{At}(\mathcal{A}) \: x \in \prescript{\omega}{}N_a, l_a(x) \leq b \}$.
-  \end{center}
-    We check that $rep$ preserves cylindrifications and diagonal elements. Let $i, j < \omega$ and $a \in \mathcal{A}$:
-    \begin{enumerate}
-    \item Suppose $x \in rep(c_i a)$, then there exists an atom $b$ such that $x \in \prescript{\omega}{}N_b$ with $l_b(x) \leq c_i a$.
-Then there exists $y \equiv_i x$ with $l_b(y) \leq a$, so $x \in {\bf C}_i(rep(a))$.
-
-    If $x \in {\bf C}_i(rep(a))$, then there exists $y \equiv_i x$ such that $y \in rep(a)$, that is, there exists an atom $b$ such that $y \in \prescript{\omega}{}N_b$ and $l_b(y) \leq a$.
-    \item If $x \in rep(d_{ij})$, so there exists an atom $b$ with $x \in \prescript{\omega}{}N_a$ and $l_b(x) \leq d_{ij}$, then $x_i = x_j$, then $x \in D_{ij}$.
-    \end{enumerate}
-  \end{enumerate}
-\end{proof}
-
-\begin{theorem} Let $\mathcal{A}$ be a BAO of type ${\bf CA}_{\omega}$:
-
-\begin{enumerate}
-  \item $\exists$ has a winning strategy in $\mathcal{G}_{m}(\mathcal{A})$ ($m < \omega$), then $\exists$ has a winning strategy in $\mathcal{G}_{\omega}(\Pi_U \mathcal{A})$, where $\Pi_U \mathcal{A}$ is the non-principal ultrapower of $\mathcal{A}$ modulo ultrafilter $U$ over $\omega$.
-  \item $\exists$ has a winning strategy in $\mathcal{G}_{m}(\mathcal{A})$ (for every $m < \omega$) iff $\mathcal{A}$ is elementarily equivalent to a completely representable cylindric algebra of dimension $\omega$.
-\end{enumerate}
-\end{theorem}
-
-\begin{proof}
-  $ $
-
-  The argument uses \L o\'s's Theorem, see \cite[Theorem 8.5.3]{hodges1997shorter}.
-
-  \begin{enumerate}
-    \item
-    \item
-  \end{enumerate}
-\end{proof}
-
-\section{The result itself}
-
-\begin{lemma}\label{Neat}
-  Let $\mathcal{A}$ be a BAO of type ${\bf CA}_{\alpha}$ and $\mathcal{B}$ be a $\beta$-dimensional cylindric algebra such that $\beta \leq \alpha$ and $\mathcal{A}$ neatly embeds to $\mathcal{B}$ by a complete embedding.
-
-  \begin{enumerate}
-  \item $\mathcal{A}^{+}$ neatly embeds to $\mathcal{B}^{+}$ by a complete embedding.
-  \item $\mathcal{A}$ is atomic.
-  \end{enumerate}
-\end{lemma}
-
-\begin{proof}
-$ $
-
-\begin{enumerate}
-  \item See \cite[Remark 2.7.25]{henkin1971cylindric}.
-  \item Is it true?
-\end{enumerate}
-\end{proof}
-
-\begin{theorem}[This assumption is by Ian Hodkinson]\label{Neat2}
-  $ $
-
-  Let $\mathcal{A}$ be a BAO of type ${\bf CA}_{\omega}$ such that $\mathcal{A}$ neatly embeds into ${\bf CA}_{\omega + \omega}$ by a complete embedding.
-Then $\mathcal{A}$ is completely representable as ${\bf CA}_{\omega}$.
-\end{theorem}
-
-\begin{proof}
-  Suppose $\mathcal{A} \subseteq \mathfrak{Nr}_{\omega} \mathcal{B}$, where $\mathcal{B} \in {\bf RCA}_{\omega + \omega}$ and the inclusion map $\rho : \mathcal{A} \hookrightarrow \mathfrak{Nr}_{\omega} \mathcal{B}$ is a complete embedding, that is:
-  \begin{center}
-    $\rho (\sum \limits_{i \in I} a_i) = \sum \limits_{i \in I} (\rho a_i)$, if $\sum \limits_{i \in I} a_i$ exists.
-  \end{center}
-
-  Let us show that $\mathcal{A}$ is atomic.
-
-  Consider $\rho(\mathcal{A})$. Let us show that $\exists$ has a winning strategy on $\mathcal{G}^{\kappa + \omega}(\rho(\mathcal{A}))$
-\end{proof}
-
-Lemma~\ref{Neat} and Theorem~\ref{Neat2} imply the following theorem.
-
-\begin{theorem}
-  Let $\mathcal{C} \in {\bf RCA}_{\omega}$, then $\mathcal{C}^{+} \in {\bf RCA}_{\omega}$. That is, ${\bf RCA}_{\omega}$ is closed under canonical extensions.
-\end{theorem}
-
-\begin{proof}
-
-\end{proof}
-
-\section{(Lack of) canonical axiomatisation of ${\bf CA}_{\omega}$}
-
-Here we are going to show that ${\bf CA}_{\omega}$ fails to have a canonical axiomatisation, the similar results for ${\bf RRA}$ and ${\bf RCA}_n$ for finite $n \geq 3$ have been shown by Hodkinson and Venema \cite{hodkinson2005canonical} and by Bulian and Hodkinson respectively \cite{bulian2013bare}.
-
-\section{Notes on the canonicity of ${\bf RRA}$}
-
-\begin{defin}
-  $ $
-
-    A relation algebra is an algebra $\mathcal{R} = \langle R, 0, 1, +, -, ;, {}^{\smile}, {\bf 1 }' \rangle$ such that $\langle R, 0, 1, +, - \rangle$ is a Boolean algebra and the following
-    equations hold, for each $a, b, c \in R$:
-    \begin{enumerate}
-      \item $a ; (b ; c) = (a ; b) ; c$
-      \item $(a + b) ; c = (a ; c) + (b ; c)$
-      \item $a ; {\bf 1}' = a$
-      \item $a^{\smile \smile} = a$
-      \item $(a + b)^{\smile} = a^{\smile} + b^{\smile}$
-      \item $(a ; b)^{\smile} = b^{\smile} ; a^{\smile}$
-      \item $a^{\smile} ; (- (a ; b)) \leq - b$
-    \end{enumerate}
-    where $a \leq b$ iff $a + b = b$. ${\bf RA}$ denotes the class of all relation
-    algebras.
-\end{defin}
-
-We will adapting the following proof of the fact that ${\bf RRA}$ is canonical \footnote{This idea is by Ian Hodkinson} to our case.
-This proof is due to Monk, but that was describe in McKenzie's thesis \cite{mckenzie1968representation}.
-
-\begin{enumerate}
-  \item A relation algebra $\mathcal{A}$ is representable iff $\mathcal{A}$ neatly embeds to some $\omega$-dimensional cylinric algebra,
-  \item If $\mathcal{A}$ neatly embeds in $\mathcal{A}$ then $\mathcal{A}^{+}$ neatly embeds in $\mathcal{B}^{+}$,
-  \item ${\bf CA}_{\alpha}$ is closed under canonical extensions,
-  \item Voil\'{a}.
-\end{enumerate}
-
-\begin{defin}
-  Let $\mathcal{C} \in {\bf CA}_{\alpha}$, where $\alpha \geq 3$. The relation algebra reduct of $\mathcal{C}$, written as $\mathfrak{Ra}(\mathcal{C})$, is the algebra having the form
-
-  \begin{center}
-    $\langle \operatorname{dom}(\mathfrak{Nr}_2(\mathcal{C})), 0, 1, +, -, {\bf 1}', \smile, ; \rangle$
-  \end{center}
-  where:
-
-  \begin{enumerate}
-    \item $+$, $-$, $0$, and $1$ are defined as usual in $\mathcal{C}$,
-    \item ${\bf 1}' = d_{01} \in \mathfrak{Nr}_2(\mathcal{C})$,
-    \item $r^{\smile} = s^{2}_{0} s^{0}_{1} s^{1}_{2} r$ for $r \in \mathfrak{Nr}_2(\mathcal{C})$,
-    \item Let $r, s \in \mathfrak{Nr}_2(\mathcal{C})$, then $r ; s = c_2 (s^{1}_{2}r \cdot s^{0}_{2}s)$
-  \end{enumerate}
-\end{defin}
-
-Moreover, $\mathfrak{Nr}_{\beta}(\mathcal{C})$ and $\mathfrak{Ra}(\mathcal{C})$ are closed under these operations.
-There is also the following fact by due to Henkin, Monk, and Tarski \cite{Henkin1988-HENCAP-4}:
-\begin{theorem}
-  Let $\mathcal{C} \in {\bf CA}_{\alpha}$ for $\alpha \geq 4$, then $\mathfrak{Ra}(\mathcal{C})$ is a relation algebra.
-\end{theorem}
-
-The following characterisation results are by Henkin, Monk, and Tarski \cite[5.3.13, 5.3.16]{Henkin1988-HENCAP-4} as well:
-\begin{theorem} \label{char}
-  $ $
-
-  \begin{enumerate}
-    \item ${\bf RA} = {\bf S} \mathfrak{Ra} {\bf CA}_4$,
-    \item ${\bf RRA} = \bigcap \limits_{3 \leq n < \omega} {\bf S} \mathfrak{Ra} {\bf CA}_n =  {\bf S} \mathfrak{Ra} {\bf CA}_{\alpha}$, where $\alpha$ is an infinite ordinal.
-  \end{enumerate}
-\end{theorem}
-
-Let $\mathcal{C} \in \mathcal{CA}_{\alpha}$, then $\mathcal{R} \in {\bf RA}$ neatly embeds to $\mathcal{C}$, if $\mathcal{R}$ is isomorphic to some subalgebra of $\mathfrak{Ra}(\mathcal{C})$.
-
-\begin{theorem}
-  ${\bf RRA}$ is closed under canonical extensions.
-\end{theorem}
-
-\begin{proof} Let ${\bf R} \in {\bf RRA}$.
-  By the second item of~\ref{char}, every representable relation algebra is isomorphic to some subalgbera of the relation algebra reduct $\mathfrak{Ra}\mathcal{C}$ for some $\mathcal{C} \in {\bf CA}_{\omega}$.
-But neat embeddings repsect canonical extensions, so if ${\bf R} \hookrightarrow_{n} \mathcal{C}$, so is ${\bf R}^{+} \hookrightarrow_{n} \mathcal{C}^{+}$.
-  ${\bf CA}_{\alpha}$ is closed under canonical extensions, so is ${\bf RRA}$.
-\end{proof}
+\section{Representability games}
 

### 2021-11-08 frp-paper/Paper.tex
cylindric notes, wip

@@
 \date{}
-\title{The finite representation property for some reducts of relation algebras}
+\title{Reducts of relation algebras: finite representability and recursively enumerable axiomatisability}
 \providecommand{\keywords}[1]{\textbf{\textit{Keywords---}} #1}
@@
 We also show that the class of representable join semilattice-ordered semigroups has a recursively
-enumerable axiomatisation and the finite representation property.
+enumerable axiomatisation using back-and-forth games.
 \end{abstract}
@@
 
-Relation algebras are the kind of Boolean algebras with operators that provide algebraisation
+Relation algebras are a kind of Boolean algebras with operators that provide algebraisation
 of binary relations \cite{jonsson1951boolean}. The class of all relation algebras,
-denoted as ${\bf RA}$, consists of algebras the signature of which is $\{ 0, 1, +, -, ;, {}^{\smile}, {\bf 1}' \}$ obeying certain axioms.
+denoted as ${\bf RA}$, consists of algebras of the signature of which is $\{ 0, 1, +, -, ;, {}^{\smile}, {\bf 1}' \}$, and all those algebras obey certain axioms.
 The class of representable relation algebras, ${\bf RRA}$, consists of algebras isomorphic to set relation algebras. ${\bf RRA}$ is a subclass of ${\bf RA}$, but the converse inclusion does not hold.
 That is, there exist relation algebras with no representation as set relation algebras
-\cite{lyndon1950representation}. Moreover, the class ${\bf RRA}$ is not finitely axiomatisable \cite{monk1964representable}. The problem of determining whether a given finite relation algebra
+\cite{lyndon1950representation}.
+Moreover, the class ${\bf RRA}$ is not a finitely axiomatisable variety \cite{monk1964representable} with neither Sahlqvist \cite{venema1997atom} nor axiomatisation \cite{hodkinson2005canonical}.
+The problem of determining whether a given finite relation algebra
 is representable is undecidable, see \cite{hirsch2001representability}.
 
-In other words, relation algebras are quite badly behaved.
-The study of such reducts is mostly motivated by such ``bad behaviour'' in order to avoid these restrictions and determine the possible reasons for them.
+In other words, relation algebras are quite badly behaved. For this reason, we are interested in reduct since they might behave better.
 
-There are several results on reducts of relation algebras having no finite axiomatisation such as ordered monoids \cite{hirsch2005class}, distributive residuated lattices \cite{andreka1994lambek}, join semilattice-ordered semigroups \cite{andreka2011axiomatizability}, algebras whose signature contains composition, meet, and converse \cite{hodkinson2000axiomatizability}, etc.
+There are several results on reducts of relation algebras that have no finite axiomatisation such as ordered monoids \cite{hirsch2005class}, distributive residuated lattices \cite{andreka1994lambek}, join semilattice-ordered semigroups \cite{andreka2011axiomatizability}, algebras whose signature contains composition, meet, and converse \cite{hodkinson2000axiomatizability}, etc.
 
@@
 already known that this class is not finitely axiomatisable \cite{andreka2011axiomatizability}.
-First, we show that ${\bf R}(;,+)$ has a recursively enumerable axiomatisation. For that,
+We show that ${\bf R}(;,+)$ has a recursively enumerable axiomatisation. For that,
 we define networks and representability games for them. This class is axiomatised with the axioms of join semilattice-ordered semigroups plus the countable set of universal formulas claiming that
-$\exists$ has a winning strategy on every finitary step.
-
-After that, we show that ${\bf R}(;,+)$ has the finite representation property.
-Moreover, we extract the explicit recursively enumerable axiomatisation of ${\bf R}(;,+)$.
-As a corollary, the representability problem for finite join semilattice-ordered semigroup is decidable.
+$\exists$ has a winning strategy on every finitary step. The question of finite representability for this class remains open \cite{vsemrl2021domain}.
 
@@
 
-Let us the basic definitions related to relation algebras \cite{hirsch2002relation}.
+Let us introduce the basic definitions related to relation algebras. See \cite[Section 3]{hirsch2002relation} to have more details.
 \begin{definition} A relation algebra is an algebra $\mathcal{R} = \langle R, 0, 1, +, -, ;, {}^{\smile}, {\bf 1 }\rangle$ such that $\langle R, 0, 1, +, - \rangle$ is a Boolean algebra and the following equations hold, for each $a, b, c \in R$:
@@
 ${\bf R}(\tau)$ has the finite representation property if every $\mathcal{A} \in {\bf R}(\tau)$ has a finite representation over a finite base.
+
+The finite representation property and recursively enumerable axiomatisability imply decidability for finite structures in the following sense.
+One may think of that as a counterpart of Harrop's theorem from universal algebra \cite{harrop1958existence}.
+
+\begin{proposition}\label{dec}
+  Let $\tau$ be a signature of operations and predicates definable in ${\bf RA}$, then, if ${\bf R}(\tau)$ has a recursively enumerable axiomatisation and the finite representation property, then the membership problem for finite $\tau$-structures is decidable.
+\end{proposition}
+
+\begin{proof}
+  Suppose ${\bf R}(\tau)$ has a recursively enumerable axiomatisation and every finite
+  $\mathcal{A} \in {\bf R}(\tau)$ is representable over a finite base.
+
+  Let $\mathcal{A}$ be a $\tau$-structure. We check axioms of $\mathcal{A} \in {\bf R}(\tau)$  with the algorithm $S$ that yields ``no'' and halts, if $\mathcal{A}$ is not representable.
+
+  In parallel, we check whether $\mathcal{A}$ has a finite representation.
+If there exists such $0 < n < \omega$ such that $\mathcal{A}$ is representable over size $n$, then the algorithm $T$ does output ``yes'' and terminates.
+
+  So, one has either $S(\mathcal{A}) = \text{no}$ and $T(\mathcal{A}) = \text{yes}$.
+\end{proof}
 
@@
 \begin{center}
-$(\mathcal{P}(A))_m = \{ X \in \mathcal{P}(S) \: | \: m X = X\}$ )
+$(\mathcal{P}(A))_m = \{ X \in \mathcal{P}(S) \: | \: m X = X\}$
 \end{center}
@@
 
-\section{Join semilattice-ordered semigroups: the explicit axiomatisation and the finite representation property}
+\section{Join semilattice-ordered semigroups: the explicit axiomatisation}
 
@@
 That makes ${\bf R}(;,+)$ not finitely axiomatisable.
-Thus, one needs to show the decidability of membership in a different way.
-We have the following fact:
+Although, this class has a universal recursively enumerable axiomatisation.
 
-\begin{proposition}\label{dec}
-  Let $\tau$ be a signature of operations and predicates definable in ${\bf RA}$, then, if ${\bf R}(\tau)$ has a recursively enumerable axiomatisation and the finite representation property, then the membership problem for finite $\tau$-structures is decidable.
-\end{proposition}
-
-\begin{proof}
-  Suppose ${\bf R}(\tau)$ has a recursively enumerable axiomatisation and every finite
-  $\mathcal{A} \in {\bf R}(\tau)$ is representable over a finite base.
-
-  Let $\mathcal{A}$ be a $\tau$-structure. We check axioms of $\mathcal{A} \in {\bf R}(\tau)$  with the algorithm $S$ that yields ``no'' and halts, if $\mathcal{A}$ is not representable.
-
-  In parallel, we check whether $\mathcal{A}$ has a finite representation.
-If there exists such $0 < n < \omega$ such that $\mathcal{A}$ is representable over size $n$, then the algorithm $T$ does output ``yes'' and terminates.
-
-  So, one has either $S(\mathcal{A}) = \text{no}$ and $T(\mathcal{A}) = \text{yes}$.
-\end{proof}
-
-In this section, we show that ${\bf R}(+, ;)$ has a recursively enumerable universal axiomatisation.
-For that, we characterise representability using Ehrenfeucht-Fra\"{i}ss\'{e} games on networks.
-After that, we show that every finite $\mathcal{A} \in {\bf R}(+, ;)$ is representable by factorising its network through an equivalence relation having a finite index.
-That is, an equivalence relation whose quotient is finite.
+In this section, we show that ${\bf R}(+, ;)$ has a recursively enumerable universal axiomatisation. For that, we characterise representability using Ehrenfeucht-Fra\"{i}ss\'{e} games on networks.
 
@@
 
-\subsection{The finite representation property for ${\bf R}(+, ;)$}
-
-Let us show finite representability for ${\bf R}(+, ;)$.
-Suppose that a finite $\mathcal{A}$ is representable, so $\exists$ has a winning strategy on $\mathcal{G}^{\omega}(\mathcal{A})$.
-Let us consider $\mathcal{N} = \langle V, E, l \rangle$, the network constructed during the $\mathcal{G}^{\omega}(\mathcal{A})$ game.
-
-\begin{definition}
-  Let $\mathcal{A}$ be a join semilattice-ordered semigroup and let $\mathcal{N}_1$ and $\mathcal{N}_2$ be $\mathcal{A}$-networks.
-A strong homomorphism from $\mathcal{N}_1$ onto $\mathcal{N}_2$ is a map $f : \mathcal{N}_1 \twoheadrightarrow \mathcal{N}_2$ such that:
-  \begin{enumerate}
-    \item $(x, y) \in E_2$ implies $(f(x), f(y)) \in E_1$,
-    \item If $(x', y') \in E_2$, then there exists $x, y \in E_1$ such that $f(x) = x'$ and $f(y) = y'$ and $(x, y) \in E_1$,
-    \item $l_{\mathcal{N}_1}(x, y) = l_{\mathcal{N}_2}(f(x),f(y))$
-  \end{enumerate}
-  By $\mathcal{N}_1 \twoheadrightarrow \mathcal{N}_2$ we mean that there exists a strong homomorphism of $\mathcal{N}_1$ onto $\mathcal{N}_2$.
-\end{definition}
-
-\begin{lemma}
-  Let $\mathcal{A}$ be a representable join semilattice-ordered semigroup and $\mathcal{N} = \langle V, E, l \rangle$ and $\mathcal{N}' = \langle V', E', l' \rangle$ $\mathcal{A}$-networks.
-
-  Suppose $\mathcal{N} \twoheadrightarrow \mathcal{N}'$. If $\mathcal{N}^{*}$ is a representation of $\mathcal{A}$, so is $ $
-\end{lemma}
-
-\begin{lemma}
-\end{lemma}
-
 \section{Acknowledgements}
 
-The author is sincerely grateful to Robin Hirsch, Ian Hodkinson, and his supervisor Ilya Shapirovsky for valuable comments.
-The previous version of this paper contained an error.
-Many thanks to Ja\v{s} \v{S}emrl for noticing that.
+The author is sincerely grateful to Robin Hirsch, Ja\v{s} \v{S}emrl, and his supervisor Ilya Shapirovsky for valuable comments.
 

### 2021-11-10 cylindric-notes/Text.tex
updates in cylindric notes

@@
 \date{}
-\title{Representable cylindric algebras of dimension $\omega$}
+\title{Complete representability for canonical extensions of representable cylindric algebras}
 
@@
 
-Suppose $\mathcal{C} \in {\bf RCA}_{\omega}$, whether $\mathcal{C}^{+}$ has a complete, $\omega$-dimensional representation? \cite{hirsch2002relation}
+Given $\mathcal{C} \in {\bf RCA}_{\omega}$, whether $\mathcal{C}^{+}$ has a complete, $\omega$-dimensional representation? \cite{hirsch2002relation}
 
@@
 A generalised cylindric set algebra of dimension $\alpha$ is a subdirect of cylindric algebras that have dimension $\alpha$.
+   ${\bf Cs}_{\alpha}$ denotes the class of all cylindric set algebras of dimension $\alpha$.
 \end{defin}
@@
 
+We are interested in the special case of cylindric $\omega$-frames called Cartesian structure of dimension $\omega$. To be more precise:
+
+\begin{defin}
+Let $U$ be a set and $V \subseteq \prescript{\omega}{}U$ be a non-empty subset of the full Cartesian space of dimension $\omega$, then an $\alpha$-dimension Cartesian structure generated by $V$ is an $\omega$-frame $\mathfrak{G}(V) = \langle V, (R_{i})_{i < \omega}, (E_{ij})_{i, j < \omega} \rangle$ such that:
+\begin{enumerate}
+\item $R_i = \{ (w, v) \: | \: w, v \in V, w_k = w_k, k < \omega, i \neq k \}$
+\item $E_{ij} = \{ w \in V \: | \: w_i = w_j \}$
+\end{enumerate}
+$\mathfrak{G}(\prescript{\omega}{}U)$ is the full $\omega$-dimensional Cartesian structure.
+$\mathcal{F} \mathfrak{c}\mathfrak{t}_{\omega}$ is the class of all full $\omega$-dimensional Cartesian structures.
+\end{defin}
+
+Every cylindric set algebra is a subalgebra of some complex algebra induced by an $\omega$-dimensional Cartesian structure. That is, we have the following characterisation:
+
+\begin{lemma}
+${\bf I}{\bf Cs}_{\omega} = {\bf S} \mathfrak{Cm} \mathcal{F} \mathfrak{c}\mathfrak{t}_{\omega}$.
+\end{lemma}
+
 In this section, we reproduce the results related to characterisation ${\bf RCA}_{\omega}$. The following results are due to Goldblatt \cite{goldblatt1995elementary}.
@@
 \begin{lemma}
-${\bf RCA}_{\omega} = {\bf S} \: \mathfrak{Cm} {\bf Ud} \mathcal{F} \mathfrak{ct}_{\alpha}$
+${\bf RCA}_{\omega} = {\bf S} \: \mathfrak{Cm} {\bf Ud} \mathcal{F} \mathfrak{ct}_{\omega}$
 \end{lemma}

### 2021-11-21 cylindric-notes/Text.tex
more updates on RCA_omega

@@
 
-Given $\mathcal{C} \in {\bf RCA}_{\omega}$, whether $\mathcal{C}^{+}$ has a complete, $\omega$-dimensional representation? \cite{hirsch2002relation}
-
+\begin{itemize}
+\item Given $\mathcal{C} \in {\bf RCA}_{\omega}$, whether $\mathcal{C}^{+}$ has a complete, $\omega$-dimensional representation? The conjecture is yes. \cite{hirsch2002relation}
+\item Whether ${\bf RCA}_{\omega}$ is barely canonical. The conjecture is yes.
+\end{itemize}
 \section{Atomic Representations}
@@
 \begin{defin} Let $\mathcal{F} = \langle W, (R_{i})_{i \in I} \rangle$ is an inner substructure \footnote{Or alternatively, a generated subframe} of $\mathcal{F}' = \langle W', ({R'}_{i})_{i \in I} \rangle$, if $W \subseteq W'$ and the embedding $\mathcal{F} \hookrightarrow \mathcal{F}'$ is a bounded morphism.
+Let $\mathbb{F}$ be a class atom structures, then $\mathbb{S}(\mathbb{F})$ is the closure of $\mathbb{F}$ under generated subframes.
 \end{defin}
@@
 
-Let $\alpha$ be an ordinal. Let $\prescript{\alpha}{}U$ be the set of all functions mapping $\alpha$ to a non-empty set $U$. We denote $x(i) = x_i$ for
+Let $\alpha$ be an ordinal. Denote $\{ f \: | \: f \alpha \to U\}$ as $\prescript{\alpha}{}U$. $x_i$ stands for $x(i)$, where
 $x \in \prescript{\alpha}{}U$ and $i < \alpha$.
@@
 Let $\mathcal{A} \in {\bf CA}_{\omega}$, then $\mathcal{A}$ has a \emph{complete representation}, if its representation preserves all existing suprema.
-In other words, $\mathcal{A}$ is completely representable.
+In other words, $\mathcal{A}$ is \emph{completely representable}.
 
@@
   \end{enumerate}
+  $\mathcal{C}\mathfrak{a}_{\omega}$ is the class of all $\omega$-frames.
+
+  If $\mathcal{F} \in \mathcal{C}\mathfrak{a}_{\omega}$ and $x \in \mathcal{F}$, then $\mathcal{F}^{x}$ is a generated subframe generated by $x$, which is defined standardly.
+Generally, $\mathcal{F}_1$ is a generated subframe of $\mathcal{F}_2$, if $\underline{\mathcal{F}_1} \subseteq \underline{\mathcal{F}_2}$ and $\underline{\mathcal{F}_1}$ is closed under under ${R_i}_2$ equivalences for every $i < \omega$.
+That is:
+  \begin{center}
+  For all $i < \omega$ and $x \in \mathcal{F}_1$, we have ${R_i}_2(x) \subseteq \mathcal{F}_1$ and, thus, ${R_i}_1(x) = {R_i}_2(x)$.
+  \end{center}
 \end{defin}
+
+We have the following connection betweens $\omega$-frames and their generated subframes, which is standard for modal logic:
+\begin{prop}
+Let $\mathcal{F} \in \mathcal{C}\mathfrak{a}_{\omega}$, then
+\begin{enumerate}
+\item $\mathcal{F} = \coprod \limits_{x \in \mathcal{F}} \mathcal{F}^{x}$,
+\item $\mathfrak{Cm}(\mathcal{F}) \cong \prod \limits_{x \in \mathcal{F}} \mathfrak{Cm}(\mathcal{F}^{x})$,
+\item $\mathfrak{Cm}(\mathcal{F}^{x})$ is subdirectly irreducible.
+\end{enumerate}
+\end{prop}
+
+It is known that $\mathcal{C}\mathfrak{a}_{\omega}$ forms an elementary class, since one can express the conditions of an $\omega$-frame with the first-order language.
 
@@
 \begin{defin}
-Let $U$ be a set and $V \subseteq \prescript{\omega}{}U$ be a non-empty subset of the full Cartesian space of dimension $\omega$, then an $\alpha$-dimension Cartesian structure generated by $V$ is an $\omega$-frame $\mathfrak{G}(V) = \langle V, (R_{i})_{i < \omega}, (E_{ij})_{i, j < \omega} \rangle$ such that:
+Let $U$ be a set and $V \subseteq \prescript{\omega}{}U$ be a non-empty subset of the full Cartesian space of dimension $\omega$, then an $\alpha$-dimension Cartesian structure generated by $V$ is an $\omega$-frame $\mathfrak{S}(V) = \langle V, (R_{i})_{i < \omega}, (E_{ij})_{i, j < \omega} \rangle$ such that:
 \begin{enumerate}
@@
 \end{enumerate}
-$\mathfrak{G}(\prescript{\omega}{}U)$ is the full $\omega$-dimensional Cartesian structure.
+$\mathfrak{S}(\prescript{\omega}{}U)$ is the full $\omega$-dimensional Cartesian structure.
 $\mathcal{F} \mathfrak{c}\mathfrak{t}_{\omega}$ is the class of all full $\omega$-dimensional Cartesian structures.
@@
 
-Every cylindric set algebra is a subalgebra of some complex algebra induced by an $\omega$-dimensional Cartesian structure. That is, we have the following characterisation:
+Clearly $\mathcal{F} \mathfrak{c}\mathfrak{t}_{\omega} \subseteq \mathcal{C}\mathfrak{a}_{\omega}$.
+
+We have the following connection between ${\bf RCA}_{\omega}$, ${\bf IGs}_{\omega}$, and complex algebras of full Cartesian structures:
+\begin{center}
+${\bf RCA}_{\omega} = {\bf IGs}_{\omega} = {\bf S} \mathfrak{Cm} {\bf Ud} \mathcal{F} \mathfrak{c}\mathfrak{t}_{\omega} = {\bf S P} \mathfrak{Cm}  \mathcal{F} \mathfrak{c}\mathfrak{t}_{\omega}$.
+\end{center}
+This follows from the fact that ${\bf Cs}_{\omega} = \mathfrak{Cm} \mathcal{F} \mathfrak{c}\mathfrak{t}_{\omega}$.
+Every generalised cylindric set algebra is a subdirect product of cylindric set algebras, thus, a generalised cylindric set algebra is a complex algebra of disjoint union of some full Cartesian spaces.
+But ${\bf RCA}_{\omega}$ is the closure of ${\bf Cs}_{\omega}$ under isomorphism.
+
+\begin{defin}
+The weak Cartesian space with base $U$ and dimension $\omega$ determined by $x \in \prescript{\omega}{}U$ is the set:
+\begin{center}
+$\prescript{\omega}{}U^{(x)} = \{ y \in \prescript{\omega}{}U \: | \: |\{ k < \omega \: | \: x_k \neq y_k \}| < \aleph_0 \}$
+\end{center}
+$\mathfrak{S}(\prescript{\omega}{}U^{(x)})$ is a weak Cartesian structure of dimension $\omega$.
+$\mathcal{W} \mathfrak{ct}_{\omega}$ is the class of all weak Cartesian structure of dimension $\omega$ up to isomorphism.
+\end{defin}
+Note that we have $\mathcal{W} \mathfrak{ct}_{\omega} \subseteq \mathcal{C}\mathfrak{a}_{\omega}$.
+
+Every cylindric set algebra is a subalgebra of some complex algebra induced by an $\omega$-dimensional Cartesian structure. In other words,
 
@@
 
-In this section, we reproduce the results related to characterisation ${\bf RCA}_{\omega}$. The following results are due to Goldblatt \cite{goldblatt1995elementary}.
+In this section, we reproduce the results related to characterisation ${\bf RCA}_{\omega}$.
+The following results are due to Goldblatt \cite{goldblatt1995elementary}.
+This denotes that a cylindric algebra of dimension algebra is representable iff it is isomorphic to a subalgebra of the complex algebra of disjoint sum of some full $\omega$-dimensional Cartesian structure.
+Assuming the duality, this is equivalent to the standard definition of representability formulated in terms of sublagebras of subdirect products.
 
-\begin{lemma}
-${\bf RCA}_{\omega} = {\bf S} \: \mathfrak{Cm} {\bf Ud} \mathcal{F} \mathfrak{ct}_{\omega}$
+\begin{lemma} \label{rcachar}
+${\bf RCA}_{\omega} = {\bf S} \: \mathfrak{Cm} \mathbb{S} {\bf Ud} \mathcal{F} \mathfrak{ct}_{\omega} = {\bf S} \: \mathfrak{Cm} \mathbb{S} {\bf Ud} \mathcal{W} \mathfrak{ct}_{\omega} = {\bf IGws}_{\omega}$
 \end{lemma}
+
+\begin{proof}
+
+\end{proof}
+
+Here we use the following fact related to canonical varieties generated by some class of complex algebras. Let ${\bf K}$ be an elementary class of relational structures, then:
+\begin{center}
+If ${\bf K}$ is closed under $p$-morphic images, generated subframes, and disjoint unious, \\ then ${\bf S} \mathfrak{Cm} {\bf K}$ is a canonical variety.
+\end{center}
+One may think of this fact a more abstract version of Fine's theorem which claims that every elementary modal logic is canonical \cite{fine1975some}.
+This version denotes the same fact, but it is formulated in terms of varieties BAOs generated by complex algebras of some atom structures.
+
+This is a specialised version of \cite[Theorem 4.4]{goldblatt1995elementary} formulated for dimension $\omega$.
 
@@
 \end{theorem}
+
+\begin{proof}
+By Lemma~\ref{rcachar},
+\end{proof}
 

### 2021-11-27 cylindric-notes/Text.tex
RCA_n (n < omega) and its canonicity

@@
 \newtheorem{theorem}{Theorem}
+
+\theoremstyle{claim}
+\newtheorem{claim}{Claim}
 
@@
 
-\begin{lemma} \label{rcachar}
+\begin{lemma} \ref{rcachar}
 ${\bf RCA}_{\omega} = {\bf S} \: \mathfrak{Cm} \mathbb{S} {\bf Ud} \mathcal{F} \mathfrak{ct}_{\omega} = {\bf S} \: \mathfrak{Cm} \mathbb{S} {\bf Ud} \mathcal{W} \mathfrak{ct}_{\omega} = {\bf IGws}_{\omega}$
@@
 This version denotes the same fact, but it is formulated in terms of varieties BAOs generated by complex algebras of some atom structures.
+We provide a more precise formulation of the fact above.
+
+Let ${\bf K}$ be a class of frames, denote the closure of ${\bf K}$ under ultraproducts as ${\bf Pu} {\bf K}$.
+
+\begin{prop}
+Let ${\bf K}$ be a class of frames, then ${\bf Pu}{\bf K} \subseteq \mathbb{H}\mathbb{S}{\bf Ud} {\bf K}$ implies that ${\bf S} \mathfrak{Cm} \mathbb{S}{\bf Ud} {\bf K}$ is a canonical variety.
+\end{prop}
 
@@
 \begin{proof}
-By Lemma~\ref{rcachar},
+We have ${\bf RCA}_{\omega} = {\bf S} \mathfrak{Cm} \mathbb{S} {\bf Ud} \mathcal{F} \mathfrak{ct}_{\omega}$.
+That's enough to show that ${\bf Pu} \mathcal{F} \mathfrak{ct}_{\omega} \subseteq \mathbb{H}\mathbb{S}{\bf Ud} \mathcal{F} \mathfrak{ct}_{\omega}$.
+For that, we need the following claim:
+\begin{claim}\label{claim1}
+${\bf Pu} \mathcal{F} \mathfrak{ct}_{\omega} \subseteq$
+\end{claim}
+\end{proof}
+
+\section{Canonicity of ${\bf RCA}_n$ for finite $n$}
+
+In this section we consider classes ${\bf RCA}_n$, where $n < \omega$ is finite.
+
+We provide the complete proof of the following theorem \cite[Theorem 3.4.3]{hirsch2013completions}.
+\begin{theorem}\label{finitecanon}
+Let $\mathcal{A} \in {\bf CA}_n$, then $\mathcal{A}$ is representable iff $\mathcal{A}^{+}$ is completely representable.
+\end{theorem}
+
+For that we need such model theoretic notions as saturation and types, see \cite[Section 6.3]{hodges1993model}.
+
+\begin{defin} Let $\mathcal{M}$ be a first-order structure of a signature $L$ and $S \subseteq \mathcal{M}$.
+Let $L(S)$ be an extension of $L$ with copies of elements from $S$ as additional constants.
+We assume that $Cnst(L)$ and $S$ are disjoint.
+
+\begin{enumerate}
+\item Let $n < \omega$, an $n$-type over $S$ is a set $\mathcal{T}$ of $L(S)$ formulas $A(\overline{x})$, where $\overline{x}$ is a fixed $n$-tuple of elements from $S$.
+Notation: $\mathcal{T}(\overline{x})$.
+A type is an $n$-type for some $n < \omega$.
+\item An $n$-type $\mathcal{T}(\overline{x})$ is realised in $\mathcal{M}$, if there exists $\overline{m} \in \mathcal{M}^n$ such that $\mathcal{M} \models A(\overline{m})$ for every $A \in \mathcal{T}(\overline{x})$.
+$\mathcal{M}$ omits $\mathcal{T}(\overline{x})$, if $\mathcal{T}(\overline{x})$ is not realised in $\mathcal{M}$.
+\item $\mathcal{T}(\overline{x})$ is finitely satisfied in $\mathcal{M}$, if every finite subtype $\mathcal{T}_0(\overline{x}) \subseteq \mathcal{T}(\overline{x})$ is realised in $\mathcal{M}$.
+We can reformulate that as $\mathcal{M} \models \exists \overline{a} \bigwedge \limits_{A \in \mathcal{T}_0} A(\overline{a})$.
+\item Let $T$ be a theory, then a type $\mathcal{T}$ over the empty set of constants is $T$-consistent, if there exists a model $\mathcal{M} \models T$ such that $\mathcal{T}$ is finitely satisfied in $\mathcal{M}$.
+\item Let $\kappa$ be a cardinal, then $\mathcal{M}$ is $\kappa$-saturated, if for every $S \subseteq \mathcal{M}$ with $|S| < \kappa$ every finitely satisfied $1$-type $\mathcal{T}$ is realised in $\mathcal{M}$.
+\end{enumerate}
+\end{defin}
+
+By default, a saturated model is an $\omega$-saturated model for us.
+
+The useful facts, they are from \cite{chang1990model} and \cite{hodges1993model}:
+
+\begin{fact} Let $\mathcal{M}$ be an FO-structue and $\kappa$ a cardinal, then:
+\begin{enumerate}
+\item $\mathcal{M}$ is $\kappa$-saturated, iff every finitely satisfiable $\alpha$-type (an arbitrary $\alpha \leq \kappa$) with fewer than $\kappa$ parameters is realised in  $\mathcal{M}$.
+\item If $\mathcal{M}$ is $\kappa$-saturated, then $\mathcal{M}$ is $\lambda$-saturated for every $\lambda < \kappa$.
+\item \label{saturation} Every consistent theory has a $\kappa$-saturated model and every model has an elementary $\kappa$-saturated extension.
+\item Let $(\mathcal{M}_i)_{i < \omega}$ a family of structures of the (at most) countable signature and $D$ a non-principal ultrafilter over $\omega$, then $\Pi_D \mathcal{M}_i$ is $\omega_1$-saturated.
+\end{enumerate}
+\end{fact}
+
+\subsection{Proof of Theorem~\ref{finitecanon}}
+
+Let $\mathcal{A} \in {\bf CA}_n$, then if $\mathcal{A}$ is completely representable, then $h$, a complete representation of $\mathcal{A}$, is atomic.
+That is, $(a_1, \dots, a_n) \in h(1)$, then $(a_1, \dots, a_n) \in h(y)$ for some $y \in \operatorname{At}(\mathcal{A})$.
+
+\begin{defin} \label{theory} Let $\mathcal{A}$ be a cylindric algebra of dimension $n < \omega$.
+$L(\mathcal{A})$ is the first-order language that consists of equality plus $n$-ary predicate letters $(R^n_a)_{a \in \mathcal{A}}$.
+The $L(\mathcal{A})$-theory $T_{\mathcal{A}}$ consists of the following sentences:
+\begin{enumerate}
+\item $A_+(a,b,c) := \forall x_1, \dots, x_n \: (R_a(x_1, \dots, x_n) \leftrightarrow R_b(x_1, \dots, x_n) \lor R_c(x_1, \dots, x_n))$. Informally, that means $\mathcal{A} \models a = b + c$.
+\item $A_{-}(a,b) := \forall x_1, \dots, x_n \: (R_a(x_1, \dots, x_n) \leftrightarrow \neg R_b(x_1, \dots, x_n))$. That is, $\mathcal{A} \models a = - b$.
+\item $A_{\neq 0}(a) := \exists x_1, \dots, x_n R_a(x_1, \dots, x_n)$. That is, $\mathcal{A} \models a \neq 0$.
+\item $A_{c_i}(a) := \forall x_1, \dots, x_n (R_{c_i a}(x_1, \dots, x_n) \leftrightarrow \exists y_1, \dots y_n (R_a(y_1, \dots, y_n) \land x_i = y_j))$, for $i < n$ and $j < n$ such that $i \neq j$.
+Informally, $\mathcal{A} \models {c_i} a = 1$.
+\item $A_{d_{ij}} := \forall x_1, \dots, x_n (R_{d_{ij}}(x_1, \dots, x_n) \leftrightarrow x_i = x_j)$, for $i, j < n$.
+\end{enumerate}
+\end{defin}
+
+In fact, we need to show the following implication:
+\begin{center}
+If $\mathcal{A}$ is representable, then $A^{+}$ is completely representable.
+\end{center}
+
+Assume that $\mathcal{A}$ is representable, then the theory $T(\mathcal{A})$ is consistent, then it has an $\omega$-saturated model $\mathcal{M}$ by Fact~\ref{saturation}. We have the following claim:
+\begin{claim}
+The set $U_{x_1,\dots,x_n} = \{ a \in \mathcal{A} \: | \: \mathcal{M} \models R_a(x_1,\dots,x_n)\}$ is an ultrafilter of $\mathcal{A}$, for $x_1,\dots,x_n \in \mathcal{M}$ with $R_{1}(x_1,\dots,x_n)$.
+\end{claim}
+Those $U_{x_1,\dots,x_n}$'s allow us to represent atoms of $\mathcal{A}^{+}$.
+
+We define a representation of $\mathcal{A}^{+}$ as a map $h : \mathcal{A}^{+} \to 2^{\mathcal{M}^n}$ such that:
+\begin{center}
+$h : S \mapsto \{ (x_1, \dots, x_n) \in 1^{\mathcal{M}} \: | \: U_{x_1,\dots,x_n} \in S \}$, for $S \in \operatorname{Spec}(\mathcal{A})$.
+\end{center}
+
+\begin{claim} Let $A_1, A_2 \in \operatorname{Spec}(\mathcal{A})$
+\begin{enumerate}
+\item $h(0^{\mathcal{A}^{+}}) = \emptyset$
+\item $h(- A_1) = - h(A_1)$
+\item $h(1^{\mathcal{A}^{+}}) = 1^{\mathcal{M}}$
+\item If $S \subseteq \operatorname{Spec}(\mathcal{A})$, then $h(\bigcup S) = \bigcup \limits_{U \in S} h(U)$
+\end{enumerate}
+In particular, $h$ is a Boolean homomorphism.
+\end{claim}
+
+\begin{proof}
+$ $
+
+\begin{enumerate}
+\item $h(0^{\mathcal{A}^{+}}) = h(\emptyset) = \emptyset$.
+\item From the definiton of $h$.
+\item $h(- A_1) = - h(A_1)$
+
+Let $x_1, \dots, x_n \in 1^{\mathcal{M}}$, then we have:
+
+\begin{center}
+$(x_1, \dots, x_n) \in h(- A_1)$ iff $U_{x_1, \dots, x_n} \in - A_1$ iff $U_{x_1, \dots, x_n} \notin A_1$ iff $(x_1, \dots, x_n) \notin h(A_1)$
+\end{center}
+\item Let $S = \bigcup \limits_{i \in I} S_i$, where $S_i \in \operatorname{Spec}(\mathcal{A})$ for every $i \in I$.
+Let $(x_1, \dots, x_n) \in 1^{\mathcal{M}}$, then we have:
+\begin{center}
+$(x_1, \dots, x_n) \in h(\bigcup \limits_{i \in I} S_i)$ iff $f_{x_1, \dots, x_n} \in \bigcup \limits_{i \in I} S_i$ iff $\exists i \in I \:\: f_{x_1, \dots, x_n} \in S_i$ iff $\exists i \in I \:\: (x_1, \dots, x_n) \in h(S_i)$ iff $(x_1, \dots, x_n) \in \bigcup \limits_{i \in I} S_i$
+\end{center}
+\end{enumerate}
+\end{proof}
+
+\begin{claim}
+$h$ is injective.
+\end{claim}
+
+\begin{proof}
+Let $U \in \operatorname{Spec}(\mathcal{A})$. The first is to show that $h(U)$ is non-empty. The following $n$-type:
+\begin{center}
+$T(x_1, \dots, x_n) = \{ R_a(x_1, \dots, x_n) \: | \: a \in U \}$
+\end{center}
+if finitely satisfied in $\mathcal{M}$.
+
+Consider $T_0 = \{ R_{a_1}(x_1, \dots, x_n), \dots, R_{a_k}(x_1, \dots, x_n) \} \subseteq T$.
+Then $a_1, \dots, a_k \in U$ and $a = a_1 \cdot \dots \cdot a_k \in U$.
+By the instance of the $A_{\neq 0}(a)$-axiom, we have $\mathcal{M} \models \exists x_1, \dots, x_n R_a(x_1, \dots, x_n)$.
+$a \leq a_i$ for $i \leq k$, so we have $\mathcal{M} \models \exists x_1, \dots, x_n R_{a_i}(x_1, \dots, x_n)$ for every $a_i$ with $i \leq k$ by the instance of the $A_{+}(a_i, a, a)$-axiom.
+That makes every finite subtype of $T$ satisfiable, thus the whole type is finitely satifiable in $\mathcal{M}$.
+$\mathcal{M}$ is $\omega$-saturated, then $T$ is realised in $\mathcal{M}$ by some $(x_1, \dots, x_n) \in \mathcal{M}^n$ and, moreover, $\mathcal{M} \models 1(x_1, \dots, x_n)$.
+As we have already said, $U_{x_1, \dots, x_n}$ is an ultrafilter, but $U_{x_1, \dots, x_n} \subseteq U$, thus $U = U_{x_1, \dots, x_n}$, so $(x_1, \dots, x_n) \in h(U)$.
+
+That makes $h$ one-to-one.
+\end{proof}
+
+\begin{claim}
+$ $
+
+\begin{enumerate}
+\item $h(c_i{^{\mathcal{A}^{+}}} U) = C_i (h(U))$
+\item $h(d_{ij}^{\mathcal{A}^{+}}) = D_{ij} \subseteq \operatorname{Spec}(\mathcal{A})$
+\end{enumerate}
+\end{claim}
+
+\begin{proof}
+$ $
+
+\begin{enumerate}
+\item Let $\overline{x} = (x_1, \dots, x_n) \in \mathcal{M}^{n}$ and $S \subseteq \operatorname{Spec}(\mathcal{A})$. Assume $(x_1, \dots, x_n) \in h(c_i{^{\mathcal{A}^{+}}} S)$.
+
+Let us show that $\overline{x} \in C_i (h(S))$, that is, there exists $\overline{y} = (y_1, \dots, y_n) \in h(S)$ such that $\overline{x} \equiv_{i} \overline{y}$.
+
+Then $\mathcal{M} \models 1(x_1, \dots, x_n)$ and $U_{x_1, \dots x_n} \in c_i{^{\mathcal{A}^{+}}} S$. But $\mathcal{A}^{+}$ is the complex algebra of the ultrafilter frame $\mathcal{F}_{\mathcal{A}}$.
+Then we have:
+\begin{center}
+$c_i{^{\mathcal{A}^{+}}} S = \{ U_1 \in \operatorname{Spec}(\mathcal{A}) \: | \: \exists U' \in S  \: U_1 R_{i} U' \}$
+\end{center}
+Then there must be an ultrafilter $U' \in S$ such that $U_{x_1, \dots x_n} R_{i} U'$, that is, $c_i a \in U_{x_1, \dots x_n}$ whenever $a \in U'$.
+Hence $\mathcal{M} \models R_{c_i}(x_1, \dots x_n)$. By the $A_{c_i}(a)$-axiom, we have
+\begin{center}
+$\mathcal{M} \models \exists z_1, \dots, z_n (R_a(z_1, \dots, z_n) \land x_i = z_j)$ for $i < n$ and $j < n$ such that $i \neq j$.
+\end{center}
+Consider the following $n$-type with free variables $z_1, \dots, z_n$ and parameters $x_1, \dots, x_n \in \mathcal{M}$:
+\begin{center}
+$T(z_1, \dots, z_n) = \{ R_a(z_1, \dots, z_n) \land x_i = z_j \: | \: i < n, j < n, i \neq j, a \in U' \}$.
+\end{center}
+Let us show that $T(z_1, \dots, z_n)$ is finitely satifsiable in $\mathcal{M}$.
+Consider a finite subset of $T$, say $T_0 = \{ R_{b_k}(z_1, \dots, z_n) \land x_i = y_j \: | \: i < n, j < n, i \neq j, b_k \in U', k < \omega \}$.
+We put $p = p_1 \cdot \dots \cdot p_k$ and $p \in U'$ since $U'$ is a filter. Then we have:
+\begin{center}
+$\mathcal{M} \models \exists z_1, \dots, z_n (R_b(z_1, \dots, z_n) \land x_i = z_j)$ for $i < n$ and $j < n$ such that $i \neq j$
+\end{center}
+Thus, we have, as required:
+\begin{center}
+$\mathcal{M} \models \exists z_1, \dots, z_n \bigwedge \limits_{i = 1}^{k} (R_{b_k}(z_1, \dots, z_n) \land x_i = z_j)$ for $i < n$ and $j < n$ such that $i \neq j$.
+\end{center}
+As above, using $\omega$-saturation, we conclude that $T$ is realised in $\mathcal{M}$ at an $n$-tuple $(y_1, \dots, y_n) = \overline{y}$.
+Then we have:
+\begin{center}
+$\mathcal{M} \models 1(\overline{y})$, $\overline{x} \equiv_i \overline{y}$, $U_{\overline{y}} \supseteq U'$
+\end{center}
+Then $U_{\overline{y}} = U'$, then $\overline{y} \in h(S)$. Then $\overline{x} \in C_i (h (S))$.
+
+Suppose for the converse, $\overline{x} = (x_1, \dots, x_n) \in C_i (h(S))$. We need $\overline{x} \in h(c_i(S))$.
+Then there exists $\overline{y} = (y_1, \dots, y_n)$ such that $\overline{x} \equiv_i \overline{y}$ and $\overline{y} \in h(S)$.
+Then there exists an ultrafilter $U_{y_1, \dots, y_n} \in S$.
+Let us show that $\mathcal{M} \models 1(x_1, \dots, x_n)$ and $U_{x_1, \dots, x_n} \in c_i U_{y_1, \dots, y_n}$.
+Let $a \in U_{y_1, \dots, y_n}$.
+Then we have $\mathcal{M} \models R_a(y_1, \models, y_n)$.
+By the $A_{c_i}(a)$ axiom, we have $\mathcal{M} \models R_{c_i a}(x_1, \dots, x_n)$.
+Then $\mathcal{M} \models 1(x_1, \dots, x_n)$ and $c_i a \in U_{x_1, \dots, x_n}$, thus, $\overline{x} \in h(c_i(S))$.
+\item Let us show that $h$ preserves cylindrifications.
+
+Let $(x_1, \dots, x_n) \in \mathcal{M}^n$.
+Then $(x_1, \dots, x_n) \in D_{ij}$ iff $\mathcal{M} \models 1(x_1, \dots, x_n)$ and $x_i = x_j$ iff $U_{x_1, \dots, x_n} \in d_{ij}^{\mathcal{A}^{+}} = \{ U \in \operatorname{Spec}(\mathcal{A}) \: | \: d_ij \in U \}$ iff $\mathcal{M} \models d_{ij}^{\mathcal{M}}(x_1, \dots, x_n)$.
+\end{enumerate}
 \end{proof}

### 2023-01-08 varieties-reducts/Text.tex
add discriminator stuff

\documentclass[a4paper]{article}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{bussproofs}
\usepackage{mathtools}
\usepackage{verbatim}
\usepackage{dsfont}
\usepackage{mathabx}
\usepackage[all, 2cell]{xy}
\usepackage[all]{xy}
\usepackage{wasysym}
\usepackage{rotating}
\usepackage{geometry}
\usepackage{trfsigns}
\usepackage{cmll}
\usepackage{authblk}
\usepackage{hyperref}
\usepackage{cleveref}
\usepackage{lipsum}
\usepackage{extpfeil}
\usepackage{soul}
\usepackage{graphicx}

\newcommand\mapsfrom{\mathrel{\reflectbox{\ensuremath{\mapsto}}}}

\theoremstyle{defin}
\newtheorem{definition}{Definition}

\theoremstyle{theorem}
\newtheorem{theorem}{Theorem}

\theoremstyle{claim}
\newtheorem{claim}{Claim}

\theoremstyle{prop}
\newtheorem{prop}{Proposition}

\theoremstyle{lemma}
\newtheorem{lemma}{Lemma}

\theoremstyle{fact}
\newtheorem{fact}{Fact}

\theoremstyle{ex}
\newtheorem{ex}{Example}

\theoremstyle{col}
\newtheorem{col}{Corollary}

\let\strokeL\L
\DeclareRobustCommand{\L}{\ifmmode\mathbf{L}\else\strokeL\fi}

\author{Daniel Rogozin}
\date{}
\title{Varieties of representable relation algebra reducts}

\begin{document}
\maketitle

\nocite{*}

\section{Varieties and discriminators}

Standardly, a class of algebras is called \emph{variety}, if it can be determined by some equational theory, or, equivalently, it is closed under homomorphic images, subalgebras and direct products. Given a class of algebras $\mathcal{K}$, ${\bf V}(\mathcal{K})$ is a variety generated by $\mathcal{K}$ or, equivalently, ${\bf HSP}(\mathcal{K})$, the closure of $\mathcal{K}$ under homomorphic images, subalgebras and direct products.

Let $\{ A_i \: | \: i \in I \}$ be an indexed family of algebras, then a subalgebra $A \subseteq \Pi_{i \in I} A_i$ is a subdirect product if $\pi_i(A) = A_i$. An embedding $\alpha : A \to \Pi_{i \in I} A_i$ is subdirect if $\alpha(A)$ is subdirect product. An algebra is subdirectly irreducible if for every subdirect embedding $\alpha : A \to \Pi_{i \in I} A_i$ there exists $i \in I$ such that $\pi_i \circ \alpha : A \to A_i$ is an isomorphism.

An equivalence relation $\theta$ on an algebra $A$ is called \emph{congruence}, if $\theta$ respects any operation. ${\bf Con}(A)$ is the set of all congruences on $A$. An algebra is called \emph{simple}, if ${\bf Con}(A/\theta) = \{ \Delta, \nabla \}$, where $\Delta$ and $\nabla$ are trivial congruences. One can obtain a simple algebra by factorising it through the maximal congruence \cite[Theorem 8]{sankappanavar1981course}.

One can equivalently define subdirectly irreducible algebras using congruences as follows. an algebra is subdirectly irreducible iff either $A$ is trivial or there exists a minimal congruence in ${\bf Con}(A) \setminus \{ \Delta \}$.

Recall that a Stone space is a compact Hausdorff zero-dimensional topological space. A subdirect product $A \subseteq \Pi_{x \in X} A_x$ over a Stone space $X$ if
\begin{enumerate}
\item for all $a, b \in A$ $\{ x \in X \: | \: a(x) = b(x) \}$ is clopen.
\item for all $a, b \in A$ and for all clopen $Y \subseteq X$ $a \upharpoonright_{Y} \cup a \upharpoonright_{X \setminus Y} \in A$.
\end{enumerate}

A variety $\mathcal{V}$ is \emph{arithmetical}, if it is congruence-permutable and congruence-distributive, or, equivalently, there exists a ternary term $p$ such that:
\begin{equation}
\mathcal{V} \models p(x,y,x) \approx p(x,y,y) \approx p(y,y,x) \approx x
\end{equation}

A ternary term $t(x, y, z)$ for an algebra $A$ if, for all $a, b, c \in A$:

\begin{equation}
t(a,b,c) =
  \begin{cases}
  c & \text{if $a = b$} \\
  a & \text{otherwise}
  \end{cases}
\end{equation}

A variety $\mathcal{V}$ is called discriminator if there exists a class $\mathcal{K}$ such that $\mathcal{V} = {\bf V}(\mathcal{K})$ and there exists a term $t(x,y,z)$, which is a discriminator term for every member of $\mathcal{K}$. It is known that if an algebra $A$ has a discriminator term, then $A$ is simple \cite[Lemma 9.2]{sankappanavar1981course}. Moreover, we have the following property of discriminator terms, see \cite[Theorem 9.4]{sankappanavar1981course}.

\begin{theorem} Let $t(x,y,z)$ be a discriminator term for every member of a class $\mathcal{K}$:

\begin{enumerate}
\item ${\bf V}(\mathcal{K})$ is an arithmetical variety.
\item Every indecomposable member of ${\bf V}(\mathcal{K})$ is simple.
\item Simple algebras are precisely members of ${ISP}_U(\mathcal{K}_+)$.
\item Every member of ${\bf V}(\mathcal{K})$ is isomorphic to a Boolean product of simple algebras.
\end{enumerate}
\end{theorem}

\section{BAOs, relation algebras and their reducts}

\subsection{Discriminator varieties of BAOs}

Let $B$ be a Boolean algebra, an operator is an $n$-ary function $f : A^n \to A$ such that, for all $x_1, \dots, x_n, x, y \in B$:

\begin{itemize}
\item $f(x_1, \dots, x + y, \dots, x_n) = f(x_1, \dots, x, \dots, x_n) + f(x_1, \dots, y, \dots, x_n)$
\item $f(x_1, \dots, 0, \dots, x_n) = 0$
\end{itemize}

A Boolean algebra with operators is an algebra $M = (B, (f_i)_{i \in I})$, where each $f_i$ is an operator.

In the case of BAOs, one can define discriminator simpler, as an unary term $d(x)$ such that, for all $a \in M$, where $M$ is a BAO:
\begin{equation}
d(x) =
\begin{cases}
  0 & \text{if $x = 0$} \\
  1 & \text{otherwise}
\end{cases}
\end{equation}

One can characterise discriminator varieties as follows, see \cite[Lemma 2.1]{andreka1998notions}:
\begin{theorem} Let $\mathcal{V}$ be a variety of BAOs and $d(x)$ a unary term, then the following are equivalent:

\begin{enumerate}
\item $d$ is a discriminator variety.
\item The following equations are valid in $\mathcal{V}$:
\begin{enumerate}
\item $x \leq d(x)$
\item $d(d(x)) \leq d(x)$
\item $d(- d(x)) \leq - d(x)$
\item $f(x_0, \dots, x_{n-1}) \leq d(x_i)$ for all $n > 0$ and for every operator $f$ of $M$
\end{enumerate}
\end{enumerate}
\end{theorem}

\subsection{Relation algebras and their reducts}

In this subsection, we consider relation algebras, a kind of BAOs.

\begin{definition}
  $ $

    A relation algebra is an algebra $\mathcal{R} = (R, 0, 1, +, -, ;, {}^{\smile}, {\bf 1'})$ such that $(R, 0, 1, +, -)$ is a Boolean algebra and the following hold:
    \begin{enumerate}
      \item $(R, ;, {\bf 1'})$ is a monoid
      \item $(a + b) ; c = (a ; c) + (b ; c)$
      \item $a^{\smile \smile} = a$
      \item $(a + b)^{\smile} = a^{\smile} + b^{\smile}$
      \item $(a ; b)^{\smile} = b^{\smile} ; a^{\smile}$
      \item $a^{\smile} ; (- (a ; b)) \leq - b$
    \end{enumerate}
    where $a \leq b$ iff $a + b = b$. ${\bf RA}$ denotes the class of all relation algebras.
\end{definition}

\begin{definition}
    A proper relation algebra is an algebra $\mathcal{R} = (R, \emptyset,W, \cup, -, |, {}^{\smile}, {\bf 1 })$ such that $R \subseteq \mathcal{P}(W)$, where $W \subseteq X \times X$ is an equivalence relation; $|$ is relation composition, ${}^{\smile}$ is relation converse, ${\bf Id}$ is a
    diagonal subset of $W$, that is:
    \begin{enumerate}
    \item $a | b = \{ (x, z) \: | \: \exists y \: (x, y) \in a \: \& \: (y, z) \in b \}$
    \item $a^{\smile} = \{ (x, y) \: | \: (y, x) \in a \}$
    \item ${\bf Id} = \{ (x, y) \: | \: x = y \}$
    \end{enumerate}
      The class of all proper relation algebras is denoted as ${\bf PRA}$. ${\bf Rs}$ is the class of all relation set algebras, proper relation algebra with a diagonal subrelation as an identity. ${\bf RRA}$ is the class of all representable relation algebras, that is, the closure of ${\bf PRA}$ under isomorphic copies. That is,
      ${\bf RRA} = {\bf IPRA}$.
\end{definition}

\subsection{Varieties}
\subsection{Non-varieties}
\subsection{Unknown}

\section{Decidability aspects}
\subsection{Current results}
\subsection{Problems}

\bibliographystyle{alpha}
\bibliography{Text}
\end{document}

### 2023-02-05 varieties-reducts/Text.tex
residuated BAOs

@@
 \nocite{*}
+
+\section{Question}
+
+Let $\tau$ be a subsignature of operations expressible in the signature of $RA$. ${\bf R}(\tau)$ is a variety iff ??.
+
+Investigate discriminators for representable relation algebra reducts.
 
@@
 ${\bf RRA}$ is the class of all representable relation algebras, that is, the closure of ${\bf PRA}$ under isomorphic copies.
-That is,
-      ${\bf RRA} = {\bf IPRA}$.
 \end{definition}
@@
 \subsection{Varieties}
+
+\subsubsection{Boolean algebras with residuated operators}
+
+Let $\mathcal{A}_0$ be a Boolean algebra, a unary operation on $f$ is called \emph{residuated} if there exists a \emph{residual} operation $g$ such that for all $a, b \in \mathcal{A}_0$:
+\begin{center}
+$f(a) \leq b$ iff $a \leq f(b)$
+\end{center}
+Equivalently, $f$ is residuated if there exists a conjugate operation $h$ such that for all $a, b \in \mathcal{A}_0$:
+\begin{center}
+$f(a) \cdot b = 0$ iff $a \land h(b) = 0$
+\end{center}
+$f$ is \emph{self-conjugate} whenever $f$ is equal to its conjugate operation.
+
+\begin{theorem}
+Let $f$ be a residuated operator on $\mathcal{A}_0$, then
+
+\begin{enumerate}
+\item $f$ is normal and completely additive,
+\item $f$ and $h$ are conjugate operations on $\mathcal{A}_0$ iff they are normal and the following holds for all $a, b \in \mathcal{A}_0$
+\begin{center}
+$f(a) \cdot b \leq f(a \cdot h(b))$ and $a \cdot h(b) \leq h(f(a) \cdot b)$
+\end{center}
+\end{enumerate}
+\end{theorem}
+Two $n$-ary operators $f$ and $h$ are conjugate in the $i$-th argument if $f_{\vec{a}, i}$ is conjugate to $g_{\vec{a}, i}$.
+Let $\mathcal{A}$ be a BAO, $\mathcal{A}$ is called a \emph{Boolean algebra with residuated operators} if for each operator $f$ of $\mathcal{A}$ and for each $i < ar(f)$ there exists an $ar(f)$-ry term $t$, which is conjugate to $f$ in the $i$-th argument.
+${\bf RA}$ is an example, where $\smile$ is self-conjugate and composition has conjugate terms $x ; y^{\smile}$ and $x^{\smile} ; y$.
+
+An $r$-algebra is a Boolean algebra with three residuated binary operations $\bullet$, $\triangleright$ and $\triangleleft$, where $\triangleright$ and $\triangleleft$ are right and left conjugates to $\bullet$, that is, the following conditions are equivalent:
+\begin{enumerate}
+\item $(x \bullet y) \cdot z = 0$
+\item $(x \triangleright z) \cdot y = 0$
+\item $(z \triangleleft y) \cdot x = 0$
+\end{enumerate}
+
+\subsubsection{Residuated monoids}
+
+\subsubsection{Positive relation algebras}
+
+\subsubsection{Domain algebras}
+
 \subsection{Non-varieties}

### 2023-02-06 residuated-semigroups/Text.tex
lambek and residuated stuff

\documentclass[a4paper]{article}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{bussproofs}
\usepackage{mathtools}
\usepackage{verbatim}
\usepackage{dsfont}
\usepackage{mathabx}
\usepackage[all, 2cell]{xy}
\usepackage[all]{xy}
\usepackage{wasysym}
\usepackage{rotating}
\usepackage{geometry}
\usepackage{trfsigns}
\usepackage{cmll}
\usepackage{authblk}
\usepackage{hyperref}
\usepackage{cleveref}
\usepackage{lipsum}
\usepackage{extpfeil}
\usepackage{soul}
\usepackage{graphicx}

\newcommand\mapsfrom{\mathrel{\reflectbox{\ensuremath{\mapsto}}}}

\theoremstyle{defin}
\newtheorem{definition}{Definition}

\theoremstyle{theorem}
\newtheorem{theorem}{Theorem}

\theoremstyle{claim}
\newtheorem{claim}{Claim}

\theoremstyle{prop}
\newtheorem{prop}{Proposition}

\theoremstyle{lemma}
\newtheorem{lemma}{Lemma}

\theoremstyle{fact}
\newtheorem{fact}{Fact}

\theoremstyle{ex}
\newtheorem{ex}{Example}

\theoremstyle{col}
\newtheorem{col}{Corollary}

\let\strokeL\L
\DeclareRobustCommand{\L}{\ifmmode\mathbf{L}\else\strokeL\fi}

\author{Daniel Rogozin}
\date{}
\title{Finite model property for residuated semigroups and related remarks}

\begin{document}
\maketitle

\nocite{*}

\section{Finite networks for atomic formulas}

Let $PV = \{ p_i \: | \: i < \omega \}$ be the set of propositional variables (or atomic types).
The set of formulas is generated by the following grammar:
\begin{center}
$\varphi, \psi ::= p \: | \: (\varphi \bullet \psi) \: | \: (\varphi \backslash \psi) \: | \: (\varphi / \psi)$
\end{center}

The Lambek calculus is defined as a Gentzen-style sequent calculus:

\begin{prooftree}
\AxiomC{$ $}
\RightLabel{\scriptsize{ax}}
\UnaryInfC{$p \rightarrow p$}
\end{prooftree}

\begin{minipage}{0.5\textwidth}
\begin{flushleft}
      \begin{prooftree}
    \AxiomC{$\Gamma \rightarrow \varphi$}
    \AxiomC{$\Delta, \psi, \Theta \rightarrow \theta$}
    \RightLabel{$\backslash \rightarrow$}
    \BinaryInfC{$\Delta, \Gamma, \varphi \backslash \psi, \Theta \rightarrow \theta$}
  \end{prooftree}

  \begin{prooftree}
    \AxiomC{$\Gamma \rightarrow \varphi$}
    \AxiomC{$\Delta, \psi, \Theta \rightarrow \theta$}
    \RightLabel{$/ \rightarrow$}
    \BinaryInfC{$\Delta, \psi / \varphi, \Gamma, \Theta \rightarrow \theta$}
  \end{prooftree}

  \begin{prooftree}
    \AxiomC{$\Gamma, \varphi, \psi, \Delta \rightarrow \theta$}
    \RightLabel{$\bullet \rightarrow$}
    \UnaryInfC{$\Gamma, \varphi \bullet \psi, \Delta \rightarrow \theta$}
  \end{prooftree}
\end{flushleft}
\end{minipage}\hfill
\begin{minipage}{0.5\textwidth}
\begin{flushright}
  \begin{prooftree}
    \AxiomC{$\varphi, \Pi \rightarrow \psi$}
    \RightLabel{$\rightarrow \backslash, \Pi \text{ is non-empty}$}
    \UnaryInfC{$\Pi \rightarrow \varphi \backslash \psi$}
  \end{prooftree}

  \begin{prooftree}
    \AxiomC{$\Pi, \varphi \rightarrow \psi$}
    \RightLabel{$\rightarrow /, \Pi \text{ is non-empty}$}
    \UnaryInfC{$\Pi \rightarrow \psi / \varphi$}
  \end{prooftree}

  \begin{prooftree}
    \AxiomC{$\Gamma \rightarrow \varphi$}
    \AxiomC{$\Delta \rightarrow \varphi$}
    \RightLabel{$\rightarrow \bullet$}
    \BinaryInfC{$\Gamma, \Delta \rightarrow \varphi \bullet \psi$}
  \end{prooftree}
\end{flushright}
\end{minipage}

\begin{theorem}
Let $RS$ be the class of all residuated semigroups, then $\Gamma \rightarrow \varphi$ iff $RS \models \Gamma \rightarrow \varphi$
\end{theorem}

\section{Representability networks for at most countable residuated semigroups}

Let $\mathcal{A}$ be a residuated semigroup, an $\mathcal{A}$-network is a triple $\mathcal{N} = (V, E, l)$, where
\begin{itemize}
\item $(V, E)$ is a directed graph with no loops and $E$ is transitive
\item $l : E \to \mathcal{A}$ such that
\begin{itemize}
\item $l(x, z) \leq l(x, y) ; l(y,z)$, whenever $(x, y), (y, z) \in E$, for all $x, y, z \in V$.
\item 
\end{itemize}
\end{itemize}

Let $n \leq \omega$

\section{Games for the FMP}

\bibliographystyle{alpha}
\bibliography{Text}
\end{document}

### 2023-02-15 residuated-semigroups/Text.tex
representability criterion for residuated semigroups and a half of the definition of a game

@@
 
+\subsection{Completeness}
+
 \begin{theorem}
@@
 
-Let $\mathcal{A}$ be a residuated semigroup, an $\mathcal{A}$-network is a triple $\mathcal{N} = (V, E, l)$, where
+\subsection{Relational residuated semigroups as Kripke models}
+
+One can introduce Kripke-style relational semantics for the Lambek calculus as follows.
+Let $W$ be a non-empty set and $R$ a transitive relation on $W$.
+We consider models of the kind $M = (R, \vartheta)$, where $\vartheta : PV \to 2^R$.
+The truth definition is inductive:
 \begin{itemize}
-\item $(V, E)$ is a directed graph with no loops and $E$ is transitive
-\item $l : E \to \mathcal{A}$ such that
-\begin{itemize}
-\item $l(x, z) \leq l(x, y) ; l(y,z)$, whenever $(x, y), (y, z) \in E$, for all $x, y, z \in V$.
-\item 
-\end{itemize}
+\item $\mathcal{M}, (x,y) \models p_i$ iff $(x, y) \in \vartheta(p_i)$,
+\item $\mathcal{M}, (x,y) \models \varphi \bullet \psi$ iff there exists $z \in W$ such that
+$(x,z), (z, y) \in R$ and $\mathcal{M}, (x, z) \models \varphi$ and $\mathcal{M}, (z, y) \models \phi$
+\item $\mathcal{M}, (x, y) \models \varphi \setminus \psi$ iff for all $z \in W$ such that if $(z, x) \in R$ and $\mathcal{M}, (z, x) \models \varphi$, then $\mathcal{M}, (z, y) \models \psi$
+\item $\mathcal{M}, (x, y) \models \varphi / \psi$ iff for all $z \in W$ such that if $(y, z) \in R$ and $\mathcal{M}, (y, z) \models \psi$, then $\mathcal{M}, (x, z) \models \varphi$
+\item $\mathcal{M}, (x, y) \models \varphi_1, \varphi_2, \dots, \varphi_n \rightarrow \varphi$ iff $\mathcal{M}, (x, y) \models \varphi_1 \bullet \varphi_2 \bullet \dots \bullet \varphi_n$ implies $\mathcal{M}, (x, y) \models \varphi$.
 \end{itemize}
 
-Let $n \leq \omega$
+According to the definition above, to refute a sequent $\varphi_1, \varphi_2, \dots, \varphi_n \rightarrow \varphi$, we have to find a transitive binary relation $R$, some valuation $\vartheta : PV \to 2^R$ and $(x, y) \in R$ such that
+$\mathcal{M}, (x, y) \models \varphi_1 \bullet \varphi_2 \bullet \dots \bullet \varphi_n$, but $\mathcal{M}, (x, y) \nvDash \varphi$. Alternatively, one can reformulate that condition as
+\begin{center}
+$(x, y) \in ||\varphi_1|| ; ||\varphi_2|| ; \dots ; ||\varphi_n||$, but $(x, y) \notin ||(x, y)||$
+\end{center}
+
+\subsection{Relational representation of residuated semigroups: a game-theoretic approach}
+
+Let $\mathcal{A}$ be a residuated semigroup, an $\mathcal{A}$-prenetwork is a triple $\mathcal{N} = (V, E, l)$, where where $(V, E)$ is a directed graph and $l : E \to \mathcal{A}$ is a labelling function.
+A prenetwork is a network if
+the following conditions hold:
+\begin{itemize}
+\item $E$ has no loops and it is transitive,
+\item $l(x, z) \leq l(x, y) ; l(y,z)$, whenever $(x, y), (y, z) \in E$, for all $x, y, z \in V$,
+\item For all $a \in \mathcal{A}$, for all $x \in U$, there is some $u \in U$ such that $l(u, x) = a$,
+\item For all $a \in \mathcal{A}$, for all $y \in U$, there is some $v \in U$ such that $l(y, v) = a$,
+\item For all $a, b, c \in \mathcal{A}$, for all $x, y \in U$, if $c \leq a ; b$, $(x, y) \in E$ and $l(x, y) = c$, then there exists $z \in U$ such that $l(x,y) = a$ and $l(y, z) = b$.
+\end{itemize}
+
+Let $n \leq \omega$, define a game $\mathcal{G}(\mathcal{A})_n$ for two players $\forall$ and $\exists$ by induction on $n$.
+\begin{enumerate}
+\item {\bf step} 0
+
+$\forall$ picks a pair of elements $a, b \in \mathcal{A}$ such that $a \not\leq b$. $\exists$ must respond with a network $\mathcal{N}_0 = (\{ x, y\}, \{(x,y)\}, l_0 : (x,y) \mapsto a)$:
+
+\xymatrix{
+&&&&&& x \ar[rr]^{a} && y
+}
+
+\item {\bf step} $n + 1 < \omega$
+
+Suppose the networks:
+\begin{center}
+$\mathcal{N}_0 \subseteq \mathcal{N}_1 \subseteq \dots \subseteq \mathcal{N}_n$
+\end{center}
+have been already constructed.
+
+There are four different options:
+\begin{enumerate}
+\item {\bf Composition move}
+
+$\forall$ picks $x, y, z \in \mathcal{N}_n$ such that $b = l_n(x, y)$ and $c = l_n(y, z)$:
+
+\xymatrix{
+&&&&&&& y \ar[dr]^{c} \\
+&&&&&& x \ar[ur]^{b} && z
+}
+
+$\exists$ has to respond with $\mathcal{N}_{n + 1} = (V_n, E_n \cup \{ (x,z)\}, l_{n+1})$ where $l_{n+1}(x,z) = b ; c$ and $l_{n+1}(x',y') = l_n(x',y')$ for $(x', y') \in E_n$.
+
+\xymatrix{
+&&&&&&& y \ar[dr]^{c} \\
+&&&&&& x \ar@{-->}[rr]_{b ; c} \ar[ur]^{b} && z
+}
+
+\item {\bf Witness move}
+$\forall$ picks $(x,y) \in E_n$ such that $l_n(x,y) = a$ and $a \leq b ; c$:
+
+\xymatrix{
+&&&&&& x \ar[rr]^{a} && y
+}
+
+$\exists$ has to respond with $\mathcal{N}_{n + 1} = (V_n \cup \{ z \}, E_n \cup \{ (x,z), (z, y) \}, l_{n + 1})$, where
+\begin{center}
+$l_{n + 1}(x, z) = b$
+
+$l_{n + 1}(y, z) = c$
+
+$l_{n + 1}(p) = l_n$ for others $p \in E_n$
+\end{center}
+
+The latter can be visualised with the following triangle:
+
+\xymatrix{
+&&&&&& x \ar[rr]^{a} \ar@{-->}[dr]_{b} && y \\
+&&&&&&& z \ar@{-->}[ur]_{c}
+}
+
+\item {\bf Left redisual move}
+\item {\bf Right residual move}
+\end{enumerate}
+\end{enumerate}
+
+\begin{theorem}
+
+Let $\mathcal{A}$ be a at most countable residuated semigroup, then
+
+\begin{enumerate}
+\item $\exists$ has a winning stragery in $\mathcal{G}_{\omega}(\mathcal{A})$
+\item $\mathcal{A}$ is representable.
+\end{enumerate}
+\end{theorem}
+
+TODO: check if the representability class is closed under products, subalgebras and ultraproducts. Check the criterion for the Horn formulas. Closed under H? can't be defined by equations?
 

### 2023-03-12 residuated-semigroups/Text.tex
Please enter the commit message for your changes. Lines starting
consider two particular examples

@@
 \item {\bf Left redisual move}
+
 \item {\bf Right residual move}
+
 \end{enumerate}
@@
 \item $\exists$ has a winning stragery in $\mathcal{G}_{\omega}(\mathcal{A})$
+
+\begin{center}
+$rep(a) = \{ (x,y) \: | \: l(x, y) \leq a \}$
+\end{center}
+
+\begin{enumerate}
+\item $rep(a ; b) = rep(a) ; rep(b)$
+\item $rep(a \setminus b) = rep(a) \setminus rep(b)$
+\item $rep(a / b) = rep(a) / rep(b)$
+\end{enumerate}
+
 \item $\mathcal{A}$ is representable.
@@
 
-\section{Games for the FMP}
+\section{Step-by-step construction for the FMP}
+
+\subsection{Example 1}
+
+Consider a sequent $\varphi_1 \bullet \varphi_2 \rightarrow \varphi_1 \bullet (\psi_1 \setminus \psi_2)$, which is known to unprovable in the Lambek calculus.
+We would like to construct a finite model on a network such that there exists $(u, v) \nvDash \varphi_1 \bullet \varphi_2 \rightarrow \varphi_1 \bullet (\psi_1 \setminus \psi_2)$, that is, $(u, v) \in ||\varphi_1|| ; ||\varphi_2||$ and $(u, v) \notin ||\varphi_1|| ; (||\psi_1|| \setminus ||\psi_2||)$.
+
+The construction is step-by-step:
+\begin{enumerate}
+\item {\bf Step 0}
+
+We construct the edge $(u, v)$ labelled by a pair of sets $\lambda_0$ and $\delta_0$:
+
+\centerline{
+\xymatrix{
+u \ar[rr]^{\lambda_0}_{\delta_0} && v
+}
+}
+
+\begin{center}
+$\lambda_{0}(u, v) = \{ \varphi_1 \bullet \varphi_2, -(\varphi_1 \bullet (\psi_1 \setminus \psi_2)) \}$
+
+$\delta_{0}(u, v) = \{ \varphi_1 \bullet \varphi_2\}$
+\end{center}
+
+\item {\bf Step 1}
+
+We have to create a witness for $\varphi_1 \bullet \varphi_2 \in \delta_{0}(u, v)$ by adding a vertice $w$:
+
+\centerline{
+\xymatrix{
+& w \ar[ddr]^{\lambda_1}_{\delta_1} \\
+\\
+u \ar[uur]^{\lambda_1}_{\delta_1}
+\ar[rr]^{\lambda_0}_{\delta_0} && v
+}
+}
+
+\begin{center}
+$\lambda_1(u, w) = \{ \varphi_1 \}$
+
+$\lambda_1(w, v) = \{ \varphi_2, -(\psi_1 \setminus \psi_2)\}$
+
+$\delta_1(u, w) = \emptyset$
+
+$\delta_1(w, v) = \{ - (\psi_1 \setminus \psi_2) \}$
+\end{center}
+
+\item {\bf Step 2}
+
+We create a witness $w'$ for $- (\psi_1 \setminus \psi_2) \in \delta_1(w, v)$
+
+\centerline{
+\xymatrix{
+w \ar[ddrr]^{\lambda_1}_{\delta_1} && w' \ar[ll]_{\lambda_2}^{\delta_2} \ar[dd]_{\lambda_2}^{\delta_2} \\
+\\
+u \ar[uu]^{\lambda_1}_{\delta_1}
+\ar[rr]^{\lambda_0}_{\delta_0} && v
+}
+}
+
+By putting:
+\begin{center}
+$\lambda_2(w', w) = \{ \psi_1 \}$
+
+$\lambda_2(w', v) = \{ -\psi_2 \}$
+
+$\delta_2(w', w) = \emptyset$
+
+$\delta_2(w', v) = \emptyset$
+\end{center}
+
+\end{enumerate}
+
+Now let us describe $||\varphi_1|| ; ||\varphi_2||$ as follows:
+
+TODO:
+
+\subsection{Example 2}
+
+Let us consider a sequent $LC \not\vdash \varphi_1 \setminus (\varphi_2 \bullet \varphi_3) \rightarrow \varphi_1 \setminus \psi$
+
+\begin{enumerate}
+\item {\bf Step 0}
+
+Introduce a wedge $(u, v)$ with labelling $\lambda_0$ and $\delta_0$
+
+\centerline{
+\xymatrix{
+u \ar[rr]^{\lambda_0}_{\delta_0} && v
+}
+}
+
+where
+
+\begin{center}
+$\lambda_0(u, v) = \{\varphi_1 \setminus (\varphi_2 \bullet \varphi_3), -(\varphi_1 \setminus \psi)\}$
+
+$\delta_0(u, v) = \{ -(\varphi_1 \setminus \psi) \}$
+\end{center}
+
+\item {\bf Step 1}
+
+We create a witness $w$ for $-(\varphi_1 \setminus \psi) \in \delta_0(u, v)$
+
+\centerline{
+\xymatrix{
+& w \ar[ddr]^{\lambda_1}_{\delta_1} \ar[ddl]_{\lambda_1}^{\delta_1} \\
+\\
+u \ar[rr]^{\lambda_0}_{\delta_0} && v
+}
+}
+where
+\begin{center}
+$\lambda_1(w, u) = \{ \varphi_1 \}$
+
+$\lambda_1(w, v) = \{ -\psi, \varphi_1 \bullet (\varphi_1 \setminus (\varphi_2 \bullet \varphi_3)), \varphi_2 \bullet \varphi_3\}$
+
+$\delta_1(w, u) = \emptyset$
+
+$\delta_1(w, v) = \{ \varphi_2 \bullet \varphi_3 \}$
+
+\end{center}
+
+\item {\bf Step 2}
+
+We have to create a witness $w'$ for $\varphi_2 \bullet \varphi_3 \in \delta_1(w, v)$
+
+\centerline{
+\xymatrix{
+w \ar[ddrr]^{\lambda_1}_{\delta_1} \ar[dd]_{\lambda_1}^{\delta_1} \ar[rr]^{\lambda_2}_{\delta_2} && w' \ar[dd]_{\lambda_2}^{\delta_2} \\
+\\
+u \ar[rr]^{\lambda_0}_{\delta_0} && v
+}
+}
+
+\begin{center}
+$\lambda_2(w, w') = \{ \varphi_2 \}$
+
+$\lambda_2(w', v) = \{ \varphi_3 \}$
+
+$\delta_2(w, w') = \emptyset$
+
+$\delta_2(w', v) = \emptyset$
+\end{center}
+
+\end{enumerate}
+
+\subsection{General case}
+
+Consider a sequent $\Gamma \rightarrow \varphi$ such that $LC \not\vdash \Gamma \rightarrow \varphi$.
+We have to construct a finite relational model over some network such that $(u, v) \not\models \Gamma \rightarrow \varphi$, that is, $(u, v) \models \Gamma$ and $(u, v) \not\models \varphi$.
 

### 2023-03-29 finitely-axiomatisable-reducts/Text.tex
initialise a new project

\documentclass[a4paper]{article}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{bussproofs}
\usepackage{mathtools}
\usepackage{verbatim}
\usepackage{dsfont}
\usepackage{mathabx}
\usepackage[all, 2cell]{xy}
\usepackage[all]{xy}
\usepackage{wasysym}
\usepackage{rotating}
\usepackage{geometry}
\usepackage{trfsigns}
\usepackage{cmll}
\usepackage{authblk}
\usepackage{hyperref}
\usepackage{cleveref}
\usepackage{lipsum}
\usepackage{extpfeil}
\usepackage{soul}
\usepackage{graphicx}

\newcommand\mapsfrom{\mathrel{\reflectbox{\ensuremath{\mapsto}}}}

\theoremstyle{defin}
\newtheorem{definition}{Definition}

\theoremstyle{theorem}
\newtheorem{theorem}{Theorem}

\theoremstyle{claim}
\newtheorem{claim}{Claim}

\theoremstyle{prop}
\newtheorem{prop}{Proposition}

\theoremstyle{lemma}
\newtheorem{lemma}{Lemma}

\theoremstyle{fact}
\newtheorem{fact}{Fact}

\theoremstyle{ex}
\newtheorem{ex}{Example}

\theoremstyle{col}
\newtheorem{col}{Corollary}

\let\strokeL\L
\DeclareRobustCommand{\L}{\ifmmode\mathbf{L}\else\strokeL\fi}

\author{Daniel Rogozin}
\date{}
\title{Finitely axiomatisable varieties generated by representable relation algebra reducts}

\begin{document}
\maketitle

\nocite{*}

\section{Questions}

\begin{enumerate}
\item Is the variety ${\bf V}({\bf R}(\cdot, ;, {\bf 1}))$ finitely axiomatisable?
\item Is the variety ${\bf V}({\bf R}(\cdot, +, ;, {\bf 1}))$ finitely axiomatisable?
\end{enumerate}

\section{Definitions}

A structure $\mathcal{M} = (M, \cdot, ;, {\bf 1})$ is called a \emph{lower-semilattice ordered monoid} if the following axioms hold:
\begin{itemize}
\item $(M, \cdot)$ is a meet-semilattice,
\item $(M, ;, {\bf 1})$ is a monoid,
\item $({\bf 1} \cdot x) ; ({\bf 1} \cdot y) = {\bf 1} \cdot x \cdot y$,
\item $({\bf 1} \cdot x) ; (y \cdot z) = ({\bf 1} \cdot x) ; y \cdot z$,
\item $(x \cdot y) ; ({\bf 1} \cdot z) = x \cdot y ; ({\bf 1} \cdot z)$.
\end{itemize}
${\bf LSMod}$ is the class (variety) of all lower-semilattice ordered monoids.

A lower-semilattice ordered monoid $\mathcal{M}$ is \emph{representable} if there exists a map $f : \mathcal{M} \to 2^{X}$ for some $X \neq \emptyset$ such that:
\begin{itemize}
\item $f$ is one-to-one
\item $f(a ; b) = f(a) | f(b)$
\item $f({\bf 1}) = {\bf Id}$
\item $f(a \cdot b) = f(a) \cap f(b)$
\end{itemize}
${\bf R}(\cdot, ;, {\bf 1})$ is the class of all representable lower-semilattice ordered monoids

\subsection{Stone-style representation for lower-semilattices}

Let $\mathcal{L}$ be a lower semilattice.

A subset $F \subseteq \mathcal{F}(\omega)$ is a filter if $F$ is a downward closed meet-subsemilattice. $F$ is principal if there is some $a \in F$ such that $F = \uparrow a$. Let $A \subseteq \mathcal{F}(\omega)$, then $\langle A \rangle = \cup_{a \in A} \uparrow a$, the filter generated by $F$.

A subset $I \subseteq \mathcal{L}$ is an \emph{ideal} if it is upward closed and updirected, that is, $a, b \in I$ implies there exists some $c$ such that $a, b \leq c$. It is known that $F$ is a prime filter iff $L \setminus F$ is a prime ideal.

\begin{theorem}
Let $\mathcal{L}$ be a lower semilattice. Then $\mathcal{L} \hookrightarrow (2^{\operatorname{Spec}(\mathcal{L})})$ whenever $\mathcal{L}$ is distributive.
\end{theorem}

\subsection{Finite axiomatisability}

\begin{theorem}
${\bf LSMod} = {\bf V}({\bf R}(\cdot, ;, {\bf 1}))$.
\end{theorem}

The right-to-left inclusion is obvious. To show the left-to-right inclusion, one needs to show that the free lower-semilattice ordered monoid with $\omega$ generators $\mathcal{F}(\omega)$ is representable.

A network is a structure $\mathcal{N} = (V, E, l)$, where $V$ is a set of vertices, $E$ is a set of edges and $l : E \to 2^{\mathcal{F}(\omega)}$ is a labelling function with the following data:
\begin{itemize}
\item Each $l(x,y)$ is a filter,
\item $l(x,y) ; l(y, z) \subseteq l(y, z)$,
\item ${\bf 1'} \in l(x, y)$ implies $x = y$,
\item
\end{itemize}

We define a back-and-forth representability game $\mathcal{G}_{\omega}(\mathcal{F}(\omega))$ with two players $\forall$ and $\exists$.

\section{The distributive lattice case}

\begin{theorem}
Let $\mathcal{L}$ be a distributive lattice, then $\mathcal{L} \hookrightarrow 2^{\operatorname{Spec}(\mathcal{L})}$.
\end{theorem}

Given a distributive-lattice ordered monoid $\mathcal{M}$, its canonical extension is a structure $\mathcal{M}_{+} = (\operatorname{Spec}(\mathcal{M}), \subseteq, R, E)$, where
\begin{itemize}
\item
\item
\item
\end{itemize}

\begin{definition}
Join-irreducibles...
\end{definition}

TODO: complete representation vs atomic representation vs representations via join-irreducibles

TODO: Birkhoff representation

TODO: Raney representation

\begin{theorem}
$(\mathcal{F}_{\omega})_+^+$ is completely representable.
\end{theorem}

\bibliographystyle{alpha}
\bibliography{Text}

\end{document}

### 2023-04-01 duality/Text.tex
initialise a new project

--- 2023-03-29 finitely-axiomatisable-reducts/Text.tex
@@
 \date{}
-\title{Finitely axiomatisable varieties generated by representable relation algebra reducts}
+\title{Characterising representable positive relation algebras via Priestley duality}
 
@@
 
-\section{Questions}
+\section{Priestley duality}
 
-\begin{enumerate}
-\item Is the variety ${\bf V}({\bf R}(\cdot, ;, {\bf 1}))$ finitely axiomatisable?
-\item Is the variety ${\bf V}({\bf R}(\cdot, +, ;, {\bf 1}))$ finitely axiomatisable?
-\end{enumerate}
+\section{Positive relation algebras and their representatition}
 
-\section{Definitions}
+\section{Spectral spaces for positive relation algebras}
 
-A structure $\mathcal{M} = (M, \cdot, ;, {\bf 1})$ is called a \emph{lower-semilattice ordered monoid} if the following axioms hold:
-\begin{itemize}
-\item $(M, \cdot)$ is a meet-semilattice,
-\item $(M, ;, {\bf 1})$ is a monoid,
-\item $({\bf 1} \cdot x) ; ({\bf 1} \cdot y) = {\bf 1} \cdot x \cdot y$,
-\item $({\bf 1} \cdot x) ; (y \cdot z) = ({\bf 1} \cdot x) ; y \cdot z$,
-\item $(x \cdot y) ; ({\bf 1} \cdot z) = x \cdot y ; ({\bf 1} \cdot z)$.
-\end{itemize}
-${\bf LSMod}$ is the class (variety) of all lower-semilattice ordered monoids.
+\section{Complete representability}
 
-A lower-semilattice ordered monoid $\mathcal{M}$ is \emph{representable} if there exists a map $f : \mathcal{M} \to 2^{X}$ for some $X \neq \emptyset$ such that:
-\begin{itemize}
-\item $f$ is one-to-one
-\item $f(a ; b) = f(a) | f(b)$
-\item $f({\bf 1}) = {\bf Id}$
-\item $f(a \cdot b) = f(a) \cap f(b)$
-\end{itemize}
-${\bf R}(\cdot, ;, {\bf 1})$ is the class of all representable lower-semilattice ordered monoids
+\subsection{Completely representable distibutive lattices}
 
-\subsection{Stone-style representation for lower-semilattices}
+\subsection{Completely representable positive relation algebras}
 
-Let $\mathcal{L}$ be a lower semilattice.
-
-A subset $F \subseteq \mathcal{F}(\omega)$ is a filter if $F$ is a downward closed meet-subsemilattice.
-$F$ is principal if there is some $a \in F$ such that $F = \uparrow a$.
-Let $A \subseteq \mathcal{F}(\omega)$, then $\langle A \rangle = \cup_{a \in A} \uparrow a$, the filter generated by $F$.
-
-A subset $I \subseteq \mathcal{L}$ is an \emph{ideal} if it is upward closed and updirected, that is, $a, b \in I$ implies there exists some $c$ such that $a, b \leq c$.
-It is known that $F$ is a prime filter iff $L \setminus F$ is a prime ideal.
+\section{The main result}
 
 \begin{theorem}
-Let $\mathcal{L}$ be a lower semilattice. Then $\mathcal{L} \hookrightarrow (2^{\operatorname{Spec}(\mathcal{L})})$ whenever $\mathcal{L}$ is distributive.
+Let $\mathcal{R}$ be a positive relation algebra, then $\mathcal{R}$ is representable iff $(\mathcal{R}_{+})^{+}$ is completely representable.
 \end{theorem}
 
-\subsection{Finite axiomatisability}
-
 \begin{theorem}
-${\bf LSMod} = {\bf V}({\bf R}(\cdot, ;, {\bf 1}))$.
+${\bf RPRA}$ is a canonical variety.
 \end{theorem}
 
-The right-to-left inclusion is obvious.
-To show the left-to-right inclusion, one needs to show that the free lower-semilattice ordered monoid with $\omega$ generators $\mathcal{F}(\omega)$ is representable.
-
-A network is a structure $\mathcal{N} = (V, E, l)$, where $V$ is a set of vertices, $E$ is a set of edges and $l : E \to 2^{\mathcal{F}(\omega)}$ is a labelling function with the following data:
-\begin{itemize}
-\item Each $l(x,y)$ is a filter,
-\item $l(x,y) ; l(y, z) \subseteq l(y, z)$,
-\item ${\bf 1'} \in l(x, y)$ implies $x = y$,
-\item
-\end{itemize}
-
-We define a back-and-forth representability game $\mathcal{G}_{\omega}(\mathcal{F}(\omega))$ with two players $\forall$ and $\exists$.
-
-\section{The distributive lattice case}
-
-\begin{theorem}
-Let $\mathcal{L}$ be a distributive lattice, then $\mathcal{L} \hookrightarrow 2^{\operatorname{Spec}(\mathcal{L})}$.
-\end{theorem}
-
-Given a distributive-lattice ordered monoid $\mathcal{M}$, its canonical extension is a structure $\mathcal{M}_{+} = (\operatorname{Spec}(\mathcal{M}), \subseteq, R, E)$, where
-\begin{itemize}
-\item
-\item
-\item
-\end{itemize}
-
-\begin{definition}
-Join-irreducibles...
-\end{definition}
-
-TODO: complete representation vs atomic representation vs representations via join-irreducibles
-
-TODO: Birkhoff representation
-
-TODO: Raney representation
-
-\begin{theorem}
-$(\mathcal{F}_{\omega})_+^+$ is completely representable.
-\end{theorem}
-
-\bibliographystyle{alpha}
-\bibliography{Text}
-
 \end{document}

### 2023-04-01 duality/Text.tex
a couple of lemmas

@@
 
-\section{Priestley duality}
+\section{Distributive lattice representation and priestley duality}
+
+\begin{lemma}
+Let $h : \mathcal{L} \to \mathcal{R}$ be a representation, then then
+\begin{center}
+$h^{-1}[x] = \{ a \in \mathcal{L} \: | \: x \in h(a) \}$
+\end{center}
+is prime filter in $\mathcal{L}$.
+\end{lemma}
+
+\begin{proof}
+Let $c \in h^{-1}[x]$ and $c \leq d$, then $x \in h(c)$, but $h$ is order-preserving, so $x \in h(d)$.
+If $c, d \in h^{-1}[x]$, then $x \in h(c)$ and $x \in h(d)$, so $x \in h(c) \cap h(d)$, then $x \in h(c \cdot d)$, so $c \cdot d \in h^{-1}[x]$.
+Let $c + d \in h^{-1}[x]$, then $x \in h(c + d) = h(c) \cup h(d)$, so either $x \in h(c)$ or $x \in h(d)$, so either $c \in h^{-1}[x]$ or $d \in h^{-1}[x]$.
+\end{proof}
 
@@
 \section{Complete representability}
+
+An element $a$ of a lattice $\mathcal{L}$ is \emph{completely join-irreducible} if for every $A \subseteq \mathcal{L}$ such that $a \leq \bigvee A$ there exists $b \in A$ such that $a \leq b$.
+
+A distributive lattice $\mathcal{L}$ is \emph{completely representable} if there exists a representation $h : \mathcal{L} \to \mathcal{R}$ such that
+\begin{center}
+$f(\Sigma A) = \bigcup \{ f(a) \: | \: a \in A\}$
+
+$f(\Pi A) = \bigcap \{ f(a) \: | \: a \in A \}$
+\end{center}
+for each $A \subseteq \mathcal{L}$ such that $\Sigma A$ and $\Pi A$ exist.
+
+A representation $h : \mathcal{L} \to \mathcal{R}$ is \emph{refined} if for all $a \in \mathcal{L}$ there exists a completely join-irreducible $j$ such that $a \in h(j)$.
+
+\begin{lemma}
+Let $h : \mathcal{L} \to \mathcal{R}$ be a \emph{refined} representation, then each $h^{-1}[x]$ is principal and is generated by completely join-irreducible element.
+\end{lemma}
+
+\begin{theorem}
+Let $\mathcal{L}$ be a distributive lattice and let $h : \mathcal{L} \to \mathcal{R}$ be representation of $\mathcal{L}$ over the base set $X$, then $h$ is refined iff it is complete.
+\end{theorem}
+
+\begin{proof}
+Suppose $h$ is refined. Take any $b \in \mathcal{L}$, then
+\begin{center}
+$h(b) = \bigcup \{ h(j) \: | \: j \in \mathcal{JR}(\mathcal{L}) \: \& \: j \leq a \}$
+\end{center}
+
+Take $A \subseteq \mathcal{L}$ such that $\Sigma A$ and $\Pi A$ exist.
+
+TODO:
+
+For converse, suppose $h$ is complete
+\end{proof}
 
@@
 \begin{theorem}
-Let $\mathcal{R}$ be a positive relation algebra, then $\mathcal{R}$ is representable iff $(\mathcal{R}_{+})^{+}$ is completely representable.
+Let $\mathcal{A}$ be a positive relation algebra, then $\mathcal{R}$ is representable iff $(\mathcal{R}_{+})^{+}$ is completely representable.
 \end{theorem}

### 2023-04-02 duality/Text.tex
some definitions

--- 2023-04-01 duality/Text.tex
@@
 
-\section{Priestley duality}
+\section{Distributive lattice representation and Priestley duality}
 
-\section{Positive relation algebras and their representatition}
+\begin{lemma}
+Let $h : \mathcal{L} \to \mathcal{R}$ be a representation, then then
+\begin{center}
+$h^{-1}[x] = \{ a \in \mathcal{L} \: | \: x \in h(a) \}$
+\end{center}
+is prime filter in $\mathcal{L}$.
+\end{lemma}
+
+\begin{proof}
+Let $c \in h^{-1}[x]$ and $c \leq d$, then $x \in h(c)$, but $h$ is order-preserving, so $x \in h(d)$.
+If $c, d \in h^{-1}[x]$, then $x \in h(c)$ and $x \in h(d)$, so $x \in h(c) \cap h(d)$, then $x \in h(c \cdot d)$, so $c \cdot d \in h^{-1}[x]$.
+Let $c + d \in h^{-1}[x]$, then $x \in h(c + d) = h(c) \cup h(d)$, so either $x \in h(c)$ or $x \in h(d)$, so either $c \in h^{-1}[x]$ or $d \in h^{-1}[x]$.
+\end{proof}
+
+\subsection{Priestley duality}
+
+\section{Representatiting positive relation algebras}
 
@@
 
+TODO: read \cite{egrot2012completely}
+
 \subsection{Completely representable positive relation algebras}
@@
 
+For that we need such model theoretic notions as saturation and types, see \cite[Section 6.3]{hodges1993model}.
+
+\begin{definition} Let $\mathcal{M}$ be a first-order structure of a signature $L$ and $S \subseteq \mathcal{M}$.
+Let $L(S)$ be an extension of $L$ with copies of elements from $S$ as additional constants.
+We assume that $Cnst(L)$ and $S$ are disjoint.
+
+\begin{enumerate}
+\item Let $n < \omega$, an $n$-type over $S$ is a set $\mathcal{T}$ of $L(S)$ formulas $A(\overline{x})$, where $\overline{x}$ is a fixed $n$-tuple of elements from $S$.
+Notation: $\mathcal{T}(\overline{x})$.
+A type is an $n$-type for some $n < \omega$.
+\item An $n$-type $\mathcal{T}(\overline{x})$ is realised in $\mathcal{M}$, if there exists $\overline{m} \in \mathcal{M}^n$ such that $\mathcal{M} \models A(\overline{m})$ for every $A \in \mathcal{T}(\overline{x})$.
+$\mathcal{M}$ omits $\mathcal{T}(\overline{x})$, if $\mathcal{T}(\overline{x})$ is not realised in $\mathcal{M}$.
+\item $\mathcal{T}(\overline{x})$ is finitely satisfied in $\mathcal{M}$, if every finite subtype $\mathcal{T}_0(\overline{x}) \subseteq \mathcal{T}(\overline{x})$ is realised in $\mathcal{M}$.
+We can reformulate that as $\mathcal{M} \models \exists \overline{a} \bigwedge \limits_{A \in \mathcal{T}_0} A(\overline{a})$.
+\item Let $T$ be a theory, then a type $\mathcal{T}$ over the empty set of constants is $T$-consistent, if there exists a model $\mathcal{M} \models T$ such that $\mathcal{T}$ is finitely satisfied in $\mathcal{M}$.
+\item Let $\kappa$ be a cardinal, then $\mathcal{M}$ is $\kappa$-saturated, if for every $S \subseteq \mathcal{M}$ with $|S| < \kappa$ every finitely satisfied $1$-type $\mathcal{T}$ is realised in $\mathcal{M}$.
+\end{enumerate}
+\end{definition}
+
+By default, a saturated model is an $\omega$-saturated model for us.
+
+The useful facts, they are from \cite{chang1990model} and \cite{hodges1993model}:
+
+\begin{fact} Let $\mathcal{M}$ be an FO-structue and $\kappa$ a cardinal, then:
+\begin{enumerate}
+\item $\mathcal{M}$ is $\kappa$-saturated iff every finitely satisfiable $\alpha$-type (an arbitrary $\alpha \leq \kappa$) with fewer than $\kappa$ parameters is realised in  $\mathcal{M}$.
+\item If $\mathcal{M}$ is $\kappa$-saturated, then $\mathcal{M}$ is $\lambda$-saturated for every $\lambda < \kappa$.
+\item \label{saturation} Every consistent theory has a $\kappa$-saturated model and every model has an elementary $\kappa$-saturated extension.
+\item Let $(\mathcal{M}_i)_{i < \omega}$ a family of structures of the (at most) countable signature and $D$ a non-principal ultrafilter over $\omega$, then $\Pi_D \mathcal{M}_i$ is $\omega_1$-saturated.
+\end{enumerate}
+\end{fact}
+
+Let $\mathcal{A}$ be a positive relation algebra, define the first-order relational language of the form
+\begin{center}
+$\mathcal{L}(\mathcal{A}) = (=, \{ R^2_a \}_{a \in \mathcal{A}})$
+\end{center}
+
+The $\mathcal{L}(\mathcal{A})$-theory ${T}_{\mathcal{A}}$ consists of the following statements:
+\begin{itemize}
+\item $\sigma_1 = \forall x \forall y ({\bf 1}'(x, y) \leftrightarrow (x = y))$
+\item $\sigma_+(R, S, T) = \forall x \forall y (R(x, y) \leftrightarrow S(x, y) \lor T (x, y))$
+\item $\sigma_{\cdot}(R, S, T) = \forall x \forall y (R(x, y) \leftrightarrow S(x, y) \land T (x, y))$
+\item $\sigma_{;}(R, S, T) = \forall x \forall y (R(x, y) \leftrightarrow \exists z (S(x, z) \land T(z, y)))$
+\item $\sigma_{\smile}(R, S) = \forall x \forall y (R(x, y) \leftrightarrow S(y, x))$
+\item $\sigma_{\neq 0} = \exists x \exists y R(x, y)$ for any $R_a$ such that $a \neq 0$
+\item $\sigma_{0} = \neg \exists x \exists y 0(x, y)$
+\item $\sigma_{{\bf 1}} = \forall x \forall y (R(x, y) \rightarrow {\bf 1}(x, y))$
+\end{itemize}
+
+\begin{prop}
+$T_{\mathcal{A}}$ is satisfiable whenever $\mathcal{A}$ is representable.
+\end{prop}
+
 \begin{theorem}
-Let $\mathcal{R}$ be a positive relation algebra, then $\mathcal{R}$ is representable iff $(\mathcal{R}_{+})^{+}$ is completely representable.
+Let $\mathcal{A}$ be a positive relation algebra, then $\mathcal{R}$ is representable iff $(\mathcal{R}_{+})^{+}$ is completely representable.
 \end{theorem}
@@
 
+\bibliographystyle{alpha}
+\bibliography{Text}
 \end{document}

### 2023-04-02 duality/Text.tex
PRA-spaces

@@
 
-\begin{lemma}
+Given a bounded distributive lattice $\mathcal{L}$, a proper subset $F \subset \mathcal{L}$ is said to be a \emph{filter} if it is upward closed and closer under finite infima.
+A filter $F$ is \emph{prime} if $a + b \in F$ implies either $a \in F$ or $b \in F$.
+The spectrum of $\mathcal{L}$, denoted as $\operatorname{Spec}(\mathcal{L})$, is the set of all prime filters.
+
+A filter is \emph{complete} if whenever $\Pi T$ exists for $T \subseteq F$, then $\Pi T \in F$.
+A filter is \emph{completely prime} if whenever $\Sigma T$ exists for $T \subseteq F$, then there exists $t \in T$ such that $t \in F$.
+The dual definitions are for ideals.
+
+\begin{prop}
 Let $h : \mathcal{L} \to \mathcal{R}$ be a representation, then then
 \begin{center}
-$h^{-1}[x] = \{ a \in \mathcal{L} \: | \: x \in h(a) \}$
+$h^{-1}[x] = \{ a \in \mathcal{L} \: | \: x \in h(a) \} \in \operatorname{Spec}(\mathcal{L})$
 \end{center}
-is prime filter in $\mathcal{L}$.
-\end{lemma}
+\end{prop}
 
-\begin{proof}
-Let $c \in h^{-1}[x]$ and $c \leq d$, then $x \in h(c)$, but $h$ is order-preserving, so $x \in h(d)$.
-If $c, d \in h^{-1}[x]$, then $x \in h(c)$ and $x \in h(d)$, so $x \in h(c) \cap h(d)$, then $x \in h(c \cdot d)$, so $c \cdot d \in h^{-1}[x]$.
-Let $c + d \in h^{-1}[x]$, then $x \in h(c + d) = h(c) \cup h(d)$, so either $x \in h(c)$ or $x \in h(d)$, so either $c \in h^{-1}[x]$ or $d \in h^{-1}[x]$.
-\end{proof}
+Recall that a \emph{Priestley space} is a triple $\mathcal{X} = (X, \tau, \leq)$ such that $(X, \tau)$ is a compact topological space, $(X, \leq)$ is a bounded poset such that if $x \not\leq y$, then there exists a clopen $U$ such that $x \in U$ and $y \notin U$.
+Given a bounded distributive lattice $\mathcal{L}$, define the map $\phi : \mathcal{L} \to 2^{\operatorname{Spec}(\mathcal{L})}$ such that
+\begin{center}
+$\phi : a \mapsto \{ F \in \operatorname{Spec}(\mathcal{L}) \: | \: a \in F \}$
+\end{center}
 
-\subsection{Priestley duality}
+\begin{fact}
+
+\begin{enumerate}
+\item The sets $\phi(a)$ and $- \phi(a)$ form the subbasis of the topology $\tau$ on $\operatorname{Spec}(\mathcal{L})$.
+\item $(\operatorname{Spec}(\mathcal{L}), \tau, \subseteq)$ is a Priestley space.
+\end{enumerate}
+\end{fact}
+
+Given a Priestley space $\mathcal{X} = (X, \tau, \leq)$, the set $\operatorname{ClOp}(\mathcal{X})$ consists of all clopens of $\mathcal{X}$.
+The structure $(\operatorname{ClOp}(\mathcal{X}), \cap, \cup, \emptyset, X)$ is a distributive lattice.
+
+\begin{fact}
+
+\begin{enumerate} Let $\mathcal{L}$ be a distributive lattice and let $\mathcal{X}$ be a Priestley space:
+\item $\mathcal{L} \hookrightarrow \mathcal{L}^{+} = (2^{\operatorname{Spec}(\mathcal{L})}, \cap, \cup, \emptyset, \operatorname{Spec}(\mathcal{L}))$,
+\item $\mathcal{L} \cong \operatorname{ClOp}(\operatorname{Spec}(\mathcal{L}))$,
+\item $\mathcal{X} \cong \operatorname{Spec}(\operatorname{ClOp}(\mathcal{X}))$,
+\item The categories of Priestley spaces and bounded distributive lattices are dually equivalent.
+\end{enumerate}
+\end{fact}
 
@@
 
+\begin{definition}
+
+A positive relation algebra is a algebra $\mathcal{R} = (R, \cdot, +, ;, \smile, 0, 1, {\bf 1}')$ such that
+\begin{enumerate}
+\item $(R, \cdot, +, 0, 1)$ is a bounded distributive lattice,
+\item $(R, ;, {\bf 1}')$ is a monoid,
+\item for all $a, b, c \in R$
+\begin{enumerate}
+\item $a ; (b + c) = a ; b + a ; c$,
+\item $a^{\smile \smile} = a$,
+\item $(a + b)^{\smile} = a^{\smile} + b^{\smile}$,
+\item $(a ; b)^{\smile} = b^{\smile} ; a^{\smile}$,
+\item $(a ; b) \cdot c^{\smile} = 0 \leftrightarrow (b ; c) \cdot c^{\smile} = 0$.
+\end{enumerate}
+\end{enumerate}
+\end{definition}
+
+A positive relation algebra $\mathcal{R}$ is \emph{representable} if there exists a one-to-one function $h : \mathcal{R} \to 2^{X \times X}$ over the base set $X \neq \emptyset$ such that:
+\begin{itemize}
+\item $f(a \cdot b) = f(a) \cap f(b)$,
+\item $f(a + b) = f(a) + f(b)$,
+\item $f(0) = \emptyset$,
+\item $f(1) = \bigcup \limits_{a \in \mathcal{R}} f(a)$,
+\item $f({\bf 1}') = \Delta_X$,
+\item $f(a ; b) = \{ (x, z) \: | \: \exists y \in X \: ((x, y) \in f(a) \: \& \: (y,z) \in f(b))\} = f(a) | f(b)$,
+\item $f(a^{\smile}) = \{ (y, x) \: | \: (x, y) \in f(a) \}$.
+\end{itemize}
+
 \section{Spectral spaces for positive relation algebras}
+
+A PRA-space is a structure $(X, \tau, \leq, R, I, E)$ where $X = (X, \tau, \leq)$ is a Priestley space and $R \subseteq X^3$, $I \subseteq X^2$ and $E \subseteq X$ such that:
+\begin{itemize}
+\item For all $x, y, z, w \in X$ there exists $u \in X$ such that $R(x, y, u)$ and $R(u, z, w)$ iff there exists $v \in X$ such that $R(y,z,v)$ and $R(x, v, w)$,
+\item If $A, B \subseteq X$ are upward closed, so is $R[A, B, \underline{\:\:\:}]$,
+\item $I(A)$ is upward closed clopen whenever $A$ is upward closed clopen,
+\item $I(x)$ is closed for each $x \in X$,
+\item For all $x, y, z \in X$, $x \leq y$ and $I(x, z)$ imply $I(y, z)$,
+\item For all $x, y \in X$ there exists $z \in X$ such that $x = y$ iff $I(z, y)$ and $I(x, z)$,
+\item For all $x, y, z \in X$ there exists $u \in X$ such that $I(u, z)$ and $R(x, y, u)$ iff there exist $u, w \in X$ such that $R(v, w, z)$, $I(y,v)$ and $I(x, w)$.
+\item For all $x,y,u, v \in X$, $R(u,v,y)$ and $I(x, u)$ implies $R(x,y,v)$,
+\item $E$ is upward closed clopen such that for each clopen $A \subseteq X$ one has
+\begin{center}
+$R[A, E, \overline{\:\:\:}] = R[E, A, \overline{\:\:\:}] = A$
+\end{center}
+\end{itemize}
+
+\begin{lemma} Let $\mathcal{R}$ be a positive relation algebra, then
+
+\begin{enumerate}
+\item
+\item 
+\end{enumerate}
+\end{lemma}
 
@@
 \subsection{Completely representable distibutive lattices}
+
+Let $\mathcal{L}$ be a bounded distributive lattice, then a set $S \subseteq 2^{\mathcal{L}}$ is said to be distinguishing if for every $a, b \in \mathcal{L}$ such that $a \neq b$ there exists $s \in S$ such that either $a \in s$ and $b \notin b$ or vice versa.
+
+\begin{theorem} Let $\mathcal{L}$ be a bounded distributive lattice, then
+
+\begin{enumerate}
+\item $\mathcal{L}$ is completely representable iff $\mathcal{L}$ has a distinguishing set of complete, completely prime filters,
+\item $(\mathcal{L}_+)^{+}$ is completely representable.
+\end{enumerate}
+
+\end{theorem}
 

### 2023-04-02 duality/Text.tex
proof sketch with a couple of todos

@@
 \date{}
-\title{Characterising representable positive relation algebras via Priestley duality}
+\title{Characterising representable positive relation algebras with Priestley duality}
 
@@
 
+$ $
+
 \begin{enumerate}
@@
 
-\begin{fact}
+\begin{fact} Let $\mathcal{L}$ be a distributive lattice and let $\mathcal{X}$ be a Priestley space:
 
-\begin{enumerate} Let $\mathcal{L}$ be a distributive lattice and let $\mathcal{X}$ be a Priestley space:
+\begin{enumerate}
 \item $\mathcal{L} \hookrightarrow \mathcal{L}^{+} = (2^{\operatorname{Spec}(\mathcal{L})}, \cap, \cup, \emptyset, \operatorname{Spec}(\mathcal{L}))$,
@@
 \end{fact}
+
+\subsection{Completely representable distibutive lattices}
+
+Let $\mathcal{L}$ be a bounded distributive lattice, then a set $S \subseteq 2^{\mathcal{L}}$ is said to be distinguishing if for every $a, b \in \mathcal{L}$ such that $a \neq b$ there exists $s \in S$ such that either $a \in s$ and $b \notin b$ or vice versa.
+
+\begin{theorem} Let $\mathcal{L}$ be a bounded distributive lattice, then
+
+\begin{enumerate}
+\item $\mathcal{L}$ is completely representable iff $\mathcal{L}$ has a distinguishing set of complete, completely prime filters,
+\item $(\mathcal{L}_+)^{+}$ is completely representable.
+\end{enumerate}
+\end{theorem}
+
+TODO: read \cite{egrot2012completely}
 
@@
 
-\section{Spectral spaces for positive relation algebras}
+A positive relation algebra is \emph{completely representable} if it is representable and its bounded distributive lattice reduct is completely representable.
+
+\section{Priestley spaces for positive relation algebras}
 
@@
 
-\begin{lemma} Let $\mathcal{R}$ be a positive relation algebra, then
+\begin{fact} Let $\mathcal{R}$ be a positive relation algebra, then
 
 \begin{enumerate}
-\item
-\item 
+\item If $F \subseteq \mathcal{R}$ is a filter, then for each $X, Y \subseteq \mathcal{R}$ $X ; Y \subseteq F$ iff $\uparrow (X ; Y) \subseteq F$
+\item If $F_1, F_2 \subseteq \mathcal{R}$ are filters, so is $\uparrow (F_1 ; F_2)$,
+\item Let $F_1, F_2$ be filters and $F_3 \in \operatorname{Spec}(\mathcal{R})$ such that $\uparrow (F_1 ; F_2) \subseteq F_3$, then there are $F_1', F_2' \in \operatorname{Spec}(\mathcal{R})$ such that $F_1 \subseteq F_1'$, $F_2 \subseteq F_2'$ and $\uparrow (F_1' ; F_2') \subseteq F_3$.
 \end{enumerate}
-\end{lemma}
 
-\section{Complete representability}
+\end{fact}
 
-\subsection{Completely representable distibutive lattices}
-
-Let $\mathcal{L}$ be a bounded distributive lattice, then a set $S \subseteq 2^{\mathcal{L}}$ is said to be distinguishing if for every $a, b \in \mathcal{L}$ such that $a \neq b$ there exists $s \in S$ such that either $a \in s$ and $b \notin b$ or vice versa.
-
-\begin{theorem} Let $\mathcal{L}$ be a bounded distributive lattice, then
+\begin{theorem} Let $\mathcal{R}$ be a positive relation algebra and $\mathcal{X}$ a PRA-space, then
 
 \begin{enumerate}
-\item $\mathcal{L}$ is completely representable iff $\mathcal{L}$ has a distinguishing set of complete, completely prime filters,
-\item $(\mathcal{L}_+)^{+}$ is completely representable.
+\item $\mathcal{R} \hookrightarrow (2^{\operatorname{Spec}(R)}, \cup, \cap, \emptyset, \operatorname{Spec}(R), \bullet, \iota, \epsilon)$,
+\item $\mathcal{R} \cong \operatorname{ClOp}(\operatorname{Spec}(\mathcal{R}))$,
+\item $\mathcal{X} \cong \operatorname{Spec}(\operatorname{ClOp}(\mathcal{X}))$,
+\item The categories of positive relation algebras and PRA-spaces are dually equivalent.
 \end{enumerate}
-
 \end{theorem}
-
-TODO: read \cite{egrot2012completely}
-
-\subsection{Completely representable positive relation algebras}
 
@@
 \begin{theorem}
-Let $\mathcal{A}$ be a positive relation algebra, then $\mathcal{R}$ is representable iff $(\mathcal{R}_{+})^{+}$ is completely representable.
+Let $\mathcal{A}$ be a positive relation algebra, then $\mathcal{R}$ is representable iff $\mathcal{R}^{+}$ is completely representable.
 \end{theorem}
 
+\begin{proof}
+The right-to-left implication is easy.
+If $\mathcal{R}^{+}$ is representable (no completeness needed here), so is $\operatorname{ClOp}(\operatorname{Spec}(\mathcal{L}))$ as a subalgebra of $\mathcal{R}^{+}$.
+But, by Priestley duality for positive relation algebras, $\operatorname{ClOp}(\operatorname{Spec}(\mathcal{L})) \cong \mathcal{L}$, so $\mathcal{L}$ is representable.
+
+Assume that $\mathcal{R}$ is representable, then $T_{\mathcal{R}}$ is satisfiable, let $M \models T_{\mathcal{R}}$ and $M$ is $\omega$-saturated. Define a map $h : \mathcal{R}^+ \to 2^{M \times M}$ as
+\begin{center}
+$h : F \mapsto \{ (x, y) \in 1^M \: | \: f_{x, y} \in F \}$
+\end{center}
+where
+\begin{center}
+$f_{x, y} = \{ a \in \mathcal{R} \: | \: M \models R_a(x,y) \}$
+\end{center}
+
+\begin{claim}
+$f_{x, y} \in \operatorname{Spec}(\mathcal{R})$ whenever $M \models 1(x, y)$.
+\end{claim}
+
+\begin{proof}
+Let $a \in f_{x, y}$ and $a \leq b$.
+
+Then $M \models R_a(x, y)$. But $a \leq b$ iff $a \cdot b = a$, so, from the axiom $R_{\cdot}(R_a, R_a, R_b)$ we have $M \models R_b(x, y)$.
+
+Let $a, b \in f_{x, y}$, so $M \models R_a(x, y) \land R_b(x, y)$, so $R_{a \cdot b}(x, y)$ by the axiom $R_{\cdot}(R_a, R_a, R_b)$ again.
+
+Let $a + b \in f_{x, y}$, so $M \models R_{a + b}(x, y)$, then we have either $M \models R_a(x, y)$ or $M \models R_b(x, y)$, so either $a \in f_{x,y}$ or $b \in f_{x,y}$.
+\end{proof}
+
+\begin{claim}
+$h$ is one-to-one.
+\end{claim}
+
+\begin{claim}
+$h$ is a complete representation of the bounded lattice reduct.
+\end{claim}
+
+\begin{proof}
+Follows from the fact that $\operatorname{Spec}(\mathcal{L})$ is a Hausdorff space. (????)
+\end{proof}
+
+\begin{claim}
+$h$ preserves the structure of an involutive monoid.
+\end{claim}
+\end{proof}
+
 \begin{theorem}
-${\bf RPRA}$ is a canonical variety.
+${\bf RPRA}$ is a canonical variety. (is it a variety at all?)
 \end{theorem}
+
+\section{Union-free reducts of positive relation algebras}
+
+TODO: \cite{bezhanishvili2011priestley}.
 

### 2023-04-03 duality/Text.tex
minor updates and todos

@@
 
-\section{Priestley spaces for positive relation algebras}
+\subsection{Priestley duality for positive relation algebras}
+
+Given a positive relation algebra $\mathcal{R} = (R, \cdot, +, ;, \smile, {\bf 1}', 0, 1)$, let $A, B \subseteq \mathcal{R}$, define the usual pointwise product operation:
+\begin{center}
+$A ; B = \{ a ; b \: | \: a \in A, b \in B \}$
+\end{center}
+We modify this operation with the upward closure:
+\begin{center}
+$A \bullet B = \uparrow (A ; B)$
+\end{center}
+
+We use the $\bullet$ operation to analyse the following properties of filters in positive relation algebras:
+\begin{fact} Let $\mathcal{R}$ be a positive relation algebra, then
+
+\begin{enumerate}
+\item If $F \subseteq \mathcal{R}$ is a filter, then for each $X, Y \subseteq \mathcal{R}$ $X ; Y \subseteq F$ iff $X \bullet Y \subseteq F$
+\item If $F_1, F_2 \subseteq \mathcal{R}$ are filters, so is $F_1 \bullet F_2$,
+\item Let $F_1, F_2$ be filters and $F_3 \in \operatorname{Spec}(\mathcal{R})$ such that $F_1 \bullet F_2 \subseteq F_3$, then there are $F_1', F_2' \in \operatorname{Spec}(\mathcal{R})$ such that $F_1 \subseteq F_1'$, $F_2 \subseteq F_2'$ and $F_1' \bullet F_2' \subseteq F_3$.
+\end{enumerate}
+\end{fact}
+
+First of all, given a positive relation algebra $\mathcal{R} = (R, \cdot, +, ;, \smile, 0, 1, {\bf 1}')$, we define its canonical extension $\mathcal{R}^{+}$ on the subsets of the spectrum $\operatorname{Spec}(\mathcal{R})$ by piggybacking Priestley representation of bounded distibutive lattices.
+
+TODO:
 
 A PRA-space is a structure $(X, \tau, \leq, R, I, E)$ where $X = (X, \tau, \leq)$ is a Priestley space and $R \subseteq X^3$, $I \subseteq X^2$ and $E \subseteq X$ such that:
+\begin{enumerate}
+\item For all $x, y, z \in X$ such that $R(x, y, z)$:
 \begin{itemize}
+\item $x \geq x'$ for $x' \in X$ implies $R(x', y, z)$,
+\item $y \leq y'$ for $y' \in X$ implies $R(x, y', z)$,
+\item $z \leq z'$ for $z' \in X$ implies $R(x, y, z')$.
+\end{itemize}
 \item For all $x, y, z, w \in X$ there exists $u \in X$ such that $R(x, y, u)$ and $R(u, z, w)$ iff there exists $v \in X$ such that $R(y,z,v)$ and $R(x, v, w)$,
@@
 \end{center}
-\end{itemize}
-
-\begin{fact} Let $\mathcal{R}$ be a positive relation algebra, then
-
-\begin{enumerate}
-\item If $F \subseteq \mathcal{R}$ is a filter, then for each $X, Y \subseteq \mathcal{R}$ $X ; Y \subseteq F$ iff $\uparrow (X ; Y) \subseteq F$
-\item If $F_1, F_2 \subseteq \mathcal{R}$ are filters, so is $\uparrow (F_1 ; F_2)$,
-\item Let $F_1, F_2$ be filters and $F_3 \in \operatorname{Spec}(\mathcal{R})$ such that $\uparrow (F_1 ; F_2) \subseteq F_3$, then there are $F_1', F_2' \in \operatorname{Spec}(\mathcal{R})$ such that $F_1 \subseteq F_1'$, $F_2 \subseteq F_2'$ and $\uparrow (F_1' ; F_2') \subseteq F_3$.
 \end{enumerate}
-
-\end{fact}
 
@@
 \begin{enumerate}
-\item $\mathcal{R} \hookrightarrow (2^{\operatorname{Spec}(R)}, \cup, \cap, \emptyset, \operatorname{Spec}(R), \bullet, \iota, \epsilon)$,
+\item $\mathcal{R} \hookrightarrow \mathcal{R}^{+} = (2^{\operatorname{Spec}(R)}, \cup, \cap, \emptyset, \operatorname{Spec}(R), \bullet, \iota, \epsilon)$,
 \item $\mathcal{R} \cong \operatorname{ClOp}(\operatorname{Spec}(\mathcal{R}))$,
@@
 \end{enumerate}
+
+\begin{proof} $ $
+
+\begin{enumerate}
+\item We check only the cases of $\bullet$, $\iota$ and $\epsilon$.
+\item Note that the dual space of a positive relation algebra $\mathcal{R}$ is a PRA-space.
+
+TODO: check the claim above
+
+\item The dual space of a PRA-space is a positive relation algebra.
+
+TODO: check the claim above
+
+\item
+\end{enumerate}
+\end{proof}
+
 \end{theorem}
